### 2017-01-31 deforestmap.Rmd
Initial commit

---
title: "Sixty-five years of deforestation and forest degradation in Madagascar"
bibliography: bib/biblio.bib
csl: bib/conservation-letters.csl
output: 
  html_document:
    fig_caption: yes
  pdf_document:
    fig_caption: true
    keep_tex: true
---

```{r load_data}
load("deforestmap.rda")
```

#### Ghislain Vieilledent<sup>*,1,2</sup>, Clovis Grinand<sup>3</sup>, Fety A. Rakotomalala<sup>3</sup>, Jean-Roger Rakotoarijaona<sup>5</sup>, and Frédéric Achard<sup>1</sup>

(*) /Email: ghislain.vieilledent@cirad.fr, /Phone: +39.329.457.2273

1. Joint Research Center of the European Commission, Bio-economy Unit (JRC.D.1), I-21027 Ispra (VA), ITALY
2. Cirad, UPR Forêts et Sociétés, F-34398 Montpellier, FRANCE
3. ETC Terra, F-75016 Paris, FRANCE
4. Wildlife Conservation Society, 101 Antananarivo, MADAGASCAR
5. Office National de l'Environnement, 101 Antananarivo, MADAGASCAR

## Summary

1. xxx

2. xxx

**Keywords**: biodiversity, climate-change, deforestation, Landsat, protected areas, Madagascar, tropical forest, tree cover.

**Potential journals** (2015 IF): Conservation Letters (7.126), Conservation Biology (4.257), Environmental Research Letters (4.134), Biological Conservation (3.985), Landscape Ecology (3.657), Ambio (2.555), Environmental Conservation (2.235), Biotropica (1.944), Tropical Conservation Science (1.55).

## 1. Introduction

Tropical forests provide important ecosystem services at the global scale, such as biodiversity conservation and climate regulation [@Costanza1997] but also many at the regional scale for local populations. At the local scale, tropical forests supply clean water, can preserve populations from flood [@Bradshaw2007] and provide wood for construction and fuel and other non-woody products for food and health. On the island of Madagascar, three types of tropical forests, covering about 9 millions of hectares (15% of the country area) can be found: the moist forest in the East, the dry forest in the West and the spiny forest in the South [@Vieilledent2016]. Among all the ecosystem services cited previously, Malagasy forests are particularly important for the unique biodiversity they shelter, both in term of number of species and endemism in many taxonomic groups [@Goodman2005].

- Harper,
- Hansen,
- Kim,
- ONE/PERR-FH (http://www.etcterra.org/fr/projets/perrfh)

## 2. Materials and Methods

### 2.1 Historical deforestation maps

We derived 30m cloud-free forest maps for Madagascar for the years 1990, 2000, 2010 and 2014 combining the 1990 and 2000 forest maps by @Harper2007 and the 2000-2014 tree percent cover map by @Hansen2013. On Harper's map, ~200,000 ha clouds are present over the ~4.2 million ha moist forest ecoregion (4.8% of clouds). To remove these clouds, we used the 2000 cloud-free tree percent cover map provided by @Hansen2013 (also at 30~m resolution) and we chose a threshold of 75% of tree cover to decide whether to replace cloud pixels with forest or non-forest pixels [@Achard2014]. We thus obtained a cloud-free year 2000 forest map for Madagascar. From this map, using tree cover loss data from 2000 to 2014 by @Hansen2013, we obtained cloud-free forest maps in 2010 and 2014 at 30~m resolution. Finally, we corrected the cloudy 1990 Harper forest map assuming that if forest was present in 2000, it was also present in 1990. Doing so, all clouds were removed and we obtained a cloud-free 1990 forest map for Madagascar. These maps are freely available at http:/bioscenemada.cirad.fr/forestmaps.

### 2.2 Deforestation intensity

### 2.3 Fragmentation

## 3. Results

## 4. Discussion

## Acknowledgements

This study is part of the BioSceneMada project (http://bioscenemada.cirad.fr) and the ReCaREDD project (http://roadlessforests.org). The BioSceneMada project is funded by FRB (Fondation pour la Recherche sur la Biodiversité) and the FFEM (Fond Français pour l'Environnement Mondial) under the project agreement AAP-SCEN-2013 I. The ReCaREDD project is funded by the European Commission.

## Data availability statement

All the data and codes used for this study are made publicly available in the `deforestmap` GitHub repository (https://github.com/ghislainv/deforestmap.git). The results and the manuscript are fully reproducible running the R script `deforestmap.R` located inside the `deforestmap` repository.

## Supplementary materials

## References

### 2017-01-31 deforestmap.R

@@ -1,0 +1,65 @@
+##=========================================================================
+## Sixty-five years of deforestation and forest fragmentation in Madagascar
+## Ghislain Vieilledent <ghislain.vieilledent@cirad.fr>
+## January 2017
+##=========================================================================
+
+##= Libraries
+pkg <- c("broom","sp","rgdal","raster","ggplot2","gridExtra",
+         "rasterVis","knitr","rmarkdown","rgeos")
+## broom: to convert map into data-frame with tiny()
+## gridExtra: to combine several ggplots
+## rasterVis: for gplot()
+## rgeos: for crop()
+load.pkg <- function(x) {
+  if(!require(x, character.only = T)) {
+    install.packages(x)
+    require(x, character.only = T)
+  }
+}
+loaded <- lapply(pkg,load.pkg)
+## Remove useless objects
+rm(pkg,load.pkg,loaded)
+
+##======================================================================
+## Download data (263.6 Mo): will have to be done from a Zenodo repository
+d <- "http://bioscenemada.cirad.fr/githubdata/deforestmap/deforestmap_data.zip"
+#download.file(url=d,destfile="menabe_data.zip",method="wget",quiet=TRUE)
+#unzip("menabe_data.zip")
+
+##===========================================================
+## Create new directories to save figures and new raster data
+dir.create("figs")
+dir.create("rast")
+
+##================================================
+
+x <- 1
+
+##========================
+## Save objects
+# load("menabe.rda")
+save(list=ls(all.names=TRUE),file="deforestmap.rda")
+
+##========================
+## Knit the document
+
+## Set knitr chunk default options
+opts_chunk$set(echo=FALSE, cache=FALSE,
+               results="hide", warning=FALSE,
+               message=FALSE, highlight=TRUE,
+               fig.show="hide", size="small",
+               tidy=FALSE)
+
+## Knit and translate to html and pdf
+dir.create("report")
+## Main document
+render("deforestmap.Rmd",output_dir="report") # html output
+## Cover letter
+render("coverletter.md",output_format=c("pdf_document"),output_dir="report") # pdf output
+#render("deforestmap.Rmd",output_format=c("html_document","pdf_document","word_document"),output_dir="report")
+
+##===========================================================================
+## End of script
+##===========================================================================
+

### 2017-01-31 deforestmap.sh

@@ -1,0 +1,6 @@
+#!/bin/sh
+echo "Executing deforestmap.R script in the background"
+Rscript --vanilla deforestmap.R > deforestmap.log 2>&1 &
+echo "Check the progress with command 'tail -f deforestmap.log'"
+echo "Check the processor usage with command 'top'"
+## End of script

### 2017-01-31 deforestmap.R
Cleaning

@@ -4,9 +4,23 @@
 ## January 2017
 ##=========================================================================
 
+## gdal library is needed to run this script
+## http://www.gdal.org/
+
+## GRASS GIS 7.x.x is also needed to run this script
+## https://grass.osgeo.org/
+
+## Read argument for download
+## Set "down" to TRUE if you want to download the sources. Otherwise, the data already provided in the gisdata repository will be used.
+arg <- commandArgs(trailingOnly=TRUE)
+down <- FALSE
+if (length(arg)>0) {
+  down <- arg[1]
+}
+
 ##= Libraries
 pkg <- c("broom","sp","rgdal","raster","ggplot2","gridExtra",
-         "rasterVis","knitr","rmarkdown","rgeos")
+         "rasterVis","knitr","rmarkdown","rgeos","rgdal","rgrass7")
 ## broom: to convert map into data-frame with tiny()
 ## gridExtra: to combine several ggplots
 ## rasterVis: for gplot()
@@ -21,20 +35,309 @@ loaded <- lapply(pkg,load.pkg)
 ## Remove useless objects
 rm(pkg,load.pkg,loaded)
 
+##=====================================
+## Create new grass location in UTM 38S
+#dir.create("grassdata")
+#system("grass72 -c epsg:32738 grassdata/deforestmap")  # Ignore errors
+## Connect R to grass location
+initGRASS(gisBase="/usr/local/grass-7.2.0",home=tempdir(), 
+          gisDbase="grassdata",
+          location="deforestmap",mapset="PERMANENT",
+          override=TRUE)
+
 ##======================================================================
 ## Download data (263.6 Mo): will have to be done from a Zenodo repository
-d <- "http://bioscenemada.cirad.fr/githubdata/deforestmap/deforestmap_data.zip"
-#download.file(url=d,destfile="menabe_data.zip",method="wget",quiet=TRUE)
-#unzip("menabe_data.zip")
+if (down) {
+  d <- "http://bioscenemada.cirad.fr/githubdata/deforestmap/deforestmap_data.zip"
+  download.file(url=d,destfile="deforestmap_data.zip",method="wget",quiet=TRUE)
+  unzip("deforest_data.zip")
+}
 
 ##===========================================================
 ## Create new directories to save figures and new raster data
 dir.create("figs")
[…]

### 2017-02-01 deforestmap.R
Run all

@@ -7,7 +7,7 @@
 ## gdal library is needed to run this script
 ## http://www.gdal.org/
 
-## GRASS GIS 7.x.x is also needed to run this script
+## GRASS GIS 7.2.x is also needed to run this script
 ## https://grass.osgeo.org/
 
 ## Read argument for download
@@ -47,16 +47,15 @@ initGRASS(gisBase="/usr/local/grass-7.2.0",home=tempdir(),
 
 ##======================================================================
 ## Download data (263.6 Mo): will have to be done from a Zenodo repository
-if (down) {
-  d <- "http://bioscenemada.cirad.fr/githubdata/deforestmap/deforestmap_data.zip"
+#if (down) {
+#  d <- "http://bioscenemada.cirad.fr/githubdata/deforestmap/deforestmap_data.zip"
   download.file(url=d,destfile="deforestmap_data.zip",method="wget",quiet=TRUE)
   unzip("deforest_data.zip")
 }
 
 ##===========================================================
-## Create new directories to save figures and new raster data
-dir.create("figs")
-dir.create("rast")
+## Create new directories to save outputs
+dir.create("outputs")
 
 #= Projection UTM38S (EPSG:32738)
 
@@ -69,23 +68,23 @@ nodat <- "0"
 proj.s <- "EPSG:32638"
 proj.t <- "EPSG:32738"
 Input <- "gisdata/rasters/harper/905_dsp_nosea.img"
-Output <- "rast/Forest_CI_905_30m.tif"
+Output <- "outputs/harper.tif"
 # gdalwarp
 system(paste0("gdalwarp -overwrite -dstnodata ",nodat," -s_srs ",proj.s," -t_srs ",proj.t," -te ",Extent,
              " -tr ",Res," ",Res," -r near -co 'COMPRESS=LZW' -co 'PREDICTOR=2' ",Input," ",Output))
 
-#= Import Forest_CI_905_30m.tif into grass location
-system("r.in.gdal --o input=rast/Forest_CI_905_30m.tif output=Forest_CI_905_30m")
+#= Import harper into grass location
+system("r.in.gdal --o input=outputs/harper.tif output=harper")
 # Set region
-system("g.region rast=Forest_CI_905_30m -ap")
+system("g.region rast=harper -ap")
 
 #= Typology of CI map
-system("r.info Forest_CI_905_30m")
-system("r.describe Forest_CI_905_30m") 
[…]

### 2017-02-01 deforestmap.sh
New bash

@@ -1,6 +1,8 @@
 #!/bin/sh
 echo "Executing deforestmap.R script in the background"
 Rscript --vanilla deforestmap.R > deforestmap.log 2>&1 &
+# If you wan't to download the source data do:
+# Rscript --vanilla deforestmap.R TRUE > deforestmap.log 2>&1 &
 echo "Check the progress with command 'tail -f deforestmap.log'"
 echo "Check the processor usage with command 'top'"
 ## End of script

### 2017-02-02 deforestmap.Rmd
Writing methodology

@@
 These maps are freely available at http:/bioscenemada.cirad.fr/forestmaps.
+On the basis of the forest maps, we computed the mean annual deforestation intensity (in hectares of forest and in percentage) for the five period of time from 1990 to 2014.
 
-### 2.2 Deforestation intensity
+### 2.2 Validation
+
+We estimated the accuracy of our approach combining Harper's and Hansen's products to derive cloud-free forest maps after the year 2000 for Madagascar.
+To do so, we used a deforestation map produced on the period 2005-2013 by a supervised classification of Landsat images at multiple dates using the Random Forest algorithm [@Grinand2013, @Rakotomalala2015].
+Following our methodology, we produced a deforestation map on the same period of time (2005-2013) and compute several accuracy indices (overall accuracy, Cohen's Kappa and True Skill Statistics) on the basis of a confusion matrix computed from the two deforestation maps at dates 2005 and 2013.
 
 ### 2.3 Fragmentation
+
+We report an analysis of the evolution of forest fragmentation based on the forest-cover maps for Madagascar for the years 1953, 1973, 1990, 2000, 2010 and 2014.
+We used the approach developed by @Riitters2000 which uses a moving window to characterize the fragmentation around each forested pixel.
+We used a moving window of 3x3 pixels (0.81 km<sup2</sup>).
+Six categories of fragmentation are identified from the amount of forest and its occurrence as adjacent forest pixels: interior, perforated, edge, transitional, patch, and undetermined.
+The interior category can be interpreted as intact forest (see @Potapov2017).
+We reported the area of forest falling in each category for the 6 years and analysed the evolution of the fragmentation.
 

### 2017-02-02 deforestmap.R

@@ -47,8 +47,8 @@ initGRASS(gisBase="/usr/local/grass-7.2.0",home=tempdir(),
 
 ##======================================================================
 ## Download data (263.6 Mo): will have to be done from a Zenodo repository
-#if (down) {
-#  d <- "http://bioscenemada.cirad.fr/githubdata/deforestmap/deforestmap_data.zip"
+if (down) {
+  d <- "http://bioscenemada.cirad.fr/githubdata/deforestmap/deforestmap_data.zip"
   download.file(url=d,destfile="deforestmap_data.zip",method="wget",quiet=TRUE)
   unzip("deforest_data.zip")
 }
@@ -201,27 +201,6 @@ system("r.out.gdal --o input=for2010 createopt='compress=lzw,predictor=2' type=B
 system("r.out.gdal --o input=for2005 createopt='compress=lzw,predictor=2' type=Byte output=outputs/for2005.tif")
 system("r.out.gdal --o input=for2000 createopt='compress=lzw,predictor=2' type=Byte output=outputs/for2000.tif")
 
-##= Compute fordefor2010. For test of function .sample() in  Python deforestprob module
-system("r.mapcalc --o 'fordefor2010 = if(isnull(for2010) && for2000 == 1, 0, for2010)'")
-system("echo '1 26:152:80 \n 0 red' | r.colors fordefor2010 rules=-")
-system("r.out.gdal --o input=fordefor2010 createopt='compress=lzw,predictor=2' type=Byte output=outputs/fordefor2010.tif")
-
-#= Deforestation statistics
-defor.df <- data.frame(Year=2000:2014,ncells=NA,area=NA,theta=NA)
-Year <- as.character(2000:2014)
-for (i in 1:length(Year)) {
-  # Message
-  cat(paste("Year: ",i,"\n",sep=""))
-  # Computation
-  statcell <- system(paste("r.stats -c for",Year[i],sep=""), intern=TRUE)
-  defor.df$ncells[i] <- as.numeric(strsplit(statcell[1],split=" ")[[1]][2])
-  defor.df$area[i] <- round(defor.df$ncells[i]*(as.numeric(Res)^2)/10000)
-}
-for (i in 2:length(Year)) {
-  defor.df$theta[i] <- round(100*(defor.df$area[i-1]-defor.df$area[i])/defor.df$area[i-1],2)  
-}
-write.table(defor.df,"outputs/defor.txt",row.names=FALSE,sep="\t")
-
[…]

### 2017-02-02 deforestmap.Rmd
Authors
Title

@@
 ---
-title: "Sixty-five years of deforestation and forest degradation in Madagascar"
+title: "Sixty years of deforestation and forest degradation in Madagascar"
 bibliography: bib/biblio.bib
@@
 
-#### Ghislain Vieilledent<sup>*,1,2</sup>, Clovis Grinand<sup>3</sup>, Fety A. Rakotomalala<sup>3</sup>, Jean-Roger Rakotoarijaona<sup>5</sup>, and Frédéric Achard<sup>1</sup>
+#### Ghislain Vieilledent<sup>*,1,2</sup>, Clovis Grinand<sup>3</sup>, Fety A.
+Rakotomalala<sup>3</sup>, Rija Ranaivosoa<sup>5</sup>, Jean-Roger Rakotoarijaona<sup>5</sup>, Thomas F.
+Allnutt<sup>4</sup>, and Frédéric Achard<sup>1</sup>
 

### 2017-02-02 deforestmap.Rmd
Compute fragmentation

@@
 
+Regarding the gains: we did not consider regrowth:
+- few potential regrowth in Madagascar.
+- regrowth are usually more easilly re-burnt for agriculture on the long term.
+- young secondary forests do not provide the same ecosystem services (in terms of biodiversity and carbon storage) as old growth natural forests.
+
 ### 2.2 Validation
@@
 We report an analysis of the evolution of forest fragmentation based on the forest-cover maps for Madagascar for the years 1953, 1973, 1990, 2000, 2010 and 2014.
-We used the approach developed by @Riitters2000 which uses a moving window to characterize the fragmentation around each forested pixel.
+We used the method developed by @Riitters2000 which uses a moving window to characterize the fragmentation around each forested pixel.
+Computation were done using function `r.forestfrag` from GRASS GIS software.
+Six categories of fragmentation are identified from the amount of forest and its occurrence as adjacent forest pixels: interior, perforated, edge, transitional, patch, and undetermined.
 We used a moving window of 3x3 pixels (0.81 km<sup2</sup>).
-Six categories of fragmentation are identified from the amount of forest and its occurrence as adjacent forest pixels: interior, perforated, edge, transitional, patch, and undetermined.
-The interior category can be interpreted as intact forest (see @Potapov2017).
-We reported the area of forest falling in each category for the 6 years and analysed the evolution of the fragmentation.
+Using this window size, forest edge had a width of about 100m.
+The interior category can be interpreted as the most intact forest (sensus @Potapov2017).
+We reported the area of forest in each fragmentation category for the 6 years and analysed the evolution of the fragmentation with time.
 
@@
 
+### 3.1 Validation
+
+### 3.2 Evolution of forest cover, deforestation intensity and fragmentation with time
+
 ## 4. Discussion
+
+### 4.1 Interest of such an approach
+
+- simple
+- easilly reproducible
+- no need to define a tree cover threshold to
 

### 2017-02-02 deforestmap.R

@@ -326,6 +326,22 @@ system("convert -delay 100 -loop 0 outputs/*.gif outputs/defor_Mada.gif")
 #= See also OpenShot app for movie and transition effects
 
 ##==========================
+## Forest fragmentation
+##==========================
+
+Year <- c(1953,1973,1990,2000,2010,2014)
+system("g.region rast=harper -ap")
+system("g.region rast=harper w=340000 e=412000 s=7420000 n=7500000 -ap")
+for (i in 1:length(Year)) {
+  # Message
+  cat(paste("Year: ",Year[i],"\n",sep=""))
+  # Computation
+  system(paste0("r.mapcalc --o 'for",Year[i],"_0 = if(!isnull(for",Year[i],"), 1, 0)'"))
+  system(paste0("r.forestfrag input=for",Year[i],"_0 output=frag",Year[i]," size=7 --overwrite"))
+  system(paste0("r.out.gdal --o input=frag",Year[i]," createopt='compress=lzw,predictor=2' type=Byte output=outputs/frag",Year[i],".tif"))
+}
+
+##==========================
 ## Statistics
 ##==========================
 
@@ -349,6 +365,22 @@ for (i in 2:length(Year)) {
   defor.df$theta[i] <- round(100*(1-(1-theta.prim)^(1/Y)),2)
 }
 write.table(defor.df,"outputs/defor.txt",row.names=FALSE,sep="\t")
+
+#= Fragmentation statistics
+Year <- c(1953,1973,1990,2000,2010,2014)
+frag.df <- data.frame(Year=Year,forest=NA,patch=NA,transitional=NA,
+                      edge=NA,perforated=NA,interior=NA,undetermined=NA)
+# Areas
+for (i in 1:length(Year)) {
+  # Message
+  cat(paste("Year: ",Year[i],"\n",sep=""))
+  # Computation
+  statcell <- system(paste("r.stats -c frag",Year[i],sep=""), intern=TRUE)
+  ncells <- as.numeric(matrix(unlist(strsplit(statcell,split=" ")),ncol=2,byrow=TRUE)[-1,2])
+  frag.df$forest[i] <- round(sum(ncells)*(as.numeric(Res)^2)/10000)
+  frag.df[i,c(3:8)] <- round(100*ncells/sum(ncells),2)
+}
+write.table(frag.df,"outputs/frag.txt",row.names=FALSE,sep="\t")
 
 ##========================
 ## Save objects

### 2017-02-03 deforestmap.R
Validation

@@ -1,5 +1,5 @@
 ##=========================================================================
-## Sixty-five years of deforestation and forest fragmentation in Madagascar
+## Sixty years of deforestation and forest fragmentation in Madagascar
 ## Ghislain Vieilledent <ghislain.vieilledent@cirad.fr>
 ## January 2017
 ##=========================================================================
@@ -80,7 +80,7 @@ system("g.region rast=harper -ap")
 
 #= Typology of CI map
 system("r.info harper")
-system("r.describe harper") 
+system("r.describe harper")
 # * 111 112 115 122 152 155 222 444 555 755 775 777
 # 1=forest, 2=nonforest, 5=cloud, 4=water, 7=mangrove
 
@@ -256,7 +256,7 @@ system("r.stats -c for1990")
 ## * 1241360825
 
 theta.prim <- (158236269-119582639)/158236269
-Y <- 1990-1973 
+Y <- 1990-1973
 theta <- 1-(1-theta.prim)^(1/Y) # 1.6 %.yr-1 (ok, in line with Harper et al. 2007)
 
 #= Export
@@ -303,7 +303,7 @@ system("echo '1 26:152:80' | r.colors for2000 rules=-")
 system("echo '1 26:152:80' | r.colors for2010 rules=-")
 system("echo '1 26:152:80' | r.colors for2014 rules=-")
 
-#= PNG
+# #= PNG
 system("g.region res=1000")
 system("r.out.png input=for1953 output=outputs/for1953.png")
 system("r.out.png input=for1973 output=outputs/for1973.png")
@@ -326,12 +326,209 @@ system("convert -delay 100 -loop 0 outputs/*.gif outputs/defor_Mada.gif")
 #= See also OpenShot app for movie and transition effects
 
 ##==========================
+## Validation
+##==========================
+
+# Import perr-fh map
+system("r.in.gdal --o input=gisdata/rasters/perr-fh/FCC_051013_MG_VF_38s.tif output=perrfh")
+
+# Classification
+system("r.describe perrfh")
+# * 7 111 113 119 133 183 222 223 228 233 288 333 444 533 553 555 666 733 773 777 778 788 888 1000
[…]

### 2017-02-06 deforestmap.Rmd
Update introduction

@@
 Among all the ecosystem services cited previously, Malagasy forests are particularly important for the unique biodiversity they shelter, both in term of number of species and endemism in many taxonomic groups [@Goodman2005].
+Deforestation threatens species survival and diminishes biodiversity by destroying forest habitat, creating forest fragments too small to maintain viable populations and increasing 'edge effects' at forest/non-forest interfaces (Harris 1984).
+Edge effects due to fragmentation typically affect an area several times larger than the forest destruction itself (Harris 1984; Skole & Tucker 1993), affecting micro-meteorology over short distances (Kapos 1989), and increasing exposure to damaging winds (Ferreira & Laurance 1997), fire
+frequency (Cochrane 2001), and access for livestock, other non-forest animals and hunters (Enserink 1999; Cullen et al.
+2000).
+To assess the impact of tropical deforestation on biological diversity, not only the area deforested, but also the isolation of forest patches and the area of edge habitat must be determined.
 
-- Harper,
-- Hansen,
-- Kim,
-- ONE/PERR-FH (http://www.etcterra.org/fr/projets/perrfh)
+Accurate maps of forest cover and forest cover change are necessary to monitore carbon emissions and loss of biodiversity associated to deforestation, assess the efficiency of present conservation strategies [@Eklund2016], and implement new strategies for the future [@Vieilledent2013a, @Vieilledent2016]].
+Unfortunately, accurate and cloud-free forest cover maps are not available for Madagascar for the last decade (2005-2015).
+In their paper untitled _"Fifty years of deforestation and forest degradation in Madagascar"_, @Harper2007 published maps of forest cover change for Madagascar from 1950s to 2000 (for the years _c._ 1953, _c._ 1973, 1990 and 2000).
+The 1950s forest map was derived from the visual interpretation of aerial photographies and gives a coarse representation (scale 1/1,000,000) of the forest extent for Madagascar in the 1950s.
+Forest maps for the years _c._ 1973, 1990 and 2000, were obtained from a supervised classification of Landsat satellite images at 30m and can be used to derive much more accurate estimates of the forest cover at these three dates.
+Notheless, maps provided by @Harper2007 were not free of clouds (11 244 km2 of unknown cover type for the year 2000 for example).
+Using the same supervised classification approach as in @Harper2007, more recent maps have been released for the periods 2000-2005-2010 by national institutions, with the technical support of international environmental NGOs [@MEFT2009, @ONE2013].
+Another recent map using more advanced statistical tool for classification (the Random Forest classifier, see @Grinand2013 and @Rakotomalala2015) has been released for the periods 2005-2010-2013 [@ONE2015].
+However, these maps might (i) be too old to give recent estimates of deforestation [@MEFT2009, @ONE2013], (ii) figure large areas of clouds (images with more than 20% of clouds in 2010 for the map provided by [@ONE2013]), or (iii) show large errors of classification in specific areas, especially in the dry and spiny forests for which the spectral answer varies strongly due to deciduousness associated with the higher seasonality than in moist forests (Kappa<0.6 and overall accuracy<0.8 for the dry and spiny forests in @ONE2015).
+Moreover, deriving such forest maps from supervised classification requires a huge amount of time and work, especially regarding image selection to reduce cloud cover and visual interpretation of the numerous observation polygons used to train the classification algorithm.
+In addition, visual interpretation might be subjective and is potentially a source of errors.
+All this work of image selection and visual interpretation will have to be repeated to produce new forest maps in the future.
+
+Global forest products have also been published recently and could be used to derive recent forest cover map at the national scale for Madagascar.
+@Kim2015 produced a global Landsat-based forest-cover change map from 1990 to 2000 but their approach was not reproduced to obtain more recent maps of forest-cover, after 2000.
+Also, the approach used in @Kim2015 did not allow to correctly map the forest in the dry and spiny ecosystems.
+Another global study is the work by @Hansen2013 who identified annual tree cover, forest loss and forest gain from 2000 to 2012.
+Meanwhile, results have been updated and are available up to the year 2014 at http://earthenginepartners.appspot.com/science-2013-global-forest.
+One problem with using the @Hansen2013 product to map forest cover is that a tree cover threshold must be defined to transform the tree cover data into a forest/non-forest map.
+Identifying such threshold is not straightforward as forests cannot be defined only by a tree cover value.
+Also, the tree cover threshold might strongly vary between forest types, mean tree cover values being substantially lower for dry forests than for moist forests.
+Also, a tree cover gives no information on land-use.
+Therefore, tree plantations such as oil palm or eucalyptus plantations cannot be differenciated from natural forests, which provide very different ecosystem services particularly in terms of biodiversity [@Tropek2014].
+Thus, using @Hansen2013 alone can thus only give a first-approach estimate of the forest cover.
+But, tree cover is not the only information provided by @Hansen2013.
+The forest loss, i.e.
+the complete loss of forest tree cover (which shifts from a positive value over 10% to zero) is also reported.
+
+In this study, we present a simple approach combining maps from @Harper2007 and products from @Hansen2013 to derive annual and cloud-free forest-cover change map for Madagascar for the period 2000-2014.
+Basically, we used the forest map provided by @Harper2007 for the year 2000 (defining the land-use) with the tree cover loss product provided by @Hansen2013 that we apllied only in forest areas.
+Following the work done by Harper2007, we studied in details the change in deforestation rates and the evolution of forest fragmentation on six decades from _c._ 1953 to 2014.
+We showed that despite all the conservation efforts done in Madagascar since the mid 1980s, deforestation has increased in speed since 2010 and that fragmentation has dramatically increased leading to few remaining intact forest landscapes.
+The approach described in this study could be easilly extended to all tropical countries with a forest map obtained at one date between 2000 and 2014, and could be repeated in the future to obtain future forest-cover change maps, as and when the @Hansen2013 product is updated.
 
@@
 - young secondary forests do not provide the same ecosystem services (in terms of biodiversity and carbon storage) as old growth natural forests.
+- Harper intro: "we saw little evidence of forest regeneration in Madagascar".
 
@@
 
-We estimated the accuracy of our approach combining Harper's and Hansen's products to derive cloud-free forest maps after the year 2000 for Madagascar.
-To do so, we used a deforestation map produced on the period 2005-2013 by a supervised classification of Landsat images at multiple dates using the Random Forest algorithm [@Grinand2013, @Rakotomalala2015].
-Following our methodology, we produced a deforestation map on the same period of time (2005-2013) and compute several accuracy indices (overall accuracy, Cohen's Kappa and True Skill Statistics) on the basis of a confusion matrix computed from the two deforestation maps at dates 2005 and 2013.
+We estimated the accuracy of our approach combining Harper's and Hansen's products to derive cloud-free forest-cover change maps after the year 2000 in Madagascar.
+To do so, we used a forest-cover change map obtained from a regional study on the humid ecoregion on the period 2005-2013 [@Rakotomalala2015].
+@Rakotomalala2015 performed a supervised classification of Landsat images at multiple dates using the Random Forest classifier [@Grinand2013].
+The map by @Rakotomalala2015 is considered rather accurate on the moist ecoregion with an external validation on nearly 10,854 visual interpretation points leading to rather high Kappa and overall-accuracy values (0.79 and 0.89 respectively).
+Following our methodology, we produced a deforestation map on the same period of time (2005-2013) and compared the forest-cover (in ha) and mean annual deforestation rates (in ha and % per year) between the two maps.
 
@@
 
-We report an analysis of the evolution of forest fragmentation based on the forest-cover maps for Madagascar for the years 1953, 1973, 1990, 2000, 2010 and 2014.
+We report an analysis of the evolution of forest fragmentation based on the forest-cover maps for Madagascar for the years _c._ 1953, _c._ 1973, 1990, 2000, 2010 and 2014.
 We used the method developed by @Riitters2000 which uses a moving window to characterize the fragmentation around each forested pixel.
-Computation were done using function `r.forestfrag` from GRASS GIS software.
+Computation were done using function `r.forestfrag` from GRASS GIS software [@Neteler2008].
 Six categories of fragmentation are identified from the amount of forest and its occurrence as adjacent forest pixels: interior, perforated, edge, transitional, patch, and undetermined.
-We used a moving window of 3x3 pixels (0.81 km<sup2</sup>).
-Using this window size, forest edge had a width of about 100m.
+We used a moving window of 7x7 pixels (4.41 km<sup>2</sup>).
+Using this window size, forest edge had a width of about 100 m.
 The interior category can be interpreted as the most intact forest (sensus @Potapov2017).
-We reported the area of forest in each fragmentation category for the 6 years and analysed the evolution of the fragmentation with time.
+We reported the area of forest in each fragmentation category for the six years and analysed the evolution of the fragmentation with time.
 
@@
 
-### 4.1 Interest of such an approach
+### 4.1 Evolution of deforestation and forest fragmentation on the last 60 years in Madagascar
+
+### 4.2 Interest of such an approach and limitations
 
@@
 - no need to define a tree cover threshold to
+
+## 5. Tables
+
+## 6. Figure captions
+
+## 7. Figures
 
@@
 
-## Supplementary materials
-
 ## References

### 2017-02-06 deforestmap.Rmd
Add tables

@@
 ---
-title: "Sixty years of deforestation and forest degradation in Madagascar"
+title: "Six decades of deforestation and forest fragmentation in Madagascar"
 bibliography: bib/biblio.bib
@@
 Unfortunately, accurate and cloud-free forest cover maps are not available for Madagascar for the last decade (2005-2015).
-In their paper untitled _"Fifty years of deforestation and forest degradation in Madagascar"_, @Harper2007 published maps of forest cover change for Madagascar from 1950s to 2000 (for the years _c._ 1953, _c._ 1973, 1990 and 2000).
-The 1950s forest map was derived from the visual interpretation of aerial photographies and gives a coarse representation (scale 1/1,000,000) of the forest extent for Madagascar in the 1950s.
+In their paper untitled _"Fifty years of deforestation and forest fragmentation in Madagascar"_, @Harper2007 published maps of forest cover change for Madagascar from 1950s to 2000 (for the years _c._ 1953, _c._ 1973, 1990 and 2000).
+The 1950s forest map was derived from the visual interpretation of aerial photographies and gives a coarse representation (scale 1/1,000,000) of the forest extent in Madagascar for the 1950s.
 Forest maps for the years _c._ 1973, 1990 and 2000, were obtained from a supervised classification of Landsat satellite images at 30m and can be used to derive much more accurate estimates of the forest cover at these three dates.
@@
 Global forest products have also been published recently and could be used to derive recent forest cover map at the national scale for Madagascar.
-@Kim2015 produced a global Landsat-based forest-cover change map from 1990 to 2000 but their approach was not reproduced to obtain more recent maps of forest-cover, after 2000.
-Also, the approach used in @Kim2015 did not allow to correctly map the forest in the dry and spiny ecosystems.
+@Kim2014 produced a global Landsat-based forest-cover change map from 1990 to 2000 but their approach was not reproduced to obtain more recent maps of forest-cover, after 2000.
+Also, the approach used in @Kim2014 did not allow to correctly map the forest in the dry and spiny ecosystems.
 Another global study is the work by @Hansen2013 who identified annual tree cover, forest loss and forest gain from 2000 to 2012.
@@
 
-### 3.2 Evolution of forest cover, deforestation intensity and fragmentation with time
+Difference might be due to post-processing of @Rakotomalala2015 map.
+
+### 3.2 Evolution of forest cover and deforestation intensity
+
+### 3.3 Evolution of forest fragmentation with time
 
@@
 
-### 4.1 Evolution of deforestation and forest fragmentation on the last 60 years in Madagascar
+### 4.1 Evolution of deforestation and forest fragmentation in Madagascar on the last 60 years
 
@@
 ## 5. Tables
+
+```{r comp, results="asis"}
+tab.1 <- comp_moist.df
+mean.diff <- round(100*(sum(tab.1$BSM)-sum(tab.1$PERR_FH))/sum(tab.1$PERR_FH))
+names(tab.1) <- c("Year","Vieil (ha)", "Rakot (ha)", "Rakot_NA (ha)", "ad_Vieil (ha/yr)", "ad_Rakot (ha/yr)", "theta_Vieil (%/yr)", "theta_Rakoto (%/yr)", "diff.area (%)")
+tab.1 <- tab.1[,c(1:4,9,5:8)]
+kable(tab.1)
+```
+
+Table 1: **Comparison of the forest cover and deforestation rates between our study and the study by @Rakotomalala2015**.
+This comparison is done only for the humid ecoregion for the periods 2005-2010-2013, which is the extent and time-periods of the @Rakotomalala2015 map.
+@Rakotomalala2015 map has been validated using 10,854 independant visual interpretation points leading to a rather high accuracy (Kappa=0.79 and overall-accuracy=0.89).
+Our study (suffix _"Vieil"_) estimated a forest cover greater by about `r mean.diff`% compared to @Rakotomalala2015 (suffix _"Rakot"_).
+But the difference (_"diff.area"_) did not vary much between dates (between 3.56 and 6.18%).
+Our study also estimated higher annual deforestation (_"ad"_) and deforestation rates (_"theta"_) but the same trend for the deforestation was observed between the two studies, with a doubling of the deforestation for the period 2010-2013 compared to 2005-2010.
+The differences between the two studies could be attributed to the forest definition which might be different in the two studies and to the post-processing in @Rakotomalala2015 map.
+
+```{r defor, results="asis"}
+tab.2 <- defor.df[,-c(2)]
+kable(tab.2)
+```
+
+Table 2: **Evolution of the forest cover and deforestation intensity from 1950s to 2014 in Madagascar**.
 

### 2017-02-06 deforestmap.R

@@ -1,5 +1,5 @@
 ##=========================================================================
-## Sixty years of deforestation and forest fragmentation in Madagascar
+## Six decades of deforestation and forest fragmentation in Madagascar
 ## Ghislain Vieilledent <ghislain.vieilledent@cirad.fr>
 ## January 2017
 ##=========================================================================

### 2017-02-07 deforestmap.Rmd
Tables

@@
 We used a moving window of 7x7 pixels (4.41 km<sup>2</sup>).
-Using this window size, forest edge had a width of about 100 m.
+Using this window size, forest edge had a width of about 90 m [@Riitters2000].
 The interior category can be interpreted as the most intact forest (sensus @Potapov2017).
@@
 ```{r defor, results="asis"}
-tab.2 <- defor.df[,-c(2)]
+tab.2 <- defor.df
+names(tab.2) <- c("Year","Forest (ha)", "Annual defor. (ha/yr)", "theta (%/yr)")
 kable(tab.2)
@@
 Table 2: **Evolution of the forest cover and deforestation intensity from 1950s to 2014 in Madagascar**.
+Madagascar has lost 37% of its natural forest in 40 years between _c._ 1973 and 2014 (and roughly 56% in 60 years since _c._ 1953).
+If the deforestation has progressively slowed down from _c._ 1953 to 2010 (from 1.76 to 0.59%/yr), most likely due to a decrease of available arable lands under the remaining forests and to a decrease of forest accessibility, the deforestation has recently increased, with a doubling deforestation on the period 2010-2014 compared to 2000-2010.
+This increase could be attributed to both political instability (political crisis in 2009) and to an explosion of the demography (according to the @UN2015, population has increased from about 16M inhabitants in 2000 to 24M in 2014).
+Natural forest cover 8.9 Mha in 2014 and current (period 2000-2014) mean annual deforestation rate is of about 70,000 ha/yr.
+
+```{r frag, results="asis"}
+tab.3 <- frag.df
+names(tab.3) <- c("Year","Forest (ha)", "patch (%)", "transitional (%)", "edge (%)", "perforated (%)", "interior (%)", "NA (%)")
+kable(tab.3)
+```
+
+Table 3: **Evolution of the forest fragmentation from 1950s to 2014 in Madagascar**.
+Forest fragmentation has progressively increased from _c._ 1953.
+The percentage of forest belonging to the interior category (the most intact forest) has fallen from 94 in _c._ 1973 to 48% in 2014.
+More than 15% of the forest belong to the patch and transitional categories (isolated small forest patches) in 2014.
+This increase in forest fragmentation has a great impact on both carbon emissions [@Putz2014] and biodiversity loss [@Gibson2013, @Murcia1995].
 

### 2017-02-07 deforestmap.R

@@ -533,7 +533,12 @@ for (i in 1:length(Year)) {
   # Message
   cat(paste("Year: ",Year[i],"\n",sep=""))
   # Computation
-  system(paste0("r.mapcalc --o 'for",Year[i],"_0 = if(!isnull(for",Year[i],"), 1, 0)'"))
+  if (Year[i]==1973) {  # ! 1975 with 5 (clouds) 
+    system(paste0("r.mapcalc --o 'for1973_0 = if(!isnull(for1973) && for1973==1, 1, 0)'"))
+  }
+  else {
+    system(paste0("r.mapcalc --o 'for",Year[i],"_0 = if(!isnull(for",Year[i],"), 1, 0)'"))
+  }
   system(paste0("r.forestfrag input=for",Year[i],"_0 output=frag",Year[i]," size=7 --overwrite"))
   system(paste0("r.out.gdal --o input=frag",Year[i]," createopt='compress=lzw,predictor=2' type=Byte output=outputs/frag",Year[i],".tif"))
 }
@@ -544,15 +549,15 @@ for (i in 1:length(Year)) {
 
 #= Deforestation statistics
 Year <- c(1953,1973,1990,2000,2010,2014)
-defor.df <- data.frame(Year=Year,ncells=NA,area=NA,ann.defor=NA,theta=NA)
+defor.df <- data.frame(Year=Year,area=NA,ann.defor=NA,theta=NA)
 # Areas
 for (i in 1:length(Year)) {
   # Message
   cat(paste("Year: ",Year[i],"\n",sep=""))
   # Computation
   statcell <- system(paste0("r.stats -c for",Year[i]), intern=TRUE)
-  defor.df$ncells[i] <- as.numeric(strsplit(statcell[1],split=" ")[[1]][2])
-  defor.df$area[i] <- round(defor.df$ncells[i]*(as.numeric(Res)^2)/10000)
+  ncells <- as.numeric(strsplit(statcell[1],split=" ")[[1]][2])
+  defor.df$area[i] <- round(ncells*(as.numeric(Res)^2)/10000)
 }
 # Annual rates
 for (i in 2:length(Year)) {
@@ -573,7 +578,7 @@ for (i in 1:length(Year)) {
   cat(paste("Year: ",Year[i],"\n",sep=""))
   # Computation
   statcell <- system(paste0("r.stats -c frag",Year[i]), intern=TRUE)
[…]

### 2017-02-08 deforestmap.R
Plot fcc map

@@ -585,6 +585,59 @@ for (i in 1:length(Year)) {
 write.table(frag.df,"outputs/frag.txt",row.names=FALSE,sep="\t")
 
 ##========================
+## Forest-cover change map
+##========================
+
+## Raster of water-bodies over Madagascar
+# Region
+Extent <- "298440 7155900 1100820 8682420"
+Res <- "30"
+proj.s <- "EPSG:4326"
+proj.t <- "EPSG:32738"
+Input <- "outputs/water.vrt"
+Output <- "outputs/water.tif"
+# gdalbuildvrt
+system("gdalbuildvrt -overwrite outputs/water.vrt gisdata/rasters/waterbodies/*.tif")
+# gdalwarp
+system(paste0("gdalwarp -overwrite -s_srs ",proj.s," -t_srs ",proj.t," -te ",Extent,
+              " -tr ",Res," ",Res," -r near -ot Byte -co 'COMPRESS=LZW' -co 'PREDICTOR=2' ",Input," ",Output))
+# Import
+system("r.in.gdal --o input=outputs/water.tif output=water")
+
+## FCC map
+
+# Mask on Harper map
+system("r.null map=harper setnull=0 --verbose")
+system("r.mask --o raster=harper")
+# Compute fcc
+system(paste0("r.mapcalc --o 'fcc = if(!isnull(for2014),24,if(!isnull(for2000),23,if(!isnull(for1990),22, \\
+       if(!isnull(for1973) &&& for1973==1,21,if(!isnull(water) &&& water>0,water, \\
+              if(!isnull(for1953),20,if(!isnull(for1973) &&& for1973==5,215,0)))))))'"))
+# Color palette
+system("r.colors map=fcc rules=- << EOF
+0 243:243:220
+1 153:217:234
+12 0:0:170
+20 0:0:0
+21 255:255:0
+215 184:184:184
+22 255:165:0
+23 255:0:0
+24 34:139:34
+nv 255:255:255
+EOF")
+# Export
+system("r.out.gdal --o input=fcc createopt='compress=lzw,predictor=2' type=Byte output=outputs/fcc.tif")
+fcc <- raster("outputs/fcc.tif")
+plot(fcc)
+for1953 <- raster("outputs/for1953.tif")
+plot(for1953)
+
+water <- raster("outputs/water.tif")
+plot(water)
+plot(MASK)
+
+##========================
 ## Save objects
 ##========================
 

### 2017-02-10 deforestmap.R
Map fcc_for1953

@@ -277,14 +277,13 @@ Output <- "outputs/harper1953.tif"
 system(paste0("gdalwarp -overwrite -dstnodata ",nodat," -s_srs ",proj.s," -t_srs ",proj.t," -te ",Extent,
               " -tr ",Res," ",Res," -r near -co 'COMPRESS=LZW' -co 'PREDICTOR=2' ",Input," ",Output))
 
-#= Import for1953_orig_30m.tif into grass location
+#= Import harper1953.tif into grass location
 system("r.in.gdal --o input=outputs/harper1953.tif output=harper1953")
 system("r.info harper1953")
 system("r.stats -c harper1953")
 
-#= Corrections from for1973
-system("r.mapcalc --o 'for1953 = if(!isnull(for1973) &&& (for1973==1 || for1973==7), 1, harper1953)'")
-system("r.mapcalc --o 'for1953 = if(for1953!=1,null(),1)'")
+#= Compute for1933
+system("r.mapcalc --o 'for1953 = if(harper1953!=1,null(),1)'")
 system("r.stats -c for1953")
 ## 1 225792265
 ## * 1135151199
@@ -613,14 +612,26 @@ system("r.mask --o raster=harper")
 system(paste0("r.mapcalc --o 'fcc = if(!isnull(for2014),24,if(!isnull(for2000),23,if(!isnull(for1990),22, \\
        if(!isnull(for1973) &&& for1973==1,21,if(!isnull(water) &&& water>0,water, \\
               if(!isnull(for1953),20,if(!isnull(for1973) &&& for1973==5,215,0)))))))'"))
+# fcc_for1953
+system(paste0("r.mapcalc --o 'fcc_for1953 = if(!isnull(for1953),20,0)'"))
 # Color palette
+# 0 243:243:220
+# 1 153:217:234
+# 12 0:0:170
+# 20 0:0:0
+# 21 255:255:0
+# 215 184:184:184
+# 22 255:165:0
+# 23 255:0:0
+# 24 34:139:34
+# nv 255:255:255
 system("r.colors map=fcc rules=- << EOF
 0 243:243:220
 1 153:217:234
 12 0:0:170
-20 0:0:0
-21 255:255:0
-215 184:184:184
+20 243:243:220
+21 255:255:102
+215 243:243:220
 22 255:165:0
 23 255:0:0
 24 34:139:34
@@ -628,14 +639,46 @@ nv 255:255:255
 EOF")
 # Export
[…]

### 2017-02-10 deforestmap.R
Nice fcc map

@@ -19,7 +19,7 @@ if (length(arg)>0) {
 }
 
 ##= Libraries
-pkg <- c("broom","sp","rgdal","raster","ggplot2","gridExtra",
+pkg <- c("broom","sp","rgdal","raster","ggplot2","gridExtra","RColorBrewer",
          "rasterVis","knitr","rmarkdown","rgeos","rgdal","rgrass7")
 ## broom: to convert map into data-frame with tiny()
 ## gridExtra: to combine several ggplots
@@ -663,22 +663,41 @@ theme_base <- theme(axis.line=element_blank(),
                     panel.grid.minor=element_blank(),
                     panel.border=element_blank())
 
-## colors
-forest.col = rgb(34/255,139/255,34/255)
-nonforest.col = rgb(243/255,243/255,220/255)
-
+## Colors
+# Discrete values: 0,1:12,20,21,215,22,23,24,215
+for2014.col = rgb(34/255,139/255,34/255) # forest green
+nofor73.col = rgb(243/255,243/255,220/255) # light yellow
+defor73_90.col = rgb(255/255,255/255,102/255) # yellow-orange
+defor90_00.col = rgb(255/255,165/255,0/255) # orange
+defor00_14.col = rgb(255/255,0/255,0/255) # red
+w1.col = rgb(153/255,217/255,234/255)
+w2.col = rgb(0/255,0/255,170/255)
+water.col <-colorRampPalette(c(w1.col,w2.col))(12)
+# Building palette
+fcc.col <- c(nofor73.col,water.col,nofor73.col,
+              defor73_90.col,nofor73.col,defor90_00.col,defor00_14.col,for2014.col)
+fcc.name <- c(0,1:12,20,21,215,22,23,24)
+fcc.palette <- paste0("c(",paste0("\"",fcc.name,"\"","=","\"",fcc.col,"\"",collapse=","),")")
+fcc.palette <- eval(parse(text=fcc.palette))
 
 ## Forest in 1953 on a separate plot
 for1953 <- raster("outputs/fcc_for1953.tif")
 plot.for1953 <- gplot(for1953,maxpixels=res.rast) +
   geom_raster(aes(fill=factor(value))) +
-  scale_fill_manual(values=c(nonforest.col,green.col),na.value="transparent") +
+  scale_fill_manual(values=c("0"=nonforest.col,"20"=green.col),na.value="transparent") +
[…]

### 2017-02-22 deforestmap.Rmd
Add figure

@@
 ---
-title: "Six decades of deforestation and forest fragmentation in Madagascar"
+title: "Combining global tree cover loss data with historical forest-cover maps to look at six decades of deforestation and forest fragmentation in Madagascar"
 bibliography: bib/biblio.bib
@@
 
-Tropical forests provide important ecosystem services at the global scale, such as biodiversity conservation and climate regulation [@Costanza1997] but also many at the regional scale for local populations.
+Tropical forests provide important ecosystem services at the global scale, such as biodiversity conservation and climate regulation [@Costanza1997] but also many at the local scale for indigeneous populations.
 At the local scale, tropical forests supply clean water, can preserve populations from flood [@Bradshaw2007] and provide wood for construction and fuel and other non-woody products for food and health.
@@
 Deforestation threatens species survival and diminishes biodiversity by destroying forest habitat, creating forest fragments too small to maintain viable populations and increasing 'edge effects' at forest/non-forest interfaces (Harris 1984).
-Edge effects due to fragmentation typically affect an area several times larger than the forest destruction itself (Harris 1984; Skole & Tucker 1993), affecting micro-meteorology over short distances (Kapos 1989), and increasing exposure to damaging winds (Ferreira & Laurance 1997), fire
-frequency (Cochrane 2001), and access for livestock, other non-forest animals and hunters (Enserink 1999; Cullen et al.
+Edge effects due to fragmentation typically affect an area several times larger than the forest destruction itself (Harris 1984; Skole & Tucker 1993), affecting micro-meteorology over short distances (Kapos 1989), and increasing exposure to damaging winds (Ferreira & Laurance 1997), fire frequency (Cochrane 2001), and access for livestock, other non-forest animals and hunters (Enserink 1999; Cullen et al.
 2000).
@@
 
-Accurate maps of forest cover and forest cover change are necessary to monitore carbon emissions and loss of biodiversity associated to deforestation, assess the efficiency of present conservation strategies [@Eklund2016], and implement new strategies for the future [@Vieilledent2013a, @Vieilledent2016]].
-Unfortunately, accurate and cloud-free forest cover maps are not available for Madagascar for the last decade (2005-2015).
+Accurate maps of forest-cover and forest-cover change are necessary to monitore carbon emissions and loss of biodiversity associated to deforestation, assess the efficiency of present conservation strategies [@Eklund2016], and implement new strategies for the future [@Vieilledent2013a, @Vieilledent2016].
+Unfortunately, accurate and cloud-free forest-cover maps are not available for Madagascar for the last decade (2005-2015).
 In their paper untitled _"Fifty years of deforestation and forest fragmentation in Madagascar"_, @Harper2007 published maps of forest cover change for Madagascar from 1950s to 2000 (for the years _c._ 1953, _c._ 1973, 1990 and 2000).
-The 1950s forest map was derived from the visual interpretation of aerial photographies and gives a coarse representation (scale 1/1,000,000) of the forest extent in Madagascar for the 1950s.
+The _c._ 1953 forest map was derived from the visual interpretation of aerial photographies and gives a coarse representation (scale 1/1,000,000) of the forest extent in Madagascar for the 1950s.
 Forest maps for the years _c._ 1973, 1990 and 2000, were obtained from a supervised classification of Landsat satellite images at 30m and can be used to derive much more accurate estimates of the forest cover at these three dates.
-Notheless, maps provided by @Harper2007 were not free of clouds (11 244 km2 of unknown cover type for the year 2000 for example).
+Maps provided by @Harper2007 were not free of clouds (11 244 km2 of unknown cover type for the year 2000 for example).
 Using the same supervised classification approach as in @Harper2007, more recent maps have been released for the periods 2000-2005-2010 by national institutions, with the technical support of international environmental NGOs [@MEFT2009, @ONE2013].
 Another recent map using more advanced statistical tool for classification (the Random Forest classifier, see @Grinand2013 and @Rakotomalala2015) has been released for the periods 2005-2010-2013 [@ONE2015].
-However, these maps might (i) be too old to give recent estimates of deforestation [@MEFT2009, @ONE2013], (ii) figure large areas of clouds (images with more than 20% of clouds in 2010 for the map provided by [@ONE2013]), or (iii) show large errors of classification in specific areas, especially in the dry and spiny forests for which the spectral answer varies strongly due to deciduousness associated with the higher seasonality than in moist forests (Kappa<0.6 and overall accuracy<0.8 for the dry and spiny forests in @ONE2015).
+However, these maps might (i) be too old to give recent estimates of deforestation [@MEFT2009, @ONE2013], (ii) figure large areas of clouds (images with more than 20% of clouds in 2010 for the map provided by [@ONE2013]), or (iii) show large errors of classification in specific areas, especially in the dry and spiny forests for which the spectral answer varies strongly due to deciduousness associated with the higher seasonality than in moist forests (Kappa<0.6 and overall accuracy<0.8 for the dry and spiny forests for the maps produced by @ONE2015).
 Moreover, deriving such forest maps from supervised classification requires a huge amount of time and work, especially regarding image selection to reduce cloud cover and visual interpretation of the numerous observation polygons used to train the classification algorithm.
 In addition, visual interpretation might be subjective and is potentially a source of errors.
-All this work of image selection and visual interpretation will have to be repeated to produce new forest maps in the future.
+All this work of image selection and visual interpretation have to be repeated to produce new forest maps in the future.
 
@@
 Also, the tree cover threshold might strongly vary between forest types, mean tree cover values being substantially lower for dry forests than for moist forests.
-Also, a tree cover gives no information on land-use.
+Moreover, a tree cover gives no information on land-use.
 Therefore, tree plantations such as oil palm or eucalyptus plantations cannot be differenciated from natural forests, which provide very different ecosystem services particularly in terms of biodiversity [@Tropek2014].
-Thus, using @Hansen2013 alone can thus only give a first-approach estimate of the forest cover.
+Thus, using @Hansen2013 alone can only give a first-approach estimate of the forest cover.
 But, tree cover is not the only information provided by @Hansen2013.
@@
 In this study, we present a simple approach combining maps from @Harper2007 and products from @Hansen2013 to derive annual and cloud-free forest-cover change map for Madagascar for the period 2000-2014.
-Basically, we used the forest map provided by @Harper2007 for the year 2000 (defining the land-use) with the tree cover loss product provided by @Hansen2013 that we apllied only in forest areas.
-Following the work done by Harper2007, we studied in details the change in deforestation rates and the evolution of forest fragmentation on six decades from _c._ 1953 to 2014.
+Basically, we used the forest map provided by @Harper2007 for the year 2000 (defining the land-use) with the tree cover loss product provided by @Hansen2013 that we applied only inside forest areas identified by @Harper2007.
+Following the work done by @Harper2007, we studied in details the change in deforestation rates and the evolution of forest fragmentation on six decades from _c._ 1953 to 2014.
 We showed that despite all the conservation efforts done in Madagascar since the mid 1980s, deforestation has increased in speed since 2010 and that fragmentation has dramatically increased leading to few remaining intact forest landscapes.
@@
 On Harper's map, ~200,000 ha clouds are present over the ~4.2 million ha moist forest ecoregion (4.8% of clouds).
-To remove these clouds, we used the 2000 cloud-free tree percent cover map provided by @Hansen2013 (also at 30~m resolution) and we chose a threshold of 75% of tree cover to decide whether to replace cloud pixels with forest or non-forest pixels [@Achard2014].
+To remove these clouds, we used the 2000 cloud-free tree percent cover map provided by @Hansen2013 (also at 30~m resolution) and we chose a threshold of 75% of tree cover [@Achard2014] to decide whether to replace cloud pixels with forest or non-forest pixels.
 We thus obtained a cloud-free year 2000 forest map for Madagascar.
-From this map, using tree cover loss data from 2000 to 2014 by @Hansen2013, we obtained cloud-free forest maps in 2010 and 2014 at 30~m resolution.
+From this map, using tree cover loss data from 2000 to 2014 provided by @Hansen2013, we obtained cloud-free forest maps in 2010 and 2014 at 30~m resolution.
 Finally, we corrected the cloudy 1990 Harper forest map assuming that if forest was present in 2000, it was also present in 1990.
@@
 These maps are freely available at http:/bioscenemada.cirad.fr/forestmaps.
-On the basis of the forest maps, we computed the mean annual deforestation intensity (in hectares of forest and in percentage) for the five period of time from 1990 to 2014.
+On the basis of the forest-cover maps, we computed the mean annual deforestation intensity (in hectares of forest and in percentage) on six decades for 5 time periods from _c._ 1953 to 2014 (_c._ 1953-_c._ 1973-1990-2000-2010-2014).
 
-Regarding the gains: we did not consider regrowth:
-- few potential regrowth in Madagascar.
-- regrowth are usually more easilly re-burnt for agriculture on the long term.
-- young secondary forests do not provide the same ecosystem services (in terms of biodiversity and carbon storage) as old growth natural forests.
-- Harper intro: "we saw little evidence of forest regeneration in Madagascar".
-
-### 2.2 Validation
+### 2.2 Comparison with other historical deforestation maps
 
@@
 ### 4.1 Evolution of deforestation and forest fragmentation in Madagascar on the last 60 years
+
+Regarding the potential forest regrowth, we made the same choice as @Harper2007 and did not consider it in our study, although @Hansen2013 are also providing tree cover gain data over the period 2010-2014.
+We made this choice for several reasons.
+First, there is little evidence of natural forest regeneration in Madagascar on the long term [@Harper2007, @Grouzis2001].
+This can be explained by several ecological mechanisms following burning practice such as soil erosion [@Grinand2017] and reduced seed bank due to fire and soil loss [@Grouzis2001].
+Moreover, in the areas where forest regeneration is ecologically possible, young regrowth forests are more easilly re-burnt for agriculture and pasture.
+Finally, young secondary forests do not provide the same ecosystem services (in terms of biodiversity and carbon storage) as old-growth natural forests.
 
@@
 
-## 6. Figure captions
+## 6. Figures
 
-## 7. Figures
+![Figure 1: **Forest-cover change on six decades from _c._ 1953 to 2014 in Madagascar.** Forest cover changes from _c._ 1973 to 2014 are shown in the main figure, and forest cover in _c._ 1953 is shown in the lower-right inset.
+Two zooms in the western dry (left) and eastern moist (right) regions present more detailed views of ](outputs/fig_fcc.png)
 
@@
 
-This study is part of the BioSceneMada project (http://bioscenemada.cirad.fr) and the ReCaREDD project (http://roadlessforests.org).
+This study is part of the BioSceneMada project (https://bioscenemada.cirad.fr) and the ReCaREDD project (http://roadlessforests.org).
 The BioSceneMada project is funded by FRB (Fondation pour la Recherche sur la Biodiversité) and the FFEM (Fond Français pour l'Environnement Mondial) under the project agreement AAP-SCEN-2013 I.

### 2017-02-22 deforestmap.Rmd
Update fcc figure

@@
 ---
-title: "Combining global tree cover loss data with historical forest-cover maps to look at six decades of deforestation and forest fragmentation in Madagascar"
+title: "Combining global tree cover loss data with national historical forest-cover maps to look at six decades of deforestation and forest fragmentation in Madagascar"
 bibliography: bib/biblio.bib
@@
 
-![Figure 1: **Forest-cover change on six decades from _c._ 1953 to 2014 in Madagascar.** Forest cover changes from _c._ 1973 to 2014 are shown in the main figure, and forest cover in _c._ 1953 is shown in the lower-right inset.
-Two zooms in the western dry (left) and eastern moist (right) regions present more detailed views of ](outputs/fig_fcc.png)
+![Figure 1: **Forest-cover change on six decades from _c._ 1953 to 2014 in Madagascar.** Forest cover changes from _c._ 1973 to 2014 are shown in the main figure, and forest cover in _c._ 1953 is shown in the bottom-right inset.
+Two zooms in the western dry (left part) and eastern moist (right part) ecoregions present more detailed views of (from top to bottom): forest-cover in 1950s, forest-cover change from _c._ 1973 to 2014, forest fragmentation in 2014 and distance to forest edge in 2014.
+Data on water bodies (blue) and water seasonality (light blue for seasonal water to dark blue for permanent water) has been extracted from @Pekel2016.](outputs/fig_fcc.png)
 

### 2017-02-22 deforestmap.R

@@ -552,6 +552,25 @@ for (i in 1:length(Year)) {
 system("r.mask -r")
 
 ##==========================
+## Distance to forest edge
+##==========================
+
+Year <- c(1953,1973,1990,2000,2010,2014)
+system("g.region rast=harper -ap")
+for (i in 1:length(Year)) {
+  i=1
+  # Message
+  cat(paste("Year: ",Year[i],"\n",sep=""))
+  # Export
+  system(paste0("r.out.gdal --o input=for",Year[i],"_0 createopt='compress=lzw,predictor=2' type=Byte output=outputs/for",Year[i],"_0.tif"))
+  # Computation
+  system(paste0("gdal_proximity.py outputs/for",Year[i],"_0.tif outputs/dist_edge_",Year[i],".tif -co 'COMPRESS=LZW' -co 'PREDICTOR=2' \
+    -values 0 -ot UInt32 -distunits GEO"))
+  system("gdal_translate -a_nodata 0 -co 'COMPRESS=LZW' -co 'PREDICTOR=2' outputs/_dist_edge_",Year[i],".tif outputs/dist_edge_",Year[i],".tif")
+  system(paste0("r.in.gdal --o input=outputs/dist_edge_",Year[i],".tif output=dist_edge_",Year[i],".tif"))
+}
+
+##==========================
 ## Statistics
 ##==========================
 
@@ -656,16 +675,6 @@ system("r.mask -r")
 ## =======================================
 ## Plot raster with gplot() from rasterVis
 source(file="R/plotfcc.R")
-
-
-
-
-
-## FCC
-
-plot.fcc
-ggsave(filename="outputs/fcc.png",plot=plot.fcc,
-       width=15,height=10,unit="cm")
 
 
 ##========================

### 2017-02-23 deforestmap.R
Correct dist to forest edge

@@ -533,6 +533,9 @@ system("r.mask -r")
 ## Forest fragmentation
 ##==========================
 
+## The GRASS Add-on r.forestfrag must be installed
+## https://grass.osgeo.org/grass72/manuals/addons/r.forestfrag.html
+
 Year <- c(1953,1973,1990,2000,2010,2014)
 system("g.region rast=harper -ap")
 system("r.mask --o raster=harper")
@@ -541,7 +544,7 @@ for (i in 1:length(Year)) {
   cat(paste("Year: ",Year[i],"\n",sep=""))
   # Computation
   if (Year[i]==1973) {  # ! 1975 with 5 (clouds) 
-    system(paste0("r.mapcalc --o 'for1973_0 = if(!isnull(for1973) && for1973==1, 1, 0)'"))
+    system(paste0("r.mapcalc --o 'for1973_0 = if(!isnull(for1973) &&& for1973==1, 1, 0)'"))
   }
   else {
     system(paste0("r.mapcalc --o 'for",Year[i],"_0 = if(!isnull(for",Year[i],"), 1, 0)'"))
@@ -558,15 +561,14 @@ system("r.mask -r")
 Year <- c(1953,1973,1990,2000,2010,2014)
 system("g.region rast=harper -ap")
 for (i in 1:length(Year)) {
-  i=1
   # Message
   cat(paste("Year: ",Year[i],"\n",sep=""))
-  # Export
-  system(paste0("r.out.gdal --o input=for",Year[i],"_0 createopt='compress=lzw,predictor=2' type=Byte output=outputs/for",Year[i],"_0.tif"))
   # Computation
-  system(paste0("gdal_proximity.py outputs/for",Year[i],"_0.tif outputs/dist_edge_",Year[i],".tif -co 'COMPRESS=LZW' -co 'PREDICTOR=2' \
-    -values 0 -ot UInt32 -distunits GEO"))
-  system("gdal_translate -a_nodata 0 -co 'COMPRESS=LZW' -co 'PREDICTOR=2' outputs/_dist_edge_",Year[i],".tif outputs/dist_edge_",Year[i],".tif")
+  system(paste0("gdal_proximity.py outputs/for",Year[i],".tif outputs/_dist_edge_",Year[i],".tif \\
+                -co 'COMPRESS=LZW' -co 'PREDICTOR=2' -values 255 -ot UInt32 -distunits GEO"))
[…]

### 2017-02-23 R/plotfcc.R
Update readme

@@ -19,17 +19,17 @@ zooms <- list(zoom.w,zoom.e)
 
 # Compute zooms
 for (z in 1:length(zooms)) {
-  Extent1 <- paste(zooms[[z]]$xmin,zooms[[z]]$ymin,zooms[[z]]$xmax,zooms[[z]]$ymax)
-  Input <- c("fcc.tif","fcc_for1953.tif","frag2014.tif")
+  ExtentZ <- paste(zooms[[z]]$xmin,zooms[[z]]$ymin,zooms[[z]]$xmax,zooms[[z]]$ymax)
+  Input <- c("fcc.tif","fcc_for1953.tif","frag2014.tif","dist_edge_2014.tif")
   Output <- paste0(sub(".tif","",Input),"_zoom",z,".tif")
   for (i in 1:length(Input)) {
-    system(paste0("gdalwarp -overwrite -srcnodata 255 -dstnodata 255 -te ",Extent1," -co 'COMPRESS=LZW' -co 'PREDICTOR=2' ",
+    system(paste0("gdalwarp -overwrite -srcnodata 255 -dstnodata 255 -te ",ExtentZ," -co 'COMPRESS=LZW' -co 'PREDICTOR=2' ",
                   paste0("outputs/",Input[i])," ",paste0("outputs/",Output[i])))
   }
 }
 
 # Resolution
-res <- 10e5
+res <- 10e3
 
 # get_legend()
 get_legend <- function(myggplot) {
@@ -94,6 +94,18 @@ plot_zoom <- function(rast_file,res,palette) {
   return(pzoom)
 }
 
+# plot_zoom_dist()
+plot_zoom_dist <- function(rast_file,res) {
+  # Function to plot zoom
+  r <- raster(rast_file)
+  pzoom <- gplot(r,maxpixels=res) +
+    geom_raster(aes(fill=value)) +
+    theme_bw() + theme_zoom +
+    coord_equal()
+  return(pzoom)
+}
+
+
 
 ## Palette
 # fcc
@@ -115,6 +127,13 @@ frag.name <- c(0:6)
 frag.palette <- paste0("c(",paste0("\"",frag.name,"\"","=","\"",
                                    frag.col,"\"",collapse=","),")")
 frag.palette <- eval(parse(text=frag.palette))
+# distance to forest edge
+cf <- colorRampPalette(c("darkred","orange","darkgreen"))(255)
+dist.col <- c(nofor73.col,cf)
[…]

### 2017-02-23 R/plotfcc.R
Improving fcc plot

@@ -40,6 +40,7 @@ get_legend <- function(myggplot) {
   return(legend)
 }
 
+# ======
 # Colors
 for2014.col = rgb(34/255,139/255,34/255) # forest green
 defor00_14.col = rgb(255/255,0/255,0/255) # red
@@ -50,6 +51,34 @@ w1.col = rgb(153/255,217/255,234/255) # light blue
 w2.col = rgb(0/255,0/255,170/255) # dark blue
 water.col <-colorRampPalette(c(w1.col,w2.col))(12)
 
+# =======
+# Palette
+# fcc (discrete)
+fcc.col <- c(nofor73.col,water.col,nofor73.col,defor73_90.col,nofor73.col,
+             defor90_00.col,defor00_14.col,for2014.col)
+fcc.name <- c(0,1:12,20,21,215,22,23,24)
+fcc.palette <- paste0("c(",paste0("\"",fcc.name,"\"","=","\"",fcc.col,"\"",collapse=","),")")
+fcc.palette <- eval(parse(text=fcc.palette))
+# fcc_for1953 (discrete)
+fcc_for1953.col <- c(nofor73.col,water.col,for2014.col)
+fcc_for1953.name <- c(0,1:12,20)
+fcc_for1953.palette <- paste0("c(",paste0("\"",fcc_for1953.name,"\"","=","\"",
+                                          fcc_for1953.col,"\"",collapse=","),")")
+fcc_for1953.palette <- eval(parse(text=fcc_for1953.palette))
+# fragmentation (discrete)
+cf <- colorRampPalette(c("darkred","orange","darkgreen"))(5)
+frag.col <- c(nofor73.col,cf)
+frag.name <- c(0:6)
+frag.palette <- paste0("c(",paste0("\"",frag.name,"\"","=","\"",
+                                   frag.col,"\"",collapse=","),")")
+frag.palette <- eval(parse(text=frag.palette))
+# distance to forest edge (continuous)
+dist.v <- c(50,100,500,1000,3000)
+dist.vr <- c(0,dist.v)/3000
+cf <- colorRampPalette(c("darkred","orange","darkgreen"))(5)
+dist.palette <- c(nofor73.col,cf)
+
+# ======
 # Themes
 theme_base <- theme(axis.line=element_blank(),
                     axis.text.x=element_blank(),
@@ -82,59 +111,55 @@ theme_zoom <- theme(axis.line=element_blank(),
                     panel.grid.minor=element_blank(),
[…]

### 2017-02-23 deforestmap.R

@@ -569,7 +569,13 @@ for (i in 1:length(Year)) {
   system(paste0("gdal_translate -a_nodata 0 -co 'COMPRESS=LZW' -co 'PREDICTOR=2' \\
                 outputs/_dist_edge_",Year[i],".tif outputs/dist_edge_",Year[i],".tif"))
   file.remove(paste0("outputs/_dist_edge_",Year[i],".tif"))
-  system(paste0("r.in.gdal --o input=outputs/dist_edge_",Year[i],".tif output=dist_edge_",Year[i],".tif"))
+  system(paste0("r.in.gdal --o input=outputs/dist_edge_",Year[i],".tif output=dist_edge_",Year[i]))
+  if (Year[i]==2014) {  # To mask outside Mada for plots
+    system("r.mask --o raster=harper")
+    system(paste0("r.out.gdal --o input=dist_edge_",Year[i]," createopt='compress=lzw,predictor=2' \\
+                  output=outputs/dist_edge_",Year[i],".tif"))
+    system("r.mask -r")
+  }
 }
 
 ##==========================

### 2017-02-24 deforestmap.Rmd
Comparison with previous studies

@@
 
-We derived 30m cloud-free forest maps for Madagascar for the years 1990, 2000, 2010 and 2014 combining the 1990 and 2000 forest maps by @Harper2007 and the 2000-2014 tree percent cover map by @Hansen2013.
-On Harper's map, ~200,000 ha clouds are present over the ~4.2 million ha moist forest ecoregion (4.8% of clouds).
-To remove these clouds, we used the 2000 cloud-free tree percent cover map provided by @Hansen2013 (also at 30~m resolution) and we chose a threshold of 75% of tree cover [@Achard2014] to decide whether to replace cloud pixels with forest or non-forest pixels.
+We derived 30m cloud-free forest maps for Madagascar for the years 1990, 2000, 2010 and 2014 combining the 1990 and 2000 forest maps by @Harper2007 and the 2000-2014 tree cover and tree cover loss maps by @Hansen2013.
+On the 2000 Harper's map, ~200,000 ha clouds are present over the ~4.2 million ha moist forest ecoregion (4.8% of clouds).
+To remove these clouds, we used the 2000 cloud-free tree cover map provided by @Hansen2013 (also at 30~m resolution) and we chose a threshold of 75% of tree cover [@Achard2014] to decide whether to replace cloud pixels with forest or non-forest pixels.
 We thus obtained a cloud-free year 2000 forest map for Madagascar.
@@
 Doing so, all clouds were removed and we obtained a cloud-free 1990 forest map for Madagascar.
+For all forest cover maps, we also removed the isolated non-forest pixels if all the height surrounding pixels were covered by forest, assuming that small-scaled deforestation on a 30m square pixel inside the forest is unlikely.
+This was important in order not to over-estimate forest fragmentation.
 These maps are freely available at http:/bioscenemada.cirad.fr/forestmaps.
@@
 
-### 3.1 Validation
+### 3.1 Evolution of forest cover and deforestation intensity
+
+### 3.2 Comparison with previous forest-cover change studies in Madagascar
 
 Difference might be due to post-processing of @Rakotomalala2015 map.
-
-### 3.2 Evolution of forest cover and deforestation intensity
 
@@
 Finally, young secondary forests do not provide the same ecosystem services (in terms of biodiversity and carbon storage) as old-growth natural forests.
+
+Should be used to actualize FAO report.
+
+There has been a long debate about whether Madagascar has lost 90% of its forest since the arrival of the humans on the island [@Virah-Sawmy2009, @Agarwal2005].
+What it sure now, is that Madgascar has lost around 50% of its natural forest in six decades since _c._ 1953.
 

### 2017-02-24 R/comp_old_deprecated.R

@@ -1,0 +1,193 @@
+# Import perr-fh map
+system("r.in.gdal --o input=gisdata/rasters/perr-fh/FCC_051013_MG_VF_38s.tif output=perrfh")
+
+# Classification
+system("r.describe perrfh")
+# * 7 111 113 119 133 183 222 223 228 233 288 333 444 533 553 555 666 733 773 777 778 788 888 1000
+# 1=moist_forest, 2=dry_forest, 3=no_forest, 4=water, 5=spiny_forest, 6=building/soil/rock,
+# 7=mangroves, 8=cloud/shadow, 9=cloud, 1000=artefact
+
+# 111	Forêt humide 2013
+# 113	Déforestation 2010-2013
+# 119	Nuage 2013
+# 133	Déforestation 2005-2013
+# 183	Ombre 2010
+# 222	Forêt sèche 2013
+# 223	Déforestation forêt séche 2010-2013
+# 228	Ombre 2013
+# 233	Déforestation forêt séche 2005-2010
+# 288	Ombre 2010
+# 333	Mosaic de culture, jachère et savane
+# 444	Zone en eau
+# 533	Déforestation forêt épineuse 2005-2010
+# 553	Déforestation forêt épineuse 2010-2013
+# 555	Forêt épineuse
+# 666	Etablissement, sol et roche nue
+# 733	Déforestation forêt de mangrove 2005-2010
+# 773	Déforestation forêt de Mangrove 2010-2013
+# 777	Forêt Mangrove 2013
+# 778	Nuage 2010
+# 788	Nuage 2010
+# 888	Nuage
+
+# Report
+system("r.report perrfh units=h")
+
+# +-----------------------------------------------------------------------------+
+# |                         RASTER MAP CATEGORY REPORT                          |
+# |LOCATION: deforestmap                                Fri Feb  3 11:25:54 2017|
+# |-----------------------------------------------------------------------------|
+# |          north: 8682420    east: 1100820                                    |
+# |REGION    south: 7155900    west:  298440                                    |
+# |          res:        30    res:       30                                    |
+# |-----------------------------------------------------------------------------|
[…]

### 2017-02-24 deforestmap.R

@@ -296,7 +296,27 @@ system("r.stats -c for1953")
 ## * 1135151199
 
 #= Export
-system("r.out.gdal --o input=for1953 createopt='compress=lzw,predictor=2' type=Byte output=outputs/for1953.tif")
+system("r.out.gdal --o input=for1953 createopt='compress=lzw,predictor=2' \\
+       type=Byte output=outputs/for1953.tif")
+
+# =======================================
+# Remove salt and pepper
+# =======================================
+
+Year <- c(1953,1973,1990,2000,2010,2014)
+system("g.region rast=harper -ap")
+system("r.mask --o raster=harper")
+for (i in 1:length(Year)) {
+  # Message
+  cat(paste0("Year: ",Year[i],"\n"))
+  # Computation  
+  system(paste0("r.neighbors --o input=for",Year[i]," output=for",Year[i],"_sum method=sum size=3"))
+  system(paste0("r.mapcalc --o 'for",Year[i]," = if(isnull(for",Year[i],") &&& for",Year[i],"_sum==8, 1, for",Year[i],")'"))
+  # Export
+  system(paste0("r.out.gdal --o input=for",Year[i]," createopt='compress=lzw,predictor=2' \\
+         type=Byte output=outputs/for",Year[i],".tif"))
+}
+system("r.mask -r")
 
 #====================================================================================
 # Export images
@@ -309,14 +329,14 @@ system("echo '1 26:152:80' | r.colors for2000 rules=-")
 system("echo '1 26:152:80' | r.colors for2010 rules=-")
 system("echo '1 26:152:80' | r.colors for2014 rules=-")
 
-# #= PNG
+#= PNG
 system("g.region res=1000")
-system("r.out.png input=for1953 output=outputs/for1953.png")
-system("r.out.png input=for1973 output=outputs/for1973.png")
-system("r.out.png input=for1990 output=outputs/for1990.png")
-system("r.out.png input=for2000 output=outputs/for2000.png")
-system("r.out.png input=for2010 output=outputs/for2010.png")
-system("r.out.png input=for2014 output=outputs/for2014.png")
+system("r.out.png --o input=for1953 output=outputs/for1953.png")
[…]

### 2017-02-25 deforestmap.R
Add _test for computation

@@ -425,6 +425,27 @@ for (i in 2:length(Year)) {
 }
 write.table(defor.df,"outputs/defor.txt",row.names=FALSE,sep="\t")
 
+#= Deforestation statistics for comparison
+Year <- c(1953,1973,1990,2000,2005,2010,2013)  # Here we add 2005 and 2013
+defor.df <- data.frame(Year=Year,area=NA,ann.defor=NA,theta=NA)
+# Areas
+for (i in 1:length(Year)) {
+  # Message
+  cat(paste("Year: ",Year[i],"\n",sep=""))
+  # Computation
+  statcell <- system(paste0("r.stats -c for",Year[i]), intern=TRUE)
+  ncells <- as.numeric(strsplit(statcell[1],split=" ")[[1]][2])
+  defor.df$area[i] <- round(ncells*(as.numeric(Res)^2)/10000)
+}
+# Annual rates
+for (i in 2:length(Year)) {
+  theta.prim <- (defor.df$area[i-1]-defor.df$area[i])/defor.df$area[i-1]
+  Y <- defor.df$Year[i]-defor.df$Year[i-1]
+  defor.df$ann.defor[i] <- round((defor.df$area[i-1]-defor.df$area[i])/Y)
+  defor.df$theta[i] <- round(100*(1-(1-theta.prim)^(1/Y)),2)
+}
+write.table(defor.df,"outputs/defor_for_comp.txt",row.names=FALSE,sep="\t")
+
 #= Fragmentation statistics
 Year <- c(1953,1973,1990,2000,2010,2014)
 frag.df <- data.frame(Year=Year,forest=NA,patch=NA,transitional=NA,
@@ -493,6 +514,30 @@ for (j in 1:3) {
     defor.df$ann.defor[i] <- round((defor.df$area[i-1]-defor.df$area[i])/Y)
     defor.df$theta[i] <- round(100*(1-(1-theta.prim)^(1/Y)),2)
   }
+  # Export
+  write.table(defor.df,paste0("outputs/defor_",ecoregion[j],".txt"),row.names=FALSE,sep="\t")
+  # ==========================================
+  # B. Deforestation statistics for comparison
+  Year <- c(1953,1973,1990,2000,2005,2010,2013)  # Here we add 2005 and 2013 for comparison
+  defor.df <- data.frame(Year=Year,area=NA,ann.defor=NA,theta=NA)
+  # Areas
+  for (i in 1:length(Year)) {
+    # Message
[…]

### 2017-02-25 deforestmap_test.R

@@ -1,0 +1,283 @@
+#!/usr/bin/Rscript
+
+# ==============================================================================
+# author          :Ghislain Vieilledent
+# email           :ghislain.vieilledent@cirad.fr, ghislainv@gmail.com
+# web             :https://ghislainv.github.io
+# license         :GPLv3
+# ==============================================================================
+
+##=========================================================================
+## Six decades of deforestation and forest fragmentation in Madagascar
+##=========================================================================
+
+## gdal library is needed to run this script
+## http://www.gdal.org/
+
+## GRASS GIS 7.2.x is also needed to run this script
+## https://grass.osgeo.org/
+
+## Read argument for download
+## Set "down" to TRUE if you want to download the sources. Otherwise, the data already provided in the gisdata repository will be used.
+arg <- commandArgs(trailingOnly=TRUE)
+down <- FALSE
+if (length(arg)>0) {
+  down <- arg[1]
+}
+
+##= Libraries
+pkg <- c("broom","sp","rgdal","raster","ggplot2","gridExtra","RColorBrewer",
+         "rasterVis","knitr","rmarkdown","rgeos","rgdal","rgrass7")
+## broom: to convert map into data-frame with tiny()
+## gridExtra: to combine several ggplots
+## rasterVis: for gplot()
+## rgeos: for crop()
+load.pkg <- function(x) {
+  if(!require(x, character.only = T)) {
+    install.packages(x)
+    require(x, character.only = T)
+  }
+}
+loaded <- lapply(pkg,load.pkg)
+## Remove useless objects
+rm(pkg,load.pkg,loaded)
+
+##=====================================
+## Create new grass location in UTM 38S
+#dir.create("grassdata")
+#system("grass72 -c epsg:32738 grassdata/deforestmap")  # Ignore errors
+## Connect R to grass location
+initGRASS(gisBase="/usr/local/grass-7.2.0",home=tempdir(), 
+          gisDbase="grassdata",
[…]

### 2017-02-25 deforestmap.sh
Add _test for computation

@@ -1,6 +1,6 @@
 #!/bin/sh
 echo "Executing deforestmap.R script in the background"
-Rscript --vanilla deforestmap.R > deforestmap.log 2>&1 &
+Rscript --vanilla deforestmap_test.R > deforestmap.log 2>&1 &
 # If you wan't to download the source data do:
 # Rscript --vanilla deforestmap.R TRUE > deforestmap.log 2>&1 &
 echo "Check the progress with command 'tail -f deforestmap.log'"

### 2017-02-25 deforestmap_test.R
Add Res <- 30

@@ -106,6 +106,8 @@ for (i in 1:length(Year)) {
 ## Statistics whole forest
 ##========================
 
+Res <- 30
+
 #= Deforestation statistics
 Year <- c(1953,1973,1990,2000,2010,2014)
 defor.df <- data.frame(Year=Year,area=NA,ann.defor=NA,theta=NA)

### 2017-02-27 deforestmap.Rmd
Add Tab. 2

@@
 
-```{r comp, results="asis"}
-tab.1 <- comp_moist.df
-mean.diff <- round(100*(sum(tab.1$BSM)-sum(tab.1$PERR_FH))/sum(tab.1$PERR_FH))
-names(tab.1) <- c("Year","Vieil (ha)", "Rakot (ha)", "Rakot_NA (ha)", "ad_Vieil (ha/yr)", "ad_Rakot (ha/yr)", "theta_Vieil (%/yr)", "theta_Rakoto (%/yr)", "diff.area (%)")
-tab.1 <- tab.1[,c(1:4,9,5:8)]
+```{r defor, results="asis"}
+perc.1 <- round(100*(defor.df$area[defor.df$Year==1973]-defor.df$area[defor.df$Year==2014])/defor.df$area[defor.df$Year==1973])
+perc.2 <- round(100*(defor.df$area[defor.df$Year==1953]-defor.df$area[defor.df$Year==2014])/defor.df$area[defor.df$Year==1953])
+d.1 <- defor.df$theta[defor.df$Year==1990]
+d.2 <- defor.df$theta[defor.df$Year==2010]
+tab.1 <- defor.df
+names(tab.1) <- c("Year","Forest (ha)", "Annual defor. (ha/yr)", "theta (%/yr)")
 kable(tab.1)
@@
 
-Table 1: **Comparison of the forest cover and deforestation rates between our study and the study by @Rakotomalala2015**.
-This comparison is done only for the humid ecoregion for the periods 2005-2010-2013, which is the extent and time-periods of the @Rakotomalala2015 map.
-@Rakotomalala2015 map has been validated using 10,854 independant visual interpretation points leading to a rather high accuracy (Kappa=0.79 and overall-accuracy=0.89).
-Our study (suffix _"Vieil"_) estimated a forest cover greater by about `r mean.diff`% compared to @Rakotomalala2015 (suffix _"Rakot"_).
-But the difference (_"diff.area"_) did not vary much between dates (between 3.56 and 6.18%).
-Our study also estimated higher annual deforestation (_"ad"_) and deforestation rates (_"theta"_) but the same trend for the deforestation was observed between the two studies, with a doubling of the deforestation for the period 2010-2013 compared to 2005-2010.
-The differences between the two studies could be attributed to the forest definition which might be different in the two studies and to the post-processing in @Rakotomalala2015 map.
+Table 1: **Evolution of the forest cover and deforestation intensity from 1950s to 2014 in Madagascar**.
+Madagascar has lost `r perc.1`% of its natural forest in 40 years between _c._ 1973 and 2014 (and about `r perc.2`% in 60 years since _c._ 1953).
+If the deforestation has progressively slowed down from _c._ 1973 to 2010 (from `r d.1` to `r d.2`%/yr), most likely due to a decrease of available arable lands under the remaining forests and to a decrease of forest accessibility, the deforestation has recently increased, with a doubling deforestation on the period 2010-2014 compared to 2000-2010.
+This increase could be attributed to both political instability (political crisis in 2009) and to an explosion of the demography (according to the @UN2015, population has increased from about 16M inhabitants in 2000 to 24M in 2014).
+Natural forest covers 8.9 Mha in 2014 and current (period 2000-2014) mean annual deforestation rate is of about 70,000 ha/yr.
 
-```{r defor, results="asis"}
-tab.2 <- defor.df
-names(tab.2) <- c("Year","Forest (ha)", "Annual defor. (ha/yr)", "theta (%/yr)")
-kable(tab.2)
+```{r comp, results="asis"}
+tab.2 <- df.comp
+tab.2 <- tab.2[,c(2,1,3:10)]
+tab.2$ForestType <- c("Total",rep("",3),"Moist",rep(" ",3),"Dry",rep(" ",3),"Spiny",rep(" ",3),"Mangroves",rep(" ",3))
+col.names <- c("Forest type","Source","1953","1973","1990","2000","2005","2010","2013","2014")
+kable(tab.2,col.names=col.names,align=c("l","l",rep("r",8)))
+#print(xtable(tab.2,col.names=col.names,align=c("llrrrrrrrrr")),type="html")
 ```
 
-Table 2: **Evolution of the forest cover and deforestation intensity from 1950s to 2014 in Madagascar**.
-Madagascar has lost 37% of its natural forest in 40 years between _c._ 1973 and 2014 (and roughly 56% in 60 years since _c._ 1953).
-If the deforestation has progressively slowed down from _c._ 1953 to 2010 (from 1.76 to 0.59%/yr), most likely due to a decrease of available arable lands under the remaining forests and to a decrease of forest accessibility, the deforestation has recently increased, with a doubling deforestation on the period 2010-2014 compared to 2000-2010.
-This increase could be attributed to both political instability (political crisis in 2009) and to an explosion of the demography (according to the @UN2015, population has increased from about 16M inhabitants in 2000 to 24M in 2014).
-Natural forest cover 8.9 Mha in 2014 and current (period 2000-2014) mean annual deforestation rate is of about 70,000 ha/yr.
+Table 2: **Comparison of Madagascar forest-cover estimates (in ha) with previous studies on the period 1953-2014**.
+We found similiar forest-cover estimates compared to previous studies [@Harper2007;@MEFT2009;@ONE2015].
+Our study extends the forest-cover monitoring on a six decades period (from 1953 to 2014) while harmonizing the data from the previous studies.
+Our method provides cloud-free forest-cover estimates from 1990 to 2014 and corrections for the forest-cover in 1973.
 
@@
 tab.3 <- frag.df
+p.1 <- round(frag.df$interior[frag.df$Year==1973])
+p.2 <- round(frag.df$interior[frag.df$Year==2014])
+p.3 <- round(frag.df$patch[frag.df$Year==2014]+frag.df$transitional[frag.df$Year==2014])
 names(tab.3) <- c("Year","Forest (ha)", "patch (%)", "transitional (%)", "edge (%)", "perforated (%)", "interior (%)", "NA (%)")
@@
 Forest fragmentation has progressively increased from _c._ 1953.
-The percentage of forest belonging to the interior category (the most intact forest) has fallen from 94 in _c._ 1973 to 48% in 2014.
-More than 15% of the forest belong to the patch and transitional categories (isolated small forest patches) in 2014.
+The percentage of forest belonging to the "interior"" category (the most intact forest) has fallen from `r p.1`% in _c._ 1973 to `r p.2`% in 2014.
+More than `r p.3`% of the forest belong to the "patch"" and "transitional" categories (isolated small forest patches) in 2014.
 This increase in forest fragmentation has a great impact on both carbon emissions [@Putz2014] and biodiversity loss [@Gibson2013, @Murcia1995].

### 2017-02-27 deforestmap.R

@@ -27,7 +27,7 @@ if (length(arg)>0) {
 
 ##= Libraries
 pkg <- c("broom","sp","rgdal","raster","ggplot2","gridExtra","RColorBrewer",
-         "rasterVis","knitr","rmarkdown","rgeos","rgdal","rgrass7")
+         "rasterVis","knitr","rmarkdown","rgeos","rgdal","rgrass7","xtable")
 ## broom: to convert map into data-frame with tiny()
 ## gridExtra: to combine several ggplots
 ## rasterVis: for gplot()
@@ -580,7 +580,34 @@ for (j in 1:3) {
 ## Comparison with previous studies on forest-cover change
 ##========================================================
 
+# Historical data
 df.comp <- read.csv("data/fcc_comp.csv",header=TRUE)
+# This study
+defor_for_comp <- read.table("outputs/defor_for_comp.txt",header=TRUE)
+defor_moist_for_comp <- read.table("outputs/defor_moist_for_comp.txt",header=TRUE)
+defor_dry_for_comp <- read.table("outputs/defor_dry_for_comp.txt",header=TRUE)
+defor_spiny_for_comp <- read.table("outputs/defor_spiny_for_comp.txt",header=TRUE)
+defor_mangroves_for_comp <- read.table("outputs/defor_mangroves_for_comp.txt",header=TRUE)
+# Complete data
+df.comp[df.comp$ForestType=="Total" & df.comp$Source=="this study", 3:9] <- defor_for_comp$area
+df.comp[df.comp$ForestType=="Moist" & df.comp$Source=="this study", 3:9] <- defor_moist_for_comp$area
+df.comp[df.comp$ForestType=="Dry" & df.comp$Source=="this study", 3:9] <- defor_dry_for_comp$area
+df.comp[df.comp$ForestType=="Spiny" & df.comp$Source=="this study", 3:9] <- defor_spiny_for_comp$area
+df.comp[df.comp$ForestType=="Mangroves" & df.comp$Source=="this study", 3:9] <- defor_mangroves_for_comp$area
+# This study for year 2014
+defor <- read.table("outputs/defor.txt",header=TRUE)
+defor_moist <- read.table("outputs/defor_moist.txt",header=TRUE)
+defor_dry <- read.table("outputs/defor_dry.txt",header=TRUE)
+defor_spiny <- read.table("outputs/defor_spiny.txt",header=TRUE)
[…]

### 2017-02-27 deforestmap.R
Plot dist to forest edge

@@ -464,17 +464,19 @@ write.table(frag.df,"outputs/frag.txt",row.names=FALSE,sep="\t")
 
 #= Distance to forest edge statistics
 Year <- c(1953,1973,1990,2000,2010,2014)
-dist.df <- data.frame(Year=Year,min=NA,max=NA,mean=NA,sd=NA)
+dist.df <- data.frame(Year=Year,min=NA,max=NA,mean=NA,sd=NA,q1=NA,q2=NA)
 # Areas
 for (i in 1:length(Year)) {
   # Message
   cat(paste("Year: ",Year[i],"\n",sep=""))
   # Computation
-  statcell <- system(paste0("r.univar dist_edge_",Year[i]), intern=TRUE)
+  statcell <- system(paste0("r.univar -e percentile=5,95 map=dist_edge_",Year[i]), intern=TRUE)
   dist.df$min[i] <- unlist(strsplit(statcell[7],split=" "))[2]
   dist.df$max[i] <- unlist(strsplit(statcell[8],split=" "))[2]
   dist.df$mean[i] <- unlist(strsplit(statcell[10],split=" "))[2]
-  dist.df$sd[i] <- unlist(strsplit(statcell[12],split=" "))[2]
+  dist.df$sd[i] <- unlist(strsplit(statcell[12],split=" "))[3]
+  dist.df$q1[i] <- unlist(strsplit(statcell[19],split=" "))[3]
+  dist.df$q2[i] <- unlist(strsplit(statcell[20],split=" "))[3]
 }
 write.table(dist.df,"outputs/dist.txt",row.names=FALSE,sep="\t")
 
@@ -674,6 +676,28 @@ system("r.mask -r")
 ## Plot raster with gplot() from rasterVis
 source(file="R/plotfcc.R")
 
+##==================================================
+## Plot for evolution of the distance to forest edge
+##==================================================
+
+# Data
+dist.df <- read.table(file="outputs/dist.txt", header=TRUE)
+
+# Plot
+png(file="outputs/dist.png",width=800,height=600)
+par(mar=c(4,4,0,0),cex=2,lwd=2)
+plot(NA, xlim=c(1953,2014), ylim=c(0,5000),
+     xlab="Year",
+     ylab="Distance to forest edge (km)",
+     
+     axes=FALSE)
[…]

### 2017-02-28 deforestmap.Rmd
Add tab defor.comp

@@
 
-```{r comp, results="asis"}
+```{r fcc_comp, results="asis"}
 tab.2 <- df.comp
@@
 We found similiar forest-cover estimates compared to previous studies [@Harper2007;@MEFT2009;@ONE2015].
+The increase in mangrove and spiny forest covers from _c._ 1953 to _c._ 1973 in @Harper2007 and our study is most probably due to differences in forest definition and mapping methods.
 Our study extends the forest-cover monitoring on a six decades period (from 1953 to 2014) while harmonizing the data from the previous studies.
@@
 
+```{r defor_comp, results="asis"}
+tab.3 <- defor.comp
+tab.3$ForestType <- c("Total",rep("",3),"Moist",rep(" ",3),"Dry",rep(" ",3),"Spiny",rep(" ",3),"Mangroves",rep(" ",3))
+col.names <- c("Forest type","Source","1953-1973","1973-1990","1990-2000","2000-2005","2005-2010","2010-2013")
+kable(tab.2,col.names=col.names,align=c("l","l",rep("r",6)))
+```
+
+Table 3: **Comparison of Madagascar annual deforestation rates with previous studies on the period 1953-2014**.
+Annual deforestation rates are providing in ha/yr and in %/yr (second number in parenthesis).
+Original estimates of deforestation rates might differ from original publications [@Harper2007;@MEFT2009;@ONE2015].
+In our study, we used the following formula [@Puyravaud2003] to compute the annual deforestation rates: $\theta = 100 \times [1-(1-(F_{t2}-F_{t1})/F_{t1})^{(1/(t2-t1))}]$, with $\theta$ the annual deforestation rate (in %/yr), $F_{t2}$ and $F_{t1}$ the forest covers (reported in Tab.
+2) at times t2 and t1 (in years).
+
 ```{r frag, results="asis"}
-tab.3 <- frag.df
+tab.4 <- frag.df
 p.1 <- round(frag.df$interior[frag.df$Year==1973])
@@
 p.3 <- round(frag.df$patch[frag.df$Year==2014]+frag.df$transitional[frag.df$Year==2014])
-names(tab.3) <- c("Year","Forest (ha)", "patch (%)", "transitional (%)", "edge (%)", "perforated (%)", "interior (%)", "NA (%)")
-kable(tab.3)
+names(tab.4) <- c("Year","Forest (ha)", "patch (%)", "transitional (%)", "edge (%)", "perforated (%)", "interior (%)", "NA (%)")
+kable(tab.4)
 ```
 
-Table 3: **Evolution of the forest fragmentation from 1950s to 2014 in Madagascar**.
+Table 4: **Evolution of the forest fragmentation from 1950s to 2014 in Madagascar**.
 Forest fragmentation has progressively increased from _c._ 1953.
 The percentage of forest belonging to the "interior"" category (the most intact forest) has fallen from `r p.1`% in _c._ 1973 to `r p.2`% in 2014.
-More than `r p.3`% of the forest belong to the "patch"" and "transitional" categories (isolated small forest patches) in 2014.
+More than `r p.3`% of the forest belong to the "patch" and "transitional" categories (isolated small forest patches) in 2014.
 This increase in forest fragmentation has a great impact on both carbon emissions [@Putz2014] and biodiversity loss [@Gibson2013, @Murcia1995].

### 2017-02-28 deforestmap.R

@@ -582,6 +582,7 @@ for (j in 1:3) {
 ## Comparison with previous studies on forest-cover change
 ##========================================================
 
+## Forest-cover
 # Historical data
 df.comp <- read.csv("data/fcc_comp.csv",header=TRUE)
 # This study
@@ -610,6 +611,28 @@ df.comp$y2014[df.comp$ForestType=="Spiny" & df.comp$Source=="this study"] <- def
 df.comp$y2014[df.comp$ForestType=="Mangroves" & df.comp$Source=="this study"] <- defor_mangroves$area[defor_mangroves$Year==2014]
 # Save results
 write.table(df.comp,file="outputs/fcc_comp.txt",row.names=FALSE)
+
+## Deforestation
+# Historical data
+df.comp <- read.table("outputs/fcc_comp.txt",header=TRUE)
+# Deforestation ha/yr
+defor.ha.comp <- df.comp[,c(2,1,3:8)]
+periods <- c("1953-1973","1973-1990","1990-2000","2000-2005","2005-2010","2010-2013")
+lp <- c(20,17,10,5,5,3)
+names(defor.ha.comp)[3:8] <- periods
+for (i in 1:length(periods)) {
+  defor.ha.comp[,2+i] <- round((df.comp[,2+i]-df.comp[,2+1+i])/lp[i])
+}
+# Deforestation %/yr
+defor.perc.comp <- defor.ha.comp
+for (i in 1:length(periods)) {
+  defor.perc.comp[,2+i] <- round(100*(1-(1-(df.comp[,2+i]-df.comp[,2+1+i])/df.comp[,2+i])^(1/lp[i])),2)
+}
+# Combining ha and percentages
+ha <- as.numeric(unlist(defor.ha.comp[,3:8]))
+perc <- sprintf("%.2f", as.numeric(unlist(defor.perc.comp[,3:8])))
+defor.comp[,3:8] <- data.frame(matrix(paste(ha," (",perc,")", sep=""),ncol=6))
+defor.comp[defor.comp=="NA (NA)"] <- NA
 
 ##========================
 ## Forest-cover change map

### 2017-02-28 deforestmap.sh

@@ -1,6 +1,6 @@
 #!/bin/sh
 echo "Executing deforestmap.R script in the background"
-Rscript --vanilla deforestmap_test.R > deforestmap.log 2>&1 &
+Rscript --vanilla deforestmap.R > deforestmap.log 2>&1 &
 # If you wan't to download the source data do:
 # Rscript --vanilla deforestmap.R TRUE > deforestmap.log 2>&1 &
 echo "Check the progress with command 'tail -f deforestmap.log'"

### 2017-02-28 deforestmap_test.R

@@ -1,285 +1,0 @@
-#!/usr/bin/Rscript
-
-# ==============================================================================
-# author          :Ghislain Vieilledent
-# email           :ghislain.vieilledent@cirad.fr, ghislainv@gmail.com
-# web             :https://ghislainv.github.io
-# license         :GPLv3
-# ==============================================================================
-
-##=========================================================================
-## Six decades of deforestation and forest fragmentation in Madagascar
-##=========================================================================
-
-## gdal library is needed to run this script
-## http://www.gdal.org/
-
-## GRASS GIS 7.2.x is also needed to run this script
-## https://grass.osgeo.org/
-
-## Read argument for download
-## Set "down" to TRUE if you want to download the sources. Otherwise, the data already provided in the gisdata repository will be used.
-arg <- commandArgs(trailingOnly=TRUE)
-down <- FALSE
-if (length(arg)>0) {
-  down <- arg[1]
-}
-
-##= Libraries
-pkg <- c("broom","sp","rgdal","raster","ggplot2","gridExtra","RColorBrewer",
-         "rasterVis","knitr","rmarkdown","rgeos","rgdal","rgrass7")
-## broom: to convert map into data-frame with tiny()
-## gridExtra: to combine several ggplots
-## rasterVis: for gplot()
-## rgeos: for crop()
-load.pkg <- function(x) {
-  if(!require(x, character.only = T)) {
-    install.packages(x)
-    require(x, character.only = T)
-  }
-}
-loaded <- lapply(pkg,load.pkg)
-## Remove useless objects
-rm(pkg,load.pkg,loaded)
-
-##=====================================
-## Create new grass location in UTM 38S
-#dir.create("grassdata")
-#system("grass72 -c epsg:32738 grassdata/deforestmap")  # Ignore errors
-## Connect R to grass location
-initGRASS(gisBase="/usr/local/grass-7.2.0",home=tempdir(), 
-          gisDbase="grassdata",
[…]

### 2017-03-01 deforestmap.Rmd
Update methods part

@@
 In this study, we present a simple approach combining maps from @Harper2007 and products from @Hansen2013 to derive annual and cloud-free forest-cover change map for Madagascar for the period 2000-2014.
-Basically, we used the forest map provided by @Harper2007 for the year 2000 (defining the land-use) with the tree cover loss product provided by @Hansen2013 that we applied only inside forest areas identified by @Harper2007.
+Basically, we used the natural forest-cover map provided by @Harper2007 for the year 2000 (defining the land-use) with the tree cover loss product provided by @Hansen2013 that we applied only inside forest areas identified by @Harper2007.
 Following the work done by @Harper2007, we studied in details the change in deforestation rates and the evolution of forest fragmentation on six decades from _c._ 1953 to 2014.
@@
 
-### 2.1 Historical deforestation maps
+### 2.1 Forest-cover maps for Madagascar from _c._ 1953 to 2014
 
 We derived 30m cloud-free forest maps for Madagascar for the years 1990, 2000, 2010 and 2014 combining the 1990 and 2000 forest maps by @Harper2007 and the 2000-2014 tree cover and tree cover loss maps by @Hansen2013.
-On the 2000 Harper's map, ~200,000 ha clouds are present over the ~4.2 million ha moist forest ecoregion (4.8% of clouds).
-To remove these clouds, we used the 2000 cloud-free tree cover map provided by @Hansen2013 (also at 30~m resolution) and we chose a threshold of 75% of tree cover [@Achard2014] to decide whether to replace cloud pixels with forest or non-forest pixels.
-We thus obtained a cloud-free year 2000 forest map for Madagascar.
-From this map, using tree cover loss data from 2000 to 2014 provided by @Hansen2013, we obtained cloud-free forest maps in 2010 and 2014 at 30~m resolution.
-Finally, we corrected the cloudy 1990 Harper forest map assuming that if forest was present in 2000, it was also present in 1990.
-Doing so, all clouds were removed and we obtained a cloud-free 1990 forest map for Madagascar.
-For all forest cover maps, we also removed the isolated non-forest pixels if all the height surrounding pixels were covered by forest, assuming that small-scaled deforestation on a 30m square pixel inside the forest is unlikely.
+On the 2000 Harper's map, 200,000 ha clouds are present over the 4.2 million ha moist forest ecoregion (4.8% of clouds).
+To remove these clouds, we used the 2000 cloud-free tree cover map provided by @Hansen2013 (also at 30m resolution) and we chose a threshold of 75% of tree cover [@Achard2014] to decide whether to replace clouds with forest or non-forest.
+We thus obtained a cloud-free year 2000 forest-cover map for Madagascar.
+From this map, using tree cover loss data from 2000 to 2014 provided by @Hansen2013, we obtained annual cloud-free forest-cover maps from 2010 to 2014 at 30m resolution.
+We corrected the cloudy 1990 Harper forest map assuming that if forest was present in 2000, it was also present in 1990.
+Doing so, cloud-cover was reduced to a negligeable quantity of `r epsilon_1990`ha.
+For these residual cloud areas, we assumed that forest-cover was the same as for the year 2000.
+We also corrected the cloudy 1973 Harper forest map assuming that if forest was present in 1973, it was also present in 1973.
+Contrary to the year 1990, we were not able to remove all the clouds on the 1973 map and the remaining cloud cover was of `r round(cloud_1973/10e6,2)` million ha.
+Because the 1953 forest-cover map was obtained from the digitalization of a paper map derived from aerial photographies, and thus cannot be perfectly aligned with the maps obtained in the following years, we did not performed any modifications on the original forest map of 1953 except reprojection.
+For all forest-cover maps from 1973, we removed the isolated non-forest pixels if all the surrounding pixels were covered by forest, assuming that small-scaled deforestation on a 30m square pixel inside the forest was unlikely.
 This was important in order not to over-estimate forest fragmentation.
-These maps are freely available at http:/bioscenemada.cirad.fr/forestmaps.
-On the basis of the forest-cover maps, we computed the mean annual deforestation intensity (in hectares of forest and in percentage) on six decades for 5 time periods from _c._ 1953 to 2014 (_c._ 1953-_c._ 1973-1990-2000-2010-2014).
+All these maps are made freely available at https://bioscenemada.cirad.fr/forestmaps.
 
-### 2.2 Comparison with other historical deforestation maps
+### 2.2 Computing forest-cover area and deforestation rate and comparing estimates with previous studies
 
-We estimated the accuracy of our approach combining Harper's and Hansen's products to derive cloud-free forest-cover change maps after the year 2000 in Madagascar.
-To do so, we used a forest-cover change map obtained from a regional study on the humid ecoregion on the period 2005-2013 [@Rakotomalala2015].
-@Rakotomalala2015 performed a supervised classification of Landsat images at multiple dates using the Random Forest classifier [@Grinand2013].
-The map by @Rakotomalala2015 is considered rather accurate on the moist ecoregion with an external validation on nearly 10,854 visual interpretation points leading to rather high Kappa and overall-accuracy values (0.79 and 0.89 respectively).
-Following our methodology, we produced a deforestation map on the same period of time (2005-2013) and compared the forest-cover (in ha) and mean annual deforestation rates (in ha and % per year) between the two maps.
+On the basis of the forest-cover maps we obtained, we computed the forest-cover area (in ha) and the mean annual deforestation rate (in hectares of forest and in percentage) on six decades, for five time periods from _c._ 1953 to 2014 (_c._ 1953-_c._ 1973-1990-2000-2010-2014).
+Mean annual deforestation rates in percentage were computed as follow [@Puyravaud]: $\theta = 100 \times [1-(1-(F_{t2}-F_{t1})/F_{t1})^{(1/(t2-t1))}]$, $\theta$ being the annual deforestation rate (in %/yr), $F_{t2}$ and $F_{t1}$ being the forest covers (reported in Tab.
+2) at times $t2$ and $t1$ (in years).
+Because of the large cloud-cover in 1973, the annual deforestation rates in hectares of forest for the two periods _c._ 1953-1973 and 1973-1990 are only rough estimates of the deforestation rates.
+This problem was overcome when computing the annual deforestation rates in percentage for these two periods.
+In this case, we only considered the cover change and computed the ratio $F_{t2}-F_{t1})/F_{t1}$ for pixels that were not covered by clouds in 1973.
+Computation were done at the national scale and for each forest type in Madagascar: the moist forest in the East, the dry forest in the West, the spiny forest in the South and mangroves on the West coast.
+To define the forest type, we used a map provided by the MEFT (_"Ministère de l'Environnement et des Forêts à Madagascar"_) showing the boundaries of the four ecoregions in Madagascar.
+Ecoregions were defined on the basis of climatic and vegetation criteria using the climate classification by @Cornet1974 and the vegetation classification from the 1996 IEFN national forest inventory [@IEFN1996].
+
+We compared our estimates of forest-cover and forest-cover change with three previous studies that have estimated forest-cover and forest-cover change at the national scale for Madagascar [@Harper2007; @MEFT2009; @ONE2015].
+@Harper2007 provided estimates for the periods _c._ 1953-_c._ 1973-1990-2000, @MEFT2009 provided estimates for the periods 1990-2000-2005 and @ONE2015 provided estimates for the periods 2005-2010-2013.
+The same ecoregion boundaries as in our study have been used in the @ONE2015 report but this was not the case for the @Harper2007 and @MEFT2009 reports, which can explain some of the differences between studies in the forest-cover areas and deforestation rates by ecoregions.
+W
 
@@
 We used the method developed by @Riitters2000 which uses a moving window to characterize the fragmentation around each forested pixel.
-Computation were done using function `r.forestfrag` from GRASS GIS software [@Neteler2008].
-Six categories of fragmentation are identified from the amount of forest and its occurrence as adjacent forest pixels: interior, perforated, edge, transitional, patch, and undetermined.
+Computation were done using the function `r.forestfrag` of the GRASS GIS software [@Neteler2008].
+Six categories of fragmentation are identified from the amount of forest and its occurrence as adjacent forest pixels: "interior", "perforated", "edge", "transitional", "patch", and "undetermined".
 We used a moving window of 7x7 pixels (4.41 km<sup>2</sup>).
@@
 The interior category can be interpreted as the most intact forest (sensus @Potapov2017).
-We reported the area of forest in each fragmentation category for the six years and analysed the evolution of the fragmentation with time.
+We reported the area of forest in each fragmentation category for the six years and analysed the evolution of the fragmentation on the six decades.
+We also computed the distance to forest edge (in km) for all forest pixels for the years _c._ 1953, _c._ 1973, 1990, 2000, 2010 and 2014.
+For that, we used the function `gdal_proximity.py` of the GDAL software (http://www.gdal.org/).
+We computed the mean and 95% quantiles of the distance to forest edge and looked at the evolution of these values with time.
 
@@
 ### 3.2 Comparison with previous forest-cover change studies in Madagascar
-
-Difference might be due to post-processing of @Rakotomalala2015 map.
 
@@
 col.names <- c("Forest type","Source","1953-1973","1973-1990","1990-2000","2000-2005","2005-2010","2010-2013")
-kable(tab.2,col.names=col.names,align=c("l","l",rep("r",6)))
+kable(tab.3,col.names=col.names,align=c("l","l",rep("r",6)))
 ```
@@
 
+![Figure 2: **Evolution of the mean distance to forest edge from 1953 to 2014 in Madagascar.** Black dots represent the mean distance to forest edge for each year.
+Vertical dashed segments represent the 95% quantiles of the distance to forest edge.
+We observed a continuous decrease of the mean distance to forest edge from 1953 to 2014.
+The mean distance was of about 1km in 1973 while it was of about 300m in 2014.
+Moreover, many areas were located far from forest edge (>2km and up to 4km) in 1973, while the majority of the forest was at a distance <1.2 km of the forest edge in 2014.](outputs/dist.png)
+
 ## Acknowledgements

### 2017-03-01 deforestmap.R

@@ -85,7 +85,7 @@ system("r.in.gdal --o input=outputs/harper.tif output=harper")
 # Set region
 system("g.region rast=harper -ap")
 
-#= Typology of CI map
+#= Typology of harper map 1990-2000-2005
 system("r.info harper")
 system("r.describe harper")
 # * 111 112 115 122 152 155 222 444 555 755 775 777
@@ -95,7 +95,7 @@ system("r.report harper units=h")
 
 ## +-----------------------------------------------------------------------------+
 ## |                         RASTER MAP CATEGORY REPORT                          |
-## |LOCATION: Forest90-00-10_Mada                        Fri Oct 10 09:30:06 2014|
+## |LOCATION: deforestmap                                Wed Mar  1 09:22:04 2017|
 ## |-----------------------------------------------------------------------------|
 ## |          north: 8682420    east: 1100820                                    |
 ## |REGION    south: 7155900    west:  298440                                    |
@@ -202,7 +202,7 @@ for (i in 1:14) {
   }
 }
 
-#= Export forest2010
+#= Export forest-cover for years 2000, 2005, 2010, 2014
 system("r.out.gdal --o input=for2014 createopt='compress=lzw,predictor=2' type=Byte output=outputs/for2014.tif")
 system("r.out.gdal --o input=for2010 createopt='compress=lzw,predictor=2' type=Byte output=outputs/for2010.tif")
 system("r.out.gdal --o input=for2005 createopt='compress=lzw,predictor=2' type=Byte output=outputs/for2005.tif")
@@ -214,16 +214,24 @@ system("r.out.gdal --o input=for2000 createopt='compress=lzw,predictor=2' type=B
 #= Integrating classes of forest in 1990 (111, 112, 122, 115, 152, 155, 755, 775 and 777) and uncertainties (555)
 system("r.mapcalc --o 'for1990_temp = if(harper==111 || harper==112 || harper==122 || \\
        harper==115 || harper==152 || harper==155 || harper==755 || \\
-       harper==775 || harper==777 || harper==555, 1, null())'")
+       harper==775 || harper==777, 1, if(harper==555, 5, null()))'")
[…]

### 2017-03-02 R/plotfcc.R
Plot ecoregions

@@ -159,6 +159,23 @@ plot_zoom_dist <- function(rast_file,palette=dist.palette,vr=dist.vr) {
 }
 
 # ================
+# Ecoregions
+
+# Import
+ecoregion <- readOGR(dsn="gisdata/vectors",layer="madagascar_ecoregion_redd_utm38s")
+plot(ecoregion)
+ecoregion.df <- tidy(ecoregion)
+
+# Plot
+plot.ecoregion <- ggplot(data=ecoregion.df,aes(x=long,y=lat,group=group,fill=id)) +
+  geom_polygon(colour=grey(0.4),size=0.2) +
+  theme_bw() + theme_base + theme(plot.margin=unit(c(-0.25,-0.25,-0.5,-0.5),"line")) +
+  coord_equal()
+
+## Grob
+grob.Mada <- ggplotGrob(plot.ecoregion)
+
+# ================
 # Maps
 
 ## for1953

### 2017-03-03 deforestmap.Rmd
Ecoregion plot

@@
 On the basis of the forest-cover maps we obtained, we computed the forest-cover area (in ha) and the mean annual deforestation rate (in hectares of forest and in percentage) on six decades, for five time periods from _c._ 1953 to 2014 (_c._ 1953-_c._ 1973-1990-2000-2010-2014).
-Mean annual deforestation rates in percentage were computed as follow [@Puyravaud]: $\theta = 100 \times [1-(1-(F_{t2}-F_{t1})/F_{t1})^{(1/(t2-t1))}]$, $\theta$ being the annual deforestation rate (in %/yr), $F_{t2}$ and $F_{t1}$ being the forest covers (reported in Tab.
+Mean annual deforestation rates in percentage were computed as follow [@Puyravaud2003]: $\theta = 100 \times [1-(1-(F_{t2}-F_{t1})/F_{t1})^{(1/(t2-t1))}]$, $\theta$ being the annual deforestation rate (in %/yr), $F_{t2}$ and $F_{t1}$ being the forest covers (reported in Tab.
 2) at times $t2$ and $t1$ (in years).
@@
 
-![Figure 1: **Forest-cover change on six decades from _c._ 1953 to 2014 in Madagascar.** Forest cover changes from _c._ 1973 to 2014 are shown in the main figure, and forest cover in _c._ 1953 is shown in the bottom-right inset.
+![Figure 1: **Ecoregions and forest types in Madagascar.** Madagascar can be divided into four climatic ecoregions with four forest types: the moist forest in the East (green), the dry forest in the West (orange), the spiny forest in the South (red), and the mangroves on the West coast (blue).
+Ecoregions were defined following climatic [@Cornet1974] and vegetation [@IEFN1996] criteria.
+The dark grey areas represent the remaining forest-cover for the year 2014.](outputs/ecoregion.png)
+
+![Figure 2: **Forest-cover change on six decades from _c._ 1953 to 2014 in Madagascar.** Forest cover changes from _c._ 1973 to 2014 are shown in the main figure, and forest cover in _c._ 1953 is shown in the bottom-right inset.
 Two zooms in the western dry (left part) and eastern moist (right part) ecoregions present more detailed views of (from top to bottom): forest-cover in 1950s, forest-cover change from _c._ 1973 to 2014, forest fragmentation in 2014 and distance to forest edge in 2014.

### 2017-03-03 R/plotfcc.R

@@ -161,19 +161,40 @@ plot_zoom_dist <- function(rast_file,palette=dist.palette,vr=dist.vr) {
 # ================
 # Ecoregions
 
+# devtools::install_github("paleolimbot/ggspatial")
+library(ggspatial)  # for geom_spatial()
+
 # Import
 ecoregion <- readOGR(dsn="gisdata/vectors",layer="madagascar_ecoregion_redd_utm38s")
-plot(ecoregion)
-ecoregion.df <- tidy(ecoregion)
+
+# Text
+xt <- c(850000,580000,530000,600000)
+yt <- c(7920000,8060000,7250000,8460000)
+t.df <- data.frame(text=c("Moist","Dry","Spiny","Mangroves"),x=xt,y=yt)
+
+# Segments
+seg.df <- data.frame(x=c(720000),y=c(8282000),
+                     xend=c(600000),yend=c(8405000))
+
+# Colors
+red.t <- adjustcolor("red",alpha.f=0.5)
+blue.t <- adjustcolor("blue",alpha.f=0.5)
+green.t <- adjustcolor("dark green",alpha.f=0.5)
+orange.t <- adjustcolor("orange",alpha.f=0.5)
+eco.col <- c("2"=green.t,"3"=orange.t,"0"=red.t,"1"="blue")
 
 # Plot
-plot.ecoregion <- ggplot(data=ecoregion.df,aes(x=long,y=lat,group=group,fill=id)) +
-  geom_polygon(colour=grey(0.4),size=0.2) +
-  theme_bw() + theme_base + theme(plot.margin=unit(c(-0.25,-0.25,-0.5,-0.5),"line")) +
+plot.ecoregion <- ggplot() +
+  geom_spatial(ecoregion, aes(x=long,y=lat,group=group,fill=factor(id))) +
+  scale_fill_manual(values=eco.col) +
+  geom_text(data=t.df, aes(x=x, y=y, label=text), size=9) +
+  geom_segment(data=seg.df, aes(x=x, xend=xend, y=y, yend=yend)) +
+  theme_bw() + theme_base +
+  theme(legend.position="null") +
+  scale_y_continuous(limits=c(7165000,8685000),expand=c(0,0)) +
+  scale_x_continuous(limits=c(300000,1100000),expand=c(0,0)) +
   coord_equal()
-
-## Grob
-grob.Mada <- ggplotGrob(plot.ecoregion)
[…]

### 2017-03-03 deforestmap.Rmd
Library Milano

@@
 ![Figure 1: **Ecoregions and forest types in Madagascar.** Madagascar can be divided into four climatic ecoregions with four forest types: the moist forest in the East (green), the dry forest in the West (orange), the spiny forest in the South (red), and the mangroves on the West coast (blue).
-Ecoregions were defined following climatic [@Cornet1974] and vegetation [@IEFN1996] criteria.
+Ecoregions were defined following climatic (see @Cornet1974) and vegetation (see @IEFN1996) criteria.
 The dark grey areas represent the remaining forest-cover for the year 2014.](outputs/ecoregion.png)
@@
 We observed a continuous decrease of the mean distance to forest edge from 1953 to 2014.
-The mean distance was of about 1km in 1973 while it was of about 300m in 2014.
+The mean distance was of ~1km in 1973 while it was of ~300m in 2014.
 Moreover, many areas were located far from forest edge (>2km and up to 4km) in 1973, while the majority of the forest was at a distance <1.2 km of the forest edge in 2014.](outputs/dist.png)

### 2017-03-03 R/plotfcc.R

@@ -151,7 +151,7 @@ plot_zoom_dist <- function(rast_file,palette=dist.palette,vr=dist.vr) {
                         breaks=c(1,500,1000,1500,2000),
                         labels=c(">0","0.5","1","1.5","2"),
                         guide=guide_colorbar(title="Distance (km)",title.position="top",
-                                            label.position="bottom", barwidth=10)) +
+                                            label.position="bottom",barwidth=10)) +
     theme_bw() + theme_zoom + coord_equal() +
     theme(legend.position="bottom", 
           legend.margin=margin(-1.5,-1,-1,-1,"lines"))
@@ -172,6 +172,9 @@ xt <- c(850000,580000,530000,600000)
 yt <- c(7920000,8060000,7250000,8460000)
 t.df <- data.frame(text=c("Moist","Dry","Spiny","Mangroves"),x=xt,y=yt)
 
+# Forest 2014  # If we want to add forest to ecoregions, to be done...
+# for2014 <- raster("outputs/for2014.tif")
+
 # Segments
 seg.df <- data.frame(x=c(720000),y=c(8282000),
                      xend=c(600000),yend=c(8405000))
@@ -187,14 +190,14 @@ eco.col <- c("2"=green.t,"3"=orange.t,"0"=red.t,"1"="blue")
 plot.ecoregion <- ggplot() +
   geom_spatial(ecoregion, aes(x=long,y=lat,group=group,fill=factor(id))) +
   scale_fill_manual(values=eco.col) +
-  geom_text(data=t.df, aes(x=x, y=y, label=text), size=9) +
-  geom_segment(data=seg.df, aes(x=x, xend=xend, y=y, yend=yend)) +
+  geom_text(data=t.df, aes(x=x, y=y, label=text), size=4) +
+  geom_segment(data=seg.df, aes(x=x, xend=xend, y=y, yend=yend), size=0.25) +
   theme_bw() + theme_base +
   theme(legend.position="null") +
   scale_y_continuous(limits=c(7165000,8685000),expand=c(0,0)) +
   scale_x_continuous(limits=c(300000,1100000),expand=c(0,0)) +
   coord_equal()
-ggsave("outputs/ecoregion.png", plot.ecoregion, width=5, height=9, units="in")
+ggsave("outputs/ecoregion.png", plot.ecoregion, width=5, height=9, units="cm")
 
 # ================
 # Maps

### 2017-04-19 deforestmap.R
Improve cloud stat by ecoregion

@@ -91,7 +91,7 @@ system("r.describe harper")
 # * 111 112 115 122 152 155 222 444 555 755 775 777
 # 1=forest, 2=nonforest, 5=cloud, 4=water, 7=mangrove
 
-system("r.report harper units=h")
+system("r.report --o harper units=h output='outputs/report_harper.txt'")
 
 ## +-----------------------------------------------------------------------------+
 ## |                         RASTER MAP CATEGORY REPORT                          |
@@ -125,6 +125,47 @@ system("r.report harper units=h")
 ## |TOTAL                                                            |122,484,912|
 ## +-----------------------------------------------------------------------------+
 
+# Import the GRASS report into R
+#r <- readLines("outputs/report_harper.txt")
+#r <- r[-c(1:15,28:31)]
+#writeLines(r, "outputs/report_harper.txt")
+report.harper <- read.table(file="outputs/report_harper.txt", sep="|", header=FALSE)
+report.harper <- report.harper[,c(2,4)]
+names(report.harper) <- c("code", "area")
+report.harper$area <- as.numeric(gsub(pattern=",", replacement="", report.harper$area))
+
+#====================================================================================
+# Report on clouds
+
+# Hectares of cloud in 2000 Harper map
+ha.cloud.2000 <- sum(report.harper$area[report.harper$code %in% c(152, 155, 555, 755)]) # 208192 ha
+
+# Ecoregions
+# 1: spiny, 2: mangroves, 3: moist, 4: dry
+# Rasterize ecoregions
+Extent <- "298440 7155900 1100820 8682420"
+Res <- "30"
+nodat <- "255"
+Input <- "gisdata/vectors/madagascar_ecoregion_tenaizy_38s.shp"
+Output <- "outputs/ecoregion.tif"
+# gdalwarp
+system(paste0("gdal_rasterize -a 'ecoregion' -ot Byte -a_nodata 255 -te ",Extent,
+              " -tr ",Res," ",Res," -co 'COMPRESS=LZW' -co 'PREDICTOR=2' ",Input," ",Output))
+# Import into grass
+system("r.in.gdal --o input=outputs/ecoregion.tif output=ecoregion")
+# Cross-tabulation table
+system("r.report --o map=ecoregion,harper units=h output='outputs/report_harper_ecoregion.txt'")
+
+# Moist ecoregions
+# Import the GRASS report into R
[…]

### 2017-04-19 deforestmap.R
Cloud estimates

@@ -158,11 +158,11 @@ system("r.report --o map=ecoregion,harper units=h output='outputs/report_harper_
 
 # Moist ecoregions
 # Import the GRASS report into R
-r <- readLines("outputs/report_harper_ecoregion.txt")
-r <- r[c(47:57)] # select outputs for the moist ecoregion
-writeLines(r, "outputs/report_harper.txt")
-report.harper <- read.table(file="outputs/report_harper.txt", sep="|", header=FALSE)
-report.harper <- report.harper[,c(2,4)]
+#r <- readLines("outputs/report_harper_ecoregion.txt")
+#r <- r[c(47:57)] # select outputs for the moist ecoregion
+#writeLines(r, "outputs/report_harper_ecoregion.txt")
+report.harper.moist <- read.table(file="outputs/report_harper_ecoregion.txt", sep="|", header=FALSE)
+report.harper.moist <- report.harper[,c(2,4)]
 names(report.harper) <- c("code", "area")
 report.harper$area <- as.numeric(gsub(pattern=",", replacement="", report.harper$area))
 

### 2017-04-20 deforestmap.Rmd
New table

@@
 Another recent map using more advanced statistical tool for classification (the Random Forest classifier, see @Grinand2013 and @Rakotomalala2015) has been released for the periods 2005-2010-2013 [@ONE2015].
-However, these maps might (i) be too old to give recent estimates of deforestation [@MEFT2009, @ONE2013], (ii) figure large areas of clouds (images with more than 20% of clouds in 2010 for the map provided by [@ONE2013]), or (iii) show large errors of classification in specific areas, especially in the dry and spiny forests for which the spectral answer varies strongly due to deciduousness associated with the higher seasonality than in moist forests (Kappa<0.6 and overall accuracy<0.8 for the dry and spiny forests for the maps produced by @ONE2015).
-Moreover, deriving such forest maps from supervised classification requires a huge amount of time and work, especially regarding image selection to reduce cloud cover and visual interpretation of the numerous observation polygons used to train the classification algorithm.
+However, these maps (i) are too old to give recent estimates of deforestation [@MEFT2009, @ONE2013], (ii) figure large areas of clouds (images with more than 20% of clouds in 2010 for the map provided by [@ONE2013]), or (iii) show large errors of classification in specific areas, especially in the dry and spiny forests for which the spectral answer varies strongly due to deciduousness associated with the higher seasonality than in moist forests (Kappa<0.6 and overall accuracy<0.8 for the dry and spiny forests for the maps produced by @ONE2015).
+Moreover, deriving such forest maps from supervised classification requires a huge amount of time and work, especially regarding image selection to reduce cloud cover and visual interpretation of the numerous observation polygons used to train the classification algorithm [@Rakotomalala2015].
 In addition, visual interpretation might be subjective and is potentially a source of errors.
@@
 tab.1 <- defor.df
-names(tab.1) <- c("Year","Forest (ha)", "Annual defor. (ha/yr)", "theta (%/yr)")
+names(tab.1) <- c("Year", "Forest (ha)", "Clouds (ha)", "Annual defor. (ha/yr)", "theta (%/yr)")
 kable(tab.1)
@@
 
-Table 1: **Evolution of the forest cover and deforestation intensity from 1950s to 2014 in Madagascar**.
+Table 1: **Evolution of the forest cover and deforestation rates from 1950s to 2014 in Madagascar**.
 Madagascar has lost `r perc.1`% of its natural forest in 40 years between _c._ 1973 and 2014 (and about `r perc.2`% in 60 years since _c._ 1953).
-If the deforestation has progressively slowed down from _c._ 1973 to 2010 (from `r d.1` to `r d.2`%/yr), most likely due to a decrease of available arable lands under the remaining forests and to a decrease of forest accessibility, the deforestation has recently increased, with a doubling deforestation on the period 2010-2014 compared to 2000-2010.
+Deforestation has slew down after 1990 (from `r d.1` in 1973-1990 to `r d.2`%/yr in 2000-2010), most likely due to political changes (end of the Second Republic of Madagascar in 1992), a decrease of available arable lands under the remaining forests and to a decrease of forest accessibility.
+But more recently, deforestation has increased, with a doubling deforestation on the period 2010-2014 compared to 2000-2010.
 This increase could be attributed to both political instability (political crisis in 2009) and to an explosion of the demography (according to the @UN2015, population has increased from about 16M inhabitants in 2000 to 24M in 2014).
-Natural forest covers 8.9 Mha in 2014 and current (period 2000-2014) mean annual deforestation rate is of about 70,000 ha/yr.
+Natural forests cover 8.9 Mha in 2014 and current (period 2000-2014) mean annual deforestation rate is of about 70000 ha/yr.
 
@@
 Table 2: **Comparison of Madagascar forest-cover estimates (in ha) with previous studies on the period 1953-2014**.
-We found similiar forest-cover estimates compared to previous studies [@Harper2007;@MEFT2009;@ONE2015].
-The increase in mangrove and spiny forest covers from _c._ 1953 to _c._ 1973 in @Harper2007 and our study is most probably due to differences in forest definition and mapping methods.
+Estimates of the forest-cover are given without considering the cloud cover and are then subject to uncertainty.
+We found forest-cover estimates comparable to previous studies [@Harper2007;@MEFT2009;@ONE2015].
+The increase in mangrove and spiny forest covers from _c._ 1953 to _c._ 1973 in @Harper2007 and our study is most probably due to differences in forest definition and mapping methods between the 1953 aerial-photography derived map and the 1973 Landsat image derived map.
 Our study extends the forest-cover monitoring on a six decades period (from 1953 to 2014) while harmonizing the data from the previous studies.

### 2017-04-20 deforestmap.R

@@ -43,6 +43,10 @@ loaded <- lapply(pkg,load.pkg)
 rm(pkg,load.pkg,loaded)
 
 ##=====================================
+## Name of objects to save
+SavedObjects <- c()
+
+##=====================================
 ## Create new grass location in UTM 38S
 #dir.create("grassdata")
 #system("grass72 -c epsg:32738 grassdata/deforestmap")  # Ignore errors
@@ -66,7 +70,7 @@ dir.create("outputs")
 
 #= Projection UTM38S (EPSG:32738)
 
-#= Reproject CI map
+#= Reproject Harper 1990-2000-2005 map
 
 # Region
 Extent <- "298440 7155900 1100820 8682420"
@@ -125,6 +129,10 @@ system("r.report --o harper units=h output='outputs/report_harper.txt'")
 ## |TOTAL                                                            |122,484,912|
 ## +-----------------------------------------------------------------------------+
 
+#====================================================================================
+# Report on clouds
+
+#= Whole forest
 # Import the GRASS report into R
 #r <- readLines("outputs/report_harper.txt")
 #r <- r[-c(1:15,28:31)]
@@ -134,13 +142,11 @@ report.harper <- report.harper[,c(2,4)]
 names(report.harper) <- c("code", "area")
 report.harper$area <- as.numeric(gsub(pattern=",", replacement="", report.harper$area))
 
-#====================================================================================
-# Report on clouds
-
 # Hectares of cloud in 2000 Harper map
 ha.cloud.2000 <- sum(report.harper$area[report.harper$code %in% c(152, 155, 555, 755)]) # 208192 ha
+SavedObjects <- c(SavedObjects, "ha.cloud.2000")
 
-# Ecoregions
+#= By ecoregion
 # 1: spiny, 2: mangroves, 3: moist, 4: dry
 # Rasterize ecoregions
 Extent <- "298440 7155900 1100820 8682420"
@@ -162,12 +168,17 @@ system("r.report --o map=ecoregion,harper units=h output='outputs/report_harper_
 #r <- r[c(47:57)] # select outputs for the moist ecoregion
 #writeLines(r, "outputs/report_harper_ecoregion.txt")
[…]

### 2017-04-21 deforestmap.Rmd
New tab of results

@@
 Table 2: **Comparison of Madagascar forest-cover estimates (in ha) with previous studies on the period 1953-2014**.
-Estimates of the forest-cover are given without considering the cloud cover and are then subject to uncertainty.
 We found forest-cover estimates comparable to previous studies [@Harper2007;@MEFT2009;@ONE2015].
@@
 Annual deforestation rates are providing in ha/yr and in %/yr (second number in parenthesis).
-Original estimates of deforestation rates might differ from original publications [@Harper2007;@MEFT2009;@ONE2015].
-In our study, we used the following formula [@Puyravaud2003] to compute the annual deforestation rates: $\theta = 100 \times [1-(1-(F_{t2}-F_{t1})/F_{t1})^{(1/(t2-t1))}]$, with $\theta$ the annual deforestation rate (in %/yr), $F_{t2}$ and $F_{t1}$ the forest covers (reported in Tab.
+For our study, we used the following formula [@Puyravaud2003] to compute the annual deforestation rates: $\theta = 100 \times [1-(1-(F_{t2}-F_{t1})/F_{t1})^{(1/(t2-t1))}]$, with $\theta$ the annual deforestation rate (in %/yr), $F_{t2}$ and $F_{t1}$ the forest covers (reported in Tab.
 2) at times t2 and t1 (in years).

### 2017-04-21 deforestmap.R

@@ -694,9 +694,6 @@ defor.comp[defor.comp$Source=="this study",3:8] <- defor.comp.thistudy[defor.com
 # Save results
 write.table(defor.comp,file="outputs/defor_comp.txt",row.names=FALSE)
 
-# Correction for 1953-1973 due to clouds ? TODO
-
-
 ##========================
 ## Forest-cover change map
 ##========================

### 2017-04-21 deforestmap.Rmd
Comparison

@@
 The _c._ 1953 forest map was derived from the visual interpretation of aerial photographies and gives a coarse representation (scale 1/1,000,000) of the forest extent in Madagascar for the 1950s.
-Forest maps for the years _c._ 1973, 1990 and 2000, were obtained from a supervised classification of Landsat satellite images at 30m and can be used to derive much more accurate estimates of the forest cover at these three dates.
+Forest maps for the years _c._ 1973, 1990 and 2000, were obtained from a supervised classification of Landsat satellite images at 60 (for the year 1973) and 30m (for years 1990 and 2000) and can be used to derive much more accurate estimates of the forest cover at these three dates.
 Maps provided by @Harper2007 were not free of clouds (11 244 km2 of unknown cover type for the year 2000 for example).
@@
 Another recent map using more advanced statistical tool for classification (the Random Forest classifier, see @Grinand2013 and @Rakotomalala2015) has been released for the periods 2005-2010-2013 [@ONE2015].
-However, these maps (i) are too old to give recent estimates of deforestation [@MEFT2009, @ONE2013], (ii) figure large areas of clouds (images with more than 20% of clouds in 2010 for the map provided by [@ONE2013]), or (iii) show large errors of classification in specific areas, especially in the dry and spiny forests for which the spectral answer varies strongly due to deciduousness associated with the higher seasonality than in moist forests (Kappa<0.6 and overall accuracy<0.8 for the dry and spiny forests for the maps produced by @ONE2015).
+However, these maps (i) are too old to give recent estimates of deforestation [@MEFT2009, @ONE2013], (ii) figure large areas of clouds (images with more than 20% of clouds in 2010 for the map provided by [@ONE2013]), or (iii) show large errors of classification in specific areas, especially in the dry and spiny forests, for which the spectral answer varies strongly due to deciduousness associated with the higher seasonality than in moist forests (Kappa<0.6 and overall accuracy<0.8 for the dry and spiny forests for the maps produced by @ONE2015).
 Moreover, deriving such forest maps from supervised classification requires a huge amount of time and work, especially regarding image selection to reduce cloud cover and visual interpretation of the numerous observation polygons used to train the classification algorithm [@Rakotomalala2015].
@@
 
-We derived 30m cloud-free forest maps for Madagascar for the years 1990, 2000, 2010 and 2014 combining the 1990 and 2000 forest maps by @Harper2007 and the 2000-2014 tree cover and tree cover loss maps by @Hansen2013.
-On the 2000 Harper's map, 200,000 ha clouds are present over the 4.2 million ha moist forest ecoregion (4.8% of clouds).
-To remove these clouds, we used the 2000 cloud-free tree cover map provided by @Hansen2013 (also at 30m resolution) and we chose a threshold of 75% of tree cover [@Achard2014] to decide whether to replace clouds with forest or non-forest.
+We derived 30m cloud-free annual forest maps for Madagascar for the period 2000-2014 combining the 2000 forest map by @Harper2007 and the 2000-2014 tree cover and tree cover loss maps by @Hansen2013.
+On the 2000 Harper's forest map, `r ha.cloud.2000` ha of clouds are present, mostly (`r perc.cloud.moist`) over the `r round(ha.forest.2000.moist/1e6,2)` million ha moist forest ecoregion.
+To remove these clouds, we used the 2000 cloud-free tree cover map provided by @Hansen2013 (at 30m resolution) and we chose a threshold of 75% of tree cover [@Achard2014] to decide whether to replace clouds with forest or non-forest.
 We thus obtained a cloud-free year 2000 forest-cover map for Madagascar.
 From this map, using tree cover loss data from 2000 to 2014 provided by @Hansen2013, we obtained annual cloud-free forest-cover maps from 2010 to 2014 at 30m resolution.
-We corrected the cloudy 1990 Harper forest map assuming that if forest was present in 2000, it was also present in 1990.
-Doing so, cloud-cover was reduced to a negligeable quantity of `r epsilon_1990`ha.
+We then corrected the cloudy 1990 Harper forest map assuming that if forest was present in 2000, it was also present in 1990.
+Doing so, cloud-cover was reduced to a negligeable quantity of `r residual.clouds.1990` ha.
 For these residual cloud areas, we assumed that forest-cover was the same as for the year 2000.
-We also corrected the cloudy 1973 Harper forest map assuming that if forest was present in 1973, it was also present in 1973.
-Contrary to the year 1990, we were not able to remove all the clouds on the 1973 map and the remaining cloud cover was of `r round(cloud_1973/10e6,2)` million ha.
+We also corrected the cloudy 1973 Harper forest map assuming that if forest was present in 1990, it was also present in 1973.
+Contrary to the year 1990, we were not able to remove all the clouds on the 1973 map and the remaining cloud cover was of `r round(cloud.1973/1e6,2)` million ha.
 Because the 1953 forest-cover map was obtained from the digitalization of a paper map derived from aerial photographies, and thus cannot be perfectly aligned with the maps obtained in the following years, we did not performed any modifications on the original forest map of 1953 except reprojection.
@@
 
-### 2.2 Computing forest-cover area and deforestation rate and comparing estimates with previous studies
+### 2.2 Computing forest-cover area and deforestation rate
 
-On the basis of the forest-cover maps we obtained, we computed the forest-cover area (in ha) and the mean annual deforestation rate (in hectares of forest and in percentage) on six decades, for five time periods from _c._ 1953 to 2014 (_c._ 1953-_c._ 1973-1990-2000-2010-2014).
-Mean annual deforestation rates in percentage were computed as follow [@Puyravaud2003]: $\theta = 100 \times [1-(1-(F_{t2}-F_{t1})/F_{t1})^{(1/(t2-t1))}]$, $\theta$ being the annual deforestation rate (in %/yr), $F_{t2}$ and $F_{t1}$ being the forest covers (reported in Tab.
-2) at times $t2$ and $t1$ (in years).
-Because of the large cloud-cover in 1973, the annual deforestation rates in hectares of forest for the two periods _c._ 1953-1973 and 1973-1990 are only rough estimates of the deforestation rates.
-This problem was overcome when computing the annual deforestation rates in percentage for these two periods.
-In this case, we only considered the cover change and computed the ratio $F_{t2}-F_{t1})/F_{t1}$ for pixels that were not covered by clouds in 1973.
+On the basis of the forest-cover maps we obtained, we computed the forest-cover area (in ha) and the mean annual deforestation rate (in ha/yr and in %/yr) on six decades, for five time periods from _c._ 1953 to 2014 (_c._ 1953-_c._ 1973-1990-2000-2010-2014).
+Mean annual deforestation rates in percentage were computed as follow [@Puyravaud2003; @Vieilledent2013]: $\theta = 100 \times [1-(1-(F_{t2}-F_{t1})/F_{t1})^{(1/(t2-t1))}]$.
+In this formula, $\theta$ is the annual deforestation rate (in %/yr), $F_{t2}$ and $F_{t1}$ are the forest covers free of clouds at both dates $t2$ and $t1$, and $t2-t1$ is the time-interval (in years) between the two dates.
+Because of the large cloud-cover (`r round(cloud.1973/1e6,2)` million ha) in 1973, the annual deforestation rates in hectares of forest for the two periods _c._ 1953-1973 and 1973-1990 are only rough estimates of the deforestation rates in number of hectares.
+This problem was overcome when computing the annual deforestation rates in %/yr for these two periods.
+In this case, we only considered the cover change and computed the ratio $F_{t2}-F_{t1})/F_{t1}$ considering only the pixels that were not covered by clouds in 1973.
 Computation were done at the national scale and for each forest type in Madagascar: the moist forest in the East, the dry forest in the West, the spiny forest in the South and mangroves on the West coast.
@@
 Ecoregions were defined on the basis of climatic and vegetation criteria using the climate classification by @Cornet1974 and the vegetation classification from the 1996 IEFN national forest inventory [@IEFN1996].
+Because mangrove forests are highly dynamic ecosystems that can expand or contract on decadal scales depending on environmental factors [@Armitage2015], a fixed delimitation of the mangrove ecoregion might not be appropriate.
+As a consequence, our estimates of the forest-cover and deforestation rates for mangroves in Madagascar must be considered with caution.
 
-We compared our estimates of forest-cover and forest-cover change with three previous studies that have estimated forest-cover and forest-cover change at the national scale for Madagascar [@Harper2007; @MEFT2009; @ONE2015].
-@Harper2007 provided estimates for the periods _c._ 1953-_c._ 1973-1990-2000, @MEFT2009 provided estimates for the periods 1990-2000-2005 and @ONE2015 provided estimates for the periods 2005-2010-2013.
-The same ecoregion boundaries as in our study have been used in the @ONE2015 report but this was not the case for the @Harper2007 and @MEFT2009 reports, which can explain some of the differences between studies in the forest-cover areas and deforestation rates by ecoregions.
-W
+### 2.3 Comparing forest-cover and deforestation rate estimates with previous studies
+
+We compared our estimates of forest-cover and deforestation rate with three previous studies at the national scale for Madagascar [@Harper2007; @MEFT2009; @ONE2015].
+@Harper2007 provided forest-cover estimates for the periods _c._ 1953-_c._ 1973-1990-2000, @MEFT2009 provided estimates for the periods 1990-2000-2005 and @ONE2015 provided estimates for the periods 2005-2010-2013.
+To be able to compare on the same time-periods, we thus considered three additional time-periods to compute forest-covers and deforestation rates from the results our study: 2000-2005, 2005-2010 and 2010-2013.
+All previous studies provide annual deforestation rate estimates in %/yr.
+The way annual deforestation rates in %/yr have been computed in these previous studies is not always detailed and might slightly differ from one study to another [see @Puyravaud2003].
+But these estimates always correct for the potential presences of residuals clouds on forest maps.
+@Harper2007 also provide deforestation rate estimates in ha but only for the two periods 1973-1990 and 1990-2000.
+We converted these estimates into annual deforestation rate estimates in ha/yr.
+When annual deforestation rate estimates in ha/yr were not reported (for 1953-1973 in @Harper2007 and in @MEFT2009 and @ONE2015), we computed them from the provided forest-cover estimates in each study.
+These estimates do not correct for the potential presence of residual clouds.
+Forest-cover and deforestation rates were compared between studies for the whole Madagascar and for the four ecoregions.
+The same ecoregion boundaries as in our study have been used in the @ONE2015 report but this was not the case for the @Harper2007 article and the @MEFT2009 report, which can explain some of the differences between studies in the forest-cover areas and deforestation rates by ecoregions.
 
@@
 ```{r defor, results="asis"}
+defor.df <- read.table(file="outputs/defor.txt", sep="\t", header=TRUE)
 perc.1 <- round(100*(defor.df$area[defor.df$Year==1973]-defor.df$area[defor.df$Year==2014])/defor.df$area[defor.df$Year==1973])
@@
 ```{r fcc_comp, results="asis"}
-tab.2 <- df.comp
+fcc.comp <- read.table(file="outputs/fcc_comp.txt", sep=" ", header=TRUE)
+tab.2 <- fcc.comp
 tab.2 <- tab.2[,c(2,1,3:10)]
@@
 Annual deforestation rates are providing in ha/yr and in %/yr (second number in parenthesis).
-For our study, we used the following formula [@Puyravaud2003] to compute the annual deforestation rates: $\theta = 100 \times [1-(1-(F_{t2}-F_{t1})/F_{t1})^{(1/(t2-t1))}]$, with $\theta$ the annual deforestation rate (in %/yr), $F_{t2}$ and $F_{t1}$ the forest covers (reported in Tab.
-2) at times t2 and t1 (in years).
+Exact same numbers as in scientific articles and reports from previous studies have been reported for deforestation rate in %/yr.
+The way annual deforestation rates in %/yr have been computed in these previous studies can slightly differ from one study to another but estimates always correct for the potential presences of clouds on forest maps.
+On the contrary, deforestation rates in ha/yr have been recomputed from forest-cover estimates in Tab.
+2 and do not correct for the potential presence of clouds.
 
 ```{r frag, results="asis"}
+frag.df <- read.table(file="outputs/frag.txt", sep="\t", header=TRUE)
 tab.4 <- frag.df

### 2017-04-21 deforestmap.R

@@ -352,7 +352,7 @@ system("r.out.gdal --o input=for1953 createopt='compress=lzw,predictor=2' \\
 # Remove salt and pepper
 # =======================================
 
-Year <- c(1953,1973,1990,2000,2010,2014)
+Year <- c(1953,1973,1990,2000,2005,2010,2013,2014)
 system("g.region rast=harper -ap")
 system("r.mask --o raster=harper")
 for (i in 1:length(Year)) {
@@ -637,6 +637,8 @@ for (j in 1:3) {
 ## Forest-cover
 # Historical data
 fcc.comp <- read.csv("data/fcc_comp.csv",header=TRUE)
+ha.forest.2000.moist <- fcc.comp$y2000[fcc.comp$Source=="Harper2007" & fcc.comp$ForestType=="Moist"]
+SavedObjects <- c(SavedObjects,"ha.forest.2000.moist")
 # This study
 defor_for_comp <- read.table("outputs/defor_for_comp.txt",header=TRUE)
 defor_moist_for_comp <- read.table("outputs/defor_moist_for_comp.txt",header=TRUE)
@@ -663,6 +665,7 @@ fcc.comp$y2014[fcc.comp$ForestType=="Spiny" & fcc.comp$Source=="this study"] <- 
 fcc.comp$y2014[fcc.comp$ForestType=="Mangroves" & fcc.comp$Source=="this study"] <- defor_mangroves$area[defor_mangroves$Year==2014]
 # Save results
 write.table(fcc.comp,file="outputs/fcc_comp.txt",row.names=FALSE)
+SavedObjects <- c(SavedObjects,"fcc.comp")
 
 ## Deforestation
 # Historical data
@@ -693,6 +696,7 @@ defor.comp.thistudy[defor.comp.thistudy=="NA (NA)"] <- NA
 defor.comp[defor.comp$Source=="this study",3:8] <- defor.comp.thistudy[defor.comp.thistudy$Source=="this study",3:8]
 # Save results
 write.table(defor.comp,file="outputs/defor_comp.txt",row.names=FALSE)
+SavedObjects <- c(SavedObjects,"defor.comp")
 
 ##========================
 ## Forest-cover change map
@@ -787,7 +791,7 @@ dev.off()
 ##========================
 
 # load("deforestmap.rda")
-save(list=ls(all.names=TRUE),file="deforestmap.rda")
+save(list=SavedObjects,file="deforestmap.rda")
 
 ##========================
 ## Knit the document

### 2017-05-02 deforestmap.Rmd
Check methodo

@@
 We derived 30m cloud-free annual forest maps for Madagascar for the period 2000-2014 combining the 2000 forest map by @Harper2007 and the 2000-2014 tree cover and tree cover loss maps by @Hansen2013.
-On the 2000 Harper's forest map, `r ha.cloud.2000` ha of clouds are present, mostly (`r perc.cloud.moist`) over the `r round(ha.forest.2000.moist/1e6,2)` million ha moist forest ecoregion.
+On the 2000 Harper's forest map, `r as.integer(ha.cloud.2000)` ha of clouds are present, mostly (`r round(perc.cloud.moist)`%) over the moist forest which cover `r round(ha.forest.2000.moist/1e6,2)` million ha.
 To remove these clouds, we used the 2000 cloud-free tree cover map provided by @Hansen2013 (at 30m resolution) and we chose a threshold of 75% of tree cover [@Achard2014] to decide whether to replace clouds with forest or non-forest.
@@
 On the basis of the forest-cover maps we obtained, we computed the forest-cover area (in ha) and the mean annual deforestation rate (in ha/yr and in %/yr) on six decades, for five time periods from _c._ 1953 to 2014 (_c._ 1953-_c._ 1973-1990-2000-2010-2014).
-Mean annual deforestation rates in percentage were computed as follow [@Puyravaud2003; @Vieilledent2013]: $\theta = 100 \times [1-(1-(F_{t2}-F_{t1})/F_{t1})^{(1/(t2-t1))}]$.
-In this formula, $\theta$ is the annual deforestation rate (in %/yr), $F_{t2}$ and $F_{t1}$ are the forest covers free of clouds at both dates $t2$ and $t1$, and $t2-t1$ is the time-interval (in years) between the two dates.
-Because of the large cloud-cover (`r round(cloud.1973/1e6,2)` million ha) in 1973, the annual deforestation rates in hectares of forest for the two periods _c._ 1953-1973 and 1973-1990 are only rough estimates of the deforestation rates in number of hectares.
+Mean annual deforestation rates in percentage were computed as follow [@Puyravaud2003; @Vieilledent2013]:
+
+$$\theta = 100 \times [1-(1-(F_{t_2}-F_{t_1})/F_{t_1})^{(1/(t_2-t_1))}]$$
+
+In this formula, $\theta$ is the annual deforestation rate (in %/yr), $F_{t_2}$ and $F_{t_1}$ are the forest cover free of clouds at both dates $t_2$ and $t_1$, and $t_2-t_1$ is the time-interval (in years) between the two dates.
+Because of the large cloud-cover (`r round(cloud.1973/1e6,2)` million ha) in 1973, the annual deforestation rates in hectares of forest for the two periods _c._ 1953-1973 and 1973-1990 are only coarse estimates.
 This problem was overcome when computing the annual deforestation rates in %/yr for these two periods.
-In this case, we only considered the cover change and computed the ratio $F_{t2}-F_{t1})/F_{t1}$ considering only the pixels that were not covered by clouds in 1973.
+In this case, we computed the ratio $(F_{t_2}-F_{t_1})/F_{t_1}$ considering only the forest pixels that were not covered by clouds in 1973.
 Computation were done at the national scale and for each forest type in Madagascar: the moist forest in the East, the dry forest in the West, the spiny forest in the South and mangroves on the West coast.
@@
 Ecoregions were defined on the basis of climatic and vegetation criteria using the climate classification by @Cornet1974 and the vegetation classification from the 1996 IEFN national forest inventory [@IEFN1996].
-Because mangrove forests are highly dynamic ecosystems that can expand or contract on decadal scales depending on environmental factors [@Armitage2015], a fixed delimitation of the mangrove ecoregion might not be appropriate.
+Because mangrove forests are highly dynamic ecosystems that can expand or contract on decadal scales depending on environmental factors [@Armitage2015], a fixed delimitation of the mangrove ecoregion on six decades might not be appropriate.
 As a consequence, our estimates of the forest-cover and deforestation rates for mangroves in Madagascar must be considered with caution.
@@
 @Harper2007 provided forest-cover estimates for the periods _c._ 1953-_c._ 1973-1990-2000, @MEFT2009 provided estimates for the periods 1990-2000-2005 and @ONE2015 provided estimates for the periods 2005-2010-2013.
-To be able to compare on the same time-periods, we thus considered three additional time-periods to compute forest-covers and deforestation rates from the results our study: 2000-2005, 2005-2010 and 2010-2013.
+To be able to compare the forest-cover and deforestation rate estimates on the same time-periods, we considered three additional time-periods in our study: 2000-2005, 2005-2010 and 2010-2013.
 All previous studies provide annual deforestation rate estimates in %/yr.
@@
 But these estimates always correct for the potential presences of residuals clouds on forest maps.
-@Harper2007 also provide deforestation rate estimates in ha but only for the two periods 1973-1990 and 1990-2000.
+@Harper2007 also provide deforestation rate estimates in total number of hectares but only for the two periods 1973-1990 and 1990-2000.
 We converted these estimates into annual deforestation rate estimates in ha/yr.
@@
 These estimates do not correct for the potential presence of residual clouds.
-Forest-cover and deforestation rates were compared between studies for the whole Madagascar and for the four ecoregions.
+Forest-cover and deforestation rates were compared between studies for the whole Madagascar and the four ecoregions.
 The same ecoregion boundaries as in our study have been used in the @ONE2015 report but this was not the case for the @Harper2007 article and the @MEFT2009 report, which can explain some of the differences between studies in the forest-cover areas and deforestation rates by ecoregions.

### 2017-05-02 deforestmap.R

@@ -174,7 +174,7 @@ report.harper.moist$area <- as.numeric(gsub(pattern=",", replacement="", report.
 
 # Hectares of cloud in 2000 Harper map fr moist ecoregion
 ha.cloud.2000.moist <- sum(report.harper.moist$area[report.harper.moist$code %in% c(152, 155, 555, 755)]) # 182650 ha
-perc.cloud.moist <- 100*ha.cloud.2000.moist/ha.cloud.2000 # 87.7%
+perc.cloud.moist <- 100*ha.cloud.2000.moist/ha.cloud.2000 # 88%
 SavedObjects <- c(SavedObjects, "ha.cloud.2000.moist", "perc.cloud.moist")
 
 #======================================================================================================
@@ -792,6 +792,7 @@ dev.off()
 
 # load("deforestmap.rda")
 save(list=SavedObjects,file="deforestmap.rda")
+save(list=ls(),file="deforestmap.rda")
 
 ##========================
 ## Knit the document

### 2017-05-03 deforestmap.Rmd
Start discussion part

@@
 In this case, we computed the ratio $(F_{t_2}-F_{t_1})/F_{t_1}$ considering only the forest pixels that were not covered by clouds in 1973.
-Computation were done at the national scale and for each forest type in Madagascar: the moist forest in the East, the dry forest in the West, the spiny forest in the South and mangroves on the West coast.
+Computation were done at the national scale and for each forest type in Madagascar: the moist forest in the East, the dry forest in the West, the spiny forest in the South and mangroves on the West coast (Fig.
+1).
 To define the forest type, we used a map provided by the MEFT (_"Ministère de l'Environnement et des Forêts à Madagascar"_) showing the boundaries of the four ecoregions in Madagascar.
@@
 To be able to compare the forest-cover and deforestation rate estimates on the same time-periods, we considered three additional time-periods in our study: 2000-2005, 2005-2010 and 2010-2013.
+We computed the Pearson's correlation coefficient and the root mean square error (RMSE) between our forest-cover estimates and forest-cover estimates from previous studies at the same date.
 All previous studies provide annual deforestation rate estimates in %/yr.
@@
 
+Natural forests in Madagascar covered 16.0 Mha in 1953, about 27% of the national territory of 587041 km<sup>2</sup>.
+In 2014, this number has dropped down to 8.9 Mha, about 15% of the national territory (Fig.
+2 and Tab.
+1).
+Madagascar has then lost 44% of its natural forest in 60 years since c.
+1953 and a minimum of 37% in 40 years between c.
+1973 and 2014 (Fig.
+2 and Tab.
+1).
+In 2014, the 8.9 Mha of natural forest in Madagascar were divided into 4.4 Mha (50%) of moist forest, 2.6 Mha (29%) of dry forest, 1.7 Mha of spiny forest (19%) and 176778 ha (2%) of mangrove forest (Fig.
+1 and Tab.
+2).
+Regarding the deforestation trend, we observed a progressive decrease of the deforestation rate after 1990 from 204712 ha/yr (1.63%/yr) on the period 1973-1990 to 44296 ha/yr (0.43%/yr) on the period 2000-2005 (Tab.
+1).
+Since 2005, deforestation has progressively increased in Madagascar and deforestation rate has more than doubled on the period 2010-2014 (98651 ha/yr, 1.08%/yr) compared to 2000-2005 (Tab.
+1).
+The deforestation trend characterized by a progressive decrease of the deforestation rate on the period 1990-2005 and a progressive increase of the deforestation after 2005 is observable for the four ecoregions (Tab.
+3).
+One exception for the spiny forest for the period 2010-2013 for which we estimated a lower deforestation rate compared to 2005-2010 (Tab.
+3).
+
 ### 3.2 Comparison with previous forest-cover change studies in Madagascar
+
+Forest-cover maps provided by previous studies [@Harper2007; @MEFT2009; @ONE2015] at the national scale for Madagascar were not free of clouds.
+In @Harper2007, 5003 and 11244 km2 of unknown cover type were present for the years 1990 and 2000 respectively.
+Percentage of clouds are not reported in the two other previous studies by @MEFT2009 and @ONE2015.
+With our approach, we were able to obtain cloud-free forest-cover change maps from 1990 to 2014 (Tab.
+1) and thus provide more accurate estimates of forest-cover and deforestation rates on this period.
+Forest-cover estimates in our study were similar (correlation of 0.99) to estimates from the three previous studies (Tab.
+2).
+We estimated a RMSE of 300000 ha (6% of the mean forest cover of 4.8 Mha) between our forest-cover estimates and estimates from previous studies.
+A part of these differences can be attributed to differences in ecoregion boundaries between studies.
+Despite the substantial differences in the deforestation rates estimates between studies (Tab.
+3), the same deforestation trend (decrease of the deforestation rate on the period 1990-2005 followed by a progressive increase of the deforestation after 2005) was observed accross studies.
 
@@
 
+Accompagnying deforestion, forest fragmentation has progressively increased since 1953 in Madagascar.
+We observed a continuous decrease of the mean distance to forest edge from 1953 to 2014 in Madagascar (Fig.
+3).
+The mean distance to forest edge has decreased to ~300m in 2014 while it was previously of ~1km in 1973 (Fig.
+3).
+Moreover, a large part of the natural forest was located far from forest edge (>2km and up to 4km) in 1973, while the majority of the forest was at a distance <1.2 km of the forest edge in 2014 (Fig.
+3).
+The percentage of forest that can be considered intact in Madagascar has progressively decreased since 1953.
+The percentage of forest belonging to the "interior"" category (the most intact forest) has fallen from 68% in 1973 to 50% in 2014 (Tab.
+4).
+In 2014, more than 16% of the forest belonged to the "patch" and "transitional" categories (isolated small forest patches) compared to 9.5% in 1973 (Tab.
+4).
+
 ## 4. Discussion
 
-### 4.1 Evolution of deforestation and forest fragmentation in Madagascar on the last 60 years
+### 4.1 Evolution of forest-cover in Madagascar over the last 60 years
+
+There has been a long debate about the statement saying that Madagascar has lost 90% of its forest since the arrival of the humans on the island [@Virah-Sawmy2009, @Agarwal2005].
+What it sure now, is that Madagascar has lost around 50% of its natural forest in six decades since 1953.
 
@@
 
-Should be used to actualize FAO report.
+The three previous studies have used (i) different forest definition (associated to independant visual interpretations of observation polygons to train the classifier), (ii) different classification algorithms (Maximum Likelihood Classification for @Harper2007 and @MEFT2009 vs.
+Random Forests for @ONE2015), and (iii) do not systematically correct for the presence of clouds.
+Despite these differences, the four studies gave similar forest-cover estimates and show the same deforestation trend over Madagascar, with a decrease of the deforestation rate on the period 1990-2005 followed by a progressive increase of the deforestation rate after 2005.
+Given the fact that our deforestation rate estimates have been computed from cloud-free maps, our estimates can be considered more accurate compared to estimates from previous studies, in particular for deforestation rates expressed in number of hectares per year.
+Current rates of deforestation for Madagascar (98651 ha/yr, 1.08 %/yr, if we consider the period 2010-2014) could be used to build reference scenarios for deforestation in Madagascar [@Olander2008].
+Forest-cover and deforestation rate estimates from this study could also be used to update the estimates in the upcoming FAO Forest Resources Assessment reports [@Keenan2015].
 
-There has been a long debate about whether Madagascar has lost 90% of its forest since the arrival of the humans on the island [@Virah-Sawmy2009, @Agarwal2005].
-What it sure now, is that Madgascar has lost around 50% of its natural forest in six decades since _c._ 1953.
+### 4.2 Consequences of deforestation and forest fragmentation on biodiversity and carbon emissions
 
-### 4.2 Interest of such an approach and limitations
+Increase in forest fragmentation has a great impact on both carbon emissions [@Putz2014] and biodiversity loss [@Gibson2013, @Murcia1995].
+Biodiversity loss, see @Allnutt2008.
+
+### 4.3 Interest of such an approach and limitations
 
@@
 Table 2: **Comparison of Madagascar forest-cover estimates (in ha) with previous studies on the period 1953-2014**.
-We found forest-cover estimates comparable to previous studies [@Harper2007;@MEFT2009;@ONE2015].
+We found forest-cover estimates similar to previous studies [@Harper2007;@MEFT2009;@ONE2015].
 The increase in mangrove and spiny forest covers from _c._ 1953 to _c._ 1973 in @Harper2007 and our study is most probably due to differences in forest definition and mapping methods between the 1953 aerial-photography derived map and the 1973 Landsat image derived map.
@@
 Annual deforestation rates are providing in ha/yr and in %/yr (second number in parenthesis).
-Exact same numbers as in scientific articles and reports from previous studies have been reported for deforestation rate in %/yr.
+For deforestation rate in %/yr, exact same numbers as in scientific articles and reports from previous studies have been reported.
 The way annual deforestation rates in %/yr have been computed in these previous studies can slightly differ from one study to another but estimates always correct for the potential presences of clouds on forest maps.
-On the contrary, deforestation rates in ha/yr have been recomputed from forest-cover estimates in Tab.
-2 and do not correct for the potential presence of clouds.
+Deforestation rates in ha/yr have been recomputed from forest-cover estimates in Tab.
+2 (except for @Harper2007 for the periods 1973-1990 and 1990-2000 for which annual deforestation rates in ha/yr were derived from numbers reported in the original publication, see methods) and do not correct for the potential presence of clouds.
 
@@
 Table 4: **Evolution of the forest fragmentation from 1950s to 2014 in Madagascar**.
-Forest fragmentation has progressively increased from _c._ 1953.
-The percentage of forest belonging to the "interior"" category (the most intact forest) has fallen from `r p.1`% in _c._ 1973 to `r p.2`% in 2014.
+Forest fragmentation has progressively increased since 1953.
+The percentage of forest belonging to the "interior"" category (the most intact forest) has fallen from `r p.1`% in 1973 to `r p.2`% in 2014.
 More than `r p.3`% of the forest belong to the "patch" and "transitional" categories (isolated small forest patches) in 2014.
-This increase in forest fragmentation has a great impact on both carbon emissions [@Putz2014] and biodiversity loss [@Gibson2013, @Murcia1995].
 
@@
 
-This study is part of the BioSceneMada project (https://bioscenemada.cirad.fr) and the ReCaREDD project (http://roadlessforests.org).
+This study is part of the BioSceneMada project (https://bioscenemada.cirad.fr) and the ReCaREDD project (http://forobs.jrc.ec.europa.eu/recaredd).
 The BioSceneMada project is funded by FRB (Fondation pour la Recherche sur la Biodiversité) and the FFEM (Fond Français pour l'Environnement Mondial) under the project agreement AAP-SCEN-2013 I.

### 2017-05-03 deforestmap.R

@@ -407,7 +407,7 @@ system("convert -delay 100 -loop 0 outputs/*.gif outputs/defor_Mada.gif")
 ## The GRASS Add-on r.forestfrag must be installed
 ## https://grass.osgeo.org/grass72/manuals/addons/r.forestfrag.html
 
-Year <- c(1953,1973,1990,2000,2010,2014)
+Year <- c(1953,1973,1990,2000,2005,2010,2014)
 system("g.region rast=harper -ap")
 system("r.mask --o raster=harper")
 for (i in 1:length(Year)) {
@@ -429,7 +429,7 @@ system("r.mask -r")
 ## Distance to forest edge
 ##==========================
 
-Year <- c(1953,1973,1990,2000,2010,2014)
+Year <- c(1953,1973,1990,2000,2005,2010,2014)
 system("g.region rast=harper -ap")
 for (i in 1:length(Year)) {
   # Message
@@ -454,7 +454,7 @@ for (i in 1:length(Year)) {
 ##========================
 
 #= Deforestation statistics
-Year <- c(1953,1973,1990,2000,2010,2014)
+Year <- c(1953,1973,1990,2000,2005,2010,2014)
 defor.df <- data.frame(Year=Year,area=NA,clouds=0,ann.defor=NA,theta=NA)
 # Areas
 for (i in 1:length(Year)) {
@@ -478,7 +478,7 @@ defor.df$clouds[defor.df$Year==1973] <- cloud.1973
 write.table(defor.df,"outputs/defor.txt",row.names=FALSE,sep="\t")
 
 #= Deforestation statistics for comparison
-Year <- c(1953,1973,1990,2000,2005,2010,2013)  # Here we add 2005 and 2013
+Year <- c(1953,1973,1990,2000,2005,2010,2013)  # Here we add 2013
 defor.df <- data.frame(Year=Year,area=NA,ann.defor=NA,theta=NA)
 # Areas
 for (i in 1:length(Year)) {
@@ -500,7 +500,7 @@ for (i in 2:length(Year)) {
 write.table(defor.df,"outputs/defor_for_comp.txt",row.names=FALSE,sep="\t")
 
 #= Fragmentation statistics
-Year <- c(1953,1973,1990,2000,2010,2014)
+Year <- c(1953,1973,1990,2000,2005,2010,2014)
[…]

### 2017-05-04 R/plotfcc.R
Ecoregion plot

@@ -11,6 +11,7 @@ library(ggplot2)
 library(rasterVis)
 library(gridExtra)
 library(grid)
+library(rgdal)
 
 # Zooms
 zoom.w <- list(xmin=346000,xmax=439000,ymin=7387000,ymax=7480000)
@@ -165,15 +166,15 @@ plot_zoom_dist <- function(rast_file,palette=dist.palette,vr=dist.vr) {
 library(ggspatial)  # for geom_spatial()
 
 # Import
-ecoregion <- readOGR(dsn="gisdata/vectors",layer="madagascar_ecoregion_redd_utm38s")
+ecoregion <- readOGR(dsn="gisdata/vectors",layer="madagascar_ecoregion_tenaizy_38s")
 
 # Text
 xt <- c(850000,580000,530000,600000)
 yt <- c(7920000,8060000,7250000,8460000)
 t.df <- data.frame(text=c("Moist","Dry","Spiny","Mangroves"),x=xt,y=yt)
 
-# Forest 2014  # If we want to add forest to ecoregions, to be done...
-# for2014 <- raster("outputs/for2014.tif")
+# Forest 2014
+for2014 <- raster("outputs/for2014.tif")
 
 # Segments
 seg.df <- data.frame(x=c(720000),y=c(8282000),
@@ -187,9 +188,10 @@ orange.t <- adjustcolor("orange",alpha.f=0.5)
 eco.col <- c("2"=green.t,"3"=orange.t,"0"=red.t,"1"="blue")
 
 # Plot
-plot.ecoregion <- ggplot() +
+plot.ecoregion <- gplot(for2014, maxpixels=res) +
   geom_spatial(ecoregion, aes(x=long,y=lat,group=group,fill=factor(id))) +
-  scale_fill_manual(values=eco.col) +
+  geom_raster(aes(fill=factor(value))) +
+  scale_fill_manual(values=eco.col, na.value="transparent") +
   geom_text(data=t.df, aes(x=x, y=y, label=text), size=4) +
   geom_segment(data=seg.df, aes(x=x, xend=xend, y=y, yend=yend), size=0.25) +
   theme_bw() + theme_base +

### 2017-05-04 deforestmap.Rmd
Plot ecoregion

@@
 Ecoregions were defined following climatic (see @Cornet1974) and vegetation (see @IEFN1996) criteria.
-The dark grey areas represent the remaining forest-cover for the year 2014.](outputs/ecoregion.png)
+The dark grey areas represent the remaining natural forest cover for the year 2014.](outputs/ecoregion.png)
 
@@
 
-This study is part of the BioSceneMada project (https://bioscenemada.cirad.fr) and the ReCaREDD project (http://forobs.jrc.ec.europa.eu/recaredd).
+This study is part of the Cirad's BioSceneMada project (https://bioscenemada.cirad.fr) and the Joint Research Center's ReCaREDD project (http://forobs.jrc.ec.europa.eu/recaredd).
 The BioSceneMada project is funded by FRB (Fondation pour la Recherche sur la Biodiversité) and the FFEM (Fond Français pour l'Environnement Mondial) under the project agreement AAP-SCEN-2013 I.

### 2017-05-04 R/plotfcc.R

@@ -167,6 +167,10 @@ library(ggspatial)  # for geom_spatial()
 
 # Import
 ecoregion <- readOGR(dsn="gisdata/vectors",layer="madagascar_ecoregion_tenaizy_38s")
+# Recode ecoregion
+ecoregion.data <- ecoregion@data
+ecoregion.data$code <- c("s","m","h","d") # spiny, mangroves, dry, humid
+ecoregion@data <- ecoregion.data
 
 # Text
 xt <- c(850000,580000,530000,600000)
@@ -185,21 +189,22 @@ red.t <- adjustcolor("red",alpha.f=0.5)
 blue.t <- adjustcolor("blue",alpha.f=0.5)
 green.t <- adjustcolor("dark green",alpha.f=0.5)
 orange.t <- adjustcolor("orange",alpha.f=0.5)
-eco.col <- c("2"=green.t,"3"=orange.t,"0"=red.t,"1"="blue")
+black.t <- adjustcolor("black",alpha.f=0.5)
+eco.col <- c("h"=green.t,"d"=orange.t,"s"=red.t,"m"="blue","1"=black.t)
 
 # Plot
-plot.ecoregion <- gplot(for2014, maxpixels=res) +
-  geom_spatial(ecoregion, aes(x=long,y=lat,group=group,fill=factor(id))) +
+plot.ecoregion <- gplot(for2014, maxpixels=10e5) +
+  geom_spatial(ecoregion, aes(x=long,y=lat,group=group,fill=factor(code))) +
   geom_raster(aes(fill=factor(value))) +
   scale_fill_manual(values=eco.col, na.value="transparent") +
-  geom_text(data=t.df, aes(x=x, y=y, label=text), size=4) +
+  geom_text(data=t.df, aes(x=x, y=y, label=text), size=3) +
   geom_segment(data=seg.df, aes(x=x, xend=xend, y=y, yend=yend), size=0.25) +
   theme_bw() + theme_base +
   theme(legend.position="null") +
   scale_y_continuous(limits=c(7165000,8685000),expand=c(0,0)) +
   scale_x_continuous(limits=c(300000,1100000),expand=c(0,0)) +
   coord_equal()
-ggsave("outputs/ecoregion.png", plot.ecoregion, width=5, height=9, units="cm")
+ggsave("outputs/ecoregion.png", plot.ecoregion, width=4, height=7, units="cm")
 
 # ================
 # Maps

### 2017-05-05 deforestmap.Rmd
Summary

@@
 ---
-title: "Combining global tree cover loss data with national historical forest-cover maps to look at six decades of deforestation and forest fragmentation in Madagascar"
+title: "Combining global tree cover loss data with historical national forest-cover maps to look at six decades of deforestation and forest fragmentation in Madagascar"
 bibliography: bib/biblio.bib
@@
 
-1. xxx
+1.
+The island of Madagascar has an unparralleled biodiversity, both in terms of number of species and endemism.
+This biodiversity is mainly present in the tropical forests of the island and is highly threatened by anthropogenic deforestation, associated principally to slash-and-burn agriculture and pasture.
+Scattered data from past studies on deforestation at the national level with substancial cloud cover on available forest maps prevent from analyzing the deforestation trend on the long term and estimate accurate recent annual deforestation rates for Madagascar.
 
-2. xxx
+2.
+In this study, we propose a new approach combining recent (2000-2014) global annual tree cover loss data with historical (1953-2000) national forest-cover maps to look at six decades (1953-2014) of deforestation and forest fragmentation in Madagascar.
+With this approach, we obtained cloud-free 30m-resolution forest-cover maps for Madagascar, for the year 1990 and annually from 2000 to 2014.
+
+3.
+We estimated that Madagascar has lost 44% of its natural forest on the period 1953-2014 (37% on the period 1973-2014).
+Natural forest covered 8.9 Mha in 2014 (15% of the national territory) and was divided into 4.4 Mha (50%) of moist forest, 2.6 Mha (29%) of dry forest, 1.7 Mha of spiny forest (19%) and 176778 ha (2%) of mangroves.
+Since 2005, deforestation has progressively increased in Madagascar and has more than doubled on the period 2010-2014 (98651 ha/yr, 1.08%/yr) compared to 2000-2005.
+Most of the forest is now located at a distance <1.2km of the forest edge while half of it was located far from forest edge (>1km and up to 4km) in 1973.
+The percentage of most intact forest as then decreased from 68% in 1973 to 50% in 2014.
+
+4.
+_Conclusion:_ Despite the conservation efforts in Madagascar, deforestation is dramatically increasing at the national level since 2005, with strong consequences on biodiversity loss and carbon emissions.
+This increase in the deforestation could be attributable to a rapid population growth (close to 3%/yr) and to a high political instability in the country.
+More effort, including conservation and development programs, should be done in Madagascar to save the remaining natural tropical forests, both to enhance local people livelihoods and answer to global environmental problems.
 
@@
 Following the work done by @Harper2007, we studied in details the change in deforestation rates and the evolution of forest fragmentation on six decades from _c._ 1953 to 2014.
-We showed that despite all the conservation efforts done in Madagascar since the mid 1980s, deforestation has increased in speed since 2010 and that fragmentation has dramatically increased leading to few remaining intact forest landscapes.
+We showed that despite all the conservation efforts done in Madagascar since the mid 1980s, deforestation has increased in speed since 2005 and that fragmentation has dramatically increased leading to few remaining intact forest landscapes.
 The approach described in this study could be easilly extended to all tropical countries with a forest map obtained at one date between 2000 and 2014, and could be repeated in the future to obtain future forest-cover change maps, as and when the @Hansen2013 product is updated.

### 2017-05-05 R/plotfcc.R

@@ -228,7 +228,6 @@ grob.for1953 <- ggplotGrob(for1953.plot)
 ## fcc
 fcc <- raster("outputs/fcc.tif")
 fcc.plot <- gplot(fcc,maxpixels=res) +
-  annotation_custom(grob=grob.ecoregion,xmin=810000,xmax=1110000,ymin=7800000,ymax=7800000) +
   annotation_custom(grob=grob.for1953,xmin=810000,xmax=1110000,ymin=7110000,ymax=7800000) +
   geom_raster(aes(fill=factor(value))) +
   geom_rect(aes(xmin=346000,xmax=439000,ymin=7387000,ymax=7480000),
@@ -269,6 +268,8 @@ zw.frag <- plot_zoom_frag("outputs/frag2014_zoom1.tif")
 ze.frag <- plot_zoom_frag("outputs/frag2014_zoom2.tif") + ylab("") + theme(legend.position="none")
 
 ## Zooms dist
+dedge2014 <- raster("outputs/dist_edge_2014_zoom1.tif")
+plot(dedge2014)
 zw.dist <- plot_zoom_dist("outputs/dist_edge_2014_zoom1.tif")
 ze.dist <- plot_zoom_dist("outputs/dist_edge_2014_zoom2.tif") + ylab("") + theme(legend.position="none")
 

### 2017-05-05 deforestmap.Rmd
Fcc plot

@@
 The island of Madagascar has an unparralleled biodiversity, both in terms of number of species and endemism.
-This biodiversity is mainly present in the tropical forests of the island and is highly threatened by anthropogenic deforestation, associated principally to slash-and-burn agriculture and pasture.
+This biodiversity is mainly located in the tropical forests of the island and is highly threatened by anthropogenic deforestation, associated principally to slash-and-burn agriculture and pasture.
 Scattered data from past studies on deforestation at the national level with substancial cloud cover on available forest maps prevent from analyzing the deforestation trend on the long term and estimate accurate recent annual deforestation rates for Madagascar.

### 2017-05-05 R/plotfcc.R

@@ -21,10 +21,11 @@ zooms <- list(zoom.w,zoom.e)
 # Compute zooms
 for (z in 1:length(zooms)) {
   ExtentZ <- paste(zooms[[z]]$xmin,zooms[[z]]$ymin,zooms[[z]]$xmax,zooms[[z]]$ymax)
-  Input <- c("fcc.tif","fcc_for1953.tif","frag2014.tif","dist_edge_2014.tif")
+  Input <- c("fcc.tif","fcc_for1953.tif","frag2014.tif","dist_edge_2014_mask.tif")
+  nodata <- c(rep(255,3),9999)
   Output <- paste0(sub(".tif","",Input),"_zoom",z,".tif")
   for (i in 1:length(Input)) {
-    system(paste0("gdalwarp -overwrite -srcnodata 255 -dstnodata 255 -te ",ExtentZ," -co 'COMPRESS=LZW' -co 'PREDICTOR=2' ",
+    system(paste0("gdalwarp -overwrite -srcnodata ",nodata[i]," -dstnodata ",nodata[i]," -te ",ExtentZ," -co 'COMPRESS=LZW' -co 'PREDICTOR=2' ",
                   paste0("outputs/",Input[i])," ",paste0("outputs/",Output[i])))
   }
 }
@@ -268,7 +269,7 @@ zw.frag <- plot_zoom_frag("outputs/frag2014_zoom1.tif")
 ze.frag <- plot_zoom_frag("outputs/frag2014_zoom2.tif") + ylab("") + theme(legend.position="none")
 
 ## Zooms dist
-dedge2014 <- raster("outputs/dist_edge_2014_zoom1.tif")
+dedge2014 <- raster("outputs/dist_edge_2014_mask.tif")
 plot(dedge2014)
 zw.dist <- plot_zoom_dist("outputs/dist_edge_2014_zoom1.tif")
 ze.dist <- plot_zoom_dist("outputs/dist_edge_2014_zoom2.tif") + ylab("") + theme(legend.position="none")

### 2017-05-05 deforestmap.R

@@ -441,13 +441,13 @@ for (i in 1:length(Year)) {
                 outputs/_dist_edge_",Year[i],".tif outputs/dist_edge_",Year[i],".tif"))
   file.remove(paste0("outputs/_dist_edge_",Year[i],".tif"))
   system(paste0("r.in.gdal --o input=outputs/dist_edge_",Year[i],".tif output=dist_edge_",Year[i]))
-  if (Year[i]==2014) {  # To mask outside Mada for plots
-    system("r.mask --o raster=harper")
-    system(paste0("r.out.gdal --o input=dist_edge_",Year[i]," createopt='compress=lzw,predictor=2' \\
-                  output=outputs/dist_edge_",Year[i],".tif"))
-    system("r.mask -r")
-  }
 }
+# Prepare dist_edge_2014 for figures
+system("r.mask --o raster=harper")
+system("r.mapcalc --o 'dist_edge_2014_mask = if(isnull(dist_edge_2014), 0, dist_edge_2014)'")
+system("r.out.gdal --o input=dist_edge_2014_mask nodata=9999 createopt='compress=lzw,predictor=2' \\
+       output=outputs/dist_edge_2014_mask.tif")
+system("r.mask -r")
 
 ##========================
 ## Statistics whole forest

### 2017-05-05 deforestmap.Rmd
Corrected map

@@
 Two zooms in the western dry (left part) and eastern moist (right part) ecoregions present more detailed views of (from top to bottom): forest-cover in 1950s, forest-cover change from _c._ 1973 to 2014, forest fragmentation in 2014 and distance to forest edge in 2014.
-Data on water bodies (blue) and water seasonality (light blue for seasonal water to dark blue for permanent water) has been extracted from @Pekel2016.](outputs/fig_fcc.png)
+Data on water bodies (blue) and water seasonality (light blue for seasonal water to dark blue for permanent water) has been extracted from @Pekel2016.](outputs/fig_fcc_highres.png)
 

### 2017-05-05 R/plotfcc.R

@@ -31,7 +31,7 @@ for (z in 1:length(zooms)) {
 }
 
 # Resolution
-res <- 10e3
+res <- 10e5
 
 # get_legend()
 get_legend <- function(myggplot) {
@@ -76,7 +76,7 @@ frag.palette <- paste0("c(",paste0("\"",frag.name,"\"","=","\"",
 frag.palette <- eval(parse(text=frag.palette))
 # distance to forest edge (continuous)
 dist.v <- c(50,100,500,1000,3000)
-dist.vr <- c(0,dist.v)/3000
+dist.vr <- c(0,dist.v)/3000 # rescale
 cf <- colorRampPalette(c("darkred","orange","darkgreen"))(5)
 dist.palette <- c(nofor73.col,cf)
 
@@ -149,11 +149,11 @@ plot_zoom_dist <- function(rast_file,palette=dist.palette,vr=dist.vr) {
   pzoom <- gplot(r,maxpixels=res) +
     geom_raster(aes(fill=value)) + ylab("Dist. to forest edge 2014") +
     scale_fill_gradientn(colours=palette,na.value="transparent",
-                        values=vr,
-                        breaks=c(1,500,1000,1500,2000),
-                        labels=c(">0","0.5","1","1.5","2"),
-                        guide=guide_colorbar(title="Distance (km)",title.position="top",
-                                            label.position="bottom",barwidth=10)) +
+                         values=vr,
+                         breaks=c(1,500,1000,1500,2000),
+                         labels=c(">0","0.5","1","1.5","2"),
+                         guide=guide_colorbar(title="Distance (km)",title.position="top",
+                                              label.position="bottom",barwidth=10)) +
     theme_bw() + theme_zoom + coord_equal() +
     theme(legend.position="bottom", 
           legend.margin=margin(-1.5,-1,-1,-1,"lines"))
@@ -269,10 +269,8 @@ zw.frag <- plot_zoom_frag("outputs/frag2014_zoom1.tif")
 ze.frag <- plot_zoom_frag("outputs/frag2014_zoom2.tif") + ylab("") + theme(legend.position="none")
 
 ## Zooms dist
-dedge2014 <- raster("outputs/dist_edge_2014_mask.tif")
-plot(dedge2014)
[…]

### 2017-05-11 deforestmap.Rmd
Update fig_fcc

@@
 The island of Madagascar has an unparralleled biodiversity, both in terms of number of species and endemism.
-This biodiversity is mainly located in the tropical forests of the island and is highly threatened by anthropogenic deforestation, associated principally to slash-and-burn agriculture and pasture.
+This biodiversity is mainly located in the tropical forests of the island and is highly threatened by anthropogenic deforestation, which is associated principally to slash-and-burn agriculture and pasture.
 Scattered data from past studies on deforestation at the national level with substancial cloud cover on available forest maps prevent from analyzing the deforestation trend on the long term and estimate accurate recent annual deforestation rates for Madagascar.
@@
 In this study, we propose a new approach combining recent (2000-2014) global annual tree cover loss data with historical (1953-2000) national forest-cover maps to look at six decades (1953-2014) of deforestation and forest fragmentation in Madagascar.
-With this approach, we obtained cloud-free 30m-resolution forest-cover maps for Madagascar, for the year 1990 and annually from 2000 to 2014.
+With this approach, we obtained cloud-free 30m-resolution forest-cover maps for Madagascar for the year 1990, and annually from 2000 to 2014.
 
@@
 
-There has been a long debate about the statement saying that Madagascar has lost 90% of its forest since the arrival of the humans on the island [@Virah-Sawmy2009, @Agarwal2005].
+There has been a long debate to know whether Madagascar had lost 90% of its forest since the arrival of the humans on the island [@Virah-Sawmy2009, @Agarwal2005].
 What it sure now, is that Madagascar has lost around 50% of its natural forest in six decades since 1953.

### 2017-05-11 R/plotfcc.R

@@ -104,7 +104,7 @@ theme_zoom <- theme(axis.line=element_blank(),
                     axis.title.y=element_text(size=18, margin=margin(0,-5,0,0)),
                     plot.title=element_text(size=20,face="bold",hjust=0.5,
                                             margin=margin(-0.5,0,0,0,"lines")),
-                    plot.margin=margin(0.5,0,0.5,0.25,"lines"),
+                    plot.margin=margin(0.5,0,1.0,0.25,"lines"),
                     panel.spacing=unit(c(0,0,0,0),"null"),
                     legend.position="none",
                     plot.background=element_rect(fill="transparent"),
@@ -281,6 +281,8 @@ lay <- rbind(c(NA,NA,NA,NA),
              c(NA,NA,NA,NA))
 png("outputs/fig_fcc.png",width=1000,height=1000,units="px",pointsize=12)
 grid.arrange(zw.for1953,zw.fcc,zw.frag,zw.dist,fcc.combi,
-             ze.for1953,ze.fcc,ze.frag,ze.dist,layout_matrix=lay,heights=c(0.05,1,1,1.01,1.04,0.05))
+             ze.for1953,ze.fcc,ze.frag,ze.dist,layout_matrix=lay,
+             widths=c(1,1,1,1),
+             heights=c(0.05,1.04,1.04,1.04,1.04,0.05))
 dev.off()
   

### 2017-05-11 deforestmap.Rmd
Improve discussion

@@
 
-### 4.1 Evolution of forest-cover in Madagascar over the last 60 years
+### 4.1 Interest of combining recent global annual tree cover loss data with historical national forest-cover maps
 
-There has been a long debate to know whether Madagascar had lost 90% of its forest since the arrival of the humans on the island [@Virah-Sawmy2009, @Agarwal2005].
-What it sure now, is that Madagascar has lost around 50% of its natural forest in six decades since 1953.
+In this study, we combined recent (2000-2014) global annual tree cover loss data [@Hansen2013] with historical (1953-2000) national forest-cover maps [@Harper2007] to look at six decades (1953-2014) of deforestation and forest fragmentation in Madagascar.
+With this approach, we obtained annual cloud-free 30m-resolution forest-cover maps for Madagascar for the period 2000 to 2014.
+This approach overcome the problem associated to the forest definition which is necessary to transform the 2000 global tree cover data from @Hansen2013 into a forest/non-forest map.
+Instead, an historical national or regional forest-cover map, based on a more appropriate forest definition at the national or regional scale, can be used as a mask to derive future forest-cover maps until 2014 taking into account global annual tree cover loss data by @Hansen2013.
+This approach could be easilly extended to other regions or countries with available forest-cover maps at one date on the period 2000-2014.
+Moreover, that approach can be repeated in the future to obtain future forest-cover change maps, as and when the @Hansen2013 product is updated.
+The R/GRASS code used for this study is made freely available in a GitHub repository (see Data availability statement).
+Code should be easilly transposable to other study sites or re-used for future analysis in Madagascar.
+Of course, the accuracy of the derived forest-cover change maps depends directly on the accuracy of the historical forest-cover maps and the accuracy of the tree cover loss data.
+Accuracy of the tree cover loss data [@Hansen2013] has been discussed in several recent studies [@Verhegghen2016].
 
@@
 
-The three previous studies have used (i) different forest definition (associated to independant visual interpretations of observation polygons to train the classifier), (ii) different classification algorithms (Maximum Likelihood Classification for @Harper2007 and @MEFT2009 vs.
-Random Forests for @ONE2015), and (iii) do not systematically correct for the presence of clouds.
-Despite these differences, the four studies gave similar forest-cover estimates and show the same deforestation trend over Madagascar, with a decrease of the deforestation rate on the period 1990-2005 followed by a progressive increase of the deforestation rate after 2005.
-Given the fact that our deforestation rate estimates have been computed from cloud-free maps, our estimates can be considered more accurate compared to estimates from previous studies, in particular for deforestation rates expressed in number of hectares per year.
-Current rates of deforestation for Madagascar (98651 ha/yr, 1.08 %/yr, if we consider the period 2010-2014) could be used to build reference scenarios for deforestation in Madagascar [@Olander2008].
-Forest-cover and deforestation rate estimates from this study could also be used to update the estimates in the upcoming FAO Forest Resources Assessment reports [@Keenan2015].
+### 4.2 Evolution of forest-cover in Madagascar since 1953 and current forest-cover
 
-### 4.2 Consequences of deforestation and forest fragmentation on biodiversity and carbon emissions
+We estimated that natural forests covered 8.9 Mha in Madagascar (15% of the national territory of 587041 km<sup>2</sup>) in 2014 and that Madagascar has lost 44% of its natural forest in six decades since 1953 (37% since 1973).
+There is an age-old debate about the original forest cover in Madagascar to know to which extent humans have altered the natural forest landscape since their large-scale settlement around 800 CE [@Cox2012].
+French naturalists Henri Humbert [@Humbert1927] and Henri Perrier de la Bâthie [@Perrier1921] stated that the full island was originally covered by forest, leading to the common statement that 90% of the natural forests have disappeared since the arrival of humans on the island [@Kull2000].
+More recent studies counter-balanced that point of view saying that grasslands existed in Madagascar long before human arrival and were determined by climate, natural grazing and other natural factors [@Virah-Sawmy2009, @Vorontsova2017].
+Other authors have questioned the entire narrative of extensive alteration of the landscape by early human activity which, through legislation, has severe consequences on local people [@Klein2002; @Kull2000].
+Whatever the original proportion of natural forests and grasslands in Madagascar, our results demonstrate that human activities since the 5Os have profoundly impacted the forest landscape and that recent conservation and development programs in Madagascar have failed to stop deforestation in the recent years.
+Associated to deforestation, we showed in our study that residual forest in Madagascar were highly fragmented with 95% of the forest being at a distance <1.2 km of the forest edge in 2014.
+Both deforestation and forest fragmentation has strong consequences on biodiversity and carbon emissions in Madagascar.
+Around 90% of Madagascar's endemic species are forest dependent [@Allnutt2008; @Goodman2005] and @Allnutt2008 estimated that deforestation between 1953 and 2000 had led to an extinction of 9% of the species.
+
+Increased wind speed, vapour pressure deficit and disturbance pressure in the vicinity of forest edges increase tree mortality, resulting in persistent carbon emissions long after deforestation occurs.
+Fire, wind and desiccation can penetrate up to kilometres into the forest15,19,20, altering microclimate within 100-300m of the forest edge13,21,22,23.
+Field studies have estimated relative carbon losses in forest edges (e) averaging 11% (ref.
+24) and up to 36% (ref.
+22) and 50% (ref.
+25) of aboveground biomass within the 100m edge zone, and modelling studies have shown biomass reductions of up to 70% in small fragments2
 
@@
 
-### 4.3 Interest of such an approach and limitations
+More efforts need to be done to save the few remaining natural tropical forests.
 
-- simple
-- easilly reproducible
-- no need to define a tree cover threshold to
+### 4.3 Deforestation trend and causes of the acceleration of the deforestation since 2005
+
+In our study, we have shown that the progressive decrease of the deforestation rate on the period 1990-2005 was followed by a continuous increase in the deforestation rate on the period 2005-2014.
+We showed that deforestation rate has more than doubled on the period 2010-2014 (98651 ha/yr, 1.08%/yr) compared to 2000-2005 (42296 ha.yr, 0.43%/yr).
+These results are confirmed by previous studies on deforestation at the national scale for Madagascar [@Harper2007; @MEFT2009; @ONE2015].
+Despite differences in the methodologies regarding (i) forest definition (associated to independant visual interpretations of observation polygons to train the classifier), (ii) classification algorithms (Maximum Likelihood Classification for @Harper2007 and @MEFT2009 vs.
+Random Forests for @ONE2015), (iii) deforestation rate computation, and (iii) correction for the presence of clouds, these studies gave similar forest-cover estimates and show the same deforestation trend over Madagascar.
+The increase of the deforestation rates after 2005 could be attributed to both population growth and political instability in the country.
+Nearly 90% of Madagascar's population relies on biomass for their daily energy needs [@Minten2013] and the link between population size and deforestation has previously been demonstrated in Madagascar [@Vieilledent2013, @Gorenflo2011].
+With a mean demographic growth rate of about 2.8%/yr and a population which has increased from 16 to 24 million people on the period 2000-2015 [UN2015], the increasing demand in wood-fuel and space for agriculture is likely to explain the increase in the deforestation rate.
+The political crisis of 2009, followed by several years of political instability and weak governance is also very likely to explain the increase in the deforestation rate observed on the period 2005-2014.
+Our deforestation rate estimates from 1990 to 2014 have been computed from cloud-free maps and can be considered more accurate in comparison with estimates from the previous studies, in particular for deforestation rates expressed in number of hectares per year.
+Forest-cover and deforestation rate estimates from this study could be used to update the estimates in the upcoming FAO Forest Resources Assessment reports [@Keenan2015].
+Current rates of deforestation for Madagascar (98651 ha/yr, 1.08 %/yr, if we consider the period 2010-2014) could be used to build reference scenarios for deforestation in Madagascar [@Olander2008].
 

### 2017-05-12 deforestmap.Rmd
Discussion part

@@
 Another global study is the work by @Hansen2013 who identified annual tree cover, forest loss and forest gain from 2000 to 2012.
-Meanwhile, results have been updated and are available up to the year 2014 at http://earthenginepartners.appspot.com/science-2013-global-forest.
+Meanwhile, results have been updated and are available up to the year 2014 at <http://earthenginepartners.appspot.com/science-2013-global-forest>.
 One problem with using the @Hansen2013 product to map forest cover is that a tree cover threshold must be defined to transform the tree cover data into a forest/non-forest map.
@@
 
-In this study, we combined recent (2000-2014) global annual tree cover loss data [@Hansen2013] with historical (1953-2000) national forest-cover maps [@Harper2007] to look at six decades (1953-2014) of deforestation and forest fragmentation in Madagascar.
-With this approach, we obtained annual cloud-free 30m-resolution forest-cover maps for Madagascar for the period 2000 to 2014.
-This approach overcome the problem associated to the forest definition which is necessary to transform the 2000 global tree cover data from @Hansen2013 into a forest/non-forest map.
-Instead, an historical national or regional forest-cover map, based on a more appropriate forest definition at the national or regional scale, can be used as a mask to derive future forest-cover maps until 2014 taking into account global annual tree cover loss data by @Hansen2013.
-This approach could be easilly extended to other regions or countries with available forest-cover maps at one date on the period 2000-2014.
-Moreover, that approach can be repeated in the future to obtain future forest-cover change maps, as and when the @Hansen2013 product is updated.
-The R/GRASS code used for this study is made freely available in a GitHub repository (see Data availability statement).
-Code should be easilly transposable to other study sites or re-used for future analysis in Madagascar.
+In this study, we combined recent (2001-2014) global annual tree cover loss data [@Hansen2013] with historical (1953-2000) national forest-cover maps [@Harper2007] to look at six decades (1953-2014) of deforestation and forest fragmentation in Madagascar.
+We obtained annual cloud-free 30m-resolution forest-cover maps for Madagascar for the period 2000 to 2014.
+This approach overcomes the problem associated to the forest definition which is necessary to transform the 2000 global tree cover data by @Hansen2013 into a forest/non-forest map [@Tropek2014].
+Instead, an historical national or regional forest-cover map, based on a more appropriate forest definition at the national or regional scale, can be used as a mask.
+This approach could be easilly extended to other regions or countries with available forest-cover maps at one date in the period 2000-2014.
+Moreover, that approach can be repeated in the future to obtain future forest-cover maps, as and when the @Hansen2013 product is updated.
+The R/GRASS code used for this study is made freely available in a GitHub repository (see Data availability statement) and should be easilly transposable to other study sites or re-used for future analysis in Madagascar.
 Of course, the accuracy of the derived forest-cover change maps depends directly on the accuracy of the historical forest-cover maps and the accuracy of the tree cover loss data.
-Accuracy of the tree cover loss data [@Hansen2013] has been discussed in several recent studies [@Verhegghen2016].
-
-Regarding the potential forest regrowth, we made the same choice as @Harper2007 and did not consider it in our study, although @Hansen2013 are also providing tree cover gain data over the period 2010-2014.
+Reported global accuracy for tree cover loss data on the period 2000-2012 is 99.6% (see Tab.
+S5 in @Hansen2013).
+@Verhegghen2016 also found similar annual deforestation rate estimates when comparing results derived from the tree cover loss data by @Hansen2013 on the period 2000-2012 with results derived from semi-supervized classification of Landsat satellite images on the period 2000-2010 by @Achard2014 for six countries in Central Africa.
+Consistently with the study by @Harper2007, we did not consider the potential natural forest regrowth in Madagascar, although @Hansen2013 are also providing tree cover gain data over the period 2001-2014.
 We made this choice for several reasons.
-First, there is little evidence of natural forest regeneration in Madagascar on the long term [@Harper2007, @Grouzis2001].
+First, gain data provided by @Hansen2013 can figure plantations and not natural forests [@Tropek2014].
+Second, there is little evidence of natural forest regeneration in Madagascar on the long term [@Harper2007, @Grouzis2001].
 This can be explained by several ecological mechanisms following burning practice such as soil erosion [@Grinand2017] and reduced seed bank due to fire and soil loss [@Grouzis2001].
 Moreover, in the areas where forest regeneration is ecologically possible, young regrowth forests are more easilly re-burnt for agriculture and pasture.
-Finally, young secondary forests do not provide the same ecosystem services (in terms of biodiversity and carbon storage) as old-growth natural forests.
+Third, young secondary forests do not provide the same ecosystem services (in terms of biodiversity and carbon storage) as old-growth natural forests.
 
@@
 We estimated that natural forests covered 8.9 Mha in Madagascar (15% of the national territory of 587041 km<sup>2</sup>) in 2014 and that Madagascar has lost 44% of its natural forest in six decades since 1953 (37% since 1973).
-There is an age-old debate about the original forest cover in Madagascar to know to which extent humans have altered the natural forest landscape since their large-scale settlement around 800 CE [@Cox2012].
+There is an age-old debate [@Burns2016] about the original forest cover in Madagascar to know to which extent humans have altered the natural forest landscape since their large-scale settlement around 800 CE [@Cox2012].
 French naturalists Henri Humbert [@Humbert1927] and Henri Perrier de la Bâthie [@Perrier1921] stated that the full island was originally covered by forest, leading to the common statement that 90% of the natural forests have disappeared since the arrival of humans on the island [@Kull2000].
@@
 Whatever the original proportion of natural forests and grasslands in Madagascar, our results demonstrate that human activities since the 5Os have profoundly impacted the forest landscape and that recent conservation and development programs in Madagascar have failed to stop deforestation in the recent years.
-Associated to deforestation, we showed in our study that residual forest in Madagascar were highly fragmented with 95% of the forest being at a distance <1.2 km of the forest edge in 2014.
+Associated to deforestation, we showed that residual forest in Madagascar was highly fragmented with 95% of the forest being at a distance <1.2 km of the forest edge in 2014.
 Both deforestation and forest fragmentation has strong consequences on biodiversity and carbon emissions in Madagascar.
@@
 Despite differences in the methodologies regarding (i) forest definition (associated to independant visual interpretations of observation polygons to train the classifier), (ii) classification algorithms (Maximum Likelihood Classification for @Harper2007 and @MEFT2009 vs.
-Random Forests for @ONE2015), (iii) deforestation rate computation, and (iii) correction for the presence of clouds, these studies gave similar forest-cover estimates and show the same deforestation trend over Madagascar.
+Random Forests for @ONE2015), (iii) deforestation rate computation method, and (iii) correction for the presence of clouds, these studies gave similar forest-cover estimates and show the same deforestation trend over Madagascar.
+Our deforestation rate estimates from 1990 to 2014 have been computed from cloud-free maps and can be considered more accurate in comparison with estimates from the previous studies, in particular for the deforestation rates expressed in number of hectares per year.
+Forest-cover and deforestation rate estimates from this study could be used to update the estimates in the upcoming FAO Forest Resources Assessment reports [@Keenan2015].
+Current rates of deforestation for Madagascar (98651 ha/yr, 1.08 %/yr for 2010-2014) could also be used to build reference scenarios for deforestation in Madagascar and help implement deforestation mitigation initatives such as the REDD+ [@Olander2008].
+
 The increase of the deforestation rates after 2005 could be attributed to both population growth and political instability in the country.
 Nearly 90% of Madagascar's population relies on biomass for their daily energy needs [@Minten2013] and the link between population size and deforestation has previously been demonstrated in Madagascar [@Vieilledent2013, @Gorenflo2011].
-With a mean demographic growth rate of about 2.8%/yr and a population which has increased from 16 to 24 million people on the period 2000-2015 [UN2015], the increasing demand in wood-fuel and space for agriculture is likely to explain the increase in the deforestation rate.
-The political crisis of 2009, followed by several years of political instability and weak governance is also very likely to explain the increase in the deforestation rate observed on the period 2005-2014.
-Our deforestation rate estimates from 1990 to 2014 have been computed from cloud-free maps and can be considered more accurate in comparison with estimates from the previous studies, in particular for deforestation rates expressed in number of hectares per year.
-Forest-cover and deforestation rate estimates from this study could be used to update the estimates in the upcoming FAO Forest Resources Assessment reports [@Keenan2015].
-Current rates of deforestation for Madagascar (98651 ha/yr, 1.08 %/yr, if we consider the period 2010-2014) could be used to build reference scenarios for deforestation in Madagascar [@Olander2008].
+With a mean demographic growth rate of about 2.8%/yr and a population which has increased from 16 to 24 million people on the period 2000-2015 [@UN2015], the increasing demand in wood-fuel and space for agriculture is likely to explain the increase in deforestation rates.
+The political crisis of 2009 [@Ploch2012], followed by several years of political instability and weak governance is also very likely to explain the increase in the deforestation rate observed on the period 2005-2014 [@Smith2003].
 

### 2017-05-12 deforestmap.Rmd
Finishing discussion

@@
 Other authors have questioned the entire narrative of extensive alteration of the landscape by early human activity which, through legislation, has severe consequences on local people [@Klein2002; @Kull2000].
-Whatever the original proportion of natural forests and grasslands in Madagascar, our results demonstrate that human activities since the 5Os have profoundly impacted the forest landscape and that recent conservation and development programs in Madagascar have failed to stop deforestation in the recent years.
+Whatever the original proportion of natural forests and grasslands in Madagascar, our results demonstrate that human activities since the 5Os have profoundly impacted the natural tropical forests and that recent conservation and development programs in Madagascar have failed to stop deforestation in the recent years.
+Deforestation has strong consequences on biodiversity and carbon emissions in Madagascar.
+Around 90% of Madagascar's species are forest dependent [@Allnutt2008; @Goodman2005] and @Allnutt2008 estimated that deforestation between 1953 and 2000 had led to an extinction of 9% of the species.
+The additional deforestation we observed on the period 2000-2014 (around 1Mha of deforested natural forest) worsen this result.
+Regarding carbon emissions, using the 2010 aboveground forest carbon map by @Vieilledent2016, we estimated that deforestation on the period 2010-2014 has led to the emissions of 40224 Gg (1 Gg = 10<sup>9</sup> g) of carbon in the atmosphere (10055 Gg/yr) and that the residual aboveground forest carbon stock was of 832822 Gg in 2014.
 Associated to deforestation, we showed that residual forest in Madagascar was highly fragmented with 95% of the forest being at a distance <1.2 km of the forest edge in 2014.
-Both deforestation and forest fragmentation has strong consequences on biodiversity and carbon emissions in Madagascar.
-Around 90% of Madagascar's endemic species are forest dependent [@Allnutt2008; @Goodman2005] and @Allnutt2008 estimated that deforestation between 1953 and 2000 had led to an extinction of 9% of the species.
-
-Increased wind speed, vapour pressure deficit and disturbance pressure in the vicinity of forest edges increase tree mortality, resulting in persistent carbon emissions long after deforestation occurs.
-Fire, wind and desiccation can penetrate up to kilometres into the forest15,19,20, altering microclimate within 100-300m of the forest edge13,21,22,23.
-Field studies have estimated relative carbon losses in forest edges (e) averaging 11% (ref.
-24) and up to 36% (ref.
-22) and 50% (ref.
-25) of aboveground biomass within the 100m edge zone, and modelling studies have shown biomass reductions of up to 70% in small fragments2
-
-Increase in forest fragmentation has a great impact on both carbon emissions [@Putz2014] and biodiversity loss [@Gibson2013, @Murcia1995].
-Biodiversity loss, see @Allnutt2008.
-
-More efforts need to be done to save the few remaining natural tropical forests.
+Forest fragmentation creates forest fragments too small to maintain viable populations and increases 'edge effects' at forest/non-forest interfaces having great impact on both carbon emissions [@Brinck2017] and biodiversity loss [@Gibson2013; @Murcia1995].
 
@@
 Random Forests for @ONE2015), (iii) deforestation rate computation method, and (iii) correction for the presence of clouds, these studies gave similar forest-cover estimates and show the same deforestation trend over Madagascar.
-Our deforestation rate estimates from 1990 to 2014 have been computed from cloud-free maps and can be considered more accurate in comparison with estimates from the previous studies, in particular for the deforestation rates expressed in number of hectares per year.
+Our deforestation rate estimates from 1990 to 2014 have been computed from cloud-free maps and can be considered more accurate in comparison with estimates from these previous studies, in particular for the deforestation rates expressed in number of hectares per year.
 Forest-cover and deforestation rate estimates from this study could be used to update the estimates in the upcoming FAO Forest Resources Assessment reports [@Keenan2015].
 Current rates of deforestation for Madagascar (98651 ha/yr, 1.08 %/yr for 2010-2014) could also be used to build reference scenarios for deforestation in Madagascar and help implement deforestation mitigation initatives such as the REDD+ [@Olander2008].
-
-The increase of the deforestation rates after 2005 could be attributed to both population growth and political instability in the country.
-Nearly 90% of Madagascar's population relies on biomass for their daily energy needs [@Minten2013] and the link between population size and deforestation has previously been demonstrated in Madagascar [@Vieilledent2013, @Gorenflo2011].
+The increase of the deforestation rates we observed after 2005 could be attributed to both population growth and political instability in the country.
+Nearly 90% of Madagascar's population relies on biomass for their daily energy needs [@Minten2013] and the link between population size and deforestation has previously been demonstrated in Madagascar [@Vieilledent2013; @Gorenflo2011].
 With a mean demographic growth rate of about 2.8%/yr and a population which has increased from 16 to 24 million people on the period 2000-2015 [@UN2015], the increasing demand in wood-fuel and space for agriculture is likely to explain the increase in deforestation rates.
-The political crisis of 2009 [@Ploch2012], followed by several years of political instability and weak governance is also very likely to explain the increase in the deforestation rate observed on the period 2005-2014 [@Smith2003].
+The political crisis of 2009 [@Ploch2012], followed by several years of political instability and weak governance could also explain the increase in the deforestation rate observed on the period 2005-2014 [@Smith2003].
+These results show that despite the conservation policy in Madagascar [@Freudenberger2010], deforestation has dramatically increased at the national level since 2005.
+More efforts should be done to identify the real causes and mechanisms explaining this rampant deforestation in order to implement new conservation strategies that will be effective at conserving the few remaining natural tropical forests in Madagascar.
 

### 2017-05-12 deforestmap.R

@@ -797,6 +797,23 @@ axis(2,at=seq(0,5000,by=1000),labels=seq(0,5,by=1))
 points(dist.df$Year, dist.df$mean, type="b", pch=19)
 dev.off()
 
+##==================================================
+## Carbon emissions between 2010 and 2014
+##==================================================
+
+system("r.in.gdal --o input=gisdata/rasters/carbon/ACD_RF_2010.tif output=acd")
+system("g.region rast=for2010 -ap")
+system("r.mask --o raster=for2010")
+system("r.univar map=acd")
+system("g.region rast=for2014 -ap")
+system("r.mask --o raster=for2014")
+system("r.univar map=acd")
+system("r.mask -r")
+acd_2010 <- 9700502114*30*30/10000 # 873045190
+acd_2014 <- 9253573526*30*30/10000 # 832821617
+carbon_emissions <- acd_2010-acd_2014
+
+
 ##========================
 ## Save objects
 ##========================

### 2017-05-12 deforestmap.Rmd
Update bib

@@
 
-Tropical forests provide important ecosystem services at the global scale, such as biodiversity conservation and climate regulation [@Costanza1997] but also many at the local scale for indigeneous populations.
-At the local scale, tropical forests supply clean water, can preserve populations from flood [@Bradshaw2007] and provide wood for construction and fuel and other non-woody products for food and health.
-On the island of Madagascar, three types of tropical forests, covering about 9 millions of hectares (15% of the country area) can be found: the moist forest in the East, the dry forest in the West and the spiny forest in the South [@Vieilledent2016].
-Among all the ecosystem services cited previously, Malagasy forests are particularly important for the unique biodiversity they shelter, both in term of number of species and endemism in many taxonomic groups [@Goodman2005].
-Deforestation threatens species survival and diminishes biodiversity by destroying forest habitat, creating forest fragments too small to maintain viable populations and increasing 'edge effects' at forest/non-forest interfaces (Harris 1984).
-Edge effects due to fragmentation typically affect an area several times larger than the forest destruction itself (Harris 1984; Skole & Tucker 1993), affecting micro-meteorology over short distances (Kapos 1989), and increasing exposure to damaging winds (Ferreira & Laurance 1997), fire frequency (Cochrane 2001), and access for livestock, other non-forest animals and hunters (Enserink 1999; Cullen et al.
+Separated from the African continent and the Indian plate about 165 and 85 million years ago respectively [@Ali2008], the flora and fauna of Madagascar followed its own evolutionary path.
+Combined with a high number of micro-habitats [@Pearson2009], this has led to Madagascar's exceptional biodiversity both in term of number of species and endemism in many taxonomic groups [@Crottini2012; @Goodman2005].
+Most of the biodiversity in Madagascar is concentrated in the tropical forests of the island which can be divided in three types: the moist forest in the East, the dry forest in the West and the spiny forest in the South [@Vieilledent2016].
+This unparalleled biodiversity is severely threatened by deforestation [@Harper2007; @Vieilledent2013] associated to human activities such as slash-and-burn agriculture and pasture [@Scales2011].
+Tropical forest in Madagascar also store a large amount of carbon [@Vieilledent2016] and high rates of deforestation in Madagascar are responsible for large emissions of CO2 in the atmosphere [@Achard2014].
+Deforestation threaten species survival by reducing their habitat [@Brooks2002].
+Associated to deforestation, forest and habitat fragmentation can also lead to species extinction by isolating populations from each others and creating forest patches too small to maintain viable populations [@Saunders1991].
+Edge effects due to fragmentation [@Murcia1995] typically affect an area several times larger than the forest destruction itself (Harris 1984; Skole & Tucker 1993).
+Edge effects affecting micro-meteorology over short distances (Kapos 1989), and increasing exposure to damaging winds (Ferreira & Laurance 1997), fire frequency (Cochrane 2001), and access for livestock, other non-forest animals and hunters (Enserink 1999; Cullen et al.
 2000).
 To assess the impact of tropical deforestation on biological diversity, not only the area deforested, but also the isolation of forest patches and the area of edge habitat must be determined.
+Detailed information on the extent and distribution of tropical forests and their rate of clearing are essential to estimating threats to biological diversity and carbon emissions.
 
-Accurate maps of forest-cover and forest-cover change are necessary to monitore carbon emissions and loss of biodiversity associated to deforestation, assess the efficiency of present conservation strategies [@Eklund2016], and implement new strategies for the future [@Vieilledent2013a, @Vieilledent2016].
-Unfortunately, accurate and cloud-free forest-cover maps are not available for Madagascar for the last decade (2005-2015).
+Accurate maps of forest-cover and forest-cover change are necessary to monitore loss of biodiversity and carbon emissions associated to deforestation, assess the efficiency of present conservation strategies [@Eklund2016], and implement new strategies for the future [@Vieilledent2013a, @Vieilledent2016].
+Unfortunately, accurate and cloud-free forest-cover maps are not available for Madagascar for the last fifteen years (2000-2015).
 In their paper untitled _"Fifty years of deforestation and forest fragmentation in Madagascar"_, @Harper2007 published maps of forest cover change for Madagascar from 1950s to 2000 (for the years _c._ 1953, _c._ 1973, 1990 and 2000).
 The _c._ 1953 forest map was derived from the visual interpretation of aerial photographies and gives a coarse representation (scale 1/1,000,000) of the forest extent in Madagascar for the 1950s.
-Forest maps for the years _c._ 1973, 1990 and 2000, were obtained from a supervised classification of Landsat satellite images at 60 (for the year 1973) and 30m (for years 1990 and 2000) and can be used to derive much more accurate estimates of the forest cover at these three dates.
+Forest maps for the years _c._ 1973, 1990 and 2000, were obtained from a supervised classification of Landsat satellite images at 60m (for the year 1973) and 30m (for years 1990 and 2000) and can be used to derive much more accurate estimates of the forest cover at these three dates.
 Maps provided by @Harper2007 were not free of clouds (11 244 km2 of unknown cover type for the year 2000 for example).

### 2017-05-13 deforestmap.Rmd
New bib file and intro

@@
 bibliography: bib/biblio.bib
-csl: bib/conservation-letters.csl
+csl: bib/biological-conservation.csl
 output: 
@@
 This unparalleled biodiversity is severely threatened by deforestation [@Harper2007; @Vieilledent2013] associated to human activities such as slash-and-burn agriculture and pasture [@Scales2011].
-Tropical forest in Madagascar also store a large amount of carbon [@Vieilledent2016] and high rates of deforestation in Madagascar are responsible for large emissions of CO2 in the atmosphere [@Achard2014].
-Deforestation threaten species survival by reducing their habitat [@Brooks2002].
+Tropical forests in Madagascar also store a large amount of carbon [@Vieilledent2016] and high rates of deforestation in Madagascar are responsible for large CO2 emissions in the atmosphere [@Achard2014].
+Deforestation threatens species survival by reducing their habitat [@Brooks2002; @Tidd2001].
 Associated to deforestation, forest and habitat fragmentation can also lead to species extinction by isolating populations from each others and creating forest patches too small to maintain viable populations [@Saunders1991].
-Edge effects due to fragmentation [@Murcia1995] typically affect an area several times larger than the forest destruction itself (Harris 1984; Skole & Tucker 1993).
-Edge effects affecting micro-meteorology over short distances (Kapos 1989), and increasing exposure to damaging winds (Ferreira & Laurance 1997), fire frequency (Cochrane 2001), and access for livestock, other non-forest animals and hunters (Enserink 1999; Cullen et al.
-2000).
+Edge effects due to fragmentation [@Murcia1995].
+
+Dire en quoi la déforestation (perte d'habitat) et la fragmentation (isolation des populations + microclimate on forest edge) affecte la bodiversité.
+Ainsi, afin de connaitre l'impact de la deforestation sur la biodiversité, il est nécessaire de disposer de cartes forestières spatialisées.
+Les chiffres de la déforestation et des surfaces forestières seuls, tels qu'ils peuvent être produit par la FAO [@Keenan2015], ne suffisent pas.
+De même, concernant les émissions de CO2 associées à la déforestation.
+Les stocks de carbone forestiers varient spatialement [@Vieilledent2016].
+Il est donc nécessaires de disposer de cartes de déforestation spatialement explicites.
+
 To assess the impact of tropical deforestation on biological diversity, not only the area deforested, but also the isolation of forest patches and the area of edge habitat must be determined.
@@
 
-Accurate maps of forest-cover and forest-cover change are necessary to monitore loss of biodiversity and carbon emissions associated to deforestation, assess the efficiency of present conservation strategies [@Eklund2016], and implement new strategies for the future [@Vieilledent2013a, @Vieilledent2016].
+Accurate maps of forest-cover and forest-cover change are necessary to monitore loss of biodiversity and carbon emissions associated to deforestation and forest fragmentation, assess the efficiency of present conservation strategies [@Eklund2016], and implement new strategies for the future [@Vieilledent2013a, @Vieilledent2016].
+
 Unfortunately, accurate and cloud-free forest-cover maps are not available for Madagascar for the last fifteen years (2000-2015).

### 2017-05-13 deforestmap.R

@@ -840,7 +840,7 @@ dir.create("report")
 render("deforestmap.Rmd",output_dir="report") # html output
 ## Cover letter
 render("coverletter.md",output_format=c("pdf_document"),output_dir="report") # pdf output
-#render("deforestmap.Rmd",output_format=c("html_document","pdf_document","word_document"),output_dir="report")
+#render("deforestmap.Rmd",output_format=c("pdf_document","word_document"),output_dir="report")
 
 ##===========================================================================
 ## End of script

### 2017-05-15 deforestmap.Rmd
Version coauthor

@@
     keep_tex: true
+  word_document:
+    fig_caption: yes
 ---
@@
 
-Separated from the African continent and the Indian plate about 165 and 85 million years ago respectively [@Ali2008], the flora and fauna of Madagascar followed its own evolutionary path.
+Separated from the African continent and the Indian plate about 165 and 88 million years ago respectively [@Ali2008], the flora and fauna of Madagascar followed its own evolutionary path.
 Combined with a high number of micro-habitats [@Pearson2009], this has led to Madagascar's exceptional biodiversity both in term of number of species and endemism in many taxonomic groups [@Crottini2012; @Goodman2005].
 Most of the biodiversity in Madagascar is concentrated in the tropical forests of the island which can be divided in three types: the moist forest in the East, the dry forest in the West and the spiny forest in the South [@Vieilledent2016].
-This unparalleled biodiversity is severely threatened by deforestation [@Harper2007; @Vieilledent2013] associated to human activities such as slash-and-burn agriculture and pasture [@Scales2011].
+This unparalleled biodiversity is severely threatened by anthropogenic deforestation [@Harper2007; @Vieilledent2013] associated to activities such as slash-and-burn agriculture and pasture [@Scales2011].
 Tropical forests in Madagascar also store a large amount of carbon [@Vieilledent2016] and high rates of deforestation in Madagascar are responsible for large CO2 emissions in the atmosphere [@Achard2014].
 Deforestation threatens species survival by reducing their habitat [@Brooks2002; @Tidd2001].
-Associated to deforestation, forest and habitat fragmentation can also lead to species extinction by isolating populations from each others and creating forest patches too small to maintain viable populations [@Saunders1991].
-Edge effects due to fragmentation [@Murcia1995].
-
-Dire en quoi la déforestation (perte d'habitat) et la fragmentation (isolation des populations + microclimate on forest edge) affecte la bodiversité.
-Ainsi, afin de connaitre l'impact de la deforestation sur la biodiversité, il est nécessaire de disposer de cartes forestières spatialisées.
-Les chiffres de la déforestation et des surfaces forestières seuls, tels qu'ils peuvent être produit par la FAO [@Keenan2015], ne suffisent pas.
-De même, concernant les émissions de CO2 associées à la déforestation.
-Les stocks de carbone forestiers varient spatialement [@Vieilledent2016].
-Il est donc nécessaires de disposer de cartes de déforestation spatialement explicites.
-
-To assess the impact of tropical deforestation on biological diversity, not only the area deforested, but also the isolation of forest patches and the area of edge habitat must be determined.
-Detailed information on the extent and distribution of tropical forests and their rate of clearing are essential to estimating threats to biological diversity and carbon emissions.
-
-Accurate maps of forest-cover and forest-cover change are necessary to monitore loss of biodiversity and carbon emissions associated to deforestation and forest fragmentation, assess the efficiency of present conservation strategies [@Eklund2016], and implement new strategies for the future [@Vieilledent2013a, @Vieilledent2016].
+Associated to deforestation, forest fragmentation can also lead to species extinction by isolating populations from each others and creating forest patches too small to maintain viable populations [@Saunders1991].
+Fragmentation also extents forest edge where ecological conditions (such as air temperature, light intensity and air moisture) can be dramatically modified, with consequences on the abundance and distribution of species [@Murcia1995].
+Forest fragmentation can also have substancial effects on forest carbon storage capacity, carbon stocks being much lower at the forest edge than under a closed canopy [@Brinck2017].
+Moreover, forest carbon stocks are varying spatially due to climate or soil factors [@Vieilledent2016; @Saatchi2011].
+As a consequence, accurate and spatially explicit maps of forest-cover and forest-cover change are necessary to monitore loss of biodiversity and carbon emissions associated to deforestation and forest fragmentation, assess the efficiency of present conservation strategies [@Eklund2016], and implement new strategies for the future [@Vieilledent2013a, @Vieilledent2016].
+Simple time-series of forest-cover estimates, such as those provided by the FAO Forest Resource Assessment report [@Keenan2015] are not sufficient.
 
@@
 
-![Figure 2: **Evolution of the mean distance to forest edge from 1953 to 2014 in Madagascar.** Black dots represent the mean distance to forest edge for each year.
+![Figure 3: **Evolution of the mean distance to forest edge from 1953 to 2014 in Madagascar.** Black dots represent the mean distance to forest edge for each year.
 Vertical dashed segments represent the 95% quantiles of the distance to forest edge.

### 2017-05-15 R/comp_old_deprecated.R

@@ -1,193 +1,0 @@
-# Import perr-fh map
-system("r.in.gdal --o input=gisdata/rasters/perr-fh/FCC_051013_MG_VF_38s.tif output=perrfh")
-
-# Classification
-system("r.describe perrfh")
-# * 7 111 113 119 133 183 222 223 228 233 288 333 444 533 553 555 666 733 773 777 778 788 888 1000
-# 1=moist_forest, 2=dry_forest, 3=no_forest, 4=water, 5=spiny_forest, 6=building/soil/rock,
-# 7=mangroves, 8=cloud/shadow, 9=cloud, 1000=artefact
-
-# 111	Forêt humide 2013
-# 113	Déforestation 2010-2013
-# 119	Nuage 2013
-# 133	Déforestation 2005-2013
-# 183	Ombre 2010
-# 222	Forêt sèche 2013
-# 223	Déforestation forêt séche 2010-2013
-# 228	Ombre 2013
-# 233	Déforestation forêt séche 2005-2010
-# 288	Ombre 2010
-# 333	Mosaic de culture, jachère et savane
-# 444	Zone en eau
-# 533	Déforestation forêt épineuse 2005-2010
-# 553	Déforestation forêt épineuse 2010-2013
-# 555	Forêt épineuse
-# 666	Etablissement, sol et roche nue
-# 733	Déforestation forêt de mangrove 2005-2010
-# 773	Déforestation forêt de Mangrove 2010-2013
-# 777	Forêt Mangrove 2013
-# 778	Nuage 2010
-# 788	Nuage 2010
-# 888	Nuage
-
-# Report
-system("r.report perrfh units=h")
-
-# +-----------------------------------------------------------------------------+
-# |                         RASTER MAP CATEGORY REPORT                          |
-# |LOCATION: deforestmap                                Fri Feb  3 11:25:54 2017|
-# |-----------------------------------------------------------------------------|
-# |          north: 8682420    east: 1100820                                    |
-# |REGION    south: 7155900    west:  298440                                    |
-# |          res:        30    res:       30                                    |
-# |-----------------------------------------------------------------------------|
[…]

### 2017-05-15 R/plotfcc.R

@@ -198,14 +198,14 @@ plot.ecoregion <- gplot(for2014, maxpixels=10e5) +
   geom_spatial(ecoregion, aes(x=long,y=lat,group=group,fill=factor(code))) +
   geom_raster(aes(fill=factor(value))) +
   scale_fill_manual(values=eco.col, na.value="transparent") +
-  geom_text(data=t.df, aes(x=x, y=y, label=text), size=3) +
+  geom_text(data=t.df, aes(x=x, y=y, label=text), size=5) +
   geom_segment(data=seg.df, aes(x=x, xend=xend, y=y, yend=yend), size=0.25) +
   theme_bw() + theme_base +
   theme(legend.position="null") +
   scale_y_continuous(limits=c(7165000,8685000),expand=c(0,0)) +
   scale_x_continuous(limits=c(300000,1100000),expand=c(0,0)) +
   coord_equal()
-ggsave("outputs/ecoregion.png", plot.ecoregion, width=4, height=7, units="cm")
+ggsave("outputs/ecoregion.png", plot.ecoregion, width=10, height=15, units="cm")
 
 # ================
 # Maps

### 2017-05-15 deforestmap.R

@@ -839,8 +839,8 @@ dir.create("report")
 ## Main document
 render("deforestmap.Rmd",output_dir="report") # html output
 ## Cover letter
-render("coverletter.md",output_format=c("pdf_document"),output_dir="report") # pdf output
-#render("deforestmap.Rmd",output_format=c("pdf_document","word_document"),output_dir="report")
+#render("coverletter.md",output_format=c("pdf_document"),output_dir="report") # pdf output
+render("deforestmap.Rmd",output_format=c("pdf_document","word_document"),output_dir="report")
 
 ##===========================================================================
 ## End of script

### 2017-05-29 deforestmap.Rmd
Comments by Fred

@@
 4. Wildlife Conservation Society, 101 Antananarivo, MADAGASCAR
-5. Office National de l'Environnement, 101 Antananarivo, MADAGASCAR
+5. Office National pour l'Environnement, 101 Antananarivo, MADAGASCAR
 
@@
 1.
-The island of Madagascar has an unparralleled biodiversity, both in terms of number of species and endemism.
+_Context:_ The island of Madagascar has an unparalleled biodiversity, both in terms of number of species and endemism.
 This biodiversity is mainly located in the tropical forests of the island and is highly threatened by anthropogenic deforestation, which is associated principally to slash-and-burn agriculture and pasture.
-Scattered data from past studies on deforestation at the national level with substancial cloud cover on available forest maps prevent from analyzing the deforestation trend on the long term and estimate accurate recent annual deforestation rates for Madagascar.
+Scattered forest maps from past studies at national level with substantial gaps (due to presence of cloud cover on satellite imagery) prevent from analyzing the deforestation trends on a long term perspective in Madagascar.
 
 2.
-In this study, we propose a new approach combining recent (2000-2014) global annual tree cover loss data with historical (1953-2000) national forest-cover maps to look at six decades (1953-2014) of deforestation and forest fragmentation in Madagascar.
-With this approach, we obtained cloud-free 30m-resolution forest-cover maps for Madagascar for the year 1990, and annually from 2000 to 2014.
+_Approach:_ In this study, we propose a new approach combining historical (1953-2000) national forest-cover maps with recent (2000-2014) global annual tree cover loss data to look at six decades (1953-2014) of deforestation and forest fragmentation in Madagascar.
+We also produce new forest-cover maps at 30m-resolution over the full territory of Madagascar for the year 1990, and annually from 2000 to 2014.
 
 3.
-We estimated that Madagascar has lost 44% of its natural forest on the period 1953-2014 (37% on the period 1973-2014).
-Natural forest covered 8.9 Mha in 2014 (15% of the national territory) and was divided into 4.4 Mha (50%) of moist forest, 2.6 Mha (29%) of dry forest, 1.7 Mha of spiny forest (19%) and 176778 ha (2%) of mangroves.
-Since 2005, deforestation has progressively increased in Madagascar and has more than doubled on the period 2010-2014 (98651 ha/yr, 1.08%/yr) compared to 2000-2005.
-Most of the forest is now located at a distance <1.2km of the forest edge while half of it was located far from forest edge (>1km and up to 4km) in 1973.
-The percentage of most intact forest as then decreased from 68% in 1973 to 50% in 2014.
+_Results:_ We estimate that Madagascar has lost 44% of its natural forest cover over the period 1953-2014 (including 37% over the period 1973-2014).
+Natural forests cover 8.9 Mha in 2014 (15% of the national territory) which are divided into 4.4 Mha (50%) of moist forests, 2.6 Mha (29%) of dry forests, 1.7 Mha of spiny forests (19%) and 177,000 ha (2%) of mangroves.
+Since 2005, the annual deforestation rate has progressively increased in Madagascar and has more than doubled from the period 2000-2005 to reach 99,000 ha/yr  during 2010-2014 (corresponding to a rate of 1.08%/yr).
+Most of the forests are now located at less than 1.2km from forest edges while half of them were located far from forest edges in 1973  (i.e.
+between 1km and 4km from forest edges).
+The percentage of core forest has then decreased from 68% in 1973 to 50% in 2014.
 
 4.
-_Conclusion:_ Despite the conservation efforts in Madagascar, deforestation is dramatically increasing at the national level since 2005, with strong consequences on biodiversity loss and carbon emissions.
-This increase in the deforestation could be attributable to a rapid population growth (close to 3%/yr) and to a high political instability in the country.
-More effort, including conservation and development programs, should be done in Madagascar to save the remaining natural tropical forests, both to enhance local people livelihoods and answer to global environmental problems.
+_Conclusion:_ Despite the conservation efforts in Madagascar, deforestation has continued to significantly increase at the national level since 2005, with strong consequences on biodiversity loss and carbon emissions.
+This increase in the deforestation rate can be related to a rapid population growth (close to 3%/yr) and to poor law enforcement due to political instability in the country.
+More effort, including conservation and development programs, should be done in Madagascar to protect the remaining natural tropical forests, both to enhance local people livelihoods and answer to global environmental problems.
 
@@
 Fragmentation also extents forest edge where ecological conditions (such as air temperature, light intensity and air moisture) can be dramatically modified, with consequences on the abundance and distribution of species [@Murcia1995].
-Forest fragmentation can also have substancial effects on forest carbon storage capacity, carbon stocks being much lower at the forest edge than under a closed canopy [@Brinck2017].
-Moreover, forest carbon stocks are varying spatially due to climate or soil factors [@Vieilledent2016; @Saatchi2011].
-As a consequence, accurate and spatially explicit maps of forest-cover and forest-cover change are necessary to monitore loss of biodiversity and carbon emissions associated to deforestation and forest fragmentation, assess the efficiency of present conservation strategies [@Eklund2016], and implement new strategies for the future [@Vieilledent2013a, @Vieilledent2016].
+Forest fragmentation can also have substantial effects on forest carbon storage capacity, carbon stocks being much lower at the forest edge than under a closed canopy [@Brinck2017].
+Moreover, forest carbon stocks are varying spatially due to climate or soil factors [@Saatchi2011; @Vieilledent2016].
+As a consequence, accurate and spatially explicit maps of forest-cover and forest-cover change are necessary to monitor loss of biodiversity and carbon emissions associated to deforestation and forest fragmentation, assess the efficiency of present conservation strategies [@Eklund2016], and implement new strategies for the future [@Vieilledent2013; @Vieilledent2016].
 Simple time-series of forest-cover estimates, such as those provided by the FAO Forest Resource Assessment report [@Keenan2015] are not sufficient.
 
-Unfortunately, accurate and cloud-free forest-cover maps are not available for Madagascar for the last fifteen years (2000-2015).
-In their paper untitled _"Fifty years of deforestation and forest fragmentation in Madagascar"_, @Harper2007 published maps of forest cover change for Madagascar from 1950s to 2000 (for the years _c._ 1953, _c._ 1973, 1990 and 2000).
-The _c._ 1953 forest map was derived from the visual interpretation of aerial photographies and gives a coarse representation (scale 1/1,000,000) of the forest extent in Madagascar for the 1950s.
-Forest maps for the years _c._ 1973, 1990 and 2000, were obtained from a supervised classification of Landsat satellite images at 60m (for the year 1973) and 30m (for years 1990 and 2000) and can be used to derive much more accurate estimates of the forest cover at these three dates.
-Maps provided by @Harper2007 were not free of clouds (11 244 km2 of unknown cover type for the year 2000 for example).
-Using the same supervised classification approach as in @Harper2007, more recent maps have been released for the periods 2000-2005-2010 by national institutions, with the technical support of international environmental NGOs [@MEFT2009, @ONE2013].
-Another recent map using more advanced statistical tool for classification (the Random Forest classifier, see @Grinand2013 and @Rakotomalala2015) has been released for the periods 2005-2010-2013 [@ONE2015].
-However, these maps (i) are too old to give recent estimates of deforestation [@MEFT2009, @ONE2013], (ii) figure large areas of clouds (images with more than 20% of clouds in 2010 for the map provided by [@ONE2013]), or (iii) show large errors of classification in specific areas, especially in the dry and spiny forests, for which the spectral answer varies strongly due to deciduousness associated with the higher seasonality than in moist forests (Kappa<0.6 and overall accuracy<0.8 for the dry and spiny forests for the maps produced by @ONE2015).
-Moreover, deriving such forest maps from supervised classification requires a huge amount of time and work, especially regarding image selection to reduce cloud cover and visual interpretation of the numerous observation polygons used to train the classification algorithm [@Rakotomalala2015].
-In addition, visual interpretation might be subjective and is potentially a source of errors.
-All this work of image selection and visual interpretation have to be repeated to produce new forest maps in the future.
+Unfortunately, accurate and exhaustive forest-cover maps are not available for Madagascar for the last fifteen years (2000-2015).
+@Harper2007 produced maps of forest cover and forest cover changes over Madagascar for the years c.
+1953, c.
+1973, 1990 and 2000.
+The c.
+1953 forest map was derived from the visual interpretation of aerial photography at coarse scale (1/1,000,000).
+Forest maps for the years c.
+1973, 1990 and 2000, were obtained from supervised classification of Landsat satellite images at 60m resolution (for the year 1973) or 30m resolution (for years 1990 and 2000) and can be used to derive accurate estimates (89.5% accuracy reported for the forest / non-forest map of year 2000).
+Maps provided by @Harper2007 were not exhaustive (due to the presence of clouds in the satellite imagery), e.g.
+11 244 km2 are mapped as unknown cover type for the year 2000.
+Using a similar supervised classification approach as in @Harper2007, more recent maps were produced for the periods 2000-2005-2010 by national institutions, with the technical support of international environmental NGOs [@MEFT2009; @ONE2013].
+Another set of recent forest cover maps using an advanced statistical tool for classification, the Random Forest classifier [@Grinand2013; @Rakotomalala2015], was produced for the periods 2005-2010-2013 [@ONE2015].
+However, these maps are either too old to give recent estimates of deforestation [@MEFT2009; @ONE2013], include large areas of missing information (due to images with more than 20% of clouds) for the 2010 map provided by [@ONE2013], or show large mis-classification in specific areas, especially in the dry and spiny forest domain for which the spectral answer has a strong seasonal behavior due to the deciduousness of such forests (overall accuracy is lower than 0.8 for the dry and spiny forests for the maps produced by @ONE2015).
+Moreover, the production of such forest maps from a supervised classification approach requires a significant amount of resources, especially regarding the image selection step (required to minimize cloud cover) and the training step (visual interpretation of a large number of polygons needed to train the classification algorithm) [@Rakotomalala2015].
+Most of this work of image selection and visual interpretation would need to be repeated if one would intend to produce new forest maps in the future using a similar approach.
 
-Global forest products have also been published recently and could be used to derive recent forest cover map at the national scale for Madagascar.
-@Kim2014 produced a global Landsat-based forest-cover change map from 1990 to 2000 but their approach was not reproduced to obtain more recent maps of forest-cover, after 2000.
-Also, the approach used in @Kim2014 did not allow to correctly map the forest in the dry and spiny ecosystems.
-Another global study is the work by @Hansen2013 who identified annual tree cover, forest loss and forest gain from 2000 to 2012.
-Meanwhile, results have been updated and are available up to the year 2014 at <http://earthenginepartners.appspot.com/science-2013-global-forest>.
-One problem with using the @Hansen2013 product to map forest cover is that a tree cover threshold must be defined to transform the tree cover data into a forest/non-forest map.
-Identifying such threshold is not straightforward as forests cannot be defined only by a tree cover value.
-Also, the tree cover threshold might strongly vary between forest types, mean tree cover values being substantially lower for dry forests than for moist forests.
-Moreover, a tree cover gives no information on land-use.
-Therefore, tree plantations such as oil palm or eucalyptus plantations cannot be differenciated from natural forests, which provide very different ecosystem services particularly in terms of biodiversity [@Tropek2014].
-Thus, using @Hansen2013 alone can only give a first-approach estimate of the forest cover.
-But, tree cover is not the only information provided by @Hansen2013.
-The forest loss, i.e.
-the complete loss of forest tree cover (which shifts from a positive value over 10% to zero) is also reported.
+Global forest or tree cover products were also published recently and can be tested at the national scale for Madagascar.
+[@Kim2014] produced a global forest-cover change map from 1990 to 2000 (derived from Landsat imagery).
+This product was updated to cover the period 1975-2005 (<http://glcf.umd.edu/data/landsatFCC/>) but forest-cover maps after 2005 were not produced.
+Moreover, the approach used in @Kim2014 did not allow to correctly map the forests in the dry and spiny ecosystems of Madagascar (see Fig.
+8 in @Kim2014).
+@Hansen2013 mapped tree cover percentage, annual forest loss and forest gain from 2000 to 2012 at global scale and at 30 m resolution.
+Meanwhile, this product was updated and is now available up to the year 2014 [@Hansen et al.
+2013].
+To map forest cover from the @Hansen2013 product, a tree cover threshold must be selected (to define forest cover).
+Selecting such a threshold is not straightforward as the accuracy of the global tree cover map strongly varies between forest types, being substantially lower for dry forests than for moist forests [@Bastin2017].
+Moreover, the @Hansen2013 product does not provide information on land-use.
+In particular the global tree cover map does not separate tree plantations such as oil palm or eucalyptus plantations from natural forests [@Tropek2014].
+Thus, the global tree cover map from @Hansen2013 cannot be used alone to produce a map of forest cover [@Tuykavina2017].
+In complement to the tree cover percentage provided in @Hansen2013, a layer of annual forest loss is also provided (i.e.
+complete loss of tree cover from a value higher than 10% to zero).
 
-In this study, we present a simple approach combining maps from @Harper2007 and products from @Hansen2013 to derive annual and cloud-free forest-cover change map for Madagascar for the period 2000-2014.
-Basically, we used the natural forest-cover map provided by @Harper2007 for the year 2000 (defining the land-use) with the tree cover loss product provided by @Hansen2013 that we applied only inside forest areas identified by @Harper2007.
-Following the work done by @Harper2007, we studied in details the change in deforestation rates and the evolution of forest fragmentation on six decades from _c._ 1953 to 2014.
-We showed that despite all the conservation efforts done in Madagascar since the mid 1980s, deforestation has increased in speed since 2005 and that fragmentation has dramatically increased leading to few remaining intact forest landscapes.
-The approach described in this study could be easilly extended to all tropical countries with a forest map obtained at one date between 2000 and 2014, and could be repeated in the future to obtain future forest-cover change maps, as and when the @Hansen2013 product is updated.
+In this study, we present a simple approach which combines the maps from @Harper2007 and products from @Hansen2013 to derive annual wall-to-wall forest-cover change maps over the period 2000-2014 for Madagascar.
+Similarly to the approach of @Harper2007, we assess then in details the trends in deforestation rates and forest fragmentation from c.
+1953 to 2014.
+The approach described in this study should help assessing the efficiency of present conservation strategies and implementing new strategies for the future.
+Our approach could be easily extended to other tropical countries when at least a forest cover map exist between 2000 and 2014 or repeated in the future as and when the @Hansen2013 products are updated.
 
@@
 
-### 2.1 Forest-cover maps for Madagascar from _c._ 1953 to 2014
+### 2.1 Creation of new forest-cover maps of Madagascar from 1953 to 2014
 
-We derived 30m cloud-free annual forest maps for Madagascar for the period 2000-2014 combining the 2000 forest map by @Harper2007 and the 2000-2014 tree cover and tree cover loss maps by @Hansen2013.
-On the 2000 Harper's forest map, `r as.integer(ha.cloud.2000)` ha of clouds are present, mostly (`r round(perc.cloud.moist)`%) over the moist forest which cover `r round(ha.forest.2000.moist/1e6,2)` million ha.
-To remove these clouds, we used the 2000 cloud-free tree cover map provided by @Hansen2013 (at 30m resolution) and we chose a threshold of 75% of tree cover [@Achard2014] to decide whether to replace clouds with forest or non-forest.
-We thus obtained a cloud-free year 2000 forest-cover map for Madagascar.
-From this map, using tree cover loss data from 2000 to 2014 provided by @Hansen2013, we obtained annual cloud-free forest-cover maps from 2010 to 2014 at 30m resolution.
-We then corrected the cloudy 1990 Harper forest map assuming that if forest was present in 2000, it was also present in 1990.
-Doing so, cloud-cover was reduced to a negligeable quantity of `r residual.clouds.1990` ha.
-For these residual cloud areas, we assumed that forest-cover was the same as for the year 2000.
-We also corrected the cloudy 1973 Harper forest map assuming that if forest was present in 1990, it was also present in 1973.
-Contrary to the year 1990, we were not able to remove all the clouds on the 1973 map and the remaining cloud cover was of `r round(cloud.1973/1e6,2)` million ha.
-Because the 1953 forest-cover map was obtained from the digitalization of a paper map derived from aerial photographies, and thus cannot be perfectly aligned with the maps obtained in the following years, we did not performed any modifications on the original forest map of 1953 except reprojection.
-For all forest-cover maps from 1973, we removed the isolated non-forest pixels if all the surrounding pixels were covered by forest, assuming that small-scaled deforestation on a 30m square pixel inside the forest was unlikely.
-This was important in order not to over-estimate forest fragmentation.
-All these maps are made freely available at https://bioscenemada.cirad.fr/forestmaps.
+We produced annual forest/non-forest maps at 30m resolution for the full territory of Madagascar for the period 2000-2014 by combining the forest map of year 2000 from @Harper2007 and the tree cover percentage and annual forest cover loss maps over the period 2000-2014 from @Hansen2013.
+The 2000 Harper's forest map includes 208,000 ha of unclassified areas due to the presence of clouds in input satellite imagery, mostly (88%) within the moist forest domain which covers 4.17 million ha in total.
+To provide a label (forest or non-forest) to these unclassified pixels, we use the 2000 tree cover percentage map of @Hansen2013 by selecting a threshold of 75% tree cover to define forest cover as recommended by other studies for the moist domain [@Achard2014].
+We thus obtained a forest-cover map for the year 2000 covering the full territory of Madagascar.
+We then combined this forest cover map of the year 2000 with the annual forest cover loss maps from 2001 to 2014 provided by @Hansen2013 to create annual forest cover maps from 2001 to 2014 at 30m resolution.
+In a further step we completed the Harper’s forest map of year 1990 by filling unclassified areas using our forest cover map of year 2000: we assume that if forest was present in 2000, the pixel was also forested in 1990.
+The remaining unclassified pixels were limited to a relatively small total area of c.
+8,000 ha.
+We labeled these residual pixels as non-forest as for the year 2000.
+Similarly we completed the Harper forest map of year 1973 by filling unclassified areas using our forest cover map of the year 1990 assuming that if forest was present in 1990, it was also present in 1973.
+Contrary to the year 1990, the remaining unclassified pixels for year 1973 corresponded to a significant total area of 3.32 million ha, which could not be labeled.
+The forest-cover map of year 1953 was only re-projected to a common projection.
+This map was produced by scanning a paper map derived from aerial photos, and thus could not be perfectly aligned with the other maps produced through digital processing of satellite imagery.
+Finally for all forest-cover maps from 1973, the isolated single non-forest pixels (i.e.
+fully surrounded by forest pixels) were removed, assuming that single non-forest pixels inside a forest patch were not corresponding to deforestation (they might correspond to selective logging activities).
+This allowed avoiding to include very small scale events (<0.1 ha such as selective logging) as forest fragmentation.
+All the resulting maps are freely available at <https://bioscenemada.cirad.fr/forestmaps>.
 
-### 2.2 Computing forest-cover area and deforestation rate
+### 2.2 Computing forest-cover areas and deforestation rates
 
-On the basis of the forest-cover maps we obtained, we computed the forest-cover area (in ha) and the mean annual deforestation rate (in ha/yr and in %/yr) on six decades, for five time periods from _c._ 1953 to 2014 (_c._ 1953-_c._ 1973-1990-2000-2010-2014).
-Mean annual deforestation rates in percentage were computed as follow [@Puyravaud2003; @Vieilledent2013]:
+From these new forest-cover maps  we computed the total forest-cover area for six available years (1953-1973-1990-2000-2010-2014), and the annual deforested area and annual deforestation rate for the corresponding five time periods between 1953 and 2014.
+The annual deforestation rates were computed as follow [@Puyravaud2003; @Vieilledent2013]:
 
@@
 
-In this formula, $\theta$ is the annual deforestation rate (in %/yr), $F_{t_2}$ and $F_{t_1}$ are the forest cover free of clouds at both dates $t_2$ and $t_1$, and $t_2-t_1$ is the time-interval (in years) between the two dates.
-Because of the large cloud-cover (`r round(cloud.1973/1e6,2)` million ha) in 1973, the annual deforestation rates in hectares of forest for the two periods _c._ 1953-1973 and 1973-1990 are only coarse estimates.
-This problem was overcome when computing the annual deforestation rates in %/yr for these two periods.
-In this case, we computed the ratio $(F_{t_2}-F_{t_1})/F_{t_1}$ considering only the forest pixels that were not covered by clouds in 1973.
-Computation were done at the national scale and for each forest type in Madagascar: the moist forest in the East, the dry forest in the West, the spiny forest in the South and mangroves on the West coast (Fig.
+where $\theta$ is the annual deforestation rate (in %/yr), $F_{t_2}$ and $F_{t_1}$ are the forest cover free of clouds at both dates $t_2$ and $t_1$, and $t_2-t_1$ is the time-interval (in years) between the two dates.
+
+Because of the large unclassified area (3.32 million ha) in 1973, the annual deforestation areas and rates for the two periods 1953-1973 and 1973-1990 are only indicative estimates.
+For these two periods the annual deforestation rates are computed as the ratio $(F_{t_2}-F_{t_1})/F_{t_1}$ considering only the mapped forest pixels.
+Area and rate estimates are produced at the national scale and for the four forest ecosystems present in Madagascar: moist forest in the East, dry forest in the West, spiny forest in the South, and mangroves on the Western coast (Fig.
 1).
-To define the forest type, we used a map provided by the MEFT (_"Ministère de l'Environnement et des Forêts à Madagascar"_) showing the boundaries of the four ecoregions in Madagascar.
+To define the forest domains, we used a map from the MEFT (_"Ministère de l'Environnement et des Forêts à Madagascar"_) with the boundaries of the four ecoregions in Madagascar.
 Ecoregions were defined on the basis of climatic and vegetation criteria using the climate classification by @Cornet1974 and the vegetation classification from the 1996 IEFN national forest inventory [@IEFN1996].
-Because mangrove forests are highly dynamic ecosystems that can expand or contract on decadal scales depending on environmental factors [@Armitage2015], a fixed delimitation of the mangrove ecoregion on six decades might not be appropriate.
-As a consequence, our estimates of the forest-cover and deforestation rates for mangroves in Madagascar must be considered with caution.
+Because mangrove forests are highly dynamic ecosystems that can expand or contract on decadal scales depending on changes in environmental factors [@Armitage2015], a fixed delimitation of the mangrove ecoregion on six decades might not be fully appropriate.
+As a consequence, our estimates of the forest-cover and deforestation rates for mangroves in Madagascar must be considered with this limitation.
 
-### 2.3 Comparing forest-cover and deforestation rate estimates with previous studies
+### 2.3 Comparing our forest-cover and deforestation rate estimates with previous studies
 
-We compared our estimates of forest-cover and deforestation rate with three previous studies at the national scale for Madagascar [@Harper2007; @MEFT2009; @ONE2015].
-@Harper2007 provided forest-cover estimates for the periods _c._ 1953-_c._ 1973-1990-2000, @MEFT2009 provided estimates for the periods 1990-2000-2005 and @ONE2015 provided estimates for the periods 2005-2010-2013.
-To be able to compare the forest-cover and deforestation rate estimates on the same time-periods, we considered three additional time-periods in our study: 2000-2005, 2005-2010 and 2010-2013.
-We computed the Pearson's correlation coefficient and the root mean square error (RMSE) between our forest-cover estimates and forest-cover estimates from previous studies at the same date.
-All previous studies provide annual deforestation rate estimates in %/yr.
-The way annual deforestation rates in %/yr have been computed in these previous studies is not always detailed and might slightly differ from one study to another [see @Puyravaud2003].
-But these estimates always correct for the potential presences of residuals clouds on forest maps.
-@Harper2007 also provide deforestation rate estimates in total number of hectares but only for the two periods 1973-1990 and 1990-2000.
-We converted these estimates into annual deforestation rate estimates in ha/yr.
-When annual deforestation rate estimates in ha/yr were not reported (for 1953-1973 in @Harper2007 and in @MEFT2009 and @ONE2015), we computed them from the provided forest-cover estimates in each study.
-These estimates do not correct for the potential presence of residual clouds.
-Forest-cover and deforestation rates were compared between studies for the whole Madagascar and the four ecoregions.
-The same ecoregion boundaries as in our study have been used in the @ONE2015 report but this was not the case for the @Harper2007 article and the @MEFT2009 report, which can explain some of the differences between studies in the forest-cover areas and deforestation rates by ecoregions.
+We compared our estimates of forest-cover and deforestation rates with estimates from the three existing studies at the national scale for Madagascar : (i) [@Harper2007], (ii) [@MEFT2009] and (iii) [@ONE2015].
+@Harper2007 provides forest-cover and deforestation estimates for the periods c.
+1953-c.
+1973-1990-2000.
+MEFT, USAID, and CI (2009) provides estimates for the periods 1990-2000-2005 and ONE, DGF, MNP, WCS, and Etc Terra (2015) provides estimates for the periods 2005-2010-2013.
+To compare our forest-cover and deforestation estimates over the same time periods, we consider three additional time-periods in our study (2000-2005, 2005-2010 and 2010-2013) by creating extra forest cover maps for years 2005 and 2013.
+We computed the Pearson's correlation coefficient and the root mean square error (RMSE) between our forest-cover estimates and forest-cover estimates from previous studies for all the dates and the four forest types.
+For previous studies, the computation of annual deforestation rates (in %/yr) is not always detailed and might slightly differ from one study to another [see @Puyravaud2003].
+@Harper2007 also provides total deforested areas for the two periods 1973-1990 and 1990-2000.
+We converted these values into annual deforested area estimates.
+When annual deforested areas were not reported (for 1953-1973 in @Harper2007 and in @MEFT2009 and @ONE2015), we computed them from the forest-cover estimates in each study.
+These estimates cannot be corrected from the potential bias due to the presence of residual clouds.
+Forest-cover and deforestation rates were then compared between all studies for the whole Madagascar and the four ecoregions.
+The same ecoregion boundaries as in our study were used in @ONE2015 but this was not the case for @Harper2007 and @MEFT2009, which can explain part of the differences between the estimates.
 
@@
 
-We report an analysis of the evolution of forest fragmentation based on the forest-cover maps for Madagascar for the years _c._ 1953, _c._ 1973, 1990, 2000, 2010 and 2014.
-We used the method developed by @Riitters2000 which uses a moving window to characterize the fragmentation around each forested pixel.
+An analysis of the changes in forest fragmentation was carried out from our forest-cover maps over Madagascar for the years 1953, 1973, 1990, 2000, 2010 and 2014.
+We applied the method developed by @Riitters2000 which uses a moving window to characterize the fragmentation around each forested pixel.
 Computation were done using the function `r.forestfrag` of the GRASS GIS software [@Neteler2008].
-Six categories of fragmentation are identified from the amount of forest and its occurrence as adjacent forest pixels: "interior", "perforated", "edge", "transitional", "patch", and "undetermined".
-We used a moving window of 7x7 pixels (4.41 km<sup>2</sup>).
-Using this window size, forest edge had a width of about 90 m [@Riitters2000].
-The interior category can be interpreted as the most intact forest (sensus @Potapov2017).
-We reported the area of forest in each fragmentation category for the six years and analysed the evolution of the fragmentation on the six decades.
-We also computed the distance to forest edge (in km) for all forest pixels for the years _c._ 1953, _c._ 1973, 1990, 2000, 2010 and 2014.
+Six categories of fragmentation were identified from the amount of forest and its occurrence as adjacent forest pixels: "interior", "perforated", "edge", "transitional", "patch", and "undetermined".
+We used a moving window of 7x7 pixels (4.4 ha).
+Using this window size, forest edge had a width of about 90m [@Riitters2000].
+The interior category can be interpreted as the most intact forest [@Potapov2017].
+We reported the area of forest in each fragmentation category for the six years and analyzed the dynamics of fragmentation over the six decades.
+We also computed the distance to forest edge for all forest pixels for the years 1953, 1973, 1990, 2000, 2010 and 2014.
 For that, we used the function `gdal_proximity.py` of the GDAL software (http://www.gdal.org/).
@@
 
-### 3.1 Evolution of forest cover and deforestation intensity
+### 3.1 Dynamics of forest cover and deforestation intensity
 
-Natural forests in Madagascar covered 16.0 Mha in 1953, about 27% of the national territory of 587041 km<sup>2</sup>.
-In 2014, this number has dropped down to 8.9 Mha, about 15% of the national territory (Fig.
+Natural forests in Madagascar covered 16.0 Mha in 1953, about 27% of the national territory of 587,041 km2.
+In 2014, the forest cover has dropped down to 8.9 Mha, corresponding to about 15% of the national territory (Fig.
 2 and Tab.
 1).
-Madagascar has then lost 44% of its natural forest in 60 years since c.
-1953 and a minimum of 37% in 40 years between c.
-1973 and 2014 (Fig.
+Madagascar has lost 44% and 37% of its natural forests between 1953 and 2013, and between c.
+1973 and 2014 respectively (Fig.
 2 and Tab.
 1).
-In 2014, the 8.9 Mha of natural forest in Madagascar were divided into 4.4 Mha (50%) of moist forest, 2.6 Mha (29%) of dry forest, 1.7 Mha of spiny forest (19%) and 176778 ha (2%) of mangrove forest (Fig.
+In 2014 the remaining 8.9 Mha of natural forest were distributed as: 4.4 Mha of moist forest (50% of total forest cover), 2.6 Mha of dry forest (29%), 1.7 Mha of spiny forest (19%) and 0.18 Mha (2%) of mangrove forest (Fig.
 1 and Tab.
 2).
-Regarding the deforestation trend, we observed a progressive decrease of the deforestation rate after 1990 from 204712 ha/yr (1.63%/yr) on the period 1973-1990 to 44296 ha/yr (0.43%/yr) on the period 2000-2005 (Tab.
+Regarding the deforestation trend, we observed a progressive decrease of the deforestation rate after1990 from 205,000 ha/yr (1.63%/yr) over the period 1973-1990 to 44,300 ha/yr (0.43%/yr) over the period 2000-2005 (Tab.
 1).
-Since 2005, deforestation has progressively increased in Madagascar and deforestation rate has more than doubled on the period 2010-2014 (98651 ha/yr, 1.08%/yr) compared to 2000-2005 (Tab.
+Then from 2005, the deforestation rate has progressively increased and has more than doubled over the period 2010-2014 (98,700 ha/yr, 1.08%/yr) compared to 2000-2005 (Tab.
 1).
-The deforestation trend characterized by a progressive decrease of the deforestation rate on the period 1990-2005 and a progressive increase of the deforestation after 2005 is observable for the four ecoregions (Tab.
-3).
-One exception for the spiny forest for the period 2010-2013 for which we estimated a lower deforestation rate compared to 2005-2010 (Tab.
+The deforestation trend characterized by a progressive decrease of the deforestation rate over the period 1990-2005 and a progressive increase of the deforestation after 2005 is valid for all four ecoregions (Tab.
+3), excepting the spiny forest domain for which the deforestation rate during the period 2010-2013 was lower than during 2005-2010 (Tab.
 3).
@@
 
-Forest-cover maps provided by previous studies [@Harper2007; @MEFT2009; @ONE2015] at the national scale for Madagascar were not free of clouds.
-In @Harper2007, 5003 and 11244 km2 of unknown cover type were present for the years 1990 and 2000 respectively.
-Percentage of clouds are not reported in the two other previous studies by @MEFT2009 and @ONE2015.
-With our approach, we were able to obtain cloud-free forest-cover change maps from 1990 to 2014 (Tab.
-1) and thus provide more accurate estimates of forest-cover and deforestation rates on this period.
-Forest-cover estimates in our study were similar (correlation of 0.99) to estimates from the three previous studies (Tab.
-2).
-We estimated a RMSE of 300000 ha (6% of the mean forest cover of 4.8 Mha) between our forest-cover estimates and estimates from previous studies.
-A part of these differences can be attributed to differences in ecoregion boundaries between studies.
-Despite the substantial differences in the deforestation rates estimates between studies (Tab.
-3), the same deforestation trend (decrease of the deforestation rate on the period 1990-2005 followed by a progressive increase of the deforestation after 2005) was observed accross studies.
+Forest-cover maps provided by previous studies over Madagascar were not exhaustive (unclassified areas) due to the presence of clouds in satellite imagery used to produce such maps.
+In @Harper2007, the maps of years 1990 and 2000 include 0.5 and 1.12 Mha of unknown cover type respectively.
+Proportions of unclassified areas are not reported in the two other existing studies by @MEFT2009 and @ONE2015.
+With our approach, we produced wall to wall forest-cover change maps from 1990 to 2014 for the full territory of Madagascar (Tab.
+1).
+This allows to produce more robust estimates of forest-cover and deforestation rates over this period.
+Quite logically, our forest-cover estimates over the period 1953-2013 (considering forest cover estimates at national level and by ecoregions for the 6 available dates) were well correlated  (0.99) to estimates from the three previous studies (Tab.
+2) with a RMSE of 300,000 ha (6% of the mean forest cover of 4.8 Mha when considering all dates and forest types together).
+These small differences can be partly  attributed to differences in ecoregion boundaries.
+Despite significant differences in deforestation estimates (Tab.
+3), a similar deforestation trend was observed across studies with a decrease of deforestation rates over the period 1990-2005 followed by a progressive increase of the deforestation after 2005.
 
@@
 
-Accompagnying deforestion, forest fragmentation has progressively increased since 1953 in Madagascar.
-We observed a continuous decrease of the mean distance to forest edge from 1953 to 2014 in Madagascar (Fig.
+In parallel to the dynamics of deforestation, forest fragmentation has progressively increased since 1953 in Madagascar.
+We observed a continuous decrease of the mean distance to forest edges from 1953 to 2014 in Madagascar.
+The mean distance to forest edge has decreased to c.
+300m in 2014 while it was previously c.
+1km in 1973 (Fig.
 3).
-The mean distance to forest edge has decreased to ~300m in 2014 while it was previously of ~1km in 1973 (Fig.
+Moreover, a large proportion of the natural forests were located far from forest edges in 1973 (between 2km and 4km), while the majority of forests were at less than 1.2 km from forest edges in 2014 (Fig.
 3).
-Moreover, a large part of the natural forest was located far from forest edge (>2km and up to 4km) in 1973, while the majority of the forest was at a distance <1.2 km of the forest edge in 2014 (Fig.
-3).
-The percentage of forest that can be considered intact in Madagascar has progressively decreased since 1953.
-The percentage of forest belonging to the "interior"" category (the most intact forest) has fallen from 68% in 1973 to 50% in 2014 (Tab.
-4).
+The percentage of forest that can be considered intact in Madagascar has continuously decreased since 1953: the percentage of forest belonging to the "interior" category (most intact forests) has fallen from 68% in 1973 to 50% in 2014.
 In 2014, more than 16% of the forest belonged to the "patch" and "transitional" categories (isolated small forest patches) compared to 9.5% in 1973 (Tab.

### 2017-06-01 deforestmap.Rmd
Cover letter

@@
 bibliography: bib/biblio.bib
-csl: bib/biological-conservation.csl
+csl: bib/journal-of-applied-ecology.csl
 output: 
@@
 
+## Highlights
+
+- We produced new 30m-resolution annual forest-cover maps for Madagascar for the period 2000-2014.
+- Madagascar has lost 44% of its natural forest-cover over the period 1953-2014.
+- Half of the tropical forest in Madagascar is now located at less than 100m from forest edge.
+- Annual deforestation rate has increased in Madagascar since 2005 to reach 1.08%/yr.
+- Conservation and development efforts must be intensified to save Madagascar forest and biodiversity.
+
 ## Summary
 
-1.
-_Context:_ The island of Madagascar has an unparalleled biodiversity, both in terms of number of species and endemism.
-This biodiversity is mainly located in the tropical forests of the island and is highly threatened by anthropogenic deforestation, which is associated principally to slash-and-burn agriculture and pasture.
-Scattered forest maps from past studies at national level with substantial gaps (due to presence of cloud cover on satellite imagery) prevent from analyzing the deforestation trends on a long term perspective in Madagascar.
+The island of Madagascar has an unparalleled biodiversity, mainly located in the tropical forests of the island, which is highly threatened by anthropogenic deforestation.
+Scattered forest maps from past studies at national level with substantial gaps (due to presence of cloud cover on satellite imagery) prevent from analyzing the deforestation trends on the long term in Madagascar.
 
-2.
-_Approach:_ In this study, we propose a new approach combining historical (1953-2000) national forest-cover maps with recent (2000-2014) global annual tree cover loss data to look at six decades (1953-2014) of deforestation and forest fragmentation in Madagascar.
-We also produce new forest-cover maps at 30m-resolution over the full territory of Madagascar for the year 1990, and annually from 2000 to 2014.
+In this study, we propose a new approach combining historical (1953-2000) national forest-cover maps with recent (2000-2014) global annual tree cover loss data to look at six decades (1953-2014) of deforestation and forest fragmentation in Madagascar.
+We produced new forest-cover maps at 30m-resolution over the full territory of Madagascar for the year 1990, and annually from 2000 to 2014.
 
-3.
-_Results:_ We estimate that Madagascar has lost 44% of its natural forest cover over the period 1953-2014 (including 37% over the period 1973-2014).
+We estimated that Madagascar has lost 44% of its natural forest cover over the period 1953-2014 (including 37% over the period 1973-2014).
 Natural forests cover 8.9 Mha in 2014 (15% of the national territory) which are divided into 4.4 Mha (50%) of moist forests, 2.6 Mha (29%) of dry forests, 1.7 Mha of spiny forests (19%) and 177,000 ha (2%) of mangroves.
-Since 2005, the annual deforestation rate has progressively increased in Madagascar and has more than doubled from the period 2000-2005 to reach 99,000 ha/yr  during 2010-2014 (corresponding to a rate of 1.08%/yr).
-Most of the forests are now located at less than 1.2km from forest edges while half of them were located far from forest edges in 1973  (i.e.
-between 1km and 4km from forest edges).
-The percentage of core forest has then decreased from 68% in 1973 to 50% in 2014.
+Since 2005, the annual deforestation rate has progressively increased in Madagascar to reach 99,000 ha/yr during 2010-2014 (corresponding to a rate of 1.08%/yr).
+Around half of the forest (46%) is now located at less than 100m from the forest edge.
 
-4.
-_Conclusion:_ Despite the conservation efforts in Madagascar, deforestation has continued to significantly increase at the national level since 2005, with strong consequences on biodiversity loss and carbon emissions.
-This increase in the deforestation rate can be related to a rapid population growth (close to 3%/yr) and to poor law enforcement due to political instability in the country.
+Increase in the deforestation rate can be related to rapid population growth (close to 3%/yr) and to poor law enforcement due to political instability in the country.
 More effort, including conservation and development programs, should be done in Madagascar to protect the remaining natural tropical forests, both to enhance local people livelihoods and answer to global environmental problems.
 
-**Keywords**: biodiversity, climate-change, deforestation, Landsat, protected areas, Madagascar, tropical forest, tree cover.
+**Keywords**: biodiversity, climate-change, deforestation, Landsat, Madagascar, tropical forest.
 
-**Potential journals** (2015 IF): Conservation Letters (7.126), Conservation Biology (4.257), Environmental Research Letters (4.134), Biological Conservation (3.985), Landscape Ecology (3.657), Ambio (2.555), Environmental Conservation (2.235), Biotropica (1.944), Tropical Conservation Science (1.55).
+**Potential journals** (2015 IF): Conservation Letters (7.126), Journal of Applied Ecology (5.196), Conservation Biology (4.257), Environmental Research Letters (4.134), Biological Conservation (3.985), Landscape Ecology (3.657), Ambio (2.555), Environmental Conservation (2.235), Biotropica (1.944), Tropical Conservation Science (1.55).
 
@@
 @Hansen2013 mapped tree cover percentage, annual forest loss and forest gain from 2000 to 2012 at global scale and at 30 m resolution.
-Meanwhile, this product was updated and is now available up to the year 2014 [@Hansen et al.
-2013].
+Meanwhile, this product was updated and is now available up to the year 2014 [@Hansen2013].
 To map forest cover from the @Hansen2013 product, a tree cover threshold must be selected (to define forest cover).
@@
 In particular the global tree cover map does not separate tree plantations such as oil palm or eucalyptus plantations from natural forests [@Tropek2014].
-Thus, the global tree cover map from @Hansen2013 cannot be used alone to produce a map of forest cover [@Tuykavina2017].
+Thus, the global tree cover map from @Hansen2013 cannot be used alone to produce a map of forest cover [@Tyukavina2017].
 In complement to the tree cover percentage provided in @Hansen2013, a layer of annual forest loss is also provided (i.e.
@@
 
-From these new forest-cover maps  we computed the total forest-cover area for six available years (1953-1973-1990-2000-2010-2014), and the annual deforested area and annual deforestation rate for the corresponding five time periods between 1953 and 2014.
+From these new forest-cover maps  we computed the total forest-cover area for seven available years (1953-1973-1990-2000-2010-2005-2014), and the annual deforested area and annual deforestation rate for the corresponding six time periods between 1953 and 2014.
 The annual deforestation rates were computed as follow [@Puyravaud2003; @Vieilledent2013]:
@@
 
-We compared our estimates of forest-cover and deforestation rates with estimates from the three existing studies at the national scale for Madagascar : (i) [@Harper2007], (ii) [@MEFT2009] and (iii) [@ONE2015].
+We compared our estimates of forest-cover and deforestation rates with estimates from the three existing studies at the national scale for Madagascar: (i) [@Harper2007], (ii) [@MEFT2009] and (iii) [@ONE2015].
 @Harper2007 provides forest-cover and deforestation estimates for the periods c.
@@
 MEFT, USAID, and CI (2009) provides estimates for the periods 1990-2000-2005 and ONE, DGF, MNP, WCS, and Etc Terra (2015) provides estimates for the periods 2005-2010-2013.
-To compare our forest-cover and deforestation estimates over the same time periods, we consider three additional time-periods in our study (2000-2005, 2005-2010 and 2010-2013) by creating extra forest cover maps for years 2005 and 2013.
-We computed the Pearson's correlation coefficient and the root mean square error (RMSE) between our forest-cover estimates and forest-cover estimates from previous studies for all the dates and the four forest types.
+To compare our forest-cover and deforestation estimates over the same time periods, we consider an additional time-period in our study (2010-2013) by creating an extra forest cover map for the year 2013.
+We computed the Pearson's correlation coefficient and the root mean square error (RMSE) between our forest-cover estimates and forest-cover estimates from previous studies for all the dates and forest types (including also the total forest cover estimates).
 For previous studies, the computation of annual deforestation rates (in %/yr) is not always detailed and might slightly differ from one study to another [see @Puyravaud2003].
-@Harper2007 also provides total deforested areas for the two periods 1973-1990 and 1990-2000.
+@Harper2007 provides total deforested areas for the two periods 1973-1990 and 1990-2000.
 We converted these values into annual deforested area estimates.
@@
 Using this window size, forest edge had a width of about 90m [@Riitters2000].
-The interior category can be interpreted as the most intact forest [@Potapov2017].
+The "interior" category can be interpreted as the most intact forest [@Potapov2017].
+The "patch" and "transitional" categories correspond to isolated small forest patches.
 We reported the area of forest in each fragmentation category for the six years and analyzed the dynamics of fragmentation over the six decades.
-We also computed the distance to forest edge for all forest pixels for the years 1953, 1973, 1990, 2000, 2010 and 2014.
+We also computed the distance to forest edge for all forest pixels for the years 1953, 1973, 1990, 2000, 2005, 2010 and 2014.
 For that, we used the function `gdal_proximity.py` of the GDAL software (http://www.gdal.org/).
@@
 300m in 2014 while it was previously c.
-1km in 1973 (Fig.
+1.5km in 1973 (Fig.
 3).
-Moreover, a large proportion of the natural forests were located far from forest edges in 1973 (between 2km and 4km), while the majority of forests were at less than 1.2 km from forest edges in 2014 (Fig.
+Moreover, a large proportion (73%) of the forest was located at a distance greater than 100m in 1973 while in 2014, almost half of the forest (46%) is at a distance lower than 100m from forest edge (Fig.
 3).
@@
 
-### 4.1 Interest of combining recent global annual tree cover loss data with historical national forest-cover maps
+### 4.1 Benefits of the combined use of recent global annual tree cover loss data with historical national forest-cover maps
 
 In this study, we combined recent (2001-2014) global annual tree cover loss data [@Hansen2013] with historical (1953-2000) national forest-cover maps [@Harper2007] to look at six decades (1953-2014) of deforestation and forest fragmentation in Madagascar.
-We obtained annual cloud-free 30m-resolution forest-cover maps for Madagascar for the period 2000 to 2014.
-This approach overcomes the problem associated to the forest definition which is necessary to transform the 2000 global tree cover data by @Hansen2013 into a forest/non-forest map [@Tropek2014].
-Instead, an historical national or regional forest-cover map, based on a more appropriate forest definition at the national or regional scale, can be used as a mask.
-This approach could be easilly extended to other regions or countries with available forest-cover maps at one date in the period 2000-2014.
-Moreover, that approach can be repeated in the future to obtain future forest-cover maps, as and when the @Hansen2013 product is updated.
-The R/GRASS code used for this study is made freely available in a GitHub repository (see Data availability statement) and should be easilly transposable to other study sites or re-used for future analysis in Madagascar.
-Of course, the accuracy of the derived forest-cover change maps depends directly on the accuracy of the historical forest-cover maps and the accuracy of the tree cover loss data.
-Reported global accuracy for tree cover loss data on the period 2000-2012 is 99.6% (see Tab.
+We produced annual forest-cover maps at 30m-resolution covering Madagascar for the period 2000 to 2014.
+Our study extends the forest-cover monitoring on a six decades period (from 1953 to 2014) while harmonizing the data from the previous studies.
+This approach tackles the issue associated with the forest definition which is needed to transform the 2000 global tree cover dataset from Hansen et al.
+(2013) into a forest/non-forest map [@Tropek2014].
+We demonstrate that an historical national forest-cover map, based on an national forest definition can be used as a mask.
+This approach could be easily extended to other regions or countries for which an accurate forest-cover map is available at any date within the period 2000-2014, but preferably at the beginning of the period to take profit of the full period to derive estimates of deforestation.
+Moreover, this approach can be repeated in the future when the global tree cover product will be updated.
+The R/GRASS code used for this study is made freely available in a GitHub repository (see Data availability statement) and can be applied to other study areas.
+
+The accuracy of the derived forest-cover change maps depends directly on the accuracies of the historical forest-cover maps and the tree cover loss dataset.
+The reported global accuracy of the tree cover loss dataset is 99.6% (see Tab.
 S5 in @Hansen2013).
-@Verhegghen2016 also found similar annual deforestation rate estimates when comparing results derived from the tree cover loss data by @Hansen2013 on the period 2000-2012 with results derived from semi-supervized classification of Landsat satellite images on the period 2000-2010 by @Achard2014 for six countries in Central Africa.
-Consistently with the study by @Harper2007, we did not consider the potential natural forest regrowth in Madagascar, although @Hansen2013 are also providing tree cover gain data over the period 2001-2014.
-We made this choice for several reasons.
-First, gain data provided by @Hansen2013 can figure plantations and not natural forests [@Tropek2014].
-Second, there is little evidence of natural forest regeneration in Madagascar on the long term [@Harper2007, @Grouzis2001].
-This can be explained by several ecological mechanisms following burning practice such as soil erosion [@Grinand2017] and reduced seed bank due to fire and soil loss [@Grouzis2001].
-Moreover, in the areas where forest regeneration is ecologically possible, young regrowth forests are more easilly re-burnt for agriculture and pasture.
-Third, young secondary forests do not provide the same ecosystem services (in terms of biodiversity and carbon storage) as old-growth natural forests.
+@Verhegghen2016 have compared deforestation estimates derived from the global tree cover loss dataset [@Hansen2013] with results derived from semi-automated supervised classification of Landsat satellite images [@Achard2014] for six countries in Central Africa and they found a good agreement between these two sets of estimates.
+Consistently with @Harper2007, we did not consider potential forest regrowths in Madagascar (although @Hansen2013 provided a tree cover gains layer for the period 2001-2013) for the following reasons: (i) the tree gain layer of @Hansen2013 includes and catches more easily tree plantations than natural forest regrowth [@Tropek2014]; (ii) there is little evidence of natural forest regeneration in Madagascar [@Harper2007, @Grouzis2001] - this can be explained by several ecological processes following burning practice such as soil erosion [@Grinand2017] and reduced seed bank due to fire and soil loss [@Grouzis2001], (iii) in areas where forest regeneration is ecologically possible, young forest regrowth are more easily re-burnt for agriculture and pasture; and (iv) young secondary forests provide more limited ecosystem services compared to old-growth natural forests in terms of biodiversity and carbon storage.
 
-### 4.2 Evolution of forest-cover in Madagascar since 1953 and current forest-cover
+### 4.2 Dynamics of forest-cover in Madagascar from 1953 to 2014
 
-We estimated that natural forests covered 8.9 Mha in Madagascar (15% of the national territory of 587041 km<sup>2</sup>) in 2014 and that Madagascar has lost 44% of its natural forest in six decades since 1953 (37% since 1973).
-There is an age-old debate [@Burns2016] about the original forest cover in Madagascar to know to which extent humans have altered the natural forest landscape since their large-scale settlement around 800 CE [@Cox2012].
-French naturalists Henri Humbert [@Humbert1927] and Henri Perrier de la Bâthie [@Perrier1921] stated that the full island was originally covered by forest, leading to the common statement that 90% of the natural forests have disappeared since the arrival of humans on the island [@Kull2000].
+We estimated that natural forests in Madagascar cover 8.9 Mha in 2014 (corresponding to 15% of the country) and that Madagascar has lost 44% of its natural forest since 1953 (37% since 1973).
+There is an age-old scientific debate about the original forest cover in Madagascar to know to which extent humans have altered the natural forest landscapes since their large-scale settlement around 800 CE [@Burns2016; @Cox2012].
+French naturalists stated that the full island was originally covered by forest [@Humbert1927; @Perrier1921], leading to the common statement that 90% of the natural forests have disappeared since the arrival of humans on the island [@Kull2000].
 More recent studies counter-balanced that point of view saying that grasslands existed in Madagascar long before human arrival and were determined by climate, natural grazing and other natural factors [@Virah-Sawmy2009, @Vorontsova2017].
 Other authors have questioned the entire narrative of extensive alteration of the landscape by early human activity which, through legislation, has severe consequences on local people [@Klein2002; @Kull2000].
-Whatever the original proportion of natural forests and grasslands in Madagascar, our results demonstrate that human activities since the 5Os have profoundly impacted the natural tropical forests and that recent conservation and development programs in Madagascar have failed to stop deforestation in the recent years.
+Whatever the original proportion of natural forests and grasslands in Madagascar, our results demonstrate that human activities since the 1950s have profoundly impacted the natural tropical forests and that recent conservation and development programs in Madagascar have failed to stop deforestation in the recent years.
 Deforestation has strong consequences on biodiversity and carbon emissions in Madagascar.
-Around 90% of Madagascar's species are forest dependent [@Allnutt2008; @Goodman2005] and @Allnutt2008 estimated that deforestation between 1953 and 2000 had led to an extinction of 9% of the species.
-The additional deforestation we observed on the period 2000-2014 (around 1Mha of deforested natural forest) worsen this result.
-Regarding carbon emissions, using the 2010 aboveground forest carbon map by @Vieilledent2016, we estimated that deforestation on the period 2010-2014 has led to the emissions of 40224 Gg (1 Gg = 10<sup>9</sup> g) of carbon in the atmosphere (10055 Gg/yr) and that the residual aboveground forest carbon stock was of 832822 Gg in 2014.
-Associated to deforestation, we showed that residual forest in Madagascar was highly fragmented with 95% of the forest being at a distance <1.2 km of the forest edge in 2014.
-Forest fragmentation creates forest fragments too small to maintain viable populations and increases 'edge effects' at forest/non-forest interfaces having great impact on both carbon emissions [@Brinck2017] and biodiversity loss [@Gibson2013; @Murcia1995].
+Around 90% of Madagascar's species are forest dependent [@Allnutt2008; @Goodman2005] and @Allnutt2008 estimated that deforestation between 1953 and 2000 led to an extinction of 9% of the species.
+The additional deforestation we observed over the period 2000-2014 (around 1Mha of natural forest) worsen this result.
+Regarding carbon emissions, using the 2010 aboveground forest carbon map by @Vieilledent2016, we estimated that deforestation on the period 2010-2014 has led to 40.2 Mt C of carbon emissions in the atmosphere (10 Mt C /yr) and that the remaining aboveground forest carbon stock in 2014 is 832.8 Mt C.
+Associated to deforestation, we showed that the remaining forests of Madagascar are highly fragmented with 46% of the forest being at less than 100m of the forest edge.
+Small forest fragments do not allow to maintain viable populations and 'edge effects' at forest/non-forest interfaces have impacts on both carbon emissions [@Brinck2017] and biodiversity loss [@Gibson2013; @Murcia1995].
 
@@
 In our study, we have shown that the progressive decrease of the deforestation rate on the period 1990-2005 was followed by a continuous increase in the deforestation rate on the period 2005-2014.
-We showed that deforestation rate has more than doubled on the period 2010-2014 (98651 ha/yr, 1.08%/yr) compared to 2000-2005 (42296 ha.yr, 0.43%/yr).
-These results are confirmed by previous studies on deforestation at the national scale for Madagascar [@Harper2007; @MEFT2009; @ONE2015].
-Despite differences in the methodologies regarding (i) forest definition (associated to independant visual interpretations of observation polygons to train the classifier), (ii) classification algorithms (Maximum Likelihood Classification for @Harper2007 and @MEFT2009 vs.
-Random Forests for @ONE2015), (iii) deforestation rate computation method, and (iii) correction for the presence of clouds, these studies gave similar forest-cover estimates and show the same deforestation trend over Madagascar.
-Our deforestation rate estimates from 1990 to 2014 have been computed from cloud-free maps and can be considered more accurate in comparison with estimates from these previous studies, in particular for the deforestation rates expressed in number of hectares per year.
-Forest-cover and deforestation rate estimates from this study could be used to update the estimates in the upcoming FAO Forest Resources Assessment reports [@Keenan2015].
-Current rates of deforestation for Madagascar (98651 ha/yr, 1.08 %/yr for 2010-2014) could also be used to build reference scenarios for deforestation in Madagascar and help implement deforestation mitigation initatives such as the REDD+ [@Olander2008].
-The increase of the deforestation rates we observed after 2005 could be attributed to both population growth and political instability in the country.
+We showed in particular that deforestation rate has more than doubled on the period 2010-2014 compared to 2000-2005.
+Our results are confirmed by previous studies [@Harper2007; @MEFT2009; @ONE2015] despite differences in the methodologies regarding (i) forest definition (associated to independent visual interpretations of observation polygons to train the classifier), (ii) classification algorithms, (iii) deforestation rate computation method, and (iv) correction for the presence of clouds.
+Our deforestation rate estimates from 1990 to 2014 have been computed from wall to wall maps at 30 m resolution and can be considered more accurate in comparison with estimates from these previous studies.
+Our forest-cover and deforestation rate estimates can be used as source of information for the next FAO Forest Resources Assessment project [@Keenan2015].
+Current rates of deforestation can also be used to build reference scenarios for deforestation in Madagascar and contribute to the implementation of deforestation mitigation activities in the framework of REDD+ [@Olander2008].
+
+The increase of deforestation rates after 2005 can be explained by population growth and political instability in the country.
 Nearly 90% of Madagascar's population relies on biomass for their daily energy needs [@Minten2013] and the link between population size and deforestation has previously been demonstrated in Madagascar [@Vieilledent2013; @Gorenflo2011].
 With a mean demographic growth rate of about 2.8%/yr and a population which has increased from 16 to 24 million people on the period 2000-2015 [@UN2015], the increasing demand in wood-fuel and space for agriculture is likely to explain the increase in deforestation rates.
-The political crisis of 2009 [@Ploch2012], followed by several years of political instability and weak governance could also explain the increase in the deforestation rate observed on the period 2005-2014 [@Smith2003].
+The political crisis of 2009 [@Ploch2012], followed by several years of political instability and weak governance could also explain the increase in the deforestation rate observed on the period 2005-2014 (Smith et al., 2003).
 These results show that despite the conservation policy in Madagascar [@Freudenberger2010], deforestation has dramatically increased at the national level since 2005.
@@
 tab.1 <- defor.df
-names(tab.1) <- c("Year", "Forest (ha)", "Clouds (ha)", "Annual defor. (ha/yr)", "theta (%/yr)")
+names(tab.1) <- c("Year", "Forest (ha)", "Unmap (ha)", "Annual defor. (ha/yr)", "Rate (%/yr)")
 kable(tab.1)
@@
 Table 1: **Evolution of the forest cover and deforestation rates from 1950s to 2014 in Madagascar**.
-Madagascar has lost `r perc.1`% of its natural forest in 40 years between _c._ 1973 and 2014 (and about `r perc.2`% in 60 years since _c._ 1953).
-Deforestation has slew down after 1990 (from `r d.1` in 1973-1990 to `r d.2`%/yr in 2000-2010), most likely due to political changes (end of the Second Republic of Madagascar in 1992), a decrease of available arable lands under the remaining forests and to a decrease of forest accessibility.
-But more recently, deforestation has increased, with a doubling deforestation on the period 2010-2014 compared to 2000-2010.
-This increase could be attributed to both political instability (political crisis in 2009) and to an explosion of the demography (according to the @UN2015, population has increased from about 16M inhabitants in 2000 to 24M in 2014).
-Natural forests cover 8.9 Mha in 2014 and current (period 2000-2014) mean annual deforestation rate is of about 70000 ha/yr.
+Forest map for the year 1973 has 3.3 Mha of unclassified areas due to the presence of clouds on satellite images.
+As a consequence, deforestation rates for the periods 1953-1973 and 1973-1990 are indicative.
+The two last columns indicate the deforestation rates on the previous time-period (e.g.
+1953-1973 for year 1973, 1973-1990 for year 1990, etc.).
 
@@
 
-Table 2: **Comparison of Madagascar forest-cover estimates (in ha) with previous studies on the period 1953-2014**.
-We found forest-cover estimates similar to previous studies [@Harper2007;@MEFT2009;@ONE2015].
-The increase in mangrove and spiny forest covers from _c._ 1953 to _c._ 1973 in @Harper2007 and our study is most probably due to differences in forest definition and mapping methods between the 1953 aerial-photography derived map and the 1973 Landsat image derived map.
-Our study extends the forest-cover monitoring on a six decades period (from 1953 to 2014) while harmonizing the data from the previous studies.
-Our method provides cloud-free forest-cover estimates from 1990 to 2014 and corrections for the forest-cover in 1973.
+Table 2: **Comparing our estimates of forest-cover (in ha) for Madagascar with previous studies on the period 1953-2014**.
+We compared our forest-cover with the estimates from three previous studies [@Harper2007;@MEFT2009;@ONE2015].
+We computed a The increase in mangrove and spiny forest covers from _c._ 1953 to _c._ 1973 in @Harper2007 and our study is most probably due to differences in forest definition and mapping methods between the 1953 aerial-photography derived map and the 1973 Landsat image derived map.
 
@@
 
-Table 3: **Comparison of Madagascar annual deforestation rates with previous studies on the period 1953-2014**.
-Annual deforestation rates are providing in ha/yr and in %/yr (second number in parenthesis).
-For deforestation rate in %/yr, exact same numbers as in scientific articles and reports from previous studies have been reported.
+Table 3: **Comparing our estimates of annual deforestation rates for Madagascar with previous studies on the period 1953-2014**.
+Annual deforestation areas (in ha/yr) and annual deforestation rates (second number in parenthesis, in %/yr) are provided.
+For deforestation rates in %/yr, exact same numbers as in scientific articles and reports from previous studies have been reported.
 The way annual deforestation rates in %/yr have been computed in these previous studies can slightly differ from one study to another but estimates always correct for the potential presences of clouds on forest maps.
-Deforestation rates in ha/yr have been recomputed from forest-cover estimates in Tab.
-2 (except for @Harper2007 for the periods 1973-1990 and 1990-2000 for which annual deforestation rates in ha/yr were derived from numbers reported in the original publication, see methods) and do not correct for the potential presence of clouds.
+Annual deforested areas in ha/yr have been recomputed from forest-cover estimates in Tab.
+2 (except for @Harper2007 for the periods 1973-1990 and 1990-2000 for which annual deforested ares in ha/yr were derived from numbers reported in the original publication, see methods) and do not correct for the potential presence of clouds.
 
@@
 Table 4: **Evolution of the forest fragmentation from 1950s to 2014 in Madagascar**.
-Forest fragmentation has progressively increased since 1953.
-The percentage of forest belonging to the "interior"" category (the most intact forest) has fallen from `r p.1`% in 1973 to `r p.2`% in 2014.
-More than `r p.3`% of the forest belong to the "patch" and "transitional" categories (isolated small forest patches) in 2014.
+Six categories of fragmentation were identified from the amount of forest and its occurrence as adjacent forest pixels: "interior", "perforated", "edge", "transitional", "patch", and "undetermined" [@Riitters2000].
+We used a moving window of 7x7 pixels (4.4 ha).
+Using this window size, forest edge had a width of about 90m.
+The "interior" category can be interpreted as the most intact forest.
+The "patch" and "transitional" categories correspond to isolated small forest patches.
 
@@
 ![Figure 1: **Ecoregions and forest types in Madagascar.** Madagascar can be divided into four climatic ecoregions with four forest types: the moist forest in the East (green), the dry forest in the West (orange), the spiny forest in the South (red), and the mangroves on the West coast (blue).
-Ecoregions were defined following climatic (see @Cornet1974) and vegetation (see @IEFN1996) criteria.
+Ecoregions were defined following climatic [@Cornet1974] and vegetation [@IEFN1996] criteria.
 The dark grey areas represent the remaining natural forest cover for the year 2014.](outputs/ecoregion.png)
@@
 
-![Figure 3: **Evolution of the mean distance to forest edge from 1953 to 2014 in Madagascar.** Black dots represent the mean distance to forest edge for each year.
-Vertical dashed segments represent the 95% quantiles of the distance to forest edge.
-We observed a continuous decrease of the mean distance to forest edge from 1953 to 2014.
-The mean distance was of ~1km in 1973 while it was of ~300m in 2014.
-Moreover, many areas were located far from forest edge (>2km and up to 4km) in 1973, while the majority of the forest was at a distance <1.2 km of the forest edge in 2014.](outputs/dist.png)
+![Figure 3: **Evolution of the distance to forest edge from 1953 to 2014 in Madagascar.** Black dots represent the mean distance to forest edge for each year.
+Vertical dashed segments represent the 90% quantiles (5% and 95%) of the distance to forest edge.
+Horizontal dashed grey line indicates a distance to forest edge of 100m.
+Percentages indicate the percentage of forest at a distance to forest edge lower than 100m for each year.](outputs/dist.png)
 

### 2017-06-01 deforestmap.R

@@ -515,6 +515,22 @@ for (i in 1:length(Year)) {
 }
 write.table(frag.df,"outputs/frag.txt",row.names=FALSE,sep="\t")
 
+#= Distance to forest edge statistics ## Quantiles with r.quantile
+Year <- c(1953,1973,1990,2000,2005,2010,2014)
+dist.quant.df <- data.frame(Year=Year,median=NA,q1=NA,q2=NA)
+# Loop on year
+for (i in 1:length(Year)) {
+  # Message
+  cat(paste("Year: ",Year[i],"\n",sep=""))
+  # Computation
+  statcell <- system(paste0("r.quantile percentiles=5,50,95 input=dist_edge_",Year[i]), intern=TRUE)
+  mat.statcell <- matrix(as.numeric(unlist(strsplit(statcell,split=":"))), byrow=TRUE, nrow=3)
+  dist.quant.df$median[i] <- mat.statcell[2,3]
+  dist.quant.df$q1[i] <- mat.statcell[1,3]
+  dist.quant.df$q2[i] <- mat.statcell[3,3]
+}
+write.table(dist.quant.df,"outputs/dist.quant.txt",row.names=FALSE,sep="\t")
+
 #= Distance to forest edge statistics
 Year <- c(1953,1973,1990,2000,2005,2010,2014)
 dist.df <- data.frame(Year=Year,min=NA,max=NA,mean=NA,sd=NA,q1=NA,q2=NA)
@@ -531,7 +547,74 @@ for (i in 1:length(Year)) {
   dist.df$q1[i] <- unlist(strsplit(statcell[19],split=" "))[3]
   dist.df$q2[i] <- unlist(strsplit(statcell[20],split=" "))[3]
 }
+# Replacing approximated quantiles with exact values
+dist.df$q1 <- dist.quant.df$q1
+dist.df$q2 <- dist.quant.df$q2
 write.table(dist.df,"outputs/dist.txt",row.names=FALSE,sep="\t")
+  
+#= Percentage of forest at less than 500m from the forest edge
+Year <- c(1953,1973,1990,2000,2005,2010,2014)
+perc.500m.dist.df <- data.frame(Year=Year,perc=NA)
+# Intermediate rast
+for (i in 1:length(Year)) {
+  # Message
+  cat(paste("Year: ",Year[i],"\n",sep=""))
+  # Computation
[…]

### 2017-06-01 deforestmap.R
Cover letter

@@ -927,7 +927,7 @@ dir.create("report")
 ## Main document
 render("deforestmap.Rmd",output_dir="report") # html output
 ## Cover letter
-#render("coverletter.md",output_format=c("pdf_document"),output_dir="report") # pdf output
+render("coverletter.md",output_format=c("pdf_document"),output_dir="report") # pdf output
 #render("deforestmap.Rmd",output_format=c("pdf_document","word_document"),output_dir="report")
 
 ##===========================================================================

### 2017-06-03 deforestmap.R
Frag 2005

@@ -416,8 +416,7 @@ for (i in 1:length(Year)) {
   # Computation
   if (Year[i]==1973) {  # ! 1975 with 5 (clouds) 
     system(paste0("r.mapcalc --o 'for1973_0 = if(!isnull(for1973) &&& for1973==1, 1, 0)'"))
-  }
-  else {
+  } else {
     system(paste0("r.mapcalc --o 'for",Year[i],"_0 = if(!isnull(for",Year[i],"), 1, 0)'"))
   }
   system(paste0("r.forestfrag input=for",Year[i],"_0 output=frag",Year[i]," size=7 --overwrite"))

### 2017-06-06 deforestmap.Rmd
Corrections by Tom

@@
 #### Ghislain Vieilledent<sup>*,1,2</sup>, Clovis Grinand<sup>3</sup>, Fety A.
-Rakotomalala<sup>3</sup>, Rija Ranaivosoa<sup>5</sup>, Jean-Roger Rakotoarijaona<sup>5</sup>, Thomas F.
-Allnutt<sup>4</sup>, and Frédéric Achard<sup>1</sup>
+Rakotomalala<sup>3</sup>, Rija Ranaivosoa<sup>4</sup>, Jean-Roger Rakotoarijaona<sup>4</sup>, Thomas F.
+Allnutt<sup>5,6</sup>, and Frédéric Achard<sup>1</sup>
 
@@
 3. ETC Terra, F-75016 Paris, FRANCE
-4. Wildlife Conservation Society, 101 Antananarivo, MADAGASCAR
-5. Office National pour l'Environnement, 101 Antananarivo, MADAGASCAR
+4. Office National pour l'Environnement, 101 Antananarivo, MADAGASCAR
+5. Wildlife Conservation Society, 101 Antananarivo, MADAGASCAR
+6. GreenInfo Network, Oakland, California, USA
+
+*Running headline:* Six decades of deforestation in Madagascar
 
@@
 
+1.
 The island of Madagascar has an unparalleled biodiversity, mainly located in the tropical forests of the island, which is highly threatened by anthropogenic deforestation.
-Scattered forest maps from past studies at national level with substantial gaps (due to presence of cloud cover on satellite imagery) prevent from analyzing the deforestation trends on the long term in Madagascar.
+Scattered forest maps from past studies at national level with substantial gaps (due to presence of cloud cover on satellite imagery) prevent the analyzis of long-term deforestation trends in Madagascar.
 
+2.
 In this study, we propose a new approach combining historical (1953-2000) national forest-cover maps with recent (2000-2014) global annual tree cover loss data to look at six decades (1953-2014) of deforestation and forest fragmentation in Madagascar.
@@
 
+3.
 We estimated that Madagascar has lost 44% of its natural forest cover over the period 1953-2014 (including 37% over the period 1973-2014).
@@
 
-Increase in the deforestation rate can be related to rapid population growth (close to 3%/yr) and to poor law enforcement due to political instability in the country.
-More effort, including conservation and development programs, should be done in Madagascar to protect the remaining natural tropical forests, both to enhance local people livelihoods and answer to global environmental problems.
+4.
+_Policy implications_: Increase in the deforestation rate since 2005 can be related to rapid population growth (close to 3%/yr) and to poor law enforcement due to political instability in the country.
+More effort, including conservation and development programs, is needed in Madagascar to protect the remaining natural tropical forests, both to enhance local people's livelihoods and to support global environmental efforts.
 
@@
 Separated from the African continent and the Indian plate about 165 and 88 million years ago respectively [@Ali2008], the flora and fauna of Madagascar followed its own evolutionary path.
-Combined with a high number of micro-habitats [@Pearson2009], this has led to Madagascar's exceptional biodiversity both in term of number of species and endemism in many taxonomic groups [@Crottini2012; @Goodman2005].
-Most of the biodiversity in Madagascar is concentrated in the tropical forests of the island which can be divided in three types: the moist forest in the East, the dry forest in the West and the spiny forest in the South [@Vieilledent2016].
-This unparalleled biodiversity is severely threatened by anthropogenic deforestation [@Harper2007; @Vieilledent2013] associated to activities such as slash-and-burn agriculture and pasture [@Scales2011].
-Tropical forests in Madagascar also store a large amount of carbon [@Vieilledent2016] and high rates of deforestation in Madagascar are responsible for large CO2 emissions in the atmosphere [@Achard2014].
-Deforestation threatens species survival by reducing their habitat [@Brooks2002; @Tidd2001].
-Associated to deforestation, forest fragmentation can also lead to species extinction by isolating populations from each others and creating forest patches too small to maintain viable populations [@Saunders1991].
-Fragmentation also extents forest edge where ecological conditions (such as air temperature, light intensity and air moisture) can be dramatically modified, with consequences on the abundance and distribution of species [@Murcia1995].
-Forest fragmentation can also have substantial effects on forest carbon storage capacity, carbon stocks being much lower at the forest edge than under a closed canopy [@Brinck2017].
-Moreover, forest carbon stocks are varying spatially due to climate or soil factors [@Saatchi2011; @Vieilledent2016].
-As a consequence, accurate and spatially explicit maps of forest-cover and forest-cover change are necessary to monitor loss of biodiversity and carbon emissions associated to deforestation and forest fragmentation, assess the efficiency of present conservation strategies [@Eklund2016], and implement new strategies for the future [@Vieilledent2013; @Vieilledent2016].
+Isolation combined with a high number of micro-habitats [@Pearson2009] has led to Madagascar's exceptional biodiversity both in term of number of species and endemism in many taxonomic groups [@Crottini2012; @Goodman2005].
+Most of the biodiversity in Madagascar is concentrated in the tropical forests of the island which can be divided into three types: the moist forest in the East, the dry forest in the West and the spiny forest in the South [@Vieilledent2016].
+This unparalleled biodiversity is severely threatened by anthropogenic deforestation [@Harper2007; @Vieilledent2013] associated with activities such as slash-and-burn agriculture and pasture [@Scales2011].
+Tropical forests in Madagascar also store a large amount of carbon [@Vieilledent2016] and high rates of deforestation in Madagascar are responsible for large CO<sub>2</sub> emissions in the atmosphere [@Achard2014].
+Deforestation threatens species survival by directly reducing their available habitat [@Brooks2002; @Tidd2001].
+Forest fragmentation can also lead to species extinction by isolating populations from each other and creating forest patches too small to maintain viable populations [@Saunders1991].
+Fragmentation also increases forest edge where ecological conditions (such as air temperature, light intensity and air moisture) can be dramatically modified, with consequences on the abundance and distribution of species [@Murcia1995].
+Forest fragmentation can also have substantial effects on forest carbon storage capacity, as carbon stocks are much lower at the forest edge than under a closed canopy [@Brinck2017].
+Moreover, forest carbon stocks vary spatially due to climate or soil factors [@Saatchi2011; @Vieilledent2016].
+As a consequence, accurate and spatially explicit maps of forest-cover and forest-cover change are necessary to monitor biodiversity loss and carbon emissions from deforestation and forest fragmentation, assess the efficiency of present conservation strategies [@Eklund2016], and implement new strategies for the future [@Vieilledent2013; @Vieilledent2016].
 Simple time-series of forest-cover estimates, such as those provided by the FAO Forest Resource Assessment report [@Keenan2015] are not sufficient.
@@
 11 244 km2 are mapped as unknown cover type for the year 2000.
-Using a similar supervised classification approach as in @Harper2007, more recent maps were produced for the periods 2000-2005-2010 by national institutions, with the technical support of international environmental NGOs [@MEFT2009; @ONE2013].
-Another set of recent forest cover maps using an advanced statistical tool for classification, the Random Forest classifier [@Grinand2013; @Rakotomalala2015], was produced for the periods 2005-2010-2013 [@ONE2015].
+Using a similar supervised classification approach as in @Harper2007, more recent maps have been produced for the periods 2000-2005-2010 by national institutions, with the technical support of international environmental NGOs [@MEFT2009; @ONE2013].
+Another set of recent forest-cover maps using an advanced statistical tool for classification, the Random Forest classifier [@Grinand2013; @Rakotomalala2015], was produced for the periods 2005-2010-2013 [@ONE2015].
 However, these maps are either too old to give recent estimates of deforestation [@MEFT2009; @ONE2013], include large areas of missing information (due to images with more than 20% of clouds) for the 2010 map provided by [@ONE2013], or show large mis-classification in specific areas, especially in the dry and spiny forest domain for which the spectral answer has a strong seasonal behavior due to the deciduousness of such forests (overall accuracy is lower than 0.8 for the dry and spiny forests for the maps produced by @ONE2015).
-Moreover, the production of such forest maps from a supervised classification approach requires a significant amount of resources, especially regarding the image selection step (required to minimize cloud cover) and the training step (visual interpretation of a large number of polygons needed to train the classification algorithm) [@Rakotomalala2015].
-Most of this work of image selection and visual interpretation would need to be repeated if one would intend to produce new forest maps in the future using a similar approach.
+Moreover, the production of such forest maps from a supervised classification approach requires significant resources, especially regarding the image selection step (required to minimize cloud cover) and the training step (visual interpretation of a large number of polygons needed to train the classification algorithm) [@Rakotomalala2015].
+Most of this work of image selection and visual interpretation would need to be repeated to produce new forest maps in the future using a similar approach.
 
-Global forest or tree cover products were also published recently and can be tested at the national scale for Madagascar.
+Global forest or tree cover products have also been published recently and can be tested at the national scale for Madagascar.
 [@Kim2014] produced a global forest-cover change map from 1990 to 2000 (derived from Landsat imagery).
 This product was updated to cover the period 1975-2005 (<http://glcf.umd.edu/data/landsatFCC/>) but forest-cover maps after 2005 were not produced.
-Moreover, the approach used in @Kim2014 did not allow to correctly map the forests in the dry and spiny ecosystems of Madagascar (see Fig.
+Moreover, the approach used in @Kim2014 did not accurately map the forests in the dry and spiny ecosystems of Madagascar (see Fig.
 8 in @Kim2014).
-@Hansen2013 mapped tree cover percentage, annual forest loss and forest gain from 2000 to 2012 at global scale and at 30 m resolution.
-Meanwhile, this product was updated and is now available up to the year 2014 [@Hansen2013].
-To map forest cover from the @Hansen2013 product, a tree cover threshold must be selected (to define forest cover).
-Selecting such a threshold is not straightforward as the accuracy of the global tree cover map strongly varies between forest types, being substantially lower for dry forests than for moist forests [@Bastin2017].
+@Hansen2013 mapped tree cover percentage, annual forest loss and forest gain from 2000 to 2012 at global scale at 30 m resolution.
+This product has since been updated and is now available up to the year 2014 [@Hansen2013].
+To map forest cover from the @Hansen2013 product, a tree cover threshold must be selected (that defines forest cover).
+Selecting such a threshold is not straightforward as the accuracy of the global tree cover map strongly varies between forest types, and is substantially lower for dry forests than for moist forests [@Bastin2017].
 Moreover, the @Hansen2013 product does not provide information on land-use.
@@
 Thus, the global tree cover map from @Hansen2013 cannot be used alone to produce a map of forest cover [@Tyukavina2017].
-In complement to the tree cover percentage provided in @Hansen2013, a layer of annual forest loss is also provided (i.e.
-complete loss of tree cover from a value higher than 10% to zero).
+In complement to the tree cover percentage provided in @Hansen2013, a layer of annual tree cover loss is also provided (i.e.
+complete loss of tree cover from a value higher than 10% to zero) for the period 2001-2014.
 
 In this study, we present a simple approach which combines the maps from @Harper2007 and products from @Hansen2013 to derive annual wall-to-wall forest-cover change maps over the period 2000-2014 for Madagascar.
-Similarly to the approach of @Harper2007, we assess then in details the trends in deforestation rates and forest fragmentation from c.
+We use the forest-cover map provided by @Harper2007 for the year 2000 (defining the land-use) with the tree cover loss product provided by @Hansen2013 that we apply only inside forest areas identified by @Harper2007.
+Similar to the approach of @Harper2007, we also assess trends in deforestation rates and forest fragmentation from c.
 1953 to 2014.
-The approach described in this study should help assessing the efficiency of present conservation strategies and implementing new strategies for the future.
-Our approach could be easily extended to other tropical countries when at least a forest cover map exist between 2000 and 2014 or repeated in the future as and when the @Hansen2013 products are updated.
+The approach described in this study can help assess the effectiveness of current conservation strategies, and assist the implementation of future strategies.
+Our approach could be easily extended to other tropical countries that have at least one forest-cover map between 2000 and 2014.
+This approach can easily be repeated in the future when the @Hansen2013 products are updated.
 
@@
 
-We produced annual forest/non-forest maps at 30m resolution for the full territory of Madagascar for the period 2000-2014 by combining the forest map of year 2000 from @Harper2007 and the tree cover percentage and annual forest cover loss maps over the period 2000-2014 from @Hansen2013.
+We produced annual forest/non-forest maps at 30m resolution for the full territory of Madagascar for the period 2000-2014 by combining the forest map of year 2000 from @Harper2007, and the tree cover percentage and annual forest cover loss maps over the period 2000-2014 from @Hansen2013.
 The 2000 Harper's forest map includes 208,000 ha of unclassified areas due to the presence of clouds in input satellite imagery, mostly (88%) within the moist forest domain which covers 4.17 million ha in total.
@@
 We thus obtained a forest-cover map for the year 2000 covering the full territory of Madagascar.
-We then combined this forest cover map of the year 2000 with the annual forest cover loss maps from 2001 to 2014 provided by @Hansen2013 to create annual forest cover maps from 2001 to 2014 at 30m resolution.
-In a further step we completed the Harper’s forest map of year 1990 by filling unclassified areas using our forest cover map of year 2000: we assume that if forest was present in 2000, the pixel was also forested in 1990.
+We then combined this forest-cover map of the year 2000 with the annual forest cover loss maps from 2001 to 2014 provided by @Hansen2013 to create annual forest-cover maps from 2001 to 2014 at 30m resolution.
+We also completed the Harper's forest map of year 1990 by filling unclassified areas (due to the presence of clouds on satellite images) using our forest-cover map of year 2000.
+To do so, we assumed that if forest was present in 2000, the pixel was also forested in 1990.
 The remaining unclassified pixels were limited to a relatively small total area of c.
@@
 We labeled these residual pixels as non-forest as for the year 2000.
-Similarly we completed the Harper forest map of year 1973 by filling unclassified areas using our forest cover map of the year 1990 assuming that if forest was present in 1990, it was also present in 1973.
-Contrary to the year 1990, the remaining unclassified pixels for year 1973 corresponded to a significant total area of 3.32 million ha, which could not be labeled.
-The forest-cover map of year 1953 was only re-projected to a common projection.
-This map was produced by scanning a paper map derived from aerial photos, and thus could not be perfectly aligned with the other maps produced through digital processing of satellite imagery.
+Similarly we completed the Harper forest map of year 1973 by filling unclassified areas using our forest-cover map of the year 1990 assuming that if forest was present in 1990, it was also present in 1973.
+Contrary to the year 1990, the remaining unclassified pixels for year 1973 corresponded to a significant total area of 3.32 million ha.
+We also reprojected the forest-cover map of year 1953 to a common projection in order to compare to the other eras.
+This map was produced by scanning a paper map derived from aerial photos, and thus could not be perfectly aligned with the other maps produced through digital processing of satellite imagery [@Harper2007].
 Finally for all forest-cover maps from 1973, the isolated single non-forest pixels (i.e.
 fully surrounded by forest pixels) were removed, assuming that single non-forest pixels inside a forest patch were not corresponding to deforestation (they might correspond to selective logging activities).
-This allowed avoiding to include very small scale events (<0.1 ha such as selective logging) as forest fragmentation.
+This allowed us to avoid counting very small scale events (<0.1 ha such as selective logging) as forest fragmentation.
 All the resulting maps are freely available at <https://bioscenemada.cirad.fr/forestmaps>.
@@
 
-From these new forest-cover maps  we computed the total forest-cover area for seven available years (1953-1973-1990-2000-2010-2005-2014), and the annual deforested area and annual deforestation rate for the corresponding six time periods between 1953 and 2014.
-The annual deforestation rates were computed as follow [@Puyravaud2003; @Vieilledent2013]:
+From these new forest-cover maps, we calculated the total forest-cover area for seven available years (1953-1973-1990-2000-2010-2005-2014), and the annual deforested area and annual deforestation rate for the corresponding six time periods between 1953 and 2014.
+The annual deforestation rates were calculated as follows [@Puyravaud2003; @Vieilledent2013]:
 
@@
 MEFT, USAID, and CI (2009) provides estimates for the periods 1990-2000-2005 and ONE, DGF, MNP, WCS, and Etc Terra (2015) provides estimates for the periods 2005-2010-2013.
-To compare our forest-cover and deforestation estimates over the same time periods, we consider an additional time-period in our study (2010-2013) by creating an extra forest cover map for the year 2013.
+To compare our forest-cover and deforestation estimates over the same time periods, we consider an additional time-period in our study (2010-2013) by creating an extra forest-cover map for the year 2013.
 We computed the Pearson's correlation coefficient and the root mean square error (RMSE) between our forest-cover estimates and forest-cover estimates from previous studies for all the dates and forest types (including also the total forest cover estimates).
 For previous studies, the computation of annual deforestation rates (in %/yr) is not always detailed and might slightly differ from one study to another [see @Puyravaud2003].
-@Harper2007 provides total deforested areas for the two periods 1973-1990 and 1990-2000.
+@Harper2007 also provide total deforested areas for the two periods 1973-1990 and 1990-2000.
 We converted these values into annual deforested area estimates.
@@
 These estimates cannot be corrected from the potential bias due to the presence of residual clouds.
-Forest-cover and deforestation rates were then compared between all studies for the whole Madagascar and the four ecoregions.
+Forest-cover and deforestation rates were then compared between all studies for the whole of Madagascar and the four ecoregions.
 The same ecoregion boundaries as in our study were used in @ONE2015 but this was not the case for @Harper2007 and @MEFT2009, which can explain part of the differences between the estimates.
@@
 
-An analysis of the changes in forest fragmentation was carried out from our forest-cover maps over Madagascar for the years 1953, 1973, 1990, 2000, 2010 and 2014.
+We also conducted an analysis of changes in forest fragmentation for the years 1953, 1973, 1990, 2000, 2010 and 2014.
 We applied the method developed by @Riitters2000 which uses a moving window to characterize the fragmentation around each forested pixel.
-Computation were done using the function `r.forestfrag` of the GRASS GIS software [@Neteler2008].
+Computations were done using the function `r.forestfrag` of the GRASS GIS software [@Neteler2008].
 Six categories of fragmentation were identified from the amount of forest and its occurrence as adjacent forest pixels: "interior", "perforated", "edge", "transitional", "patch", and "undetermined".
@@
 Natural forests in Madagascar covered 16.0 Mha in 1953, about 27% of the national territory of 587,041 km2.
-In 2014, the forest cover has dropped down to 8.9 Mha, corresponding to about 15% of the national territory (Fig.
+In 2014, the forest cover dropped to 8.9 Mha, corresponding to about 15% of the national territory (Fig.
 2 and Tab.
@@
 The deforestation trend characterized by a progressive decrease of the deforestation rate over the period 1990-2005 and a progressive increase of the deforestation after 2005 is valid for all four ecoregions (Tab.
-3), excepting the spiny forest domain for which the deforestation rate during the period 2010-2013 was lower than during 2005-2010 (Tab.
+3), with the exception of the spiny forest domain for which the deforestation rate during the period 2010-2013 was lower than during 2005-2010 (Tab.
 3).
@@
 1).
-This allows to produce more robust estimates of forest-cover and deforestation rates over this period.
-Quite logically, our forest-cover estimates over the period 1953-2013 (considering forest cover estimates at national level and by ecoregions for the 6 available dates) were well correlated  (0.99) to estimates from the three previous studies (Tab.
+This allowed us to produce more robust estimates of forest-cover and deforestation rates over this period.
+Our forest-cover estimates over the period 1953-2013 (considering forest cover estimates at national level and by ecoregions for the 6 available dates) were well correlated  (0.99) to estimates from the three previous studies (Tab.
 2) with a RMSE of 300,000 ha (6% of the mean forest cover of 4.8 Mha when considering all dates and forest types together).
@@
 Despite significant differences in deforestation estimates (Tab.
-3), a similar deforestation trend was observed across studies with a decrease of deforestation rates over the period 1990-2005 followed by a progressive increase of the deforestation after 2005.
+3), a similar deforestation trend was observed across studies with a decrease of deforestation rates over the period 1990-2005, followed by a progressive increase of the deforestation after 2005.
 
@@
 In parallel to the dynamics of deforestation, forest fragmentation has progressively increased since 1953 in Madagascar.
-We observed a continuous decrease of the mean distance to forest edges from 1953 to 2014 in Madagascar.
+We observed a continuous decrease of the mean distance to forest edge from 1953 to 2014 in Madagascar.
 The mean distance to forest edge has decreased to c.
-300m in 2014 while it was previously c.
-1.5km in 1973 (Fig.
+300 m in 2014 while it was previously c.
+1.5 km in 1973 (Fig.
 3).
-Moreover, a large proportion (73%) of the forest was located at a distance greater than 100m in 1973 while in 2014, almost half of the forest (46%) is at a distance lower than 100m from forest edge (Fig.
+Moreover, a large proportion (73%) of the forest was located at a distance greater than 100 m in 1973, while almost half of the forest (46%) was at a distance lower than 100 m from forest edge in 2014 (Fig.
 3).
-The percentage of forest that can be considered intact in Madagascar has continuously decreased since 1953: the percentage of forest belonging to the "interior" category (most intact forests) has fallen from 68% in 1973 to 50% in 2014.
+The percentage of forest that can be considered intact in Madagascar has continuously decreased since 1953.
+The percentage of forest belonging to the "interior" category (most intact forests) has fallen from 68% in 1973 to 50% in 2014.
 In 2014, more than 16% of the forest belonged to the "patch" and "transitional" categories (isolated small forest patches) compared to 9.5% in 1973 (Tab.
@@
 Our study extends the forest-cover monitoring on a six decades period (from 1953 to 2014) while harmonizing the data from the previous studies.
-This approach tackles the issue associated with the forest definition which is needed to transform the 2000 global tree cover dataset from Hansen et al.
+We propose a generic approach to solve the problem of forest definition which is needed to transform the 2000 global tree cover dataset from Hansen et al.
 (2013) into a forest/non-forest map [@Tropek2014].
-We demonstrate that an historical national forest-cover map, based on an national forest definition can be used as a mask.
-This approach could be easily extended to other regions or countries for which an accurate forest-cover map is available at any date within the period 2000-2014, but preferably at the beginning of the period to take profit of the full period to derive estimates of deforestation.
-Moreover, this approach can be repeated in the future when the global tree cover product will be updated.
-The R/GRASS code used for this study is made freely available in a GitHub repository (see Data availability statement) and can be applied to other study areas.
+We propose to use a historical national forest-cover map, based on a national forest definition, as a forest cover mask.
+This approach could be easily extended to other regions or countries for which an accurate forest-cover map is available at any date within the period 2000-2014, but preferably at the beginning of the period to profit from the full record and derive long-term estimates of deforestation.
+Moreover, this approach can be repeated in the future if and when the global tree cover product is updated.
+We have made the R/GRASS code used for this study freely available in a GitHub repository (see Data availability statement) to facilitate application to other study areas.
 
@@
 @Verhegghen2016 have compared deforestation estimates derived from the global tree cover loss dataset [@Hansen2013] with results derived from semi-automated supervised classification of Landsat satellite images [@Achard2014] for six countries in Central Africa and they found a good agreement between these two sets of estimates.
-Consistently with @Harper2007, we did not consider potential forest regrowths in Madagascar (although @Hansen2013 provided a tree cover gains layer for the period 2001-2013) for the following reasons: (i) the tree gain layer of @Hansen2013 includes and catches more easily tree plantations than natural forest regrowth [@Tropek2014]; (ii) there is little evidence of natural forest regeneration in Madagascar [@Harper2007, @Grouzis2001] - this can be explained by several ecological processes following burning practice such as soil erosion [@Grinand2017] and reduced seed bank due to fire and soil loss [@Grouzis2001], (iii) in areas where forest regeneration is ecologically possible, young forest regrowth are more easily re-burnt for agriculture and pasture; and (iv) young secondary forests provide more limited ecosystem services compared to old-growth natural forests in terms of biodiversity and carbon storage.
+Consistent with @Harper2007, we did not consider potential forest regrowth in Madagascar (although @Hansen2013 provided a tree cover gains layer for the period 2001-2013) for the following reasons: (i) the tree gain layer of @Hansen2013 includes and catches more easily tree plantations than natural forest regrowth [@Tropek2014]; (ii) there is little evidence of natural forest regeneration in Madagascar [@Harper2007, @Grouzis2001] - this can be explained by several ecological processes following burning practice such as soil erosion [@Grinand2017] and reduced seed bank due to fire and soil loss [@Grouzis2001], (iii) in areas where forest regeneration is ecologically possible, young forest regrowth are more easily re-burnt for agriculture and pasture; and (iv) young secondary forests provide more limited ecosystem services compared to old-growth natural forests in terms of biodiversity and carbon storage.
 
@@
 We estimated that natural forests in Madagascar cover 8.9 Mha in 2014 (corresponding to 15% of the country) and that Madagascar has lost 44% of its natural forest since 1953 (37% since 1973).
-There is an age-old scientific debate about the original forest cover in Madagascar to know to which extent humans have altered the natural forest landscapes since their large-scale settlement around 800 CE [@Burns2016; @Cox2012].
-French naturalists stated that the full island was originally covered by forest [@Humbert1927; @Perrier1921], leading to the common statement that 90% of the natural forests have disappeared since the arrival of humans on the island [@Kull2000].
-More recent studies counter-balanced that point of view saying that grasslands existed in Madagascar long before human arrival and were determined by climate, natural grazing and other natural factors [@Virah-Sawmy2009, @Vorontsova2017].
+There is ongoing scientific debate about the extent of the "original" forest cover in Madagascar, and the extent to which humans have altered the natural forest landscapes since their large-scale settlement around 800 CE [@Burns2016; @Cox2012].
+Early French naturalists stated that the full island was originally covered by forest [@Humbert1927; @Perrier1921], leading to the common statement that 90% of the natural forests have disappeared since the arrival of humans on the island [@Kull2000].
+More recent studies counter-balanced that point of view saying that extensive areas of grassland existed in Madagascar long before human arrival and were determined by climate, natural grazing and other natural factors [@Virah-Sawmy2009, @Vorontsova2017].
 Other authors have questioned the entire narrative of extensive alteration of the landscape by early human activity which, through legislation, has severe consequences on local people [@Klein2002; @Kull2000].
@@
 In our study, we have shown that the progressive decrease of the deforestation rate on the period 1990-2005 was followed by a continuous increase in the deforestation rate on the period 2005-2014.
-We showed in particular that deforestation rate has more than doubled on the period 2010-2014 compared to 2000-2005.
+In particular, we showed that deforestation rate has more than doubled on the period 2010-2014 compared to 2000-2005.
 Our results are confirmed by previous studies [@Harper2007; @MEFT2009; @ONE2015] despite differences in the methodologies regarding (i) forest definition (associated to independent visual interpretations of observation polygons to train the classifier), (ii) classification algorithms, (iii) deforestation rate computation method, and (iv) correction for the presence of clouds.
@@
 
-## 5. Tables
+## Author's contribution
+
+All the authors conceived the study; GV made the analysis and drafted the manuscript. All the authors revised the manuscript.
+
+All authors conceived the ideas and designed methodology; GV analysed the data and wrote the R/GRASS script; GV drafted the manuscript.
+All authors contributed critically to the drafts and gave final approval for publication.
+
+## Acknowledgements
+
+This study is part of the Cirad's BioSceneMada project (https://bioscenemada.cirad.fr) and the Joint Research Center's ReCaREDD project (http://forobs.jrc.ec.europa.eu/recaredd).
+The BioSceneMada project is funded by FRB (Fondation pour la Recherche sur la Biodiversité) and the FFEM (Fond Français pour l'Environnement Mondial) under the project agreement AAP-SCEN-2013 I.
+The ReCaREDD project is funded by the European Commission.
+
+## Data accessibility
+
+All the data and codes used for this study are made publicly available in the `deforestmap` GitHub repository (https://github.com/ghislainv/deforestmap.git).
+The results and the manuscript are fully reproducible running the R script `deforestmap.R` located inside the `deforestmap` repository.
+
+## Tables
 
@@
 
-Table 1: **Evolution of the forest cover and deforestation rates from 1950s to 2014 in Madagascar**.
+Table 1: **Evolution of the forest cover and deforestation rates from 1953 to 2014 in Madagascar**.
 Forest map for the year 1973 has 3.3 Mha of unclassified areas due to the presence of clouds on satellite images.
@@
 
-Table 4: **Evolution of the forest fragmentation from 1950s to 2014 in Madagascar**.
+Table 4: **Evolution of the forest fragmentation from 1953 to 2014 in Madagascar**.
 Six categories of fragmentation were identified from the amount of forest and its occurrence as adjacent forest pixels: "interior", "perforated", "edge", "transitional", "patch", and "undetermined" [@Riitters2000].
@@
 
-## 6. Figures
+## Figures
 
@@
 
-![Figure 2: **Forest-cover change on six decades from _c._ 1953 to 2014 in Madagascar.** Forest cover changes from _c._ 1973 to 2014 are shown in the main figure, and forest cover in _c._ 1953 is shown in the bottom-right inset.
+![Figure 2: **Forest-cover change on six decades from 1953 to 2014 in Madagascar.** Forest cover changes from _c._ 1973 to 2014 are shown in the main figure, and forest cover in _c._ 1953 is shown in the bottom-right inset.
 Two zooms in the western dry (left part) and eastern moist (right part) ecoregions present more detailed views of (from top to bottom): forest-cover in 1950s, forest-cover change from _c._ 1973 to 2014, forest fragmentation in 2014 and distance to forest edge in 2014.
@@
 
-## Acknowledgements
-
-This study is part of the Cirad's BioSceneMada project (https://bioscenemada.cirad.fr) and the Joint Research Center's ReCaREDD project (http://forobs.jrc.ec.europa.eu/recaredd).
-The BioSceneMada project is funded by FRB (Fondation pour la Recherche sur la Biodiversité) and the FFEM (Fond Français pour l'Environnement Mondial) under the project agreement AAP-SCEN-2013 I.
-The ReCaREDD project is funded by the European Commission.
-
-## Data availability statement
-
-All the data and codes used for this study are made publicly available in the `deforestmap` GitHub repository (https://github.com/ghislainv/deforestmap.git).
-The results and the manuscript are fully reproducible running the R script `deforestmap.R` located inside the `deforestmap` repository.
-
 ## References

### 2017-06-06 deforestmap.R

@@ -924,10 +924,9 @@ options(knitr.kable.NA = "-")
 ## Knit and translate to html and pdf
 dir.create("report")
 ## Main document
-render("deforestmap.Rmd",output_dir="report") # html output
+render("deforestmap.Rmd",output_format=c("pdf_document","html_document"),output_dir="report") # html output
 ## Cover letter
 render("coverletter.md",output_format=c("pdf_document"),output_dir="report") # pdf output
-#render("deforestmap.Rmd",output_format=c("pdf_document","word_document"),output_dir="report")
 
 ##===========================================================================
 ## End of script

### 2017-06-07 deforestmap.Rmd
Latex

@@
 output: 
-  html_document:
-    fig_caption: yes
   pdf_document:
+    citation_package: natbib
     fig_caption: true
     keep_tex: true
-  word_document:
-    fig_caption: yes
 ---

### 2017-06-07 latex/deforestmap.tex
Latex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Preambule %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass[a4paper, 12pt, leqno]{article} %leqno: numéro d'équation à gauche
\pagenumbering{arabic} % choose how to number the pages
\usepackage{a4wide}
\usepackage[utf8]{inputenc} % accents interprétés
\usepackage{graphicx}
%\usepackage{subfig}
%\usepackage[hmargin=2cm, vmargin = 2cm, noheadfoot]{geometry} %% Pour gérer le format des pages
%\usepackage{layout} %% Pour avoir la longueur des marges
\usepackage[round,sort]{natbib} %% Natbib is a popular style for formatting references.
%\usepackage{multibib}
%\newcites{secnm}{Bibliographie} 
%\usepackage{verbatim} % for multiline comments
\usepackage{amssymb} %symbole de maths
\usepackage{amsmath} %idem
%\usepackage{stmaryrd} %% Symbole flèche à l'envers
%\usepackage{amsfonts}
\usepackage[english]{babel} %% Les titres en anglais
\usepackage{array} %% Pour centrer verticalement le contenu d'un tableau, entre autres...
\setcounter{secnumdepth}{4} %% Profondeur des sections, subsections
\usepackage{setspace} %% Gère l'interligne: singlespacing, doublespacing
\usepackage{booktabs}
%\singlespacing
\usepackage{longtable}
\setlength{\LTleft}{-5cm plus 1 fill}
\setlength{\LTright}{-5cm plus 1 fill}
\usepackage[colorlinks=true,citecolor=blue]{hyperref} %% Gère les hyperliens
\usepackage{lineno} %% Numérotation des lignes
\usepackage{caption2}
%\linenumbers
\newcommand{\logit}{\text{logit}}
\newcommand{\bs}[1]{\boldsymbol{#1}}
\newcommand{\p}{\text{p}}
\newcommand{\R}{\textnormal{\sffamily\bfseries R}}
\newcommand{\pkg}[1]{{\fontseries{b}\selectfont #1}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Title %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\title{Combining global tree cover loss data with historical national
  forest-cover maps to look at six decades of deforestation and forest
  fragmentation in Madagascar}
\date{}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Document %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}
\maketitle

\vspace{-1cm}
\begin{center}
%%{\large
  Ghislain Vieilledent$^{1,2,\star}$\hspace{1cm}
  Clovis Grinand$^{3}$ \\
  \vspace{0.25cm}
  Fety A. Rakotomalala$^{3}$ \hspace{1cm}
  Rija Ranaivosoa$^{4}$ \hspace{1cm}
  Jean-Roger Rakotoarijaona$^{4}$ \\
  \vspace{0.25cm}
  Thomas F. Allnutt$^{5,6}$\hspace{1cm}and\hspace{1cm}
  Frédéric Achard$^{1}$
%%}
\end{center}

\vspace{0.3cm}

{\small
  \begin{flushleft}  
    $[1]$ \textbf{Joint Research Center of the European Commission} -- Bio-economy Unit, I-21027 Ispra (VA), ITALY\\
    $[2]$ \textbf{Cirad} -- UPR Forêts et Sociétés, F-34398 Montpellier, FRANCE\\
    $[3]$ \textbf{ETC Terra}, F-75020 Paris, FRANCE\\
    $[4]$ \textbf{Office National pour l'Environnement}, 101 Antananarivo, MADAGASCAR\\
    $[5]$ \textbf{Wildlife Conservation Society}, 101 Antananarivo, MADAGASCAR\\
    $[6]$ \textbf{GreenInfo Network}, Oakland, California, USA\\
    ~\\
    $[\star]$ \textbf{Corresponding author:}
    \textbackslash{E-mail}:~ghislain.vieilledent@cirad.fr
    \textbackslash{Phone}:~+39.033.278.3516\\
  \end{flushleft}}

\vspace{0.3cm}

\begin{center}
  \textbf{Running headline:} \emph{Six decades of deforestation in Madagascar}
\end{center}

\newpage

\linenumbers
\doublespacing

\newcommand{\keywords}[1]{\par\noindent
{\small{\em Keywords\/}: #1}}

\begin{abstract}

  1. The island of Madagascar has an unparalleled biodiversity, mainly
  located in the tropical forests of the island, which is highly
  threatened by anthropogenic deforestation. Scattered forest maps
  from past studies at national level with substantial gaps (due to
  presence of cloud cover on satellite imagery) prevent the analyzis
  of long-term deforestation trends in Madagascar.

  2. In this study, we propose a new approach combining historical
  (1953-2000) national forest-cover maps with recent (2000-2014)
  global annual tree cover loss data to look at six decades
  (1953-2014) of deforestation and forest fragmentation in
  Madagascar. We produced new forest-cover maps at 30m-resolution over
  the full territory of Madagascar for the year 1990, and annually
  from 2000 to 2014.

  3. We estimated that Madagascar has lost 44\% of its natural forest
  cover over the period 1953-2014 (including 37\% over the period
  1973-2014).  Natural forests cover 8.9 Mha in 2014 (15\% of the
  national territory) which are divided into 4.4 Mha (50\%) of moist
  forests, 2.6 Mha (29\%) of dry forests, 1.7 Mha of spiny forests
  (19\%) and 177,000 ha (2\%) of mangroves. Since 2005, the annual
  deforestation rate has progressively increased in Madagascar to
  reach 99,000 ha/yr during 2010-2014 (corresponding to a rate of
  1.08\%/yr). Around half of the forest (46\%) is now located at less
  than 100m from the forest edge.

  4. \emph{Policy implications}: Increase in the deforestation rate
  since 2005 can be related to rapid population growth (close to
  3\%/yr) and to poor law enforcement due to political instability in
  the country.  More effort, including conservation and development
  programs, is needed in Madagascar to protect the remaining natural
  tropical forests, both to enhance local people's livelihoods and to
  support global environmental efforts.

\vspace{0.5cm}

\keywords{biodiversity, climate-change, deforestation, Landsat,
  Madagascar, tropical forest}

\end{abstract}

\newpage

\section{Introduction}
\label{introduction}

Separated from the African continent and the Indian plate about 165
and 88 million years ago respectively \citep{Ali2008}, the flora and
fauna of Madagascar followed its own evolutionary path. Isolation
combined with a high number of micro-habitats \citep{Pearson2009} has
led to Madagascar's exceptional biodiversity both in term of number of
species and endemism in many taxonomic groups \citep{Crottini2012,
  Goodman2005}.  Most of the biodiversity in Madagascar is
concentrated in the tropical forests of the island which can be
divided into three types: the moist forest in the East, the dry forest
in the West and the spiny forest in the South
\citep{Vieilledent2016}. This unparalleled biodiversity is severely
threatened by anthropogenic deforestation \citep{Harper2007,
  Vieilledent2013} associated with activities such as slash-and-burn
agriculture and pasture \citep{Scales2011}. Tropical forests in
Madagascar also store a large amount of carbon \citep{Vieilledent2016}
and high rates of deforestation in Madagascar are responsible for
large CO2 emissions in the atmosphere
\citep{Achard2014}. Deforestation threatens species survival by
directly reducing their available habitat \citep{Brooks2002,
  Tidd2001}. Forest fragmentation can also lead to species extinction
by isolating populations from each other and creating forest patches
too small to maintain viable populations
\citep{Saunders1991}. Fragmentation also increases forest edge where
ecological conditions (such as air temperature, light intensity and
air moisture) can be dramatically modified, with consequences on the
abundance and distribution of species \citep{Murcia1995}. Forest
fragmentation can also have substantial effects on forest carbon
storage capacity, as carbon stocks are much lower at the forest edge
than under a closed canopy \citep{Brinck2017}.  Moreover, forest
carbon stocks vary spatially due to climate or soil factors
\citep{Saatchi2011, Vieilledent2016}. As a consequence, accurate and
spatially explicit maps of forest-cover and forest-cover change are
necessary to monitor biodiversity loss and carbon emissions from
deforestation and forest fragmentation, assess the efficiency of
present conservation strategies \citep{Eklund2016}, and implement new
strategies for the future \citep{Vieilledent2013,
  Vieilledent2016}. Simple time-series of forest-cover estimates, such
as those provided by the FAO Forest Resource Assessment report
\citep{Keenan2015} are not sufficient.

Unfortunately, accurate and exhaustive forest-cover maps are not
available for Madagascar for the last fifteen years (2000-2015).
\citet{Harper2007} produced maps of forest cover and forest cover
changes over Madagascar for the years c. 1953, c. 1973, 1990 and 2000.
The c. 1953 forest map was derived from the visual interpretation of
aerial photography at coarse scale (1/1,000,000). Forest maps for the
years c. 1973, 1990 and 2000, were obtained from supervised
classification of Landsat satellite images at 60m resolution (for the
year 1973) or 30m resolution (for years 1990 and 2000) and can be used
to derive accurate estimates (89.5\% accuracy reported for the forest
/ non-forest map of year 2000). Maps provided by \citet{Harper2007}
were not exhaustive (due to the presence of clouds in the satellite
imagery), e.g.~11 244 km2 are mapped as unknown cover type for the
year 2000.  Using a similar supervised classification approach as in
\citet{Harper2007}, more recent maps have been produced for the
periods 2000-2005-2010 by national institutions, with the technical
support of international environmental NGOs \citep{MEFT2009,
  ONE2013}. Another set of recent forest-cover maps using an advanced
statistical tool for classification, the Random Forest classifier
\citep{Grinand2013, Rakotomalala2015}, was produced for the periods
2005-2010-2013 \citep{ONE2015}. However, these maps are either too old
to give recent estimates of deforestation \citep{MEFT2009, ONE2013},
include large areas of missing information (due to images with more
than 20\% of clouds) for the 2010 map provided by \citep{ONE2013}, or
show large mis-classification in specific areas, especially in the dry
and spiny forest domain for which the spectral answer has a strong
seasonal behavior due to the deciduousness of such forests (overall
accuracy is lower than 0.8 for the dry and spiny forests for the maps
produced by \citet{ONE2015}). Moreover, the production of such forest
maps from a supervised classification approach requires significant
resources, especially regarding the image selection step (required to
minimize cloud cover) and the training step (visual interpretation of
a large number of polygons needed to train the classification
algorithm) \citep{Rakotomalala2015}. Most of this work of image
selection and visual interpretation would need to be repeated to
produce new forest maps in the future using a similar approach.

Global forest or tree cover products have also been published recently
and can be tested at the national scale for
Madagascar. \citep{Kim2014} produced a global forest-cover change map
from 1990 to 2000 (derived from Landsat imagery). This product was
updated to cover the period 1975-2005
(\url{http://glcf.umd.edu/data/landsatFCC/}) but forest-cover maps
after 2005 were not produced. Moreover, the approach used in
\citet{Kim2014} did not accurately map the forests in the dry and
spiny ecosystems of Madagascar (see Fig. 8 in \citet{Kim2014}).
\citet{Hansen2013} mapped tree cover percentage, annual forest loss
and forest gain from 2000 to 2012 at global scale at 30 m
resolution. This product has since been updated and is now available
up to the year 2014 \citep{Hansen2013}. To map forest cover from the
\citet{Hansen2013} product, a tree cover threshold must be selected
(that defines forest cover). Selecting such a threshold is not
straightforward as the accuracy of the global tree cover map strongly
varies between forest types, and is substantially lower for dry
forests than for moist forests \citep{Bastin2017}. Moreover, the
\citet{Hansen2013} product does not provide information on
land-use. In particular the global tree cover map does not separate
tree plantations such as oil palm or eucalyptus plantations from
natural forests \citep{Tropek2014}. Thus, the global tree cover map
from \citet{Hansen2013} cannot be used alone to produce a map of
forest cover \citep{Tyukavina2017}. In complement to the tree cover
percentage provided in \citet{Hansen2013}, a layer of annual tree
cover loss is also provided (i.e.~complete loss of tree cover from a
value higher than 10\% to zero) for the period 2001-2014.

In this study, we present a simple approach which combines the maps
from \citet{Harper2007} and products from \citet{Hansen2013} to derive
annual wall-to-wall forest-cover change maps over the period 2000-2014
for Madagascar. We use the forest-cover map provided by
\citet{Harper2007} for the year 2000 (defining the land-use) with the
tree cover loss product provided by \citet{Hansen2013} that we apply
only inside forest areas identified by \citet{Harper2007}. Similar to
the approach of \citet{Harper2007}, we also assess trends in
deforestation rates and forest fragmentation from c. 1953 to 2014. The
approach described in this study can help assess the effectiveness of
current conservation strategies, and assist the implementation of
future strategies. Our approach could be easily extended to other
tropical countries that have at least one forest-cover map between
2000 and 2014. This approach can easily be repeated in the future when
the \citet{Hansen2013} products are updated.

\section{Materials and Methods}
\label{materials-and-methods}

\subsection{Creation of new forest-cover maps of Madagascar from
1953 to 2014}

We produced annual forest/non-forest maps at 30m resolution for the
full territory of Madagascar for the period 2000-2014 by combining the
forest map of year 2000 from \citet{Harper2007}, and the tree cover
percentage and annual forest cover loss maps over the period 2000-2014
from \citet{Hansen2013}. The 2000 Harper's forest map includes 208,000
ha of unclassified areas due to the presence of clouds in input
satellite imagery, mostly (88\%) within the moist forest domain which
covers 4.17 million ha in total. To provide a label (forest or
non-forest) to these unclassified pixels, we use the 2000 tree cover
percentage map of \citet{Hansen2013} by selecting a threshold of 75\%
tree cover to define forest cover as recommended by other studies for
the moist domain \citep{Achard2014}. We thus obtained a forest-cover
map for the year 2000 covering the full territory of Madagascar. We
then combined this forest-cover map of the year 2000 with the annual
forest cover loss maps from 2001 to 2014 provided by
\citet{Hansen2013} to create annual forest-cover maps from 2001 to
2014 at 30m resolution. We also completed the Harper's forest map of
year 1990 by filling unclassified areas (due to the presence of clouds
on satellite images) using our forest-cover map of year 2000. To do
so, we assumed that if forest was present in 2000, the pixel was also
forested in 1990. The remaining unclassified pixels were limited to a
relatively small total area of c. 8,000 ha. We labeled these residual
pixels as non-forest as for the year 2000.  Similarly we completed the
Harper forest map of year 1973 by filling unclassified areas using our
forest-cover map of the year 1990 assuming that if forest was present
in 1990, it was also present in 1973.  Contrary to the year 1990, the
remaining unclassified pixels for year 1973 corresponded to a
significant total area of 3.32 million ha. We also reprojected the
forest-cover map of year 1953 to a common projection in order to
compare to the other eras. This map was produced by scanning a paper
map derived from aerial photos, and thus could not be perfectly
aligned with the other maps produced through digital processing of
satellite imagery \citep{Harper2007}. Finally for all forest-cover
maps from 1973, the isolated single non-forest pixels (i.e.~fully
surrounded by forest pixels) were removed, assuming that single
non-forest pixels inside a forest patch were not corresponding to
deforestation (they might correspond to selective logging activities).
This allowed us to avoid counting very small scale events
(\textless{}0.1 ha such as selective logging) as forest fragmentation.
All the resulting maps are freely available at
\url{https://bioscenemada.cirad.fr/forestmaps}.

\subsection{Computing forest-cover areas and deforestation
rates}

From these new forest-cover maps, we calculated the total forest-cover
area for seven available years (1953-1973-1990-2000-2010-2005-2014),
and the annual deforested area and annual deforestation rate for the
corresponding six time periods between 1953 and 2014. The annual
deforestation rates were calculated as follows \citep{Puyravaud2003,
  Vieilledent2013}:

\begin{equation}
  \theta = 100 \times [1-(1-(F_{t_2}-F_{t_1})/F_{t_1})^{(1/(t_2-t_1))}
\end{equation}

where $\theta$ is the annual deforestation rate (in \%/yr),
$F_{t_2}$ and $F_{t_1}$ are the forest cover free of clouds at both
dates $t_2$ and $t_1$, and $t_2-t_1$ is the time-interval (in
years) between the two dates.

Because of the large unclassified area (3.32 million ha) in 1973, the
annual deforestation areas and rates for the two periods 1953-1973 and
1973-1990 are only indicative estimates. For these two periods the
annual deforestation rates are computed as the ratio
$(F_{t_2}-F_{t_1})/F_{t_1}$ considering only the mapped forest pixels.
Area and rate estimates are produced at the national scale and for the
four forest ecosystems present in Madagascar: moist forest in the
East, dry forest in the West, spiny forest in the South, and mangroves
on the Western coast (Fig.~\ref{fig:ecoregion}). To define the forest
domains, we used a map from the MEFT (\emph{``Ministère de
  l'Environnement et des Forêts à Madagascar''}) with the boundaries
of the four ecoregions in Madagascar.  Ecoregions were defined on the
basis of climatic and vegetation criteria using the climate
classification by \citet{Cornet1974} and the vegetation classification
from the 1996 IEFN national forest inventory \citep{IEFN1996}. Because
mangrove forests are highly dynamic ecosystems that can expand or
contract on decadal scales depending on changes in environmental
factors \citep{Armitage2015}, a fixed delimitation of the mangrove
ecoregion on six decades might not be fully appropriate. As a
consequence, our estimates of the forest-cover and deforestation rates
for mangroves in Madagascar must be considered with this limitation.

\subsection{Comparing our forest-cover and deforestation rate
estimates with previous studies}

We compared our estimates of forest-cover and deforestation rates with
estimates from the three existing studies at the national scale for
Madagascar: (i) \citep{Harper2007}, (ii) \citep{MEFT2009} and (iii)
\citep{ONE2015}. \citet{Harper2007} provides forest-cover and
deforestation estimates for the periods
c. 1953-c. 1973-1990-2000. MEFT, USAID, and CI (2009) provides
estimates for the periods 1990-2000-2005 and ONE, DGF, MNP, WCS, and
Etc Terra (2015) provides estimates for the periods 2005-2010-2013. To
compare our forest-cover and deforestation estimates over the same
time periods, we consider an additional time-period in our study
(2010-2013) by creating an extra forest-cover map for the year
2013. We computed the Pearson's correlation coefficient and the root
mean square error (RMSE) between our forest-cover estimates and
forest-cover estimates from previous studies for all the dates and
forest types (including also the total forest cover estimates). For
previous studies, the computation of annual deforestation rates (in
\%/yr) is not always detailed and might slightly differ from one study
to another \citep[see][]{Puyravaud2003}. \citet{Harper2007} also
provide total deforested areas for the two periods 1973-1990 and
1990-2000. We converted these values into annual deforested area
estimates. When annual deforested areas were not reported (for
1953-1973 in \citet{Harper2007} and in \citet{MEFT2009} and
\citet{ONE2015}), we computed them from the forest-cover estimates in
each study. These estimates cannot be corrected from the potential
bias due to the presence of residual clouds. Forest-cover and
deforestation rates were then compared between all studies for the
whole of Madagascar and the four ecoregions. The same ecoregion
boundaries as in our study were used in \citet{ONE2015} but this was
not the case for \citet{Harper2007} and \citet{MEFT2009}, which can
explain part of the differences between the estimates.

\subsection{Fragmentation}

We also conducted an analysis of changes in forest fragmentation for
the years 1953, 1973, 1990, 2000, 2010 and 2014. We applied the method
developed by \citet{Riitters2000} which uses a moving window to
characterize the fragmentation around each forested
pixel. Computations were done using the function \texttt{r.forestfrag}
of the GRASS GIS software \citep{Neteler2008}. Six categories of
fragmentation were identified from the amount of forest and its
occurrence as adjacent forest pixels: ``interior'', ``perforated'',
``edge'', ``transitional'', ``patch'', and ``undetermined''. We used a
moving window of 7x7 pixels (4.4 ha). Using this window size, forest
edge had a width of about 90m \citep{Riitters2000}. The ``interior''
category can be interpreted as the most intact forest
\citep{Potapov2017}. The ``patch'' and ``transitional'' categories
correspond to isolated small forest patches.  We reported the area of
forest in each fragmentation category for the six years and analyzed
the dynamics of fragmentation over the six decades. We also computed
the distance to forest edge for all forest pixels for the years 1953,
1973, 1990, 2000, 2005, 2010 and 2014. For that, we used the function
\texttt{gdal\_proximity.py} of the GDAL software
(\url{http://www.gdal.org/}). We computed the mean and 95\% quantiles
of the distance to forest edge and looked at the evolution of these
values with time.

\section{Results}
\label{results}

\subsection{Dynamics of forest cover and deforestation
intensity}

Natural forests in Madagascar covered 16.0 Mha in 1953, about 27\% of
the national territory of 587,041 km2. In 2014, the forest cover
dropped to 8.9 Mha, corresponding to about 15\% of the national
territory (Fig.~\ref{fig:fcc} and
Tab.~\ref{tab:forest_cover}). Madagascar has lost 44\% and 37\% of its
natural forests between 1953 and 2013, and between c. 1973 and 2014
respectively (Fig.~\ref{fig:fcc} and Tab.~\ref{tab:forest_cover}). In
2014 the remaining 8.9 Mha of natural forest were distributed as: 4.4
Mha of moist forest (50\% of total forest cover), 2.6 Mha of dry
forest (29\%), 1.7 Mha of spiny forest (19\%) and 0.18 Mha (2\%) of
mangrove forest (Fig.~\ref{fig:ecoregion} and
Tab.~\ref{tab:comp_forest}). Regarding the deforestation trend, we
observed a progressive decrease of the deforestation rate after1990
from 205,000 ha/yr (1.63\%/yr) over the period 1973-1990 to 44,300
ha/yr (0.43\%/yr) over the period 2000-2005
(Tab.~\ref{tab:forest_cover}). Then from 2005, the deforestation rate
has progressively increased and has more than doubled over the period
2010-2014 (98,700 ha/yr, 1.08\%/yr) compared to 2000-2005
(Tab.~\ref{tab:forest_cover}). The deforestation trend characterized
by a progressive decrease of the deforestation rate over the period
1990-2005 and a progressive increase of the deforestation after 2005
is valid for all four ecoregions (Tab.~\ref{tab:comp_defor}), with the
exception of the spiny forest domain for which the deforestation rate
during the period 2010-2013 was lower than during 2005-2010
(Tab.~\ref{tab:comp_defor}).

\subsection{Comparison with previous forest-cover change studies
in Madagascar}

Forest-cover maps provided by previous studies over Madagascar were
not exhaustive (unclassified areas) due to the presence of clouds in
satellite imagery used to produce such maps. In \citet{Harper2007},
the maps of years 1990 and 2000 include 0.5 and 1.12 Mha of unknown
cover type respectively. Proportions of unclassified areas are not
reported in the two other existing studies by \citet{MEFT2009} and
\citet{ONE2015}.  With our approach, we produced wall to wall
forest-cover change maps from 1990 to 2014 for the full territory of
Madagascar (Tab.~\ref{tab:forest_cover}). This allowed us to produce
more robust estimates of forest-cover and deforestation rates over
this period. Our forest-cover estimates over the period 1953-2013
(considering forest cover estimates at national level and by
ecoregions for the 6 available dates) were well correlated (0.99) to
estimates from the three previous studies (Tab.~\ref{tab:comp_forest})
with a RMSE of 300,000 ha (6\% of the mean forest cover of 4.8 Mha
when considering all dates and forest types together). These small
differences can be partly attributed to differences in ecoregion
boundaries. Despite significant differences in deforestation estimates
(Tab.~\ref{tab:comp_defor}), a similar deforestation trend was
observed across studies with a decrease of deforestation rates over
the period 1990-2005, followed by a progressive increase of the
deforestation after 2005.

\subsection{Evolution of forest fragmentation with time}

In parallel to the dynamics of deforestation, forest fragmentation has
progressively increased since 1953 in Madagascar. We observed a
continuous decrease of the mean distance to forest edge from 1953 to
2014 in Madagascar. The mean distance to forest edge has decreased to
c.  300 m in 2014 while it was previously c. 1.5 km in 1973
(Fig.~\ref{fig:dist_edge}).  Moreover, a large proportion (73\%) of
the forest was located at a distance greater than 100 m in 1973, while
almost half of the forest (46\%) was at a distance lower than 100 m
from forest edge in 2014 (Fig.  3). The percentage of forest that can
be considered intact in Madagascar has continuously decreased since
1953. The percentage of forest belonging to the ``interior'' category
(most intact forests) has fallen from 68\% in 1973 to 50\% in 2014. In
2014, more than 16\% of the forest belonged to the ``patch'' and
``transitional'' categories (isolated small forest patches) compared
to 9.5\% in 1973 (Tab.~\ref{tab:frag}).

\section{Discussion}
\label{discussion}

\subsection{Benefits of the combined use of recent global annual
  tree cover loss data with historical national forest-cover maps}

In this study, we combined recent (2001-2014) global annual tree cover
loss data \citep{Hansen2013} with historical (1953-2000) national
forest-cover maps \citep{Harper2007} to look at six decades (1953-2014)
of deforestation and forest fragmentation in Madagascar. We produced
annual forest-cover maps at 30m-resolution covering Madagascar for the
period 2000 to 2014. Our study extends the forest-cover monitoring on a
six decades period (from 1953 to 2014) while harmonizing the data from
the previous studies. We propose a generic approach to solve the problem
of forest definition which is needed to transform the 2000 global tree
cover dataset from Hansen et al. (2013) into a forest/non-forest map
\citep{Tropek2014}. We propose to use a historical national forest-cover
map, based on a national forest definition, as a forest cover mask. This
approach could be easily extended to other regions or countries for
which an accurate forest-cover map is available at any date within the
period 2000-2014, but preferably at the beginning of the period to
profit from the full record and derive long-term estimates of
deforestation. Moreover, this approach can be repeated in the future if
and when the global tree cover product is updated. We have made the
R/GRASS code used for this study freely available in a GitHub repository
(see Data availability statement) to facilitate application to other
study areas.

The accuracy of the derived forest-cover change maps depends directly on
the accuracies of the historical forest-cover maps and the tree cover
loss dataset. The reported global accuracy of the tree cover loss
dataset is 99.6\% (see Tab. S5 in \citet{Hansen2013}).
\citet{Verhegghen2016} have compared deforestation estimates derived
from the global tree cover loss dataset \citep{Hansen2013} with results
derived from semi-automated supervised classification of Landsat
satellite images \citep{Achard2014} for six countries in Central Africa
and they found a good agreement between these two sets of estimates.
Consistent with \citet{Harper2007}, we did not consider potential forest
regrowth in Madagascar (although \citet{Hansen2013} provided a tree
cover gains layer for the period 2001-2013) for the following reasons:
(i) the tree gain layer of \citet{Hansen2013} includes and catches more
easily tree plantations than natural forest regrowth \citep{Tropek2014};
(ii) there is little evidence of natural forest regeneration in
Madagascar \citep[\citet{Grouzis2001}]{Harper2007} - this can be
explained by several ecological processes following burning practice
such as soil erosion \citep{Grinand2017} and reduced seed bank due to
fire and soil loss \citep{Grouzis2001}, (iii) in areas where forest
regeneration is ecologically possible, young forest regrowth are more
easily re-burnt for agriculture and pasture; and (iv) young secondary
forests provide more limited ecosystem services compared to old-growth
natural forests in terms of biodiversity and carbon storage.

\subsection{Dynamics of forest-cover in Madagascar from 1953 to 2014}

We estimated that natural forests in Madagascar cover 8.9 Mha in 2014
(corresponding to 15\% of the country) and that Madagascar has lost 44\%
of its natural forest since 1953 (37\% since 1973). There is ongoing
scientific debate about the extent of the ``original'' forest cover in
Madagascar, and the extent to which humans have altered the natural
forest landscapes since their large-scale settlement around 800 CE
\citep{Burns2016, Cox2012}. Early French naturalists stated that the
full island was originally covered by forest
\citep{Humbert1927, Perrier1921}, leading to the common statement that
90\% of the natural forests have disappeared since the arrival of humans
on the island \citep{Kull2000}. More recent studies counter-balanced
that point of view saying that extensive areas of grassland existed in
Madagascar long before human arrival and were determined by climate,
natural grazing and other natural factors
\citep[\citet{Vorontsova2017}]{Virah-Sawmy2009}. Other authors have
questioned the entire narrative of extensive alteration of the landscape
by early human activity which, through legislation, has severe
consequences on local people \citep{Klein2002, Kull2000}. Whatever the
original proportion of natural forests and grasslands in Madagascar, our
results demonstrate that human activities since the 1950s have
profoundly impacted the natural tropical forests and that recent
conservation and development programs in Madagascar have failed to stop
deforestation in the recent years. Deforestation has strong consequences
on biodiversity and carbon emissions in Madagascar. Around 90\% of
Madagascar's species are forest dependent
\citep{Allnutt2008, Goodman2005} and \citet{Allnutt2008} estimated that
deforestation between 1953 and 2000 led to an extinction of 9\% of the
species. The additional deforestation we observed over the period
2000-2014 (around 1Mha of natural forest) worsen this result. Regarding
carbon emissions, using the 2010 aboveground forest carbon map by
\citet{Vieilledent2016}, we estimated that deforestation on the period
2010-2014 has led to 40.2 Mt C of carbon emissions in the atmosphere (10
Mt C /yr) and that the remaining aboveground forest carbon stock in 2014
is 832.8 Mt C. Associated to deforestation, we showed that the remaining
forests of Madagascar are highly fragmented with 46\% of the forest
being at less than 100m of the forest edge. Small forest fragments do
not allow to maintain viable populations and ``edge effects'' at
forest/non-forest interfaces have impacts on both carbon emissions
\citep{Brinck2017} and biodiversity loss \citep{Gibson2013, Murcia1995}.

\subsection{Deforestation trend and causes of the acceleration of the
  deforestation since 2005}

In our study, we have shown that the progressive decrease of the
deforestation rate on the period 1990-2005 was followed by a continuous
increase in the deforestation rate on the period 2005-2014. In
particular, we showed that deforestation rate has more than doubled on
the period 2010-2014 compared to 2000-2005. Our results are confirmed by
previous studies \citep{Harper2007, MEFT2009, ONE2015} despite
differences in the methodologies regarding (i) forest definition
(associated to independent visual interpretations of observation
polygons to train the classifier), (ii) classification algorithms, (iii)
deforestation rate computation method, and (iv) correction for the
presence of clouds. Our deforestation rate estimates from 1990 to 2014
have been computed from wall to wall maps at 30 m resolution and can be
considered more accurate in comparison with estimates from these
previous studies. Our forest-cover and deforestation rate estimates can
be used as source of information for the next FAO Forest Resources
Assessment project \citep{Keenan2015}. Current rates of deforestation
can also be used to build reference scenarios for deforestation in
Madagascar and contribute to the implementation of deforestation
mitigation activities in the framework of REDD+ \citep{Olander2008}.

The increase of deforestation rates after 2005 can be explained by
population growth and political instability in the country. Nearly 90\%
of Madagascar's population relies on biomass for their daily energy
needs \citep{Minten2013} and the link between population size and
deforestation has previously been demonstrated in Madagascar
\citep{Vieilledent2013, Gorenflo2011}. With a mean demographic growth
rate of about 2.8\%/yr and a population which has increased from 16 to
24 million people on the period 2000-2015 \citep{UN2015}, the increasing
demand in wood-fuel and space for agriculture is likely to explain the
increase in deforestation rates. The political crisis of 2009
\citep{Ploch2012}, followed by several years of political instability
and weak governance could also explain the increase in the deforestation
rate observed on the period 2005-2014 (Smith et al., 2003). These
results show that despite the conservation policy in Madagascar
\citep{Freudenberger2010}, deforestation has dramatically increased at
the national level since 2005. More efforts should be done to identify
the real causes and mechanisms explaining this rampant deforestation in
order to implement new conservation strategies that will be effective at
conserving the few remaining natural tropical forests in Madagascar.

\newpage

\section{Author's contribution}
\label{authors-contribution}

All authors conceived the ideas and designed methodology; GV analysed
the data and wrote the {\R}/GRASS script; GV drafted the manuscript. All
authors contributed critically to the drafts and gave final approval for
publication.

\section{Acknowledgements}
\label{acknowledgements}

This study is part of the Cirad's BioSceneMada project
(\url{https://bioscenemada.cirad.fr}) and the Joint Research Center's
ReCaREDD project (\url{http://forobs.jrc.ec.europa.eu/recaredd}). The
BioSceneMada project is funded by FRB (Fondation pour la Recherche sur
la Biodiversité) and the FFEM (Fond Français pour l'Environnement
Mondial) under the project agreement AAP-SCEN-2013 I. The ReCaREDD
project is funded by the European Commission.

\section{Data accessibility}
\label{data-accessibility}

All the data and codes used for this study are made publicly available
in the \texttt{deforestmap} GitHub repository
(\url{https://github.com/ghislainv/deforestmap.git}). The results and
the manuscript are fully reproducible running the {\R} script
\texttt{deforestmap.R} located inside the \texttt{deforestmap}
repository.

\newpage
\singlespacing

\bibliographystyle{gcb}
\bibliography{bib/biblio}

\newpage

\section{Tables}
\label{tables}

\nopagebreak

\vfill
\begin{table}[!h]
  \begin{longtable}[]{@{}rrrrr@{}}
    \toprule
    Year & Forest (ha) & Unmap (ha) & Annual defor. (ha/yr) & Rate
    (\%/yr)\tabularnewline
    \midrule
    \endhead
    1953 & 15968176 & 0 & - & -\tabularnewline
    1973 & 14242592 & 3316531 & 86279 & 0.57\tabularnewline
    1990 & 10762442 & 0 & 204715 & 1.63\tabularnewline
    2000 & 9879031 & 0 & 88341 & 0.85\tabularnewline
    2005 & 9667553 & 0 & 42296 & 0.43\tabularnewline
    2010 & 9319851 & 0 & 69540 & 0.73\tabularnewline
    2014 & 8925246 & 0 & 98651 & 1.08\tabularnewline
    \bottomrule
  \end{longtable}
  \addtocounter{table}{-1}

  \caption{\textbf{Evolution of the forest cover and deforestation rates
      from 1953 to 2014 in Madagascar}. Forest map for the year 1973 has 3.3
    Mha of unclassified areas due to the presence of clouds on satellite
    images. As a consequence, deforestation rates for the periods 1953-1973
    and 1973-1990 are indicative. The two last columns indicate the
    deforestation rates on the previous time-period (e.g.~1953-1973 for year
    1973, 1973-1990 for year 1990, etc.).}

  \label{tab:forest_cover}
\end{table}
\vfill

\newpage

\vspace*{\stretch{1}}
\begin{table}[!h]
  {\footnotesize
  \begin{longtable}[]{@{}llrrrrrrrr@{}}
    \toprule
    Forest type & Source & 1953 & 1973 & 1990 & 2000 & 2005 & 2010 & 2013 &
    2014\tabularnewline
    \midrule
    \endhead
    Total & Harper2007 & 15995900 & 14173100 & 10605700 & 8982100 & - & - &
    - & -\tabularnewline
    & MEFT2009 & - & - & 10650142 & 9678402 & 9413218 & - & - &
    -\tabularnewline
    & ONE2015 & - & - & - & - & 9451350 & 8977337 & 8485509 &
    -\tabularnewline
    & this study & 15968176 & 14242592 & 10762494 & 9879031 & 9667553 &
    9319851 & 9051029 & 8925246\tabularnewline
    Moist & Harper2007 & 8765600 & 6876000 & 5234300 & 4166800 & - & - & - &
    -\tabularnewline
    & MEFT2009 & - & - & 5270599 & 4787771 & 4700430 & - & - &
    -\tabularnewline
    & ONE2015 & - & - & - & - & 4555788 & 4457184 & 4345093 &
    -\tabularnewline
    & this study & 8578299 & 6989942 & 5270169 & 4872016 & 4767876 & 4633104
    & 4470194 & 4409842\tabularnewline
    Dry & Harper2007 & 4252100 & 4027700 & 2711800 & 2457000 & - & - & - &
    -\tabularnewline
    & MEFT2009 & - & - & 3320582 & 3084976 & 3027505 & - & - &
    -\tabularnewline
    & ONE2015 & - & - & - & - & 3223028 & 2970192 & 2678640 &
    -\tabularnewline
    & this study & 4761551 & 4434871 & 3224917 & 2940970 & 2880819 & 2734639
    & 2642253 & 2595621\tabularnewline
    Spiny & Harper2007 & 2978200 & 3029800 & 2420000 & 2132200 & - & - & - &
    -\tabularnewline
    & MEFT2009 & - & - & 2123630 & 1871735 & 1756884 & - & - &
    -\tabularnewline
    & ONE2015 & - & - & - & - & 1681527 & 1558533 & 1466765 &
    -\tabularnewline
    & this study & 2462830 & 2582880 & 2054724 & 1857628 & 1810704 & 1744427
    & 1731308 & 1712731\tabularnewline
    Mangroves & Harper2007 & - & - & 239600 & 226100 & - & - & - &
    -\tabularnewline
    & MEFT2009 & - & - & - & - & - & - & - & -\tabularnewline
    & ONE2015 & - & - & - & - & 173564 & 171220 & 169877 & -\tabularnewline
    & this study & 143412 & 199853 & 181226 & 177708 & 177492 & 177149 &
    176890 & 176718\tabularnewline
    \bottomrule
  \end{longtable}}
  \addtocounter{table}{-1}

  \caption{\textbf{Comparing our estimates of forest-cover (in ha) for
      Madagascar with previous studies on the period 1953-2014}. We compared
    our forest-cover with the estimates from three previous studies
    \citep{Harper2007, MEFT2009, ONE2015}. We computed a The increase in
    mangrove and spiny forest covers from \emph{c.} 1953 to \emph{c.} 1973
    in \citet{Harper2007} and our study is most probably due to differences
    in forest definition and mapping methods between the 1953
    aerial-photography derived map and the 1973 Landsat image derived map.}

  \label{tab:comp_forest}
\end{table}
\vspace*{\stretch{1}}

\newpage

\vspace*{\stretch{1}}
\begin{table}[!h]
  {\footnotesize
  \begin{longtable}[]{@{}llrrrrrr@{}}
    \toprule
    Forest type & Source & 1953-1973 & 1973-1990 & 1990-2000 & 2000-2005 &
    2005-2010 & 2010-2013\tabularnewline
    \midrule
    \endhead
    Total & Harper2007 & 91140 (0.30) & 200206 (1.70) & 80740 (0.90) & - & -
    & -\tabularnewline
    & MEFT2009 & - & - & 97174 (0.83) & 53037 (0.53) & - & -\tabularnewline
    & ONE2015 & - & - & - & - & 94803 (1.18) & 163943 (1.50)\tabularnewline
    & this study & 86279 (0.57) & 204712 (1.63) & 88346 (0.85) & 42296
    (0.43) & 69540 (0.73) & 89607 (0.97)\tabularnewline
    Moist & Harper2007 & 94480 (0.60) & 87188 (1.70) & 32200 (0.80) & - & -
    & -\tabularnewline
    & MEFT2009 & - & - & 48283 (0.79) & 17468 (0.35) & - & -\tabularnewline
    & ONE2015 & - & - & - & - & 19721 (0.50) & 37364 (0.94)\tabularnewline
    & this study & 79418 (1.02) & 101163 (1.65) & 39815 (0.78) & 20828
    (0.43) & 26954 (0.57) & 54303 (1.19)\tabularnewline
    Dry & Harper2007 & 11220 (0.20) & 77153 (1.90) & 19820 (0.70) & - & - &
    -\tabularnewline
    & MEFT2009 & - & - & 23561 (0.67) & 11494 (0.40) & - & -\tabularnewline
    & ONE2015 & - & - & - & - & 50567 (1.80) & 97184 (2.29)\tabularnewline
    & this study & 16334 (0.35) & 71174 (1.86) & 28395 (0.92) & 12030 (0.41)
    & 29236 (1.04) & 30795 (1.14)\tabularnewline
    Spiny & Harper2007 & -2580 (-0.10) & 35865 (1.20) & 28170 (1.20) & - & -
    & -\tabularnewline
    & MEFT2009 & - & - & 25190 (1.19) & 22970 (1.23) & - & -\tabularnewline
    & ONE2015 & - & - & - & - & 24599 (1.69) & 30589 (1.66)\tabularnewline
    & this study & -6002 (-0.24) & 31068 (1.34) & 19710 (1.00) & 9385 (0.51)
    & 13255 (0.74) & 4373 (0.25)\tabularnewline
    Mangroves & Harper2007 & - & - & 550 (0.20) & - & - & -\tabularnewline
    & MEFT2009 & - & - & - & - & - & -\tabularnewline
    & ONE2015 & - & - & - & - & 469 (0.32) & 448 (0.20)\tabularnewline
    & this study & -2822 (-1.67) & 1096 (0.57) & 352 (0.20) & 43 (0.02) & 69
    (0.04) & 86 (0.05)\tabularnewline
    \bottomrule
  \end{longtable}}
  \addtocounter{table}{-1}

  \caption{\textbf{Comparing our estimates of annual deforestation
      rates for Madagascar with previous studies on the period
      1953-2014}. Annual deforestation areas (in ha/yr) and annual
    deforestation rates (second number in parenthesis, in \%/yr) are
    provided. For deforestation rates in \%/yr, exact same numbers as
    in scientific articles and reports from previous studies have been
    reported. The way annual deforestation rates in \%/yr have been
    computed in these previous studies can slightly differ from one
    study to another but estimates always correct for the potential
    presences of clouds on forest maps. Annual deforested areas in
    ha/yr have been recomputed from forest-cover estimates in
    Tab.~\ref{tab:comp_forest} (except for \citet{Harper2007} for the
    periods 1973-1990 and 1990-2000 for which annual deforested ares
    in ha/yr were derived from numbers reported in the original
    publication, see methods) and do not correct for the potential
    presence of clouds.}

  \label{tab:comp_defor}
\end{table}
\vspace*{\stretch{1}}

\newpage

\vspace*{\stretch{1}}
\begin{table}[!h]

  {\small
  \begin{longtable}[]{@{}rrrrrrrr@{}}
    \toprule
    Year & Forest (ha) & patch (\%) & transitional (\%) & edge (\%) &
    perforated (\%) & interior (\%) & NA (\%)\tabularnewline
    \midrule
    \endhead
    1953 & 15962870 & 0.01 & 1.12 & 4.46 & 0.58 & 93.83 &
    0.00\tabularnewline
    1973 & 14228217 & 2.21 & 7.25 & 19.81 & 2.86 & 67.87 &
    0.01\tabularnewline
    1990 & 10749572 & 3.00 & 8.17 & 21.28 & 3.81 & 63.73 &
    0.01\tabularnewline
    2000 & 9866145 & 3.09 & 8.37 & 22.13 & 3.92 & 62.49 &
    0.01\tabularnewline
    2005 & 9659861 & 3.51 & 8.88 & 22.56 & 6.44 & 58.59 &
    0.02\tabularnewline
    2010 & 9306528 & 4.28 & 9.72 & 22.94 & 8.52 & 54.52 &
    0.02\tabularnewline
    2014 & 8911481 & 5.18 & 10.72 & 23.25 & 10.58 & 50.24 &
    0.03\tabularnewline
    \bottomrule
  \end{longtable}}
  \addtocounter{table}{-1}

  \caption{\textbf{Evolution of the forest fragmentation from 1953 to 2014
      in Madagascar}. Six categories of fragmentation were identified from the
    amount of forest and its occurrence as adjacent forest pixels:
    ``interior'', ``perforated'', ``edge'', ``transitional'', ``patch'', and
    ``undetermined'' \citep{Riitters2000}. We used a moving window of 7x7
    pixels (4.4 ha). Using this window size, forest edge had a width of
    about 90m. The ``interior'' category can be interpreted as the most
    intact forest. The ``patch'' and ``transitional'' categories correspond
    to isolated small forest patches.}

  \label{tab:frag}
\end{table}
\vspace*{\stretch{1}}

\newpage

\section{Figures}
\label{figures}

\nopagebreak

\vfill
\begin{figure}[h!]
  \centering

    \includegraphics[width=11cm]{outputs/ecoregion.png}

    \caption{\textbf{Ecoregions and forest types in Madagascar.}
      Madagascar can be divided into four climatic ecoregions with four forest
      types: the moist forest in the East (green), the dry forest in the West
      (orange), the spiny forest in the South (red), and the mangroves on the
      West coast (blue). Ecoregions were defined following climatic
      \citep{Cornet1974} and vegetation \citep{IEFN1996} criteria. The dark
      grey areas represent the remaining natural forest cover for the year
      2014.}

    \label{fig:ecoregion}

\end{figure}
\vfill

\newpage

\vspace*{\stretch{1}}
\begin{figure}[h!]
  \centering

  \includegraphics[width=\textwidth]{outputs/fig_fcc_highres.png}

  \caption{\textbf{Forest-cover change on six decades from 1953
      to 2014 in Madagascar.} Forest cover changes from \emph{c.} 1973 to 2014
    are shown in the main figure, and forest cover in \emph{c.} 1953 is
    shown in the bottom-right inset. Two zooms in the western dry (left
    part) and eastern moist (right part) ecoregions present more detailed
    views of (from top to bottom): forest-cover in 1950s, forest-cover
    change from \emph{c.} 1973 to 2014, forest fragmentation in 2014 and
    distance to forest edge in 2014. Data on water bodies (blue) and water
    seasonality (light blue for seasonal water to dark blue for permanent
    water) has been extracted from \citet{Pekel2016}.}

  \label{fig:fcc}

\end{figure}
\vspace*{\stretch{1}}

\newpage

\vspace*{\stretch{1}}
\begin{figure}[h!]
  \centering

  \includegraphics[width=13cm]{outputs/dist.png}

  \caption{\textbf{Evolution of the distance to forest edge from
      1953 to 2014 in Madagascar.} Black dots represent the mean distance to
    forest edge for each year. Vertical dashed segments represent the 90\%
    quantiles (5\% and 95\%) of the distance to forest edge. Horizontal
    dashed grey line indicates a distance to forest edge of 100~m.
    Percentages indicate the percentage of forest at a distance to forest
    edge lower than 100~m for each year.}

  \label{fig:dist_edge}

\end{figure}
\vspace*{\stretch{1}}

\end{document}

### 2017-06-07 deforestmap.R

@@ -924,7 +924,7 @@ options(knitr.kable.NA = "-")
 ## Knit and translate to html and pdf
 dir.create("report")
 ## Main document
-render("deforestmap.Rmd",output_format=c("pdf_document","html_document"),output_dir="report") # html output
+render("deforestmap.Rmd")
 ## Cover letter
 render("coverletter.md",output_format=c("pdf_document"),output_dir="report") # pdf output
 

### 2017-06-08 latex/deforestmap.tex
Add latex

@@
 \usepackage{lineno} %% Numérotation des lignes
-\usepackage{caption2}
+\usepackage{caption}
 %\linenumbers
@@
   2. In this study, we propose a new approach combining historical
-  (1953-2000) national forest-cover maps with recent (2000-2014)
+  (1953-2000) national forest-cover maps with recent (2001-2014)
   global annual tree cover loss data to look at six decades
   (1953-2014) of deforestation and forest fragmentation in
-  Madagascar. We produced new forest-cover maps at 30m-resolution over
+  Madagascar. We produced new forest-cover maps at 30~m resolution over
   the full territory of Madagascar for the year 1990, and annually
@@
   reach 99,000 ha/yr during 2010-2014 (corresponding to a rate of
-  1.08\%/yr). Around half of the forest (46\%) is now located at less
-  than 100m from the forest edge.
-
-  4. \emph{Policy implications}: Increase in the deforestation rate
-  since 2005 can be related to rapid population growth (close to
-  3\%/yr) and to poor law enforcement due to political instability in
-  the country.  More effort, including conservation and development
-  programs, is needed in Madagascar to protect the remaining natural
-  tropical forests, both to enhance local people's livelihoods and to
-  support global environmental efforts.
+  1.08\%/yr). This increase is probably due to rapid population growth
+  (close to 3\%/yr) and to poor law enforcement in the country. Around
+  half of the forest (46\%) is now located at less than 100m from the
+  forest edge.
+
+  4. \emph{Policy implications}: Accurate forest-cover change maps can
+  be used to assess the effectiveness of past and current conservation
+  programs and implement new strategies for the future. In particular,
+  forest maps and estimates can be used in the framework of the REDD+
+  (``Reducing Emissions from Deforestation and Forest Degradation'')
+  initiative and for optimizing the current protected area network.
 
@@
 
-\keywords{biodiversity, climate-change, deforestation, Landsat,
-  Madagascar, tropical forest}
+\keywords{biodiversity, climate-change, deforestation, Madagascar, tropical forest}
 
@@
 concentrated in the tropical forests of the island which can be
-divided into three types: the moist forest in the East, the dry forest
-in the West and the spiny forest in the South
-\citep{Vieilledent2016}. This unparalleled biodiversity is severely
-threatened by anthropogenic deforestation \citep{Harper2007,
-  Vieilledent2013} associated with activities such as slash-and-burn
+divided into four types: the moist forest in the East, the dry forest
+in the West, the spiny forest in the South and the mangroves on the
+West coast \citep{Vieilledent2016}. This unparalleled biodiversity is
+severely threatened by deforestation \citep{Harper2007,
+  Vieilledent2013} associated with human activities such as slash-and-burn
 agriculture and pasture \citep{Scales2011}. Tropical forests in
@@
 and high rates of deforestation in Madagascar are responsible for
-large CO2 emissions in the atmosphere
+large CO$_2$ emissions in the atmosphere
 \citep{Achard2014}. Deforestation threatens species survival by
@@
 \citet{Harper2007} produced maps of forest cover and forest cover
-changes over Madagascar for the years c. 1953, c. 1973, 1990 and 2000.
-The c. 1953 forest map was derived from the visual interpretation of
-aerial photography at coarse scale (1/1,000,000). Forest maps for the
-years c. 1973, 1990 and 2000, were obtained from supervised
-classification of Landsat satellite images at 60m resolution (for the
-year 1973) or 30m resolution (for years 1990 and 2000) and can be used
-to derive accurate estimates (89.5\% accuracy reported for the forest
-/ non-forest map of year 2000). Maps provided by \citet{Harper2007}
-were not exhaustive (due to the presence of clouds in the satellite
-imagery), e.g.~11 244 km2 are mapped as unknown cover type for the
+changes over Madagascar for the years \emph{c.}~1953, \emph{c.}~1973,
+1990 and 2000. The \emph{c.}~1953 forest map was derived from the
+visual interpretation of aerial photography at coarse scale
+(1/1,000,000). Forest maps for the years \emph{c.}~1973, 1990, and
+2000 were obtained from supervised classification of Landsat satellite
+images at 60~m resolution (for the year 1973) or 30~m resolution (for
+years 1990 and 2000) and can be used to derive more accurate estimates
+of forest cover (89.5\% accuracy reported for the forest/non-forest
+map of year 2000). Nonetheless, maps provided by \citet{Harper2007}
+are not exhaustive (due to the presence of clouds in the satellite
+imagery), e.g.~11 244~km2 are mapped as unknown cover type for the
 year 2000. Using a similar supervised classification approach as in
@@
 to give recent estimates of deforestation \citep{MEFT2009, ONE2013},
-include large areas of missing information (due to images with more
-than 20\% of clouds) for the 2010 map provided by \citep{ONE2013}, or
-show large mis-classification in specific areas, especially in the dry
-and spiny forest domain for which the spectral answer has a strong
-seasonal behavior due to the deciduousness of such forests (overall
-accuracy is lower than 0.8 for the dry and spiny forests for the maps
-produced by \citet{ONE2015}). Moreover, the production of such forest
-maps from a supervised classification approach requires significant
-resources, especially regarding the image selection step (required to
-minimize cloud cover) and the training step (visual interpretation of
-a large number of polygons needed to train the classification
-algorithm) \citep{Rakotomalala2015}. Most of this work of image
-selection and visual interpretation would need to be repeated to
-produce new forest maps in the future using a similar approach.
+include large areas of missing information due to images with high
+percentage of cloud cover \citep{ONE2013}, or show large
+mis-classification in specific areas, especially in the dry and spiny
+forest domain for which the spectral answer has a strong seasonal
+behavior due to the deciduousness of such forests (overall accuracy is
+lower than 0.8 for the dry and spiny forests for the maps produced by
+\citet{ONE2015}). Moreover, the production of such forest maps from a
+supervised classification approach requires significant resources,
+especially regarding the image selection step (required to minimize
+cloud cover) and the training step (visual interpretation of a large
+number of polygons needed to train the classification algorithm)
+\citep{Rakotomalala2015}. Most of this work of image selection and
+visual interpretation would need to be repeated to produce new forest
+maps in the future using a similar approach.
 
@@
 and can be tested at the national scale for
-Madagascar. \citep{Kim2014} produced a global forest-cover change map
+Madagascar. \citet{Kim2014} produced a global forest-cover change map
 from 1990 to 2000 (derived from Landsat imagery). This product was
@@
 the approach of \citet{Harper2007}, we also assess trends in
-deforestation rates and forest fragmentation from c. 1953 to 2014. The
+deforestation rates and forest fragmentation from \emph{c.} 1953 to 2014. The
 approach described in this study can help assess the effectiveness of
@@
 
+\newpage
+
 \section{Materials and Methods}
@@
 
-We produced annual forest/non-forest maps at 30m resolution for the
+We produced annual forest/non-forest maps at 30~m resolution for the
 full territory of Madagascar for the period 2000-2014 by combining the
@@
 from \citet{Hansen2013}. The 2000 Harper's forest map includes 208,000
-ha of unclassified areas due to the presence of clouds in input
-satellite imagery, mostly (88\%) within the moist forest domain which
-covers 4.17 million ha in total. To provide a label (forest or
-non-forest) to these unclassified pixels, we use the 2000 tree cover
-percentage map of \citet{Hansen2013} by selecting a threshold of 75\%
-tree cover to define forest cover as recommended by other studies for
-the moist domain \citep{Achard2014}. We thus obtained a forest-cover
-map for the year 2000 covering the full territory of Madagascar. We
-then combined this forest-cover map of the year 2000 with the annual
-forest cover loss maps from 2001 to 2014 provided by
-\citet{Hansen2013} to create annual forest-cover maps from 2001 to
-2014 at 30m resolution. We also completed the Harper's forest map of
-year 1990 by filling unclassified areas (due to the presence of clouds
-on satellite images) using our forest-cover map of year 2000. To do
-so, we assumed that if forest was present in 2000, the pixel was also
-forested in 1990. The remaining unclassified pixels were limited to a
-relatively small total area of c. 8,000 ha. We labeled these residual
-pixels as non-forest as for the year 2000.  Similarly we completed the
-Harper forest map of year 1973 by filling unclassified areas using our
-forest-cover map of the year 1990 assuming that if forest was present
-in 1990, it was also present in 1973.  Contrary to the year 1990, the
-remaining unclassified pixels for year 1973 corresponded to a
-significant total area of 3.32 million ha. We also reprojected the
-forest-cover map of year 1953 to a common projection in order to
-compare to the other eras. This map was produced by scanning a paper
-map derived from aerial photos, and thus could not be perfectly
-aligned with the other maps produced through digital processing of
-satellite imagery \citep{Harper2007}. Finally for all forest-cover
-maps from 1973, the isolated single non-forest pixels (i.e.~fully
-surrounded by forest pixels) were removed, assuming that single
-non-forest pixels inside a forest patch were not corresponding to
-deforestation (they might correspond to selective logging activities).
-This allowed us to avoid counting very small scale events
-(\textless{}0.1 ha such as selective logging) as forest fragmentation.
-All the resulting maps are freely available at
+ha of unclassified areas due to the presence of clouds on satellite
+images, mostly (88\%) within the moist forest domain which covered
+4.17 Mha in total in 2000. To provide a label (forest or non-forest)
+to these unclassified pixels, we used the 2000 tree cover percentage
+map of \citet{Hansen2013} by selecting a threshold of 75\% tree cover
+to define forest cover as recommended by other studies for the moist
+domain \citep{Achard2014}. We thus obtained a forest-cover map for the
+year 2000 covering the full territory of Madagascar. We then combined
+this forest-cover map of the year 2000 with the annual tree cover loss
+maps from 2001 to 2014 provided by \citet{Hansen2013} to create annual
+forest-cover maps from 2001 to 2014 at 30~m resolution. We also
+completed the Harper's forest map of year 1990 by filling unclassified
+areas (due to the presence of clouds on satellite images) using our
+forest-cover map of year 2000. To do so, we assumed that if forest was
+present in 2000, the pixel was also forested in 1990. The remaining
+unclassified pixels were limited to a relatively small total area of
+\emph{c.} 8,000 ha. We labeled these residual pixels as non-forest as
+for the year 2000. Similarly we completed the Harper's forest map of
+year 1973 by filling unclassified areas using our forest-cover map of
+the year 1990 assuming that if forest was present in 1990, it was also
+present in 1973. Contrary to the year 1990, the remaining unclassified
+pixels for year 1973 corresponded to a significant total area of 3.32
+million ha. We also reprojected the forest-cover map of year 1953 to a
+common projection in order to compare the forest-cover area in 1953
+with forest-cover areas at the following dates. This map was produced
+by scanning a paper map derived from aerial photos, and thus could not
+be perfectly aligned with the other maps produced through digital
+processing of satellite imagery \citep{Harper2007}. Finally for all
+forest-cover maps from 1973, the isolated single non-forest pixels
+(i.e.~fully surrounded by forest pixels) were removed, assuming that
+single non-forest pixels inside a forest patch were not corresponding
+to deforestation (they might correspond to selective logging
+activities). This allowed us to avoid counting very small scale events
+(\textless{}0.1 ha such as selective logging) as forest
+fragmentation. All the resulting maps are freely available at
 \url{https://bioscenemada.cirad.fr/forestmaps}.
@@
 
-\begin{equation}
+\begin{equation*}
   \theta = 100 \times [1-(1-(F_{t_2}-F_{t_1})/F_{t_1})^{(1/(t_2-t_1))}
-\end{equation}
+\end{equation*}
 
@@
 We also conducted an analysis of changes in forest fragmentation for
-the years 1953, 1973, 1990, 2000, 2010 and 2014. We applied the method
-developed by \citet{Riitters2000} which uses a moving window to
+the years 1953, 1973, 1990, 2000, 2005, 2010 and 2014. We applied the
+method developed by \citet{Riitters2000} which uses a moving window to
 characterize the fragmentation around each forested
@@
 \texttt{gdal\_proximity.py} of the GDAL software
-(\url{http://www.gdal.org/}). We computed the mean and 95\% quantiles
-of the distance to forest edge and looked at the evolution of these
-values with time.
+(\url{http://www.gdal.org/}). We computed the mean and 90\% quantiles
+(5\% and 95\%) of the distance to forest edge and looked at the
+evolution of these values with time.
+
+\newpage
 
@@
 Tab.~\ref{tab:forest_cover}). Madagascar has lost 44\% and 37\% of its
-natural forests between 1953 and 2013, and between c. 1973 and 2014
+natural forests between 1953 and 2014, and between 1973 and 2014
 respectively (Fig.~\ref{fig:fcc} and Tab.~\ref{tab:forest_cover}). In
@@
 Tab.~\ref{tab:comp_forest}). Regarding the deforestation trend, we
-observed a progressive decrease of the deforestation rate after1990
+observed a progressive decrease of the deforestation rate after 1990
 from 205,000 ha/yr (1.63\%/yr) over the period 1973-1990 to 44,300
@@
 Forest-cover maps provided by previous studies over Madagascar were
-not exhaustive (unclassified areas) due to the presence of clouds in
-satellite imagery used to produce such maps. In \citet{Harper2007},
-the maps of years 1990 and 2000 include 0.5 and 1.12 Mha of unknown
-cover type respectively. Proportions of unclassified areas are not
-reported in the two other existing studies by \citet{MEFT2009} and
+not exhaustive (unclassified areas) due to the presence of clouds on
+satellite images used to produce such maps. In \citet{Harper2007}, the
+maps of years 1990 and 2000 include 0.5 and 1.12 Mha of unknown cover
+type respectively. Proportions of unclassified areas are not reported
+in the two other existing studies by \citet{MEFT2009} and
 \citet{ONE2015}. With our approach, we produced wall to wall
@@
 (considering forest cover estimates at national level and by
-ecoregions for the 6 available dates) were well correlated (0.99) to
-estimates from the three previous studies (Tab.~\ref{tab:comp_forest})
-with a RMSE of 300,000 ha (6\% of the mean forest cover of 4.8 Mha
-when considering all dates and forest types together). These small
-differences can be partly attributed to differences in ecoregion
-boundaries. Despite significant differences in deforestation estimates
+ecoregions for all the available dates) were well correlated
+(Pearson's correlation coefficient $=$ 0.99) to estimates from the
+three previous studies (Tab.~\ref{tab:comp_forest}) with a RMSE of
+300,000 ha (6\% of the mean forest cover of 4.8 Mha when considering
+all dates and forest types together). These small differences can be
+partly attributed to differences in ecoregion boundaries. Despite
+significant differences in deforestation estimates
 (Tab.~\ref{tab:comp_defor}), a similar deforestation trend was
@@
 2014 in Madagascar. The mean distance to forest edge has decreased to
-c.  300 m in 2014 while it was previously c. 1.5 km in 1973
+\emph{c.} 300~m in 2014 while it was previously \emph{c.} 1.5~km in 1973
 (Fig.~\ref{fig:dist_edge}). Moreover, a large proportion (73\%) of
@@
 almost half of the forest (46\%) was at a distance lower than 100 m
-from forest edge in 2014 (Fig.  3). The percentage of forest that can
-be considered intact in Madagascar has continuously decreased since
-1953. The percentage of forest belonging to the ``interior'' category
-(most intact forests) has fallen from 68\% in 1973 to 50\% in 2014. In
-2014, more than 16\% of the forest belonged to the ``patch'' and
-``transitional'' categories (isolated small forest patches) compared
-to 9.5\% in 1973 (Tab.~\ref{tab:frag}).
+from forest edge in 2014 (Fig.~\ref{fig:dist_edge}). The percentage of
+forest that can be considered intact in Madagascar has continuously
+decreased since 1953. The percentage of forest belonging to the
+``interior'' category (most intact forests) has fallen from 68\% in
+1973 to 50\% in 2014. In 2014, more than 16\% of the forest belonged
+to the ``patch'' and ``transitional'' categories (isolated small
+forest patches) compared to 9.5\% in 1973 (Tab.~\ref{tab:frag}).
+
+\newpage
 
@@
 loss data \citep{Hansen2013} with historical (1953-2000) national
-forest-cover maps \citep{Harper2007} to look at six decades (1953-2014)
-of deforestation and forest fragmentation in Madagascar. We produced
-annual forest-cover maps at 30m-resolution covering Madagascar for the
-period 2000 to 2014. Our study extends the forest-cover monitoring on a
-six decades period (from 1953 to 2014) while harmonizing the data from
-the previous studies. We propose a generic approach to solve the problem
-of forest definition which is needed to transform the 2000 global tree
-cover dataset from Hansen et al. (2013) into a forest/non-forest map
-\citep{Tropek2014}. We propose to use a historical national forest-cover
-map, based on a national forest definition, as a forest cover mask. This
-approach could be easily extended to other regions or countries for
-which an accurate forest-cover map is available at any date within the
-period 2000-2014, but preferably at the beginning of the period to
-profit from the full record and derive long-term estimates of
-deforestation. Moreover, this approach can be repeated in the future if
-and when the global tree cover product is updated. We have made the
-R/GRASS code used for this study freely available in a GitHub repository
-(see Data availability statement) to facilitate application to other
-study areas.
-
-The accuracy of the derived forest-cover change maps depends directly on
-the accuracies of the historical forest-cover maps and the tree cover
-loss dataset. The reported global accuracy of the tree cover loss
-dataset is 99.6\% (see Tab. S5 in \citet{Hansen2013}).
+forest-cover maps \citep{Harper2007} to look at six decades
+(1953-2014) of deforestation and forest fragmentation in
+Madagascar. We produced annual forest-cover maps at 30~m resolution
+covering Madagascar for the period 2000 to 2014. Our study extends the
+forest-cover monitoring on a six decades period (from 1953 to 2014)
+while harmonizing the data from previous studies \citep{Harper2007,
+  MEFT2009, ONE2015}. We propose a generic approach to solve the
+problem of forest definition which is needed to transform the 2000
+global tree cover dataset from \citet{Hansen2013} into a
+forest/non-forest map \citep{Tropek2014}. We propose to use a
+historical national forest-cover map, based on a national forest
+definition, as a forest cover mask. This approach could be easily
+extended to other regions or countries for which an accurate
+forest-cover map is available at any date within the period 2000-2014,
+but preferably at the beginning of the period to profit from the full
+record and derive long-term estimates of deforestation. Moreover, this
+approach can be repeated in the future if and when the global tree
+cover product is updated. We have made the R/GRASS code used for this
+study freely available in a GitHub repository (see Data availability
+statement) to facilitate application to other study areas or repeat
+the analysis in the future for Madagascar.
+
+The accuracy of the derived forest-cover change maps depends directly
+on the accuracies of the historical forest-cover maps and the tree
+cover loss dataset. The reported global accuracy of the tree cover
+loss dataset is 99.6\% (see Tab. S5 in \citet{Hansen2013}).
 \citet{Verhegghen2016} have compared deforestation estimates derived
-from the global tree cover loss dataset \citep{Hansen2013} with results
-derived from semi-automated supervised classification of Landsat
-satellite images \citep{Achard2014} for six countries in Central Africa
-and they found a good agreement between these two sets of estimates.
-Consistent with \citet{Harper2007}, we did not consider potential forest
-regrowth in Madagascar (although \citet{Hansen2013} provided a tree
-cover gains layer for the period 2001-2013) for the following reasons:
-(i) the tree gain layer of \citet{Hansen2013} includes and catches more
-easily tree plantations than natural forest regrowth \citep{Tropek2014};
-(ii) there is little evidence of natural forest regeneration in
-Madagascar \citep[\citet{Grouzis2001}]{Harper2007} - this can be
-explained by several ecological processes following burning practice
-such as soil erosion \citep{Grinand2017} and reduced seed bank due to
-fire and soil loss \citep{Grouzis2001}, (iii) in areas where forest
-regeneration is ecologically possible, young forest regrowth are more
-easily re-burnt for agriculture and pasture; and (iv) young secondary
-forests provide more limited ecosystem services compared to old-growth
-natural forests in terms of biodiversity and carbon storage.
+from the global tree cover loss dataset \citep{Hansen2013} with
+results derived from semi-automated supervised classification of
+Landsat satellite images \citep{Achard2014} for six countries in
+Central Africa and they found a good agreement between these two sets
+of estimates. Consistent with \citet{Harper2007}, we did not consider
+potential forest regrowth in Madagascar (although \citet{Hansen2013}
+provided a tree cover gains layer for the period 2001-2013) for
+several reasons. First, the tree gain layer of \citet{Hansen2013}
+includes and catches more easily tree plantations than natural forest
+regrowth \citep{Tropek2014}. Second, there is little evidence of
+natural forest regeneration in Madagascar \citep{Grouzis2001,
+ Harper2007}. This can be explained by several ecological processes
+following burning practice such as soil erosion \citep{Grinand2017}
+and reduced seed bank due to fire and soil loss
+\citep{Grouzis2001}. Moreover, in areas where forest regeneration is
+ecologically possible, young forest regrowth are more easily re-burnt
+for agriculture and pasture. Third, young secondary forests provide
+more limited ecosystem services compared to old-growth natural forests
+in terms of biodiversity and carbon storage.
 
@@
 We estimated that natural forests in Madagascar cover 8.9 Mha in 2014
-(corresponding to 15\% of the country) and that Madagascar has lost 44\%
-of its natural forest since 1953 (37\% since 1973). There is ongoing
-scientific debate about the extent of the ``original'' forest cover in
-Madagascar, and the extent to which humans have altered the natural
-forest landscapes since their large-scale settlement around 800 CE
-\citep{Burns2016, Cox2012}. Early French naturalists stated that the
-full island was originally covered by forest
+(corresponding to 15\% of the country) and that Madagascar has lost
+44\% of its natural forest since 1953 (37\% since 1973). There is
+ongoing scientific debate about the extent of the ``original'' forest
+cover in Madagascar, and the extent to which humans have altered the
+natural forest landscapes since their large-scale settlement around
+800 CE \citep{Burns2016, Cox2012}. Early French naturalists stated
+that the full island was originally covered by forest
 \citep{Humbert1927, Perrier1921}, leading to the common statement that
-90\% of the natural forests have disappeared since the arrival of humans
-on the island \citep{Kull2000}. More recent studies counter-balanced
-that point of view saying that extensive areas of grassland existed in
-Madagascar long before human arrival and were determined by climate,
-natural grazing and other natural factors
-\citep[\citet{Vorontsova2017}]{Virah-Sawmy2009}. Other authors have
-questioned the entire narrative of extensive alteration of the landscape
-by early human activity which, through legislation, has severe
-consequences on local people \citep{Klein2002, Kull2000}. Whatever the
-original proportion of natural forests and grasslands in Madagascar, our
+90\% of the natural forests have disappeared since the arrival of
+humans on the island \citep{Kull2000}. More recent studies
+counter-balanced that point of view saying that extensive areas of
+grassland existed in Madagascar long before human arrival and were
+determined by climate, natural grazing and other natural factors
+\citep{Vorontsova2017, Virah-Sawmy2009}. Other authors have questioned
+the entire narrative of extensive alteration of the landscape by early
+human activity which, through legislation, has severe consequences on
+local people \citep{Klein2002, Kull2000}. Whatever the original
+proportion of natural forests and grasslands in Madagascar, our
 results demonstrate that human activities since the 1950s have
-profoundly impacted the natural tropical forests and that recent
-conservation and development programs in Madagascar have failed to stop
-deforestation in the recent years. Deforestation has strong consequences
-on biodiversity and carbon emissions in Madagascar. Around 90\% of
-Madagascar's species are forest dependent
-\citep{Allnutt2008, Goodman2005} and \citet{Allnutt2008} estimated that
-deforestation between 1953 and 2000 led to an extinction of 9\% of the
-species. The additional deforestation we observed over the period
-2000-2014 (around 1Mha of natural forest) worsen this result. Regarding
-carbon emissions, using the 2010 aboveground forest carbon map by
-\citet{Vieilledent2016}, we estimated that deforestation on the period
-2010-2014 has led to 40.2 Mt C of carbon emissions in the atmosphere (10
-Mt C /yr) and that the remaining aboveground forest carbon stock in 2014
-is 832.8 Mt C. Associated to deforestation, we showed that the remaining
-forests of Madagascar are highly fragmented with 46\% of the forest
-being at less than 100m of the forest edge. Small forest fragments do
-not allow to maintain viable populations and ``edge effects'' at
-forest/non-forest interfaces have impacts on both carbon emissions
-\citep{Brinck2017} and biodiversity loss \citep{Gibson2013, Murcia1995}.
-
-\subsection{Deforestation trend and causes of the acceleration of the
-  deforestation since 2005}
+profoundly impacted the natural tropical forests and that conservation
+and development programs in Madagascar have failed to stop
+deforestation in the recent years. Deforestation has strong
+consequences on biodiversity and carbon emissions in
+Madagascar. Around 90\% of Madagascar's species are forest dependent
+\citep{Allnutt2008, Goodman2005} and \citet{Allnutt2008} estimated
+that deforestation between 1953 and 2000 led to an extinction of 9\%
+of the species. The additional deforestation we observed over the
+period 2000-2014 (around 1Mha of natural forest) worsen this
+result. Regarding carbon emissions, using the 2010 aboveground forest
+carbon map by \citet{Vieilledent2016}, we estimated that deforestation
+on the period 2010-2014 has led to 40.2 Mt C of carbon emissions in
+the atmosphere (10 Mt C /yr) and that the remaining aboveground forest
+carbon stock in 2014 is 832.8 Mt C. Associated to deforestation, we
+showed that the remaining forests of Madagascar are highly fragmented
+with 46\% of the forest being at less than 100m of the forest
+edge. Small forest fragments do not allow to maintain viable
+populations and ``edge effects'' at forest/non-forest interfaces have
+impacts on both carbon emissions \citep{Brinck2017} and biodiversity
+loss \citep{Gibson2013, Murcia1995}.
+
+\subsection{Deforestation trend and impacts on conservation and
+  development policies}
 
 In our study, we have shown that the progressive decrease of the
-deforestation rate on the period 1990-2005 was followed by a continuous
-increase in the deforestation rate on the period 2005-2014. In
-particular, we showed that deforestation rate has more than doubled on
-the period 2010-2014 compared to 2000-2005. Our results are confirmed by
-previous studies \citep{Harper2007, MEFT2009, ONE2015} despite
-differences in the methodologies regarding (i) forest definition
-(associated to independent visual interpretations of observation
-polygons to train the classifier), (ii) classification algorithms, (iii)
-deforestation rate computation method, and (iv) correction for the
-presence of clouds. Our deforestation rate estimates from 1990 to 2014
-have been computed from wall to wall maps at 30 m resolution and can be
-considered more accurate in comparison with estimates from these
-previous studies. Our forest-cover and deforestation rate estimates can
-be used as source of information for the next FAO Forest Resources
-Assessment project \citep{Keenan2015}. Current rates of deforestation
-can also be used to build reference scenarios for deforestation in
-Madagascar and contribute to the implementation of deforestation
-mitigation activities in the framework of REDD+ \citep{Olander2008}.
+deforestation rate on the period 1990-2005 was followed by a
+continuous increase in the deforestation rate on the period
+2005-2014. In particular, we showed that deforestation rate has more
+than doubled on the period 2010-2014 compared to 2000-2005. Our
+results are confirmed by previous studies \citep{Harper2007, MEFT2009,
+  ONE2015} despite differences in the methodologies regarding (i)
+forest definition (associated to independent visual interpretations of
+observation polygons to train the classifier), (ii) classification
+algorithms, (iii) deforestation rate computation method, and (iv)
+correction for the presence of clouds. Our deforestation rate
+estimates from 1990 to 2014 have been computed from wall to wall maps
+at 30~m resolution and can be considered more accurate in comparison
+with estimates from these previous studies. Our forest-cover and
+deforestation rate estimates can be used as source of information for
+the next FAO Forest Resources Assessment project
+\citep{Keenan2015}. Current rates of deforestation can also be used to
+build reference scenarios for deforestation in Madagascar and
+contribute to the implementation of deforestation mitigation
+activities in the framework of REDD+ \citep{Olander2008}.
 
 The increase of deforestation rates after 2005 can be explained by
-population growth and political instability in the country. Nearly 90\%
-of Madagascar's population relies on biomass for their daily energy
-needs \citep{Minten2013} and the link between population size and
-deforestation has previously been demonstrated in Madagascar
+population growth and political instability in the country. Nearly
+90\% of Madagascar's population relies on biomass for their daily
+energy needs \citep{Minten2013} and the link between population size
+and deforestation has previously been demonstrated in Madagascar
 \citep{Vieilledent2013, Gorenflo2011}. With a mean demographic growth
 rate of about 2.8\%/yr and a population which has increased from 16 to
-24 million people on the period 2000-2015 \citep{UN2015}, the increasing
-demand in wood-fuel and space for agriculture is likely to explain the
-increase in deforestation rates. The political crisis of 2009
-\citep{Ploch2012}, followed by several years of political instability
-and weak governance could also explain the increase in the deforestation
-rate observed on the period 2005-2014 (Smith et al., 2003). These
-results show that despite the conservation policy in Madagascar
-\citep{Freudenberger2010}, deforestation has dramatically increased at
-the national level since 2005. More efforts should be done to identify
-the real causes and mechanisms explaining this rampant deforestation in
-order to implement new conservation strategies that will be effective at
-conserving the few remaining natural tropical forests in Madagascar.
+24 million people on the period 2000-2015 \citep{UN2015}, the
+increasing demand in wood-fuel and space for agriculture is likely to
+explain the increase in deforestation rates. The political crisis of
+2009 \citep{Ploch2012}, followed by several years of political
+instability and weak governance could also explain the increase in the
+deforestation rate observed on the period 2005-2014
+\citep{Smith2003}. These results show that despite the conservation
+policy in Madagascar \citep{Freudenberger2010}, deforestation has
+dramatically increased at the national level since 2005. Results of
+this study, including recent spatially explicit forest-cover change
+maps and forest-cover estimates, should help implement new
+conservation strategies to save Madagascar natural tropical forests
+and their unique biodiversity.
 
@@
 in the \texttt{deforestmap} GitHub repository
-(\url{https://github.com/ghislainv/deforestmap.git}). The results and
-the manuscript are fully reproducible running the {\R} script
-\texttt{deforestmap.R} located inside the \texttt{deforestmap}
-repository.
+(\url{https://github.com/ghislainv/deforestmap.git}). The results are
+fully reproducible running the {\R} script \texttt{deforestmap.R}
+located inside the \texttt{deforestmap} repository.
 
@@
 
-\bibliographystyle{gcb}
-\bibliography{bib/biblio}
+%%\bibliographystyle{bib/jae}
+%%\bibliography{bib/biblio}
+
+\section{References}
+
+\renewcommand{\bibsection}{} %% Remove section name for bibliography
+\begin{thebibliography}{55}
+\providecommand{\natexlab}[1]{#1}
+
+\bibitem[{Achard \emph{et~al.}(2014)Achard, Beuchle, Mayaux, Stibig, Bodart,
+  Brink, Carboni, Desclée, Donnay, Eva, Lupi, Raši, Seliger \&
+  Simonetti}]{Achard2014}
+Achard, F., Beuchle, R., Mayaux, P., Stibig, H.J., Bodart, C., Brink, A.,
+  Carboni, S., Desclée, B., Donnay, F., Eva, H.D., Lupi, A., Raši, R.,
+  Seliger, R. \& Simonetti, D. (2014) Determination of tropical deforestation
+  rates and related carbon losses from 1990 to 2010.
+\newblock \emph{Global Change Biology}, \textbf{20}, 2540--2554.
+
+\bibitem[{Ali \& Aitchison(2008)}]{Ali2008}
+Ali, J.R. \& Aitchison, J.C. (2008) {Gondwana to Asia: Plate tectonics,
+  paleogeography and the biological connectivity of the Indian sub-continent
+  from the Middle Jurassic through latest Eocene (166--35 Ma)}.
+\newblock \emph{Earth-Science Reviews}, \textbf{88}, 145--166.
+
+\bibitem[{Allnutt \emph{et~al.}(2008)Allnutt, Ferrier, Manion, Powell,
+  Ricketts, Fisher, Harper, Irwin, Kremen, Labat, Lees, Pearce \&
+  Rakotondrainibe}]{Allnutt2008}
+Allnutt, T.F., Ferrier, S., Manion, G., Powell, G.V.N., Ricketts, T.H., Fisher,
+  B.L., Harper, G.J., Irwin, M.E., Kremen, C., Labat, J.N., Lees, D.C., Pearce,
+  T.A. \& Rakotondrainibe, F. (2008) {A method for quantifying biodiversity
+  loss and its application to a 50-year record of deforestation across
+  Madagascar}.
+\newblock \emph{Conservation Letters}, \textbf{1}, 173--181.
+
+\bibitem[{Armitage \emph{et~al.}(2015)Armitage, Highfield, Brody \&
+  Louchouarn}]{Armitage2015}
+Armitage, A.R., Highfield, W.E., Brody, S.D. \& Louchouarn, P. (2015) {The
+  contribution of mangrove expansion to salt marsh loss on the Texas Gulf
+  Coast}.
+\newblock \emph{PloS One}, \textbf{10}, e0125404.
+
+\bibitem[{Bastin \emph{et~al.}(2017)Bastin, Berrahmouni, Grainger, Maniatis,
+  Mollicone, Moore, Patriarca, Picard, Sparrow, Abraham, Aloui, Atesoglu,
+  Attore, Bass{\"u}ll{\"u}, Bey, Garzuglia, Garc{\'\i}a-Montero, Groot, Guerin,
+  Laestadius, Lowe, Mamane, Marchi, Patterson, Rezende, Ricci, Salcedo, Diaz,
+  Stolle, Surappaeva \& Castro}]{Bastin2017}
+Bastin, J.F., Berrahmouni, N., Grainger, A., Maniatis, D., Mollicone, D.,
+  Moore, R., Patriarca, C., Picard, N., Sparrow, B., Abraham, E.M., Aloui, K.,
+  Atesoglu, A., Attore, F., Bass{\"u}ll{\"u}, {\c C}., Bey, A., Garzuglia, M.,
+  Garc{\'\i}a-Montero, L.G., Groot, N., Guerin, G., Laestadius, L., Lowe, A.J.,
+  Mamane, B., Marchi, G., Patterson, P., Rezende, M., Ricci, S., Salcedo, I.,
+  Diaz, A.S.P., Stolle, F., Surappaeva, V. \& Castro, R. (2017) The extent of
+  forest in dryland biomes.
+\newblock \emph{Science}, \textbf{356}, 635--638.
+
+\bibitem[{Brinck \emph{et~al.}(2017)Brinck, Fischer, Groeneveld, Lehmann,
+  Dantas De~Paula, Pütz, Sexton, Song \& Huth}]{Brinck2017}
+Brinck, K., Fischer, R., Groeneveld, J., Lehmann, S., Dantas De~Paula, M.,
+  Pütz, S., Sexton, J.O., Song, D. \& Huth, A. (2017) High resolution analysis
+  of tropical forest fragmentation and its impact on the global carbon cycle.
+\newblock \emph{Nature Communications}, \textbf{8}, 14855.
+
+\bibitem[{Brooks \emph{et~al.}(2002)Brooks, Mittermeier, Mittermeier,
+  da~Fonseca, Rylands, Konstant, Flick, Pilgrim, Oldfield, Magin \&
+  Hilton-Taylor}]{Brooks2002}
+Brooks, T.M., Mittermeier, R.A., Mittermeier, C.G., da~Fonseca, G.A.B.,
+  Rylands, A.B., Konstant, W.R., Flick, P., Pilgrim, J., Oldfield, S., Magin,
+  G. \& Hilton-Taylor, C. (2002) Habitat loss and extinction in the hotspots of
+  biodiversity.
+\newblock \emph{Conservation Biology}, \textbf{16}, 909--923.
+
+\bibitem[{Burns \emph{et~al.}(2016)Burns, Godfrey, Faina, McGee, Hardt,
+  Ranivoharimanana \& Randrianasy}]{Burns2016}
+Burns, S.J., Godfrey, L.R., Faina, P., McGee, D., Hardt, B., Ranivoharimanana,
+  L. \& Randrianasy, J. (2016) {Rapid human-induced landscape transformation in
+  Madagascar at the end of the first millennium of the Common Era}.
+\newblock \emph{Quaternary Science Reviews}, \textbf{134}, 92 -- 99.
+
+\bibitem[{Cornet(1974)}]{Cornet1974}
+Cornet, A. (1974) {Essai de cartographie bioclimatique à Madagascar}.
+\newblock , Orstom.
+
+\bibitem[{Cox \emph{et~al.}(2012)Cox, Nelson, Tumonggor, Ricaut \&
+  Sudoyo}]{Cox2012}
+Cox, M.P., Nelson, M.G., Tumonggor, M.K., Ricaut, F.X. \& Sudoyo, H. (2012) {A
+  small cohort of Island Southeast Asian women founded Madagascar}.
+\newblock \emph{Proceedings of the Royal Society B: Biological Sciences},
+  \textbf{279}, 2761--2768.
+
+\bibitem[{Crottini \emph{et~al.}(2012)Crottini, Madsen, Poux, Strau{\ss},
+  Vieites \& Vences}]{Crottini2012}
+Crottini, A., Madsen, O., Poux, C., Strau{\ss}, A., Vieites, D.R. \& Vences, M.
+  (2012) {Vertebrate time-tree elucidates the biogeographic pattern of a major
+  biotic change around the K--T boundary in Madagascar}.
+\newblock \emph{Proceedings of the National Academy of Sciences}, \textbf{109},
+  5358--5363.
+
+\bibitem[{Eklund \emph{et~al.}(2016)Eklund, Blanchet, Nyman, Rocha, Virtanen \&
+  Cabeza}]{Eklund2016}
+Eklund, J., Blanchet, F.G., Nyman, J., Rocha, R., Virtanen, T. \& Cabeza, M.
+  (2016) {Contrasting spatial and temporal trends of protected area
+  effectiveness in mitigating deforestation in Madagascar}.
+\newblock \emph{Biological Conservation}, \textbf{203}, 290 -- 297.
+
+\bibitem[{Freudenberger(2010)}]{Freudenberger2010}
+Freudenberger, K. (2010) {Paradise Lost? Lessons from 25 years of USAID
+  environment programs in Madagascar}.
+\newblock \emph{International Resources Group, Washington DC}.
+
+\bibitem[{Gibson \emph{et~al.}(2013)Gibson, Lynam, Bradshaw, He, Bickford,
+  Woodruff, Bumrungsri \& Laurance}]{Gibson2013}
+Gibson, L., Lynam, A.J., Bradshaw, C.J., He, F., Bickford, D.P., Woodruff,
+  D.S., Bumrungsri, S. \& Laurance, W.F. (2013) Near-complete extinction of
+  native small mammal fauna 25 years after forest fragmentation.
+\newblock \emph{Science}, \textbf{341}, 1508--1510.
+
+\bibitem[{Goodman \& Benstead(2005)}]{Goodman2005}
+Goodman, S.M. \& Benstead, J.P. (2005) {Updated estimates of biotic diversity
+  and endemism for Madagascar}.
+\newblock \emph{Oryx}, \textbf{39}, 73--77.
+
+\bibitem[{Gorenflo \emph{et~al.}(2011)Gorenflo, Corson, Chomitz, Harper,
+  Honzák \& {\"O}zler}]{Gorenflo2011}
+Gorenflo, L.J., Corson, C., Chomitz, K.M., Harper, G., Honzák, M. \&
+  {\"O}zler, B. (2011) \emph{{Exploring the Association Between People and
+  Deforestation in Madagascar}}, vol. 1650.
+\newblock Springer Berlin Heidelberg.
+
+\bibitem[{Grinand \emph{et~al.}(2017)Grinand, Le~Maire, Vieilledent,
+  Razakamanarivo, Razafimbelo \& Bernoux}]{Grinand2017}
+Grinand, C., Le~Maire, G., Vieilledent, G., Razakamanarivo, H., Razafimbelo, T.
+  \& Bernoux, M. (2017) {Estimating temporal changes in soil carbon stocks at
+  ecoregional scale in Madagascar using remote-sensing}.
+\newblock \emph{International Journal of Applied Earth Observation and
+  Geoinformation}, \textbf{54}, 1--14.
+
+\bibitem[{Grinand \emph{et~al.}(2013)Grinand, Rakotomalala, Gond, Vaudry,
+  Bernoux \& Vieilledent}]{Grinand2013}
+Grinand, C., Rakotomalala, F., Gond, V., Vaudry, R., Bernoux, M. \&
+  Vieilledent, G. (2013) {Estimating deforestation in tropical humid and dry
+  forests in Madagascar from 2000 to 2010 using multi-date Landsat satellite
+  images and the Random Forests classifier}.
+\newblock \emph{Remote Sensing of Environment}, \textbf{139}, 68--80.
+
+\bibitem[{Grouzis \emph{et~al.}(2001)Grouzis, Razanaka, Le~Floc’h \&
+  Leprun}]{Grouzis2001}
+Grouzis, M., Razanaka, S., Le~Floc’h, E. \& Leprun, J.C. (2001)
+  {{\'E}volution de la v{\'e}g{\'e}tation et de quelques param{\`e}tres
+  {\'e}daphiques au cours de la phase post-culturale dans la r{\'e}gion
+  d’Analabo}.
+\newblock \emph{Soci{\'e}t{\'e}s paysannes, transitions agraires et dynamiques
+  {\'e}cologiques dans le Sud-Ouest de Madagascar, Antananarivo, IRD/CNRE}, pp.
+  327--337.
+
+\bibitem[{Hansen \emph{et~al.}(2013)Hansen, Potapov, Moore, Hancher,
+  Turubanova, Tyukavina, Thau, Stehman, Goetz, Loveland, Kommareddy, Egorov,
+  Chini, Justice \& Townshend}]{Hansen2013}
+Hansen, M.C., Potapov, P.V., Moore, R., Hancher, M., Turubanova, S.A.,
+  Tyukavina, A., Thau, D., Stehman, S.V., Goetz, S.J., Loveland, T.R.,
+  Kommareddy, A., Egorov, A., Chini, L., Justice, C.O. \& Townshend, J.R.G.
+  (2013) {High-Resolution Global Maps of 21st-Century Forest Cover Change}.
+\newblock \emph{Science}, \textbf{342}, 850--853.
+
+\bibitem[{Harper \emph{et~al.}(2007)Harper, Steininger, Tucker, Juhn \&
+  Hawkins}]{Harper2007}
+Harper, G.J., Steininger, M.K., Tucker, C.J., Juhn, D. \& Hawkins, F. (2007)
+  {Fifty years of deforestation and forest fragmentation in Madagascar}.
+\newblock \emph{Environmental Conservation}, \textbf{34}, 325--333.
+
+\bibitem[{Humbert(1927)}]{Humbert1927}
+Humbert, H. (1927) {La destruction d'une flore insulaire par le feu. Principaux
+  aspects de la végétation à Madagascar}.
+\newblock \emph{Mémoires de l'Académie Malgache}, \textbf{5}, 1--80.
+
+\bibitem[{Keenan \emph{et~al.}(2015)Keenan, Reams, Achard, de~Freitas, Grainger
+  \& Lindquist}]{Keenan2015}
+Keenan, R.J., Reams, G.A., Achard, F., de~Freitas, J.V., Grainger, A. \&
+  Lindquist, E. (2015) {Dynamics of global forest area: Results from the FAO
+  Global Forest Resources Assessment 2015}.
+\newblock \emph{Forest Ecology and Management}, \textbf{352}, 9 -- 20.
+
+\bibitem[{Kim \emph{et~al.}(2014)Kim, Sexton, Noojipady, Huang, Anand, Channan,
+  Feng \& Townshend}]{Kim2014}
+Kim, D.H., Sexton, J.O., Noojipady, P., Huang, C., Anand, A., Channan, S.,
+  Feng, M. \& Townshend, J.R. (2014) {Global, Landsat-based forest-cover change
+  from 1990 to 2000}.
+\newblock \emph{Remote Sensing of Environment}, \textbf{155}, 178--193.
+
+\bibitem[{Klein(2002)}]{Klein2002}
+Klein, J. (2002) {Deforestation in the Madagascar Highlands -- Established
+  `truth' and scientific uncertainty}.
+\newblock \emph{GeoJournal}, \textbf{56}, 191--199.
+
+\bibitem[{Kull(2000)}]{Kull2000}
+Kull, C.A. (2000) {Deforestation, erosion, and fire: degradation myths in the
+  environmental history of Madagascar}.
+\newblock \emph{Environment and History}, \textbf{6}, 423--450.
+
+\bibitem[{Perrier~de La~B{\^a}thie(1921)}]{Perrier1921}
+Perrier~de La~B{\^a}thie, H. (1921) \emph{{La v{\'e}g{\'e}tation malgache}},
+  vol.~23.
+\newblock Mus{\'e}e Colonial.
+
+\bibitem[{{MEFT, USAID, and CI}(2009)}]{MEFT2009}
+{MEFT, USAID, and CI} (2009) \emph{{Evolution de la couverture de forêts
+  naturelles à Madagascar, 1990-2000-2005}}.
+
+\bibitem[{{Ministère de l'Environnement}(1996)}]{IEFN1996}
+{Ministère de l'Environnement} (1996) {IEFN: Inventaire Ecologique Forestier
+  National}.
+\newblock {Ministère de l'Environnement de Madagascar, Direction des Eaux et
+  Forêts, DFS Deutsch Forest Service GmbH, Entreprise d'études de
+  développement rural ``Mamokatra'', FTM}.
+
+\bibitem[{Minten \emph{et~al.}(2013)Minten, Sander \& Stifel}]{Minten2013}
+Minten, B., Sander, K. \& Stifel, D. (2013) {Forest management and economic
+  rents: Evidence from the charcoal trade in Madagascar}.
+\newblock \emph{Energy for Sustainable Development}, \textbf{17}, 106 -- 115.
+
+\bibitem[{Murcia(1995)}]{Murcia1995}
+Murcia, C. (1995) Edge effects in fragmented forests: implications for
+  conservation.
+\newblock \emph{Trends in Ecology \& Evolution}, \textbf{10}, 58 -- 62.
+
+\bibitem[{Neteler \& Mitasova(2008)}]{Neteler2008}
+Neteler, M. \& Mitasova, H. (2008) \emph{{Open source GIS: a GRASS GIS
+  approach}}.
+\newblock Springer.
+
+\bibitem[{Olander \emph{et~al.}(2008)Olander, Gibbs, Steininger, Swenson \&
+  Murray}]{Olander2008}
+Olander, L.P., Gibbs, H.K., Steininger, M., Swenson, J.J. \& Murray, B.C.
+  (2008) {Reference scenarios for deforestation and forest degradation in
+  support of REDD: a review of data and methods}.
+\newblock \emph{Environmental Research Letters}, \textbf{3}, 025011.
+
+\bibitem[{{ONE, DGF, FTM, MNP, and CI}(2013)}]{ONE2013}
+{ONE, DGF, FTM, MNP, and CI} (2013) \emph{{Evolution de la couverture de
+  forêts naturelles à Madagascar 2005-2010}}.
+
+\bibitem[{{ONE, DGF, MNP, WCS, and Etc Terra}(2015)}]{ONE2015}
+{ONE, DGF, MNP, WCS, and Etc Terra} (2015) \emph{{Changement de la couverture
+  de forêts naturelles à Madagascar, 2005-2010-2013}}.
+
+\bibitem[{Pearson \& Raxworthy(2009)}]{Pearson2009}
+Pearson, R.G. \& Raxworthy, C.J. (2009) {The evolution of local endemism in
+  Madagascar: watershed versus climatic gradient hypotheses evaluated by null
+  biogeographic models}.
+\newblock \emph{Evolution}, \textbf{63}, 959--967.
+
+\bibitem[{Pekel \emph{et~al.}(2016)Pekel, Cottam, Gorelick \&
+  Belward}]{Pekel2016}
+Pekel, J.F., Cottam, A., Gorelick, N. \& Belward, A.S. (2016) {High-resolution
+  mapping of global surface water and its long-term changes}.
+\newblock \emph{Nature}, \textbf{540}, 418--422.
+
+\bibitem[{Ploch \& Cook(2012)}]{Ploch2012}
+Ploch, L. \& Cook, N. (2012) {Madagascar's Political Crisis}.
+
+\bibitem[{Potapov \emph{et~al.}(2017)Potapov, Hansen, Laestadius, Turubanova,
+  Yaroshenko, Thies, Smith, Zhuravleva, Komarova, Minnemeyer \&
+  Esipova}]{Potapov2017}
+Potapov, P., Hansen, M.C., Laestadius, L., Turubanova, S., Yaroshenko, A.,
+  Thies, C., Smith, W., Zhuravleva, I., Komarova, A., Minnemeyer, S. \&
+  Esipova, E. (2017) {The last frontiers of wilderness: Tracking loss of intact
+  forest landscapes from 2000 to 2013}.
+\newblock \emph{Science Advances}, \textbf{3}.
+
+\bibitem[{Puyravaud(2003)}]{Puyravaud2003}
+Puyravaud, J.P. (2003) Standardizing the calculation of the annual rate of
+  deforestation.
+\newblock \emph{Forest Ecology and Management}, \textbf{177}, 593--596.
+
+\bibitem[{Rakotomala \emph{et~al.}(2015)Rakotomala, Rabenandrasana,
+  Andriambahiny, Rajaonson, Andriamalala, Burren, Rakotoarijaona, Parany,
+  Vaudry, Rakotoniaina, Ranaivosoa, Rahagalala, Randrianary \&
+  Grinand}]{Rakotomalala2015}
+Rakotomala, F., Rabenandrasana, J., Andriambahiny, J., Rajaonson, R.,
+  Andriamalala, F., Burren, C., Rakotoarijaona, J., Parany, B., Vaudry, R.,
+  Rakotoniaina, S., Ranaivosoa, R., Rahagalala, P., Randrianary, T. \& Grinand,
+  C. (2015) {Estimation de la d{\'e}forestation des for{\^e}ts humides {\`a}
+  Madagascar entre 2005, 2010 et 2013: combinaison multi-date d'images LANDSAT,
+  utilisation de l'algorithme Random Forest et proc{\'e}dure de validation}.
+\newblock \emph{Revue Fran{\c{c}}aise de Photogramm{\'e}trie et de
+  T{\'e}l{\'e}d{\'e}tection}, pp. 11--23.
+
+\bibitem[{Riitters \emph{et~al.}(2000)Riitters, Wickham, O'Neill, Jones \&
+  Smith}]{Riitters2000}
+Riitters, K., Wickham, J., O'Neill, R., Jones, B. \& Smith, E. (2000)
+  Global-scale patterns of forest fragmentation.
+\newblock \emph{Conservation Ecology}, \textbf{4}, 3.
+
+\bibitem[{Saatchi \emph{et~al.}(2011)Saatchi, Harris, Brown, Lefsky, Mitchard,
+  Salas, Zutta, Buermann, Lewis, Hagen, Petrova, White, Silman \&
+  Morel}]{Saatchi2011}
+Saatchi, S.S., Harris, N.L., Brown, S., Lefsky, M., Mitchard, E.T.A., Salas,
+  W., Zutta, B.R., Buermann, W., Lewis, S.L., Hagen, S., Petrova, S., White,
+  L., Silman, M. \& Morel, A. (2011) Benchmark map of forest carbon stocks in
+  tropical regions across three continents.
+\newblock \emph{Proceedings of the National Academy of Sciences}, \textbf{108},
+  9899--9904.
+
+\bibitem[{Saunders \emph{et~al.}(1991)Saunders, Hobbs \&
+  Margules}]{Saunders1991}
+Saunders, D.A., Hobbs, R.J. \& Margules, C.R. (1991) {Biological Consequences
+  of Ecosystem Fragmentation: A Review}.
+\newblock \emph{Conservation Biology}, \textbf{5}, 18--32.
+
+\bibitem[{Scales(2011)}]{Scales2011}
+Scales, I.R. (2011) {Farming at the Forest Frontier: Land Use and Landscape
+  Change in Western Madagascar, 1896-2005}.
+  \newblock \emph{Environment and History}, \textbf{17}, 499--524.
+
+\bibitem[{Smith \emph{et~al.}(2003)Smith, Muir, Walpole, Balmford \&
+  Leader-Williams}]{Smith2003}
+Smith, R.J., Muir, R.D.J., Walpole, M.J., Balmford, A. \& Leader-Williams, N.
+  (2003) Governance and the loss of biodiversity.
+\newblock \emph{Nature}, \textbf{426}, 67--70.
+
+\bibitem[{Tidd \emph{et~al.}(2001)Tidd, Pinder \& Ferguson}]{Tidd2001}
+Tidd, S.T., Pinder, J. \& Ferguson, G.W. (2001) {Deforestation and habitat loss
+  for the Malagasy flat-tailed tortoise from 1963 through 1993}.
+\newblock \emph{Chelonian Conservation and Biology}, \textbf{4}, 59--65.
+
+\bibitem[{Tropek \emph{et~al.}(2014)Tropek, Sedl{\'a}{\v c}ek, Beck, Keil,
+  Musilov{\'a}, {\v S}{\'\i}mov{\'a} \& Storch}]{Tropek2014}
+Tropek, R., Sedl{\'a}{\v c}ek, O., Beck, J., Keil, P., Musilov{\'a}, Z., {\v
+  S}{\'\i}mov{\'a}, I. \& Storch, D. (2014) {Comment on "High-resolution global
+  maps of 21st-century forest cover change"}.
+\newblock \emph{Science}, \textbf{344}, 981--981.
+
+\bibitem[{Tyukavina \emph{et~al.}(2017)Tyukavina, Hansen, Potapov, Stehman,
+  Smith-Rodriguez, Okpa \& Aguilar}]{Tyukavina2017}
+Tyukavina, A., Hansen, M.C., Potapov, P.V., Stehman, S.V., Smith-Rodriguez, K.,
+  Okpa, C. \& Aguilar, R. (2017) {Types and rates of forest disturbance in
+  Brazilian Legal Amazon, 2000{\textendash}2013}.
+\newblock \emph{Science Advances}, \textbf{3}.
+
+\bibitem[{{United Nations}(2015)}]{UN2015}
+{United Nations} (2015) \emph{{World Population Prospects: The 2015 Revision,
+  Key Findings and Advance Tables. Working Paper No. ESA/P/WP.241.}}
+
+\bibitem[{Verhegghen \emph{et~al.}(2016)Verhegghen, Eva, Desclée \&
+  Achard}]{Verhegghen2016}
+Verhegghen, A., Eva, H., Desclée, B. \& Achard, F. (2016) {Review and
+  Combination of Recent Remote Sensing Based Products for Forest Cover Change
+  Assessments in Cameroon}.
+\newblock \emph{International Forestry Review}, \textbf{18}, 14--25.
+
+\bibitem[{Vieilledent \emph{et~al.}(2016)Vieilledent, Gardi, Grinand, Burren,
+  Andriamanjato, Camara, Gardner, Glass, Rasolohery, Rakoto~Ratsimba, Gond \&
+  Rakotoarijaona}]{Vieilledent2016}
+Vieilledent, G., Gardi, O., Grinand, C., Burren, C., Andriamanjato, M., Camara,
+  C., Gardner, C.J., Glass, L., Rasolohery, A., Rakoto~Ratsimba, H., Gond, V.
+  \& Rakotoarijaona, J.R. (2016) {Bioclimatic envelope models predict a
+  decrease in tropical forest carbon stocks with climate change in Madagascar}.
+\newblock \emph{Journal of Ecology}, \textbf{104}, 703--715.
+
+\bibitem[{Vieilledent \emph{et~al.}(2013)Vieilledent, Grinand \&
+  Vaudry}]{Vieilledent2013}
+Vieilledent, G., Grinand, C. \& Vaudry, R. (2013) {Forecasting deforestation
+  and carbon emissions in tropical developing countries facing demographic
+  expansion: a case study in Madagascar}.
+\newblock \emph{Ecology and Evolution}, \textbf{3}, 1702--1716.
+
+\bibitem[{Virah-Sawmy(2009)}]{Virah-Sawmy2009}
+Virah-Sawmy, M. (2009) {Ecosystem management in Madagascar during global
+  change}.
+\newblock \emph{Conservation Letters}, \textbf{2}, 163--170.
+
+\bibitem[{Vorontsova \emph{et~al.}(2016)Vorontsova, Besnard, Forest, Malakasi,
+  Moat, Clayton, Ficinski, Savva, Nanjarisoa, Razanatsoa, Randriatsara, Kimeu,
+  Luke, Kayombo \& Linder}]{Vorontsova2017}
+Vorontsova, M.S., Besnard, G., Forest, F., Malakasi, P., Moat, J., Clayton,
+  W.D., Ficinski, P., Savva, G.M., Nanjarisoa, O.P., Razanatsoa, J.,
+  Randriatsara, F.O., Kimeu, J.M., Luke, W.R.Q., Kayombo, C. \& Linder, H.P.
+  (2016) {Madagascar's grasses and grasslands: anthropogenic or natural?}
+\newblock \emph{Proceedings of the Royal Society of London B: Biological
+  Sciences}, \textbf{283}.
+
+\end{thebibliography}
 
@@
     \toprule
-    Year & Forest (ha) & Unmap (ha) & Annual defor. (ha/yr) & Rate
+    Year & Forest (ha) & Unmapped (ha) & Annual defor. (ha/yr) & Rate
     (\%/yr)\tabularnewline
@@
 
-  \caption{\textbf{Evolution of the forest cover and deforestation rates
-      from 1953 to 2014 in Madagascar}. Forest map for the year 1973 has 3.3
-    Mha of unclassified areas due to the presence of clouds on satellite
-    images. As a consequence, deforestation rates for the periods 1953-1973
-    and 1973-1990 are indicative. The two last columns indicate the
-    deforestation rates on the previous time-period (e.g.~1953-1973 for year
-    1973, 1973-1990 for year 1990, etc.).}
+  \caption{\textbf{Evolution of the forest cover and deforestation
+      rates from 1953 to 2014 in Madagascar}. Forest map for the year
+    1973 has 3.3 Mha of unclassified areas due to the presence of
+    clouds on satellite images. As a consequence, deforestation rates
+    for the periods 1953-1973 and 1973-1990 are indicative. The two
+    last columns indicate the annual deforested areas and annual
+    deforestation rates on the previous time-period (e.g.~1953-1973
+    for year 1973, 1973-1990 for year 1990, etc.).}
 
@@
   \caption{\textbf{Comparing our estimates of forest-cover (in ha) for
-      Madagascar with previous studies on the period 1953-2014}. We compared
-    our forest-cover with the estimates from three previous studies
-    \citep{Harper2007, MEFT2009, ONE2015}. We computed a The increase in
-    mangrove and spiny forest covers from \emph{c.} 1953 to \emph{c.} 1973
-    in \citet{Harper2007} and our study is most probably due to differences
-    in forest definition and mapping methods between the 1953
-    aerial-photography derived map and the 1973 Landsat image derived map.}
+      Madagascar with previous studies on the period 1953-2014}. We
+    compared our estimates of forest-cover with the estimates from
+    three previous studies \citep{Harper2007, MEFT2009, ONE2015}. We
+    obtained a Pearson's correlation coefficient of 0.99 between our
+    forest-cover estimates and forest-cover estimates from previous
+    studies. The increase in mangrove and spiny forest covers from
+    \emph{c.} 1953 to \emph{c.} 1973 in \citet{Harper2007} and our
+    study is most probably due to differences in forest definition and
+    mapping methods between the 1953 aerial-photography derived map
+    and the 1973 Landsat image derived map.}
 
@@
     provided. For deforestation rates in \%/yr, exact same numbers as
-    in scientific articles and reports from previous studies have been
-    reported. The way annual deforestation rates in \%/yr have been
-    computed in these previous studies can slightly differ from one
-    study to another but estimates always correct for the potential
-    presences of clouds on forest maps. Annual deforested areas in
-    ha/yr have been recomputed from forest-cover estimates in
-    Tab.~\ref{tab:comp_forest} (except for \citet{Harper2007} for the
-    periods 1973-1990 and 1990-2000 for which annual deforested ares
-    in ha/yr were derived from numbers reported in the original
-    publication, see methods) and do not correct for the potential
-    presence of clouds.}
+    in scientific articles and reports from previous studies
+    \citep{Harper2007, MEFT2009, ONE2015} have been reported. The way
+    annual deforestation rates in \%/yr have been computed in these
+    previous studies can slightly differ from one study to another, but
+    estimates always correct for the potential presences of clouds on
+    satellite images and unclassified areas on forest maps. Annual
+    deforested areas in ha/yr have been recomputed from forest-cover
+    estimates in Tab.~\ref{tab:comp_forest} (except for
+    \citet{Harper2007} for the periods 1973-1990 and 1990-2000 for
+    which annual deforested areas in ha/yr were derived from numbers
+    reported in the original publication, see methods) and do not
+    correct for the potential presence of clouds.}
 
@@
     ``undetermined'' \citep{Riitters2000}. We used a moving window of 7x7
-    pixels (4.4 ha). Using this window size, forest edge had a width of
-    about 90m. The ``interior'' category can be interpreted as the most
+    pixels (4.4~ha). Using this window size, forest edge had a width of
+    about 90~m. The ``interior'' category can be interpreted as the most
     intact forest. The ``patch'' and ``transitional'' categories correspond

### 2017-06-19 report/deforestmap.tex
Submission to Biological Conservation

--- 2017-06-08 latex/deforestmap.tex
@@
 \usepackage{lineno} %% Numérotation des lignes
-\usepackage{caption}
 %\linenumbers
+\usepackage{caption2}
 \newcommand{\logit}{\text{logit}}
@@
 
-\begin{center}
-  \textbf{Running headline:} \emph{Six decades of deforestation in Madagascar}
-\end{center}
+{\small
+   \begin{flushleft}
+   \textbf{Highlights:}\\
+    -We produced new 30 m resolution annual forest-cover maps for Madagascar for the period 2000-2014.\\
+   - Madagascar has lost 44\% of its natural forest-cover over the period 1953-2014.\\
+   - Half of the tropical forest in Madagascar is now located at less than 100m from forest edge.\\
+   - Annual deforestation rate has increased in Madagascar since 2005 to reach 1.08\%/yr.\\
+   - Conservation efforts must be intensified to save Madagascar unique forests and biodiversity.\\
+\end{flushleft}}
 
@@
 
+\doublespacing
 \linenumbers
-\doublespacing
 
@@
 
-  1. The island of Madagascar has an unparalleled biodiversity, mainly
+  The island of Madagascar has an unparalleled biodiversity, mainly
   located in the tropical forests of the island, which is highly
@@
   from past studies at national level with substantial gaps (due to
-  presence of cloud cover on satellite imagery) prevent the analyzis
-  of long-term deforestation trends in Madagascar.
-
-  2. In this study, we propose a new approach combining historical
+  presence of cloud cover on satellite imagery) prevent the analysis
+  of long-term deforestation trends in Madagascar. In this study, we propose 
+  a new approach combining historical
   (1953-2000) national forest-cover maps with recent (2001-2014)
@@
   the full territory of Madagascar for the year 1990, and annually
-  from 2000 to 2014.
-
-  3. We estimated that Madagascar has lost 44\% of its natural forest
+  from 2000 to 2014. We estimated that Madagascar has lost 44\% of its natural forest
   cover over the period 1953-2014 (including 37\% over the period
@@
   half of the forest (46\%) is now located at less than 100m from the
-  forest edge.
-
-  4. \emph{Policy implications}: Accurate forest-cover change maps can
+  forest edge. Accurate forest-cover change maps can
   be used to assess the effectiveness of past and current conservation
@@
 
-\keywords{biodiversity, climate-change, deforestation, Madagascar, tropical forest}
+\keywords{biodiversity, climate-change, deforestation, forest-fragmentation, Madagascar, tropical forest}
 
@@
 
-\bibitem[{Perrier~de La~B{\^a}thie(1921)}]{Perrier1921}
-Perrier~de La~B{\^a}thie, H. (1921) \emph{{La v{\'e}g{\'e}tation malgache}},
-  vol.~23.
-\newblock Mus{\'e}e Colonial.
-
 \bibitem[{{MEFT, USAID, and CI}(2009)}]{MEFT2009}
@@
 \newblock \emph{Nature}, \textbf{540}, 418--422.
+
+\bibitem[{Perrier~de La~B{\^a}thie(1921)}]{Perrier1921}
+Perrier~de La~B{\^a}thie, H. (1921) \emph{{La v{\'e}g{\'e}tation malgache}},
+  vol.~23.
+\newblock Mus{\'e}e Colonial.
 
@@
 
-    \includegraphics[width=11cm]{outputs/ecoregion.png}
+    \includegraphics[width=11cm]{ecoregion.png}
 
@@
 
-  \includegraphics[width=\textwidth]{outputs/fig_fcc_highres.png}
+  \includegraphics[width=\textwidth]{fig_fcc_highres.png}
 
@@
 
-  \includegraphics[width=13cm]{outputs/dist.png}
+  \includegraphics[width=13cm]{dist.png}
 

### 2017-12-15 deforestmap.tex
essd formatting

\documentclass[]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Combining global tree cover loss data with historical national forest-cover maps to look at six decades of deforestation and forest fragmentation in Madagascar},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{natbib}
\bibliographystyle{plainnat}
\usepackage{longtable,booktabs}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

%%% Change title format to be more compact
\usepackage{titling}

% Create subtitle command for use in maketitle
\newcommand{\subtitle}[1]{
  \posttitle{
    \begin{center}\large#1\end{center}
    }
}

\setlength{\droptitle}{-2em}
  \title{Combining global tree cover loss data with historical national
forest-cover maps to look at six decades of deforestation and forest
fragmentation in Madagascar}
  \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
  \author{}
  \preauthor{}\postauthor{}
  \date{}
  \predate{}\postdate{}

\begin{document}
\maketitle

\hypertarget{ghislain-vieilledent12-clovis-grinand3-fety-a.-rakotomalala3-rija-ranaivosoa4-jean-roger-rakotoarijaona4-thomas-f.-allnutt56-and-frederic-achard1}{%
\paragraph{Ghislain Vieilledent*,1,2, Clovis Grinand3, Fety A.
Rakotomalala3, Rija Ranaivosoa4, Jean-Roger Rakotoarijaona4, Thomas F.
Allnutt5,6, and Frédéric
Achard1}\label{ghislain-vieilledent12-clovis-grinand3-fety-a.-rakotomalala3-rija-ranaivosoa4-jean-roger-rakotoarijaona4-thomas-f.-allnutt56-and-frederic-achard1}}

(*) /Email:
\href{mailto:ghislain.vieilledent@cirad.fr}{\nolinkurl{ghislain.vieilledent@cirad.fr}},
/Phone: +39.329.457.2273

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Joint Research Center of the European Commission, Bio-economy Unit
  (JRC.D.1), I-21027 Ispra (VA), ITALY
\item
  Cirad, UPR Forêts et Sociétés, F-34398 Montpellier, FRANCE
\item
  ETC Terra, F-75016 Paris, FRANCE
\item
  Office National pour l'Environnement, 101 Antananarivo, MADAGASCAR
\item
  Wildlife Conservation Society, 101 Antananarivo, MADAGASCAR
\item
  GreenInfo Network, Oakland, California, USA
\end{enumerate}

\emph{Running headline:} Six decades of deforestation in Madagascar

\hypertarget{highlights}{%
\subsection{Highlights}\label{highlights}}

\begin{itemize}
\tightlist
\item
  We produced new 30m-resolution annual forest-cover maps for Madagascar
  for the period 2000-2014.
\item
  Madagascar has lost 44\% of its natural forest-cover over the period
  1953-2014.
\item
  Half of the tropical forest in Madagascar is now located at less than
  100m from forest edge.
\item
  Annual deforestation rate has increased in Madagascar since 2005 to
  reach 1.08\%/yr.
\item
  Conservation and development efforts must be intensified to save
  Madagascar forest and biodiversity.
\end{itemize}

\hypertarget{summary}{%
\subsection{Summary}\label{summary}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  The island of Madagascar has an unparalleled biodiversity, mainly
  located in the tropical forests of the island, which is highly
  threatened by anthropogenic deforestation. Scattered forest maps from
  past studies at national level with substantial gaps (due to presence
  of cloud cover on satellite imagery) prevent the analyzis of long-term
  deforestation trends in Madagascar.
\item
  In this study, we propose a new approach combining historical
  (1953-2000) national forest-cover maps with recent (2000-2014) global
  annual tree cover loss data to look at six decades (1953-2014) of
  deforestation and forest fragmentation in Madagascar. We produced new
  forest-cover maps at 30m-resolution over the full territory of
  Madagascar for the year 1990, and annually from 2000 to 2014.
\item
  We estimated that Madagascar has lost 44\% of its natural forest cover
  over the period 1953-2014 (including 37\% over the period 1973-2014).
  Natural forests cover 8.9 Mha in 2014 (15\% of the national territory)
  which are divided into 4.4 Mha (50\%) of moist forests, 2.6 Mha (29\%)
  of dry forests, 1.7 Mha of spiny forests (19\%) and 177,000 ha (2\%)
  of mangroves. Since 2005, the annual deforestation rate has
  progressively increased in Madagascar to reach 99,000 ha/yr during
  2010-2014 (corresponding to a rate of 1.08\%/yr). Around half of the
  forest (46\%) is now located at less than 100m from the forest edge.
\item
  \emph{Policy implications}: Increase in the deforestation rate since
  2005 can be related to rapid population growth (close to 3\%/yr) and
  to poor law enforcement due to political instability in the country.
  More effort, including conservation and development programs, is
  needed in Madagascar to protect the remaining natural tropical
  forests, both to enhance local people's livelihoods and to support
  global environmental efforts.
\end{enumerate}

\textbf{Keywords}: biodiversity, climate-change, deforestation, Landsat,
Madagascar, tropical forest.

\textbf{Potential journals} (2015 IF): Conservation Letters (7.126),
Journal of Applied Ecology (5.196), Conservation Biology (4.257),
Environmental Research Letters (4.134), Biological Conservation (3.985),
Landscape Ecology (3.657), Ambio (2.555), Environmental Conservation
(2.235), Biotropica (1.944), Tropical Conservation Science (1.55).

\hypertarget{introduction}{%
\subsection{1. Introduction}\label{introduction}}

Separated from the African continent and the Indian plate about 165 and
88 million years ago respectively \citep{Ali2008}, the flora and fauna
of Madagascar followed its own evolutionary path. Isolation combined
with a high number of micro-habitats \citep{Pearson2009} has led to
Madagascar's exceptional biodiversity both in term of number of species
and endemism in many taxonomic groups \citep{Crottini2012, Goodman2005}.
Most of the biodiversity in Madagascar is concentrated in the tropical
forests of the island which can be divided into three types: the moist
forest in the East, the dry forest in the West and the spiny forest in
the South \citep{Vieilledent2016}. This unparalleled biodiversity is
severely threatened by anthropogenic deforestation
\citep{Harper2007, Vieilledent2013} associated with activities such as
slash-and-burn agriculture and pasture \citep{Scales2011}. Tropical
forests in Madagascar also store a large amount of carbon
\citep{Vieilledent2016} and high rates of deforestation in Madagascar
are responsible for large CO2 emissions in the atmosphere
\citep{Achard2014}. Deforestation threatens species survival by directly
reducing their available habitat \citep{Brooks2002, Tidd2001}. Forest
fragmentation can also lead to species extinction by isolating
populations from each other and creating forest patches too small to
maintain viable populations \citep{Saunders1991}. Fragmentation also
increases forest edge where ecological conditions (such as air
temperature, light intensity and air moisture) can be dramatically
modified, with consequences on the abundance and distribution of species
\citep{Murcia1995}. Forest fragmentation can also have substantial
effects on forest carbon storage capacity, as carbon stocks are much
lower at the forest edge than under a closed canopy \citep{Brinck2017}.
Moreover, forest carbon stocks vary spatially due to climate or soil
factors \citep{Saatchi2011, Vieilledent2016}. As a consequence, accurate
and spatially explicit maps of forest-cover and forest-cover change are
necessary to monitor biodiversity loss and carbon emissions from
deforestation and forest fragmentation, assess the efficiency of present
conservation strategies \citep{Eklund2016}, and implement new strategies
for the future \citep{Vieilledent2013, Vieilledent2016}. Simple
time-series of forest-cover estimates, such as those provided by the FAO
Forest Resource Assessment report \citep{Keenan2015} are not sufficient.

Unfortunately, accurate and exhaustive forest-cover maps are not
available for Madagascar for the last fifteen years (2000-2015).
\citet{Harper2007} produced maps of forest cover and forest cover
changes over Madagascar for the years c. 1953, c. 1973, 1990 and 2000.
The c. 1953 forest map was derived from the visual interpretation of
aerial photography at coarse scale (1/1,000,000). Forest maps for the
years c. 1973, 1990 and 2000, were obtained from supervised
classification of Landsat satellite images at 60m resolution (for the
year 1973) or 30m resolution (for years 1990 and 2000) and can be used
to derive accurate estimates (89.5\% accuracy reported for the forest /
non-forest map of year 2000). Maps provided by \citet{Harper2007} were
not exhaustive (due to the presence of clouds in the satellite imagery),
e.g.~11 244 km2 are mapped as unknown cover type for the year 2000.
Using a similar supervised classification approach as in
\citet{Harper2007}, more recent maps have been produced for the periods
2000-2005-2010 by national institutions, with the technical support of
international environmental NGOs \citep{MEFT2009, ONE2013}. Another set
of recent forest-cover maps using an advanced statistical tool for
classification, the Random Forest classifier
\citep{Grinand2013, Rakotomalala2015}, was produced for the periods
2005-2010-2013 \citep{ONE2015}. However, these maps are either too old
to give recent estimates of deforestation \citep{MEFT2009, ONE2013},
include large areas of missing information (due to images with more than
20\% of clouds) for the 2010 map provided by \citep{ONE2013}, or show
large mis-classification in specific areas, especially in the dry and
spiny forest domain for which the spectral answer has a strong seasonal
behavior due to the deciduousness of such forests (overall accuracy is
lower than 0.8 for the dry and spiny forests for the maps produced by
\citet{ONE2015}). Moreover, the production of such forest maps from a
supervised classification approach requires significant resources,
especially regarding the image selection step (required to minimize
cloud cover) and the training step (visual interpretation of a large
number of polygons needed to train the classification algorithm)
\citep{Rakotomalala2015}. Most of this work of image selection and
visual interpretation would need to be repeated to produce new forest
maps in the future using a similar approach.

Global forest or tree cover products have also been published recently
and can be tested at the national scale for Madagascar. \citep{Kim2014}
produced a global forest-cover change map from 1990 to 2000 (derived
from Landsat imagery). This product was updated to cover the period
1975-2005 (\url{http://glcf.umd.edu/data/landsatFCC/}) but forest-cover
maps after 2005 were not produced. Moreover, the approach used in
\citet{Kim2014} did not accurately map the forests in the dry and spiny
ecosystems of Madagascar (see Fig. 8 in \citet{Kim2014}).
\citet{Hansen2013} mapped tree cover percentage, annual forest loss and
forest gain from 2000 to 2012 at global scale at 30 m resolution. This
product has since been updated and is now available up to the year 2014
\citep{Hansen2013}. To map forest cover from the \citet{Hansen2013}
product, a tree cover threshold must be selected (that defines forest
cover). Selecting such a threshold is not straightforward as the
accuracy of the global tree cover map strongly varies between forest
types, and is substantially lower for dry forests than for moist forests
\citep{Bastin2017}. Moreover, the \citet{Hansen2013} product does not
provide information on land-use. In particular the global tree cover map
does not separate tree plantations such as oil palm or eucalyptus
plantations from natural forests \citep{Tropek2014}. Thus, the global
tree cover map from \citet{Hansen2013} cannot be used alone to produce a
map of forest cover \citep{Tyukavina2017}. In complement to the tree
cover percentage provided in \citet{Hansen2013}, a layer of annual tree
cover loss is also provided (i.e.~complete loss of tree cover from a
value higher than 10\% to zero) for the period 2001-2014.

In this study, we present a simple approach which combines the maps from
\citet{Harper2007} and products from \citet{Hansen2013} to derive annual
wall-to-wall forest-cover change maps over the period 2000-2014 for
Madagascar. We use the forest-cover map provided by \citet{Harper2007}
for the year 2000 (defining the land-use) with the tree cover loss
product provided by \citet{Hansen2013} that we apply only inside forest
areas identified by \citet{Harper2007}. Similar to the approach of
\citet{Harper2007}, we also assess trends in deforestation rates and
forest fragmentation from c. 1953 to 2014. The approach described in
this study can help assess the effectiveness of current conservation
strategies, and assist the implementation of future strategies. Our
approach could be easily extended to other tropical countries that have
at least one forest-cover map between 2000 and 2014. This approach can
easily be repeated in the future when the \citet{Hansen2013} products
are updated.

\hypertarget{materials-and-methods}{%
\subsection{2. Materials and Methods}\label{materials-and-methods}}

\hypertarget{creation-of-new-forest-cover-maps-of-madagascar-from-1953-to-2014}{%
\subsubsection{2.1 Creation of new forest-cover maps of Madagascar from
1953 to
2014}\label{creation-of-new-forest-cover-maps-of-madagascar-from-1953-to-2014}}

We produced annual forest/non-forest maps at 30m resolution for the full
territory of Madagascar for the period 2000-2014 by combining the forest
map of year 2000 from \citet{Harper2007}, and the tree cover percentage
and annual forest cover loss maps over the period 2000-2014 from
\citet{Hansen2013}. The 2000 Harper's forest map includes 208,000 ha of
unclassified areas due to the presence of clouds in input satellite
imagery, mostly (88\%) within the moist forest domain which covers 4.17
million ha in total. To provide a label (forest or non-forest) to these
unclassified pixels, we use the 2000 tree cover percentage map of
\citet{Hansen2013} by selecting a threshold of 75\% tree cover to define
forest cover as recommended by other studies for the moist domain
\citep{Achard2014}. We thus obtained a forest-cover map for the year
2000 covering the full territory of Madagascar. We then combined this
forest-cover map of the year 2000 with the annual forest cover loss maps
from 2001 to 2014 provided by \citet{Hansen2013} to create annual
forest-cover maps from 2001 to 2014 at 30m resolution. We also completed
the Harper's forest map of year 1990 by filling unclassified areas (due
to the presence of clouds on satellite images) using our forest-cover
map of year 2000. To do so, we assumed that if forest was present in
2000, the pixel was also forested in 1990. The remaining unclassified
pixels were limited to a relatively small total area of c. 8,000 ha. We
labeled these residual pixels as non-forest as for the year 2000.
Similarly we completed the Harper forest map of year 1973 by filling
unclassified areas using our forest-cover map of the year 1990 assuming
that if forest was present in 1990, it was also present in 1973.
Contrary to the year 1990, the remaining unclassified pixels for year
1973 corresponded to a significant total area of 3.32 million ha. We
also reprojected the forest-cover map of year 1953 to a common
projection in order to compare to the other eras. This map was produced
by scanning a paper map derived from aerial photos, and thus could not
be perfectly aligned with the other maps produced through digital
processing of satellite imagery \citep{Harper2007}. Finally for all
forest-cover maps from 1973, the isolated single non-forest pixels
(i.e.~fully surrounded by forest pixels) were removed, assuming that
single non-forest pixels inside a forest patch were not corresponding to
deforestation (they might correspond to selective logging activities).
This allowed us to avoid counting very small scale events
(\textless{}0.1 ha such as selective logging) as forest fragmentation.
All the resulting maps are freely available at
\url{https://bioscenemada.cirad.fr/forestmaps}.

\hypertarget{computing-forest-cover-areas-and-deforestation-rates}{%
\subsubsection{2.2 Computing forest-cover areas and deforestation
rates}\label{computing-forest-cover-areas-and-deforestation-rates}}

From these new forest-cover maps, we calculated the total forest-cover
area for seven available years (1953-1973-1990-2000-2010-2005-2014), and
the annual deforested area and annual deforestation rate for the
corresponding six time periods between 1953 and 2014. The annual
deforestation rates were calculated as follows
\citep{Puyravaud2003, Vieilledent2013}:

\[\theta = 100 \times [1-(1-(F_{t_2}-F_{t_1})/F_{t_1})^{(1/(t_2-t_1))}]\]

where \(\theta\) is the annual deforestation rate (in \%/yr),
\(F_{t_2}\) and \(F_{t_1}\) are the forest cover free of clouds at both
dates \(t_2\) and \(t_1\), and \(t_2-t_1\) is the time-interval (in
years) between the two dates.

Because of the large unclassified area (3.32 million ha) in 1973, the
annual deforestation areas and rates for the two periods 1953-1973 and
1973-1990 are only indicative estimates. For these two periods the
annual deforestation rates are computed as the ratio
\((F_{t_2}-F_{t_1})/F_{t_1}\) considering only the mapped forest pixels.
Area and rate estimates are produced at the national scale and for the
four forest ecosystems present in Madagascar: moist forest in the East,
dry forest in the West, spiny forest in the South, and mangroves on the
Western coast (Fig. 1). To define the forest domains, we used a map from
the MEFT (\emph{``Ministère de l'Environnement et des Forêts à
Madagascar''}) with the boundaries of the four ecoregions in Madagascar.
Ecoregions were defined on the basis of climatic and vegetation criteria
using the climate classification by \citet{Cornet1974} and the
vegetation classification from the 1996 IEFN national forest inventory
\citep{IEFN1996}. Because mangrove forests are highly dynamic ecosystems
that can expand or contract on decadal scales depending on changes in
environmental factors \citep{Armitage2015}, a fixed delimitation of the
mangrove ecoregion on six decades might not be fully appropriate. As a
consequence, our estimates of the forest-cover and deforestation rates
for mangroves in Madagascar must be considered with this limitation.

\hypertarget{comparing-our-forest-cover-and-deforestation-rate-estimates-with-previous-studies}{%
\subsubsection{2.3 Comparing our forest-cover and deforestation rate
estimates with previous
studies}\label{comparing-our-forest-cover-and-deforestation-rate-estimates-with-previous-studies}}

We compared our estimates of forest-cover and deforestation rates with
estimates from the three existing studies at the national scale for
Madagascar: (i) \citep{Harper2007}, (ii) \citep{MEFT2009} and (iii)
\citep{ONE2015}. \citet{Harper2007} provides forest-cover and
deforestation estimates for the periods c. 1953-c. 1973-1990-2000. MEFT,
USAID, and CI (2009) provides estimates for the periods 1990-2000-2005
and ONE, DGF, MNP, WCS, and Etc Terra (2015) provides estimates for the
periods 2005-2010-2013. To compare our forest-cover and deforestation
estimates over the same time periods, we consider an additional
time-period in our study (2010-2013) by creating an extra forest-cover
map for the year 2013. We computed the Pearson's correlation coefficient
and the root mean square error (RMSE) between our forest-cover estimates
and forest-cover estimates from previous studies for all the dates and
forest types (including also the total forest cover estimates). For
previous studies, the computation of annual deforestation rates (in
\%/yr) is not always detailed and might slightly differ from one study
to another \citep[see][]{Puyravaud2003}. \citet{Harper2007} also provide
total deforested areas for the two periods 1973-1990 and 1990-2000. We
converted these values into annual deforested area estimates. When
annual deforested areas were not reported (for 1953-1973 in
\citet{Harper2007} and in \citet{MEFT2009} and \citet{ONE2015}), we
computed them from the forest-cover estimates in each study. These
estimates cannot be corrected from the potential bias due to the
presence of residual clouds. Forest-cover and deforestation rates were
then compared between all studies for the whole of Madagascar and the
four ecoregions. The same ecoregion boundaries as in our study were used
in \citet{ONE2015} but this was not the case for \citet{Harper2007} and
\citet{MEFT2009}, which can explain part of the differences between the
estimates.

\hypertarget{fragmentation}{%
\subsubsection{2.3 Fragmentation}\label{fragmentation}}

We also conducted an analysis of changes in forest fragmentation for the
years 1953, 1973, 1990, 2000, 2010 and 2014. We applied the method
developed by \citet{Riitters2000} which uses a moving window to
characterize the fragmentation around each forested pixel. Computations
were done using the function \texttt{r.forestfrag} of the GRASS GIS
software \citep{Neteler2008}. Six categories of fragmentation were
identified from the amount of forest and its occurrence as adjacent
forest pixels: ``interior'', ``perforated'', ``edge'', ``transitional'',
``patch'', and ``undetermined''. We used a moving window of 7x7 pixels
(4.4 ha). Using this window size, forest edge had a width of about 90m
\citep{Riitters2000}. The ``interior'' category can be interpreted as
the most intact forest \citep{Potapov2017}. The ``patch'' and
``transitional'' categories correspond to isolated small forest patches.
We reported the area of forest in each fragmentation category for the
six years and analyzed the dynamics of fragmentation over the six
decades. We also computed the distance to forest edge for all forest
pixels for the years 1953, 1973, 1990, 2000, 2005, 2010 and 2014. For
that, we used the function \texttt{gdal\_proximity.py} of the GDAL
software (\url{http://www.gdal.org/}). We computed the mean and 95\%
quantiles of the distance to forest edge and looked at the evolution of
these values with time.

\hypertarget{results}{%
\subsection{3. Results}\label{results}}

\hypertarget{dynamics-of-forest-cover-and-deforestation-intensity}{%
\subsubsection{3.1 Dynamics of forest cover and deforestation
intensity}\label{dynamics-of-forest-cover-and-deforestation-intensity}}

Natural forests in Madagascar covered 16.0 Mha in 1953, about 27\% of
the national territory of 587,041 km2. In 2014, the forest cover dropped
to 8.9 Mha, corresponding to about 15\% of the national territory (Fig.
2 and Tab. 1). Madagascar has lost 44\% and 37\% of its natural forests
between 1953 and 2013, and between c. 1973 and 2014 respectively (Fig. 2
and Tab. 1). In 2014 the remaining 8.9 Mha of natural forest were
distributed as: 4.4 Mha of moist forest (50\% of total forest cover),
2.6 Mha of dry forest (29\%), 1.7 Mha of spiny forest (19\%) and 0.18
Mha (2\%) of mangrove forest (Fig. 1 and Tab. 2). Regarding the
deforestation trend, we observed a progressive decrease of the
deforestation rate after1990 from 205,000 ha/yr (1.63\%/yr) over the
period 1973-1990 to 44,300 ha/yr (0.43\%/yr) over the period 2000-2005
(Tab. 1). Then from 2005, the deforestation rate has progressively
increased and has more than doubled over the period 2010-2014 (98,700
ha/yr, 1.08\%/yr) compared to 2000-2005 (Tab. 1). The deforestation
trend characterized by a progressive decrease of the deforestation rate
over the period 1990-2005 and a progressive increase of the
deforestation after 2005 is valid for all four ecoregions (Tab. 3), with
the exception of the spiny forest domain for which the deforestation
rate during the period 2010-2013 was lower than during 2005-2010 (Tab.
3).

\hypertarget{comparison-with-previous-forest-cover-change-studies-in-madagascar}{%
\subsubsection{3.2 Comparison with previous forest-cover change studies
in
Madagascar}\label{comparison-with-previous-forest-cover-change-studies-in-madagascar}}

Forest-cover maps provided by previous studies over Madagascar were not
exhaustive (unclassified areas) due to the presence of clouds in
satellite imagery used to produce such maps. In \citet{Harper2007}, the
maps of years 1990 and 2000 include 0.5 and 1.12 Mha of unknown cover
type respectively. Proportions of unclassified areas are not reported in
the two other existing studies by \citet{MEFT2009} and \citet{ONE2015}.
With our approach, we produced wall to wall forest-cover change maps
from 1990 to 2014 for the full territory of Madagascar (Tab. 1). This
allowed us to produce more robust estimates of forest-cover and
deforestation rates over this period. Our forest-cover estimates over
the period 1953-2013 (considering forest cover estimates at national
level and by ecoregions for the 6 available dates) were well correlated
(0.99) to estimates from the three previous studies (Tab. 2) with a RMSE
of 300,000 ha (6\% of the mean forest cover of 4.8 Mha when considering
all dates and forest types together). These small differences can be
partly attributed to differences in ecoregion boundaries. Despite
significant differences in deforestation estimates (Tab. 3), a similar
deforestation trend was observed across studies with a decrease of
deforestation rates over the period 1990-2005, followed by a progressive
increase of the deforestation after 2005.

\hypertarget{evolution-of-forest-fragmentation-with-time}{%
\subsubsection{3.3 Evolution of forest fragmentation with
time}\label{evolution-of-forest-fragmentation-with-time}}

In parallel to the dynamics of deforestation, forest fragmentation has
progressively increased since 1953 in Madagascar. We observed a
continuous decrease of the mean distance to forest edge from 1953 to
2014 in Madagascar. The mean distance to forest edge has decreased to c.
300 m in 2014 while it was previously c. 1.5 km in 1973 (Fig. 3).
Moreover, a large proportion (73\%) of the forest was located at a
distance greater than 100 m in 1973, while almost half of the forest
(46\%) was at a distance lower than 100 m from forest edge in 2014 (Fig.
3). The percentage of forest that can be considered intact in Madagascar
has continuously decreased since 1953. The percentage of forest
belonging to the ``interior'' category (most intact forests) has fallen
from 68\% in 1973 to 50\% in 2014. In 2014, more than 16\% of the forest
belonged to the ``patch'' and ``transitional'' categories (isolated
small forest patches) compared to 9.5\% in 1973 (Tab. 4).

\hypertarget{discussion}{%
\subsection{4. Discussion}\label{discussion}}

\hypertarget{benefits-of-the-combined-use-of-recent-global-annual-tree-cover-loss-data-with-historical-national-forest-cover-maps}{%
\subsubsection{4.1 Benefits of the combined use of recent global annual
tree cover loss data with historical national forest-cover
maps}\label{benefits-of-the-combined-use-of-recent-global-annual-tree-cover-loss-data-with-historical-national-forest-cover-maps}}

In this study, we combined recent (2001-2014) global annual tree cover
loss data \citep{Hansen2013} with historical (1953-2000) national
forest-cover maps \citep{Harper2007} to look at six decades (1953-2014)
of deforestation and forest fragmentation in Madagascar. We produced
annual forest-cover maps at 30m-resolution covering Madagascar for the
period 2000 to 2014. Our study extends the forest-cover monitoring on a
six decades period (from 1953 to 2014) while harmonizing the data from
the previous studies. We propose a generic approach to solve the problem
of forest definition which is needed to transform the 2000 global tree
cover dataset from Hansen et al. (2013) into a forest/non-forest map
\citep{Tropek2014}. We propose to use a historical national forest-cover
map, based on a national forest definition, as a forest cover mask. This
approach could be easily extended to other regions or countries for
which an accurate forest-cover map is available at any date within the
period 2000-2014, but preferably at the beginning of the period to
profit from the full record and derive long-term estimates of
deforestation. Moreover, this approach can be repeated in the future if
and when the global tree cover product is updated. We have made the
R/GRASS code used for this study freely available in a GitHub repository
(see Data availability statement) to facilitate application to other
study areas.

The accuracy of the derived forest-cover change maps depends directly on
the accuracies of the historical forest-cover maps and the tree cover
loss dataset. The reported global accuracy of the tree cover loss
dataset is 99.6\% (see Tab. S5 in \citet{Hansen2013}).
\citet{Verhegghen2016} have compared deforestation estimates derived
from the global tree cover loss dataset \citep{Hansen2013} with results
derived from semi-automated supervised classification of Landsat
satellite images \citep{Achard2014} for six countries in Central Africa
and they found a good agreement between these two sets of estimates.
Consistent with \citet{Harper2007}, we did not consider potential forest
regrowth in Madagascar (although \citet{Hansen2013} provided a tree
cover gains layer for the period 2001-2013) for the following reasons:
(i) the tree gain layer of \citet{Hansen2013} includes and catches more
easily tree plantations than natural forest regrowth \citep{Tropek2014};
(ii) there is little evidence of natural forest regeneration in
Madagascar \citep[\citet{Grouzis2001}]{Harper2007} - this can be
explained by several ecological processes following burning practice
such as soil erosion \citep{Grinand2017} and reduced seed bank due to
fire and soil loss \citep{Grouzis2001}, (iii) in areas where forest
regeneration is ecologically possible, young forest regrowth are more
easily re-burnt for agriculture and pasture; and (iv) young secondary
forests provide more limited ecosystem services compared to old-growth
natural forests in terms of biodiversity and carbon storage.

\hypertarget{dynamics-of-forest-cover-in-madagascar-from-1953-to-2014}{%
\subsubsection{4.2 Dynamics of forest-cover in Madagascar from 1953 to
2014}\label{dynamics-of-forest-cover-in-madagascar-from-1953-to-2014}}

We estimated that natural forests in Madagascar cover 8.9 Mha in 2014
(corresponding to 15\% of the country) and that Madagascar has lost 44\%
of its natural forest since 1953 (37\% since 1973). There is ongoing
scientific debate about the extent of the ``original'' forest cover in
Madagascar, and the extent to which humans have altered the natural
forest landscapes since their large-scale settlement around 800 CE
\citep{Burns2016, Cox2012}. Early French naturalists stated that the
full island was originally covered by forest
\citep{Humbert1927, Perrier1921}, leading to the common statement that
90\% of the natural forests have disappeared since the arrival of humans
on the island \citep{Kull2000}. More recent studies counter-balanced
that point of view saying that extensive areas of grassland existed in
Madagascar long before human arrival and were determined by climate,
natural grazing and other natural factors
\citep[\citet{Vorontsova2017}]{Virah-Sawmy2009}. Other authors have
questioned the entire narrative of extensive alteration of the landscape
by early human activity which, through legislation, has severe
consequences on local people \citep{Klein2002, Kull2000}. Whatever the
original proportion of natural forests and grasslands in Madagascar, our
results demonstrate that human activities since the 1950s have
profoundly impacted the natural tropical forests and that recent
conservation and development programs in Madagascar have failed to stop
deforestation in the recent years. Deforestation has strong consequences
on biodiversity and carbon emissions in Madagascar. Around 90\% of
Madagascar's species are forest dependent
\citep{Allnutt2008, Goodman2005} and \citet{Allnutt2008} estimated that
deforestation between 1953 and 2000 led to an extinction of 9\% of the
species. The additional deforestation we observed over the period
2000-2014 (around 1Mha of natural forest) worsen this result. Regarding
carbon emissions, using the 2010 aboveground forest carbon map by
\citet{Vieilledent2016}, we estimated that deforestation on the period
2010-2014 has led to 40.2 Mt C of carbon emissions in the atmosphere (10
Mt C /yr) and that the remaining aboveground forest carbon stock in 2014
is 832.8 Mt C. Associated to deforestation, we showed that the remaining
forests of Madagascar are highly fragmented with 46\% of the forest
being at less than 100m of the forest edge. Small forest fragments do
not allow to maintain viable populations and `edge effects' at
forest/non-forest interfaces have impacts on both carbon emissions
\citep{Brinck2017} and biodiversity loss \citep{Gibson2013, Murcia1995}.

\hypertarget{deforestation-trend-and-causes-of-the-acceleration-of-the-deforestation-since-2005}{%
\subsubsection{4.3 Deforestation trend and causes of the acceleration of
the deforestation since
2005}\label{deforestation-trend-and-causes-of-the-acceleration-of-the-deforestation-since-2005}}

In our study, we have shown that the progressive decrease of the
deforestation rate on the period 1990-2005 was followed by a continuous
increase in the deforestation rate on the period 2005-2014. In
particular, we showed that deforestation rate has more than doubled on
the period 2010-2014 compared to 2000-2005. Our results are confirmed by
previous studies \citep{Harper2007, MEFT2009, ONE2015} despite
differences in the methodologies regarding (i) forest definition
(associated to independent visual interpretations of observation
polygons to train the classifier), (ii) classification algorithms, (iii)
deforestation rate computation method, and (iv) correction for the
presence of clouds. Our deforestation rate estimates from 1990 to 2014
have been computed from wall to wall maps at 30 m resolution and can be
considered more accurate in comparison with estimates from these
previous studies. Our forest-cover and deforestation rate estimates can
be used as source of information for the next FAO Forest Resources
Assessment project \citep{Keenan2015}. Current rates of deforestation
can also be used to build reference scenarios for deforestation in
Madagascar and contribute to the implementation of deforestation
mitigation activities in the framework of REDD+ \citep{Olander2008}.

The increase of deforestation rates after 2005 can be explained by
population growth and political instability in the country. Nearly 90\%
of Madagascar's population relies on biomass for their daily energy
needs \citep{Minten2013} and the link between population size and
deforestation has previously been demonstrated in Madagascar
\citep{Vieilledent2013, Gorenflo2011}. With a mean demographic growth
rate of about 2.8\%/yr and a population which has increased from 16 to
24 million people on the period 2000-2015 \citep{UN2015}, the increasing
demand in wood-fuel and space for agriculture is likely to explain the
increase in deforestation rates. The political crisis of 2009
\citep{Ploch2012}, followed by several years of political instability
and weak governance could also explain the increase in the deforestation
rate observed on the period 2005-2014 (Smith et al., 2003). These
results show that despite the conservation policy in Madagascar
\citep{Freudenberger2010}, deforestation has dramatically increased at
the national level since 2005. More efforts should be done to identify
the real causes and mechanisms explaining this rampant deforestation in
order to implement new conservation strategies that will be effective at
conserving the few remaining natural tropical forests in Madagascar.

\hypertarget{authors-contribution}{%
\subsection{Author's contribution}\label{authors-contribution}}

All the authors conceived the study; GV made the analysis and drafted
the manuscript. All the authors revised the manuscript.

All authors conceived the ideas and designed methodology; GV analysed
the data and wrote the R/GRASS script; GV drafted the manuscript. All
authors contributed critically to the drafts and gave final approval for
publication.

\hypertarget{acknowledgements}{%
\subsection{Acknowledgements}\label{acknowledgements}}

This study is part of the Cirad's BioSceneMada project
(\url{https://bioscenemada.cirad.fr}) and the Joint Research Center's
ReCaREDD project (\url{http://forobs.jrc.ec.europa.eu/recaredd}). The
BioSceneMada project is funded by FRB (Fondation pour la Recherche sur
la Biodiversité) and the FFEM (Fond Français pour l'Environnement
Mondial) under the project agreement AAP-SCEN-2013 I. The ReCaREDD
project is funded by the European Commission.

\hypertarget{data-accessibility}{%
\subsection{Data accessibility}\label{data-accessibility}}

All the data and codes used for this study are made publicly available
in the \texttt{deforestmap} GitHub repository
(\url{https://github.com/ghislainv/deforestmap.git}). The results and
the manuscript are fully reproducible running the R script
\texttt{deforestmap.R} located inside the \texttt{deforestmap}
repository.

\hypertarget{tables}{%
\subsection{Tables}\label{tables}}

\begin{longtable}[]{@{}lrrrr@{}}
\toprule
Year & Forest (Kha) & Unmap (Kha) & Annual defor. (Kha/yr) & Rate
(\%/yr)\tabularnewline
\midrule
\endhead
1953 & 15,968 & 0 & - & -\tabularnewline
1973 & 14,243 & 3,317 & 86 & 0.6\tabularnewline
1990 & 10,762 & 0 & 205 & 1.6\tabularnewline
2000 & 9,879 & 0 & 88 & 0.8\tabularnewline
2005 & 9,668 & 0 & 42 & 0.4\tabularnewline
2010 & 9,320 & 0 & 70 & 0.7\tabularnewline
2014 & 8,925 & 0 & 99 & 1.1\tabularnewline
\bottomrule
\end{longtable}

Table 1: \textbf{Evolution of the forest cover and deforestation rates
from 1953 to 2014 in Madagascar}. Forest map for the year 1973 has 3.3
Mha of unclassified areas due to the presence of clouds on satellite
images. As a consequence, deforestation rates for the periods 1953-1973
and 1973-1990 are indicative. The two last columns indicate the
deforestation rates on the previous time-period (e.g.~1953-1973 for year
1973, 1973-1990 for year 1990, etc.).

\begin{longtable}[]{@{}llrrrrrrrr@{}}
\toprule
Forest type & Source & 1953 & 1973 & 1990 & 2000 & 2005 & 2010 & 2013 &
2014\tabularnewline
\midrule
\endhead
Total & Harper2007 & 15,996 & 14,173 & 10,606 & 8,982 & - & - & - &
-\tabularnewline
& MEFT2009 & - & - & 10,650 & 9,678 & 9,413 & - & - & -\tabularnewline
& ONE2015 & - & - & - & - & 9,451 & 8,977 & 8,486 & -\tabularnewline
& this study & 15,968 & 14,243 & 10,762 & 9,879 & 9,668 & 9,320 & 9,051
& 8,925\tabularnewline
Moist & Harper2007 & 8,766 & 6,876 & 5,234 & 4,167 & - & - & - &
-\tabularnewline
& MEFT2009 & - & - & 5,271 & 4,788 & 4,700 & - & - & -\tabularnewline
& ONE2015 & - & - & - & - & 4,556 & 4,457 & 4,345 & -\tabularnewline
& this study & 8,578 & 6,990 & 5,270 & 4,872 & 4,768 & 4,633 & 4,470 &
4,410\tabularnewline
Dry & Harper2007 & 4,252 & 4,028 & 2,712 & 2,457 & - & - & - &
-\tabularnewline
& MEFT2009 & - & - & 3,321 & 3,085 & 3,028 & - & - & -\tabularnewline
& ONE2015 & - & - & - & - & 3,223 & 2,970 & 2,679 & -\tabularnewline
& this study & 4,762 & 4,435 & 3,225 & 2,941 & 2,881 & 2,735 & 2,642 &
2,596\tabularnewline
Spiny & Harper2007 & 2,978 & 3,030 & 2,420 & 2,132 & - & - & - &
-\tabularnewline
& MEFT2009 & - & - & 2,124 & 1,872 & 1,757 & - & - & -\tabularnewline
& ONE2015 & - & - & - & - & 1,682 & 1,559 & 1,467 & -\tabularnewline
& this study & 2,463 & 2,583 & 2,055 & 1,858 & 1,811 & 1,744 & 1,731 &
1,713\tabularnewline
Mangroves & Harper2007 & - & - & 240 & 226 & - & - & - &
-\tabularnewline
& MEFT2009 & - & - & - & - & - & - & - & -\tabularnewline
& ONE2015 & - & - & - & - & 174 & 171 & 170 & -\tabularnewline
& this study & 143 & 200 & 181 & 178 & 177 & 177 & 177 &
177\tabularnewline
\bottomrule
\end{longtable}

Table 2: \textbf{Comparing our estimates of forest-cover (in ha) for
Madagascar with previous studies on the period 1953-2014}. We compared
our forest-cover with the estimates from three previous studies
\citep{Harper2007, MEFT2009, ONE2015}. We computed a The increase in
mangrove and spiny forest covers from \emph{c.} 1953 to \emph{c.} 1973
in \citet{Harper2007} and our study is most probably due to differences
in forest definition and mapping methods between the 1953
aerial-photography derived map and the 1973 Landsat image derived map.

\begin{longtable}[]{@{}llrrrrrr@{}}
\toprule
Forest type & Source & 1953-1973 & 1973-1990 & 1990-2000 & 2000-2005 &
2005-2010 & 2010-2013\tabularnewline
\midrule
\endhead
Total & Harper2007 & 91 (0.3) & 200 (1.7) & 81 (0.9) & - & - &
-\tabularnewline
& MEFT2009 & - & - & 97 (0.8) & 53 (0.5) & - & -\tabularnewline
& ONE2015 & - & - & - & - & 95 (1.2) & 164 (1.5)\tabularnewline
& this study & 86 (0.6) & 205 (1.6) & 88 (0.9) & 42 (0.4) & 70 (0.7) &
90 (1.0)\tabularnewline
Moist & Harper2007 & 94 (0.6) & 87 (1.7) & 32 (0.8) & - & - &
-\tabularnewline
& MEFT2009 & - & - & 48 (0.8) & 17 (0.4) & - & -\tabularnewline
& ONE2015 & - & - & - & - & 20 (0.5) & 37 (0.9)\tabularnewline
& this study & 79 (1.0) & 101 (1.6) & 40 (0.8) & 21 (0.4) & 27 (0.6) &
54 (1.2)\tabularnewline
Dry & Harper2007 & 11 (0.2) & 77 (1.9) & 20 (0.7) & - & - &
-\tabularnewline
& MEFT2009 & - & - & 24 (0.7) & 11 (0.4) & - & -\tabularnewline
& ONE2015 & - & - & - & - & 51 (1.8) & 97 (2.3)\tabularnewline
& this study & 16 (0.4) & 71 (1.9) & 28 (0.9) & 12 (0.4) & 29 (1.0) & 31
(1.1)\tabularnewline
Spiny & Harper2007 & -3 (-0.1) & 36 (1.2) & 28 (1.2) & - & - &
-\tabularnewline
& MEFT2009 & - & - & 25 (1.2) & 23 (1.2) & - & -\tabularnewline
& ONE2015 & - & - & - & - & 25 (1.7) & 31 (1.7)\tabularnewline
& this study & -6 (-0.2) & 31 (1.3) & 20 (1.0) & 9 (0.5) & 13 (0.7) & 4
(0.3)\tabularnewline
Mangroves & Harper2007 & - & - & 1 (0.2) & - & - & -\tabularnewline
& MEFT2009 & - & - & - & - & - & -\tabularnewline
& ONE2015 & - & - & - & - & 0 (0.3) & 0 (0.2)\tabularnewline
& this study & -3 (-1.7) & 1 (0.6) & 0 (0.2) & 0 (0.0) & 0 (0.0) & 0
(0.0)\tabularnewline
\bottomrule
\end{longtable}

Table 3: \textbf{Comparing our estimates of annual deforestation rates
for Madagascar with previous studies on the period 1953-2014}. Annual
deforestation areas (in ha/yr) and annual deforestation rates (second
number in parenthesis, in \%/yr) are provided. For deforestation rates
in \%/yr, exact same numbers as in scientific articles and reports from
previous studies have been reported. The way annual deforestation rates
in \%/yr have been computed in these previous studies can slightly
differ from one study to another but estimates always correct for the
potential presences of clouds on forest maps. Annual deforested areas in
ha/yr have been recomputed from forest-cover estimates in Tab. 2 (except
for \citet{Harper2007} for the periods 1973-1990 and 1990-2000 for which
annual deforested ares in ha/yr were derived from numbers reported in
the original publication, see methods) and do not correct for the
potential presence of clouds.

\begin{longtable}[]{@{}lrrrrrr@{}}
\toprule
Year & Forest (ha) & patch (\%) & transitional (\%) & edge (\%) &
perforated (\%) & interior (\%)\tabularnewline
\midrule
\endhead
1953 & 15,963 & 0 & 1 & 4 & 1 & 94\tabularnewline
1973 & 14,228 & 2 & 7 & 20 & 3 & 68\tabularnewline
1990 & 10,750 & 3 & 8 & 21 & 4 & 64\tabularnewline
2000 & 9,866 & 3 & 8 & 22 & 4 & 62\tabularnewline
2005 & 9,660 & 4 & 9 & 23 & 6 & 59\tabularnewline
2010 & 9,307 & 4 & 10 & 23 & 9 & 55\tabularnewline
2014 & 8,911 & 5 & 11 & 23 & 11 & 50\tabularnewline
\bottomrule
\end{longtable}

Table 4: \textbf{Evolution of the forest fragmentation from 1953 to 2014
in Madagascar}. Six categories of fragmentation were identified from the
amount of forest and its occurrence as adjacent forest pixels:
``interior'', ``perforated'', ``edge'', ``transitional'', ``patch'', and
``undetermined'' \citep{Riitters2000}. We used a moving window of 7x7
pixels (4.4 ha). Using this window size, forest edge had a width of
about 90m. The ``interior'' category can be interpreted as the most
intact forest. The ``patch'' and ``transitional'' categories correspond
to isolated small forest patches.

\hypertarget{figures}{%
\subsection{Figures}\label{figures}}

\begin{figure}
\centering
\includegraphics{outputs/ecoregion.png}
\caption{Figure 1: \textbf{Ecoregions and forest types in Madagascar.}
Madagascar can be divided into four climatic ecoregions with four forest
types: the moist forest in the East (green), the dry forest in the West
(orange), the spiny forest in the South (red), and the mangroves on the
West coast (blue). Ecoregions were defined following climatic
\citep{Cornet1974} and vegetation \citep{IEFN1996} criteria. The dark
grey areas represent the remaining natural forest cover for the year
2014.}
\end{figure}

\begin{figure}
\centering
\includegraphics{outputs/fig_fcc_highres.png}
\caption{Figure 2: \textbf{Forest-cover change on six decades from 1953
to 2014 in Madagascar.} Forest cover changes from \emph{c.} 1973 to 2014
are shown in the main figure, and forest cover in \emph{c.} 1953 is
shown in the bottom-right inset. Two zooms in the western dry (left
part) and eastern moist (right part) ecoregions present more detailed
views of (from top to bottom): forest-cover in 1950s, forest-cover
change from \emph{c.} 1973 to 2014, forest fragmentation in 2014 and
distance to forest edge in 2014. Data on water bodies (blue) and water
seasonality (light blue for seasonal water to dark blue for permanent
water) has been extracted from \citet{Pekel2016}.}
\end{figure}

\begin{figure}
\centering
\includegraphics{outputs/dist.png}
\caption{Figure 3: \textbf{Evolution of the distance to forest edge from
1953 to 2014 in Madagascar.} Black dots represent the mean distance to
forest edge for each year. Vertical dashed segments represent the 90\%
quantiles (5\% and 95\%) of the distance to forest edge. Horizontal
dashed grey line indicates a distance to forest edge of 100m.
Percentages indicate the percentage of forest at a distance to forest
edge lower than 100m for each year.}
\end{figure}

\hypertarget{references}{%
\subsection{References}\label{references}}

\bibliography{bib/biblio.bib}

\end{document}

### 2017-12-15 deforestmap.R

@@ -51,7 +51,9 @@ SavedObjects <- c()
 #dir.create("grassdata")
 #system("grass72 -c epsg:32738 grassdata/deforestmap")  # Ignore errors
 ## Connect R to grass location
-initGRASS(gisBase="/usr/local/grass-7.2.0",home=tempdir(), 
+## Make sure that /usr/lib/grass72/lib is in your PATH in RStudio
+Sys.setenv(LD_LIBRARY_PATH=paste("/usr/lib/grass72/lib", Sys.getenv("LD_LIBRARY_PATH"),sep=":"))
+initGRASS(gisBase="/usr/lib/grass72",home=tempdir(), 
           gisDbase="grassdata",
           location="deforestmap",mapset="PERMANENT",
           override=TRUE)
@@ -399,6 +401,44 @@ system("convert -pointsize 72 -gravity North -draw \"text 0,0 '2014'\" outputs/f
 #= Animated gif
 system("convert -delay 100 -loop 0 outputs/*.gif outputs/defor_Mada.gif")
 #= See also OpenShot app for movie and transition effects
+
+##==========================
+## Validation
+##==========================
+
+#= Import validation points
+valpts <- read.csv("gisdata/vectors/validation_points.csv", stringsAsFactors=FALSE)
+valpts$Obs13[valpts$Obs13==9] <- "P" # Correction
+valpts.sp <- SpatialPointsDataFrame(valpts[,1:2], valpts[3:4])
+crs(valpts.sp) <- CRS("+init=epsg:4326")
+
+#= Reproject in UTM38S
+newcrs <- CRS("+init=epsg:32738")
+valpts.38s <- spTransform(valpts.sp, newcrs)
+
+#= Export 
+writeOGR(valpts.38s, dsn="outputs", layer="valpts_38s", driver="ESRI Shapefile")
+
+#= Import into GRASS
+system("v.in.ogr --o input=outputs/valpts_38s.shp output=valpts")
+
+#= Get values from for2010 and for 2013
+system("v.what.rast map=valpts raster=for2010 column=for2010")
+system("v.what.rast map=valpts raster=for2013 column=for2013")
+
+#= Export to output
[…]

### 2017-12-18 deforestmap.R
Sorting files

@@ -971,6 +971,13 @@ render("deforestmap.Rmd")
 ## Cover letter
 render("coverletter.md",output_format=c("pdf_document"),output_dir="report") # pdf output
 
+##========================
+## Pandoc for word
+##========================
+
+system("pandoc manuscript/manuscript_essd.tex --bibliography=manuscript/biblio.bib \\
+       --csl=manuscript/bib/methods-in-ecology-and-evolution.csl -o manuscript/manuscript_essd.docx")
+
 ##===========================================================================
 ## End of script
 ##===========================================================================

### 2017-12-18 deforestmap.sh

@@ -1,8 +1,0 @@
-#!/bin/sh
-echo "Executing deforestmap.R script in the background"
-Rscript --vanilla deforestmap.R > deforestmap.log 2>&1 &
-# If you wan't to download the source data do:
-# Rscript --vanilla deforestmap.R TRUE > deforestmap.log 2>&1 &
-echo "Check the progress with command 'tail -f deforestmap.log'"
-echo "Check the processor usage with command 'top'"
-## End of script

### 2017-12-18 manuscript/manuscript.Rmd
Reorganizing files

--- 2017-06-07 deforestmap.Rmd
@@
 tab.1 <- defor.df
-names(tab.1) <- c("Year", "Forest (ha)", "Unmap (ha)", "Annual defor. (ha/yr)", "Rate (%/yr)")
-kable(tab.1)
+tab.1[,c(2:4)] <- round(tab.1[,c(2:4)]/1000)
+tab.1$Year <- as.character(tab.1$Year)
+names(tab.1) <- c("Year", "Forest (Kha)", "Unmap (Kha)", "Annual defor. (Kha/yr)", "Rate (%/yr)")
+kable(tab.1, format.args = list(big.mark=","), digits=1)
 ```
@@
 tab.2 <- tab.2[,c(2,1,3:10)]
+tab.2[,c(3:10)] <- round(tab.2[,c(3:10)]/1000)
 tab.2$ForestType <- c("Total",rep("",3),"Moist",rep(" ",3),"Dry",rep(" ",3),"Spiny",rep(" ",3),"Mangroves",rep(" ",3))
 col.names <- c("Forest type","Source","1953","1973","1990","2000","2005","2010","2013","2014")
-kable(tab.2,col.names=col.names,align=c("l","l",rep("r",8)))
-#print(xtable(tab.2,col.names=col.names,align=c("llrrrrrrrrr")),type="html")
+kable(tab.2,col.names=col.names,align=c("l","l",rep("r",8)), format.args = list(big.mark=","))
 ```
@@
 ```{r defor_comp, results="asis"}
-tab.3 <- defor.comp
+tab.3 <- read.table(file="outputs/defor_comp.txt", sep=" ", header=TRUE)
 tab.3$ForestType <- c("Total",rep("",3),"Moist",rep(" ",3),"Dry",rep(" ",3),"Spiny",rep(" ",3),"Mangroves",rep(" ",3))
@@
 p.3 <- round(frag.df$patch[frag.df$Year==2014]+frag.df$transitional[frag.df$Year==2014])
-names(tab.4) <- c("Year","Forest (ha)", "patch (%)", "transitional (%)", "edge (%)", "perforated (%)", "interior (%)", "NA (%)")
-kable(tab.4)
+tab.4 <- tab.4[,-c(8)]
+tab.4$Year <- as.character(tab.4$Year)
+tab.4$forest <- round(tab.4$forest/1000)
+names(tab.4) <- c("Year","Forest (ha)", "patch (%)", "transitional (%)", "edge (%)", "perforated (%)", "interior (%)")
+kable(tab.4, digits=0, format.args=list(big.mark=","))
 ```

### 2017-12-18 manuscript/manuscript.tex
Reorganizing files

--- 2017-06-07 latex/deforestmap.tex
@@
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Preambule %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\documentclass[a4paper, 12pt, leqno]{article} %leqno: numéro d'équation à gauche
-\pagenumbering{arabic} % choose how to number the pages
-\usepackage{a4wide}
-\usepackage[utf8]{inputenc} % accents interprétés
-\usepackage{graphicx}
+%% Copernicus Publications Manuscript Preparation Template for LaTeX Submissions
+%% ---------------------------------
+%% This template should be used for copernicus.cls
+%% The class file and some style files are bundled in the Copernicus Latex Package, which can be downloaded from the different journal webpages.
+%% For further assistance please contact Copernicus Publications at: production@copernicus.org
+%% https://publications.copernicus.org/for_authors/manuscript_preparation.html
+
+%% Please use the following documentclass and journal abbreviations for discussion papers and final revised papers.
+
+%% 2-column papers and discussion papers
+\documentclass[essd, classical]{copernicus}
+
+%% Journal abbreviations (please use the same for discussion papers and final revised papers)
+
+% Archives Animal Breeding (aab)
+% Atmospheric Chemistry and Physics (acp)
+% Advances in Geosciences (adgeo)
+% Advances in Statistical Climatology, Meteorology and Oceanography (ascmo)
+% Annales Geophysicae (angeo)
+% ASTRA Proceedings (ap)
+% Atmospheric Measurement Techniques (amt)
+% Advances in Radio Science (ars)
+% Advances in Science and Research (asr)
+% Biogeosciences (bg)
+% Climate of the Past (cp)
+% Drinking Water Engineering and Science (dwes)
+% Earth System Dynamics (esd)
+% Earth Surface Dynamics (esurf)
+% Earth System Science Data (essd)
+% Fossil Record (fr)
+% Geographica Helvetica (gh)
+% Geoscientific Instrumentation, Methods and Data Systems (gi)
+% Geoscientific Model Development (gmd)
+% Hydrology and Earth System Sciences (hess)
+% History of Geo- and Space Sciences (hgss)
+% Journal of Micropalaeontology (jm)
+% Journal of Sensors and Sensor Systems (jsss)
+% Mechanical Sciences (ms)
+% Natural Hazards and Earth System Sciences (nhess)
+% Nonlinear Processes in Geophysics (npg)
+% Ocean Science (os)
+% Proceedings of the International Association of Hydrological Sciences (piahs)
+% Primate Biology (pb)
+% Scientific Drilling (sd)
+% SOIL (soil)
+% Solid Earth (se)
+% The Cryosphere (tc)
+% Web Ecology (we)
+% Wind Energy Science (wes)
+
+%% \usepackage commands included in the copernicus.cls:
+%\usepackage[german, english]{babel}
+\usepackage{tabularx}
+%\usepackage{cancel}
+%\usepackage{multirow}
+\usepackage{supertabular}
+%\usepackage{algorithmic}
+%\usepackage{algorithm}
+%\usepackage{amsthm}
+%\usepackage{float}
 %\usepackage{subfig}
-%\usepackage[hmargin=2cm, vmargin = 2cm, noheadfoot]{geometry} %% Pour gérer le format des pages
-%\usepackage{layout} %% Pour avoir la longueur des marges
-\usepackage[round,sort]{natbib} %% Natbib is a popular style for formatting references.
-%\usepackage{multibib}
-%\newcites{secnm}{Bibliographie} 
-%\usepackage{verbatim} % for multiline comments
+%\usepackage{rotating}
+
+%% My packages
 \usepackage{amssymb} %symbole de maths
 \usepackage{amsmath} %idem
-%\usepackage{stmaryrd} %% Symbole flèche à l'envers
-%\usepackage{amsfonts}
-\usepackage[english]{babel} %% Les titres en anglais
-\usepackage{array} %% Pour centrer verticalement le contenu d'un tableau, entre autres...
-\setcounter{secnumdepth}{4} %% Profondeur des sections, subsections
-\usepackage{setspace} %% Gère l'interligne: singlespacing, doublespacing
-\usepackage{booktabs}
-%\singlespacing
 \usepackage{longtable}
@@
 \setlength{\LTright}{-5cm plus 1 fill}
-\usepackage[colorlinks=true,citecolor=blue]{hyperref} %% Gère les hyperliens
-\usepackage{lineno} %% Numérotation des lignes
-\usepackage{caption2}
-%\linenumbers
+%% New commands
 \newcommand{\logit}{\text{logit}}
 \newcommand{\bs}[1]{\boldsymbol{#1}}
-\newcommand{\p}{\text{p}}
-\newcommand{\R}{\textnormal{\sffamily\bfseries R}}
+\newcommand{\R}{\textnormal{\sffamily{\bfseries{R}}}}
 \newcommand{\pkg}[1]{{\fontseries{b}\selectfont #1}}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Title %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{document}
 
@@
   fragmentation in Madagascar}
-\date{}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Document %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{document}
+
+% \Author[affil]{given_name}{surname}
+
+\Author[1, 2, 3]{Ghislain}{Vieilledent}
+\Author[4]{Clovis}{Grinand}
+\Author[4]{Fety A.}{Rakotomalala}
+\Author[5]{Rija}{Ranaivosoa}
+\Author[5]{Jean-Roger}{Rakotoarijaona}
+\Author[6, 7]{Thomas F.}{Allnutt}
+\Author[1]{Frédéric}{Achard}
+
+\affil[1]{Joint Research Center of the European Commission, Bio-economy Unit, I-21027 Ispra (VA), ITALY}
+\affil[2]{Cirad, UPR Forêts et Sociétés, F-34398 Montpellier, FRANCE}
+\affil[3]{Forêts et Sociétés, Univ. Montpellier, Cirad, Montpellier, FRANCE}
+\affil[4]{ETC Terra, F-75020 Paris, FRANCE}
+\affil[5]{Office National pour l'Environnement, 101 Antananarivo, MADAGASCAR}
+\affil[6]{Wildlife Conservation Society, 101 Antananarivo, MADAGASCAR}
+\affil[7]{GreenInfo Network, Oakland, California, USA}
+
+%% The [] brackets identify the author with the corresponding affiliation. 1, 2, 3, etc. should be inserted.
+
+\runningtitle{Six decades of deforestation in Madagascar}
+
+\runningauthor{G. Vieilledent}
+
+\correspondence{Ghislain Vieilledent (ghislain.vieilledent@cirad.fr)}
+
+\received{}
+\pubdiscuss{} %% only important for two-stage journals
+\revised{}
+\accepted{}
+\published{}
+
+%% These dates will be inserted by Copernicus Publications during the typesetting process.
+
+\firstpage{1}
 \maketitle
 
-\vspace{-1cm}
-\begin{center}
-%%{\large
-  Ghislain Vieilledent$^{1,2,\star}$\hspace{1cm}
-  Clovis Grinand$^{3}$ \\
-  \vspace{0.25cm}
-  Fety A. Rakotomalala$^{3}$ \hspace{1cm}
-  Rija Ranaivosoa$^{4}$ \hspace{1cm}
-  Jean-Roger Rakotoarijaona$^{4}$ \\
-  \vspace{0.25cm}
-  Thomas F. Allnutt$^{5,6}$\hspace{1cm}and\hspace{1cm}
-  Frédéric Achard$^{1}$
-%%}
-\end{center}
-
-\vspace{0.3cm}
-
-{\small
-  \begin{flushleft}  
-    $[1]$ \textbf{Joint Research Center of the European Commission} -- Bio-economy Unit, I-21027 Ispra (VA), ITALY\\
-    $[2]$ \textbf{Cirad} -- UPR Forêts et Sociétés, F-34398 Montpellier, FRANCE\\
-    $[3]$ \textbf{ETC Terra}, F-75020 Paris, FRANCE\\
-    $[4]$ \textbf{Office National pour l'Environnement}, 101 Antananarivo, MADAGASCAR\\
-    $[5]$ \textbf{Wildlife Conservation Society}, 101 Antananarivo, MADAGASCAR\\
-    $[6]$ \textbf{GreenInfo Network}, Oakland, California, USA\\
-    ~\\
-    $[\star]$ \textbf{Corresponding author:}
-    \textbackslash{E-mail}:~ghislain.vieilledent@cirad.fr
-    \textbackslash{Phone}:~+39.033.278.3516\\
-  \end{flushleft}}
-
-\vspace{0.3cm}
-
-\begin{center}
-  \textbf{Running headline:} \emph{Six decades of deforestation in Madagascar}
-\end{center}
-
-\newpage
-
-\linenumbers
-\doublespacing
-
-\newcommand{\keywords}[1]{\par\noindent
-{\small{\em Keywords\/}: #1}}
-
 \begin{abstract}
 
-  1. The island of Madagascar has an unparalleled biodiversity, mainly
+  The island of Madagascar has an unparalleled biodiversity, mainly
   located in the tropical forests of the island, which is highly
@@
   from past studies at national level with substantial gaps (due to
-  presence of cloud cover on satellite imagery) prevent the analyzis
-  of long-term deforestation trends in Madagascar.
-
-  2. In this study, we propose a new approach combining historical
-  (1953-2000) national forest-cover maps with recent (2000-2014)
-  global annual tree cover loss data to look at six decades
-  (1953-2014) of deforestation and forest fragmentation in
-  Madagascar. We produced new forest-cover maps at 30m-resolution over
-  the full territory of Madagascar for the year 1990, and annually
-  from 2000 to 2014.
-
-  3. We estimated that Madagascar has lost 44\% of its natural forest
-  cover over the period 1953-2014 (including 37\% over the period
-  1973-2014).  Natural forests cover 8.9 Mha in 2014 (15\% of the
-  national territory) which are divided into 4.4 Mha (50\%) of moist
-  forests, 2.6 Mha (29\%) of dry forests, 1.7 Mha of spiny forests
-  (19\%) and 177,000 ha (2\%) of mangroves. Since 2005, the annual
-  deforestation rate has progressively increased in Madagascar to
-  reach 99,000 ha/yr during 2010-2014 (corresponding to a rate of
-  1.08\%/yr). Around half of the forest (46\%) is now located at less
-  than 100m from the forest edge.
-
-  4. \emph{Policy implications}: Increase in the deforestation rate
-  since 2005 can be related to rapid population growth (close to
-  3\%/yr) and to poor law enforcement due to political instability in
-  the country.  More effort, including conservation and development
-  programs, is needed in Madagascar to protect the remaining natural
-  tropical forests, both to enhance local people's livelihoods and to
-  support global environmental efforts.
-
-\vspace{0.5cm}
-
-\keywords{biodiversity, climate-change, deforestation, Landsat,
-  Madagascar, tropical forest}
+  presence of cloud cover on satellite imagery) prevent the analysis
+  of long-term deforestation trends in Madagascar. In this study, we
+  propose a new approach combining historical (1953-2000) national
+  forest-cover maps with recent (2001-2014) global annual tree cover
+  loss data to look at six decades (1953-2014) of deforestation and
+  forest fragmentation in Madagascar. We produced new forest-cover
+  maps at 30~m resolution over the full territory of Madagascar for
+  the year 1990, and annually from 2000 to 2014. We estimated that
+  Madagascar has lost 44\% of its natural forest cover over the period
+  1953-2014 (including 37\% over the period 1973-2014). Natural
+  forests cover 8.9 Mha in 2014 (15\% of the national territory) which
+  are divided into 4.4 Mha (50\%) of moist forests, 2.6 Mha (29\%) of
+  dry forests, 1.7 Mha of spiny forests (19\%) and 177,000 ha (2\%) of
+  mangroves. Since 2005, the annual deforestation rate has
+  progressively increased in Madagascar to reach 99,000 ha/yr during
+  2010-2014 (corresponding to a rate of 1.1\%/yr). This increase is
+  probably due to rapid population growth (close to 3\%/yr) and to
+  poor law enforcement in the country. Around half of the forest
+  (46\%) is now located at less than 100~m from the forest
+  edge. Accurate forest-cover change maps can be used to assess the
+  effectiveness of past and current conservation programs and
+  implement new strategies for the future. In particular, forest maps
+  and estimates can be used in the REDD+ framework which aims at
+  ``Reducing Emissions from Deforestation and Forest Degradation'' and
+  for optimizing the current protected area network.
 
@@
 
-\newpage
+\copyrightstatement{The article is distributed under the Creative Commons Attribution 4.0 License.}
 
@@
 spiny ecosystems of Madagascar (see Fig. 8 in \citet{Kim2014}).
-\citet{Hansen2013} mapped tree cover percentage, annual forest loss
-and forest gain from 2000 to 2012 at global scale at 30 m
+\citet{Hansen2013} mapped tree cover percentage, annual tree cover
+loss and gain from 2000 to 2012 at global scale at 30 m
 resolution. This product has since been updated and is now available
@@
 from \citet{Hansen2013} cannot be used alone to produce a map of
-forest cover \citep{Tyukavina2017}. In complement to the tree cover
-percentage provided in \citet{Hansen2013}, a layer of annual tree
-cover loss is also provided (i.e.~complete loss of tree cover from a
-value higher than 10\% to zero) for the period 2001-2014.
+forest cover \citep{Tyukavina2017}.
 
@@
 2000 and 2014. This approach can easily be repeated in the future when
-the \citet{Hansen2013} products are updated.
+tree cover loss products will be updated.
 
@@
 
-We produced annual forest/non-forest maps at 30m resolution for the
+We produced annual forest/non-forest maps at 30~m resolution for the
 full territory of Madagascar for the period 2000-2014 by combining the
 forest map of year 2000 from \citet{Harper2007}, and the tree cover
-percentage and annual forest cover loss maps over the period 2000-2014
+percentage and annual tree cover loss maps over the period 2000-2014
 from \citet{Hansen2013}. The 2000 Harper's forest map includes 208,000
-ha of unclassified areas due to the presence of clouds in input
-satellite imagery, mostly (88\%) within the moist forest domain which
-covers 4.17 million ha in total. To provide a label (forest or
-non-forest) to these unclassified pixels, we use the 2000 tree cover
-percentage map of \citet{Hansen2013} by selecting a threshold of 75\%
-tree cover to define forest cover as recommended by other studies for
-the moist domain \citep{Achard2014}. We thus obtained a forest-cover
-map for the year 2000 covering the full territory of Madagascar. We
-then combined this forest-cover map of the year 2000 with the annual
-forest cover loss maps from 2001 to 2014 provided by
+ha of unclassified areas due to the presence of clouds on satellite
+images, mostly (88\%) within the moist forest domain which covered
+4.17 Mha in total in 2000. To provide a label (forest or non-forest)
+to these unclassified pixels, we used the 2000 tree cover percentage
+map of \citet{Hansen2013} by selecting a threshold of 75\% tree cover
+to define forest cover as recommended by other studies for the moist
+domain \citep{Achard2014, Aleman2017}. To do so, the Hansen's 2000
+tree cover map was resampled on the same grid as the original Harper's
+map at 30~m resolution using a bilinear interpolation. We thus obtained
+a forest-cover map for the year 2000 covering the full territory of
+Madagascar. We then combined this forest-cover map of the year 2000
+with the annual tree cover loss maps from 2001 to 2014 provided by
 \citet{Hansen2013} to create annual forest-cover maps from 2001 to
-2014 at 30m resolution. We also completed the Harper's forest map of
-year 1990 by filling unclassified areas (due to the presence of clouds
-on satellite images) using our forest-cover map of year 2000. To do
-so, we assumed that if forest was present in 2000, the pixel was also
-forested in 1990. The remaining unclassified pixels were limited to a
-relatively small total area of c. 8,000 ha. We labeled these residual
-pixels as non-forest as for the year 2000.  Similarly we completed the
-Harper forest map of year 1973 by filling unclassified areas using our
-forest-cover map of the year 1990 assuming that if forest was present
-in 1990, it was also present in 1973.  Contrary to the year 1990, the
-remaining unclassified pixels for year 1973 corresponded to a
-significant total area of 3.32 million ha. We also reprojected the
-forest-cover map of year 1953 to a common projection in order to
-compare to the other eras. This map was produced by scanning a paper
-map derived from aerial photos, and thus could not be perfectly
-aligned with the other maps produced through digital processing of
-satellite imagery \citep{Harper2007}. Finally for all forest-cover
-maps from 1973, the isolated single non-forest pixels (i.e.~fully
-surrounded by forest pixels) were removed, assuming that single
-non-forest pixels inside a forest patch were not corresponding to
-deforestation (they might correspond to selective logging activities).
-This allowed us to avoid counting very small scale events
-(\textless{}0.1 ha such as selective logging) as forest fragmentation.
-All the resulting maps are freely available at
+2014 at 30~m resolution. To do so, Hansen's tree cover loss maps were
+resampled on the same grid as the original Harper's map at 30~m
+resolution using a nearest-neighbor interpolation. We also completed
+the Harper's forest map of year 1990 by filling unclassified areas
+(due to the presence of clouds on satellite images) using our
+forest-cover map of year 2000. To do so, we assumed that if forest was
+present in 2000, the pixel was also forested in 1990. The remaining
+unclassified pixels were limited to a relatively small total area of
+\emph{c.} 8,000 ha. We labeled these residual pixels as non-forest, as
+for the year 2000. Similarly we completed the Harper's forest map of
+year 1973 by filling unclassified areas using our forest-cover map of
+the year 1990 assuming that if forest was present in 1990, it was also
+present in 1973. Contrary to the year 1990, the remaining unclassified
+pixels for year 1973 corresponded to a significant total area of 3.3
+million ha. We also reprojected the forest-cover map of year 1953 to a
+common projection in order to compare the forest-cover area in 1953
+with forest-cover areas at the following dates. This map was produced
+by scanning a paper map derived from aerial photos, and thus could not
+be perfectly aligned with the other maps produced through digital
+processing of satellite imagery \citep{Harper2007}. Finally for all
+forest-cover maps from 1973, the isolated single non-forest pixels
+(i.e.~fully surrounded by forest pixels) were removed, assuming that
+single non-forest pixels inside a forest patch were not corresponding
+to deforestation (they might correspond to selective logging
+activities). This allowed us to avoid counting very small scale events
+(\textless 0.1 ha, such as selective logging) as forest
+fragmentation. All the resulting maps are freely available at
 \url{https://bioscenemada.cirad.fr/forestmaps}.
@@
 From these new forest-cover maps, we calculated the total forest-cover
-area for seven available years (1953-1973-1990-2000-2010-2005-2014),
+area for seven available years (1953-1973-1990-2000-2005-2010-2014),
 and the annual deforested area and annual deforestation rate for the
 corresponding six time periods between 1953 and 2014. The annual
-deforestation rates were calculated as follows \citep{Puyravaud2003,
-  Vieilledent2013}:
+deforestation rates were calculated using Eq.~\ref{eq:theta}
+\citep{Puyravaud2003, Vieilledent2013}:
 
 \begin{equation}
+  \label{eq:theta}
   \theta = 100 \times [1-(1-(F_{t_2}-F_{t_1})/F_{t_1})^{(1/(t_2-t_1))}
@@
 
-where $\theta$ is the annual deforestation rate (in \%/yr),
+In Eq.~\ref{eq:theta}, $\theta$ is the annual deforestation rate (in \%/yr),
 $F_{t_2}$ and $F_{t_1}$ are the forest cover free of clouds at both
@@
 
-Because of the large unclassified area (3.32 million ha) in 1973, the
+Because of the large unclassified area (3.3 million ha) in 1973, the
 annual deforestation areas and rates for the two periods 1953-1973 and
-1973-1990 are only indicative estimates. For these two periods the
-annual deforestation rates are computed as the ratio
-$(F_{t_2}-F_{t_1})/F_{t_1}$ considering only the mapped forest pixels.
-Area and rate estimates are produced at the national scale and for the
-four forest ecosystems present in Madagascar: moist forest in the
-East, dry forest in the West, spiny forest in the South, and mangroves
-on the Western coast (Fig.~\ref{fig:ecoregion}). To define the forest
-domains, we used a map from the MEFT (\emph{``Ministère de
-  l'Environnement et des Forêts à Madagascar''}) with the boundaries
-of the four ecoregions in Madagascar.  Ecoregions were defined on the
-basis of climatic and vegetation criteria using the climate
-classification by \citet{Cornet1974} and the vegetation classification
-from the 1996 IEFN national forest inventory \citep{IEFN1996}. Because
-mangrove forests are highly dynamic ecosystems that can expand or
-contract on decadal scales depending on changes in environmental
-factors \citep{Armitage2015}, a fixed delimitation of the mangrove
-ecoregion on six decades might not be fully appropriate. As a
-consequence, our estimates of the forest-cover and deforestation rates
-for mangroves in Madagascar must be considered with this limitation.
+1973-1990 are only partial estimates computed on the basis of the
+available forest extent. Area and rate estimates are produced at the
+national scale and for the four forest ecosystems present in
+Madagascar: moist forest in the East, dry forest in the West, spiny
+forest in the South, and mangroves on the Western coast
+(Fig.~\ref{fig:ecoregion}). To define the forest domains, we used a
+map from the MEFT (\emph{``Ministère de l'Environnement et des Forêts
+  à Madagascar''}) with the boundaries of the four ecoregions in
+Madagascar. Ecoregions were defined on the basis of climatic and
+vegetation criteria using the climate classification by
+\citet{Cornet1974} and the vegetation classification from the 1996
+IEFN national forest inventory \citep{IEFN1996}. Because mangrove
+forests are highly dynamic ecosystems that can expand or contract on
+decadal scales depending on changes in environmental factors
+\citep{Armitage2015}, a fixed delimitation of the mangrove ecoregion
+on six decades might not be fully appropriate. As a consequence, our
+estimates of the forest-cover and deforestation rates for mangroves in
+Madagascar must be considered with this limitation.
+
+\begin{figure}[h]
+  \centering
+
+    \includegraphics[width=8.3cm]{ecoregion.png}
+
+    \caption{\textbf{Ecoregions and forest types in Madagascar.}
+      Madagascar can be divided into four climatic ecoregions with four forest
+      types: the moist forest in the East (green), the dry forest in the West
+      (orange), the spiny forest in the South (red), and the mangroves on the
+      West coast (blue). Ecoregions were defined following climatic
+      \citep{Cornet1974} and vegetation \citep{IEFN1996} criteria. The dark
+      grey areas represent the remaining natural forest cover for the year
+      2014.}
+
+    \label{fig:ecoregion}
+
+\end{figure}
 
@@
 not the case for \citet{Harper2007} and \citet{MEFT2009}, which can
-explain part of the differences between the estimates.
+explain a part of the differences between the estimates.
 
@@
 We also conducted an analysis of changes in forest fragmentation for
-the years 1953, 1973, 1990, 2000, 2010 and 2014. We applied the method
-developed by \citet{Riitters2000} which uses a moving window to
-characterize the fragmentation around each forested
-pixel. Computations were done using the function \texttt{r.forestfrag}
-of the GRASS GIS software \citep{Neteler2008}. Six categories of
-fragmentation were identified from the amount of forest and its
-occurrence as adjacent forest pixels: ``interior'', ``perforated'',
-``edge'', ``transitional'', ``patch'', and ``undetermined''. We used a
-moving window of 7x7 pixels (4.4 ha). Using this window size, forest
-edge had a width of about 90m \citep{Riitters2000}. The ``interior''
-category can be interpreted as the most intact forest
+the years 1953, 1973, 1990, 2000, 2005, 2010 and 2014 at 30~m
+resolution. We applied the method developed by \citet{Riitters2000}
+which uses a moving window to characterize the fragmentation around
+each forested pixel. Computations were done using the function
+\texttt{r.forestfrag} of the GRASS GIS software
+\citep{Neteler2008}. Six categories of fragmentation were identified
+from the amount of forest and its occurrence as adjacent forest
+pixels: ``interior'', ``perforated'', ``edge'', ``transitional'',
+``patch'', and ``undetermined''. Previous studies have shown that
+micro-habitats were mainly altered within the first 100~m of the
+forest edge \citep{Brinck2017, Gibson2013, Murcia1995,
+  Broadbent2008}. To characterize fragmentation, we have used a moving
+window of 7x7 pixels (4.4~ha). Using this window size, forest edge had
+a width of about 90~m \citep{Riitters2000}, which is close to the
+value of 100~m and conservative regarding edge effects. The
+``interior'' category can be interpreted as the most intact forest
 \citep{Potapov2017}. The ``patch'' and ``transitional'' categories
@@
 1973, 1990, 2000, 2005, 2010 and 2014. For that, we used the function
-\texttt{gdal\_proximity.py} of the GDAL software
-(\url{http://www.gdal.org/}). We computed the mean and 95\% quantiles
-of the distance to forest edge and looked at the evolution of these
-values with time.
-
-\section{Results}
-\label{results}
-
-\subsection{Dynamics of forest cover and deforestation
-intensity}
-
-Natural forests in Madagascar covered 16.0 Mha in 1953, about 27\% of
-the national territory of 587,041 km2. In 2014, the forest cover
-dropped to 8.9 Mha, corresponding to about 15\% of the national
-territory (Fig.~\ref{fig:fcc} and
-Tab.~\ref{tab:forest_cover}). Madagascar has lost 44\% and 37\% of its
-natural forests between 1953 and 2013, and between c. 1973 and 2014
-respectively (Fig.~\ref{fig:fcc} and Tab.~\ref{tab:forest_cover}). In
-2014 the remaining 8.9 Mha of natural forest were distributed as: 4.4
-Mha of moist forest (50\% of total forest cover), 2.6 Mha of dry
-forest (29\%), 1.7 Mha of spiny forest (19\%) and 0.18 Mha (2\%) of
-mangrove forest (Fig.~\ref{fig:ecoregion} and
-Tab.~\ref{tab:comp_forest}). Regarding the deforestation trend, we
-observed a progressive decrease of the deforestation rate after1990
-from 205,000 ha/yr (1.63\%/yr) over the period 1973-1990 to 44,300
-ha/yr (0.43\%/yr) over the period 2000-2005
-(Tab.~\ref{tab:forest_cover}). Then from 2005, the deforestation rate
-has progressively increased and has more than doubled over the period
-2010-2014 (98,700 ha/yr, 1.08\%/yr) compared to 2000-2005
-(Tab.~\ref{tab:forest_cover}). The deforestation trend characterized
-by a progressive decrease of the deforestation rate over the period
-1990-2005 and a progressive increase of the deforestation after 2005
-is valid for all four ecoregions (Tab.~\ref{tab:comp_defor}), with the
-exception of the spiny forest domain for which the deforestation rate
-during the period 2010-2013 was lower than during 2005-2010
-(Tab.~\ref{tab:comp_defor}).
-
-\subsection{Comparison with previous forest-cover change studies
-in Madagascar}
-
-Forest-cover maps provided by previous studies over Madagascar were
-not exhaustive (unclassified areas) due to the presence of clouds in
-satellite imagery used to produce such maps. In \citet{Harper2007},
-the maps of years 1990 and 2000 include 0.5 and 1.12 Mha of unknown
-cover type respectively. Proportions of unclassified areas are not
-reported in the two other existing studies by \citet{MEFT2009} and
-\citet{ONE2015}.  With our approach, we produced wall to wall
-forest-cover change maps from 1990 to 2014 for the full territory of
-Madagascar (Tab.~\ref{tab:forest_cover}). This allowed us to produce
-more robust estimates of forest-cover and deforestation rates over
-this period. Our forest-cover estimates over the period 1953-2013
-(considering forest cover estimates at national level and by
-ecoregions for the 6 available dates) were well correlated (0.99) to
-estimates from the three previous studies (Tab.~\ref{tab:comp_forest})
-with a RMSE of 300,000 ha (6\% of the mean forest cover of 4.8 Mha
-when considering all dates and forest types together). These small
-differences can be partly attributed to differences in ecoregion
-boundaries. Despite significant differences in deforestation estimates
-(Tab.~\ref{tab:comp_defor}), a similar deforestation trend was
-observed across studies with a decrease of deforestation rates over
-the period 1990-2005, followed by a progressive increase of the
-deforestation after 2005.
-
-\subsection{Evolution of forest fragmentation with time}
-
-In parallel to the dynamics of deforestation, forest fragmentation has
-progressively increased since 1953 in Madagascar. We observed a
-continuous decrease of the mean distance to forest edge from 1953 to
-2014 in Madagascar. The mean distance to forest edge has decreased to
-c.  300 m in 2014 while it was previously c. 1.5 km in 1973
-(Fig.~\ref{fig:dist_edge}).  Moreover, a large proportion (73\%) of
-the forest was located at a distance greater than 100 m in 1973, while
-almost half of the forest (46\%) was at a distance lower than 100 m
-from forest edge in 2014 (Fig.  3). The percentage of forest that can
-be considered intact in Madagascar has continuously decreased since
-1953. The percentage of forest belonging to the ``interior'' category
-(most intact forests) has fallen from 68\% in 1973 to 50\% in 2014. In
-2014, more than 16\% of the forest belonged to the ``patch'' and
-``transitional'' categories (isolated small forest patches) compared
-to 9.5\% in 1973 (Tab.~\ref{tab:frag}).
-
-\section{Discussion}
-\label{discussion}
-
-\subsection{Benefits of the combined use of recent global annual
-  tree cover loss data with historical national forest-cover maps}
-
-In this study, we combined recent (2001-2014) global annual tree cover
-loss data \citep{Hansen2013} with historical (1953-2000) national
-forest-cover maps \citep{Harper2007} to look at six decades (1953-2014)
-of deforestation and forest fragmentation in Madagascar. We produced
-annual forest-cover maps at 30m-resolution covering Madagascar for the
-period 2000 to 2014. Our study extends the forest-cover monitoring on a
-six decades period (from 1953 to 2014) while harmonizing the data from
-the previous studies. We propose a generic approach to solve the problem
-of forest definition which is needed to transform the 2000 global tree
-cover dataset from Hansen et al. (2013) into a forest/non-forest map
-\citep{Tropek2014}. We propose to use a historical national forest-cover
-map, based on a national forest definition, as a forest cover mask. This
-approach could be easily extended to other regions or countries for
-which an accurate forest-cover map is available at any date within the
-period 2000-2014, but preferably at the beginning of the period to
-profit from the full record and derive long-term estimates of
-deforestation. Moreover, this approach can be repeated in the future if
-and when the global tree cover product is updated. We have made the
-R/GRASS code used for this study freely available in a GitHub repository
-(see Data availability statement) to facilitate application to other
-study areas.
-
-The accuracy of the derived forest-cover change maps depends directly on
-the accuracies of the historical forest-cover maps and the tree cover
-loss dataset. The reported global accuracy of the tree cover loss
-dataset is 99.6\% (see Tab. S5 in \citet{Hansen2013}).
-\citet{Verhegghen2016} have compared deforestation estimates derived
-from the global tree cover loss dataset \citep{Hansen2013} with results
-derived from semi-automated supervised classification of Landsat
-satellite images \citep{Achard2014} for six countries in Central Africa
-and they found a good agreement between these two sets of estimates.
-Consistent with \citet{Harper2007}, we did not consider potential forest
-regrowth in Madagascar (although \citet{Hansen2013} provided a tree
-cover gains layer for the period 2001-2013) for the following reasons:
-(i) the tree gain layer of \citet{Hansen2013} includes and catches more
-easily tree plantations than natural forest regrowth \citep{Tropek2014};
-(ii) there is little evidence of natural forest regeneration in
-Madagascar \citep[\citet{Grouzis2001}]{Harper2007} - this can be
-explained by several ecological processes following burning practice
-such as soil erosion \citep{Grinand2017} and reduced seed bank due to
-fire and soil loss \citep{Grouzis2001}, (iii) in areas where forest
-regeneration is ecologically possible, young forest regrowth are more
-easily re-burnt for agriculture and pasture; and (iv) young secondary
-forests provide more limited ecosystem services compared to old-growth
-natural forests in terms of biodiversity and carbon storage.
-
-\subsection{Dynamics of forest-cover in Madagascar from 1953 to 2014}
-
-We estimated that natural forests in Madagascar cover 8.9 Mha in 2014
-(corresponding to 15\% of the country) and that Madagascar has lost 44\%
-of its natural forest since 1953 (37\% since 1973). There is ongoing
-scientific debate about the extent of the ``original'' forest cover in
-Madagascar, and the extent to which humans have altered the natural
-forest landscapes since their large-scale settlement around 800 CE
-\citep{Burns2016, Cox2012}. Early French naturalists stated that the
-full island was originally covered by forest
-\citep{Humbert1927, Perrier1921}, leading to the common statement that
-90\% of the natural forests have disappeared since the arrival of humans
-on the island \citep{Kull2000}. More recent studies counter-balanced
-that point of view saying that extensive areas of grassland existed in
-Madagascar long before human arrival and were determined by climate,
-natural grazing and other natural factors
-\citep[\citet{Vorontsova2017}]{Virah-Sawmy2009}. Other authors have
-questioned the entire narrative of extensive alteration of the landscape
-by early human activity which, through legislation, has severe
-consequences on local people \citep{Klein2002, Kull2000}. Whatever the
-original proportion of natural forests and grasslands in Madagascar, our
-results demonstrate that human activities since the 1950s have
-profoundly impacted the natural tropical forests and that recent
-conservation and development programs in Madagascar have failed to stop
-deforestation in the recent years. Deforestation has strong consequences
-on biodiversity and carbon emissions in Madagascar. Around 90\% of
-Madagascar's species are forest dependent
-\citep{Allnutt2008, Goodman2005} and \citet{Allnutt2008} estimated that
-deforestation between 1953 and 2000 led to an extinction of 9\% of the
-species. The additional deforestation we observed over the period
-2000-2014 (around 1Mha of natural forest) worsen this result. Regarding
-carbon emissions, using the 2010 aboveground forest carbon map by
-\citet{Vieilledent2016}, we estimated that deforestation on the period
-2010-2014 has led to 40.2 Mt C of carbon emissions in the atmosphere (10
-Mt C /yr) and that the remaining aboveground forest carbon stock in 2014
-is 832.8 Mt C. Associated to deforestation, we showed that the remaining
-forests of Madagascar are highly fragmented with 46\% of the forest
-being at less than 100m of the forest edge. Small forest fragments do
-not allow to maintain viable populations and ``edge effects'' at
-forest/non-forest interfaces have impacts on both carbon emissions
-\citep{Brinck2017} and biodiversity loss \citep{Gibson2013, Murcia1995}.
-
-\subsection{Deforestation trend and causes of the acceleration of the
-  deforestation since 2005}
-
-In our study, we have shown that the progressive decrease of the
-deforestation rate on the period 1990-2005 was followed by a continuous
-increase in the deforestation rate on the period 2005-2014. In
-particular, we showed that deforestation rate has more than doubled on
-the period 2010-2014 compared to 2000-2005. Our results are confirmed by
-previous studies \citep{Harper2007, MEFT2009, ONE2015} despite
-differences in the methodologies regarding (i) forest definition
-(associated to independent visual interpretations of observation
-polygons to train the classifier), (ii) classification algorithms, (iii)
-deforestation rate computation method, and (iv) correction for the
-presence of clouds. Our deforestation rate estimates from 1990 to 2014
-have been computed from wall to wall maps at 30 m resolution and can be
-considered more accurate in comparison with estimates from these
-previous studies. Our forest-cover and deforestation rate estimates can
-be used as source of information for the next FAO Forest Resources
-Assessment project \citep{Keenan2015}. Current rates of deforestation
-can also be used to build reference scenarios for deforestation in
-Madagascar and contribute to the implementation of deforestation
-mitigation activities in the framework of REDD+ \citep{Olander2008}.
-
-The increase of deforestation rates after 2005 can be explained by
-population growth and political instability in the country. Nearly 90\%
-of Madagascar's population relies on biomass for their daily energy
-needs \citep{Minten2013} and the link between population size and
-deforestation has previously been demonstrated in Madagascar
-\citep{Vieilledent2013, Gorenflo2011}. With a mean demographic growth
-rate of about 2.8\%/yr and a population which has increased from 16 to
-24 million people on the period 2000-2015 \citep{UN2015}, the increasing
-demand in wood-fuel and space for agriculture is likely to explain the
-increase in deforestation rates. The political crisis of 2009
-\citep{Ploch2012}, followed by several years of political instability
-and weak governance could also explain the increase in the deforestation
-rate observed on the period 2005-2014 (Smith et al., 2003). These
-results show that despite the conservation policy in Madagascar
-\citep{Freudenberger2010}, deforestation has dramatically increased at
-the national level since 2005. More efforts should be done to identify
-the real causes and mechanisms explaining this rampant deforestation in
-order to implement new conservation strategies that will be effective at
-conserving the few remaining natural tropical forests in Madagascar.
-
-\newpage
-
-\section{Author's contribution}
-\label{authors-contribution}
-
-All authors conceived the ideas and designed methodology; GV analysed
-the data and wrote the {\R}/GRASS script; GV drafted the manuscript. All
-authors contributed critically to the drafts and gave final approval for
-publication.
-
-\section{Acknowledgements}
-\label{acknowledgements}
-
-This study is part of the Cirad's BioSceneMada project
-(\url{https://bioscenemada.cirad.fr}) and the Joint Research Center's
-ReCaREDD project (\url{http://forobs.jrc.ec.europa.eu/recaredd}). The
-BioSceneMada project is funded by FRB (Fondation pour la Recherche sur
-la Biodiversité) and the FFEM (Fond Français pour l'Environnement
-Mondial) under the project agreement AAP-SCEN-2013 I. The ReCaREDD
-project is funded by the European Commission.
-
-\section{Data accessibility}
-\label{data-accessibility}
-
-All the data and codes used for this study are made publicly available
-in the \texttt{deforestmap} GitHub repository
-(\url{https://github.com/ghislainv/deforestmap.git}). The results and
-the manuscript are fully reproducible running the {\R} script
-\texttt{deforestmap.R} located inside the \texttt{deforestmap}
-repository.
-
-\newpage
-\singlespacing
-
-\bibliographystyle{gcb}
-\bibliography{bib/biblio}
-
-\newpage
-
-\section{Tables}
-\label{tables}
-
-\nopagebreak
-
-\vfill
-\begin{table}[!h]
-  \begin{longtable}[]{@{}rrrrr@{}}
-    \toprule
-    Year & Forest (ha) & Unmap (ha) & Annual defor. (ha/yr) & Rate
-    (\%/yr)\tabularnewline
-    \midrule
-    \endhead
-    1953 & 15968176 & 0 & - & -\tabularnewline
-    1973 & 14242592 & 3316531 & 86279 & 0.57\tabularnewline
-    1990 & 10762442 & 0 & 204715 & 1.63\tabularnewline
-    2000 & 9879031 & 0 & 88341 & 0.85\tabularnewline
-    2005 & 9667553 & 0 & 42296 & 0.43\tabularnewline
-    2010 & 9319851 & 0 & 69540 & 0.73\tabularnewline
-    2014 & 8925246 & 0 & 98651 & 1.08\tabularnewline
-    \bottomrule
-  \end{longtable}
-  \addtocounter{table}{-1}
-
-  \caption{\textbf{Evolution of the forest cover and deforestation rates
-      from 1953 to 2014 in Madagascar}. Forest map for the year 1973 has 3.3
-    Mha of unclassified areas due to the presence of clouds on satellite
-    images. As a consequence, deforestation rates for the periods 1953-1973
-    and 1973-1990 are indicative. The two last columns indicate the
-    deforestation rates on the previous time-period (e.g.~1953-1973 for year
-    1973, 1973-1990 for year 1990, etc.).}
-
-  \label{tab:forest_cover}
-\end{table}
-\vfill
-
-\newpage
-
-\vspace*{\stretch{1}}
-\begin{table}[!h]
-  {\footnotesize
-  \begin{longtable}[]{@{}llrrrrrrrr@{}}
-    \toprule
-    Forest type & Source & 1953 & 1973 & 1990 & 2000 & 2005 & 2010 & 2013 &
-    2014\tabularnewline
-    \midrule
-    \endhead
-    Total & Harper2007 & 15995900 & 14173100 & 10605700 & 8982100 & - & - &
-    - & -\tabularnewline
-    & MEFT2009 & - & - & 10650142 & 9678402 & 9413218 & - & - &
-    -\tabularnewline
-    & ONE2015 & - & - & - & - & 9451350 & 8977337 & 8485509 &
-    -\tabularnewline
-    & this study & 15968176 & 14242592 & 10762494 & 9879031 & 9667553 &
-    9319851 & 9051029 & 8925246\tabularnewline
-    Moist & Harper2007 & 8765600 & 6876000 & 5234300 & 4166800 & - & - & - &
-    -\tabularnewline
-    & MEFT2009 & - & - & 5270599 & 4787771 & 4700430 & - & - &
-    -\tabularnewline
-    & ONE2015 & - & - & - & - & 4555788 & 4457184 & 4345093 &
-    -\tabularnewline
-    & this study & 8578299 & 6989942 & 5270169 & 4872016 & 4767876 & 4633104
-    & 4470194 & 4409842\tabularnewline
-    Dry & Harper2007 & 4252100 & 4027700 & 2711800 & 2457000 & - & - & - &
-    -\tabularnewline
-    & MEFT2009 & - & - & 3320582 & 3084976 & 3027505 & - & - &
-    -\tabularnewline
-    & ONE2015 & - & - & - & - & 3223028 & 2970192 & 2678640 &
-    -\tabularnewline
-    & this study & 4761551 & 4434871 & 3224917 & 2940970 & 2880819 & 2734639
-    & 2642253 & 2595621\tabularnewline
-    Spiny & Harper2007 & 2978200 & 3029800 & 2420000 & 2132200 & - & - & - &
-    -\tabularnewline
-    & MEFT2009 & - & - & 2123630 & 1871735 & 1756884 & - & - &
-    -\tabularnewline
-    & ONE2015 & - & - & - & - & 1681527 & 1558533 & 1466765 &
-    -\tabularnewline
-    & this study & 2462830 & 2582880 & 2054724 & 1857628 & 1810704 & 1744427
-    & 1731308 & 1712731\tabularnewline
-    Mangroves & Harper2007 & - & - & 239600 & 226100 & - & - & - &
-    -\tabularnewline
-    & MEFT2009 & - & - & - & - & - & - & - & -\tabularnewline
-    & ONE2015 & - & - & - & - & 173564 & 171220 & 169877 & -\tabularnewline
-    & this study & 143412 & 199853 & 181226 & 177708 & 177492 & 177149 &
-    176890 & 176718\tabularnewline
-    \bottomrule
-  \end{longtable}}
-  \addtocounter{table}{-1}
-
-  \caption{\textbf{Comparing our estimates of forest-cover (in ha) for
-      Madagascar with previous studies on the period 1953-2014}. We compared
-    our forest-cover with the estimates from three previous studies
-    \citep{Harper2007, MEFT2009, ONE2015}. We computed a The increase in
-    mangrove and spiny forest covers from \emph{c.} 1953 to \emph{c.} 1973
-    in \citet{Harper2007} and our study is most probably due to differences
-    in forest definition and mapping methods between the 1953
-    aerial-photography derived map and the 1973 Landsat image derived map.}
-
-  \label{tab:comp_forest}
-\end{table}
-\vspace*{\stretch{1}}
-
-\newpage
-
-\vspace*{\stretch{1}}
-\begin{table}[!h]
-  {\footnotesize
-  \begin{longtable}[]{@{}llrrrrrr@{}}
-    \toprule
-    Forest type & Source & 1953-1973 & 1973-1990 & 1990-2000 & 2000-2005 &
-    2005-2010 & 2010-2013\tabularnewline
-    \midrule
-    \endhead
-    Total & Harper2007 & 91140 (0.30) & 200206 (1.70) & 80740 (0.90) & - & -
-    & -\tabularnewline
-    & MEFT2009 & - & - & 97174 (0.83) & 53037 (0.53) & - & -\tabularnewline
-    & ONE2015 & - & - & - & - & 94803 (1.18) & 163943 (1.50)\tabularnewline
-    & this study & 86279 (0.57) & 204712 (1.63) & 88346 (0.85) & 42296
-    (0.43) & 69540 (0.73) & 89607 (0.97)\tabularnewline
-    Moist & Harper2007 & 94480 (0.60) & 87188 (1.70) & 32200 (0.80) & - & -
-    & -\tabularnewline
-    & MEFT2009 & - & - & 48283 (0.79) & 17468 (0.35) & - & -\tabularnewline
-    & ONE2015 & - & - & - & - & 19721 (0.50) & 37364 (0.94)\tabularnewline
-    & this study & 79418 (1.02) & 101163 (1.65) & 39815 (0.78) & 20828
-    (0.43) & 26954 (0.57) & 54303 (1.19)\tabularnewline
-    Dry & Harper2007 & 11220 (0.20) & 77153 (1.90) & 19820 (0.70) & - & - &
-    -\tabularnewline
-    & MEFT2009 & - & - & 23561 (0.67) & 11494 (0.40) & - & -\tabularnewline
-    & ONE2015 & - & - & - & - & 50567 (1.80) & 97184 (2.29)\tabularnewline
-    & this study & 16334 (0.35) & 71174 (1.86) & 28395 (0.92) & 12030 (0.41)
-    & 29236 (1.04) & 30795 (1.14)\tabularnewline
-    Spiny & Harper2007 & -2580 (-0.10) & 35865 (1.20) & 28170 (1.20) & - & -
-    & -\tabularnewline
-    & MEFT2009 & - & - & 25190 (1.19) & 22970 (1.23) & - & -\tabularnewline
-    & ONE2015 & - & - & - & - & 24599 (1.69) & 30589 (1.66)\tabularnewline
-    & this study & -6002 (-0.24) & 31068 (1.34) & 19710 (1.00) & 9385 (0.51)
-    & 13255 (0.74) & 4373 (0.25)\tabularnewline
-    Mangroves & Harper2007 & - & - & 550 (0.20) & - & - & -\tabularnewline
-    & MEFT2009 & - & - & - & - & - & -\tabularnewline
-    & ONE2015 & - & - & - & - & 469 (0.32) & 448 (0.20)\tabularnewline
-    & this study & -2822 (-1.67) & 1096 (0.57) & 352 (0.20) & 43 (0.02) & 69
-    (0.04) & 86 (0.05)\tabularnewline
-    \bottomrule
-  \end{longtable}}
-  \addtocounter{table}{-1}
-
-  \caption{\textbf{Comparing our estimates of annual deforestation
-      rates for Madagascar with previous studies on the period
-      1953-2014}. Annual deforestation areas (in ha/yr) and annual
-    deforestation rates (second number in parenthesis, in \%/yr) are
-    provided. For deforestation rates in \%/yr, exact same numbers as
-    in scientific articles and reports from previous studies have been
-    reported. The way annual deforestation rates in \%/yr have been
-    computed in these previous studies can slightly differ from one
-    study to another but estimates always correct for the potential
-    presences of clouds on forest maps. Annual deforested areas in
-    ha/yr have been recomputed from forest-cover estimates in
-    Tab.~\ref{tab:comp_forest} (except for \citet{Harper2007} for the
-    periods 1973-1990 and 1990-2000 for which annual deforested ares
-    in ha/yr were derived from numbers reported in the original
-    publication, see methods) and do not correct for the potential
-    presence of clouds.}
-
-  \label{tab:comp_defor}
-\end{table}
-\vspace*{\stretch{1}}
-
-\newpage
-
-\vspace*{\stretch{1}}
-\begin{table}[!h]
-
-  {\small
-  \begin{longtable}[]{@{}rrrrrrrr@{}}
-    \toprule
-    Year & Forest (ha) & patch (\%) & transitional (\%) & edge (\%) &
-    perforated (\%) & interior (\%) & NA (\%)\tabularnewline
-    \midrule
-    \endhead
-    1953 & 15962870 & 0.01 & 1.12 & 4.46 & 0.58 & 93.83 &
-    0.00\tabularnewline
-    1973 & 14228217 & 2.21 & 7.25 & 19.81 & 2.86 & 67.87 &
-    0.01\tabularnewline
-    1990 & 10749572 & 3.00 & 8.17 & 21.28 & 3.81 & 63.73 &
-    0.01\tabularnewline
-    2000 & 9866145 & 3.09 & 8.37 & 22.13 & 3.92 & 62.49 &
-    0.01\tabularnewline
-    2005 & 9659861 & 3.51 & 8.88 & 22.56 & 6.44 & 58.59 &
-    0.02\tabularnewline
-    2010 & 9306528 & 4.28 & 9.72 & 22.94 & 8.52 & 54.52 &
-    0.02\tabularnewline
-    2014 & 8911481 & 5.18 & 10.72 & 23.25 & 10.58 & 50.24 &
-    0.03\tabularnewline
-    \bottomrule
-  \end{longtable}}
-  \addtocounter{table}{-1}
-
-  \caption{\textbf{Evolution of the forest fragmentation from 1953 to 2014
-      in Madagascar}. Six categories of fragmentation were identified from the
-    amount of forest and its occurrence as adjacent forest pixels:
-    ``interior'', ``perforated'', ``edge'', ``transitional'', ``patch'', and
-    ``undetermined'' \citep{Riitters2000}. We used a moving window of 7x7
-    pixels (4.4 ha). Using this window size, forest edge had a width of
-    about 90m. The ``interior'' category can be interpreted as the most
-    intact forest. The ``patch'' and ``transitional'' categories correspond
-    to isolated small forest patches.}
-
-  \label{tab:frag}
-\end{table}
-\vspace*{\stretch{1}}
-
-\newpage
-
-\section{Figures}
-\label{figures}
-
-\nopagebreak
-
-\vfill
-\begin{figure}[h!]
+\texttt{gdal\_proximity.py} of the GDAL library
+(\url{http://www.gdal.org/}). We computed the mean and 90\% quantiles
+(5\% and 95\%) of the distance to forest edge and looked at the
+evolution of these values with time.
+
+\begin{figure*}[h]
   \centering
 
-    \includegraphics[width=11cm]{outputs/ecoregion.png}
-
-    \caption{\textbf{Ecoregions and forest types in Madagascar.}
-      Madagascar can be divided into four climatic ecoregions with four forest
-      types: the moist forest in the East (green), the dry forest in the West
-      (orange), the spiny forest in the South (red), and the mangroves on the
-      West coast (blue). Ecoregions were defined following climatic
-      \citep{Cornet1974} and vegetation \citep{IEFN1996} criteria. The dark
-      grey areas represent the remaining natural forest cover for the year
-      2014.}
-
-    \label{fig:ecoregion}
-
-\end{figure}
-\vfill
-
-\newpage
-
-\vspace*{\stretch{1}}
-\begin{figure}[h!]
-  \centering
-
-  \includegraphics[width=\textwidth]{outputs/fig_fcc_highres.png}
+  \includegraphics[width=\textwidth]{fig_fcc_highres.png}
 
@@
 
+\end{figure*}
+
+\section{Results}
+\label{results}
+
+\subsection{Dynamics of forest cover and deforestation
+intensity}
+
+Natural forests in Madagascar covered 16.0 Mha in 1953, about 27\% of
+the national territory of 587,041 km2. In 2014, the forest cover
+dropped to 8.9 Mha, corresponding to about 15\% of the national
+territory (Fig.~\ref{fig:fcc} and
+Tab.~\ref{tab:forest_cover}). Madagascar has lost 44\% and 37\% of its
+natural forests between 1953 and 2014, and between 1973 and 2014
+respectively (Fig.~\ref{fig:fcc} and Tab.~\ref{tab:forest_cover}). In
+2014 the remaining 8.9 Mha of natural forest were distributed as follow: 4.4
+Mha of moist forest (50\% of total forest cover), 2.6 Mha of dry
+forest (29\%), 1.7 Mha of spiny forest (19\%) and 0.18 Mha (2\%) of
+mangrove forest (Fig.~\ref{fig:ecoregion} and
+Tab.~\ref{tab:comp_forest}). Regarding the deforestation trend, we
+observed a progressive decrease of the deforestation rate after 1990
+from 205,000 ha/yr (1.6\%/yr) over the period 1973-1990 to 42,000
+ha/yr (0.4\%/yr) over the period 2000-2005
+(Tab.~\ref{tab:forest_cover}). Then from 2005, the deforestation rate
+has progressively increased and has more than doubled over the period
+2010-2014 (99,000 ha/yr, 1.1\%/yr) compared to 2000-2005
+(Tab.~\ref{tab:forest_cover}). The deforestation trend characterized
+by a progressive decrease of the deforestation rate over the period
+1990-2005 and a progressive increase of the deforestation after 2005
+is valid for all four ecoregions (Tab.~\ref{tab:comp_defor}), with the
+exception of the spiny forest domain for which the deforestation rate
+during the period 2010-2013 was lower than during 2005-2010
+(Tab.~\ref{tab:comp_defor}).
+
+\begin{table}[h]
+
+  \caption{\textbf{Evolution of forest cover and deforestation
+      rates from 1953 to 2014 in Madagascar}.}
+  \label{tab:forest_cover}
+
+  \begin{tabular}{@{}lrrrr@{}}
+    \tophline
+    Year & Forest & Unmapped & Annual defor. & Rate \\
+    & (Kha) & (Kha) & (Kha/yr) & (\%/yr) \\
+    \middlehline
+    1953 & 15,968 & 0 & - & - \\
+    1973 & 14,243 & 3,317 & 86 & 0.6 \\
+    1990 & 10,762 & 0 & 205 & 1.6 \\
+    2000 & 9,879 & 0 & 88 & 0.8 \\
+    2005 & 9,668 & 0 & 42 & 0.4 \\
+    2010 & 9,320 & 0 & 70 & 0.7 \\
+    2014 & 8,925 & 0 & 99 & 1.1 \\
+    \bottomhline
+  \end{tabular}
+
+  \belowtable{Areas are provided in thousands of hectares
+    (Kha). Forest map for the year 1973 has 3.3 Mha of unclassified
+    areas due to the presence of clouds on satellite images. As a
+    consequence, partial deforestation rates for the periods 1953-1973
+    and 1973-1990 are computed based on the available forest
+    extent. The last two columns indicate the annual deforested areas
+    and annual deforestation rates on the previous time-period
+    (e.g.~1953-1973 for year 1973, 1973-1990 for year 1990, etc.).}
+
+\end{table}
+
+\subsection{Comparison with previous forest-cover change studies
+in Madagascar}
+
+Forest-cover maps provided by previous studies over Madagascar were
+not exhaustive (unclassified areas) due to the presence of clouds on
+satellite images used to produce such maps. In \citet{Harper2007}, the
+maps of years 1990 and 2000 include 0.5 and 1.12 Mha of unknown cover
+type respectively. Proportions of unclassified areas are not reported
+in the two other existing studies at the national level by
+\citet{MEFT2009} and \citet{ONE2015}. With our approach, we produced
+wall to wall forest-cover change maps from 1990 to 2014 for the full
+territory of Madagascar (Tab.~\ref{tab:forest_cover}). This allowed us
+to produce more robust estimates of forest-cover and deforestation
+rates over this period. Our forest-cover estimates over the period
+1953-2013 (considering forest cover estimates at national level and by
+ecoregions for all the available dates) were well correlated
+(Pearson's correlation coefficient $=$ 0.99) to estimates from the
+three previous studies (Tab.~\ref{tab:comp_forest}) with a RMSE of
+300,000 ha (6\% of the mean forest cover of 4.8 Mha when considering
+all dates and forest types together). These small differences can be
+partly attributed to differences in ecoregion boundaries. Despite
+significant differences in deforestation estimates
+(Tab.~\ref{tab:comp_defor}), a similar deforestation trend was
+observed across studies with a decrease of deforestation rates over
+the period 1990-2005, followed by a progressive increase of the
+deforestation after 2005.
+
+\begin{table*}[h]
+
+  \caption{\textbf{Comparing Madagascar forest-cover estimates (in Kha)
+      with previous studies on the period 1953-2014}.}
+  \label{tab:comp_forest}
+
+  \begin{tabular}{@{}llrrrrrrrr@{}}
+    \tophline
+    Forest type & Source & 1953 & 1973 & 1990 & 2000 & 2005 & 2010 & 2013 &
+                                                                            2014 \\
+    \middlehline
+    Total & Harper2007 & 15,996 & 14,173 & 10,606 & 8,982 & - & - & - & - \\
+          & MEFT2009 & - & - & 10,650 & 9,678 & 9,413 & - & - & - \\
+                & ONE2015 & - & - & - & - & 9,451 & 8,977 & 8,486 & - \\
+                & this study & 15,968 & 14,243 & 10,762 & 9,879 & 9,668 & 9,320 & 9,051
+                                                                          & 8,925 \\
+    Moist & Harper2007 & 8,766 & 6,876 & 5,234 & 4,167 & - & - & - &
+                                                                     - \\
+                & MEFT2009 & - & - & 5,271 & 4,788 & 4,700 & - & - & - \\
+                & ONE2015 & - & - & - & - & 4,556 & 4,457 & 4,345 & - \\
+                & this study & 8,578 & 6,990 & 5,270 & 4,872 & 4,768 & 4,633 & 4,470 &
+                                                                                       4,410 \\
+    Dry & Harper2007 & 4,252 & 4,028 & 2,712 & 2,457 & - & - & - &
+                                                                   - \\
+                & MEFT2009 & - & - & 3,321 & 3,085 & 3,028 & - & - & - \\
+                & ONE2015 & - & - & - & - & 3,223 & 2,970 & 2,679 & - \\
+                & this study & 4,762 & 4,435 & 3,225 & 2,941 & 2,881 & 2,735 & 2,642 &
+                                                                                       2,596 \\
+    Spiny & Harper2007 & 2,978 & 3,030 & 2,420 & 2,132 & - & - & - &
+                                                                     - \\
+                & MEFT2009 & - & - & 2,124 & 1,872 & 1,757 & - & - & - \\
+                & ONE2015 & - & - & - & - & 1,682 & 1,559 & 1,467 & - \\
+                & this study & 2,463 & 2,583 & 2,055 & 1,858 & 1,811 & 1,744 & 1,731 &
+                                                                                       1,713 \\
+    Mangroves & Harper2007 & - & - & 240 & 226 & - & - & - &
+                                                             - \\
+                & MEFT2009 & - & - & - & - & - & - & - & - \\
+                & ONE2015 & - & - & - & - & 174 & 171 & 170 & - \\
+                & this study & 143 & 200 & 181 & 178 & 177 & 177 & 177 &
+                                                                         177 \\
+    \bottomhline
+  \end{tabular}
+
+  \belowtable{We compared our estimates of forest-cover with the
+    estimates from three previous studies \citep{Harper2007, MEFT2009,
+      ONE2015}. Areas are provided in thousands of hectares (Kha). We
+    obtained a Pearson's correlation coefficient of 0.99 between our
+    forest-cover estimates and forest-cover estimates from previous
+    studies. The increase in mangrove and spiny forest covers from
+    \emph{c.} 1953 to \emph{c.} 1973 in \citet{Harper2007} and our
+    study is most probably due to differences in forest definition and
+    mapping methods between the 1953 aerial-photography derived map
+    and the 1973 Landsat image derived map.}
+
+\end{table*}
+
+\begin{table*}[h]
+
+  \caption{\textbf{Comparing Madagascar annual deforestation rates
+      with previous studies on the period 1953-2014}.}
+  \label{tab:comp_defor}
+
+  \begin{tabular}{@{}llrrrrrr@{}}
+    \tophline
+    Forest type & Source & 1953-1973 & 1973-1990 & 1990-2000 & 2000-2005 &
+                                                                           2005-2010 & 2010-2013\\
+    \middlehline
+    Total & Harper2007 & 91 (0.3) & 200 (1.7) & 81 (0.9) & - & - &
+                                                                   - \\
+                & MEFT2009 & - & - & 97 (0.8) & 53 (0.5) & - & - \\
+                & ONE2015 & - & - & - & - & 95 (1.2) & 164 (1.5) \\
+                & this study & 86 (0.6) & 205 (1.6) & 88 (0.9) & 42 (0.4) & 70 (0.7) &
+                                                                                       90 (1.0) \\
+    Moist & Harper2007 & 94 (0.6) & 87 (1.7) & 32 (0.8) & - & - &
+                                                                  - \\
+                & MEFT2009 & - & - & 48 (0.8) & 17 (0.4) & - & - \\
+                & ONE2015 & - & - & - & - & 20 (0.5) & 37 (0.9) \\
+                & this study & 79 (1.0) & 101 (1.6) & 40 (0.8) & 21 (0.4) & 27 (0.6) &
+                                                                                       54 (1.2) \\
+    Dry & Harper2007 & 11 (0.2) & 77 (1.9) & 20 (0.7) & - & - &
+                                                                - \\
+                & MEFT2009 & - & - & 24 (0.7) & 11 (0.4) & - & - \\
+                & ONE2015 & - & - & - & - & 51 (1.8) & 97 (2.3) \\
+                & this study & 16 (0.4) & 71 (1.9) & 28 (0.9) & 12 (0.4) & 29 (1.0) & 31
+                                                                                      (1.1) \\
+    Spiny & Harper2007 & -3 (-0.1) & 36 (1.2) & 28 (1.2) & - & - &
+                                                                   - \\
+                & MEFT2009 & - & - & 25 (1.2) & 23 (1.2) & - & - \\
+                & ONE2015 & - & - & - & - & 25 (1.7) & 31 (1.7) \\
+                & this study & -6 (-0.2) & 31 (1.3) & 20 (1.0) & 9 (0.5) & 13 (0.7) & 4
+                                                                                      (0.3) \\
+    Mangroves & Harper2007 & - & - & 1 (0.2) & - & - & - \\
+                & MEFT2009 & - & - & - & - & - & - \\
+                & ONE2015 & - & - & - & - & 0 (0.3) & 0 (0.2) \\
+                & this study & -3 (-1.7) & 1 (0.6) & 0 (0.2) & 0 (0.0) & 0 (0.0) & 0
+                                                                                   (0.0) \\
+    \bottomhline
+  \end{tabular}
+
+  \belowtable{Annual deforestated areas (in Kha/yr) and annual
+    deforestation rates (second number in parenthesis, in \%/yr) are
+    provided. For deforestation rates in \%/yr, exact same numbers as
+    in scientific articles and reports from previous studies
+    \citep{Harper2007, MEFT2009, ONE2015} have been reported. The way
+    annual deforestation rates in \%/yr have been computed in these
+    previous studies can slightly differ from one study to another, but
+    estimates always correct for the potential presences of clouds on
+    satellite images and unclassified areas on forest maps. Annual
+    deforested areas in Kha/yr have been recomputed from forest-cover
+    estimates in Tab.~\ref{tab:comp_forest} (except for
+    \citet{Harper2007} for the periods 1973-1990 and 1990-2000 for
+    which annual deforested areas in Kha/yr were derived from numbers
+    reported in the original publication, see methods) and do not
+    correct for the potential presence of clouds.}
+
+\end{table*}
+
+\subsection{Evolution of forest fragmentation with time}
+
+In parallel to the dynamics of deforestation, forest fragmentation has
+progressively increased since 1953 in Madagascar. We observed a
+continuous decrease of the mean distance to forest edge from 1953 to
+2014 in Madagascar. The mean distance to forest edge has decreased to
+\emph{c.} 300~m in 2014 while it was previously \emph{c.} 1.5~km in 1973
+(Fig.~\ref{fig:dist_edge}). Moreover, a large proportion (73\%) of
+the forest was located at a distance greater than 100 m in 1973, while
+almost half of the forest (46\%) was at a distance lower than 100 m
+from forest edge in 2014 (Fig.~\ref{fig:dist_edge}). The percentage of
+forest that can be considered intact in Madagascar has continuously
+decreased since 1953. The percentage of forest belonging to the
+``interior'' category (most intact forests) has fallen from 68\% in
+1973 to 50\% in 2014. In 2014, more than 16\% of the forest belonged
+to the ``patch'' and ``transitional'' categories (isolated small
+forest patches) compared to 9\% in 1973 (Tab.~\ref{tab:frag}).
+
+\begin{figure}[h]
+  \centering
+
+  \includegraphics[width=8.3cm]{dist.png}
+
+  \caption{\textbf{Evolution of the distance to forest edge from 1953
+      to 2014 in Madagascar.} Black dots represent the mean distance
+    to forest edge for each year. Vertical dashed segments represent
+    the 90\% quantiles (5\% and 95\%) of the distance to forest
+    edge. Horizontal dashed grey line indicates a distance to forest
+    edge of 100~m.  Numbers at the bottom of each vertical segments are
+    the percentage of forest at a distance to forest edge lower than
+    100~m for each year.}
+
+  \label{fig:dist_edge}
+
 \end{figure}
-\vspace*{\stretch{1}}
-
-\newpage
-
-\vspace*{\stretch{1}}
-\begin{figure}[h!]
-  \centering
-
-  \includegraphics[width=13cm]{outputs/dist.png}
-
-  \caption{\textbf{Evolution of the distance to forest edge from
-      1953 to 2014 in Madagascar.} Black dots represent the mean distance to
-    forest edge for each year. Vertical dashed segments represent the 90\%
-    quantiles (5\% and 95\%) of the distance to forest edge. Horizontal
-    dashed grey line indicates a distance to forest edge of 100~m.
-    Percentages indicate the percentage of forest at a distance to forest
-    edge lower than 100~m for each year.}
-
-  \label{fig:dist_edge}
-
-\end{figure}
-\vspace*{\stretch{1}}
+
+\section{Discussion}
+\label{discussion}
+
+\subsection{Benefits of the combined use of recent global annual
+  tree cover loss data with historical national forest-cover maps}
+
+In this study, we combined recent (2001-2014) global annual tree cover
+loss data \citep{Hansen2013} with historical (1953-2000) national
+forest-cover maps \citep{Harper2007} to look at six decades
+(1953-2014) of deforestation and forest fragmentation in
+Madagascar. We produced annual forest-cover maps at 30~m resolution
+covering Madagascar for the period 2000 to 2014. Our study extends the
+forest-cover monitoring on a six decades period (from 1953 to 2014)
+while harmonizing the data from previous studies \citep{Harper2007,
+  MEFT2009, ONE2015}. We propose a generic approach to solve the
+problem of forest definition which is needed to transform the 2000
+global tree cover dataset from \citet{Hansen2013} into a
+forest/non-forest map \citep{Tropek2014}. We propose to use a
+historical national forest-cover map, based on a national forest
+definition, as a forest cover mask. This approach could be easily
+extended to other regions or countries for which an accurate
+forest-cover map is available at any date within the period 2000-2014,
+but preferably at the beginning of the period to profit from the full
+record and derive long-term estimates of deforestation. Moreover, this
+approach can be repeated in the future if and when the global tree
+cover product is updated. We have made the \R{}/GRASS code used for this
+study freely available in a GitHub repository (see Data availability
+statement) to facilitate application to other study areas or repeat
+the analysis in the future for Madagascar.
+
+The accuracy of the derived forest-cover change maps depends directly
+on the accuracies of the historical forest-cover maps and the tree
+cover loss dataset. Using visual-interpretation of aerial images in
+342 areas distributed among all forest types, \citet{Harper2007}
+estimated an overall 89.5\% accuracy in identifying forest/non-forest
+classes for the year 2000. The accuracy assessment of the tree cover
+loss dataset for the tropical biome reported 13\% of false positives
+and 16.9\% of false negatives (see Tab. S5 in
+\citet{Hansen2013}). These numbers rise at 20.7\% and 20.6\%
+respectively for the subtropical biome. In the subtropical biome, the
+lower density tree cover canopy makes it difficult to detect change
+from tree cover to bare ground. For six countries in Central Africa,
+with a majority of moist dense forest, \citet{Verhegghen2016} have
+compared deforestation estimates derived from the global tree cover
+loss dataset \citep{Hansen2013} with results derived from
+semi-automated supervised classification of Landsat satellite images
+\citep{Achard2014} and they found a good agreement between the two
+sets of estimates. Therefore, our forest-cover change maps after 2000
+might be more accurate for the dense moist forest than for the dry and
+spiny forest. In another study assessing the accuracy of the tree
+cover loss product accross the tropics \citep{Tyukavina2015}, authors
+reported 4\% of false positives and 48\% of false negatives in
+Sub-Saharian Africa. They showed that 85\% of missing loss occured on
+the edges of other loss patches. This means that tree cover loss might
+be underestimated in Sub-Saharian Africa, probably due to the
+prevalence of small-scale disturbance which is hard to map at 30~m,
+but that areas of large-scale deforestation are well identified and
+spatial variability of the deforestation is well represented. A proper
+accuracy assessment of our forest-cover change maps should be
+performed to better estimate the uncertainty surrounding our
+forest-cover change estimates in Madagascar from year 2000
+\citep{Olofsson2013,Olofsson2014}. Despite this limitation, we have
+shown that the deforestation trend we observed for Madagascar, with a
+doubling deforestation on the period 2010-2014 compared to 2000-2005,
+was consistent with the other studies at the national scale
+\citep{ONE2015, MEFT2009}.
+
+Consistent with \citet{Harper2007}, we did not consider potential
+forest regrowth in Madagascar (although \citet{Hansen2013} provided a
+tree cover gains layer for the period 2001-2013) for several
+reasons. First, the tree gain layer of \citet{Hansen2013} includes and
+catches more easily tree plantations than natural forest regrowth
+\citep{Tropek2014}. Second, there is little evidence of natural forest
+regeneration in Madagascar \citep{Grouzis2001, Harper2007}. This can
+be explained by several ecological processes following burning
+practice such as soil erosion \citep{Grinand2017} and reduced seed
+bank due to fire and soil loss \citep{Grouzis2001}. Moreover, in areas
+where forest regeneration is ecologically possible, young forest
+regrowth are more easily re-burnt for agriculture and pasture. Third,
+young secondary forests provide more limited ecosystem services
+compared to old-growth natural forests in terms of biodiversity and
+carbon storage.
+
+\subsection{Dynamics of forest-cover in Madagascar from 1953 to 2014}
+
+We estimated that natural forests in Madagascar cover 8.9 Mha in 2014
+(corresponding to 15\% of the country) and that Madagascar has lost
+44\% of its natural forest since 1953 (37\% since 1973). There is
+ongoing scientific debate about the extent of the ``original'' forest
+cover in Madagascar, and the extent to which humans have altered the
+natural forest landscapes since their large-scale settlement around
+800 CE \citep{Burns2016, Cox2012}. Early French naturalists stated
+that the full island was originally covered by forest
+\citep{Humbert1927, Perrier1921}, leading to the common statement that
+90\% of the natural forests have disappeared since the arrival of
+humans on the island \citep{Kull2000}. More recent studies
+counter-balanced that point of view saying that extensive areas of
+grassland existed in Madagascar long before human arrival and were
+determined by climate, natural grazing and other natural factors
+\citep{Vorontsova2017, Virah-Sawmy2009}. Other authors have questioned
+the entire narrative of extensive alteration of the landscape by early
+human activity which, through legislation, has severe consequences on
+local people \citep{Klein2002, Kull2000}. Whatever the original
+proportion of natural forests and grasslands in Madagascar, our
+results demonstrate that human activities since the 1950s have
+profoundly impacted the natural tropical forests and that conservation
+and development programs in Madagascar have failed to stop
+deforestation in the recent years. Deforestation has strong
+consequences on biodiversity and carbon emissions in
+Madagascar. Around 90\% of Madagascar's species are forest dependent
+\citep{Allnutt2008, Goodman2005} and \citet{Allnutt2008} estimated
+that deforestation between 1953 and 2000 led to an extinction of 9\%
+of the species. The additional deforestation we observed over the
+period 2000-2014 (around 1Mha of natural forest) worsen this
+result. Regarding carbon emissions, using the 2010 aboveground forest
+carbon map by \citet{Vieilledent2016}, we estimated that deforestation
+on the period 2010-2014 has led to 40.2 Mt C of carbon emissions in
+the atmosphere (10 Mt C /yr) and that the remaining aboveground forest
+carbon stock in 2014 is 832.8 Mt C. Associated to deforestation, we
+showed that the remaining forests of Madagascar are highly fragmented
+with 46\% of the forest being at less than 100~m of the forest
+edge. Small forest fragments do not allow to maintain viable
+populations and ``edge effects'' at forest/non-forest interfaces have
+impacts on both carbon emissions \citep{Brinck2017} and biodiversity
+loss \citep{Gibson2013, Murcia1995}.
+
+\begin{table}[h]
+
+  \caption{\textbf{Evolution of the forest fragmentation from 1953 to 2014
+      in Madagascar}.}
+  \label{tab:frag}
+
+  \begin{tabular}{@{}lrrrrrr@{}}
+    \tophline
+    Year & Forest & patch & transit. & edge & perfor. & interior \\
+    & (Kha) & (\%) & (\%) & (\%) & (\%) & (\%) \\
+    \middlehline
+    1953 & 15,963 & 0 & 1 & 4 & 1 & 94 \\
+    1973 & 14,228 & 2 & 7 & 20 & 3 & 68 \\
+    1990 & 10,750 & 3 & 8 & 21 & 4 & 64 \\
+    2000 & 9,866 & 3 & 8 & 22 & 4 & 62 \\
+    2005 & 9,660 & 4 & 9 & 23 & 6 & 59 \\
+    2010 & 9,307 & 4 & 10 & 23 & 9 & 55 \\
+    2014 & 8,911 & 5 & 11 & 23 & 11 & 50 \\
+    \bottomhline
+  \end{tabular}
+
+  \belowtable{Five categories of fragmentation were identified from
+    the amount of forest and its occurrence as adjacent forest pixels:
+    ``interior'', ``perforated'', ``edge'', ``transitional'', and
+    ``patch'' \citep{Riitters2000}. We used a moving window of 7x7
+    pixels (4.4~ha). Using this window size, forest edge had a width
+    of about 90~m. The ``interior'' category can be interpreted as the
+    most intact forest. The ``patch'' and ``transitional'' categories
+    correspond to isolated small forest patches. Forest areas are
+    provided in thousands of hectares (Kha).}
+
+\end{table}
+
+\subsection{Deforestation trend and impacts on conservation and
+  development policies}
+
+In our study, we have shown that the progressive decrease of the
+deforestation rate on the period 1990-2005 was followed by a
+continuous increase in the deforestation rate on the period
+2005-2014. In particular, we showed that deforestation rate has more
+than doubled on the period 2010-2014 compared to 2000-2005. Our
+results are confirmed by previous studies \citep{Harper2007, MEFT2009,
+  ONE2015} despite differences in the methodologies regarding (i)
+forest definition (associated to independent visual interpretations of
+observation polygons to train the classifier), (ii) classification
+algorithms, (iii) deforestation rate computation method, and (iv)
+correction for the presence of clouds. Our deforestation rate
+estimates from 1990 to 2014 have been computed from wall to wall maps
+at 30~m resolution and can be considered more accurate in comparison
+with estimates from these previous studies. Our forest-cover and
+deforestation rate estimates can be used as source of information for
+the next FAO Forest Resources Assessment \citep{Keenan2015}. Current
+rates of deforestation can also be used to build reference scenarios
+for deforestation in Madagascar and contribute to the implementation
+of deforestation mitigation activities in the framework of REDD+
+\citep{Olander2008}.
+
+The increase of deforestation rates after 2005 can be explained by
+population growth and political instability in the country. Nearly
+90\% of Madagascar's population relies on biomass for their daily
+energy needs \citep{Minten2013} and the link between population size
+and deforestation has previously been demonstrated in Madagascar
+\citep{Vieilledent2013, Gorenflo2011}. With a mean demographic growth
+rate of about 2.8\%/yr and a population which has increased from 16 to
+24 million people on the period 2000-2015 \citep{UN2015}, the
+increasing demand in wood-fuel and space for agriculture is likely to
+explain the increase in deforestation rates. The political crisis of
+2009 \citep{Ploch2012}, followed by several years of political
+instability and weak governance could also explain the increase in the
+deforestation rate observed on the period 2005-2014
+\citep{Smith2003}. These results show that despite the conservation
+policy in Madagascar \citep{Freudenberger2010}, deforestation has
+dramatically increased at the national level since 2005. Results of
+this study, including recent spatially explicit forest-cover change
+maps and forest-cover estimates, should help implement new
+conservation strategies to save Madagascar natural tropical forests
+and their unique biodiversity.
+
+%% The following commands are for the statements about the availability of data sets and/or software code corresponding to the manuscript.
+%% It is strongly recommended to make use of these sections in case data sets and/or software code have been part of your research the article is based on.
+
+%%\codeavailability{TEXT} %% use this section when having only software code available
+
+%%\dataavailability{TEXT} %% use this section when having only data sets available
+
+\codedataavailability{All the data and codes used for this study are
+  made publicly available in the \texttt{deforestmap} GitHub
+  repository (\url{https://github.com/ghislainv/deforestmap.git}). The
+  results are fully reproducible running the \R{} script
+  \texttt{deforestmap.R} located inside the \texttt{deforestmap}
+  repository.} %% use this section when having data sets and software code available
+
+%\appendix
+%\section{}    %% Appendix A
+
+%\subsection{}     %% Appendix A1, A2, etc.
+
+%\noappendix       %% use this to mark the end of the appendix section
+
+%% Regarding figures and tables in appendices, the following two options are possible depending on your general handling of figures and tables in the manuscript environment:
+
+%% Option 1: If you sorted all figures and tables into the sections of the text, please also sort the appendix figures and appendix tables into the respective appendix sections.
+%% They will be correctly named automatically.
+
+%% Option 2: If you put all figures after the reference list, please insert appendix tables and figures after the normal tables and figures.
+%% To rename them correctly to A1, A2, etc., please add the following commands in front of them:
+
+%\appendixfigures  %% needs to be added in front of appendix figures
+
+%\appendixtables   %% needs to be added in front of appendix tables
+
+%% Please add \clearpage between each table and/or figure. Further guidelines on figures and tables can be found below.
+
+\authorcontribution{All authors conceived the ideas and designed methodology; GV analysed
+the data and wrote the \R{}/GRASS script; GV drafted the manuscript. All
+authors contributed critically to the drafts and gave final approval for
+publication.} %% optional section
+
+\competinginterests{Authors declare that they have no conflict of interest.}
+%% this section is mandatory even if you declare that no competing interests are present
+
+%% \disclaimer{TEXT} %% optional section
+
+\begin{acknowledgements}
+
+  Authors thank Jean-François Bastin for useful comments on a previous
+  version of the manuscript. This study is part of the Cirad's
+  BioSceneMada project (\url{https://bioscenemada.cirad.fr}) and the
+  Joint Research Center's ReCaREDD project
+  (\url{http://forobs.jrc.ec.europa.eu/recaredd}). The BioSceneMada
+  project is funded by FRB (Fondation pour la Recherche sur la
+  Biodiversité) and the FFEM (Fond Français pour l'Environnement
+  Mondial) under the project agreement AAP-SCEN-2013 I. The ReCaREDD
+  project is funded by the European Commission.
+
+\end{acknowledgements}
+
+%% REFERENCES
+
+%% The reference list is compiled as follows:
+
+% \begin{thebibliography}{}
+
+% \bibitem[AUTHOR(YEAR)]{LABEL}
+% REFERENCE 1
+
+% \bibitem[AUTHOR(YEAR)]{LABEL}
+% REFERENCE 2
+
+% \end{thebibliography}
+
+%% Since the Copernicus LaTeX package includes the BibTeX style file copernicus.bst,
+%% authors experienced with BibTeX only have to include the following two lines:
+%%
+\bibliographystyle{copernicus}
+\bibliography{biblio}
+%%
+%% URLs and DOIs can be entered in your BibTeX file as:
+%%
+%% URL = {http://www.xyz.org/~jones/idx_g.htm}
+%% DOI = {10.5194/xyz}
+
+%% LITERATURE CITATIONS
+%%
+%% command                        & example result
+%% \citet{jones90}|               & Jones et al. (1990)
+%% \citep{jones90}|               & (Jones et al., 1990)
+%% \citep{jones90,jones93}|       & (Jones et al., 1990, 1993)
+%% \citep[p.~32]{jones90}|        & (Jones et al., 1990, p.~32)
+%% \citep[e.g.,][]{jones90}|      & (e.g., Jones et al., 1990)
+%% \citep[e.g.,][p.~32]{jones90}| & (e.g., Jones et al., 1990, p.~32)
+%% \citeauthor{jones90}|          & Jones et al.
+%% \citeyear{jones90}|            & 1990
+
+%% FIGURES
+
+%% When figures and tables are placed at the end of the MS (article in one-column style), please add \clearpage
+%% between bibliography and first table and/or figure as well as between each table and/or figure.
+
+%% ONE-COLUMN FIGURES
+
+%%f
+%\begin{figure}[t]
+%\includegraphics[width=8.3cm]{FILE NAME}
+%\caption{TEXT}
+%\end{figure}
+%
+%%% TWO-COLUMN FIGURES
+%
+%%f
+%\begin{figure*}[t]
+%\includegraphics[width=12cm]{FILE NAME}
+%\caption{TEXT}
+%\end{figure*}
+%
+%
+
+%%% TABLES
+%%%
+%%% The different columns must be seperated with a & command and should
+%%% end with \\ to identify the column brake.
+%
+%%% ONE-COLUMN TABLE
+%
+%%t
+%\begin{table}[t]
+%\caption{TEXT}
+%\begin{tabular}{column = lcr}
+%\tophline
+%
+%\middlehline
+%
+%\bottomhline
+%\end{tabular}
+%\belowtable{} % Table Footnotes
+%\end{table}
+
+%%% TWO-COLUMN TABLE
+%
+%%t
+%\begin{table*}[t]
+%\caption{TEXT}
+%\begin{tabular}{column = lcr}
+%\tophline
+%
+%\middlehline
+%
+%\bottomhline
+%\end{tabular}
+%\belowtable{} % Table Footnotes
+%\end{table*}
+
+%%% MATHEMATICAL EXPRESSIONS
+%
+%%% All papers typeset by Copernicus Publications follow the math typesetting regulations
+%%% given by the IUPAC Green Book (IUPAC: Quantities, Units and Symbols in Physical Chemistry,
+%%% 2nd Edn., Blackwell Science, available at: http://old.iupac.org/publications/books/gbook/green_book_2ed.pdf, 1993).
+%%%
+%%% Physical quantities/variables are typeset in italic font (t for time, T for Temperature)
+%%% Indices which are not defined are typeset in italic font (x, y, z, a, b, c)
+%%% Items/objects which are defined are typeset in roman font (Car A, Car B)
+%%% Descriptions/specifications which are defined by itself are typeset in roman font (abs, rel, ref, tot, net, ice)
+%%% Abbreviations from 2 letters are typeset in roman font (RH, LAI)
+%%% Vectors are identified in bold italic font using \vec{x}
+%%% Matrices are identified in bold roman font
+%%% Multiplication signs are typeset using the LaTeX commands \times (for vector products, grids, and exponential notations) or \cdot
+%%% The character * should not be applied as mutliplication sign
+%
+%
+%%% EQUATIONS
+%
+%%% Single-row equation
+%
+%\begin{equation}
+%
+%\end{equation}
+%
+%%% Multiline equation
+%
+%\begin{align}
+%& 3 + 5 = 8\\
+%& 3 + 5 = 8\\
+%& 3 + 5 = 8
+%\end{align}
+%
+%
+%%% MATRICES
+%
+%\begin{matrix}
+%x & y & z\\
+%x & y & z\\
+%x & y & z\\
+%\end{matrix}
+%
+%
+%%% ALGORITHM
+%
+%\begin{algorithm}
+%\caption{?}
+%\label{a1}
+%\begin{algorithmic}
+%?
+%\end{algorithmic}
+%\end{algorithm}
+%
+%
+%%% CHEMICAL FORMULAS AND REACTIONS
+%
+%%% For formulas embedded in the text, please use \chem{}
+%
+%%% The reaction environment creates labels including the letter R, i.e. (R1), (R2), etc.
+%
+%\begin{reaction}
+%%% \rightarrow should be used for normal (one-way) chemical reactions
+%%% \rightleftharpoons should be used for equilibria
+%%% \leftrightarrow should be used for resonance structures
+%\end{reaction}
+%
+%
+%%% PHYSICAL UNITS
+%%%
+%%% Please use \unit{} and apply the exponential notation
 

### 2017-12-18 deforestmap.R

@@ -49,7 +49,7 @@ SavedObjects <- c()
 ##=====================================
 ## Create new grass location in UTM 38S
 #dir.create("grassdata")
-#system("grass72 -c epsg:32738 grassdata/deforestmap")  # Ignore errors
+#system2("grass72 -c epsg:32738 grassdata/deforestmap")  # Ignore errors
 ## Connect R to grass location
 ## Make sure that /usr/lib/grass72/lib is in your PATH in RStudio
 Sys.setenv(LD_LIBRARY_PATH=paste("/usr/lib/grass72/lib", Sys.getenv("LD_LIBRARY_PATH"),sep=":"))
@@ -83,21 +83,21 @@ proj.t <- "EPSG:32738"
 Input <- "gisdata/rasters/harper/905_dsp_nosea.img"
 Output <- "outputs/harper.tif"
 # gdalwarp
-system(paste0("gdalwarp -overwrite -dstnodata ",nodat," -s_srs ",proj.s," -t_srs ",proj.t," -te ",Extent,
+system2(paste0("gdalwarp -overwrite -dstnodata ",nodat," -s_srs ",proj.s," -t_srs ",proj.t," -te ",Extent,
              " -tr ",Res," ",Res," -r near -co 'COMPRESS=LZW' -co 'PREDICTOR=2' ",Input," ",Output))
 
 #= Import harper into grass location
-system("r.in.gdal --o input=outputs/harper.tif output=harper")
+system2("r.in.gdal --o input=outputs/harper.tif output=harper")
 # Set region
-system("g.region rast=harper -ap")
+system2("g.region rast=harper -ap")
 
 #= Typology of harper map 1990-2000-2005
-system("r.info harper")
-system("r.describe harper")
+system2("r.info harper")
+system2("r.describe harper")
 # * 111 112 115 122 152 155 222 444 555 755 775 777
 # 1=forest, 2=nonforest, 5=cloud, 4=water, 7=mangrove
 
-system("r.report --o harper units=h output='outputs/report_harper.txt'")
+system2("r.report --o harper units=h output='outputs/report_harper.txt'")
 
 ## +-----------------------------------------------------------------------------+
 ## |                         RASTER MAP CATEGORY REPORT                          |
@@ -157,12 +157,12 @@ nodat <- "255"
[…]

### 2017-12-19 deforestmap.R
Update Readme

@@ -50,9 +50,16 @@ SavedObjects <- c()
 ## Create new grass location in UTM 38S
 #dir.create("grassdata")
 #system2("grass72 -c epsg:32738 grassdata/deforestmap")  # Ignore errors
+
 ## Connect R to grass location
 ## Make sure that /usr/lib/grass72/lib is in your PATH in RStudio
+## On Linux, find the path to GRASS GIS with: $ grass72 --config path
+## It should be somethin like: "/usr/lib/grass72"
+## On Windows, find the path to GRASS GIS with: C:\>grass72.bat --config path
+## If you use OSGeo4W, it should be: "C:\OSGeo4W\apps\grass\grass-7.2"
 Sys.setenv(LD_LIBRARY_PATH=paste("/usr/lib/grass72/lib", Sys.getenv("LD_LIBRARY_PATH"),sep=":"))
+
+## Initialize GRASS
 initGRASS(gisBase="/usr/lib/grass72",home=tempdir(), 
           gisDbase="grassdata",
           location="deforestmap",mapset="PERMANENT",

### 2017-12-19 deforestmap.R
Update README

@@ -65,12 +65,12 @@ initGRASS(gisBase="/usr/lib/grass72",home=tempdir(),
           location="deforestmap",mapset="PERMANENT",
           override=TRUE)
 
-##======================================================================
-## Download data (263.6 Mo): will have to be done from a Zenodo repository
+##==================================================================
+## Download data from Zenodo: https://doi.org/10.5281/zenodo.1118955
 if (down) {
-  d <- "http://bioscenemada.cirad.fr/githubdata/deforestmap/deforestmap_data.zip"
-  download.file(url=d,destfile="deforestmap_data.zip",method="wget",quiet=TRUE)
-  unzip("deforest_data.zip")
+  d <- "https://zenodo.org/record/1118956/files/gisdata.zip"
+  download.file(url=d,destfile="gisdata.zip",method="curl",quiet=FALSE)
+  unzip("gisdata.zip", exdir="test")
 }
 
 ##===========================================================

### 2017-12-21 deforestmap.R
Forest density

@@ -49,7 +49,7 @@ SavedObjects <- c()
 ##=====================================
 ## Create new grass location in UTM 38S
 #dir.create("grassdata")
-#system2("grass72 -c epsg:32738 grassdata/deforestmap")  # Ignore errors
+#system("grass72 -c epsg:32738 grassdata/deforestmap")  # Ignore errors
 
 ## Connect R to grass location
 ## Make sure that /usr/lib/grass72/lib is in your PATH in RStudio
@@ -90,21 +90,21 @@ proj.t <- "EPSG:32738"
 Input <- "gisdata/rasters/harper/905_dsp_nosea.img"
 Output <- "outputs/harper.tif"
 # gdalwarp
-system2(paste0("gdalwarp -overwrite -dstnodata ",nodat," -s_srs ",proj.s," -t_srs ",proj.t," -te ",Extent,
+system(paste0("gdalwarp -overwrite -dstnodata ",nodat," -s_srs ",proj.s," -t_srs ",proj.t," -te ",Extent,
              " -tr ",Res," ",Res," -r near -co 'COMPRESS=LZW' -co 'PREDICTOR=2' ",Input," ",Output))
 
 #= Import harper into grass location
-system2("r.in.gdal --o input=outputs/harper.tif output=harper")
+system("r.in.gdal --o input=outputs/harper.tif output=harper")
 # Set region
-system2("g.region rast=harper -ap")
+system("g.region rast=harper -ap")
 
 #= Typology of harper map 1990-2000-2005
-system2("r.info harper")
-system2("r.describe harper")
+system("r.info harper")
+system("r.describe harper")
 # * 111 112 115 122 152 155 222 444 555 755 775 777
 # 1=forest, 2=nonforest, 5=cloud, 4=water, 7=mangrove
 
-system2("r.report --o harper units=h output='outputs/report_harper.txt'")
+system("r.report --o harper units=h output='outputs/report_harper.txt'")
 
 ## +-----------------------------------------------------------------------------+
 ## |                         RASTER MAP CATEGORY REPORT                          |
@@ -164,12 +164,12 @@ nodat <- "255"
 Input <- "gisdata/vectors/madagascar_ecoregion_tenaizy_38s.shp"
 Output <- "outputs/ecoregion.tif"
 # gdalwarp
[…]

### 2018-01-11 R/plotfcc.R
Forest density

@@ -21,7 +21,7 @@ zooms <- list(zoom.w,zoom.e)
 # Compute zooms
 for (z in 1:length(zooms)) {
   ExtentZ <- paste(zooms[[z]]$xmin,zooms[[z]]$ymin,zooms[[z]]$xmax,zooms[[z]]$ymax)
-  Input <- c("fcc.tif","fcc_for1953.tif","frag2014.tif","dist_edge_2014_mask.tif")
+  Input <- c("fcc.tif","fcc_for1953.tif","fordens2014_class_mask.tif","dist_edge_2014_mask.tif")
   nodata <- c(rep(255,3),9999)
   Output <- paste0(sub(".tif","",Input),"_zoom",z,".tif")
   for (i in 1:length(Input)) {
@@ -265,8 +265,8 @@ zw.fcc <- plot_zoom_fcc("outputs/fcc_zoom1.tif",fcc.palette) + ylab("Cover 1973-
 ze.fcc <- plot_zoom_fcc("outputs/fcc_zoom2.tif",fcc.palette) + ylab("")
 
 ## Zooms frag
-zw.frag <- plot_zoom_frag("outputs/frag2014_zoom1.tif")
-ze.frag <- plot_zoom_frag("outputs/frag2014_zoom2.tif") + ylab("") + theme(legend.position="none")
+zw.frag <- plot_zoom_frag("outputs/fordens2014_class_mask_zoom1.tif")
+ze.frag <- plot_zoom_frag("outputs/fordens2014_class_mask_zoom2.tif") + ylab("") + theme(legend.position="none")
 
 ## Zooms dist
 zw.dist <- plot_zoom_dist("outputs/dist_edge_2014_mask_zoom1.tif")

### 2018-01-11 deforestmap.R

@@ -379,7 +379,7 @@ system("r.mask -r")
 #= We rename for1973 (with cloud class 5) as for1973_5
 #= for1973 will be without clouds
 system("g.rename raster=for1973,for1973_5")
-system("r.mapcalc 'for1973 = if(!isnull(for1973_5) &&& for1973_5==1, 1, null())'")
+system("r.mapcalc --o 'for1973 = if(!isnull(for1973_5) &&& for1973_5==1, 1, null())'")
 
 #====================================================================================
 # Export images as .png and .gif
@@ -430,16 +430,20 @@ for (i in 1:length(Year)) {
   # Mask on forest
   system(paste0("r.mask --o raster=for",Year[i]))
   # Forest density (takes several hours to run)
-  system(paste0("r.neighbors input=for",Year[i]," output=forcount",Year[i]," method=count size=51"))
+  system(paste0("r.neighbors --o input=for",Year[i]," output=forcount",Year[i]," method=count size=51"))
   system(paste0("r.mapcalc --o 'fordens",Year[i]," = round(100*forcount",Year[i],"/2601)'"))
   # Reclassify
-  system(paste0("r.reclass input=fordens",Year[i]," output=fordens",Year[i],"_class rules=outputs/densrules.txt"))
+  system(paste0("r.reclass --o input=fordens",Year[i]," output=fordens",Year[i],"_class rules=outputs/densrules.txt"))
   # Export
   system(paste0("r.out.gdal --o input=fordens",Year[i]," createopt='compress=lzw,predictor=2' type=Byte output=outputs/fordens",Year[i],".tif"))
   system(paste0("r.out.gdal --o input=fordens",Year[i],"_class createopt='compress=lzw,predictor=2' type=Byte output=outputs/fordens",Year[i],"_class.tif"))
 }
+# Prepare fordens2014_class for figures
+system("r.mask --o raster=harper")
+system("r.mapcalc --o 'fordens2014_class_mask = if(isnull(fordens2014_class), 0, fordens2014_class)'")
+system("r.out.gdal --o input=fordens2014_class_mask nodata=255 type=Byte createopt='compress=lzw,predictor=2' \\
+       output=outputs/fordens2014_class_mask.tif")
 system("r.mask -r")
-
 
 ##==========================
[…]

### 2018-01-11 R/plotfcc.R
ggspatial R package

@@ -12,6 +12,7 @@ library(rasterVis)
 library(gridExtra)
 library(grid)
 library(rgdal)
+library(ggspatial)  # for geom_spatial()
 
 # Zooms
 zoom.w <- list(xmin=346000,xmax=439000,ymin=7387000,ymax=7480000)
@@ -163,9 +164,6 @@ plot_zoom_dist <- function(rast_file,palette=dist.palette,vr=dist.vr) {
 # ================
 # Ecoregions
 
-# devtools::install_github("paleolimbot/ggspatial")
-library(ggspatial)  # for geom_spatial()
-
 # Import
 ecoregion <- readOGR(dsn="gisdata/vectors",layer="madagascar_ecoregion_tenaizy_38s")
 # Recode ecoregion
@@ -195,7 +193,7 @@ eco.col <- c("h"=green.t,"d"=orange.t,"s"=red.t,"m"="blue","1"=black.t)
 
 # Plot
 plot.ecoregion <- gplot(for2014, maxpixels=10e5) +
-  geom_spatial(ecoregion, aes(x=long,y=lat,group=group,fill=factor(code))) +
+  geom_spatial(ecoregion, aes(x=long,y=lat,group=group,fill=factor(code)), crsto=32738) +
   geom_raster(aes(fill=factor(value))) +
   scale_fill_manual(values=eco.col, na.value="transparent") +
   geom_text(data=t.df, aes(x=x, y=y, label=text), size=5) +

### 2018-01-12 manuscript/manuscript.Rmd
New article version

@@
 
-1.
-The island of Madagascar has an unparalleled biodiversity, mainly located in the tropical forests of the island, which is highly threatened by anthropogenic deforestation.
-Scattered forest maps from past studies at national level with substantial gaps (due to presence of cloud cover on satellite imagery) prevent the analyzis of long-term deforestation trends in Madagascar.
-
-2.
-In this study, we propose a new approach combining historical (1953-2000) national forest-cover maps with recent (2000-2014) global annual tree cover loss data to look at six decades (1953-2014) of deforestation and forest fragmentation in Madagascar.
-We produced new forest-cover maps at 30m-resolution over the full territory of Madagascar for the year 1990, and annually from 2000 to 2014.
-
-3.
-We estimated that Madagascar has lost 44% of its natural forest cover over the period 1953-2014 (including 37% over the period 1973-2014).
-Natural forests cover 8.9 Mha in 2014 (15% of the national territory) which are divided into 4.4 Mha (50%) of moist forests, 2.6 Mha (29%) of dry forests, 1.7 Mha of spiny forests (19%) and 177,000 ha (2%) of mangroves.
-Since 2005, the annual deforestation rate has progressively increased in Madagascar to reach 99,000 ha/yr during 2010-2014 (corresponding to a rate of 1.08%/yr).
-Around half of the forest (46%) is now located at less than 100m from the forest edge.
-
-4.
-_Policy implications_: Increase in the deforestation rate since 2005 can be related to rapid population growth (close to 3%/yr) and to poor law enforcement due to political instability in the country.
-More effort, including conservation and development programs, is needed in Madagascar to protect the remaining natural tropical forests, both to enhance local people's livelihoods and to support global environmental efforts.
+The island of Madagascar has an unparalleled biodiversity, mainly located in the tropical forests of the island, which is highly
+  threatened by anthropogenic deforestation. Scattered forest maps
+  from past studies at national level with substantial gaps (due to
+  presence of cloud cover on satellite imagery) prevent the analysis
+  of long-term deforestation trends in Madagascar. In this study, we
+  propose a new approach combining historical (1953-2000) national
+  forest-cover maps with recent (2001-2014) global annual tree cover
+  loss data to look at six decades (1953-2014) of deforestation and
+  forest fragmentation in Madagascar. We produced new forest-cover
+  maps at 30~m resolution over the full territory of Madagascar for
+  the year 1990, and annually from 2000 to 2014. We estimated that
+  Madagascar has lost 44\% of its natural forest cover over the period
+  1953-2014 (including 37\% over the period 1973-2014). Natural
+  forests cover 8.9 Mha in 2014 (15\% of the national territory) which
+  are divided into 4.4 Mha (50\%) of moist forests, 2.6 Mha (29\%) of
+  dry forests, 1.7 Mha of spiny forests (19\%) and 177,000 ha (2\%) of
+  mangroves. Since 2005, the annual deforestation rate has
+  progressively increased in Madagascar to reach 99,000 ha/yr during
+  2010-2014 (corresponding to a rate of 1.1\%/yr). This increase is
+  probably due to rapid population growth (close to 3\%/yr) and to
+  poor law enforcement in the country. Around half of the forest
+  (46\%) is now located at less than 100~m from the forest
+  edge. Accurate forest-cover change maps can be used to assess the
+  effectiveness of past and current conservation programs and
+  implement new strategies for the future. In particular, forest maps
+  and estimates can be used in the REDD+ framework which aims at
+  ``Reducing Emissions from Deforestation and Forest Degradation'' and
+  for optimizing the current protected area network.
 
 **Keywords**: biodiversity, climate-change, deforestation, Landsat, Madagascar, tropical forest.
-
-**Potential journals** (2015 IF): Conservation Letters (7.126), Journal of Applied Ecology (5.196), Conservation Biology (4.257), Environmental Research Letters (4.134), Biological Conservation (3.985), Landscape Ecology (3.657), Ambio (2.555), Environmental Conservation (2.235), Biotropica (1.944), Tropical Conservation Science (1.55).
 
@@
 ```{r frag, results="asis"}
-frag.df <- read.table(file="outputs/frag.txt", sep="\t", header=TRUE)
+frag.df <- read.table(file="outputs/fordens.txt", sep="\t", header=TRUE)
 tab.4 <- frag.df
@@
 tab.4$forest <- round(tab.4$forest/1000)
-names(tab.4) <- c("Year","Forest (ha)", "patch (%)", "transitional (%)", "edge (%)", "perforated (%)", "interior (%)")
+names(tab.4) <- c("Year","Forest (ha)", "0-20", "21-40", "41-60", "61-80", "81-100")
 kable(tab.4, digits=0, format.args=list(big.mark=","))
@@
 Ecoregions were defined following climatic [@Cornet1974] and vegetation [@IEFN1996] criteria.
-The dark grey areas represent the remaining natural forest cover for the year 2014.](outputs/ecoregion.png)
+The dark grey areas represent the remaining natural forest cover for the year 2014.](figs/ecoregion.png)
 
@@
 Two zooms in the western dry (left part) and eastern moist (right part) ecoregions present more detailed views of (from top to bottom): forest-cover in 1950s, forest-cover change from _c._ 1973 to 2014, forest fragmentation in 2014 and distance to forest edge in 2014.
-Data on water bodies (blue) and water seasonality (light blue for seasonal water to dark blue for permanent water) has been extracted from @Pekel2016.](outputs/fig_fcc_highres.png)
+Data on water bodies (blue) and water seasonality (light blue for seasonal water to dark blue for permanent water) has been extracted from @Pekel2016.](figs/fig_fcc.png)
 
@@
 Horizontal dashed grey line indicates a distance to forest edge of 100m.
-Percentages indicate the percentage of forest at a distance to forest edge lower than 100m for each year.](outputs/dist.png)
+Percentages indicate the percentage of forest at a distance to forest edge lower than 100m for each year.](figs/dist.png)
 

### 2018-01-12 manuscript/manuscript.tex
Update Readme file
New article version

--- 2017-06-19 report/deforestmap.tex
@@
 \begin{center}
-%%{\large
-  Ghislain Vieilledent$^{1,2,\star}$\hspace{1cm}
-  Clovis Grinand$^{3}$ \\
-  \vspace{0.25cm}
-  Fety A. Rakotomalala$^{3}$ \hspace{1cm}
-  Rija Ranaivosoa$^{4}$ \hspace{1cm}
-  Jean-Roger Rakotoarijaona$^{4}$ \\
-  \vspace{0.25cm}
-  Thomas F. Allnutt$^{5,6}$\hspace{1cm}and\hspace{1cm}
-  Frédéric Achard$^{1}$
-%%}
+{\large
+  Ghislain Vieilledent$^{1,2,\star}$, Clovis Grinand$^{3}$,
+  Fety A. Rakotomalala$^{3}$,\\
+  Rija Ranaivosoa$^{4}$, Jean-Roger Rakotoarijaona$^{4}$,\\
+  Thomas F. Allnutt$^{5,6}$, and Frédéric Achard$^{1}$\\
+}
 \end{center}
@@
      \textbf{Highlights:}\\
-    -We produced new 30 m resolution annual forest-cover maps for Madagascar for the period 2000-2014.\\
+     - We produced new 30 m resolution forest-cover maps for Madagascar for the period 2000-2014.\\
      - Madagascar has lost 44\% of its natural forest-cover over the period 1953-2014.\\
-   - Half of the tropical forest in Madagascar is now located at less than 100m from forest edge.\\
-   - Annual deforestation rate has increased in Madagascar since 2005 to reach 1.08\%/yr.\\
+     - Annual deforestation rate has increased in Madagascar since 2005 to reach 1.1\%/yr in 2014.\\
+     - Half of the tropical forest in Madagascar is now located at less than 100~m from forest edge.\\
      - Conservation efforts must be intensified to save Madagascar unique forests and biodiversity.\\
-\end{flushleft}}
+   \end{flushleft}
+ }
 
@@
   presence of cloud cover on satellite imagery) prevent the analysis
-  of long-term deforestation trends in Madagascar. In this study, we propose 
-  a new approach combining historical
-  (1953-2000) national forest-cover maps with recent (2001-2014)
-  global annual tree cover loss data to look at six decades
-  (1953-2014) of deforestation and forest fragmentation in
-  Madagascar. We produced new forest-cover maps at 30~m resolution over
-  the full territory of Madagascar for the year 1990, and annually
-  from 2000 to 2014. We estimated that Madagascar has lost 44\% of its natural forest
-  cover over the period 1953-2014 (including 37\% over the period
-  1973-2014). Natural forests cover 8.9 Mha in 2014 (15\% of the
-  national territory) which are divided into 4.4 Mha (50\%) of moist
-  forests, 2.6 Mha (29\%) of dry forests, 1.7 Mha of spiny forests
-  (19\%) and 177,000 ha (2\%) of mangroves. Since 2005, the annual
-  deforestation rate has progressively increased in Madagascar to
-  reach 99,000 ha/yr during 2010-2014 (corresponding to a rate of
-  1.08\%/yr). This increase is probably due to rapid population growth
-  (close to 3\%/yr) and to poor law enforcement in the country. Around
-  half of the forest (46\%) is now located at less than 100m from the
-  forest edge. Accurate forest-cover change maps can
-  be used to assess the effectiveness of past and current conservation
-  programs and implement new strategies for the future. In particular,
-  forest maps and estimates can be used in the framework of the REDD+
-  (``Reducing Emissions from Deforestation and Forest Degradation'')
-  initiative and for optimizing the current protected area network.
+  of long-term deforestation trends in Madagascar. In this study, we
+  propose a new approach combining historical (1953-2000) national
+  forest-cover maps with recent (2001-2014) global annual tree cover
+  loss data to look at six decades (1953-2014) of deforestation and
+  forest fragmentation in Madagascar. We produced new forest-cover
+  maps at 30~m resolution over the full territory of Madagascar for
+  the year 1990, and annually from 2000 to 2014. We estimated that
+  Madagascar has lost 44\% of its natural forest cover over the period
+  1953-2014 (including 37\% over the period 1973-2014). Natural
+  forests cover 8.9 Mha in 2014 (15\% of the national territory) which
+  are divided into 4.4 Mha (50\%) of moist forests, 2.6 Mha (29\%) of
+  dry forests, 1.7 Mha of spiny forests (19\%) and 177,000 ha (2\%) of
+  mangroves. Since 2005, the annual deforestation rate has
+  progressively increased in Madagascar to reach 99,000 ha/yr during
+  2010-2014 (corresponding to a rate of 1.1\%/yr). This increase is
+  probably due to rapid population growth (close to 3\%/yr) and to
+  poor law enforcement in the country. Around half of the forest
+  (46\%) is now located at less than 100~m from the forest
+  edge. Accurate forest-cover change maps can be used to assess the
+  effectiveness of past and current conservation programs and
+  implement new strategies for the future. In particular, forest maps
+  and estimates can be used in the REDD+ framework which aims at
+  ``Reducing Emissions from Deforestation and Forest Degradation'' and
+  for optimizing the current protected area network.
 
@@
 from \citet{Hansen2013} cannot be used alone to produce a map of
-forest cover \citep{Tyukavina2017}. In complement to the tree cover
-percentage provided in \citet{Hansen2013}, a layer of annual tree
-cover loss is also provided (i.e.~complete loss of tree cover from a
-value higher than 10\% to zero) for the period 2001-2014.
-
-In this study, we present a simple approach which combines the maps
-from \citet{Harper2007} and products from \citet{Hansen2013} to derive
-annual wall-to-wall forest-cover change maps over the period 2000-2014
-for Madagascar. We use the forest-cover map provided by
-\citet{Harper2007} for the year 2000 (defining the land-use) with the
-tree cover loss product provided by \citet{Hansen2013} that we apply
-only inside forest areas identified by \citet{Harper2007}. Similar to
-the approach of \citet{Harper2007}, we also assess trends in
-deforestation rates and forest fragmentation from \emph{c.} 1953 to 2014. The
-approach described in this study can help assess the effectiveness of
-current conservation strategies, and assist the implementation of
-future strategies. Our approach could be easily extended to other
-tropical countries that have at least one forest-cover map between
-2000 and 2014. This approach can easily be repeated in the future when
-the \citet{Hansen2013} products are updated.
+forest cover \citep{Tyukavina2017}.
+
+In this study, we present a simple approach which combines the
+historical forest maps from \citet{Harper2007} and more recent global
+products from \citet{Hansen2013} to derive annual wall-to-wall
+forest-cover change maps over the period 2000-2014 for Madagascar. We
+use the forest-cover map provided by \citet{Harper2007} for the year
+2000 (defining the land-use) with the tree cover loss product provided
+by \citet{Hansen2013} that we apply only inside forest areas
+identified by \citet{Harper2007}. Similar to the approach of
+\citet{Harper2007}, we also assess trends in deforestation rates and
+forest fragmentation from \emph{c.} 1953 to 2014. We finally discuss
+the possibility to extend our approach to other tropical countries or
+repeat it in the future. We also discuss how our results could help
+assess the effectiveness of current conservation strategies, and
+implementation new conservation strategies for the future in
+Madagascar.
 
@@
 forest map of year 2000 from \citet{Harper2007}, and the tree cover
-percentage and annual forest cover loss maps over the period 2000-2014
+percentage and annual tree cover loss maps over the period 2000-2014
 from \citet{Hansen2013}. The 2000 Harper's forest map includes 208,000
@@
 to define forest cover as recommended by other studies for the moist
-domain \citep{Achard2014}. We thus obtained a forest-cover map for the
-year 2000 covering the full territory of Madagascar. We then combined
-this forest-cover map of the year 2000 with the annual tree cover loss
-maps from 2001 to 2014 provided by \citet{Hansen2013} to create annual
-forest-cover maps from 2001 to 2014 at 30~m resolution. We also
+domain \citep{Achard2014, Aleman2017}. To do so, the Hansen's 2000
+tree cover map was resampled on the same grid as the original Harper's
+map at 30~m resolution using a bilinear interpolation. We thus
+obtained a forest-cover map for the year 2000 covering the full
+territory of Madagascar. We then combined this forest-cover map of the
+year 2000 with the annual tree cover loss maps from 2001 to 2014
+provided by \citet{Hansen2013} to create annual forest-cover maps from
+2001 to 2014 at 30~m resolution. To do so, Hansen's tree cover loss
+maps were resampled on the same grid as the original Harper's map at
+30~m resolution using a nearest-neighbor interpolation. We also
 completed the Harper's forest map of year 1990 by filling unclassified
@@
 unclassified pixels were limited to a relatively small total area of
-\emph{c.} 8,000 ha. We labeled these residual pixels as non-forest as
+\emph{c.} 8,000 ha. We labeled these residual pixels as non-forest, as
 for the year 2000. Similarly we completed the Harper's forest map of
@@
 present in 1973. Contrary to the year 1990, the remaining unclassified
-pixels for year 1973 corresponded to a significant total area of 3.32
+pixels for year 1973 corresponded to a significant total area of 3.3
 million ha. We also reprojected the forest-cover map of year 1953 to a
@@
 activities). This allowed us to avoid counting very small scale events
-(\textless{}0.1 ha such as selective logging) as forest
-fragmentation. All the resulting maps are freely available at
-\url{https://bioscenemada.cirad.fr/forestmaps}.
-
-\subsection{Computing forest-cover areas and deforestation
-rates}
+(\textless 0.1 ha, such as selective logging) as forest
+fragmentation. All the resulting maps have been made permanently and
+freely available on Zenodo (a research data repository) at
+\url{https://doi.org/10.5281/zenodo.1145785}.
+
+\subsection{Computing forest-cover areas and deforestation rates}
 
 From these new forest-cover maps, we calculated the total forest-cover
-area for seven available years (1953-1973-1990-2000-2010-2005-2014),
+area for seven available years (1953-1973-1990-2000-2005-2010-2014),
 and the annual deforested area and annual deforestation rate for the
 corresponding six time periods between 1953 and 2014. The annual
-deforestation rates were calculated as follows \citep{Puyravaud2003,
-  Vieilledent2013}:
-
-\begin{equation*}
-  \theta = 100 \times [1-(1-(F_{t_2}-F_{t_1})/F_{t_1})^{(1/(t_2-t_1))}
-\end{equation*}
-
-where $\theta$ is the annual deforestation rate (in \%/yr),
+deforestation rates were calculated using Eq.~\ref{eq:theta}
+\citep{Puyravaud2003, Vieilledent2013}:
+
+\begin{equation}
+  \label{eq:theta}
+  \theta = 100 \times [1-(1-(F_{t_2}-F_{t_1})/F_{t_1})^{(1/(t_2-t_1))}]
+\end{equation}
+
+In Eq.~\ref{eq:theta}, $\theta$ is the annual deforestation rate (in \%/yr),
 $F_{t_2}$ and $F_{t_1}$ are the forest cover free of clouds at both
@@
 
-Because of the large unclassified area (3.32 million ha) in 1973, the
+Because of the large unclassified area (3.3 million ha) in 1973, the
 annual deforestation areas and rates for the two periods 1953-1973 and
-1973-1990 are only indicative estimates. For these two periods the
-annual deforestation rates are computed as the ratio
-$(F_{t_2}-F_{t_1})/F_{t_1}$ considering only the mapped forest pixels.
-Area and rate estimates are produced at the national scale and for the
-four forest ecosystems present in Madagascar: moist forest in the
-East, dry forest in the West, spiny forest in the South, and mangroves
-on the Western coast (Fig.~\ref{fig:ecoregion}). To define the forest
-domains, we used a map from the MEFT (\emph{``Ministère de
-  l'Environnement et des Forêts à Madagascar''}) with the boundaries
-of the four ecoregions in Madagascar. Ecoregions were defined on the
-basis of climatic and vegetation criteria using the climate
-classification by \citet{Cornet1974} and the vegetation classification
-from the 1996 IEFN national forest inventory \citep{IEFN1996}. Because
-mangrove forests are highly dynamic ecosystems that can expand or
-contract on decadal scales depending on changes in environmental
-factors \citep{Armitage2015}, a fixed delimitation of the mangrove
-ecoregion on six decades might not be fully appropriate. As a
-consequence, our estimates of the forest-cover and deforestation rates
-for mangroves in Madagascar must be considered with this limitation.
+1973-1990 are only partial estimates computed on the basis of the
+available forest extent. Area and rate estimates are produced at the
+national scale and for the four forest ecosystems present in
+Madagascar: moist forest in the East, dry forest in the West, spiny
+forest in the South, and mangroves on the Western coast
+(Fig.~\ref{fig:ecoregion}). To define the forest domains, we used a
+map from the MEFT (\emph{``Ministère de l'Environnement et des Forêts
+  à Madagascar''}) with the boundaries of the four ecoregions in
+Madagascar. Ecoregions were defined on the basis of climatic and
+vegetation criteria using the climate classification by
+\citet{Cornet1974} and the vegetation classification from the 1996
+IEFN national forest inventory \citep{IEFN1996}. Because mangrove
+forests are highly dynamic ecosystems that can expand or contract on
+decadal scales depending on changes in environmental factors
+\citep{Armitage2015}, a fixed delimitation of the mangrove ecoregion
+on six decades might not be fully appropriate. As a consequence, our
+estimates of the forest-cover and deforestation rates for mangroves in
+Madagascar must be considered with this limitation.
 
@@
 We also conducted an analysis of changes in forest fragmentation for
-the years 1953, 1973, 1990, 2000, 2005, 2010 and 2014. We applied the
-method developed by \citet{Riitters2000} which uses a moving window to
-characterize the fragmentation around each forested
-pixel. Computations were done using the function \texttt{r.forestfrag}
-of the GRASS GIS software \citep{Neteler2008}. Six categories of
-fragmentation were identified from the amount of forest and its
-occurrence as adjacent forest pixels: ``interior'', ``perforated'',
-``edge'', ``transitional'', ``patch'', and ``undetermined''. We used a
-moving window of 7x7 pixels (4.4 ha). Using this window size, forest
-edge had a width of about 90m \citep{Riitters2000}. The ``interior''
-category can be interpreted as the most intact forest
-\citep{Potapov2017}. The ``patch'' and ``transitional'' categories
-correspond to isolated small forest patches. We reported the area of
-forest in each fragmentation category for the six years and analyzed
-the dynamics of fragmentation over the six decades. We also computed
-the distance to forest edge for all forest pixels for the years 1953,
-1973, 1990, 2000, 2005, 2010 and 2014. For that, we used the function
-\texttt{gdal\_proximity.py} of the GDAL software
-(\url{http://www.gdal.org/}). We computed the mean and 90\% quantiles
-(5\% and 95\%) of the distance to forest edge and looked at the
-evolution of these values with time.
+the years 1953, 1973, 1990, 2000, 2005, 2010 and 2014 at 30~m
+resolution. We used a moving window of $51 \times 51$ pixels centered
+on each forest pixel to compute the percentage of forest pixels in the
+neighborhood. We used this percentage as an indication of the forest
+fragmentation. The size of the moving windows was based on a
+compromise: a sufficiently high number of cells (here 2601) had to be
+considered to be able to compute a percentage and a reasonably low
+number of cells had to be choosen to have a local estimate of the
+fragmentation. Computations were done using the function
+\texttt{r.neighbors} of the GRASS GIS software
+\citep{Neteler2008}. Using the density of forest in the neighborhood,
+we defined five forest fragmentation classes: 0-20\% (highly
+fragmented), 21-40\%, 41-60\%, 61-80\% and 81-100\% (lowly
+fragmented). We reported the percentage of forest falling in each
+fragmentation class for the six years and analyzed the dynamics of
+fragmentation over the six decades.
+
+We also computed the distance to forest edge for all forest
+pixels for the years 1953, 1973, 1990, 2000, 2005, 2010 and 2014. For
+that, we used the function \texttt{gdal\_proximity.py} of the GDAL
+library (\url{http://www.gdal.org/}). We computed the mean and 90\%
+quantiles (5\% and 95\%) of the distance to forest edge and looked at
+the evolution of these values with time. Previous studies have shown
+that forest micro-habitats were mainly altered within the first 100~m
+of the forest edge \citep{Brinck2017, Gibson2013, Murcia1995,
+  Broadbent2008}. Consequently, we also estimated the percentage of
+forest within the first 100~m of the forest edge for each year and
+looked at the evolution of this percentage over the six decades.
 
@@
 respectively (Fig.~\ref{fig:fcc} and Tab.~\ref{tab:forest_cover}). In
-2014 the remaining 8.9 Mha of natural forest were distributed as: 4.4
+2014 the remaining 8.9 Mha of natural forest were distributed as follow: 4.4
 Mha of moist forest (50\% of total forest cover), 2.6 Mha of dry
@@
 observed a progressive decrease of the deforestation rate after 1990
-from 205,000 ha/yr (1.63\%/yr) over the period 1973-1990 to 44,300
-ha/yr (0.43\%/yr) over the period 2000-2005
+from 205,000 ha/yr (1.6\%/yr) over the period 1973-1990 to 42,000
+ha/yr (0.4\%/yr) over the period 2000-2005
 (Tab.~\ref{tab:forest_cover}). Then from 2005, the deforestation rate
 has progressively increased and has more than doubled over the period
-2010-2014 (98,700 ha/yr, 1.08\%/yr) compared to 2000-2005
+2010-2014 (99,000 ha/yr, 1.1\%/yr) compared to 2000-2005
 (Tab.~\ref{tab:forest_cover}). The deforestation trend characterized
@@
 type respectively. Proportions of unclassified areas are not reported
-in the two other existing studies by \citet{MEFT2009} and
-\citet{ONE2015}. With our approach, we produced wall to wall
-forest-cover change maps from 1990 to 2014 for the full territory of
-Madagascar (Tab.~\ref{tab:forest_cover}). This allowed us to produce
-more robust estimates of forest-cover and deforestation rates over
-this period. Our forest-cover estimates over the period 1953-2013
-(considering forest cover estimates at national level and by
+in the two other existing studies at the national level by
+\citet{MEFT2009} and \citet{ONE2015}. With our approach, we produced
+wall to wall forest-cover change maps from 1990 to 2014 for the full
+territory of Madagascar (Tab.~\ref{tab:forest_cover}). This allowed us
+to produce more robust estimates of forest-cover and deforestation
+rates over this period. Our forest-cover estimates over the period
+1953-2013 (considering forest cover estimates at national level and by
 ecoregions for all the available dates) were well correlated
@@
 2014 in Madagascar. The mean distance to forest edge has decreased to
-\emph{c.} 300~m in 2014 while it was previously \emph{c.} 1.5~km in 1973
-(Fig.~\ref{fig:dist_edge}). Moreover, a large proportion (73\%) of
-the forest was located at a distance greater than 100 m in 1973, while
-almost half of the forest (46\%) was at a distance lower than 100 m
-from forest edge in 2014 (Fig.~\ref{fig:dist_edge}). The percentage of
-forest that can be considered intact in Madagascar has continuously
-decreased since 1953. The percentage of forest belonging to the
-``interior'' category (most intact forests) has fallen from 68\% in
-1973 to 50\% in 2014. In 2014, more than 16\% of the forest belonged
-to the ``patch'' and ``transitional'' categories (isolated small
-forest patches) compared to 9.5\% in 1973 (Tab.~\ref{tab:frag}).
+\emph{c.} 300~m in 2014 while it was previously \emph{c.} 1.5~km in
+1973 (Fig.~\ref{fig:dist_edge}). Moreover, a large proportion (73\%)
+of the forest was located at a distance greater than 100~m in 1973,
+while almost half of the forest (46\%) is now at a distance lower than
+100~m from forest edge in 2014 (Fig.~\ref{fig:dist_edge}). The
+percentage of lowly fragmented forest in Madagascar has continuously
+decreased since 1953. The percentage of forest belonging to the lowly
+fragmented category has fallen from 57\% in 1973 to 44\% in 2014. In
+2014, 22\% of the forest belonged to the two highest
+fragmented forest classes (less than 40\% of forest cover in the
+neighborhood) while 15\% of the forest belonged to these two
+classes in 1973 (Tab.~\ref{tab:frag}).
 
@@
 approach can be repeated in the future if and when the global tree
-cover product is updated. We have made the R/GRASS code used for this
+cover product is updated. We have made the \R{}/GRASS code used for this
 study freely available in a GitHub repository (see Data availability
@@
 on the accuracies of the historical forest-cover maps and the tree
-cover loss dataset. The reported global accuracy of the tree cover
-loss dataset is 99.6\% (see Tab. S5 in \citet{Hansen2013}).
-\citet{Verhegghen2016} have compared deforestation estimates derived
-from the global tree cover loss dataset \citep{Hansen2013} with
-results derived from semi-automated supervised classification of
-Landsat satellite images \citep{Achard2014} for six countries in
-Central Africa and they found a good agreement between these two sets
-of estimates. Consistent with \citet{Harper2007}, we did not consider
-potential forest regrowth in Madagascar (although \citet{Hansen2013}
-provided a tree cover gains layer for the period 2001-2013) for
-several reasons. First, the tree gain layer of \citet{Hansen2013}
-includes and catches more easily tree plantations than natural forest
-regrowth \citep{Tropek2014}. Second, there is little evidence of
-natural forest regeneration in Madagascar \citep{Grouzis2001,
- Harper2007}. This can be explained by several ecological processes
-following burning practice such as soil erosion \citep{Grinand2017}
-and reduced seed bank due to fire and soil loss
-\citep{Grouzis2001}. Moreover, in areas where forest regeneration is
-ecologically possible, young forest regrowth are more easily re-burnt
-for agriculture and pasture. Third, young secondary forests provide
-more limited ecosystem services compared to old-growth natural forests
-in terms of biodiversity and carbon storage.
+cover loss dataset. Using visual-interpretation of aerial images in
+342 areas distributed among all forest types, \citet{Harper2007}
+estimated an overall 89.5\% accuracy in identifying forest/non-forest
+classes for the year 2000. The accuracy assessment of the tree cover
+loss dataset for the tropical biome reported 13\% of false positives
+and 16.9\% of false negatives (see Tab. S5 in
+\citet{Hansen2013}). These numbers rise at 20.7\% and 20.6\%
+respectively for the subtropical biome. In the subtropical biome, the
+lower density tree cover canopy makes it difficult to detect change
+from tree cover to bare ground. For six countries in Central Africa,
+with a majority of moist dense forest, \citet{Verhegghen2016} have
+compared deforestation estimates derived from the global tree cover
+loss dataset \citep{Hansen2013} with results derived from
+semi-automated supervised classification of Landsat satellite images
+\citep{Achard2014} and they found a good agreement between the two
+sets of estimates. Therefore, our forest-cover change maps after 2000
+might be more accurate for the dense moist forest than for the dry and
+spiny forest. In another study assessing the accuracy of the tree
+cover loss product accross the tropics \citep{Tyukavina2015}, authors
+reported 4\% of false positives and 48\% of false negatives in
+Sub-Saharian Africa. They showed that 85\% of missing loss occured on
+the edges of other loss patches. This means that tree cover loss might
+be underestimated in Sub-Saharian Africa, probably due to the
+prevalence of small-scale disturbance which is hard to map at 30~m,
+but that areas of large-scale deforestation are well identified and
+spatial variability of the deforestation is well represented. A proper
+accuracy assessment of our forest-cover change maps should be
+performed to better estimate the uncertainty surrounding our
+forest-cover change estimates in Madagascar from year 2000
+\citep{Olofsson2013,Olofsson2014}. Despite this limitation, we have
+shown that the deforestation trend we observed for Madagascar, with a
+doubling deforestation on the period 2010-2014 compared to 2000-2005,
+was consistent with the other studies at the national scale
+\citep{ONE2015, MEFT2009}.
+
+Consistent with \citet{Harper2007}, we did not consider potential
+forest regrowth in Madagascar (although \citet{Hansen2013} provided a
+tree cover gains layer for the period 2001-2013) for several
+reasons. First, the tree gain layer of \citet{Hansen2013} includes and
+catches more easily tree plantations than natural forest regrowth
+\citep{Tropek2014}. Second, there is little evidence of natural forest
+regeneration in Madagascar \citep{Grouzis2001, Harper2007}. This can
+be explained by several ecological processes following burning
+practice such as soil erosion \citep{Grinand2017} and reduced seed
+bank due to fire and soil loss \citep{Grouzis2001}. Moreover, in areas
+where forest regeneration is ecologically possible, young forest
+regrowth are more easily re-burnt for agriculture and pasture. Third,
+young secondary forests provide more limited ecosystem services
+compared to old-growth natural forests in terms of biodiversity and
+carbon storage.
 
@@
 showed that the remaining forests of Madagascar are highly fragmented
-with 46\% of the forest being at less than 100m of the forest
+with 46\% of the forest being at less than 100~m of the forest
 edge. Small forest fragments do not allow to maintain viable
@@
 deforestation rate estimates can be used as source of information for
-the next FAO Forest Resources Assessment project
-\citep{Keenan2015}. Current rates of deforestation can also be used to
-build reference scenarios for deforestation in Madagascar and
-contribute to the implementation of deforestation mitigation
-activities in the framework of REDD+ \citep{Olander2008}.
+the next FAO Forest Resources Assessment \citep{Keenan2015}. Current
+rates of deforestation can also be used to build reference scenarios
+for deforestation in Madagascar and contribute to the implementation
+of deforestation mitigation activities in the framework of REDD+
+\citep{Olander2008}.
 
@@
 
-This study is part of the Cirad's BioSceneMada project
+Authors thank Jean-François Bastin for useful comments on a previous
+version of the manuscript and Peter Vogt for fruitfull discussion on
+which metric to use to compute forest fragmentation. This study is
+part of the Cirad's BioSceneMada project
 (\url{https://bioscenemada.cirad.fr}) and the Joint Research Center's
@@
 
-All the data and codes used for this study are made publicly available
-in the \texttt{deforestmap} GitHub repository
-(\url{https://github.com/ghislainv/deforestmap.git}). The results are
-fully reproducible running the {\R} script \texttt{deforestmap.R}
-located inside the \texttt{deforestmap} repository.
+All the data and the script used for this study have been made permanently and publicly available
+on Zenodo (a research data repository) so that the results are entirely reproducible:
+\begin{itemize}
+\item Input data: \url{https://doi.org/10.5281/zenodo.1118955}
+\item Script: \url{https://doi.org/10.5281/zenodo.1118484}
+\item Output data: \url{https://doi.org/10.5281/zenodo.1145785}
+\end{itemize}
 
@@
 
-%%\bibliographystyle{bib/jae}
-%%\bibliography{bib/biblio}
-
-\section{References}
-
-\renewcommand{\bibsection}{} %% Remove section name for bibliography
-\begin{thebibliography}{55}
+%%\bibliographystyle{besjournals}
+%%\bibliography{biblio}
+
+\begin{thebibliography}{60}
 \providecommand{\natexlab}[1]{#1}
@@
   Seliger, R. \& Simonetti, D. (2014) Determination of tropical deforestation
-  rates and related carbon losses from 1990 to 2010.
-\newblock \emph{Global Change Biology}, \textbf{20}, 2540--2554.
+  rates and related carbon losses from 1990 to 2010. \emph{Global Change
+  Biology} \textbf{20}, 2540--2554.
+
+\bibitem[{Aleman \emph{et~al.}(2017)Aleman, Jarzyna \& Staver}]{Aleman2017}
+Aleman, J.C., Jarzyna, M.A. \& Staver, A.C. (2017) Forest extent and
+  deforestation in tropical africa since 1900. \emph{Nature Ecology \&
+  Evolution} .
 
@@
   from the Middle Jurassic through latest Eocene (166--35 Ma)}.
-\newblock \emph{Earth-Science Reviews}, \textbf{88}, 145--166.
+  \emph{Earth-Science Reviews} \textbf{88}, 145--166.
 
@@
   loss and its application to a 50-year record of deforestation across
-  Madagascar}.
-\newblock \emph{Conservation Letters}, \textbf{1}, 173--181.
+  Madagascar}. \emph{Conservation Letters} \textbf{1}, 173--181.
 
@@
   contribution of mangrove expansion to salt marsh loss on the Texas Gulf
-  Coast}.
-\newblock \emph{PloS One}, \textbf{10}, e0125404.
+  Coast}. \emph{PloS One} \textbf{10}, e0125404.
 
@@
   Diaz, A.S.P., Stolle, F., Surappaeva, V. \& Castro, R. (2017) The extent of
-  forest in dryland biomes.
-\newblock \emph{Science}, \textbf{356}, 635--638.
+  forest in dryland biomes. \emph{Science} \textbf{356}, 635--638.
 
@@
   of tropical forest fragmentation and its impact on the global carbon cycle.
-\newblock \emph{Nature Communications}, \textbf{8}, 14855.
+  \emph{Nature Communications} \textbf{8}, 14855.
+
+\bibitem[{Broadbent \emph{et~al.}(2008)Broadbent, Asner, Keller, Knapp,
+  Oliveira \& Silva}]{Broadbent2008}
+Broadbent, E.N., Asner, G.P., Keller, M., Knapp, D.E., Oliveira, P.J. \& Silva,
+  J.N. (2008) Forest fragmentation and edge effects from deforestation and
+  selective logging in the brazilian amazon. \emph{Biological Conservation}
+  \textbf{141}, 1745 -- 1757.
 
@@
   G. \& Hilton-Taylor, C. (2002) Habitat loss and extinction in the hotspots of
-  biodiversity.
-\newblock \emph{Conservation Biology}, \textbf{16}, 909--923.
+  biodiversity. \emph{Conservation Biology} \textbf{16}, 909--923.
 
@@
   Madagascar at the end of the first millennium of the Common Era}.
-\newblock \emph{Quaternary Science Reviews}, \textbf{134}, 92 -- 99.
+  \emph{Quaternary Science Reviews} \textbf{134}, 92 -- 99.
 
 \bibitem[{Cornet(1974)}]{Cornet1974}
-Cornet, A. (1974) {Essai de cartographie bioclimatique à Madagascar}.
-\newblock , Orstom.
+Cornet, A. (1974) {Essai de cartographie bioclimatique à Madagascar}. Tech.
+  rep., Orstom.
 
@@
   small cohort of Island Southeast Asian women founded Madagascar}.
-\newblock \emph{Proceedings of the Royal Society B: Biological Sciences},
-  \textbf{279}, 2761--2768.
+  \emph{Proceedings of the Royal Society B: Biological Sciences} \textbf{279},
+  2761--2768.
 
@@
   (2012) {Vertebrate time-tree elucidates the biogeographic pattern of a major
-  biotic change around the K--T boundary in Madagascar}.
-\newblock \emph{Proceedings of the National Academy of Sciences}, \textbf{109},
-  5358--5363.
+  biotic change around the K--T boundary in Madagascar}. \emph{Proceedings of
+  the National Academy of Sciences} \textbf{109}, 5358--5363.
 
@@
   (2016) {Contrasting spatial and temporal trends of protected area
-  effectiveness in mitigating deforestation in Madagascar}.
-\newblock \emph{Biological Conservation}, \textbf{203}, 290 -- 297.
+  effectiveness in mitigating deforestation in Madagascar}. \emph{Biological
+  Conservation} \textbf{203}, 290 -- 297.
 
@@
 Freudenberger, K. (2010) {Paradise Lost? Lessons from 25 years of USAID
-  environment programs in Madagascar}.
-\newblock \emph{International Resources Group, Washington DC}.
+  environment programs in Madagascar}. \emph{International Resources Group,
+  Washington DC} .
 
@@
   D.S., Bumrungsri, S. \& Laurance, W.F. (2013) Near-complete extinction of
-  native small mammal fauna 25 years after forest fragmentation.
-\newblock \emph{Science}, \textbf{341}, 1508--1510.
+  native small mammal fauna 25 years after forest fragmentation. \emph{Science}
+  \textbf{341}, 1508--1510.
 
@@
 Goodman, S.M. \& Benstead, J.P. (2005) {Updated estimates of biotic diversity
-  and endemism for Madagascar}.
-\newblock \emph{Oryx}, \textbf{39}, 73--77.
+  and endemism for Madagascar}. \emph{Oryx} \textbf{39}, 73--77.
 
@@
   {\"O}zler, B. (2011) \emph{{Exploring the Association Between People and
-  Deforestation in Madagascar}}, vol. 1650.
-\newblock Springer Berlin Heidelberg.
+  Deforestation in Madagascar}}, vol. 1650. Springer Berlin Heidelberg.
 
@@
   \& Bernoux, M. (2017) {Estimating temporal changes in soil carbon stocks at
-  ecoregional scale in Madagascar using remote-sensing}.
-\newblock \emph{International Journal of Applied Earth Observation and
-  Geoinformation}, \textbf{54}, 1--14.
+  ecoregional scale in Madagascar using remote-sensing}. \emph{International
+  Journal of Applied Earth Observation and Geoinformation} \textbf{54}, 1--14.
 
@@
   forests in Madagascar from 2000 to 2010 using multi-date Landsat satellite
-  images and the Random Forests classifier}.
-\newblock \emph{Remote Sensing of Environment}, \textbf{139}, 68--80.
+  images and the Random Forests classifier}. \emph{Remote Sensing of
+  Environment} \textbf{139}, 68--80.
 
@@
   {\'e}daphiques au cours de la phase post-culturale dans la r{\'e}gion
-  d’Analabo}.
-\newblock \emph{Soci{\'e}t{\'e}s paysannes, transitions agraires et dynamiques
-  {\'e}cologiques dans le Sud-Ouest de Madagascar, Antananarivo, IRD/CNRE}, pp.
-  327--337.
+  d’Analabo}. \emph{Soci{\'e}t{\'e}s paysannes, transitions agraires et
+  dynamiques {\'e}cologiques dans le Sud-Ouest de Madagascar, Antananarivo,
+  IRD/CNRE} pp. 327--337.
 
@@
   (2013) {High-Resolution Global Maps of 21st-Century Forest Cover Change}.
-\newblock \emph{Science}, \textbf{342}, 850--853.
+  \emph{Science} \textbf{342}, 850--853.
 
@@
   {Fifty years of deforestation and forest fragmentation in Madagascar}.
-\newblock \emph{Environmental Conservation}, \textbf{34}, 325--333.
+  \emph{Environmental Conservation} \textbf{34}, 325--333.
 
@@
 Humbert, H. (1927) {La destruction d'une flore insulaire par le feu. Principaux
-  aspects de la végétation à Madagascar}.
-\newblock \emph{Mémoires de l'Académie Malgache}, \textbf{5}, 1--80.
+  aspects de la végétation à Madagascar}. \emph{Mémoires de l'Académie
+  Malgache} \textbf{5}, 1--80.
 
@@
   Lindquist, E. (2015) {Dynamics of global forest area: Results from the FAO
-  Global Forest Resources Assessment 2015}.
-\newblock \emph{Forest Ecology and Management}, \textbf{352}, 9 -- 20.
+  Global Forest Resources Assessment 2015}. \emph{Forest Ecology and
+  Management} \textbf{352}, 9 -- 20, changes in Global Forest Resources from
+  1990 to 2015.
 
@@
   Feng, M. \& Townshend, J.R. (2014) {Global, Landsat-based forest-cover change
-  from 1990 to 2000}.
-\newblock \emph{Remote Sensing of Environment}, \textbf{155}, 178--193.
+  from 1990 to 2000}. \emph{Remote Sensing of Environment} \textbf{155},
+  178--193.
 
@@
 Klein, J. (2002) {Deforestation in the Madagascar Highlands -- Established
-  `truth' and scientific uncertainty}.
-\newblock \emph{GeoJournal}, \textbf{56}, 191--199.
+  `truth' and scientific uncertainty}. \emph{GeoJournal} \textbf{56}, 191--199.
 
@@
 Kull, C.A. (2000) {Deforestation, erosion, and fire: degradation myths in the
-  environmental history of Madagascar}.
-\newblock \emph{Environment and History}, \textbf{6}, 423--450.
-
-\bibitem[{{MEFT, USAID, and CI}(2009)}]{MEFT2009}
+  environmental history of Madagascar}. \emph{Environment and History}
+  \textbf{6}, 423--450.
+
+\bibitem[{MEFT \emph{et~al.}(2009)}]{MEFT2009}
 {MEFT, USAID, and CI} (2009) \emph{{Evolution de la couverture de forêts
@@
 {Ministère de l'Environnement} (1996) {IEFN: Inventaire Ecologique Forestier
-  National}.
-\newblock {Ministère de l'Environnement de Madagascar, Direction des Eaux et
-  Forêts, DFS Deutsch Forest Service GmbH, Entreprise d'études de
-  développement rural ``Mamokatra'', FTM}.
+  National}. Tech. rep., {Ministère de l'Environnement de Madagascar,
+  Direction des Eaux et Forêts, DFS Deutsch Forest Service GmbH, Entreprise
+  d'études de développement rural "Mamokatra", FTM}.
 
@@
 Minten, B., Sander, K. \& Stifel, D. (2013) {Forest management and economic
-  rents: Evidence from the charcoal trade in Madagascar}.
-\newblock \emph{Energy for Sustainable Development}, \textbf{17}, 106 -- 115.
+  rents: Evidence from the charcoal trade in Madagascar}. \emph{Energy for
+  Sustainable Development} \textbf{17}, 106 -- 115, special Issue on Charcoal.
 
@@
 Murcia, C. (1995) Edge effects in fragmented forests: implications for
-  conservation.
-\newblock \emph{Trends in Ecology \& Evolution}, \textbf{10}, 58 -- 62.
+  conservation. \emph{Trends in Ecology \& Evolution} \textbf{10}, 58 -- 62.
 
@@
 Neteler, M. \& Mitasova, H. (2008) \emph{{Open source GIS: a GRASS GIS
-  approach}}.
-\newblock Springer.
+  approach}}. Springer.
 
@@
   (2008) {Reference scenarios for deforestation and forest degradation in
-  support of REDD: a review of data and methods}.
-\newblock \emph{Environmental Research Letters}, \textbf{3}, 025011.
-
-\bibitem[{{ONE, DGF, FTM, MNP, and CI}(2013)}]{ONE2013}
+  support of REDD: a review of data and methods}. \emph{Environmental Research
+  Letters} \textbf{3}, 025011.
+
+\bibitem[{Olofsson \emph{et~al.}(2014)Olofsson, Foody, Herold, Stehman,
+  Woodcock \& Wulder}]{Olofsson2014}
+Olofsson, P., Foody, G.M., Herold, M., Stehman, S.V., Woodcock, C.E. \& Wulder,
+  M.A. (2014) Good practices for estimating area and assessing accuracy of land
+  change. \emph{Remote Sensing of Environment} \textbf{148}, 42 -- 57.
+
+\bibitem[{Olofsson \emph{et~al.}(2013)Olofsson, Foody, Stehman \&
+  Woodcock}]{Olofsson2013}
+Olofsson, P., Foody, G.M., Stehman, S.V. \& Woodcock, C.E. (2013) Making better
+  use of accuracy data in land change studies: Estimating accuracy and area and
+  quantifying uncertainty using stratified estimation. \emph{Remote Sensing of
+  Environment} \textbf{129}, 122 -- 131.
+
+\bibitem[{ONE \emph{et~al.}(2013)}]{ONE2013}
 {ONE, DGF, FTM, MNP, and CI} (2013) \emph{{Evolution de la couverture de
-  forêts naturelles à Madagascar 2005-2010}}.
-
-\bibitem[{{ONE, DGF, MNP, WCS, and Etc Terra}(2015)}]{ONE2015}
+  forêts naturelles à Madagascar 2005-2010}}. \url{https://www.pnae.mg/couverture-de-forets-naturelles-2005-2010}
+
+\bibitem[{ONE \emph{et~al.}(2015)}]{ONE2015}
 {ONE, DGF, MNP, WCS, and Etc Terra} (2015) \emph{{Changement de la couverture
-  de forêts naturelles à Madagascar, 2005-2010-2013}}.
+  de forêts naturelles à Madagascar, 2005-2010-2013}}. \url{https://www.pnae.mg/couverture-de-forets-naturelles-2005-2010-2013}
 
@@
   Madagascar: watershed versus climatic gradient hypotheses evaluated by null
-  biogeographic models}.
-\newblock \emph{Evolution}, \textbf{63}, 959--967.
+  biogeographic models}. \emph{Evolution} \textbf{63}, 959--967.
 
@@
 Pekel, J.F., Cottam, A., Gorelick, N. \& Belward, A.S. (2016) {High-resolution
-  mapping of global surface water and its long-term changes}.
-\newblock \emph{Nature}, \textbf{540}, 418--422.
+  mapping of global surface water and its long-term changes}. \emph{Nature}
+  \textbf{540}, 418--422.
 
@@
 Perrier~de La~B{\^a}thie, H. (1921) \emph{{La v{\'e}g{\'e}tation malgache}},
-  vol.~23.
-\newblock Mus{\'e}e Colonial.
+  vol.~23. Mus{\'e}e Colonial.
 
@@
   Esipova, E. (2017) {The last frontiers of wilderness: Tracking loss of intact
-  forest landscapes from 2000 to 2013}.
-\newblock \emph{Science Advances}, \textbf{3}.
+  forest landscapes from 2000 to 2013}. \emph{Science Advances} \textbf{3}.
 
@@
 Puyravaud, J.P. (2003) Standardizing the calculation of the annual rate of
-  deforestation.
-\newblock \emph{Forest Ecology and Management}, \textbf{177}, 593--596.
+  deforestation. \emph{Forest Ecology and Management} \textbf{177}, 593--596.
 
@@
   utilisation de l'algorithme Random Forest et proc{\'e}dure de validation}.
-\newblock \emph{Revue Fran{\c{c}}aise de Photogramm{\'e}trie et de
-  T{\'e}l{\'e}d{\'e}tection}, pp. 11--23.
+  \emph{Revue Fran{\c{c}}aise de Photogramm{\'e}trie et de
+  T{\'e}l{\'e}d{\'e}tection} pp. 11--23.
 
@@
 Riitters, K., Wickham, J., O'Neill, R., Jones, B. \& Smith, E. (2000)
-  Global-scale patterns of forest fragmentation.
-\newblock \emph{Conservation Ecology}, \textbf{4}, 3.
+  Global-scale patterns of forest fragmentation. \emph{Conservation Ecology}
+  \textbf{4}, 3.
 
@@
   L., Silman, M. \& Morel, A. (2011) Benchmark map of forest carbon stocks in
-  tropical regions across three continents.
-\newblock \emph{Proceedings of the National Academy of Sciences}, \textbf{108},
-  9899--9904.
+  tropical regions across three continents. \emph{Proceedings of the National
+  Academy of Sciences} \textbf{108}, 9899--9904.
 
@@
 Saunders, D.A., Hobbs, R.J. \& Margules, C.R. (1991) {Biological Consequences
-  of Ecosystem Fragmentation: A Review}.
-\newblock \emph{Conservation Biology}, \textbf{5}, 18--32.
+  of Ecosystem Fragmentation: A Review}. \emph{Conservation Biology}
+  \textbf{5}, 18--32.
 
@@
 Scales, I.R. (2011) {Farming at the Forest Frontier: Land Use and Landscape
-  Change in Western Madagascar, 1896-2005}.
-  \newblock \emph{Environment and History}, \textbf{17}, 499--524.
+  Change in Western Madagascar, 1896-2005}. \emph{Environment and History}
+  \textbf{17}, 499--524.
 
@@
 Smith, R.J., Muir, R.D.J., Walpole, M.J., Balmford, A. \& Leader-Williams, N.
-  (2003) Governance and the loss of biodiversity.
-\newblock \emph{Nature}, \textbf{426}, 67--70.
+  (2003) Governance and the loss of biodiversity. \emph{Nature} \textbf{426},
+  67--70.
 
@@
   for the Malagasy flat-tailed tortoise from 1963 through 1993}.
-\newblock \emph{Chelonian Conservation and Biology}, \textbf{4}, 59--65.
+  \emph{Chelonian Conservation and Biology} \textbf{4}, 59--65.
 
@@
   S}{\'\i}mov{\'a}, I. \& Storch, D. (2014) {Comment on "High-resolution global
-  maps of 21st-century forest cover change"}.
-\newblock \emph{Science}, \textbf{344}, 981--981.
+  maps of 21st-century forest cover change"}. \emph{Science} \textbf{344},
+  981--981.
+
+\bibitem[{Tyukavina \emph{et~al.}(2015)Tyukavina, Baccini, Hansen, Potapov,
+  Stehman, Houghton, Krylov, Turubanova \& Goetz}]{Tyukavina2015}
+Tyukavina, A., Baccini, A., Hansen, M.C., Potapov, P.V., Stehman, S.V.,
+  Houghton, R.A., Krylov, A.M., Turubanova, S. \& Goetz, S.J. (2015)
+  Aboveground carbon loss in natural and managed tropical forests from 2000 to
+  2012. \emph{Environmental Research Letters} \textbf{10}, 074002.
 
@@
   Okpa, C. \& Aguilar, R. (2017) {Types and rates of forest disturbance in
-  Brazilian Legal Amazon, 2000{\textendash}2013}.
-\newblock \emph{Science Advances}, \textbf{3}.
+  Brazilian Legal Amazon, 2000{\textendash}2013}. \emph{Science Advances}
+  \textbf{3}.
 
@@
   Combination of Recent Remote Sensing Based Products for Forest Cover Change
-  Assessments in Cameroon}.
-\newblock \emph{International Forestry Review}, \textbf{18}, 14--25.
+  Assessments in Cameroon}. \emph{International Forestry Review} \textbf{18},
+  14--25.
 
@@
   decrease in tropical forest carbon stocks with climate change in Madagascar}.
-\newblock \emph{Journal of Ecology}, \textbf{104}, 703--715.
+  \emph{Journal of Ecology} \textbf{104}, 703--715.
 
@@
   and carbon emissions in tropical developing countries facing demographic
-  expansion: a case study in Madagascar}.
-\newblock \emph{Ecology and Evolution}, \textbf{3}, 1702--1716.
+  expansion: a case study in Madagascar}. \emph{Ecology and Evolution}
+  \textbf{3}, 1702--1716.
 
@@
 Virah-Sawmy, M. (2009) {Ecosystem management in Madagascar during global
-  change}.
-\newblock \emph{Conservation Letters}, \textbf{2}, 163--170.
+  change}. \emph{Conservation Letters} \textbf{2}, 163--170.
 
@@
   (2016) {Madagascar's grasses and grasslands: anthropogenic or natural?}
-\newblock \emph{Proceedings of the Royal Society of London B: Biological
-  Sciences}, \textbf{283}.
+  \emph{Proceedings of the Royal Society of London B: Biological Sciences}
+  \textbf{283}.
 
@@
     \toprule
-    Year & Forest (ha) & Unmapped (ha) & Annual defor. (ha/yr) & Rate
-    (\%/yr)\tabularnewline
+    Year & Forest (ha) & Unmapped (ha) & Defor. area (ha/yr) & Rate
+    (\%/yr) \\
     \midrule
     \endhead
-    1953 & 15968176 & 0 & - & -\tabularnewline
-    1973 & 14242592 & 3316531 & 86279 & 0.57\tabularnewline
-    1990 & 10762442 & 0 & 204715 & 1.63\tabularnewline
-    2000 & 9879031 & 0 & 88341 & 0.85\tabularnewline
-    2005 & 9667553 & 0 & 42296 & 0.43\tabularnewline
-    2010 & 9319851 & 0 & 69540 & 0.73\tabularnewline
-    2014 & 8925246 & 0 & 98651 & 1.08\tabularnewline
+    1953 & 15,968 & 0 & - & - \\
+    1973 & 14,243 & 3,317 & 86 & 0.6 \\
+    1990 & 10,762 & 0 & 205 & 1.6 \\
+    2000 & 9,879 & 0 & 88 & 0.8 \\
+    2005 & 9,668 & 0 & 42 & 0.4 \\
+    2010 & 9,320 & 0 & 70 & 0.7 \\
+    2014 & 8,925 & 0 & 99 & 1.1 \\
     \bottomrule
@@
 
-  \caption{\textbf{Evolution of the forest cover and deforestation
-      rates from 1953 to 2014 in Madagascar}. Forest map for the year
-    1973 has 3.3 Mha of unclassified areas due to the presence of
-    clouds on satellite images. As a consequence, deforestation rates
-    for the periods 1953-1973 and 1973-1990 are indicative. The two
-    last columns indicate the annual deforested areas and annual
-    deforestation rates on the previous time-period (e.g.~1953-1973
-    for year 1973, 1973-1990 for year 1990, etc.).}
+  \caption{\textbf{Evolution of forest cover and deforestation rates
+      from 1953 to 2014 in Madagascar}. Areas are provided in
+    thousands of hectares (Kha). Forest map for the year 1973 has 3.3
+    Mha of unclassified areas due to the presence of clouds on
+    satellite images. As a consequence, partial deforestation rates
+    for the periods 1953-1973 and 1973-1990 are computed based on the
+    available forest extent. The last two columns indicate the annual
+    deforested areas and annual deforestation rates on the previous
+    time-period (e.g.~1953-1973 for year 1973, 1973-1990 for year
+    1990, etc.).}
 
@@
 \begin{table}[!h]
-  {\footnotesize
+  %{\small
   \begin{longtable}[]{@{}llrrrrrrrr@{}}
@@
     Forest type & Source & 1953 & 1973 & 1990 & 2000 & 2005 & 2010 & 2013 &
-    2014\tabularnewline
+    2014 \\
     \midrule
     \endhead
-    Total & Harper2007 & 15995900 & 14173100 & 10605700 & 8982100 & - & - &
-    - & -\tabularnewline
-    & MEFT2009 & - & - & 10650142 & 9678402 & 9413218 & - & - &
-    -\tabularnewline
-    & ONE2015 & - & - & - & - & 9451350 & 8977337 & 8485509 &
-    -\tabularnewline
-    & this study & 15968176 & 14242592 & 10762494 & 9879031 & 9667553 &
-    9319851 & 9051029 & 8925246\tabularnewline
-    Moist & Harper2007 & 8765600 & 6876000 & 5234300 & 4166800 & - & - & - &
-    -\tabularnewline
-    & MEFT2009 & - & - & 5270599 & 4787771 & 4700430 & - & - &
-    -\tabularnewline
-    & ONE2015 & - & - & - & - & 4555788 & 4457184 & 4345093 &
-    -\tabularnewline
-    & this study & 8578299 & 6989942 & 5270169 & 4872016 & 4767876 & 4633104
-    & 4470194 & 4409842\tabularnewline
-    Dry & Harper2007 & 4252100 & 4027700 & 2711800 & 2457000 & - & - & - &
-    -\tabularnewline
-    & MEFT2009 & - & - & 3320582 & 3084976 & 3027505 & - & - &
-    -\tabularnewline
-    & ONE2015 & - & - & - & - & 3223028 & 2970192 & 2678640 &
-    -\tabularnewline
-    & this study & 4761551 & 4434871 & 3224917 & 2940970 & 2880819 & 2734639
-    & 2642253 & 2595621\tabularnewline
-    Spiny & Harper2007 & 2978200 & 3029800 & 2420000 & 2132200 & - & - & - &
-    -\tabularnewline
-    & MEFT2009 & - & - & 2123630 & 1871735 & 1756884 & - & - &
-    -\tabularnewline
-    & ONE2015 & - & - & - & - & 1681527 & 1558533 & 1466765 &
-    -\tabularnewline
-    & this study & 2462830 & 2582880 & 2054724 & 1857628 & 1810704 & 1744427
-    & 1731308 & 1712731\tabularnewline
-    Mangroves & Harper2007 & - & - & 239600 & 226100 & - & - & - &
-    -\tabularnewline
-    & MEFT2009 & - & - & - & - & - & - & - & -\tabularnewline
-    & ONE2015 & - & - & - & - & 173564 & 171220 & 169877 & -\tabularnewline
-    & this study & 143412 & 199853 & 181226 & 177708 & 177492 & 177149 &
-    176890 & 176718\tabularnewline
+    Total & Harper2007 & 15,996 & 14,173 & 10,606 & 8,982 & - & - & - & - \\
+          & MEFT2009 & - & - & 10,650 & 9,678 & 9,413 & - & - & - \\
+                & ONE2015 & - & - & - & - & 9,451 & 8,977 & 8,486 & - \\
+                & this study & 15,968 & 14,243 & 10,762 & 9,879 & 9,668 & 9,320 & 9,051
+                                                                          & 8,925 \\
+    Moist & Harper2007 & 8,766 & 6,876 & 5,234 & 4,167 & - & - & - &
+                                                                     - \\
+                & MEFT2009 & - & - & 5,271 & 4,788 & 4,700 & - & - & - \\
+                & ONE2015 & - & - & - & - & 4,556 & 4,457 & 4,345 & - \\
+                & this study & 8,578 & 6,990 & 5,270 & 4,872 & 4,768 & 4,633 & 4,470 &
+                                                                                       4,410 \\
+    Dry & Harper2007 & 4,252 & 4,028 & 2,712 & 2,457 & - & - & - &
+                                                                   - \\
+                & MEFT2009 & - & - & 3,321 & 3,085 & 3,028 & - & - & - \\
+                & ONE2015 & - & - & - & - & 3,223 & 2,970 & 2,679 & - \\
+                & this study & 4,762 & 4,435 & 3,225 & 2,941 & 2,881 & 2,735 & 2,642 &
+                                                                                       2,596 \\
+    Spiny & Harper2007 & 2,978 & 3,030 & 2,420 & 2,132 & - & - & - &
+                                                                     - \\
+                & MEFT2009 & - & - & 2,124 & 1,872 & 1,757 & - & - & - \\
+                & ONE2015 & - & - & - & - & 1,682 & 1,559 & 1,467 & - \\
+                & this study & 2,463 & 2,583 & 2,055 & 1,858 & 1,811 & 1,744 & 1,731 &
+                                                                                       1,713 \\
+    Mangroves & Harper2007 & - & - & 240 & 226 & - & - & - &
+                                                             - \\
+                & MEFT2009 & - & - & - & - & - & - & - & - \\
+                & ONE2015 & - & - & - & - & 174 & 171 & 170 & - \\
+                & this study & 143 & 200 & 181 & 178 & 177 & 177 & 177 &
+                                                                         177 \\
     \bottomrule
-  \end{longtable}}
+  \end{longtable}%}
   \addtocounter{table}{-1}
 
-  \caption{\textbf{Comparing our estimates of forest-cover (in ha) for
-      Madagascar with previous studies on the period 1953-2014}. We
-    compared our estimates of forest-cover with the estimates from
-    three previous studies \citep{Harper2007, MEFT2009, ONE2015}. We
-    obtained a Pearson's correlation coefficient of 0.99 between our
-    forest-cover estimates and forest-cover estimates from previous
-    studies. The increase in mangrove and spiny forest covers from
-    \emph{c.} 1953 to \emph{c.} 1973 in \citet{Harper2007} and our
-    study is most probably due to differences in forest definition and
-    mapping methods between the 1953 aerial-photography derived map
-    and the 1973 Landsat image derived map.}
+  \caption{\textbf{Comparing Madagascar forest-cover estimates (in
+      Kha) with previous studies on the period 1953-2014}. We compared
+    our estimates of forest-cover with the estimates from three
+    previous studies \citep{Harper2007, MEFT2009, ONE2015}. Areas are
+    provided in thousands of hectares (Kha). We obtained a Pearson's
+    correlation coefficient of 0.99 between our forest-cover estimates
+    and forest-cover estimates from previous studies. The increase in
+    mangrove and spiny forest covers from \emph{c.} 1953 to \emph{c.}
+    1973 in \citet{Harper2007} and our study is most probably due to
+    differences in forest definition and mapping methods between the
+    1953 aerial-photography derived map and the 1973 Landsat image
+    derived map.}
 
@@
 \begin{table}[!h]
-  {\footnotesize
+  %{\small
   \begin{longtable}[]{@{}llrrrrrr@{}}
@@
     Forest type & Source & 1953-1973 & 1973-1990 & 1990-2000 & 2000-2005 &
-    2005-2010 & 2010-2013\tabularnewline
+    2005-2010 & 2010-2013 \\
     \midrule
     \endhead
-    Total & Harper2007 & 91140 (0.30) & 200206 (1.70) & 80740 (0.90) & - & -
-    & -\tabularnewline
-    & MEFT2009 & - & - & 97174 (0.83) & 53037 (0.53) & - & -\tabularnewline
-    & ONE2015 & - & - & - & - & 94803 (1.18) & 163943 (1.50)\tabularnewline
-    & this study & 86279 (0.57) & 204712 (1.63) & 88346 (0.85) & 42296
-    (0.43) & 69540 (0.73) & 89607 (0.97)\tabularnewline
-    Moist & Harper2007 & 94480 (0.60) & 87188 (1.70) & 32200 (0.80) & - & -
-    & -\tabularnewline
-    & MEFT2009 & - & - & 48283 (0.79) & 17468 (0.35) & - & -\tabularnewline
-    & ONE2015 & - & - & - & - & 19721 (0.50) & 37364 (0.94)\tabularnewline
-    & this study & 79418 (1.02) & 101163 (1.65) & 39815 (0.78) & 20828
-    (0.43) & 26954 (0.57) & 54303 (1.19)\tabularnewline
-    Dry & Harper2007 & 11220 (0.20) & 77153 (1.90) & 19820 (0.70) & - & - &
-    -\tabularnewline
-    & MEFT2009 & - & - & 23561 (0.67) & 11494 (0.40) & - & -\tabularnewline
-    & ONE2015 & - & - & - & - & 50567 (1.80) & 97184 (2.29)\tabularnewline
-    & this study & 16334 (0.35) & 71174 (1.86) & 28395 (0.92) & 12030 (0.41)
-    & 29236 (1.04) & 30795 (1.14)\tabularnewline
-    Spiny & Harper2007 & -2580 (-0.10) & 35865 (1.20) & 28170 (1.20) & - & -
-    & -\tabularnewline
-    & MEFT2009 & - & - & 25190 (1.19) & 22970 (1.23) & - & -\tabularnewline
-    & ONE2015 & - & - & - & - & 24599 (1.69) & 30589 (1.66)\tabularnewline
-    & this study & -6002 (-0.24) & 31068 (1.34) & 19710 (1.00) & 9385 (0.51)
-    & 13255 (0.74) & 4373 (0.25)\tabularnewline
-    Mangroves & Harper2007 & - & - & 550 (0.20) & - & - & -\tabularnewline
-    & MEFT2009 & - & - & - & - & - & -\tabularnewline
-    & ONE2015 & - & - & - & - & 469 (0.32) & 448 (0.20)\tabularnewline
-    & this study & -2822 (-1.67) & 1096 (0.57) & 352 (0.20) & 43 (0.02) & 69
-    (0.04) & 86 (0.05)\tabularnewline
+    Total & Harper2007 & 91 (0.3) & 200 (1.7) & 81 (0.9) & - & - &
+                                                                   - \\
+                & MEFT2009 & - & - & 97 (0.8) & 53 (0.5) & - & - \\
+                & ONE2015 & - & - & - & - & 95 (1.2) & 164 (1.5) \\
+                & this study & 86 (0.6) & 205 (1.6) & 88 (0.9) & 42 (0.4) & 70 (0.7) &
+                                                                                       90 (1.0) \\
+    Moist & Harper2007 & 94 (0.6) & 87 (1.7) & 32 (0.8) & - & - &
+                                                                  - \\
+                & MEFT2009 & - & - & 48 (0.8) & 17 (0.4) & - & - \\
+                & ONE2015 & - & - & - & - & 20 (0.5) & 37 (0.9) \\
+                & this study & 79 (1.0) & 101 (1.6) & 40 (0.8) & 21 (0.4) & 27 (0.6) &
+                                                                                       54 (1.2) \\
+    Dry & Harper2007 & 11 (0.2) & 77 (1.9) & 20 (0.7) & - & - &
+                                                                - \\
+                & MEFT2009 & - & - & 24 (0.7) & 11 (0.4) & - & - \\
+                & ONE2015 & - & - & - & - & 51 (1.8) & 97 (2.3) \\
+                & this study & 16 (0.4) & 71 (1.9) & 28 (0.9) & 12 (0.4) & 29 (1.0) & 31
+                                                                                      (1.1) \\
+    Spiny & Harper2007 & -3 (-0.1) & 36 (1.2) & 28 (1.2) & - & - &
+                                                                   - \\
+                & MEFT2009 & - & - & 25 (1.2) & 23 (1.2) & - & - \\
+                & ONE2015 & - & - & - & - & 25 (1.7) & 31 (1.7) \\
+                & this study & -6 (-0.2) & 31 (1.3) & 20 (1.0) & 9 (0.5) & 13 (0.7) & 4
+                                                                                      (0.3) \\
+    Mangroves & Harper2007 & - & - & 1 (0.2) & - & - & - \\
+                & MEFT2009 & - & - & - & - & - & - \\
+                & ONE2015 & - & - & - & - & 0 (0.3) & 0 (0.2) \\
+                & this study & -3 (-1.7) & 1 (0.6) & 0 (0.2) & 0 (0.0) & 0 (0.0) & 0
+                                                                                   (0.0) \\
     \bottomrule
-  \end{longtable}}
+  \end{longtable}%}
   \addtocounter{table}{-1}
 
-  \caption{\textbf{Comparing our estimates of annual deforestation
-      rates for Madagascar with previous studies on the period
-      1953-2014}. Annual deforestation areas (in ha/yr) and annual
-    deforestation rates (second number in parenthesis, in \%/yr) are
-    provided. For deforestation rates in \%/yr, exact same numbers as
-    in scientific articles and reports from previous studies
-    \citep{Harper2007, MEFT2009, ONE2015} have been reported. The way
-    annual deforestation rates in \%/yr have been computed in these
-    previous studies can slightly differ from one study to another, but
+  \caption{\textbf{Comparing Madagascar annual deforestation rates
+      with previous studies on the period 1953-2014}. Annual
+    deforestated areas (in Kha/yr) and annual deforestation rates
+    (second number in parenthesis, in \%/yr) are provided. For
+    deforestation rates in \%/yr, exact same numbers as in scientific
+    articles and reports from previous studies \citep{Harper2007,
+      MEFT2009, ONE2015} have been reported. The way annual
+    deforestation rates in \%/yr have been computed in these previous
+    studies can slightly differ from one study to another, but
     estimates always correct for the potential presences of clouds on
     satellite images and unclassified areas on forest maps. Annual
-    deforested areas in ha/yr have been recomputed from forest-cover
+    deforested areas in Kha/yr have been recomputed from forest-cover
     estimates in Tab.~\ref{tab:comp_forest} (except for
     \citet{Harper2007} for the periods 1973-1990 and 1990-2000 for
-    which annual deforested areas in ha/yr were derived from numbers
+    which annual deforested areas in Kha/yr were derived from numbers
     reported in the original publication, see methods) and do not
-    correct for the potential presence of clouds.}
+    correct for the potential presence of clouds.  }
 
@@
 
-  {\small
+  %{\small
   \begin{longtable}[]{@{}rrrrrrrr@{}}
     \toprule
-    Year & Forest (ha) & patch (\%) & transitional (\%) & edge (\%) &
-    perforated (\%) & interior (\%) & NA (\%)\tabularnewline
+    Year & Forest (Kha) & 0-20\% & 21-40\% & 41-60\% &
+    61-80\% & 81-100\% \\
     \midrule
     \endhead
-    1953 & 15962870 & 0.01 & 1.12 & 4.46 & 0.58 & 93.83 &
-    0.00\tabularnewline
-    1973 & 14228217 & 2.21 & 7.25 & 19.81 & 2.86 & 67.87 &
-    0.01\tabularnewline
-    1990 & 10749572 & 3.00 & 8.17 & 21.28 & 3.81 & 63.73 &
-    0.01\tabularnewline
-    2000 & 9866145 & 3.09 & 8.37 & 22.13 & 3.92 & 62.49 &
-    0.01\tabularnewline
-    2005 & 9659861 & 3.51 & 8.88 & 22.56 & 6.44 & 58.59 &
-    0.02\tabularnewline
-    2010 & 9306528 & 4.28 & 9.72 & 22.94 & 8.52 & 54.52 &
-    0.02\tabularnewline
-    2014 & 8911481 & 5.18 & 10.72 & 23.25 & 10.58 & 50.24 &
-    0.03\tabularnewline
+    1953  & 15,968  & 0  & 1  & 8  & 12  & 78 \\
+    1973  & 14,243  & 6  & 9  & 12  & 16  & 57 \\
+    1990  & 10,762  & 7  & 10  & 13  & 17  & 53 \\
+    2000  & 9,879  & 7  & 11  & 14  & 17  & 51 \\
+    2005  & 9,673  & 8  & 11  & 14  & 18  & 49 \\
+    2010  & 9,320  & 8  & 12  & 15  & 18  & 47 \\
+    2014  & 8,925  & 9  & 13  & 16  & 19  & 44 \\
     \bottomrule
-  \end{longtable}}
+  \end{longtable}%}
   \addtocounter{table}{-1}
 
-  \caption{\textbf{Evolution of the forest fragmentation from 1953 to 2014
-      in Madagascar}. Six categories of fragmentation were identified from the
-    amount of forest and its occurrence as adjacent forest pixels:
-    ``interior'', ``perforated'', ``edge'', ``transitional'', ``patch'', and
-    ``undetermined'' \citep{Riitters2000}. We used a moving window of 7x7
-    pixels (4.4~ha). Using this window size, forest edge had a width of
-    about 90~m. The ``interior'' category can be interpreted as the most
-    intact forest. The ``patch'' and ``transitional'' categories correspond
-    to isolated small forest patches.}
+  \caption{\textbf{Evolution of the forest fragmentation from 1953 to
+      2014 in Madagascar}. Five forest fragmentation classes, based on
+    the percentage of forest in the neighborhood, are defined: 0-20\%
+    (highly fragmented), 21-40\%, 41-60\%, 61-80\% and 81-100\% (lowly
+    fragmented). The percentage of forest falling in each forest
+    fragmentation class is reported for each year. Forest areas are
+    provided in thousands of hectares (Kha).}
 
@@
 
-    \includegraphics[width=11cm]{ecoregion.png}
+    \includegraphics[width=11cm]{figs/ecoregion.png}
 
     \caption{\textbf{Ecoregions and forest types in Madagascar.}
-      Madagascar can be divided into four climatic ecoregions with four forest
-      types: the moist forest in the East (green), the dry forest in the West
-      (orange), the spiny forest in the South (red), and the mangroves on the
-      West coast (blue). Ecoregions were defined following climatic
-      \citep{Cornet1974} and vegetation \citep{IEFN1996} criteria. The dark
-      grey areas represent the remaining natural forest cover for the year
-      2014.}
+      Madagascar can be divided into four climatic ecoregions with
+      four forest types: the moist forest in the East (green), the dry
+      forest in the West (orange), the spiny forest in the South
+      (red), and the mangroves on the West coast (blue). Ecoregions
+      were defined following climatic \citep{Cornet1974} and
+      vegetation \citep{IEFN1996} criteria. The dark grey areas
+      represent the remaining natural forest cover for the year 2014.}
 
@@
 
-  \includegraphics[width=\textwidth]{fig_fcc_highres.png}
-
-  \caption{\textbf{Forest-cover change on six decades from 1953
-      to 2014 in Madagascar.} Forest cover changes from \emph{c.} 1973 to 2014
-    are shown in the main figure, and forest cover in \emph{c.} 1953 is
-    shown in the bottom-right inset. Two zooms in the western dry (left
-    part) and eastern moist (right part) ecoregions present more detailed
-    views of (from top to bottom): forest-cover in 1950s, forest-cover
-    change from \emph{c.} 1973 to 2014, forest fragmentation in 2014 and
-    distance to forest edge in 2014. Data on water bodies (blue) and water
-    seasonality (light blue for seasonal water to dark blue for permanent
-    water) has been extracted from \citet{Pekel2016}.}
+  \includegraphics[width=\textwidth]{figs/fig_fcc.png}
+
+  \caption{\textbf{Forest-cover change on six decades from 1953 to
+      2014 in Madagascar.} Forest cover changes from \emph{c.} 1973 to
+    2014 are shown in the main figure, and forest cover in \emph{c.}
+    1953 is shown in the bottom-right inset. Two zooms in the western
+    dry (left part) and eastern moist (right part) ecoregions present
+    more detailed views of (from top to bottom): forest-cover in
+    1950s, forest-cover change from \emph{c.} 1973 to 2014, forest
+    fragmentation in 2014 and distance to forest edge in 2014. Data on
+    water bodies (blue) and water seasonality (light blue for seasonal
+    water to dark blue for permanent water) has been extracted from
+    \citet{Pekel2016}.}
 
@@
 
-  \includegraphics[width=13cm]{dist.png}
-
-  \caption{\textbf{Evolution of the distance to forest edge from
-      1953 to 2014 in Madagascar.} Black dots represent the mean distance to
-    forest edge for each year. Vertical dashed segments represent the 90\%
-    quantiles (5\% and 95\%) of the distance to forest edge. Horizontal
-    dashed grey line indicates a distance to forest edge of 100~m.
-    Percentages indicate the percentage of forest at a distance to forest
-    edge lower than 100~m for each year.}
+  \includegraphics[width=12cm]{figs/dist.png}
+
+  \caption{\textbf{Evolution of the distance to forest edge from 1953
+      to 2014 in Madagascar.} Black dots represent the mean distance
+    to forest edge for each year. Vertical dashed segments represent
+    the 90\% quantiles (5\% and 95\%) of the distance to forest
+    edge. Horizontal dashed grey line indicates a distance to forest
+    edge of 100~m. Numbers at the bottom of each vertical segments are
+    the percentage of forest at a distance to forest edge lower than
+    100~m for each year.}
 

### 2018-01-12 deforestmap.R

@@ -996,8 +996,8 @@ render("coverletter.md",output_format=c("pdf_document"),output_dir="report") # p
 ## Pandoc for word
 ##========================
 
-system("pandoc manuscript/manuscript_essd.tex --bibliography=manuscript/biblio.bib \\
-       --csl=manuscript/bib/methods-in-ecology-and-evolution.csl -o manuscript/manuscript_essd.docx")
+system("pandoc manuscript/manuscript.tex --bibliography=manuscript/bib/biblio.bib \\
+       --csl=manuscript/bib/journal-of-applied-ecology.csl -o manuscript/manuscript_essd.md")
 
 ##===========================================================================
 ## End of script

### 2018-01-15 manuscript/manuscript.Rmd
Use of bookdown

@@
 title: "Combining global tree cover loss data with historical national forest-cover maps to look at six decades of deforestation and forest fragmentation in Madagascar"
+date: '`r format(Sys.time(), "%d %B, %Y")`'
+output:
+  bookdown::tufte_html2:
+    number_sections: no
+    toc: yes
+  bookdown::word_document2: null
+  bookdown::pdf_document2:
+    #includes:
+    #  before_body: doc_prefix.tex
+    #  in_header: preamble.tex
+    keep_tex: yes
+    number_sections: yes
+    toc: no
 bibliography: bib/biblio.bib
+fontsize: 12pt
+link-citations: yes
 csl: bib/journal-of-applied-ecology.csl
-output: 
-  pdf_document:
-    citation_package: natbib
-    fig_caption: true
-    keep_tex: true
 ---
@@
 
-The island of Madagascar has an unparalleled biodiversity, mainly located in the tropical forests of the island, which is highly
-  threatened by anthropogenic deforestation. Scattered forest maps
-  from past studies at national level with substantial gaps (due to
-  presence of cloud cover on satellite imagery) prevent the analysis
-  of long-term deforestation trends in Madagascar. In this study, we
-  propose a new approach combining historical (1953-2000) national
-  forest-cover maps with recent (2001-2014) global annual tree cover
-  loss data to look at six decades (1953-2014) of deforestation and
-  forest fragmentation in Madagascar. We produced new forest-cover
-  maps at 30~m resolution over the full territory of Madagascar for
-  the year 1990, and annually from 2000 to 2014. We estimated that
-  Madagascar has lost 44\% of its natural forest cover over the period
-  1953-2014 (including 37\% over the period 1973-2014). Natural
-  forests cover 8.9 Mha in 2014 (15\% of the national territory) which
-  are divided into 4.4 Mha (50\%) of moist forests, 2.6 Mha (29\%) of
-  dry forests, 1.7 Mha of spiny forests (19\%) and 177,000 ha (2\%) of
-  mangroves. Since 2005, the annual deforestation rate has
-  progressively increased in Madagascar to reach 99,000 ha/yr during
-  2010-2014 (corresponding to a rate of 1.1\%/yr). This increase is
-  probably due to rapid population growth (close to 3\%/yr) and to
-  poor law enforcement in the country. Around half of the forest
-  (46\%) is now located at less than 100~m from the forest
-  edge. Accurate forest-cover change maps can be used to assess the
-  effectiveness of past and current conservation programs and
-  implement new strategies for the future. In particular, forest maps
-  and estimates can be used in the REDD+ framework which aims at
-  ``Reducing Emissions from Deforestation and Forest Degradation'' and
-  for optimizing the current protected area network.
+The island of Madagascar has a unique biodiversity, mainly located in the tropical forests of the island.
+This biodiversity is highly threatened by anthropogenic deforestation.
+Existing historical forest maps at national level are scattered and have substantial gaps which prevent an exhaustive assessment of long-term deforestation trends in Madagascar.
+In this study, we combine historical national forest cover maps (covering the period 1953-2000) with a recent global annual tree cover loss dataset (2001-2014) to look at six decades of deforestation and forest fragmentation in Madagascar (from 1953 to 2014).
+We produced new forest cover maps at 30 m resolution for the year 1990 and annually from 2000 to 2014 over the full territory of Madagascar.
+We estimated that Madagascar has lost 44% of its natural forest cover over the period 1953-2014 (including 37% over the period 1973-2014).
+Natural forests cover 8.9 Mha in 2014 (15% of the national territory) and include 4.4 Mha (50%) of moist forests, 2.6 Mha (29%) of dry forests, 1.7 Mha of spiny forests (19%) and 177,000 ha (2%) of mangroves.
+Since 2005, the annual deforestation rate has progressively increased in Madagascar to reach 99,000 ha/yr during 2010-2014 (corresponding to a rate of 1.1%/yr).
+This increase is probably due to rapid population growth (close to 3%/yr) and to poor law enforcement in the country.
+Around half of the forest (46%) is now located at less than 100 m from the forest edge.
+Accurate forest cover change maps can be used to assess the effectiveness of past and current conservation programs and implement new strategies for the future.
+In particular, forest maps and estimates can be used in the REDD+ framework which aims at "Reducing Emissions from Deforestation and Forest Degradation" and for optimizing the current protected area network.
 
-**Keywords**: biodiversity, climate-change, deforestation, Landsat, Madagascar, tropical forest.
+**Keywords**: biodiversity, climate-change, deforestation, forest-fragmentation, Madagascar, tropical forest.
 
-## 1. Introduction
+## Introduction
 
-Separated from the African continent and the Indian plate about 165 and 88 million years ago respectively [@Ali2008], the flora and fauna of Madagascar followed its own evolutionary path.
-Isolation combined with a high number of micro-habitats [@Pearson2009] has led to Madagascar's exceptional biodiversity both in term of number of species and endemism in many taxonomic groups [@Crottini2012; @Goodman2005].
-Most of the biodiversity in Madagascar is concentrated in the tropical forests of the island which can be divided into three types: the moist forest in the East, the dry forest in the West and the spiny forest in the South [@Vieilledent2016].
-This unparalleled biodiversity is severely threatened by anthropogenic deforestation [@Harper2007; @Vieilledent2013] associated with activities such as slash-and-burn agriculture and pasture [@Scales2011].
-Tropical forests in Madagascar also store a large amount of carbon [@Vieilledent2016] and high rates of deforestation in Madagascar are responsible for large CO<sub>2</sub> emissions in the atmosphere [@Achard2014].
-Deforestation threatens species survival by directly reducing their available habitat [@Brooks2002; @Tidd2001].
-Forest fragmentation can also lead to species extinction by isolating populations from each other and creating forest patches too small to maintain viable populations [@Saunders1991].
-Fragmentation also increases forest edge where ecological conditions (such as air temperature, light intensity and air moisture) can be dramatically modified, with consequences on the abundance and distribution of species [@Murcia1995].
-Forest fragmentation can also have substantial effects on forest carbon storage capacity, as carbon stocks are much lower at the forest edge than under a closed canopy [@Brinck2017].
-Moreover, forest carbon stocks vary spatially due to climate or soil factors [@Saatchi2011; @Vieilledent2016].
-As a consequence, accurate and spatially explicit maps of forest-cover and forest-cover change are necessary to monitor biodiversity loss and carbon emissions from deforestation and forest fragmentation, assess the efficiency of present conservation strategies [@Eklund2016], and implement new strategies for the future [@Vieilledent2013; @Vieilledent2016].
-Simple time-series of forest-cover estimates, such as those provided by the FAO Forest Resource Assessment report [@Keenan2015] are not sufficient.
+Separated from the African continent and the Indian plate about 165 and
+88 million years ago respectively [@Ali2008], the flora and fauna of
+Madagascar followed its own evolutionary path. Isolation combined with a
+high number of micro-habitats [@Pearson2009] has led to Madagascar's
+exceptional biodiversity both in term of number of species and endemism
+in many taxonomic groups [@Crottini2012; @Goodman2005]. Most of the
+biodiversity in Madagascar is concentrated in the tropical forests of
+the island which can be divided into four types: the moist forest in the
+East, the dry forest in the West, the spiny forest in the South and the
+mangroves on the West coast [@Vieilledent2016]. This unparalleled
+biodiversity is severely threatened by deforestation
+[@Harper2007; @Vieilledent2013] associated with human activities such as
+slash-and-burn agriculture and pasture [@Scales2011]. Tropical forests
+in Madagascar also store a large amount of carbon [@Vieilledent2016] and
+high rates of deforestation in Madagascar are responsible for large
+CO$_2$ emissions in the atmosphere [@Achard2014]. Deforestation
+threatens species survival by directly reducing their available habitat
+[@Brooks2002; @Tidd2001]. Forest fragmentation can also lead to species
+extinction by isolating populations from each other and creating forest
+patches too small to maintain viable populations [@Saunders1991].
+Fragmentation also increases forest edge where ecological conditions
+(such as air temperature, light intensity and air moisture) can be
+dramatically modified, with consequences on the abundance and
+distribution of species [@Murcia1995]. Forest fragmentation can also
+have substantial effects on forest carbon storage capacity, as carbon
+stocks are much lower at the forest edge than under a closed canopy
+[@Brinck2017]. Moreover, forest carbon stocks vary spatially due to
+climate or soil factors [@Saatchi2011; @Vieilledent2016]. As a
+consequence, accurate and spatially explicit maps of forest cover and
+forest cover change are necessary to monitor biodiversity loss and
+carbon emissions from deforestation and forest fragmentation, assess the
+efficiency of present conservation strategies [@Eklund2016], and
+implement new strategies for the future
+[@Vieilledent2013; @Vieilledent2016]. Simple time-series of forest cover
+estimates, such as those provided by the FAO Forest Resource Assessment
+report [@Keenan2015] are not sufficient.
 
-Unfortunately, accurate and exhaustive forest-cover maps are not available for Madagascar for the last fifteen years (2000-2015).
-@Harper2007 produced maps of forest cover and forest cover changes over Madagascar for the years c.
-1953, c.
-1973, 1990 and 2000.
-The c.
-1953 forest map was derived from the visual interpretation of aerial photography at coarse scale (1/1,000,000).
-Forest maps for the years c.
-1973, 1990 and 2000, were obtained from supervised classification of Landsat satellite images at 60m resolution (for the year 1973) or 30m resolution (for years 1990 and 2000) and can be used to derive accurate estimates (89.5% accuracy reported for the forest / non-forest map of year 2000).
-Maps provided by @Harper2007 were not exhaustive (due to the presence of clouds in the satellite imagery), e.g.
-11 244 km2 are mapped as unknown cover type for the year 2000.
-Using a similar supervised classification approach as in @Harper2007, more recent maps have been produced for the periods 2000-2005-2010 by national institutions, with the technical support of international environmental NGOs [@MEFT2009; @ONE2013].
-Another set of recent forest-cover maps using an advanced statistical tool for classification, the Random Forest classifier [@Grinand2013; @Rakotomalala2015], was produced for the periods 2005-2010-2013 [@ONE2015].
-However, these maps are either too old to give recent estimates of deforestation [@MEFT2009; @ONE2013], include large areas of missing information (due to images with more than 20% of clouds) for the 2010 map provided by [@ONE2013], or show large mis-classification in specific areas, especially in the dry and spiny forest domain for which the spectral answer has a strong seasonal behavior due to the deciduousness of such forests (overall accuracy is lower than 0.8 for the dry and spiny forests for the maps produced by @ONE2015).
-Moreover, the production of such forest maps from a supervised classification approach requires significant resources, especially regarding the image selection step (required to minimize cloud cover) and the training step (visual interpretation of a large number of polygons needed to train the classification algorithm) [@Rakotomalala2015].
-Most of this work of image selection and visual interpretation would need to be repeated to produce new forest maps in the future using a similar approach.
+Unfortunately, accurate and exhaustive forest cover maps are not
+available for Madagascar after year 2000. @Harper2007 produced maps of
+forest cover and forest cover changes over Madagascar for the years
+_c._ 1953, _c._ 1973, 1990 and 2000. The _c._ 1953 forest map was
+derived from the visual interpretation of aerial photography at coarse
+scale (1/1,000,000). Forest maps for the years _c._ 1973, 1990, and 2000
+were obtained from supervised classification of Landsat satellite images
+at 60 m resolution (for the year 1973) or 30 m resolution (for years
+1990 and 2000) and can be used to derive more accurate estimates of
+forest cover (89.5% accuracy reported for the forest/non-forest map of
+year 2000). Nonetheless, maps provided by @Harper2007 are not exhaustive
+(due to the presence of clouds in the satellite imagery), e.g. 11
+244 km2 are mapped as unknown cover type for the year 2000. Using a
+similar supervised classification approach as in @Harper2007, more
+recent maps have been produced for the periods 2000-2005-2010 by
+national institutions, with the technical support of international
+environmental NGOs [@MEFT2009; @ONE2013]. Another set of recent forest
+cover maps using an advanced statistical tool for classification, the
+Random Forest classifier [@Grinand2013; @Rakotomalala2015], was produced
+for the periods 2005-2010-2013 [@ONE2015]. However, these maps are
+either too old to give recent estimates of deforestation
+[@MEFT2009; @ONE2013], include large areas of missing information due to
+images with high percentage of cloud cover [@ONE2013], or show large
+mis-classification in specific areas, especially in the dry and spiny
+forest domain for which the spectral answer has a strong seasonal
+behavior due to the deciduousness of such forests (overall accuracy is
+lower than 0.8 for the dry and spiny forests for the maps produced by
+@ONE2015). Moreover, the production of such forest maps from a
+supervised classification approach requires significant resources,
+especially regarding the image selection step (required to minimize
+cloud cover) and the training step (visual interpretation of a large
+number of polygons needed to train the classification algorithm)
+[@Rakotomalala2015]. Most of this work of image selection and visual
+interpretation would need to be repeated to produce new forest maps in
+the future using a similar approach.
 
-Global forest or tree cover products have also been published recently and can be tested at the national scale for Madagascar.
-[@Kim2014] produced a global forest-cover change map from 1990 to 2000 (derived from Landsat imagery).
-This product was updated to cover the period 1975-2005 (<http://glcf.umd.edu/data/landsatFCC/>) but forest-cover maps after 2005 were not produced.
-Moreover, the approach used in @Kim2014 did not accurately map the forests in the dry and spiny ecosystems of Madagascar (see Fig.
-8 in @Kim2014).
-@Hansen2013 mapped tree cover percentage, annual forest loss and forest gain from 2000 to 2012 at global scale at 30 m resolution.
-This product has since been updated and is now available up to the year 2014 [@Hansen2013].
-To map forest cover from the @Hansen2013 product, a tree cover threshold must be selected (that defines forest cover).
-Selecting such a threshold is not straightforward as the accuracy of the global tree cover map strongly varies between forest types, and is substantially lower for dry forests than for moist forests [@Bastin2017].
-Moreover, the @Hansen2013 product does not provide information on land-use.
-In particular the global tree cover map does not separate tree plantations such as oil palm or eucalyptus plantations from natural forests [@Tropek2014].
-Thus, the global tree cover map from @Hansen2013 cannot be used alone to produce a map of forest cover [@Tyukavina2017].
-In complement to the tree cover percentage provided in @Hansen2013, a layer of annual tree cover loss is also provided (i.e.
-complete loss of tree cover from a value higher than 10% to zero) for the period 2001-2014.
+Global forest or tree cover products have also been published recently
+and can be tested at the national scale for Madagascar. @Kim2014
+produced a global forest cover change map from 1990 to 2000 (derived
+from Landsat imagery). This product was updated to cover the period
+1975-2005 (<http://glcf.umd.edu/data/landsatFCC/>) but forest cover maps
+after 2005 were not produced. Moreover, the approach used in @Kim2014
+did not accurately map the forests in the dry and spiny ecosystems of
+Madagascar (see Fig. 8 in @Kim2014). @Hansen2013 mapped tree cover
+percentage, annual tree cover loss and gain from 2000 to 2012 at global
+scale at 30 m resolution. This product has since been updated and is now
+available up to the year 2014 [@Hansen2013]. To map forest cover from
+the @Hansen2013 product, a tree cover threshold must be selected (that
+defines forest cover). Selecting such a threshold is not straightforward
+as the accuracy of the global tree cover map strongly varies between
+forest types, and is substantially lower for dry forests than for moist
+forests [@Bastin2017]. Moreover, the @Hansen2013 product does not
+provide information on land-use. In particular the global tree cover map
+does not separate tree plantations such as oil palm or eucalyptus
+plantations from natural forests [@Tropek2014]. Thus, the global tree
+cover map from @Hansen2013 cannot be used alone to produce a map of
+forest cover [@Tyukavina2017].
 
-In this study, we present a simple approach which combines the maps from @Harper2007 and products from @Hansen2013 to derive annual wall-to-wall forest-cover change maps over the period 2000-2014 for Madagascar.
-We use the forest-cover map provided by @Harper2007 for the year 2000 (defining the land-use) with the tree cover loss product provided by @Hansen2013 that we apply only inside forest areas identified by @Harper2007.
-Similar to the approach of @Harper2007, we also assess trends in deforestation rates and forest fragmentation from c.
-1953 to 2014.
-The approach described in this study can help assess the effectiveness of current conservation strategies, and assist the implementation of future strategies.
-Our approach could be easily extended to other tropical countries that have at least one forest-cover map between 2000 and 2014.
-This approach can easily be repeated in the future when the @Hansen2013 products are updated.
+In this study, we present a simple approach which combines the
+historical forest maps from @Harper2007 and more recent global products
+from @Hansen2013 to derive annual wall-to-wall forest cover change maps
+over the period 2000-2014 for Madagascar. We use the forest cover map
+provided by @Harper2007 for the year 2000 (defining the land-use) with
+the tree cover loss product provided by @Hansen2013 that we apply only
+inside forest areas identified by @Harper2007. Similar to the approach
+of @Harper2007, we also assess trends in deforestation rates and forest
+fragmentation from _c._ 1953 to 2014. We finally discuss the possibility
+to extend our approach to other tropical countries or repeat it in the
+future. We also discuss how our results could help assess the
+effectiveness of current conservation strategies, and implementation new
+conservation strategies for the future in Madagascar.
 
-## 2. Materials and Methods
+## Materials and Methods
 
-### 2.1 Creation of new forest-cover maps of Madagascar from 1953 to 2014
+### Creation of new forest cover maps of Madagascar from 1953 to 2014
 
-We produced annual forest/non-forest maps at 30m resolution for the full territory of Madagascar for the period 2000-2014 by combining the forest map of year 2000 from @Harper2007, and the tree cover percentage and annual forest cover loss maps over the period 2000-2014 from @Hansen2013.
-The 2000 Harper's forest map includes 208,000 ha of unclassified areas due to the presence of clouds in input satellite imagery, mostly (88%) within the moist forest domain which covers 4.17 million ha in total.
-To provide a label (forest or non-forest) to these unclassified pixels, we use the 2000 tree cover percentage map of @Hansen2013 by selecting a threshold of 75% tree cover to define forest cover as recommended by other studies for the moist domain [@Achard2014].
-We thus obtained a forest-cover map for the year 2000 covering the full territory of Madagascar.
-We then combined this forest-cover map of the year 2000 with the annual forest cover loss maps from 2001 to 2014 provided by @Hansen2013 to create annual forest-cover maps from 2001 to 2014 at 30m resolution.
-We also completed the Harper's forest map of year 1990 by filling unclassified areas (due to the presence of clouds on satellite images) using our forest-cover map of year 2000.
-To do so, we assumed that if forest was present in 2000, the pixel was also forested in 1990.
-The remaining unclassified pixels were limited to a relatively small total area of c.
-8,000 ha.
-We labeled these residual pixels as non-forest as for the year 2000.
-Similarly we completed the Harper forest map of year 1973 by filling unclassified areas using our forest-cover map of the year 1990 assuming that if forest was present in 1990, it was also present in 1973.
-Contrary to the year 1990, the remaining unclassified pixels for year 1973 corresponded to a significant total area of 3.32 million ha.
-We also reprojected the forest-cover map of year 1953 to a common projection in order to compare to the other eras.
-This map was produced by scanning a paper map derived from aerial photos, and thus could not be perfectly aligned with the other maps produced through digital processing of satellite imagery [@Harper2007].
-Finally for all forest-cover maps from 1973, the isolated single non-forest pixels (i.e.
-fully surrounded by forest pixels) were removed, assuming that single non-forest pixels inside a forest patch were not corresponding to deforestation (they might correspond to selective logging activities).
-This allowed us to avoid counting very small scale events (<0.1 ha such as selective logging) as forest fragmentation.
-All the resulting maps are freely available at <https://bioscenemada.cirad.fr/forestmaps>.
+We produced annual forest/non-forest maps at 30 m resolution for the
+full territory of Madagascar for the period 2000-2014 by combining the
+forest map of year 2000 from @Harper2007, and the tree cover percentage
+and annual tree cover loss maps over the period 2000-2014 from
+@Hansen2013. The 2000 Harper's forest map includes 208,000 ha of
+unclassified areas due to the presence of clouds on satellite images,
+mostly (88%) within the moist forest domain which covered 4.17 Mha in
+total in 2000. To provide a label (forest or non-forest) to these
+unclassified pixels, we used the 2000 tree cover percentage map of
+@Hansen2013 by selecting a threshold of 75% tree cover to define forest
+cover as recommended by other studies for the moist domain
+[@Achard2014; @Aleman2017]. To do so, the Hansen's 2000 tree cover map
+was resampled on the same grid as the original Harper's map at 30 m
+resolution using a bilinear interpolation. We thus obtained a forest
+cover map for the year 2000 covering the full territory of Madagascar.
+We then combined this forest cover map of the year 2000 with the annual
+tree cover loss maps from 2001 to 2014 provided by @Hansen2013 to create
+annual forest cover maps from 2001 to 2014 at 30 m resolution. To do so,
+Hansen's tree cover loss maps were resampled on the same grid as the
+original Harper's map at 30 m resolution using a nearest-neighbor
+interpolation. We also completed the Harper's forest map of year 1990 by
+filling unclassified areas (due to the presence of clouds on satellite
+images) using our forest cover map of year 2000. To do so, we assumed
+that if forest was present in 2000, the pixel was also forested in 1990.
+The remaining unclassified pixels were limited to a relatively small
+total area of _c._ 8,000 ha. We labeled these residual pixels as
+non-forest, as for the year 2000. Similarly we completed the Harper's
+forest map of year 1973 by filling unclassified areas using our forest
+cover map of the year 1990 assuming that if forest was present in 1990,
+it was also present in 1973. Contrary to the year 1990, the remaining
+unclassified pixels for year 1973 corresponded to a significant total
+area of 3.3 million ha. We also reprojected the forest cover map of year
+1953 to a common projection in order to compare the forest cover area in
+1953 with forest cover areas at the following dates. This map was
+produced by scanning a paper map derived from aerial photos, and thus
+could not be perfectly aligned with the other maps produced through
+digital processing of satellite imagery [@Harper2007]. Finally for all
+forest cover maps from 1973, the isolated single non-forest pixels
+(i.e. fully surrounded by forest pixels) were removed, assuming that
+single non-forest pixels inside a forest patch were not corresponding to
+deforestation (they might correspond to selective logging activities).
+This allowed us to avoid counting very small scale events (\<0.1 ha,
+such as selective logging) as forest fragmentation. All the resulting
+maps have been made permanently and freely available on the Zenodo
+research data repository at <https://doi.org/10.5281/zenodo.1145785>.
 
-### 2.2 Computing forest-cover areas and deforestation rates
+### Computing forest cover areas and deforestation rates
 
-From these new forest-cover maps, we calculated the total forest-cover area for seven available years (1953-1973-1990-2000-2010-2005-2014), and the annual deforested area and annual deforestation rate for the corresponding six time periods between 1953 and 2014.
-The annual deforestation rates were calculated as follows [@Puyravaud2003; @Vieilledent2013]:
+From these new forest cover maps, we calculated the total forest cover
+area for seven available years (1953-1973-1990-2000-2005-2010-2014), and
+the annual deforested area and annual deforestation rate for the
+corresponding six time periods between 1953 and 2014. The annual
+deforestation rates were calculated using Eq. \@ref(eq:theta) [@Puyravaud2003; @Vieilledent2013]:
 
-$$\theta = 100 \times [1-(1-(F_{t_2}-F_{t_1})/F_{t_1})^{(1/(t_2-t_1))}]$$
+\begin{equation}
+  \theta = 100 \times [1-(1-(F_{t_2}-F_{t_1})/F_{t_1})^{(1/(t_2-t_1))}]
+  (\#eq:theta)
+\end{equation}
 
-where $\theta$ is the annual deforestation rate (in %/yr), $F_{t_2}$ and $F_{t_1}$ are the forest cover free of clouds at both dates $t_2$ and $t_1$, and $t_2-t_1$ is the time-interval (in years) between the two dates.
+In Eq. \ref{eq:theta}, $\theta$ is the annual deforestation rate (in
+%/yr), $F_{t_2}$ and $F_{t_1}$ are the forest cover free of clouds at
+both dates $t_2$ and $t_1$, and $t_2-t_1$ is the time-interval (in
+years) between the two dates.
 
-Because of the large unclassified area (3.32 million ha) in 1973, the annual deforestation areas and rates for the two periods 1953-1973 and 1973-1990 are only indicative estimates.
-For these two periods the annual deforestation rates are computed as the ratio $(F_{t_2}-F_{t_1})/F_{t_1}$ considering only the mapped forest pixels.
-Area and rate estimates are produced at the national scale and for the four forest ecosystems present in Madagascar: moist forest in the East, dry forest in the West, spiny forest in the South, and mangroves on the Western coast (Fig.
-1).
-To define the forest domains, we used a map from the MEFT (_"Ministère de l'Environnement et des Forêts à Madagascar"_) with the boundaries of the four ecoregions in Madagascar.
-Ecoregions were defined on the basis of climatic and vegetation criteria using the climate classification by @Cornet1974 and the vegetation classification from the 1996 IEFN national forest inventory [@IEFN1996].
-Because mangrove forests are highly dynamic ecosystems that can expand or contract on decadal scales depending on changes in environmental factors [@Armitage2015], a fixed delimitation of the mangrove ecoregion on six decades might not be fully appropriate.
-As a consequence, our estimates of the forest-cover and deforestation rates for mangroves in Madagascar must be considered with this limitation.
+Because of the large unclassified area (3.3 million ha) in 1973, the
+annual deforestation areas and rates for the two periods 1953-1973 and
+1973-1990 are only partial estimates computed on the basis of the
+available forest extent. Area and rate estimates are produced at the
+national scale and for the four forest ecosystems present in Madagascar:
+moist forest in the East, dry forest in the West, spiny forest in the
+South, and mangroves on the Western coast
+(Fig. \ref{fig:ecoregions}). To define the forest domains, we used a map
+from the MEFT (*"Ministère de l'Environnement et des Forêts à
+Madagascar"*) with the boundaries of the four ecoregions in Madagascar.
+Ecoregions were defined on the basis of climatic and vegetation criteria
+using the climate classification by @Cornet1974 and the vegetation
+classification from the 1996 IEFN national forest inventory [@IEFN1996].
+Because mangrove forests are highly dynamic ecosystems that can expand
+or contract on decadal scales depending on changes in environmental
+factors [@Armitage2015], a fixed delimitation of the mangrove ecoregion
+on six decades might not be fully appropriate. As a consequence, our
+estimates of the forest cover and deforestation rates for mangroves in
+Madagascar must be considered with this limitation.
 
-### 2.3 Comparing our forest-cover and deforestation rate estimates with previous studies
+### Comparing our forest cover and deforestation rate estimates with previous studies
 
-We compared our estimates of forest-cover and deforestation rates with estimates from the three existing studies at the national scale for Madagascar: (i) [@Harper2007], (ii) [@MEFT2009] and (iii) [@ONE2015].
-@Harper2007 provides forest-cover and deforestation estimates for the periods c.
-1953-c.
-1973-1990-2000.
-MEFT, USAID, and CI (2009) provides estimates for the periods 1990-2000-2005 and ONE, DGF, MNP, WCS, and Etc Terra (2015) provides estimates for the periods 2005-2010-2013.
-To compare our forest-cover and deforestation estimates over the same time periods, we consider an additional time-period in our study (2010-2013) by creating an extra forest-cover map for the year 2013.
-We computed the Pearson's correlation coefficient and the root mean square error (RMSE) between our forest-cover estimates and forest-cover estimates from previous studies for all the dates and forest types (including also the total forest cover estimates).
-For previous studies, the computation of annual deforestation rates (in %/yr) is not always detailed and might slightly differ from one study to another [see @Puyravaud2003].
-@Harper2007 also provide total deforested areas for the two periods 1973-1990 and 1990-2000.
-We converted these values into annual deforested area estimates.
-When annual deforested areas were not reported (for 1953-1973 in @Harper2007 and in @MEFT2009 and @ONE2015), we computed them from the forest-cover estimates in each study.
-These estimates cannot be corrected from the potential bias due to the presence of residual clouds.
-Forest-cover and deforestation rates were then compared between all studies for the whole of Madagascar and the four ecoregions.
-The same ecoregion boundaries as in our study were used in @ONE2015 but this was not the case for @Harper2007 and @MEFT2009, which can explain part of the differences between the estimates.
+We compared our estimates of forest cover and deforestation rates with
+estimates from the three existing studies at the national scale for
+Madagascar: (i) [@Harper2007], (ii) [@MEFT2009] and (iii) [@ONE2015].
+@Harper2007 provides forest cover and deforestation estimates for the
+periods c. 1953-c. 1973-1990-2000. @MEFT2009 provides estimates for the
+periods 1990-2000-2005 and @ONE2015 provides estimates for the periods
+2005-2010-2013. To compare our forest cover and deforestation estimates
+over the same time periods, we consider an additional time-period in our
+study (2010-2013) by creating an extra forest cover map for the year
+2013. We computed the Pearson's correlation coefficient and the root
+mean square error (RMSE) between our forest cover estimates and forest
+cover estimates from previous studies for all the dates and forest types
+(including also the total forest cover estimates). For previous studies,
+the computation of annual deforestation rates (in %/yr) is not always
+detailed and might slightly differ from one study to another [see @Puyravaud2003]. @Harper2007 also provide total deforested areas for the
+two periods 1973-1990 and 1990-2000. We converted these values into
+annual deforested area estimates. When annual deforested areas were not
+reported (for 1953-1973 in @Harper2007 and in @MEFT2009 and @ONE2015),
+we computed them from the forest cover estimates in each study. These
+estimates cannot be corrected from the potential bias due to the
+presence of residual clouds. forest cover and deforestation rates were
+then compared between all studies for the whole of Madagascar and the
+four ecoregions. The same ecoregion boundaries as in our study were used
+in @ONE2015 but this was not the case for @Harper2007 and @MEFT2009,
+which can explain a part of the differences between the estimates.
 
-### 2.3 Fragmentation
+### Fragmentation
 
-We also conducted an analysis of changes in forest fragmentation for the years 1953, 1973, 1990, 2000, 2010 and 2014.
-We applied the method developed by @Riitters2000 which uses a moving window to characterize the fragmentation around each forested pixel.
-Computations were done using the function `r.forestfrag` of the GRASS GIS software [@Neteler2008].
-Six categories of fragmentation were identified from the amount of forest and its occurrence as adjacent forest pixels: "interior", "perforated", "edge", "transitional", "patch", and "undetermined".
-We used a moving window of 7x7 pixels (4.4 ha).
-Using this window size, forest edge had a width of about 90m [@Riitters2000].
-The "interior" category can be interpreted as the most intact forest [@Potapov2017].
-The "patch" and "transitional" categories correspond to isolated small forest patches.
-We reported the area of forest in each fragmentation category for the six years and analyzed the dynamics of fragmentation over the six decades.
-We also computed the distance to forest edge for all forest pixels for the years 1953, 1973, 1990, 2000, 2005, 2010 and 2014.
-For that, we used the function `gdal_proximity.py` of the GDAL software (http://www.gdal.org/).
-We computed the mean and 95% quantiles of the distance to forest edge and looked at the evolution of these values with time.
+We also conducted an analysis of changes in forest fragmentation for the
+years 1953, 1973, 1990, 2000, 2005, 2010 and 2014 at 30 m resolution. We
+used a moving window of $51 \times 51$ pixels centered on each forest
+pixel to compute the percentage of forest pixels in the neighborhood. We
+used this percentage as an indication of the forest fragmentation. The
+size of the moving windows was based on a compromise: a sufficiently
+high number of cells (here 2601) had to be considered to be able to
+compute a percentage and a reasonably low number of cells had to be
+choosen to have a local estimate of the fragmentation. Computations were
+done using the function `r.neighbors` of the GRASS GIS software
+[@Neteler2008]. Using the density of forest in the neighborhood, we
+defined five forest fragmentation classes: 0-20% (highly fragmented),
+21-40%, 41-60%, 61-80% and 81-100% (lowly fragmented). We reported the
+percentage of forest falling in each fragmentation class for the six
+years and analyzed the dynamics of fragmentation over the six decades.
 
-## 3. Results
+We also computed the distance to forest edge for all forest pixels for
+the years 1953, 1973, 1990, 2000, 2005, 2010 and 2014. For that, we used
+the function `gdal_proximity.py` of the GDAL library
+(<http://www.gdal.org/>). We computed the mean and 90% quantiles (5% and
+95%) of the distance to forest edge and looked at the evolution of these
+values with time. Previous studies have shown that forest micro-habitats
+were mainly altered within the first 100 m of the forest edge
+[@Brinck2017; @Gibson2013; @Murcia1995; @Broadbent2008]. Consequently,
+we also estimated the percentage of forest within the first 100 m of the
+forest edge for each year and looked at the evolution of this percentage
+over the six decades.
 
-### 3.1 Dynamics of forest cover and deforestation intensity
+## Results
 
-Natural forests in Madagascar covered 16.0 Mha in 1953, about 27% of the national territory of 587,041 km2.
-In 2014, the forest cover dropped to 8.9 Mha, corresponding to about 15% of the national territory (Fig.
-2 and Tab.
-1).
-Madagascar has lost 44% and 37% of its natural forests between 1953 and 2013, and between c.
-1973 and 2014 respectively (Fig.
-2 and Tab.
-1).
-In 2014 the remaining 8.9 Mha of natural forest were distributed as: 4.4 Mha of moist forest (50% of total forest cover), 2.6 Mha of dry forest (29%), 1.7 Mha of spiny forest (19%) and 0.18 Mha (2%) of mangrove forest (Fig.
-1 and Tab.
-2).
-Regarding the deforestation trend, we observed a progressive decrease of the deforestation rate after1990 from 205,000 ha/yr (1.63%/yr) over the period 1973-1990 to 44,300 ha/yr (0.43%/yr) over the period 2000-2005 (Tab.
-1).
-Then from 2005, the deforestation rate has progressively increased and has more than doubled over the period 2010-2014 (98,700 ha/yr, 1.08%/yr) compared to 2000-2005 (Tab.
-1).
-The deforestation trend characterized by a progressive decrease of the deforestation rate over the period 1990-2005 and a progressive increase of the deforestation after 2005 is valid for all four ecoregions (Tab.
-3), with the exception of the spiny forest domain for which the deforestation rate during the period 2010-2013 was lower than during 2005-2010 (Tab.
-3).
+### Forest cover change and deforestation rates
 
-### 3.2 Comparison with previous forest-cover change studies in Madagascar
+Natural forests in Madagascar covered 16.0 Mha in 1953, about 27% of
+the national territory of 587,041 km2. In 2014, the forest cover
+dropped to 8.9 Mha, corresponding to about 15% of the national
+territory (Fig.~\ref{fig:fcc} and
+Tab. \@ref(tab:forest_cover)). Madagascar has lost 44% and 37% of its
+natural forests between 1953 and 2014, and between 1973 and 2014
+respectively (Fig.~\ref{fig:fcc} and Tab. \@ref(tab:forest_cover)). In
+2014 the remaining 8.9 Mha of natural forest were distributed as
+follow: 4.4 Mha of moist forest (50% of total forest cover), 2.6 Mha
+of dry forest (29%), 1.7 Mha of spiny forest (19%) and 0.18 Mha
+(2%) of mangrove forest (Fig.~\ref{fig:ecoregion} and
+Tab. \@ref(tab:comp_forest)). Regarding the deforestation trend, we
+observed a progressive decrease of the deforestation rate after 1990
+from 205,000 ha/yr (1.6%/yr) over the period 1973-1990 to 42,000
+ha/yr (0.4%/yr) over the period 2000-2005
+(Tab. \@ref(tab:forest_cover)). Then from 2005, the deforestation rate
+has progressively increased and has more than doubled over the period
+2010-2014 (99,000 ha/yr, 1.1%/yr) compared to 2000-2005
+(Tab. \@ref(tab:forest_cover)). The deforestation trend, characterized
+by a progressive decrease of the deforestation rate over the period
+1990-2005 and a progressive increase of the deforestation after 2005,
+is valid for all four forest types except the spiny forest
+(Tab. \@ref(tab:comp_defor)). For the spiny forest, the deforestation
+rate during the period 2010-2013 was lower than on the period
+2005-2010 (Tab. \@ref(tab:comp_defor)).
 
-Forest-cover maps provided by previous studies over Madagascar were not exhaustive (unclassified areas) due to the presence of clouds in satellite imagery used to produce such maps.
-In @Harper2007, the maps of years 1990 and 2000 include 0.5 and 1.12 Mha of unknown cover type respectively.
-Proportions of unclassified areas are not reported in the two other existing studies by @MEFT2009 and @ONE2015.
-With our approach, we produced wall to wall forest-cover change maps from 1990 to 2014 for the full territory of Madagascar (Tab.
-1).
-This allowed us to produce more robust estimates of forest-cover and deforestation rates over this period.
-Our forest-cover estimates over the period 1953-2013 (considering forest cover estimates at national level and by ecoregions for the 6 available dates) were well correlated  (0.99) to estimates from the three previous studies (Tab.
-2) with a RMSE of 300,000 ha (6% of the mean forest cover of 4.8 Mha when considering all dates and forest types together).
-These small differences can be partly attributed to differences in ecoregion boundaries.
-Despite significant differences in deforestation estimates (Tab.
-3), a similar deforestation trend was observed across studies with a decrease of deforestation rates over the period 1990-2005, followed by a progressive increase of the deforestation after 2005.
+### Comparison with previous forest cover change studies in Madagascar
 
-### 3.3 Evolution of forest fragmentation with time
+Forest cover maps provided by previous studies over Madagascar were
+not exhaustive (unclassified areas) due to the presence of clouds on
+satellite images used to produce such maps. In @Harper2007, the
+maps of years 1990 and 2000 include 0.5 and 1.12 Mha of unknown cover
+type respectively. Proportions of unclassified areas are not reported
+in the two other existing studies at the national level by
+@MEFT2009 and @ONE2015. With our approach, we produced
+wall-to-wall forest cover change maps from 1990 to 2014 for the full
+territory of Madagascar (Tab. \@ref(tab:forest_cover)). This allowed us
+to produce more robust estimates of forest cover and deforestation
+rates over this period. Our forest cover estimates over the period
+1953-2013 (considering forest cover estimates at national level and by
+ecoregions for all the available dates) were well correlated
+(Pearson's correlation coefficient = 0.99) to estimates from the
+three previous studies (Tab. \@ref(tab:comp_forest)) with a RMSE of
+300,000 ha (6\% of the mean forest cover of 4.8 Mha when considering
+all dates and forest types together). These small differences can be
+partly attributed to differences in ecoregion boundaries. Despite
+significant differences in deforestation estimates
+(Tab. \@ref(tab:comp_defor)), a similar deforestation trend was
+observed across studies with a decrease of deforestation rates over
+the period 1990-2005, followed by a progressive increase of the
+deforestation after 2005.
 
-In parallel to the dynamics of deforestation, forest fragmentation has progressively increased since 1953 in Madagascar.
-We observed a continuous decrease of the mean distance to forest edge from 1953 to 2014 in Madagascar.
-The mean distance to forest edge has decreased to c.
-300 m in 2014 while it was previously c.
-1.5 km in 1973 (Fig.
-3).
-Moreover, a large proportion (73%) of the forest was located at a distance greater than 100 m in 1973, while almost half of the forest (46%) was at a distance lower than 100 m from forest edge in 2014 (Fig.
-3).
-The percentage of forest that can be considered intact in Madagascar has continuously decreased since 1953.
-The percentage of forest belonging to the "interior" category (most intact forests) has fallen from 68% in 1973 to 50% in 2014.
-In 2014, more than 16% of the forest belonged to the "patch" and "transitional" categories (isolated small forest patches) compared to 9.5% in 1973 (Tab.
-4).
+### Evolution of forest fragmentation with time
 
-## 4. Discussion
+In parallel to the dynamics of deforestation, forest fragmentation has
+progressively increased since 1953 in Madagascar. We observed a
+continuous decrease of the mean distance to forest edge from 1953 to
+2014 in Madagascar. The mean distance to forest edge has decreased to
+_c._ 300 m in 2014 while it was previously _c._ 1.5 km in
+1973 (Fig.~\ref{fig:dist_edge}). Moreover, a large proportion (73%)
+of the forest was located at a distance greater than 100 m in 1973,
+while almost half of the forest (46%) is now at a distance lower than
+100~m from forest edge in 2014 (Fig.~\ref{fig:dist_edge}). The
+percentage of lowly fragmented forest in Madagascar has continuously
+decreased since 1953. The percentage of forest belonging to the lowly
+fragmented category has fallen from 57% in 1973 to 44% in 2014. In
+2014, 22% of the forest belonged to the two highest
+fragmented forest classes (less than 40% of forest cover in the
+neighborhood) while 15% of the forest belonged to these two
+classes in 1973 (Tab. \@ref(tab:frag)).
 
-### 4.1 Benefits of the combined use of recent global annual tree cover loss data with historical national forest-cover maps
+## Discussion
 
-In this study, we combined recent (2001-2014) global annual tree cover loss data [@Hansen2013] with historical (1953-2000) national forest-cover maps [@Harper2007] to look at six decades (1953-2014) of deforestation and forest fragmentation in Madagascar.
-We produced annual forest-cover maps at 30m-resolution covering Madagascar for the period 2000 to 2014.
-Our study extends the forest-cover monitoring on a six decades period (from 1953 to 2014) while harmonizing the data from the previous studies.
-We propose a generic approach to solve the problem of forest definition which is needed to transform the 2000 global tree cover dataset from Hansen et al.
-(2013) into a forest/non-forest map [@Tropek2014].
-We propose to use a historical national forest-cover map, based on a national forest definition, as a forest cover mask.
-This approach could be easily extended to other regions or countries for which an accurate forest-cover map is available at any date within the period 2000-2014, but preferably at the beginning of the period to profit from the full record and derive long-term estimates of deforestation.
-Moreover, this approach can be repeated in the future if and when the global tree cover product is updated.
-We have made the R/GRASS code used for this study freely available in a GitHub repository (see Data availability statement) to facilitate application to other study areas.
+### Advantages of combining recent global annual tree cover loss data with historical national forest cover maps
 
-The accuracy of the derived forest-cover change maps depends directly on the accuracies of the historical forest-cover maps and the tree cover loss dataset.
-The reported global accuracy of the tree cover loss dataset is 99.6% (see Tab.
-S5 in @Hansen2013).
-@Verhegghen2016 have compared deforestation estimates derived from the global tree cover loss dataset [@Hansen2013] with results derived from semi-automated supervised classification of Landsat satellite images [@Achard2014] for six countries in Central Africa and they found a good agreement between these two sets of estimates.
-Consistent with @Harper2007, we did not consider potential forest regrowth in Madagascar (although @Hansen2013 provided a tree cover gains layer for the period 2001-2013) for the following reasons: (i) the tree gain layer of @Hansen2013 includes and catches more easily tree plantations than natural forest regrowth [@Tropek2014]; (ii) there is little evidence of natural forest regeneration in Madagascar [@Harper2007, @Grouzis2001] - this can be explained by several ecological processes following burning practice such as soil erosion [@Grinand2017] and reduced seed bank due to fire and soil loss [@Grouzis2001], (iii) in areas where forest regeneration is ecologically possible, young forest regrowth are more easily re-burnt for agriculture and pasture; and (iv) young secondary forests provide more limited ecosystem services compared to old-growth natural forests in terms of biodiversity and carbon storage.
+In this study, we combined recent (2001-2014) global annual tree cover
+loss data [@Hansen2013] with historical (1953-2000) national forest
+cover maps [@Harper2007] to look at six decades (1953-2014) of
+deforestation and forest fragmentation in Madagascar. We produced annual
+forest cover maps at 30 m resolution covering Madagascar for the period
+2000 to 2014. Our study extends the forest cover monitoring on a six
+decades period (from 1953 to 2014) while harmonizing the data from
+previous studies [@Harper2007; @MEFT2009; @ONE2015]. We propose a
+generic approach to solve the problem of forest definition which is
+needed to transform the 2000 global tree cover dataset from @Hansen2013
+into a forest/non-forest map [@Tropek2014]. We propose to use a
+historical national forest cover map, based on a national forest
+definition, as a forest cover mask. This approach could be easily
+extended to other regions or countries for which an accurate forest
+cover map is available at any date within the period 2000-2014, but
+preferably at the beginning of the period to profit from the full record
+and derive long-term estimates of deforestation. Moreover, this approach
+can be repeated in the future if and when the global tree cover product
+is updated. We have made the `R/GRASS` code used for this
+study freely available in a GitHub repository (see Data availability
+statement) to facilitate application to other study areas or repeat the
+analysis in the future for Madagascar.
 
-### 4.2 Dynamics of forest-cover in Madagascar from 1953 to 2014
+The accuracy of the derived forest cover change maps depends directly on
+the accuracies of the historical forest cover maps and the tree cover
+loss dataset. Using visual-interpretation of aerial images in 342 areas
+distributed among all forest types, @Harper2007 estimated an overall
+89.5% accuracy in identifying forest/non-forest classes for the year
+2000. The accuracy assessment of the tree cover loss dataset for the
+tropical biome reported 13% of false positives and 16.9% of false
+negatives (see Tab. S5 in @Hansen2013). These numbers rise at 20.7% and
+20.6% respectively for the subtropical biome. In the subtropical biome,
+the lower density tree cover canopy makes it difficult to detect change
+from tree cover to bare ground. For six countries in Central Africa,
+with a majority of moist dense forest, @Verhegghen2016 have compared
+deforestation estimates derived from the global tree cover loss dataset
+[@Hansen2013] with results derived from semi-automated supervised
+classification of Landsat satellite images [@Achard2014] and they found
+a good agreement between the two sets of estimates. Therefore, our
+forest cover change maps after 2000 might be more accurate for the dense
+moist forest than for the dry and spiny forest. In another study
+assessing the accuracy of the tree cover loss product accross the
+tropics [@Tyukavina2015], authors reported 4% of false positives and 48%
+of false negatives in Sub-Saharian Africa. They showed that 85% of
+missing loss occured on the edges of other loss patches. This means that
+tree cover loss might be underestimated in Sub-Saharian Africa, probably
+due to the prevalence of small-scale disturbance which is hard to map at
+30 m, but that areas of large-scale deforestation are well identified
+and spatial variability of the deforestation is well represented. A
+proper accuracy assessment of our forest cover change maps should be
+performed to better estimate the uncertainty surrounding our forest
+cover change estimates in Madagascar from year 2000
+[@Olofsson2013; @Olofsson2014]. Despite this limitation, we have shown
+that the deforestation trend we observed for Madagascar, with a doubling
+deforestation on the period 2010-2014 compared to 2000-2005, was
+consistent with the other studies at the national scale
+[@ONE2015; @MEFT2009].
 
-We estimated that natural forests in Madagascar cover 8.9 Mha in 2014 (corresponding to 15% of the country) and that Madagascar has lost 44% of its natural forest since 1953 (37% since 1973).
-There is ongoing scientific debate about the extent of the "original" forest cover in Madagascar, and the extent to which humans have altered the natural forest landscapes since their large-scale settlement around 800 CE [@Burns2016; @Cox2012].
-Early French naturalists stated that the full island was originally covered by forest [@Humbert1927; @Perrier1921], leading to the common statement that 90% of the natural forests have disappeared since the arrival of humans on the island [@Kull2000].
-More recent studies counter-balanced that point of view saying that extensive areas of grassland existed in Madagascar long before human arrival and were determined by climate, natural grazing and other natural factors [@Virah-Sawmy2009, @Vorontsova2017].
-Other authors have questioned the entire narrative of extensive alteration of the landscape by early human activity which, through legislation, has severe consequences on local people [@Klein2002; @Kull2000].
-Whatever the original proportion of natural forests and grasslands in Madagascar, our results demonstrate that human activities since the 1950s have profoundly impacted the natural tropical forests and that recent conservation and development programs in Madagascar have failed to stop deforestation in the recent years.
-Deforestation has strong consequences on biodiversity and carbon emissions in Madagascar.
-Around 90% of Madagascar's species are forest dependent [@Allnutt2008; @Goodman2005] and @Allnutt2008 estimated that deforestation between 1953 and 2000 led to an extinction of 9% of the species.
-The additional deforestation we observed over the period 2000-2014 (around 1Mha of natural forest) worsen this result.
-Regarding carbon emissions, using the 2010 aboveground forest carbon map by @Vieilledent2016, we estimated that deforestation on the period 2010-2014 has led to 40.2 Mt C of carbon emissions in the atmosphere (10 Mt C /yr) and that the remaining aboveground forest carbon stock in 2014 is 832.8 Mt C.
-Associated to deforestation, we showed that the remaining forests of Madagascar are highly fragmented with 46% of the forest being at less than 100m of the forest edge.
-Small forest fragments do not allow to maintain viable populations and 'edge effects' at forest/non-forest interfaces have impacts on both carbon emissions [@Brinck2017] and biodiversity loss [@Gibson2013; @Murcia1995].
+Consistent with @Harper2007, we did not consider potential forest
+regrowth in Madagascar (although @Hansen2013 provided a tree cover gains
+layer for the period 2001-2013) for several reasons. First, the tree
+gain layer of @Hansen2013 includes and catches more easily tree
+plantations than natural forest regrowth [@Tropek2014]. Second, there is
+little evidence of natural forest regeneration in Madagascar
+[@Grouzis2001; @Harper2007]. This can be explained by several ecological
+processes following burning practice such as soil erosion [@Grinand2017]
+and reduced seed bank due to fire and soil loss [@Grouzis2001].
+Moreover, in areas where forest regeneration is ecologically possible,
+young forest regrowth are more easily re-burnt for agriculture and
+pasture. Third, young secondary forests provide more limited ecosystem
+services compared to old-growth natural forests in terms of biodiversity
+and carbon storage.
 
-### 4.3 Deforestation trend and causes of the acceleration of the deforestation since 2005
+### Natural forest cover change in Madagascar from 1953 to 2014
 
-In our study, we have shown that the progressive decrease of the deforestation rate on the period 1990-2005 was followed by a continuous increase in the deforestation rate on the period 2005-2014.
-In particular, we showed that deforestation rate has more than doubled on the period 2010-2014 compared to 2000-2005.
-Our results are confirmed by previous studies [@Harper2007; @MEFT2009; @ONE2015] despite differences in the methodologies regarding (i) forest definition (associated to independent visual interpretations of observation polygons to train the classifier), (ii) classification algorithms, (iii) deforestation rate computation method, and (iv) correction for the presence of clouds.
-Our deforestation rate estimates from 1990 to 2014 have been computed from wall to wall maps at 30 m resolution and can be considered more accurate in comparison with estimates from these previous studies.
-Our forest-cover and deforestation rate estimates can be used as source of information for the next FAO Forest Resources Assessment project [@Keenan2015].
-Current rates of deforestation can also be used to build reference scenarios for deforestation in Madagascar and contribute to the implementation of deforestation mitigation activities in the framework of REDD+ [@Olander2008].
+We estimated that natural forest in Madagascar covers 8.9 Mha in 2014
+(corresponding to 15% of the country) and that Madagascar has lost 44%
+of its natural forest since 1953 (37% since 1973). There is ongoing
+scientific debate about the extent of the "original" forest cover in
+Madagascar, and the extent to which humans have altered the natural
+forest landscapes since their large-scale settlement around 800 CE
+[@Burns2016; @Cox2012]. Early French naturalists stated that the full
+island was originally covered by forest [@Humbert1927; @Perrier1921],
+leading to the common statement that 90% of the natural forests have
+disappeared since the arrival of humans on the island [@Kull2000]. More
+recent studies counter-balanced that point of view saying that extensive
+areas of grassland existed in Madagascar long before human arrival and
+were determined by climate, natural grazing and other natural factors
+[@Vorontsova2017; @Virah-Sawmy2009]. Other authors have questioned the
+entire narrative of extensive alteration of the landscape by early human
+activity which, through legislation, has severe consequences on local
+people [@Klein2002; @Kull2000]. Whatever the original proportion of
+natural forests and grasslands in Madagascar, our results demonstrate
+that human activities since the 1950s have profoundly impacted the
+natural tropical forests and that conservation and development programs
+in Madagascar have failed to stop deforestation in the recent years.
+Deforestation has strong consequences on biodiversity and carbon
+emissions in Madagascar. Around 90% of Madagascar's species are forest
+dependent [@Allnutt2008; @Goodman2005] and @Allnutt2008 estimated that
+deforestation between 1953 and 2000 led to an extinction of 9% of the
+species. The additional deforestation we observed over the period
+2000-2014 (around 1Mha of natural forest) worsen this result. Regarding
+carbon emissions, using the 2010 aboveground forest carbon map by
+@Vieilledent2016, we estimated that deforestation on the period
+2010-2014 has led to 40.2 Mt C of carbon emissions in the atmosphere (10
+Mt C /yr) and that the remaining aboveground forest carbon stock in 2014
+is 832.8 Mt C. Associated to deforestation, we showed that the remaining
+forests of Madagascar are highly fragmented with 46% of the forest being
+at less than 100 m of the forest edge. Small forest fragments do not
+allow to maintain viable populations and "edge effects" at
+forest/non-forest interfaces have impacts on both carbon emissions
+[@Brinck2017] and biodiversity loss [@Gibson2013; @Murcia1995].
 
-The increase of deforestation rates after 2005 can be explained by population growth and political instability in the country.
-Nearly 90% of Madagascar's population relies on biomass for their daily energy needs [@Minten2013] and the link between population size and deforestation has previously been demonstrated in Madagascar [@Vieilledent2013; @Gorenflo2011].
-With a mean demographic growth rate of about 2.8%/yr and a population which has increased from 16 to 24 million people on the period 2000-2015 [@UN2015], the increasing demand in wood-fuel and space for agriculture is likely to explain the increase in deforestation rates.
-The political crisis of 2009 [@Ploch2012], followed by several years of political instability and weak governance could also explain the increase in the deforestation rate observed on the period 2005-2014 (Smith et al., 2003).
-These results show that despite the conservation policy in Madagascar [@Freudenberger2010], deforestation has dramatically increased at the national level since 2005.
-More efforts should be done to identify the real causes and mechanisms explaining this rampant deforestation in order to implement new conservation strategies that will be effective at conserving the few remaining natural tropical forests in Madagascar.
+### Deforestation trend and impacts on conservation and development policies
+
+In our study, we have shown that the progressive decrease of the
+deforestation rate on the period 1990-2005 was followed by a continuous
+increase in the deforestation rate on the period 2005-2014. In
+particular, we showed that deforestation rate has more than doubled on
+the period 2010-2014 compared to 2000-2005. Our results are confirmed by
+previous studies [@Harper2007; @MEFT2009; @ONE2015] despite differences
+in the methodologies regarding (i) forest definition (associated to
+independent visual interpretations of observation polygons to train the
+classifier), (ii) classification algorithms, (iii) deforestation rate
+computation method, and (iv) correction for the presence of clouds. Our
+deforestation rate estimates from 1990 to 2014 have been computed from
+wall-to-wall maps at 30 m resolution and can be considered more accurate
+in comparison with estimates from these previous studies. Our forest
+cover and deforestation rate estimates can be used as source of
+information for the next FAO Forest Resources Assessment [@Keenan2015].
+Current rates of deforestation can also be used to build reference
+scenarios for deforestation in Madagascar and contribute to the
+implementation of deforestation mitigation activities in the framework
+of REDD+ [@Olander2008].
+
+The increase of deforestation rates after 2005 can be explained by
+population growth and political instability in the country. Nearly 90%
+of Madagascar's population relies on biomass for their daily energy
+needs [@Minten2013] and the link between population size and
+deforestation has previously been demonstrated in Madagascar
+[@Vieilledent2013; @Gorenflo2011]. With a mean demographic growth rate
+of about 2.8%/yr and a population which has increased from 16 to 24
+million people on the period 2000-2015 [@UN2015], the increasing demand
+in wood-fuel and space for agriculture is likely to explain the increase
+in deforestation rates. The political crisis of 2009 [@Ploch2012],
+followed by several years of political instability and weak governance
+could also explain the increase in the deforestation rate observed on
+the period 2005-2014 [@Smith2003]. These results show that despite the
+conservation policy in Madagascar [@Freudenberger2010], deforestation
+has dramatically increased at the national level since 2005. Results of
+this study, including recent spatially explicit forest cover change maps
+and forest cover estimates, should help implement new conservation
+strategies to save Madagascar natural tropical forests and their unique
+biodiversity.
 
@@
 
-All the authors conceived the study; GV made the analysis and drafted the manuscript. All the authors revised the manuscript.
-
-All authors conceived the ideas and designed methodology; GV analysed the data and wrote the R/GRASS script; GV drafted the manuscript.
-All authors contributed critically to the drafts and gave final approval for publication.
+All authors conceived the ideas and designed methodology; GV analysed
+the data and wrote the `R/GRASS` script; GV drafted the
+manuscript. All authors contributed critically to the drafts and gave
+final approval for publication.
 
@@
 
-This study is part of the Cirad's BioSceneMada project (https://bioscenemada.cirad.fr) and the Joint Research Center's ReCaREDD project (http://forobs.jrc.ec.europa.eu/recaredd).
-The BioSceneMada project is funded by FRB (Fondation pour la Recherche sur la Biodiversité) and the FFEM (Fond Français pour l'Environnement Mondial) under the project agreement AAP-SCEN-2013 I.
-The ReCaREDD project is funded by the European Commission.
+The authors thank Jean-François Bastin for useful comments on a previous
+version of the manuscript and Peter Vogt for useful advices on which
+metric to use to estimate forest fragmentation. This study is part of
+the Cirad's BioSceneMada project (<https://bioscenemada.cirad.fr>) and
+the Joint Research Center's ReCaREDD project
+(<http://forobs.jrc.ec.europa.eu/recaredd>). The BioSceneMada project is
+funded by FRB (Fondation pour la Recherche sur la Biodiversité) and the
+FFEM (Fond Français pour l'Environnement Mondial) under the project
+agreement AAP-SCEN-2013 I. The ReCaREDD project is funded by the
+European Commission. The authors declare that there are no conflicts of
+interest related to this article.
 
@@
 
-All the data and codes used for this study are made publicly available in the `deforestmap` GitHub repository (https://github.com/ghislainv/deforestmap.git).
-The results and the manuscript are fully reproducible running the R script `deforestmap.R` located inside the `deforestmap` repository.
+All the data and the script used for this study have been made
+permanently and publicly available on the Zenodo research data
+repository so that the results are entirely reproducible:
+
+-   Input data: <https://doi.org/10.5281/zenodo.1118955>
+-   Script: <https://doi.org/10.5281/zenodo.1118484>
+-   Output data: <https://doi.org/10.5281/zenodo.1145785>
 
@@
 
-```{r defor, results="asis"}
-defor.df <- read.table(file="outputs/defor.txt", sep="\t", header=TRUE)
+```{r forest_cover, results="asis"}
+defor.df <- read.table(file="tabs/defor.txt", sep="\t", header=TRUE)
 perc.1 <- round(100*(defor.df$area[defor.df$Year==1973]-defor.df$area[defor.df$Year==2014])/defor.df$area[defor.df$Year==1973])
@@
 
-Table 1: **Evolution of the forest cover and deforestation rates from 1953 to 2014 in Madagascar**.
+Table 1: **Evolution of natural forest cover and deforestation rates from 1953 to 2014 in Madagascar.** Areas are provided in thousands of hectares (Kha).
 Forest map for the year 1973 has 3.3 Mha of unclassified areas due to the presence of clouds on satellite images.
-As a consequence, deforestation rates for the periods 1953-1973 and 1973-1990 are indicative.
-The two last columns indicate the deforestation rates on the previous time-period (e.g.
-1953-1973 for year 1973, 1973-1990 for year 1990, etc.).
+As a consequence, partial deforestation rates for the periods 1953-1973 and 1973-1990 are computed based on the available forest extent.
+The last two columns indicate the annual deforested areas and annual deforestation rates on the previous time-period (e.g. 1953-1973 for year 1973, 1973-1990 for year 1990, etc.).
 
-```{r fcc_comp, results="asis"}
-fcc.comp <- read.table(file="outputs/fcc_comp.txt", sep=" ", header=TRUE)
+```{r comp_forest, results="asis"}
+fcc.comp <- read.table(file="tabs/fcc_comp.txt", sep=" ", header=TRUE)
 tab.2 <- fcc.comp
@@
 
-Table 2: **Comparing our estimates of forest-cover (in ha) for Madagascar with previous studies on the period 1953-2014**.
-We compared our forest-cover with the estimates from three previous studies [@Harper2007;@MEFT2009;@ONE2015].
-We computed a The increase in mangrove and spiny forest covers from _c._ 1953 to _c._ 1973 in @Harper2007 and our study is most probably due to differences in forest definition and mapping methods between the 1953 aerial-photography derived map and the 1973 Landsat image derived map.
+Table 2: **Comparing Madagascar forest cover estimates with previous studies on the period 1953-2014.** We compared our estimates of forest cover with the estimates from three previous studies [@Harper2007; @MEFT2009; @ONE2015].
+Areas are provided in thousands of hectares (Kha).
+We obtained a Pearson's correlation coefficient of 0.99 between our forest cover estimates and forest cover estimates from previous studies.
+The increase in mangrove and spiny forest covers from _c._ 1953 to _c._ 1973 in @Harper2007 and our study is most probably due to differences in forest definition and mapping methods between the 1953 aerial-photography derived map and the 1973 Landsat image derived map.
 
-```{r defor_comp, results="asis"}
-tab.3 <- read.table(file="outputs/defor_comp.txt", sep=" ", header=TRUE)
+```{r comp_defor, results="asis"}
+tab.3 <- read.table(file="tabs/defor_comp.txt", sep=" ", header=TRUE)
 tab.3$ForestType <- c("Total",rep("",3),"Moist",rep(" ",3),"Dry",rep(" ",3),"Spiny",rep(" ",3),"Mangroves",rep(" ",3))
@@
 
-Table 3: **Comparing our estimates of annual deforestation rates for Madagascar with previous studies on the period 1953-2014**.
-Annual deforestation areas (in ha/yr) and annual deforestation rates (second number in parenthesis, in %/yr) are provided.
-For deforestation rates in %/yr, exact same numbers as in scientific articles and reports from previous studies have been reported.
-The way annual deforestation rates in %/yr have been computed in these previous studies can slightly differ from one study to another but estimates always correct for the potential presences of clouds on forest maps.
-Annual deforested areas in ha/yr have been recomputed from forest-cover estimates in Tab.
-2 (except for @Harper2007 for the periods 1973-1990 and 1990-2000 for which annual deforested ares in ha/yr were derived from numbers reported in the original publication, see methods) and do not correct for the potential presence of clouds.
+Table 3: **Comparing Madagascar annual deforestation rates with previous studies on the period 1953-2014.** Annual deforested areas (in thousands of hectares per year, Kha/yr) and annual deforestation rates (second number in parenthesis, in %/yr) are provided.
+For deforestation rates in %/yr, exact same numbers as in scientific articles and reports from previous studies [@Harper2007; @MEFT2009; @ONE2015] have been reported.
+The way annual deforestation rates in %/yr have been computed in these previous studies can slightly differ from one study to another, but estimates always correct for the potential presences of clouds on satellite images and unclassified areas on forest maps.
+Annual deforested areas in Kha/yr have been recomputed from forest cover estimates in Tab. \ref{tab:comp_forest} (except for @Harper2007 for the periods 1973-1990 and 1990-2000 for which annual deforested areas in Kha/yr were derived from numbers reported in the original publication, see methods) and do not correct for the potential presence of clouds.
 
@@
 
-Table 4: **Evolution of the forest fragmentation from 1953 to 2014 in Madagascar**.
-Six categories of fragmentation were identified from the amount of forest and its occurrence as adjacent forest pixels: "interior", "perforated", "edge", "transitional", "patch", and "undetermined" [@Riitters2000].
-We used a moving window of 7x7 pixels (4.4 ha).
-Using this window size, forest edge had a width of about 90m.
-The "interior" category can be interpreted as the most intact forest.
-The "patch" and "transitional" categories correspond to isolated small forest patches.
+Table 4: **Evolution of the forest fragmentation from 1953 to 2014 in Madagascar.** Five forest fragmentation classes, based on the percentage of forest in the neighborhood, are defined: 0-20% (highly fragmented), 21-40%, 41-60%, 61-80% and 81-100% (lowly fragmented).
+The percentage of forest falling in each forest fragmentation class is reported for each year.
+Forest areas are provided in thousands of hectares (Kha).
 
@@
 
-![Figure 1: **Ecoregions and forest types in Madagascar.** Madagascar can be divided into four climatic ecoregions with four forest types: the moist forest in the East (green), the dry forest in the West (orange), the spiny forest in the South (red), and the mangroves on the West coast (blue).
+![**Ecoregions and forest types in Madagascar.** Madagascar can be divided into four climatic ecoregions with four forest types: the moist forest in the East (green), the dry forest in the West (orange), the spiny forest in the South (red), and the mangroves on the West coast (blue).
 Ecoregions were defined following climatic [@Cornet1974] and vegetation [@IEFN1996] criteria.
-The dark grey areas represent the remaining natural forest cover for the year 2014.](figs/ecoregion.png)
+The dark grey areas represent the
+remaining natural forest cover for the year 2014.](figs/ecoregion.png)
 
-![Figure 2: **Forest-cover change on six decades from 1953 to 2014 in Madagascar.** Forest cover changes from _c._ 1973 to 2014 are shown in the main figure, and forest cover in _c._ 1953 is shown in the bottom-right inset.
-Two zooms in the western dry (left part) and eastern moist (right part) ecoregions present more detailed views of (from top to bottom): forest-cover in 1950s, forest-cover change from _c._ 1973 to 2014, forest fragmentation in 2014 and distance to forest edge in 2014.
+![**Forest cover change on six decades from 1953 to 2014 in Madagascar.** Forest cover changes from _c._ 1973 to 2014 are shown in the main figure, and forest cover in _c._ 1953 is shown in the bottom-right inset.
+Two zooms in the western dry (left part) and eastern moist (right part) ecoregions present more detailed views of (from top to bottom): forest cover in 1950s, forest cover change from _c._ 1973 to 2014, forest fragmentation in 2014 and distance to forest edge in 2014.
 Data on water bodies (blue) and water seasonality (light blue for seasonal water to dark blue for permanent water) has been extracted from @Pekel2016.](figs/fig_fcc.png)
 
-![Figure 3: **Evolution of the distance to forest edge from 1953 to 2014 in Madagascar.** Black dots represent the mean distance to forest edge for each year.
+![**Evolution of the distance to forest edge from 1953 to 2014 in Madagascar.** Black dots represent the mean distance to forest edge for each year.
 Vertical dashed segments represent the 90% quantiles (5% and 95%) of the distance to forest edge.
-Horizontal dashed grey line indicates a distance to forest edge of 100m.
-Percentages indicate the percentage of forest at a distance to forest edge lower than 100m for each year.](figs/dist.png)
+Horizontal dashed grey line indicates a distance to forest edge of 100 m.
+Numbers at the bottom of each vertical segments are the percentage of forest at a distance to forest edge lower than 100 m for each year.](figs/dist.png)
 

### 2018-01-15 deforestmap.R

@@ -964,6 +964,7 @@ system("r.mask -r")
 acd_2010 <- 9700502114*30*30/10000 # 873045190
 acd_2014 <- 9253573526*30*30/10000 # 832821617
 carbon_emissions <- acd_2010-acd_2014
+SavedObjects <- c(SavedObjects,"carbon_emissions")
 
 ##========================
 ## Save objects
@@ -971,7 +972,6 @@ carbon_emissions <- acd_2010-acd_2014
 
 # load("deforestmap.rda")
 save(list=SavedObjects,file="deforestmap.rda")
-save(list=ls(),file="deforestmap.rda")
 
 ##========================
 ## Knit the document
@@ -997,7 +997,7 @@ render("coverletter.md",output_format=c("pdf_document"),output_dir="report") # p
 ##========================
 
 system("pandoc manuscript/manuscript.tex --bibliography=manuscript/bib/biblio.bib \\
-       --csl=manuscript/bib/journal-of-applied-ecology.csl -o manuscript/manuscript_essd.md")
+       --csl=manuscript/bib/journal-of-applied-ecology.csl -o manuscript/manuscript.md")
 
 ##===========================================================================
 ## End of script

### 2018-01-16 manuscript/manuscript.Rmd
Bookdown

@@
 ---
-title: "Combining global tree cover loss data with historical national forest-cover maps to look at six decades of deforestation and forest fragmentation in Madagascar"
-date: '`r format(Sys.time(), "%d %B, %Y")`'
+title: Combining global tree cover loss data with historical national forest-cover
+  maps to look at six decades of deforestation and forest fragmentation in Madagascar
+
 output:
-  bookdown::tufte_html2:
-    number_sections: no
-    toc: yes
-  bookdown::word_document2: null
   bookdown::pdf_document2:
-    #includes:
-    #  before_body: doc_prefix.tex
-    #  in_header: preamble.tex
     keep_tex: yes
+    latex_engine: pdflatex
     number_sections: yes
     toc: no
+
+link-citations: yes
 bibliography: bib/biblio.bib
-fontsize: 12pt
-link-citations: yes
 csl: bib/journal-of-applied-ecology.csl
@@
 
-```{r load_data}
-load("deforestmap.rda")
+```{r options, echo=FALSE, results="hide"}
+library(bookdown)
+library(knitr)
+library(kableExtra)
+opts_chunk$set(echo=FALSE, cache=FALSE,
+               #results="hide", 
+               warning=FALSE,
+               message=FALSE, highlight=TRUE,
+               fig.show="hide", size="small",
+               tidy=FALSE)
+options(knitr.kable.NA="-")
 ```
 
-#### Ghislain Vieilledent<sup>*,1,2</sup>, Clovis Grinand<sup>3</sup>, Fety A.
-Rakotomalala<sup>3</sup>, Rija Ranaivosoa<sup>4</sup>, Jean-Roger Rakotoarijaona<sup>4</sup>, Thomas F.
-Allnutt<sup>5,6</sup>, and Frédéric Achard<sup>1</sup>
-
-(*) /Email: ghislain.vieilledent@cirad.fr, /Phone: +39.329.457.2273
+**Ghislain Vieilledent$^{1,2,\star}$, Clovis Grinand$^{3}$, Fety A. Rakotomalala$^{3}$, Rija Ranaivosoa$^{4}$, Jean-Roger Rakotoarijaona$^{4}$, Thomas F. Allnutt$^{5,6}$, and Frédéric Achard$^{1}$**
 
@@
 
+(*) /Email: ghislain.vieilledent@cirad.fr, /Phone: +39.329.457.2273
+
 *Running headline:* Six decades of deforestation in Madagascar
 
-## Highlights
+# Highlights
 
@@
 
-## Summary
+# Summary
 
@@
 
-## Introduction
+# Introduction
 
@@
 
-## Materials and Methods
+# Materials and Methods
 
-### Creation of new forest cover maps of Madagascar from 1953 to 2014
+## Creation of new forest cover maps of Madagascar from 1953 to 2014
 
@@
 
-### Computing forest cover areas and deforestation rates
+## Computing forest cover areas and deforestation rates
 
@@
 \begin{equation}
+  (\#eq:theta)
   \theta = 100 \times [1-(1-(F_{t_2}-F_{t_1})/F_{t_1})^{(1/(t_2-t_1))}]
-  (\#eq:theta)
 \end{equation}
 
-In Eq. \ref{eq:theta}, $\theta$ is the annual deforestation rate (in
+In Eq. \@ref(eq:theta), $\theta$ is the annual deforestation rate (in
 %/yr), $F_{t_2}$ and $F_{t_1}$ are the forest cover free of clouds at
@@
 
-### Comparing our forest cover and deforestation rate estimates with previous studies
+## Comparing our forest cover and deforestation rate estimates with previous studies
 
@@
 estimates from the three existing studies at the national scale for
-Madagascar: (i) [@Harper2007], (ii) [@MEFT2009] and (iii) [@ONE2015].
+Madagascar: (i) @Harper2007, (ii) @MEFT2009 and (iii) @ONE2015.
 @Harper2007 provides forest cover and deforestation estimates for the
@@
 estimates cannot be corrected from the potential bias due to the
-presence of residual clouds. forest cover and deforestation rates were
+presence of residual clouds. Forest cover and deforestation rates were
 then compared between all studies for the whole of Madagascar and the
@@
 
-### Fragmentation
+## Fragmentation
 
@@
 
-## Results
+# Results
 
-### Forest cover change and deforestation rates
+## Forest cover change and deforestation rates
 
@@
 
-### Comparison with previous forest cover change studies in Madagascar
+## Comparison with previous forest cover change studies in Madagascar
 
@@
 
-### Evolution of forest fragmentation with time
+## Evolution of forest fragmentation with time
 
@@
 
-## Discussion
+# Discussion
 
-### Advantages of combining recent global annual tree cover loss data with historical national forest cover maps
+## Advantages of combining recent global annual tree cover loss data with historical national forest cover maps
 
@@
 
-### Natural forest cover change in Madagascar from 1953 to 2014
+## Natural forest cover change in Madagascar from 1953 to 2014
 
@@
 
-### Deforestation trend and impacts on conservation and development policies
+## Deforestation trend and impacts on conservation and development policies
 
@@
 
-## Author's contribution
+# Author's contribution
 
@@
 
-## Acknowledgements
+# Acknowledgements
 
@@
 
-## Data accessibility
+# Data accessibility
 
@@
 
-## Tables
+# Tables
 
-```{r forest_cover, results="asis"}
+```{r forest_cover}
 defor.df <- read.table(file="tabs/defor.txt", sep="\t", header=TRUE)
@@
 names(tab.1) <- c("Year", "Forest (Kha)", "Unmap (Kha)", "Annual defor. (Kha/yr)", "Rate (%/yr)")
-kable(tab.1, format.args = list(big.mark=","), digits=1)
+# Caption
+#cap <- "**Evolution of natural forest cover and deforestation rates from 1953 to 2014 in Madagascar.** Areas are provided in thousands of hectares (Kha).
+Forest map for the year 1973 has 3.3 Mha of unclassified areas due to the presence of clouds on satellite images.
+As a consequence, partial deforestation rates for the periods 1953-1973 and 1973-1990 are computed based on the available forest extent.
+The last two columns indicate the annual deforested areas and annual deforestation rates on the previous time-period (e.g. 1953-1973 for year 1973, 1973-1990 for year 1990, etc.)."
+# Kable
+knitr::kable(tab.1, format.args=list(big.mark=","), digits=1, booktabs=TRUE, caption="Test")
 ```
 
-Table 1: **Evolution of natural forest cover and deforestation rates from 1953 to 2014 in Madagascar.** Areas are provided in thousands of hectares (Kha).
-Forest map for the year 1973 has 3.3 Mha of unclassified areas due to the presence of clouds on satellite images.
-As a consequence, partial deforestation rates for the periods 1953-1973 and 1973-1990 are computed based on the available forest extent.
-The last two columns indicate the annual deforested areas and annual deforestation rates on the previous time-period (e.g. 1953-1973 for year 1973, 1973-1990 for year 1990, etc.).
-
-```{r comp_forest, results="asis"}
+```{r comp_forest}
 fcc.comp <- read.table(file="tabs/fcc_comp.txt", sep=" ", header=TRUE)
@@
 col.names <- c("Forest type","Source","1953","1973","1990","2000","2005","2010","2013","2014")
-kable(tab.2,col.names=col.names,align=c("l","l",rep("r",8)), format.args = list(big.mark=","))
+# Caption
+#cap <- "**Comparing Madagascar forest cover estimates with previous studies on the period 1953-2014.** We compared our estimates of forest cover with the estimates from three previous studies [@Harper2007; @MEFT2009; @ONE2015].
+Areas are provided in thousands of hectares (Kha).
+We obtained a Pearson's correlation coefficient of 0.99 between our forest cover estimates and forest cover estimates from previous studies.
+The increase in mangrove and spiny forest covers from _c._ 1953 to _c._ 1973 in @Harper2007 and our study is most probably due to differences in forest definition and mapping methods between the 1953 aerial-photography derived map and the 1973 Landsat image derived map."
+# Kable
+knitr::kable(tab.2, col.names=col.names,align=c("l","l",rep("r",8)), format.args=list(big.mark=","), booktabs=TRUE, caption="Test")
 ```
 
-Table 2: **Comparing Madagascar forest cover estimates with previous studies on the period 1953-2014.** We compared our estimates of forest cover with the estimates from three previous studies [@Harper2007; @MEFT2009; @ONE2015].
-Areas are provided in thousands of hectares (Kha).
-We obtained a Pearson's correlation coefficient of 0.99 between our forest cover estimates and forest cover estimates from previous studies.
-The increase in mangrove and spiny forest covers from _c._ 1953 to _c._ 1973 in @Harper2007 and our study is most probably due to differences in forest definition and mapping methods between the 1953 aerial-photography derived map and the 1973 Landsat image derived map.
-
-```{r comp_defor, results="asis"}
+```{r comp_defor}
 tab.3 <- read.table(file="tabs/defor_comp.txt", sep=" ", header=TRUE)
@@
 col.names <- c("Forest type","Source","1953-1973","1973-1990","1990-2000","2000-2005","2005-2010","2010-2013")
-kable(tab.3,col.names=col.names,align=c("l","l",rep("r",6)))
+# Caption
+# cap <- "**Comparing Madagascar annual deforestation rates with previous studies on the period 1953-2014.** Annual deforested areas (in thousands of hectares per year, Kha/yr) and annual deforestation rates (second number in parenthesis, in %/yr) are provided.
+For deforestation rates in %/yr, exact same numbers as in scientific articles and reports from previous studies [@Harper2007; @MEFT2009; @ONE2015] have been reported.
+The way annual deforestation rates in %/yr have been computed in these previous studies can slightly differ from one study to another, but estimates always correct for the potential presences of clouds on satellite images and unclassified areas on forest maps.
+Annual deforested areas in Kha/yr have been recomputed from forest cover estimates in Tab. \\@ref(tab:comp_forest) (except for @Harper2007 for the periods 1973-1990 and 1990-2000 for which annual deforested areas in Kha/yr were derived from numbers reported in the original publication, see methods) and do not correct for the potential presence of clouds."
+# Kable
+knitr::kable(tab.3, col.names=col.names, align=c("l","l",rep("r",6)), booktabs=TRUE, caption="Test")
 ```
 
-Table 3: **Comparing Madagascar annual deforestation rates with previous studies on the period 1953-2014.** Annual deforested areas (in thousands of hectares per year, Kha/yr) and annual deforestation rates (second number in parenthesis, in %/yr) are provided.
-For deforestation rates in %/yr, exact same numbers as in scientific articles and reports from previous studies [@Harper2007; @MEFT2009; @ONE2015] have been reported.
-The way annual deforestation rates in %/yr have been computed in these previous studies can slightly differ from one study to another, but estimates always correct for the potential presences of clouds on satellite images and unclassified areas on forest maps.
-Annual deforested areas in Kha/yr have been recomputed from forest cover estimates in Tab. \ref{tab:comp_forest} (except for @Harper2007 for the periods 1973-1990 and 1990-2000 for which annual deforested areas in Kha/yr were derived from numbers reported in the original publication, see methods) and do not correct for the potential presence of clouds.
-
-```{r frag, results="asis"}
-frag.df <- read.table(file="outputs/fordens.txt", sep="\t", header=TRUE)
+```{r frag}
+# Import and format
+frag.df <- read.table(file="tabs/fordens.txt", sep="\t", header=TRUE)
 tab.4 <- frag.df
-p.1 <- round(frag.df$interior[frag.df$Year==1973])
-p.2 <- round(frag.df$interior[frag.df$Year==2014])
-p.3 <- round(frag.df$patch[frag.df$Year==2014]+frag.df$transitional[frag.df$Year==2014])
-tab.4 <- tab.4[,-c(8)]
 tab.4$Year <- as.character(tab.4$Year)
 tab.4$forest <- round(tab.4$forest/1000)
-names(tab.4) <- c("Year","Forest (ha)", "0-20", "21-40", "41-60", "61-80", "81-100")
-kable(tab.4, digits=0, format.args=list(big.mark=","))
+names(tab.4) <- c("Year","Forest (Kha)", "0-20", "21-40", "41-60", "61-80", "81-100")
+# Caption
+#cap <- "**Evolution of the forest fragmentation from 1953 to 2014 in Madagascar.** Five forest fragmentation classes, based on the percentage of forest in the neighborhood, are defined: 0-20% (highly fragmented), 21-40%, 41-60%, 61-80% and 81-100% (lowly fragmented).
+The percentage of forest falling in each forest fragmentation class is reported for each year.
+Forest areas are provided in thousands of hectares (Kha)."
+# Kable
+knitr::kable(tab.4, digits=0, format.args=list(big.mark=","), booktabs=TRUE, caption="Test")
 ```
 
-Table 4: **Evolution of the forest fragmentation from 1953 to 2014 in Madagascar.** Five forest fragmentation classes, based on the percentage of forest in the neighborhood, are defined: 0-20% (highly fragmented), 21-40%, 41-60%, 61-80% and 81-100% (lowly fragmented).
-The percentage of forest falling in each forest fragmentation class is reported for each year.
-Forest areas are provided in thousands of hectares (Kha).
-
-##  Figures
+#  Figures
 
@@
 
-## References
+# References

### 2018-01-16 manuscript/manuscript.tex
Bookdown

--- 2017-12-18 manuscript/manuscript.tex
@@
-%% Copernicus Publications Manuscript Preparation Template for LaTeX Submissions
-%% ---------------------------------
-%% This template should be used for copernicus.cls
-%% The class file and some style files are bundled in the Copernicus Latex Package, which can be downloaded from the different journal webpages.
-%% For further assistance please contact Copernicus Publications at: production@copernicus.org
-%% https://publications.copernicus.org/for_authors/manuscript_preparation.html
-
-%% Please use the following documentclass and journal abbreviations for discussion papers and final revised papers.
-
-%% 2-column papers and discussion papers
-\documentclass[essd, classical]{copernicus}
-
-%% Journal abbreviations (please use the same for discussion papers and final revised papers)
-
-% Archives Animal Breeding (aab)
-% Atmospheric Chemistry and Physics (acp)
-% Advances in Geosciences (adgeo)
-% Advances in Statistical Climatology, Meteorology and Oceanography (ascmo)
-% Annales Geophysicae (angeo)
-% ASTRA Proceedings (ap)
-% Atmospheric Measurement Techniques (amt)
-% Advances in Radio Science (ars)
-% Advances in Science and Research (asr)
-% Biogeosciences (bg)
-% Climate of the Past (cp)
-% Drinking Water Engineering and Science (dwes)
-% Earth System Dynamics (esd)
-% Earth Surface Dynamics (esurf)
-% Earth System Science Data (essd)
-% Fossil Record (fr)
-% Geographica Helvetica (gh)
-% Geoscientific Instrumentation, Methods and Data Systems (gi)
-% Geoscientific Model Development (gmd)
-% Hydrology and Earth System Sciences (hess)
-% History of Geo- and Space Sciences (hgss)
-% Journal of Micropalaeontology (jm)
-% Journal of Sensors and Sensor Systems (jsss)
-% Mechanical Sciences (ms)
-% Natural Hazards and Earth System Sciences (nhess)
-% Nonlinear Processes in Geophysics (npg)
-% Ocean Science (os)
-% Proceedings of the International Association of Hydrological Sciences (piahs)
-% Primate Biology (pb)
-% Scientific Drilling (sd)
-% SOIL (soil)
-% Solid Earth (se)
-% The Cryosphere (tc)
-% Web Ecology (we)
-% Wind Energy Science (wes)
-
-%% \usepackage commands included in the copernicus.cls:
-%\usepackage[german, english]{babel}
-\usepackage{tabularx}
-%\usepackage{cancel}
-%\usepackage{multirow}
-\usepackage{supertabular}
-%\usepackage{algorithmic}
-%\usepackage{algorithm}
-%\usepackage{amsthm}
-%\usepackage{float}
-%\usepackage{subfig}
-%\usepackage{rotating}
-
-%% My packages
-\usepackage{amssymb} %symbole de maths
-\usepackage{amsmath} %idem
-\usepackage{longtable}
-\setlength{\LTleft}{-5cm plus 1 fill}
-\setlength{\LTright}{-5cm plus 1 fill}
-%% New commands
-\newcommand{\logit}{\text{logit}}
-\newcommand{\bs}[1]{\boldsymbol{#1}}
-\newcommand{\R}{\textnormal{\sffamily{\bfseries{R}}}}
-\newcommand{\pkg}[1]{{\fontseries{b}\selectfont #1}}
-
-\begin{document}
-
+\documentclass[]{article}
+\usepackage{lmodern}
+\usepackage{amssymb,amsmath}
+\usepackage{ifxetex,ifluatex}
+\usepackage{fixltx2e} % provides \textsubscript
+\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
+  \usepackage[T1]{fontenc}
+  \usepackage[utf8]{inputenc}
+\else % if luatex or xelatex
+  \ifxetex
+    \usepackage{mathspec}
+  \else
+    \usepackage{fontspec}
+  \fi
+  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
+\fi
+% use upquote if available, for straight quotes in verbatim environments
+\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
+% use microtype if available
+\IfFileExists{microtype.sty}{%
+\usepackage{microtype}
+\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
+}{}
+\usepackage[margin=1in]{geometry}
+\usepackage{hyperref}
+\hypersetup{unicode=true,
+            pdftitle={Combining global tree cover loss data with historical national forest-cover maps to look at six decades of deforestation and forest fragmentation in Madagascar},
+            pdfborder={0 0 0},
+            breaklinks=true}
+\urlstyle{same}  % don't use monospace font for urls
+\usepackage{longtable,booktabs}
+\usepackage{graphicx,grffile}
+\makeatletter
+\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
+\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
+\makeatother
+% Scale images if necessary, so that they will not overflow the page
+% margins by default, and it is still possible to overwrite the defaults
+% using explicit options in \includegraphics[width, height, ...]{}
+\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
+\IfFileExists{parskip.sty}{%
+\usepackage{parskip}
+}{% else
+\setlength{\parindent}{0pt}
+\setlength{\parskip}{6pt plus 2pt minus 1pt}
+}
+\setlength{\emergencystretch}{3em}  % prevent overfull lines
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
+\setcounter{secnumdepth}{5}
+% Redefines (sub)paragraphs to behave more like sections
+\ifx\paragraph\undefined\else
+\let\oldparagraph\paragraph
+\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
+\fi
+\ifx\subparagraph\undefined\else
+\let\oldsubparagraph\subparagraph
+\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
+\fi
+
+%%% Use protect on footnotes to avoid problems with footnotes in titles
+\let\rmarkdownfootnote\footnote%
+\def\footnote{\protect\rmarkdownfootnote}
+
+%%% Change title format to be more compact
+\usepackage{titling}
+
+% Create subtitle command for use in maketitle
+\newcommand{\subtitle}[1]{
+  \posttitle{
+    \begin{center}\large#1\end{center}
+    }
+}
+
+\setlength{\droptitle}{-2em}
   \title{Combining global tree cover loss data with historical national
@@
 fragmentation in Madagascar}
-
-% \Author[affil]{given_name}{surname}
-
-\Author[1, 2, 3]{Ghislain}{Vieilledent}
-\Author[4]{Clovis}{Grinand}
-\Author[4]{Fety A.}{Rakotomalala}
-\Author[5]{Rija}{Ranaivosoa}
-\Author[5]{Jean-Roger}{Rakotoarijaona}
-\Author[6, 7]{Thomas F.}{Allnutt}
-\Author[1]{Frédéric}{Achard}
-
-\affil[1]{Joint Research Center of the European Commission, Bio-economy Unit, I-21027 Ispra (VA), ITALY}
-\affil[2]{Cirad, UPR Forêts et Sociétés, F-34398 Montpellier, FRANCE}
-\affil[3]{Forêts et Sociétés, Univ. Montpellier, Cirad, Montpellier, FRANCE}
-\affil[4]{ETC Terra, F-75020 Paris, FRANCE}
-\affil[5]{Office National pour l'Environnement, 101 Antananarivo, MADAGASCAR}
-\affil[6]{Wildlife Conservation Society, 101 Antananarivo, MADAGASCAR}
-\affil[7]{GreenInfo Network, Oakland, California, USA}
-
-%% The [] brackets identify the author with the corresponding affiliation. 1, 2, 3, etc. should be inserted.
-
-\runningtitle{Six decades of deforestation in Madagascar}
-
-\runningauthor{G. Vieilledent}
-
-\correspondence{Ghislain Vieilledent (ghislain.vieilledent@cirad.fr)}
-
-\received{}
-\pubdiscuss{} %% only important for two-stage journals
-\revised{}
-\accepted{}
-\published{}
-
-%% These dates will be inserted by Copernicus Publications during the typesetting process.
-
-\firstpage{1}
+  \pretitle{\vspace{\droptitle}\centering\huge}
+  \posttitle{\par}
+  \author{}
+  \preauthor{}\postauthor{}
+  \date{}
+  \predate{}\postdate{}
+
+\usepackage{booktabs}
+\usepackage{longtable}
+\usepackage{array}
+\usepackage{multirow}
+\usepackage[table]{xcolor}
+\usepackage{wrapfig}
+\usepackage{float}
+\usepackage{colortbl}
+\usepackage{pdflscape}
+\usepackage{tabu}
+\usepackage{threeparttable}
+\usepackage[normalem]{ulem}
+
+\usepackage{amsthm}
+\newtheorem{theorem}{Theorem}[section]
+\newtheorem{lemma}{Lemma}[section]
+\theoremstyle{definition}
+\newtheorem{definition}{Definition}[section]
+\newtheorem{corollary}{Corollary}[section]
+\newtheorem{proposition}{Proposition}[section]
+\theoremstyle{definition}
+\newtheorem{example}{Example}[section]
+\theoremstyle{definition}
+\newtheorem{exercise}{Exercise}[section]
+\theoremstyle{remark}
+\newtheorem*{remark}{Remark}
+\newtheorem*{solution}{Solution}
+\begin{document}
 \maketitle
 
-\begin{abstract}
-
-  The island of Madagascar has an unparalleled biodiversity, mainly
-  located in the tropical forests of the island, which is highly
-  threatened by anthropogenic deforestation. Scattered forest maps
-  from past studies at national level with substantial gaps (due to
-  presence of cloud cover on satellite imagery) prevent the analysis
-  of long-term deforestation trends in Madagascar. In this study, we
-  propose a new approach combining historical (1953-2000) national
-  forest-cover maps with recent (2001-2014) global annual tree cover
-  loss data to look at six decades (1953-2014) of deforestation and
-  forest fragmentation in Madagascar. We produced new forest-cover
-  maps at 30~m resolution over the full territory of Madagascar for
-  the year 1990, and annually from 2000 to 2014. We estimated that
-  Madagascar has lost 44\% of its natural forest cover over the period
-  1953-2014 (including 37\% over the period 1973-2014). Natural
-  forests cover 8.9 Mha in 2014 (15\% of the national territory) which
-  are divided into 4.4 Mha (50\%) of moist forests, 2.6 Mha (29\%) of
-  dry forests, 1.7 Mha of spiny forests (19\%) and 177,000 ha (2\%) of
-  mangroves. Since 2005, the annual deforestation rate has
-  progressively increased in Madagascar to reach 99,000 ha/yr during
-  2010-2014 (corresponding to a rate of 1.1\%/yr). This increase is
-  probably due to rapid population growth (close to 3\%/yr) and to
-  poor law enforcement in the country. Around half of the forest
-  (46\%) is now located at less than 100~m from the forest
-  edge. Accurate forest-cover change maps can be used to assess the
-  effectiveness of past and current conservation programs and
-  implement new strategies for the future. In particular, forest maps
-  and estimates can be used in the REDD+ framework which aims at
-  ``Reducing Emissions from Deforestation and Forest Degradation'' and
-  for optimizing the current protected area network.
-
-\end{abstract}
-
-\copyrightstatement{The article is distributed under the Creative Commons Attribution 4.0 License.}
-
-\section{Introduction}
-\label{introduction}
-
-Separated from the African continent and the Indian plate about 165
-and 88 million years ago respectively \citep{Ali2008}, the flora and
-fauna of Madagascar followed its own evolutionary path. Isolation
-combined with a high number of micro-habitats \citep{Pearson2009} has
-led to Madagascar's exceptional biodiversity both in term of number of
-species and endemism in many taxonomic groups \citep{Crottini2012,
-  Goodman2005}. Most of the biodiversity in Madagascar is
-concentrated in the tropical forests of the island which can be
-divided into four types: the moist forest in the East, the dry forest
-in the West, the spiny forest in the South and the mangroves on the
-West coast \citep{Vieilledent2016}. This unparalleled biodiversity is
-severely threatened by deforestation \citep{Harper2007,
-  Vieilledent2013} associated with human activities such as slash-and-burn
-agriculture and pasture \citep{Scales2011}. Tropical forests in
-Madagascar also store a large amount of carbon \citep{Vieilledent2016}
-and high rates of deforestation in Madagascar are responsible for
-large CO$_2$ emissions in the atmosphere
-\citep{Achard2014}. Deforestation threatens species survival by
-directly reducing their available habitat \citep{Brooks2002,
-  Tidd2001}. Forest fragmentation can also lead to species extinction
-by isolating populations from each other and creating forest patches
-too small to maintain viable populations
-\citep{Saunders1991}. Fragmentation also increases forest edge where
-ecological conditions (such as air temperature, light intensity and
-air moisture) can be dramatically modified, with consequences on the
-abundance and distribution of species \citep{Murcia1995}. Forest
-fragmentation can also have substantial effects on forest carbon
-storage capacity, as carbon stocks are much lower at the forest edge
-than under a closed canopy \citep{Brinck2017}. Moreover, forest
-carbon stocks vary spatially due to climate or soil factors
-\citep{Saatchi2011, Vieilledent2016}. As a consequence, accurate and
-spatially explicit maps of forest-cover and forest-cover change are
-necessary to monitor biodiversity loss and carbon emissions from
-deforestation and forest fragmentation, assess the efficiency of
-present conservation strategies \citep{Eklund2016}, and implement new
-strategies for the future \citep{Vieilledent2013,
-  Vieilledent2016}. Simple time-series of forest-cover estimates, such
-as those provided by the FAO Forest Resource Assessment report
-\citep{Keenan2015} are not sufficient.
-
-Unfortunately, accurate and exhaustive forest-cover maps are not
-available for Madagascar for the last fifteen years (2000-2015).
-\citet{Harper2007} produced maps of forest cover and forest cover
-changes over Madagascar for the years \emph{c.}~1953, \emph{c.}~1973,
-1990 and 2000. The \emph{c.}~1953 forest map was derived from the
-visual interpretation of aerial photography at coarse scale
-(1/1,000,000). Forest maps for the years \emph{c.}~1973, 1990, and
-2000 were obtained from supervised classification of Landsat satellite
-images at 60~m resolution (for the year 1973) or 30~m resolution (for
-years 1990 and 2000) and can be used to derive more accurate estimates
-of forest cover (89.5\% accuracy reported for the forest/non-forest
-map of year 2000). Nonetheless, maps provided by \citet{Harper2007}
-are not exhaustive (due to the presence of clouds in the satellite
-imagery), e.g.~11 244~km2 are mapped as unknown cover type for the
-year 2000. Using a similar supervised classification approach as in
-\citet{Harper2007}, more recent maps have been produced for the
-periods 2000-2005-2010 by national institutions, with the technical
-support of international environmental NGOs \citep{MEFT2009,
-  ONE2013}. Another set of recent forest-cover maps using an advanced
-statistical tool for classification, the Random Forest classifier
-\citep{Grinand2013, Rakotomalala2015}, was produced for the periods
-2005-2010-2013 \citep{ONE2015}. However, these maps are either too old
-to give recent estimates of deforestation \citep{MEFT2009, ONE2013},
-include large areas of missing information due to images with high
-percentage of cloud cover \citep{ONE2013}, or show large
+\textbf{Ghislain Vieilledent\(^{1,2,\star}\), Clovis Grinand\(^{3}\),
+Fety A. Rakotomalala\(^{3}\), Rija Ranaivosoa\(^{4}\), Jean-Roger
+Rakotoarijaona\(^{4}\), Thomas F. Allnutt\(^{5,6}\), and Frédéric
+Achard\(^{1}\)}
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\tightlist
+\item
+  Joint Research Center of the European Commission, Bio-economy Unit
+  (JRC.D.1), I-21027 Ispra (VA), ITALY
+\item
+  Cirad, UPR Forêts et Sociétés, F-34398 Montpellier, FRANCE
+\item
+  ETC Terra, F-75016 Paris, FRANCE
+\item
+  Office National pour l'Environnement, 101 Antananarivo, MADAGASCAR
+\item
+  Wildlife Conservation Society, 101 Antananarivo, MADAGASCAR
+\item
+  GreenInfo Network, Oakland, California, USA
+\end{enumerate}
+
+(*) /Email:
+\href{mailto:ghislain.vieilledent@cirad.fr}{\nolinkurl{ghislain.vieilledent@cirad.fr}},
+/Phone: +39.329.457.2273
+
+\emph{Running headline:} Six decades of deforestation in Madagascar
+
+\hypertarget{highlights}{%
+\section{Highlights}\label{highlights}}
+
+\begin{itemize}
+\tightlist
+\item
+  We produced new 30m-resolution annual forest-cover maps for Madagascar
+  for the period 2000-2014.
+\item
+  Madagascar has lost 44\% of its natural forest-cover over the period
+  1953-2014.
+\item
+  Half of the tropical forest in Madagascar is now located at less than
+  100m from forest edge.
+\item
+  Annual deforestation rate has increased in Madagascar since 2005 to
+  reach 1.08\%/yr.
+\item
+  Conservation and development efforts must be intensified to save
+  Madagascar forest and biodiversity.
+\end{itemize}
+
+\hypertarget{summary}{%
+\section{Summary}\label{summary}}
+
+The island of Madagascar has a unique biodiversity, mainly located in
+the tropical forests of the island. This biodiversity is highly
+threatened by anthropogenic deforestation. Existing historical forest
+maps at national level are scattered and have substantial gaps which
+prevent an exhaustive assessment of long-term deforestation trends in
+Madagascar. In this study, we combine historical national forest cover
+maps (covering the period 1953-2000) with a recent global annual tree
+cover loss dataset (2001-2014) to look at six decades of deforestation
+and forest fragmentation in Madagascar (from 1953 to 2014). We produced
+new forest cover maps at 30~m resolution for the year 1990 and annually
+from 2000 to 2014 over the full territory of Madagascar. We estimated
+that Madagascar has lost 44\% of its natural forest cover over the
+period 1953-2014 (including 37\% over the period 1973-2014). Natural
+forests cover 8.9 Mha in 2014 (15\% of the national territory) and
+include 4.4 Mha (50\%) of moist forests, 2.6 Mha (29\%) of dry forests,
+1.7 Mha of spiny forests (19\%) and 177,000 ha (2\%) of mangroves. Since
+2005, the annual deforestation rate has progressively increased in
+Madagascar to reach 99,000 ha/yr during 2010-2014 (corresponding to a
+rate of 1.1\%/yr). This increase is probably due to rapid population
+growth (close to 3\%/yr) and to poor law enforcement in the country.
+Around half of the forest (46\%) is now located at less than 100~m from
+the forest edge. Accurate forest cover change maps can be used to assess
+the effectiveness of past and current conservation programs and
+implement new strategies for the future. In particular, forest maps and
+estimates can be used in the REDD+ framework which aims at ``Reducing
+Emissions from Deforestation and Forest Degradation'' and for optimizing
+the current protected area network.
+
+\textbf{Keywords}: biodiversity, climate-change, deforestation,
+forest-fragmentation, Madagascar, tropical forest.
+
+\hypertarget{introduction}{%
+\section{Introduction}\label{introduction}}
+
+Separated from the African continent and the Indian plate about 165 and
+88 million years ago respectively (Ali \& Aitchison
+\protect\hyperlink{ref-Ali2008}{2008}), the flora and fauna of
+Madagascar followed its own evolutionary path. Isolation combined with a
+high number of micro-habitats (Pearson \& Raxworthy
+\protect\hyperlink{ref-Pearson2009}{2009}) has led to Madagascar's
+exceptional biodiversity both in term of number of species and endemism
+in many taxonomic groups (Goodman \& Benstead
+\protect\hyperlink{ref-Goodman2005}{2005}; Crottini \emph{et al.}
+\protect\hyperlink{ref-Crottini2012}{2012}). Most of the biodiversity in
+Madagascar is concentrated in the tropical forests of the island which
+can be divided into four types: the moist forest in the East, the dry
+forest in the West, the spiny forest in the South and the mangroves on
+the West coast (Vieilledent \emph{et al.}
+\protect\hyperlink{ref-Vieilledent2016}{2016}). This unparalleled
+biodiversity is severely threatened by deforestation (Harper \emph{et
+al.} \protect\hyperlink{ref-Harper2007}{2007}; Vieilledent, Grinand \&
+Vaudry \protect\hyperlink{ref-Vieilledent2013}{2013}) associated with
+human activities such as slash-and-burn agriculture and pasture (Scales
+\protect\hyperlink{ref-Scales2011}{2011}). Tropical forests in
+Madagascar also store a large amount of carbon (Vieilledent \emph{et
+al.} \protect\hyperlink{ref-Vieilledent2016}{2016}) and high rates of
+deforestation in Madagascar are responsible for large CO\(_2\) emissions
+in the atmosphere (Achard \emph{et al.}
+\protect\hyperlink{ref-Achard2014}{2014}). Deforestation threatens
+species survival by directly reducing their available habitat (Tidd,
+Pinder \& Ferguson \protect\hyperlink{ref-Tidd2001}{2001}; Brooks
+\emph{et al.} \protect\hyperlink{ref-Brooks2002}{2002}). Forest
+fragmentation can also lead to species extinction by isolating
+populations from each other and creating forest patches too small to
+maintain viable populations (Saunders, Hobbs \& Margules
+\protect\hyperlink{ref-Saunders1991}{1991}). Fragmentation also
+increases forest edge where ecological conditions (such as air
+temperature, light intensity and air moisture) can be dramatically
+modified, with consequences on the abundance and distribution of species
+(Murcia \protect\hyperlink{ref-Murcia1995}{1995}). Forest fragmentation
+can also have substantial effects on forest carbon storage capacity, as
+carbon stocks are much lower at the forest edge than under a closed
+canopy (Brinck \emph{et al.} \protect\hyperlink{ref-Brinck2017}{2017}).
+Moreover, forest carbon stocks vary spatially due to climate or soil
+factors (Saatchi \emph{et al.}
+\protect\hyperlink{ref-Saatchi2011}{2011}; Vieilledent \emph{et al.}
+\protect\hyperlink{ref-Vieilledent2016}{2016}). As a consequence,
+accurate and spatially explicit maps of forest cover and forest cover
+change are necessary to monitor biodiversity loss and carbon emissions
+from deforestation and forest fragmentation, assess the efficiency of
+present conservation strategies (Eklund \emph{et al.}
+\protect\hyperlink{ref-Eklund2016}{2016}), and implement new strategies
+for the future (Vieilledent, Grinand \& Vaudry
+\protect\hyperlink{ref-Vieilledent2013}{2013}; Vieilledent \emph{et al.}
+\protect\hyperlink{ref-Vieilledent2016}{2016}). Simple time-series of
+forest cover estimates, such as those provided by the FAO Forest
+Resource Assessment report (Keenan \emph{et al.}
+\protect\hyperlink{ref-Keenan2015}{2015}) are not sufficient.
+
+Unfortunately, accurate and exhaustive forest cover maps are not
+available for Madagascar after year 2000. Harper \emph{et al.}
+(\protect\hyperlink{ref-Harper2007}{2007}) produced maps of forest cover
+and forest cover changes over Madagascar for the years \emph{c.}~1953,
+\emph{c.}~1973, 1990 and 2000. The \emph{c.}~1953 forest map was derived
+from the visual interpretation of aerial photography at coarse scale
+(1/1,000,000). Forest maps for the years \emph{c.}~1973, 1990, and 2000
+were obtained from supervised classification of Landsat satellite images
+at 60~m resolution (for the year 1973) or 30~m resolution (for years
+1990 and 2000) and can be used to derive more accurate estimates of
+forest cover (89.5\% accuracy reported for the forest/non-forest map of
+year 2000). Nonetheless, maps provided by Harper \emph{et al.}
+(\protect\hyperlink{ref-Harper2007}{2007}) are not exhaustive (due to
+the presence of clouds in the satellite imagery), e.g.~11 244~km2 are
+mapped as unknown cover type for the year 2000. Using a similar
+supervised classification approach as in Harper \emph{et al.}
+(\protect\hyperlink{ref-Harper2007}{2007}), more recent maps have been
+produced for the periods 2000-2005-2010 by national institutions, with
+the technical support of international environmental NGOs (MEFT, USAID,
+and CI \protect\hyperlink{ref-MEFT2009}{2009}; ONE, DGF, FTM, MNP, and
+CI \protect\hyperlink{ref-ONE2013}{2013}). Another set of recent forest
+cover maps using an advanced statistical tool for classification, the
+Random Forest classifier (Grinand \emph{et al.}
+\protect\hyperlink{ref-Grinand2013}{2013}; Rakotomala \emph{et al.}
+\protect\hyperlink{ref-Rakotomalala2015}{2015}), was produced for the
+periods 2005-2010-2013 (ONE, DGF, MNP, WCS, and Etc Terra
+\protect\hyperlink{ref-ONE2015}{2015}). However, these maps are either
+too old to give recent estimates of deforestation (MEFT, USAID, and CI
+\protect\hyperlink{ref-MEFT2009}{2009}; ONE, DGF, FTM, MNP, and CI
+\protect\hyperlink{ref-ONE2013}{2013}), include large areas of missing
+information due to images with high percentage of cloud cover (ONE, DGF,
+FTM, MNP, and CI \protect\hyperlink{ref-ONE2013}{2013}), or show large
 mis-classification in specific areas, especially in the dry and spiny
@@
 lower than 0.8 for the dry and spiny forests for the maps produced by
-\citet{ONE2015}). Moreover, the production of such forest maps from a
-supervised classification approach requires significant resources,
-especially regarding the image selection step (required to minimize
-cloud cover) and the training step (visual interpretation of a large
-number of polygons needed to train the classification algorithm)
-\citep{Rakotomalala2015}. Most of this work of image selection and
-visual interpretation would need to be repeated to produce new forest
-maps in the future using a similar approach.
+ONE, DGF, MNP, WCS, and Etc Terra
+(\protect\hyperlink{ref-ONE2015}{2015})). Moreover, the production of
+such forest maps from a supervised classification approach requires
+significant resources, especially regarding the image selection step
+(required to minimize cloud cover) and the training step (visual
+interpretation of a large number of polygons needed to train the
+classification algorithm) (Rakotomala \emph{et al.}
+\protect\hyperlink{ref-Rakotomalala2015}{2015}). Most of this work of
+image selection and visual interpretation would need to be repeated to
+produce new forest maps in the future using a similar approach.
 
 Global forest or tree cover products have also been published recently
-and can be tested at the national scale for
-Madagascar. \citet{Kim2014} produced a global forest-cover change map
-from 1990 to 2000 (derived from Landsat imagery). This product was
-updated to cover the period 1975-2005
-(\url{http://glcf.umd.edu/data/landsatFCC/}) but forest-cover maps
-after 2005 were not produced. Moreover, the approach used in
-\citet{Kim2014} did not accurately map the forests in the dry and
-spiny ecosystems of Madagascar (see Fig. 8 in \citet{Kim2014}).
-\citet{Hansen2013} mapped tree cover percentage, annual tree cover
-loss and gain from 2000 to 2012 at global scale at 30 m
-resolution. This product has since been updated and is now available
-up to the year 2014 \citep{Hansen2013}. To map forest cover from the
-\citet{Hansen2013} product, a tree cover threshold must be selected
-(that defines forest cover). Selecting such a threshold is not
-straightforward as the accuracy of the global tree cover map strongly
-varies between forest types, and is substantially lower for dry
-forests than for moist forests \citep{Bastin2017}. Moreover, the
-\citet{Hansen2013} product does not provide information on
-land-use. In particular the global tree cover map does not separate
-tree plantations such as oil palm or eucalyptus plantations from
-natural forests \citep{Tropek2014}. Thus, the global tree cover map
-from \citet{Hansen2013} cannot be used alone to produce a map of
-forest cover \citep{Tyukavina2017}.
-
-In this study, we present a simple approach which combines the maps
-from \citet{Harper2007} and products from \citet{Hansen2013} to derive
-annual wall-to-wall forest-cover change maps over the period 2000-2014
-for Madagascar. We use the forest-cover map provided by
-\citet{Harper2007} for the year 2000 (defining the land-use) with the
-tree cover loss product provided by \citet{Hansen2013} that we apply
-only inside forest areas identified by \citet{Harper2007}. Similar to
-the approach of \citet{Harper2007}, we also assess trends in
-deforestation rates and forest fragmentation from \emph{c.} 1953 to 2014. The
-approach described in this study can help assess the effectiveness of
-current conservation strategies, and assist the implementation of
-future strategies. Our approach could be easily extended to other
-tropical countries that have at least one forest-cover map between
-2000 and 2014. This approach can easily be repeated in the future when
-tree cover loss products will be updated.
-
-\section{Materials and Methods}
-\label{materials-and-methods}
-
-\subsection{Creation of new forest-cover maps of Madagascar from
-1953 to 2014}
+and can be tested at the national scale for Madagascar. Kim \emph{et
+al.} (\protect\hyperlink{ref-Kim2014}{2014}) produced a global forest
+cover change map from 1990 to 2000 (derived from Landsat imagery). This
+product was updated to cover the period 1975-2005
+(\url{http://glcf.umd.edu/data/landsatFCC/}) but forest cover maps after
+2005 were not produced. Moreover, the approach used in Kim \emph{et al.}
+(\protect\hyperlink{ref-Kim2014}{2014}) did not accurately map the
+forests in the dry and spiny ecosystems of Madagascar (see Fig. 8 in Kim
+\emph{et al.} (\protect\hyperlink{ref-Kim2014}{2014})). Hansen \emph{et
+al.} (\protect\hyperlink{ref-Hansen2013}{2013}) mapped tree cover
+percentage, annual tree cover loss and gain from 2000 to 2012 at global
+scale at 30 m resolution. This product has since been updated and is now
+available up to the year 2014 (Hansen \emph{et al.}
+\protect\hyperlink{ref-Hansen2013}{2013}). To map forest cover from the
+Hansen \emph{et al.} (\protect\hyperlink{ref-Hansen2013}{2013}) product,
+a tree cover threshold must be selected (that defines forest cover).
+Selecting such a threshold is not straightforward as the accuracy of the
+global tree cover map strongly varies between forest types, and is
+substantially lower for dry forests than for moist forests (Bastin
+\emph{et al.} \protect\hyperlink{ref-Bastin2017}{2017}). Moreover, the
+Hansen \emph{et al.} (\protect\hyperlink{ref-Hansen2013}{2013}) product
+does not provide information on land-use. In particular the global tree
+cover map does not separate tree plantations such as oil palm or
+eucalyptus plantations from natural forests (Tropek \emph{et al.}
+\protect\hyperlink{ref-Tropek2014}{2014}). Thus, the global tree cover
+map from Hansen \emph{et al.} (\protect\hyperlink{ref-Hansen2013}{2013})
+cannot be used alone to produce a map of forest cover (Tyukavina
+\emph{et al.} \protect\hyperlink{ref-Tyukavina2017}{2017}).
+
+In this study, we present a simple approach which combines the
+historical forest maps from Harper \emph{et al.}
+(\protect\hyperlink{ref-Harper2007}{2007}) and more recent global
+products from Hansen \emph{et al.}
+(\protect\hyperlink{ref-Hansen2013}{2013}) to derive annual wall-to-wall
+forest cover change maps over the period 2000-2014 for Madagascar. We
+use the forest cover map provided by Harper \emph{et al.}
+(\protect\hyperlink{ref-Harper2007}{2007}) for the year 2000 (defining
+the land-use) with the tree cover loss product provided by Hansen
+\emph{et al.} (\protect\hyperlink{ref-Hansen2013}{2013}) that we apply
+only inside forest areas identified by Harper \emph{et al.}
+(\protect\hyperlink{ref-Harper2007}{2007}). Similar to the approach of
+Harper \emph{et al.} (\protect\hyperlink{ref-Harper2007}{2007}), we also
+assess trends in deforestation rates and forest fragmentation from
+\emph{c.}~1953 to 2014. We finally discuss the possibility to extend our
+approach to other tropical countries or repeat it in the future. We also
+discuss how our results could help assess the effectiveness of current
+conservation strategies, and implementation new conservation strategies
+for the future in Madagascar.
+
+\hypertarget{materials-and-methods}{%
+\section{Materials and Methods}\label{materials-and-methods}}
+
+\hypertarget{creation-of-new-forest-cover-maps-of-madagascar-from-1953-to-2014}{%
+\subsection{Creation of new forest cover maps of Madagascar from 1953 to
+2014}\label{creation-of-new-forest-cover-maps-of-madagascar-from-1953-to-2014}}
 
@@
 full territory of Madagascar for the period 2000-2014 by combining the
-forest map of year 2000 from \citet{Harper2007}, and the tree cover
+forest map of year 2000 from Harper \emph{et al.}
+(\protect\hyperlink{ref-Harper2007}{2007}), and the tree cover
 percentage and annual tree cover loss maps over the period 2000-2014
-from \citet{Hansen2013}. The 2000 Harper's forest map includes 208,000
-ha of unclassified areas due to the presence of clouds on satellite
-images, mostly (88\%) within the moist forest domain which covered
-4.17 Mha in total in 2000. To provide a label (forest or non-forest)
-to these unclassified pixels, we used the 2000 tree cover percentage
-map of \citet{Hansen2013} by selecting a threshold of 75\% tree cover
-to define forest cover as recommended by other studies for the moist
-domain \citep{Achard2014, Aleman2017}. To do so, the Hansen's 2000
+from Hansen \emph{et al.} (\protect\hyperlink{ref-Hansen2013}{2013}).
+The 2000 Harper's forest map includes 208,000 ha of unclassified areas
+due to the presence of clouds on satellite images, mostly (88\%) within
+the moist forest domain which covered 4.17 Mha in total in 2000. To
+provide a label (forest or non-forest) to these unclassified pixels, we
+used the 2000 tree cover percentage map of Hansen \emph{et al.}
+(\protect\hyperlink{ref-Hansen2013}{2013}) by selecting a threshold of
+75\% tree cover to define forest cover as recommended by other studies
+for the moist domain (Achard \emph{et al.}
+\protect\hyperlink{ref-Achard2014}{2014}; Aleman, Jarzyna \& Staver
+\protect\hyperlink{ref-Aleman2017}{2017}). To do so, the Hansen's 2000
 tree cover map was resampled on the same grid as the original Harper's
 map at 30~m resolution using a bilinear interpolation. We thus obtained
-a forest-cover map for the year 2000 covering the full territory of
-Madagascar. We then combined this forest-cover map of the year 2000
-with the annual tree cover loss maps from 2001 to 2014 provided by
-\citet{Hansen2013} to create annual forest-cover maps from 2001 to
-2014 at 30~m resolution. To do so, Hansen's tree cover loss maps were
-resampled on the same grid as the original Harper's map at 30~m
-resolution using a nearest-neighbor interpolation. We also completed
-the Harper's forest map of year 1990 by filling unclassified areas
-(due to the presence of clouds on satellite images) using our
-forest-cover map of year 2000. To do so, we assumed that if forest was
-present in 2000, the pixel was also forested in 1990. The remaining
-unclassified pixels were limited to a relatively small total area of
-\emph{c.} 8,000 ha. We labeled these residual pixels as non-forest, as
-for the year 2000. Similarly we completed the Harper's forest map of
-year 1973 by filling unclassified areas using our forest-cover map of
-the year 1990 assuming that if forest was present in 1990, it was also
-present in 1973. Contrary to the year 1990, the remaining unclassified
-pixels for year 1973 corresponded to a significant total area of 3.3
-million ha. We also reprojected the forest-cover map of year 1953 to a
-common projection in order to compare the forest-cover area in 1953
-with forest-cover areas at the following dates. This map was produced
-by scanning a paper map derived from aerial photos, and thus could not
-be perfectly aligned with the other maps produced through digital
-processing of satellite imagery \citep{Harper2007}. Finally for all
-forest-cover maps from 1973, the isolated single non-forest pixels
-(i.e.~fully surrounded by forest pixels) were removed, assuming that
-single non-forest pixels inside a forest patch were not corresponding
-to deforestation (they might correspond to selective logging
-activities). This allowed us to avoid counting very small scale events
-(\textless 0.1 ha, such as selective logging) as forest
-fragmentation. All the resulting maps are freely available at
-\url{https://bioscenemada.cirad.fr/forestmaps}.
-
-\subsection{Computing forest-cover areas and deforestation
-rates}
-
-From these new forest-cover maps, we calculated the total forest-cover
-area for seven available years (1953-1973-1990-2000-2005-2010-2014),
-and the annual deforested area and annual deforestation rate for the
+a forest cover map for the year 2000 covering the full territory of
+Madagascar. We then combined this forest cover map of the year 2000 with
+the annual tree cover loss maps from 2001 to 2014 provided by Hansen
+\emph{et al.} (\protect\hyperlink{ref-Hansen2013}{2013}) to create
+annual forest cover maps from 2001 to 2014 at 30~m resolution. To do so,
+Hansen's tree cover loss maps were resampled on the same grid as the
+original Harper's map at 30~m resolution using a nearest-neighbor
+interpolation. We also completed the Harper's forest map of year 1990 by
+filling unclassified areas (due to the presence of clouds on satellite
+images) using our forest cover map of year 2000. To do so, we assumed
+that if forest was present in 2000, the pixel was also forested in 1990.
+The remaining unclassified pixels were limited to a relatively small
+total area of \emph{c.}~8,000 ha. We labeled these residual pixels as
+non-forest, as for the year 2000. Similarly we completed the Harper's
+forest map of year 1973 by filling unclassified areas using our forest
+cover map of the year 1990 assuming that if forest was present in 1990,
+it was also present in 1973. Contrary to the year 1990, the remaining
+unclassified pixels for year 1973 corresponded to a significant total
+area of 3.3 million ha. We also reprojected the forest cover map of year
+1953 to a common projection in order to compare the forest cover area in
+1953 with forest cover areas at the following dates. This map was
+produced by scanning a paper map derived from aerial photos, and thus
+could not be perfectly aligned with the other maps produced through
+digital processing of satellite imagery (Harper \emph{et al.}
+\protect\hyperlink{ref-Harper2007}{2007}). Finally for all forest cover
+maps from 1973, the isolated single non-forest pixels (i.e.~fully
+surrounded by forest pixels) were removed, assuming that single
+non-forest pixels inside a forest patch were not corresponding to
+deforestation (they might correspond to selective logging activities).
+This allowed us to avoid counting very small scale events
+(\textless{}0.1 ha, such as selective logging) as forest fragmentation.
+All the resulting maps have been made permanently and freely available
+on the Zenodo research data repository at
+\url{https://doi.org/10.5281/zenodo.1145785}.
+
+\hypertarget{computing-forest-cover-areas-and-deforestation-rates}{%
+\subsection{Computing forest cover areas and deforestation
+rates}\label{computing-forest-cover-areas-and-deforestation-rates}}
+
+From these new forest cover maps, we calculated the total forest cover
+area for seven available years (1953-1973-1990-2000-2005-2010-2014), and
+the annual deforested area and annual deforestation rate for the
 corresponding six time periods between 1953 and 2014. The annual
-deforestation rates were calculated using Eq.~\ref{eq:theta}
-\citep{Puyravaud2003, Vieilledent2013}:
+deforestation rates were calculated using Eq.~\eqref{eq:theta} (Puyravaud
+\protect\hyperlink{ref-Puyravaud2003}{2003}; Vieilledent, Grinand \&
+Vaudry \protect\hyperlink{ref-Vieilledent2013}{2013}):
 
@@
   \label{eq:theta}
-  \theta = 100 \times [1-(1-(F_{t_2}-F_{t_1})/F_{t_1})^{(1/(t_2-t_1))}
+  \theta = 100 \times [1-(1-(F_{t_2}-F_{t_1})/F_{t_1})^{(1/(t_2-t_1))}]
 \end{equation}
 
-In Eq.~\ref{eq:theta}, $\theta$ is the annual deforestation rate (in \%/yr),
-$F_{t_2}$ and $F_{t_1}$ are the forest cover free of clouds at both
-dates $t_2$ and $t_1$, and $t_2-t_1$ is the time-interval (in
-years) between the two dates.
+In Eq.~\eqref{eq:theta}, \(\theta\) is the annual deforestation rate (in
+\%/yr), \(F_{t_2}\) and \(F_{t_1}\) are the forest cover free of clouds
+at both dates \(t_2\) and \(t_1\), and \(t_2-t_1\) is the time-interval
+(in years) between the two dates.
 
@@
 available forest extent. Area and rate estimates are produced at the
-national scale and for the four forest ecosystems present in
-Madagascar: moist forest in the East, dry forest in the West, spiny
-forest in the South, and mangroves on the Western coast
-(Fig.~\ref{fig:ecoregion}). To define the forest domains, we used a
-map from the MEFT (\emph{``Ministère de l'Environnement et des Forêts
-  à Madagascar''}) with the boundaries of the four ecoregions in
-Madagascar. Ecoregions were defined on the basis of climatic and
-vegetation criteria using the climate classification by
-\citet{Cornet1974} and the vegetation classification from the 1996
-IEFN national forest inventory \citep{IEFN1996}. Because mangrove
-forests are highly dynamic ecosystems that can expand or contract on
-decadal scales depending on changes in environmental factors
-\citep{Armitage2015}, a fixed delimitation of the mangrove ecoregion
-on six decades might not be fully appropriate. As a consequence, our
-estimates of the forest-cover and deforestation rates for mangroves in
-Madagascar must be considered with this limitation.
-
-\begin{figure}[h]
-  \centering
-
-    \includegraphics[width=8.3cm]{ecoregion.png}
-
-    \caption{\textbf{Ecoregions and forest types in Madagascar.}
-      Madagascar can be divided into four climatic ecoregions with four forest
-      types: the moist forest in the East (green), the dry forest in the West
-      (orange), the spiny forest in the South (red), and the mangroves on the
-      West coast (blue). Ecoregions were defined following climatic
-      \citep{Cornet1974} and vegetation \citep{IEFN1996} criteria. The dark
-      grey areas represent the remaining natural forest cover for the year
-      2014.}
-
-    \label{fig:ecoregion}
-
-\end{figure}
-
-\subsection{Comparing our forest-cover and deforestation rate
-estimates with previous studies}
-
-We compared our estimates of forest-cover and deforestation rates with
+national scale and for the four forest ecosystems present in Madagascar:
+moist forest in the East, dry forest in the West, spiny forest in the
+South, and mangroves on the Western coast (Fig.~\ref{fig:ecoregions}).
+To define the forest domains, we used a map from the MEFT
+(\emph{``Ministère de l'Environnement et des Forêts à Madagascar''})
+with the boundaries of the four ecoregions in Madagascar. Ecoregions
+were defined on the basis of climatic and vegetation criteria using the
+climate classification by Cornet
+(\protect\hyperlink{ref-Cornet1974}{1974}) and the vegetation
+classification from the 1996 IEFN national forest inventory (Ministère
+de l'Environnement \protect\hyperlink{ref-IEFN1996}{1996}). Because
+mangrove forests are highly dynamic ecosystems that can expand or
+contract on decadal scales depending on changes in environmental factors
+(Armitage \emph{et al.} \protect\hyperlink{ref-Armitage2015}{2015}), a
+fixed delimitation of the mangrove ecoregion on six decades might not be
+fully appropriate. As a consequence, our estimates of the forest cover
+and deforestation rates for mangroves in Madagascar must be considered
+with this limitation.
+
+\hypertarget{comparing-our-forest-cover-and-deforestation-rate-estimates-with-previous-studies}{%
+\subsection{Comparing our forest cover and deforestation rate estimates
+with previous
+studies}\label{comparing-our-forest-cover-and-deforestation-rate-estimates-with-previous-studies}}
+
+We compared our estimates of forest cover and deforestation rates with
 estimates from the three existing studies at the national scale for
-Madagascar: (i) \citep{Harper2007}, (ii) \citep{MEFT2009} and (iii)
-\citep{ONE2015}. \citet{Harper2007} provides forest-cover and
-deforestation estimates for the periods
-c. 1953-c. 1973-1990-2000. MEFT, USAID, and CI (2009) provides
-estimates for the periods 1990-2000-2005 and ONE, DGF, MNP, WCS, and
-Etc Terra (2015) provides estimates for the periods 2005-2010-2013. To
-compare our forest-cover and deforestation estimates over the same
-time periods, we consider an additional time-period in our study
-(2010-2013) by creating an extra forest-cover map for the year
-2013. We computed the Pearson's correlation coefficient and the root
-mean square error (RMSE) between our forest-cover estimates and
-forest-cover estimates from previous studies for all the dates and
+Madagascar: (i) Harper \emph{et al.}
+(\protect\hyperlink{ref-Harper2007}{2007}), (ii) MEFT, USAID, and CI
+(\protect\hyperlink{ref-MEFT2009}{2009}) and (iii) ONE, DGF, MNP, WCS,
+and Etc Terra (\protect\hyperlink{ref-ONE2015}{2015}). Harper \emph{et
+al.} (\protect\hyperlink{ref-Harper2007}{2007}) provides forest cover
+and deforestation estimates for the periods c. 1953-c. 1973-1990-2000.
+MEFT, USAID, and CI (\protect\hyperlink{ref-MEFT2009}{2009}) provides
+estimates for the periods 1990-2000-2005 and ONE, DGF, MNP, WCS, and Etc
+Terra (\protect\hyperlink{ref-ONE2015}{2015}) provides estimates for the
+periods 2005-2010-2013. To compare our forest cover and deforestation
+estimates over the same time periods, we consider an additional
+time-period in our study (2010-2013) by creating an extra forest cover
+map for the year 2013. We computed the Pearson's correlation coefficient
+and the root mean square error (RMSE) between our forest cover estimates
+and forest cover estimates from previous studies for all the dates and
 forest types (including also the total forest cover estimates). For
@@
 \%/yr) is not always detailed and might slightly differ from one study
-to another \citep[see][]{Puyravaud2003}. \citet{Harper2007} also
+to another (see Puyravaud \protect\hyperlink{ref-Puyravaud2003}{2003}).
+Harper \emph{et al.} (\protect\hyperlink{ref-Harper2007}{2007}) also
 provide total deforested areas for the two periods 1973-1990 and
 1990-2000. We converted these values into annual deforested area
-estimates. When annual deforested areas were not reported (for
-1953-1973 in \citet{Harper2007} and in \citet{MEFT2009} and
-\citet{ONE2015}), we computed them from the forest-cover estimates in
-each study. These estimates cannot be corrected from the potential
-bias due to the presence of residual clouds. Forest-cover and
-deforestation rates were then compared between all studies for the
-whole of Madagascar and the four ecoregions. The same ecoregion
-boundaries as in our study were used in \citet{ONE2015} but this was
-not the case for \citet{Harper2007} and \citet{MEFT2009}, which can
+estimates. When annual deforested areas were not reported (for 1953-1973
+in Harper \emph{et al.} (\protect\hyperlink{ref-Harper2007}{2007}) and
+in MEFT, USAID, and CI (\protect\hyperlink{ref-MEFT2009}{2009}) and ONE,
+DGF, MNP, WCS, and Etc Terra (\protect\hyperlink{ref-ONE2015}{2015})),
+we computed them from the forest cover estimates in each study. These
+estimates cannot be corrected from the potential bias due to the
+presence of residual clouds. Forest cover and deforestation rates were
+then compared between all studies for the whole of Madagascar and the
+four ecoregions. The same ecoregion boundaries as in our study were used
+in ONE, DGF, MNP, WCS, and Etc Terra
+(\protect\hyperlink{ref-ONE2015}{2015}) but this was not the case for
+Harper \emph{et al.} (\protect\hyperlink{ref-Harper2007}{2007}) and
+MEFT, USAID, and CI (\protect\hyperlink{ref-MEFT2009}{2009}), which can
 explain a part of the differences between the estimates.
 
-\subsection{Fragmentation}
-
-We also conducted an analysis of changes in forest fragmentation for
-the years 1953, 1973, 1990, 2000, 2005, 2010 and 2014 at 30~m
-resolution. We applied the method developed by \citet{Riitters2000}
-which uses a moving window to characterize the fragmentation around
-each forested pixel. Computations were done using the function
-\texttt{r.forestfrag} of the GRASS GIS software
-\citep{Neteler2008}. Six categories of fragmentation were identified
-from the amount of forest and its occurrence as adjacent forest
-pixels: ``interior'', ``perforated'', ``edge'', ``transitional'',
-``patch'', and ``undetermined''. Previous studies have shown that
-micro-habitats were mainly altered within the first 100~m of the
-forest edge \citep{Brinck2017, Gibson2013, Murcia1995,
-  Broadbent2008}. To characterize fragmentation, we have used a moving
-window of 7x7 pixels (4.4~ha). Using this window size, forest edge had
-a width of about 90~m \citep{Riitters2000}, which is close to the
-value of 100~m and conservative regarding edge effects. The
-``interior'' category can be interpreted as the most intact forest
-\citep{Potapov2017}. The ``patch'' and ``transitional'' categories
-correspond to isolated small forest patches. We reported the area of
-forest in each fragmentation category for the six years and analyzed
-the dynamics of fragmentation over the six decades. We also computed
-the distance to forest edge for all forest pixels for the years 1953,
-1973, 1990, 2000, 2005, 2010 and 2014. For that, we used the function
-\texttt{gdal\_proximity.py} of the GDAL library
+\hypertarget{fragmentation}{%
+\subsection{Fragmentation}\label{fragmentation}}
+
+We also conducted an analysis of changes in forest fragmentation for the
+years 1953, 1973, 1990, 2000, 2005, 2010 and 2014 at 30~m resolution. We
+used a moving window of \(51 \times 51\) pixels centered on each forest
+pixel to compute the percentage of forest pixels in the neighborhood. We
+used this percentage as an indication of the forest fragmentation. The
+size of the moving windows was based on a compromise: a sufficiently
+high number of cells (here 2601) had to be considered to be able to
+compute a percentage and a reasonably low number of cells had to be
+choosen to have a local estimate of the fragmentation. Computations were
+done using the function \texttt{r.neighbors} of the GRASS GIS software
+(Neteler \& Mitasova \protect\hyperlink{ref-Neteler2008}{2008}). Using
+the density of forest in the neighborhood, we defined five forest
+fragmentation classes: 0-20\% (highly fragmented), 21-40\%, 41-60\%,
+61-80\% and 81-100\% (lowly fragmented). We reported the percentage of
+forest falling in each fragmentation class for the six years and
+analyzed the dynamics of fragmentation over the six decades.
+
+We also computed the distance to forest edge for all forest pixels for
+the years 1953, 1973, 1990, 2000, 2005, 2010 and 2014. For that, we used
+the function \texttt{gdal\_proximity.py} of the GDAL library
 (\url{http://www.gdal.org/}). We computed the mean and 90\% quantiles
 (5\% and 95\%) of the distance to forest edge and looked at the
-evolution of these values with time.
-
-\begin{figure*}[h]
-  \centering
-
-  \includegraphics[width=\textwidth]{fig_fcc_highres.png}
-
-  \caption{\textbf{Forest-cover change on six decades from 1953
-      to 2014 in Madagascar.} Forest cover changes from \emph{c.} 1973 to 2014
-    are shown in the main figure, and forest cover in \emph{c.} 1953 is
-    shown in the bottom-right inset. Two zooms in the western dry (left
-    part) and eastern moist (right part) ecoregions present more detailed
-    views of (from top to bottom): forest-cover in 1950s, forest-cover
-    change from \emph{c.} 1973 to 2014, forest fragmentation in 2014 and
-    distance to forest edge in 2014. Data on water bodies (blue) and water
-    seasonality (light blue for seasonal water to dark blue for permanent
-    water) has been extracted from \citet{Pekel2016}.}
-
-  \label{fig:fcc}
-
-\end{figure*}
-
-\section{Results}
-\label{results}
-
-\subsection{Dynamics of forest cover and deforestation
-intensity}
+evolution of these values with time. Previous studies have shown that
+forest micro-habitats were mainly altered within the first 100~m of the
+forest edge (Murcia \protect\hyperlink{ref-Murcia1995}{1995}; Broadbent
+\emph{et al.} \protect\hyperlink{ref-Broadbent2008}{2008}; Gibson
+\emph{et al.} \protect\hyperlink{ref-Gibson2013}{2013}; Brinck \emph{et
+al.} \protect\hyperlink{ref-Brinck2017}{2017}). Consequently, we also
+estimated the percentage of forest within the first 100~m of the forest
+edge for each year and looked at the evolution of this percentage over
+the six decades.
+
+\hypertarget{results}{%
+\section{Results}\label{results}}
+
+\hypertarget{forest-cover-change-and-deforestation-rates}{%
+\subsection{Forest cover change and deforestation
+rates}\label{forest-cover-change-and-deforestation-rates}}
 
 Natural forests in Madagascar covered 16.0 Mha in 1953, about 27\% of
-the national territory of 587,041 km2. In 2014, the forest cover
-dropped to 8.9 Mha, corresponding to about 15\% of the national
-territory (Fig.~\ref{fig:fcc} and
-Tab.~\ref{tab:forest_cover}). Madagascar has lost 44\% and 37\% of its
-natural forests between 1953 and 2014, and between 1973 and 2014
-respectively (Fig.~\ref{fig:fcc} and Tab.~\ref{tab:forest_cover}). In
-2014 the remaining 8.9 Mha of natural forest were distributed as follow: 4.4
-Mha of moist forest (50\% of total forest cover), 2.6 Mha of dry
-forest (29\%), 1.7 Mha of spiny forest (19\%) and 0.18 Mha (2\%) of
-mangrove forest (Fig.~\ref{fig:ecoregion} and
-Tab.~\ref{tab:comp_forest}). Regarding the deforestation trend, we
+the national territory of 587,041 km2. In 2014, the forest cover dropped
+to 8.9 Mha, corresponding to about 15\% of the national territory
+(Fig.\textasciitilde{}\ref{fig:fcc} and Tab.~@ref(tab:forest\_cover)).
+Madagascar has lost 44\% and 37\% of its natural forests between 1953
+and 2014, and between 1973 and 2014 respectively
+(Fig.\textasciitilde{}\ref{fig:fcc} and Tab.~@ref(tab:forest\_cover)).
+In 2014 the remaining 8.9 Mha of natural forest were distributed as
+follow: 4.4 Mha of moist forest (50\% of total forest cover), 2.6 Mha of
+dry forest (29\%), 1.7 Mha of spiny forest (19\%) and 0.18 Mha (2\%) of
+mangrove forest (Fig.\textasciitilde{}\ref{fig:ecoregion} and
+Tab.~@ref(tab:comp\_forest)). Regarding the deforestation trend, we
 observed a progressive decrease of the deforestation rate after 1990
-from 205,000 ha/yr (1.6\%/yr) over the period 1973-1990 to 42,000
-ha/yr (0.4\%/yr) over the period 2000-2005
-(Tab.~\ref{tab:forest_cover}). Then from 2005, the deforestation rate
-has progressively increased and has more than doubled over the period
-2010-2014 (99,000 ha/yr, 1.1\%/yr) compared to 2000-2005
-(Tab.~\ref{tab:forest_cover}). The deforestation trend characterized
-by a progressive decrease of the deforestation rate over the period
-1990-2005 and a progressive increase of the deforestation after 2005
-is valid for all four ecoregions (Tab.~\ref{tab:comp_defor}), with the
-exception of the spiny forest domain for which the deforestation rate
-during the period 2010-2013 was lower than during 2005-2010
-(Tab.~\ref{tab:comp_defor}).
-
-\begin{table}[h]
-
-  \caption{\textbf{Evolution of forest cover and deforestation
-      rates from 1953 to 2014 in Madagascar}.}
-  \label{tab:forest_cover}
-
-  \begin{tabular}{@{}lrrrr@{}}
-    \tophline
-    Year & Forest & Unmapped & Annual defor. & Rate \\
-    & (Kha) & (Kha) & (Kha/yr) & (\%/yr) \\
-    \middlehline
-    1953 & 15,968 & 0 & - & - \\
-    1973 & 14,243 & 3,317 & 86 & 0.6 \\
-    1990 & 10,762 & 0 & 205 & 1.6 \\
-    2000 & 9,879 & 0 & 88 & 0.8 \\
-    2005 & 9,668 & 0 & 42 & 0.4 \\
-    2010 & 9,320 & 0 & 70 & 0.7 \\
-    2014 & 8,925 & 0 & 99 & 1.1 \\
-    \bottomhline
-  \end{tabular}
-
-  \belowtable{Areas are provided in thousands of hectares
-    (Kha). Forest map for the year 1973 has 3.3 Mha of unclassified
-    areas due to the presence of clouds on satellite images. As a
-    consequence, partial deforestation rates for the periods 1953-1973
-    and 1973-1990 are computed based on the available forest
-    extent. The last two columns indicate the annual deforested areas
-    and annual deforestation rates on the previous time-period
-    (e.g.~1953-1973 for year 1973, 1973-1990 for year 1990, etc.).}
-
-\end{table}
-
-\subsection{Comparison with previous forest-cover change studies
-in Madagascar}
-
-Forest-cover maps provided by previous studies over Madagascar were
-not exhaustive (unclassified areas) due to the presence of clouds on
-satellite images used to produce such maps. In \citet{Harper2007}, the
-maps of years 1990 and 2000 include 0.5 and 1.12 Mha of unknown cover
-type respectively. Proportions of unclassified areas are not reported
-in the two other existing studies at the national level by
-\citet{MEFT2009} and \citet{ONE2015}. With our approach, we produced
-wall to wall forest-cover change maps from 1990 to 2014 for the full
-territory of Madagascar (Tab.~\ref{tab:forest_cover}). This allowed us
-to produce more robust estimates of forest-cover and deforestation
-rates over this period. Our forest-cover estimates over the period
-1953-2013 (considering forest cover estimates at national level and by
-ecoregions for all the available dates) were well correlated
-(Pearson's correlation coefficient $=$ 0.99) to estimates from the
-three previous studies (Tab.~\ref{tab:comp_forest}) with a RMSE of
-300,000 ha (6\% of the mean forest cover of 4.8 Mha when considering
-all dates and forest types together). These small differences can be
-partly attributed to differences in ecoregion boundaries. Despite
-significant differences in deforestation estimates
-(Tab.~\ref{tab:comp_defor}), a similar deforestation trend was
-observed across studies with a decrease of deforestation rates over
-the period 1990-2005, followed by a progressive increase of the
-deforestation after 2005.
-
-\begin{table*}[h]
-
-  \caption{\textbf{Comparing Madagascar forest-cover estimates (in Kha)
-      with previous studies on the period 1953-2014}.}
-  \label{tab:comp_forest}
-
-  \begin{tabular}{@{}llrrrrrrrr@{}}
-    \tophline
-    Forest type & Source & 1953 & 1973 & 1990 & 2000 & 2005 & 2010 & 2013 &
-                                                                            2014 \\
-    \middlehline
-    Total & Harper2007 & 15,996 & 14,173 & 10,606 & 8,982 & - & - & - & - \\
-          & MEFT2009 & - & - & 10,650 & 9,678 & 9,413 & - & - & - \\
-                & ONE2015 & - & - & - & - & 9,451 & 8,977 & 8,486 & - \\
-                & this study & 15,968 & 14,243 & 10,762 & 9,879 & 9,668 & 9,320 & 9,051
-                                                                          & 8,925 \\
-    Moist & Harper2007 & 8,766 & 6,876 & 5,234 & 4,167 & - & - & - &
-                                                                     - \\
-                & MEFT2009 & - & - & 5,271 & 4,788 & 4,700 & - & - & - \\
-                & ONE2015 & - & - & - & - & 4,556 & 4,457 & 4,345 & - \\
-                & this study & 8,578 & 6,990 & 5,270 & 4,872 & 4,768 & 4,633 & 4,470 &
-                                                                                       4,410 \\
-    Dry & Harper2007 & 4,252 & 4,028 & 2,712 & 2,457 & - & - & - &
-                                                                   - \\
-                & MEFT2009 & - & - & 3,321 & 3,085 & 3,028 & - & - & - \\
-                & ONE2015 & - & - & - & - & 3,223 & 2,970 & 2,679 & - \\
-                & this study & 4,762 & 4,435 & 3,225 & 2,941 & 2,881 & 2,735 & 2,642 &
-                                                                                       2,596 \\
-    Spiny & Harper2007 & 2,978 & 3,030 & 2,420 & 2,132 & - & - & - &
-                                                                     - \\
-                & MEFT2009 & - & - & 2,124 & 1,872 & 1,757 & - & - & - \\
-                & ONE2015 & - & - & - & - & 1,682 & 1,559 & 1,467 & - \\
-                & this study & 2,463 & 2,583 & 2,055 & 1,858 & 1,811 & 1,744 & 1,731 &
-                                                                                       1,713 \\
-    Mangroves & Harper2007 & - & - & 240 & 226 & - & - & - &
-                                                             - \\
-                & MEFT2009 & - & - & - & - & - & - & - & - \\
-                & ONE2015 & - & - & - & - & 174 & 171 & 170 & - \\
-                & this study & 143 & 200 & 181 & 178 & 177 & 177 & 177 &
-                                                                         177 \\
-    \bottomhline
-  \end{tabular}
-
-  \belowtable{We compared our estimates of forest-cover with the
-    estimates from three previous studies \citep{Harper2007, MEFT2009,
-      ONE2015}. Areas are provided in thousands of hectares (Kha). We
-    obtained a Pearson's correlation coefficient of 0.99 between our
-    forest-cover estimates and forest-cover estimates from previous
-    studies. The increase in mangrove and spiny forest covers from
-    \emph{c.} 1953 to \emph{c.} 1973 in \citet{Harper2007} and our
-    study is most probably due to differences in forest definition and
-    mapping methods between the 1953 aerial-photography derived map
-    and the 1973 Landsat image derived map.}
-
-\end{table*}
-
-\begin{table*}[h]
-
-  \caption{\textbf{Comparing Madagascar annual deforestation rates
-      with previous studies on the period 1953-2014}.}
-  \label{tab:comp_defor}
-
-  \begin{tabular}{@{}llrrrrrr@{}}
-    \tophline
-    Forest type & Source & 1953-1973 & 1973-1990 & 1990-2000 & 2000-2005 &
-                                                                           2005-2010 & 2010-2013\\
-    \middlehline
-    Total & Harper2007 & 91 (0.3) & 200 (1.7) & 81 (0.9) & - & - &
-                                                                   - \\
-                & MEFT2009 & - & - & 97 (0.8) & 53 (0.5) & - & - \\
-                & ONE2015 & - & - & - & - & 95 (1.2) & 164 (1.5) \\
-                & this study & 86 (0.6) & 205 (1.6) & 88 (0.9) & 42 (0.4) & 70 (0.7) &
-                                                                                       90 (1.0) \\
-    Moist & Harper2007 & 94 (0.6) & 87 (1.7) & 32 (0.8) & - & - &
-                                                                  - \\
-                & MEFT2009 & - & - & 48 (0.8) & 17 (0.4) & - & - \\
-                & ONE2015 & - & - & - & - & 20 (0.5) & 37 (0.9) \\
-                & this study & 79 (1.0) & 101 (1.6) & 40 (0.8) & 21 (0.4) & 27 (0.6) &
-                                                                                       54 (1.2) \\
-    Dry & Harper2007 & 11 (0.2) & 77 (1.9) & 20 (0.7) & - & - &
-                                                                - \\
-                & MEFT2009 & - & - & 24 (0.7) & 11 (0.4) & - & - \\
-                & ONE2015 & - & - & - & - & 51 (1.8) & 97 (2.3) \\
-                & this study & 16 (0.4) & 71 (1.9) & 28 (0.9) & 12 (0.4) & 29 (1.0) & 31
-                                                                                      (1.1) \\
-    Spiny & Harper2007 & -3 (-0.1) & 36 (1.2) & 28 (1.2) & - & - &
-                                                                   - \\
-                & MEFT2009 & - & - & 25 (1.2) & 23 (1.2) & - & - \\
-                & ONE2015 & - & - & - & - & 25 (1.7) & 31 (1.7) \\
-                & this study & -6 (-0.2) & 31 (1.3) & 20 (1.0) & 9 (0.5) & 13 (0.7) & 4
-                                                                                      (0.3) \\
-    Mangroves & Harper2007 & - & - & 1 (0.2) & - & - & - \\
-                & MEFT2009 & - & - & - & - & - & - \\
-                & ONE2015 & - & - & - & - & 0 (0.3) & 0 (0.2) \\
-                & this study & -3 (-1.7) & 1 (0.6) & 0 (0.2) & 0 (0.0) & 0 (0.0) & 0
-                                                                                   (0.0) \\
-    \bottomhline
-  \end{tabular}
-
-  \belowtable{Annual deforestated areas (in Kha/yr) and annual
-    deforestation rates (second number in parenthesis, in \%/yr) are
-    provided. For deforestation rates in \%/yr, exact same numbers as
-    in scientific articles and reports from previous studies
-    \citep{Harper2007, MEFT2009, ONE2015} have been reported. The way
-    annual deforestation rates in \%/yr have been computed in these
-    previous studies can slightly differ from one study to another, but
-    estimates always correct for the potential presences of clouds on
-    satellite images and unclassified areas on forest maps. Annual
-    deforested areas in Kha/yr have been recomputed from forest-cover
-    estimates in Tab.~\ref{tab:comp_forest} (except for
-    \citet{Harper2007} for the periods 1973-1990 and 1990-2000 for
-    which annual deforested areas in Kha/yr were derived from numbers
-    reported in the original publication, see methods) and do not
-    correct for the potential presence of clouds.}
-
-\end{table*}
-
-\subsection{Evolution of forest fragmentation with time}
+from 205,000 ha/yr (1.6\%/yr) over the period 1973-1990 to 42,000 ha/yr
+(0.4\%/yr) over the period 2000-2005 (Tab.~@ref(tab:forest\_cover)).
+Then from 2005, the deforestation rate has progressively increased and
+has more than doubled over the period 2010-2014 (99,000 ha/yr, 1.1\%/yr)
+compared to 2000-2005 (Tab.~@ref(tab:forest\_cover)). The deforestation
+trend, characterized by a progressive decrease of the deforestation rate
+over the period 1990-2005 and a progressive increase of the
+deforestation after 2005, is valid for all four forest types except the
+spiny forest (Tab.~@ref(tab:comp\_defor)). For the spiny forest, the
+deforestation rate during the period 2010-2013 was lower than on the
+period 2005-2010 (Tab.~@ref(tab:comp\_defor)).
+
+\hypertarget{comparison-with-previous-forest-cover-change-studies-in-madagascar}{%
+\subsection{Comparison with previous forest cover change studies in
+Madagascar}\label{comparison-with-previous-forest-cover-change-studies-in-madagascar}}
+
+Forest cover maps provided by previous studies over Madagascar were not
+exhaustive (unclassified areas) due to the presence of clouds on
+satellite images used to produce such maps. In Harper \emph{et al.}
+(\protect\hyperlink{ref-Harper2007}{2007}), the maps of years 1990 and
+2000 include 0.5 and 1.12 Mha of unknown cover type respectively.
+Proportions of unclassified areas are not reported in the two other
+existing studies at the national level by MEFT, USAID, and CI
+(\protect\hyperlink{ref-MEFT2009}{2009}) and ONE, DGF, MNP, WCS, and Etc
+Terra (\protect\hyperlink{ref-ONE2015}{2015}). With our approach, we
+produced wall-to-wall forest cover change maps from 1990 to 2014 for the
+full territory of Madagascar (Tab.~@ref(tab:forest\_cover)). This
+allowed us to produce more robust estimates of forest cover and
+deforestation rates over this period. Our forest cover estimates over
+the period 1953-2013 (considering forest cover estimates at national
+level and by ecoregions for all the available dates) were well
+correlated (Pearson's correlation coefficient = 0.99) to estimates from
+the three previous studies (Tab.~@ref(tab:comp\_forest)) with a RMSE of
+300,000 ha (6\% of the mean forest cover of 4.8 Mha when considering all
+dates and forest types together). These small differences can be partly
+attributed to differences in ecoregion boundaries. Despite significant
+differences in deforestation estimates (Tab.~@ref(tab:comp\_defor)), a
+similar deforestation trend was observed across studies with a decrease
+of deforestation rates over the period 1990-2005, followed by a
+progressive increase of the deforestation after 2005.
+
+\hypertarget{evolution-of-forest-fragmentation-with-time}{%
+\subsection{Evolution of forest fragmentation with
+time}\label{evolution-of-forest-fragmentation-with-time}}
 
@@
 2014 in Madagascar. The mean distance to forest edge has decreased to
-\emph{c.} 300~m in 2014 while it was previously \emph{c.} 1.5~km in 1973
-(Fig.~\ref{fig:dist_edge}). Moreover, a large proportion (73\%) of
-the forest was located at a distance greater than 100 m in 1973, while
-almost half of the forest (46\%) was at a distance lower than 100 m
-from forest edge in 2014 (Fig.~\ref{fig:dist_edge}). The percentage of
-forest that can be considered intact in Madagascar has continuously
-decreased since 1953. The percentage of forest belonging to the
-``interior'' category (most intact forests) has fallen from 68\% in
-1973 to 50\% in 2014. In 2014, more than 16\% of the forest belonged
-to the ``patch'' and ``transitional'' categories (isolated small
-forest patches) compared to 9\% in 1973 (Tab.~\ref{tab:frag}).
-
-\begin{figure}[h]
-  \centering
-
-  \includegraphics[width=8.3cm]{dist.png}
-
-  \caption{\textbf{Evolution of the distance to forest edge from 1953
-      to 2014 in Madagascar.} Black dots represent the mean distance
-    to forest edge for each year. Vertical dashed segments represent
-    the 90\% quantiles (5\% and 95\%) of the distance to forest
-    edge. Horizontal dashed grey line indicates a distance to forest
-    edge of 100~m.  Numbers at the bottom of each vertical segments are
-    the percentage of forest at a distance to forest edge lower than
-    100~m for each year.}
-
-  \label{fig:dist_edge}
-
-\end{figure}
-
-\section{Discussion}
-\label{discussion}
-
-\subsection{Benefits of the combined use of recent global annual
-  tree cover loss data with historical national forest-cover maps}
+\emph{c.}~300~m in 2014 while it was previously \emph{c.}~1.5~km in 1973
+(Fig.\textasciitilde{}\ref{fig:dist_edge}). Moreover, a large proportion
+(73\%) of the forest was located at a distance greater than 100~m in
+1973, while almost half of the forest (46\%) is now at a distance lower
+than 100\textasciitilde{}m from forest edge in 2014
+(Fig.\textasciitilde{}\ref{fig:dist_edge}). The percentage of lowly
+fragmented forest in Madagascar has continuously decreased since 1953.
+The percentage of forest belonging to the lowly fragmented category has
+fallen from 57\% in 1973 to 44\% in 2014. In 2014, 22\% of the forest
+belonged to the two highest fragmented forest classes (less than 40\% of
+forest cover in the neighborhood) while 15\% of the forest belonged to
+these two classes in 1973 (Tab.~\ref{tab:frag}).
+
+\hypertarget{discussion}{%
+\section{Discussion}\label{discussion}}
+
+\hypertarget{advantages-of-combining-recent-global-annual-tree-cover-loss-data-with-historical-national-forest-cover-maps}{%
+\subsection{Advantages of combining recent global annual tree cover loss
+data with historical national forest cover
+maps}\label{advantages-of-combining-recent-global-annual-tree-cover-loss-data-with-historical-national-forest-cover-maps}}
 
 In this study, we combined recent (2001-2014) global annual tree cover
-loss data \citep{Hansen2013} with historical (1953-2000) national
-forest-cover maps \citep{Harper2007} to look at six decades
-(1953-2014) of deforestation and forest fragmentation in
-Madagascar. We produced annual forest-cover maps at 30~m resolution
-covering Madagascar for the period 2000 to 2014. Our study extends the
-forest-cover monitoring on a six decades period (from 1953 to 2014)
-while harmonizing the data from previous studies \citep{Harper2007,
-  MEFT2009, ONE2015}. We propose a generic approach to solve the
-problem of forest definition which is needed to transform the 2000
-global tree cover dataset from \citet{Hansen2013} into a
-forest/non-forest map \citep{Tropek2014}. We propose to use a
-historical national forest-cover map, based on a national forest
-definition, as a forest cover mask. This approach could be easily
-extended to other regions or countries for which an accurate
-forest-cover map is available at any date within the period 2000-2014,
+loss data (Hansen \emph{et al.}
+\protect\hyperlink{ref-Hansen2013}{2013}) with historical (1953-2000)
+national forest cover maps (Harper \emph{et al.}
+\protect\hyperlink{ref-Harper2007}{2007}) to look at six decades
+(1953-2014) of deforestation and forest fragmentation in Madagascar. We
+produced annual forest cover maps at 30~m resolution covering Madagascar
+for the period 2000 to 2014. Our study extends the forest cover
+monitoring on a six decades period (from 1953 to 2014) while harmonizing
+the data from previous studies (Harper \emph{et al.}
+\protect\hyperlink{ref-Harper2007}{2007}; MEFT, USAID, and CI
+\protect\hyperlink{ref-MEFT2009}{2009}; ONE, DGF, MNP, WCS, and Etc
+Terra \protect\hyperlink{ref-ONE2015}{2015}). We propose a generic
+approach to solve the problem of forest definition which is needed to
+transform the 2000 global tree cover dataset from Hansen \emph{et al.}
+(\protect\hyperlink{ref-Hansen2013}{2013}) into a forest/non-forest map
+(Tropek \emph{et al.} \protect\hyperlink{ref-Tropek2014}{2014}). We
+propose to use a historical national forest cover map, based on a
+national forest definition, as a forest cover mask. This approach could
+be easily extended to other regions or countries for which an accurate
+forest cover map is available at any date within the period 2000-2014,
 but preferably at the beginning of the period to profit from the full
 record and derive long-term estimates of deforestation. Moreover, this
-approach can be repeated in the future if and when the global tree
-cover product is updated. We have made the \R{}/GRASS code used for this
+approach can be repeated in the future if and when the global tree cover
+product is updated. We have made the \texttt{R/GRASS} code used for this
 study freely available in a GitHub repository (see Data availability
-statement) to facilitate application to other study areas or repeat
-the analysis in the future for Madagascar.
-
-The accuracy of the derived forest-cover change maps depends directly
-on the accuracies of the historical forest-cover maps and the tree
-cover loss dataset. Using visual-interpretation of aerial images in
-342 areas distributed among all forest types, \citet{Harper2007}
-estimated an overall 89.5\% accuracy in identifying forest/non-forest
-classes for the year 2000. The accuracy assessment of the tree cover
-loss dataset for the tropical biome reported 13\% of false positives
-and 16.9\% of false negatives (see Tab. S5 in
-\citet{Hansen2013}). These numbers rise at 20.7\% and 20.6\%
-respectively for the subtropical biome. In the subtropical biome, the
-lower density tree cover canopy makes it difficult to detect change
-from tree cover to bare ground. For six countries in Central Africa,
-with a majority of moist dense forest, \citet{Verhegghen2016} have
-compared deforestation estimates derived from the global tree cover
-loss dataset \citep{Hansen2013} with results derived from
+statement) to facilitate application to other study areas or repeat the
+analysis in the future for Madagascar.
+
+The accuracy of the derived forest cover change maps depends directly on
+the accuracies of the historical forest cover maps and the tree cover
+loss dataset. Using visual-interpretation of aerial images in 342 areas
+distributed among all forest types, Harper \emph{et al.}
+(\protect\hyperlink{ref-Harper2007}{2007}) estimated an overall 89.5\%
+accuracy in identifying forest/non-forest classes for the year 2000. The
+accuracy assessment of the tree cover loss dataset for the tropical
+biome reported 13\% of false positives and 16.9\% of false negatives
+(see Tab. S5 in Hansen \emph{et al.}
+(\protect\hyperlink{ref-Hansen2013}{2013})). These numbers rise at
+20.7\% and 20.6\% respectively for the subtropical biome. In the
+subtropical biome, the lower density tree cover canopy makes it
+difficult to detect change from tree cover to bare ground. For six
+countries in Central Africa, with a majority of moist dense forest,
+Verhegghen \emph{et al.} (\protect\hyperlink{ref-Verhegghen2016}{2016})
+have compared deforestation estimates derived from the global tree cover
+loss dataset (Hansen \emph{et al.}
+\protect\hyperlink{ref-Hansen2013}{2013}) with results derived from
 semi-automated supervised classification of Landsat satellite images
-\citep{Achard2014} and they found a good agreement between the two
-sets of estimates. Therefore, our forest-cover change maps after 2000
-might be more accurate for the dense moist forest than for the dry and
-spiny forest. In another study assessing the accuracy of the tree
-cover loss product accross the tropics \citep{Tyukavina2015}, authors
-reported 4\% of false positives and 48\% of false negatives in
-Sub-Saharian Africa. They showed that 85\% of missing loss occured on
-the edges of other loss patches. This means that tree cover loss might
-be underestimated in Sub-Saharian Africa, probably due to the
-prevalence of small-scale disturbance which is hard to map at 30~m,
-but that areas of large-scale deforestation are well identified and
-spatial variability of the deforestation is well represented. A proper
-accuracy assessment of our forest-cover change maps should be
-performed to better estimate the uncertainty surrounding our
-forest-cover change estimates in Madagascar from year 2000
-\citep{Olofsson2013,Olofsson2014}. Despite this limitation, we have
-shown that the deforestation trend we observed for Madagascar, with a
-doubling deforestation on the period 2010-2014 compared to 2000-2005,
-was consistent with the other studies at the national scale
-\citep{ONE2015, MEFT2009}.
-
-Consistent with \citet{Harper2007}, we did not consider potential
-forest regrowth in Madagascar (although \citet{Hansen2013} provided a
-tree cover gains layer for the period 2001-2013) for several
-reasons. First, the tree gain layer of \citet{Hansen2013} includes and
-catches more easily tree plantations than natural forest regrowth
-\citep{Tropek2014}. Second, there is little evidence of natural forest
-regeneration in Madagascar \citep{Grouzis2001, Harper2007}. This can
-be explained by several ecological processes following burning
-practice such as soil erosion \citep{Grinand2017} and reduced seed
-bank due to fire and soil loss \citep{Grouzis2001}. Moreover, in areas
-where forest regeneration is ecologically possible, young forest
-regrowth are more easily re-burnt for agriculture and pasture. Third,
-young secondary forests provide more limited ecosystem services
-compared to old-growth natural forests in terms of biodiversity and
-carbon storage.
-
-\subsection{Dynamics of forest-cover in Madagascar from 1953 to 2014}
-
-We estimated that natural forests in Madagascar cover 8.9 Mha in 2014
-(corresponding to 15\% of the country) and that Madagascar has lost
-44\% of its natural forest since 1953 (37\% since 1973). There is
-ongoing scientific debate about the extent of the ``original'' forest
-cover in Madagascar, and the extent to which humans have altered the
-natural forest landscapes since their large-scale settlement around
-800 CE \citep{Burns2016, Cox2012}. Early French naturalists stated
-that the full island was originally covered by forest
-\citep{Humbert1927, Perrier1921}, leading to the common statement that
-90\% of the natural forests have disappeared since the arrival of
-humans on the island \citep{Kull2000}. More recent studies
+(Achard \emph{et al.} \protect\hyperlink{ref-Achard2014}{2014}) and they
+found a good agreement between the two sets of estimates. Therefore, our
+forest cover change maps after 2000 might be more accurate for the dense
+moist forest than for the dry and spiny forest. In another study
+assessing the accuracy of the tree cover loss product accross the
+tropics (Tyukavina \emph{et al.}
+\protect\hyperlink{ref-Tyukavina2015}{2015}), authors reported 4\% of
+false positives and 48\% of false negatives in Sub-Saharian Africa. They
+showed that 85\% of missing loss occured on the edges of other loss
+patches. This means that tree cover loss might be underestimated in
+Sub-Saharian Africa, probably due to the prevalence of small-scale
+disturbance which is hard to map at 30~m, but that areas of large-scale
+deforestation are well identified and spatial variability of the
+deforestation is well represented. A proper accuracy assessment of our
+forest cover change maps should be performed to better estimate the
+uncertainty surrounding our forest cover change estimates in Madagascar
+from year 2000 (Olofsson \emph{et al.}
+\protect\hyperlink{ref-Olofsson2013}{2013},
+\protect\hyperlink{ref-Olofsson2014}{2014}). Despite this limitation, we
+have shown that the deforestation trend we observed for Madagascar, with
+a doubling deforestation on the period 2010-2014 compared to 2000-2005,
+was consistent with the other studies at the national scale (MEFT,
+USAID, and CI \protect\hyperlink{ref-MEFT2009}{2009}; ONE, DGF, MNP,
+WCS, and Etc Terra \protect\hyperlink{ref-ONE2015}{2015}).
+
+Consistent with Harper \emph{et al.}
+(\protect\hyperlink{ref-Harper2007}{2007}), we did not consider
+potential forest regrowth in Madagascar (although Hansen \emph{et al.}
+(\protect\hyperlink{ref-Hansen2013}{2013}) provided a tree cover gains
+layer for the period 2001-2013) for several reasons. First, the tree
+gain layer of Hansen \emph{et al.}
+(\protect\hyperlink{ref-Hansen2013}{2013}) includes and catches more
+easily tree plantations than natural forest regrowth (Tropek \emph{et
+al.} \protect\hyperlink{ref-Tropek2014}{2014}). Second, there is little
+evidence of natural forest regeneration in Madagascar (Grouzis \emph{et
+al.} \protect\hyperlink{ref-Grouzis2001}{2001}; Harper \emph{et al.}
+\protect\hyperlink{ref-Harper2007}{2007}). This can be explained by
+several ecological processes following burning practice such as soil
+erosion (Grinand \emph{et al.}
+\protect\hyperlink{ref-Grinand2017}{2017}) and reduced seed bank due to
+fire and soil loss (Grouzis \emph{et al.}
+\protect\hyperlink{ref-Grouzis2001}{2001}). Moreover, in areas where
+forest regeneration is ecologically possible, young forest regrowth are
+more easily re-burnt for agriculture and pasture. Third, young secondary
+forests provide more limited ecosystem services compared to old-growth
+natural forests in terms of biodiversity and carbon storage.
+
+\hypertarget{natural-forest-cover-change-in-madagascar-from-1953-to-2014}{%
+\subsection{Natural forest cover change in Madagascar from 1953 to
+2014}\label{natural-forest-cover-change-in-madagascar-from-1953-to-2014}}
+
+We estimated that natural forest in Madagascar covers 8.9 Mha in 2014
+(corresponding to 15\% of the country) and that Madagascar has lost 44\%
+of its natural forest since 1953 (37\% since 1973). There is ongoing
+scientific debate about the extent of the ``original'' forest cover in
+Madagascar, and the extent to which humans have altered the natural
+forest landscapes since their large-scale settlement around 800 CE (Cox
+\emph{et al.} \protect\hyperlink{ref-Cox2012}{2012}; Burns \emph{et al.}
+\protect\hyperlink{ref-Burns2016}{2016}). Early French naturalists
+stated that the full island was originally covered by forest (Perrier de
+La Bâthie \protect\hyperlink{ref-Perrier1921}{1921}; Humbert
+\protect\hyperlink{ref-Humbert1927}{1927}), leading to the common
+statement that 90\% of the natural forests have disappeared since the
+arrival of humans on the island (Kull
+\protect\hyperlink{ref-Kull2000}{2000}). More recent studies
 counter-balanced that point of view saying that extensive areas of
@@
 determined by climate, natural grazing and other natural factors
-\citep{Vorontsova2017, Virah-Sawmy2009}. Other authors have questioned
-the entire narrative of extensive alteration of the landscape by early
-human activity which, through legislation, has severe consequences on
-local people \citep{Klein2002, Kull2000}. Whatever the original
-proportion of natural forests and grasslands in Madagascar, our
-results demonstrate that human activities since the 1950s have
-profoundly impacted the natural tropical forests and that conservation
-and development programs in Madagascar have failed to stop
-deforestation in the recent years. Deforestation has strong
-consequences on biodiversity and carbon emissions in
-Madagascar. Around 90\% of Madagascar's species are forest dependent
-\citep{Allnutt2008, Goodman2005} and \citet{Allnutt2008} estimated
-that deforestation between 1953 and 2000 led to an extinction of 9\%
-of the species. The additional deforestation we observed over the
-period 2000-2014 (around 1Mha of natural forest) worsen this
-result. Regarding carbon emissions, using the 2010 aboveground forest
-carbon map by \citet{Vieilledent2016}, we estimated that deforestation
-on the period 2010-2014 has led to 40.2 Mt C of carbon emissions in
-the atmosphere (10 Mt C /yr) and that the remaining aboveground forest
-carbon stock in 2014 is 832.8 Mt C. Associated to deforestation, we
-showed that the remaining forests of Madagascar are highly fragmented
-with 46\% of the forest being at less than 100~m of the forest
-edge. Small forest fragments do not allow to maintain viable
+(Virah-Sawmy \protect\hyperlink{ref-Virah-Sawmy2009}{2009}; Vorontsova
+\emph{et al.} \protect\hyperlink{ref-Vorontsova2017}{2016}). Other
+authors have questioned the entire narrative of extensive alteration of
+the landscape by early human activity which, through legislation, has
+severe consequences on local people (Kull
+\protect\hyperlink{ref-Kull2000}{2000}; Klein
+\protect\hyperlink{ref-Klein2002}{2002}). Whatever the original
+proportion of natural forests and grasslands in Madagascar, our results
+demonstrate that human activities since the 1950s have profoundly
+impacted the natural tropical forests and that conservation and
+development programs in Madagascar have failed to stop deforestation in
+the recent years. Deforestation has strong consequences on biodiversity
+and carbon emissions in Madagascar. Around 90\% of Madagascar's species
+are forest dependent (Goodman \& Benstead
+\protect\hyperlink{ref-Goodman2005}{2005}; Allnutt \emph{et al.}
+\protect\hyperlink{ref-Allnutt2008}{2008}) and Allnutt \emph{et al.}
+(\protect\hyperlink{ref-Allnutt2008}{2008}) estimated that deforestation
+between 1953 and 2000 led to an extinction of 9\% of the species. The
+additional deforestation we observed over the period 2000-2014 (around
+1Mha of natural forest) worsen this result. Regarding carbon emissions,
+using the 2010 aboveground forest carbon map by Vieilledent \emph{et
+al.} (\protect\hyperlink{ref-Vieilledent2016}{2016}), we estimated that
+deforestation on the period 2010-2014 has led to 40.2 Mt C of carbon
+emissions in the atmosphere (10 Mt C /yr) and that the remaining
+aboveground forest carbon stock in 2014 is 832.8 Mt C. Associated to
+deforestation, we showed that the remaining forests of Madagascar are
+highly fragmented with 46\% of the forest being at less than 100~m of
+the forest edge. Small forest fragments do not allow to maintain viable
 populations and ``edge effects'' at forest/non-forest interfaces have
-impacts on both carbon emissions \citep{Brinck2017} and biodiversity
-loss \citep{Gibson2013, Murcia1995}.
-
-\begin{table}[h]
-
-  \caption{\textbf{Evolution of the forest fragmentation from 1953 to 2014
-      in Madagascar}.}
-  \label{tab:frag}
-
-  \begin{tabular}{@{}lrrrrrr@{}}
-    \tophline
-    Year & Forest & patch & transit. & edge & perfor. & interior \\
-    & (Kha) & (\%) & (\%) & (\%) & (\%) & (\%) \\
-    \middlehline
-    1953 & 15,963 & 0 & 1 & 4 & 1 & 94 \\
-    1973 & 14,228 & 2 & 7 & 20 & 3 & 68 \\
-    1990 & 10,750 & 3 & 8 & 21 & 4 & 64 \\
-    2000 & 9,866 & 3 & 8 & 22 & 4 & 62 \\
-    2005 & 9,660 & 4 & 9 & 23 & 6 & 59 \\
-    2010 & 9,307 & 4 & 10 & 23 & 9 & 55 \\
-    2014 & 8,911 & 5 & 11 & 23 & 11 & 50 \\
-    \bottomhline
-  \end{tabular}
-
-  \belowtable{Five categories of fragmentation were identified from
-    the amount of forest and its occurrence as adjacent forest pixels:
-    ``interior'', ``perforated'', ``edge'', ``transitional'', and
-    ``patch'' \citep{Riitters2000}. We used a moving window of 7x7
-    pixels (4.4~ha). Using this window size, forest edge had a width
-    of about 90~m. The ``interior'' category can be interpreted as the
-    most intact forest. The ``patch'' and ``transitional'' categories
-    correspond to isolated small forest patches. Forest areas are
-    provided in thousands of hectares (Kha).}
-
-\end{table}
-
+impacts on both carbon emissions (Brinck \emph{et al.}
+\protect\hyperlink{ref-Brinck2017}{2017}) and biodiversity loss (Murcia
+\protect\hyperlink{ref-Murcia1995}{1995}; Gibson \emph{et al.}
+\protect\hyperlink{ref-Gibson2013}{2013}).
+
+\hypertarget{deforestation-trend-and-impacts-on-conservation-and-development-policies}{%
 \subsection{Deforestation trend and impacts on conservation and
-  development policies}
+development
+policies}\label{deforestation-trend-and-impacts-on-conservation-and-development-policies}}
 
 In our study, we have shown that the progressive decrease of the
-deforestation rate on the period 1990-2005 was followed by a
-continuous increase in the deforestation rate on the period
-2005-2014. In particular, we showed that deforestation rate has more
-than doubled on the period 2010-2014 compared to 2000-2005. Our
-results are confirmed by previous studies \citep{Harper2007, MEFT2009,
-  ONE2015} despite differences in the methodologies regarding (i)
-forest definition (associated to independent visual interpretations of
-observation polygons to train the classifier), (ii) classification
-algorithms, (iii) deforestation rate computation method, and (iv)
-correction for the presence of clouds. Our deforestation rate
-estimates from 1990 to 2014 have been computed from wall to wall maps
-at 30~m resolution and can be considered more accurate in comparison
-with estimates from these previous studies. Our forest-cover and
-deforestation rate estimates can be used as source of information for
-the next FAO Forest Resources Assessment \citep{Keenan2015}. Current
-rates of deforestation can also be used to build reference scenarios
-for deforestation in Madagascar and contribute to the implementation
-of deforestation mitigation activities in the framework of REDD+
-\citep{Olander2008}.
+deforestation rate on the period 1990-2005 was followed by a continuous
+increase in the deforestation rate on the period 2005-2014. In
+particular, we showed that deforestation rate has more than doubled on
+the period 2010-2014 compared to 2000-2005. Our results are confirmed by
+previous studies (Harper \emph{et al.}
+\protect\hyperlink{ref-Harper2007}{2007}; MEFT, USAID, and CI
+\protect\hyperlink{ref-MEFT2009}{2009}; ONE, DGF, MNP, WCS, and Etc
+Terra \protect\hyperlink{ref-ONE2015}{2015}) despite differences in the
+methodologies regarding (i) forest definition (associated to independent
+visual interpretations of observation polygons to train the classifier),
+(ii) classification algorithms, (iii) deforestation rate computation
+method, and (iv) correction for the presence of clouds. Our
+deforestation rate estimates from 1990 to 2014 have been computed from
+wall-to-wall maps at 30~m resolution and can be considered more accurate
+in comparison with estimates from these previous studies. Our forest
+cover and deforestation rate estimates can be used as source of
+information for the next FAO Forest Resources Assessment (Keenan
+\emph{et al.} \protect\hyperlink{ref-Keenan2015}{2015}). Current rates
+of deforestation can also be used to build reference scenarios for
+deforestation in Madagascar and contribute to the implementation of
+deforestation mitigation activities in the framework of REDD+ (Olander
+\emph{et al.} \protect\hyperlink{ref-Olander2008}{2008}).
 
 The increase of deforestation rates after 2005 can be explained by
-population growth and political instability in the country. Nearly
-90\% of Madagascar's population relies on biomass for their daily
-energy needs \citep{Minten2013} and the link between population size
-and deforestation has previously been demonstrated in Madagascar
-\citep{Vieilledent2013, Gorenflo2011}. With a mean demographic growth
-rate of about 2.8\%/yr and a population which has increased from 16 to
-24 million people on the period 2000-2015 \citep{UN2015}, the
-increasing demand in wood-fuel and space for agriculture is likely to
-explain the increase in deforestation rates. The political crisis of
-2009 \citep{Ploch2012}, followed by several years of political
-instability and weak governance could also explain the increase in the
-deforestation rate observed on the period 2005-2014
-\citep{Smith2003}. These results show that despite the conservation
-policy in Madagascar \citep{Freudenberger2010}, deforestation has
-dramatically increased at the national level since 2005. Results of
-this study, including recent spatially explicit forest-cover change
-maps and forest-cover estimates, should help implement new
-conservation strategies to save Madagascar natural tropical forests
-and their unique biodiversity.
-
-%% The following commands are for the statements about the availability of data sets and/or software code corresponding to the manuscript.
-%% It is strongly recommended to make use of these sections in case data sets and/or software code have been part of your research the article is based on.
-
-%%\codeavailability{TEXT} %% use this section when having only software code available
-
-%%\dataavailability{TEXT} %% use this section when having only data sets available
-
-\codedataavailability{All the data and codes used for this study are
-  made publicly available in the \texttt{deforestmap} GitHub
-  repository (\url{https://github.com/ghislainv/deforestmap.git}). The
-  results are fully reproducible running the \R{} script
-  \texttt{deforestmap.R} located inside the \texttt{deforestmap}
-  repository.} %% use this section when having data sets and software code available
-
-%\appendix
-%\section{}    %% Appendix A
-
-%\subsection{}     %% Appendix A1, A2, etc.
-
-%\noappendix       %% use this to mark the end of the appendix section
-
-%% Regarding figures and tables in appendices, the following two options are possible depending on your general handling of figures and tables in the manuscript environment:
-
-%% Option 1: If you sorted all figures and tables into the sections of the text, please also sort the appendix figures and appendix tables into the respective appendix sections.
-%% They will be correctly named automatically.
-
-%% Option 2: If you put all figures after the reference list, please insert appendix tables and figures after the normal tables and figures.
-%% To rename them correctly to A1, A2, etc., please add the following commands in front of them:
-
-%\appendixfigures  %% needs to be added in front of appendix figures
-
-%\appendixtables   %% needs to be added in front of appendix tables
-
-%% Please add \clearpage between each table and/or figure. Further guidelines on figures and tables can be found below.
-
-\authorcontribution{All authors conceived the ideas and designed methodology; GV analysed
-the data and wrote the \R{}/GRASS script; GV drafted the manuscript. All
-authors contributed critically to the drafts and gave final approval for
-publication.} %% optional section
-
-\competinginterests{Authors declare that they have no conflict of interest.}
-%% this section is mandatory even if you declare that no competing interests are present
-
-%% \disclaimer{TEXT} %% optional section
-
-\begin{acknowledgements}
-
-  Authors thank Jean-François Bastin for useful comments on a previous
-  version of the manuscript. This study is part of the Cirad's
-  BioSceneMada project (\url{https://bioscenemada.cirad.fr}) and the
-  Joint Research Center's ReCaREDD project
+population growth and political instability in the country. Nearly 90\%
+of Madagascar's population relies on biomass for their daily energy
+needs (Minten, Sander \& Stifel
+\protect\hyperlink{ref-Minten2013}{2013}) and the link between
+population size and deforestation has previously been demonstrated in
+Madagascar (Gorenflo \emph{et al.}
+\protect\hyperlink{ref-Gorenflo2011}{2011}; Vieilledent, Grinand \&
+Vaudry \protect\hyperlink{ref-Vieilledent2013}{2013}). With a mean
+demographic growth rate of about 2.8\%/yr and a population which has
+increased from 16 to 24 million people on the period 2000-2015 (United
+Nations \protect\hyperlink{ref-UN2015}{2015}), the increasing demand in
+wood-fuel and space for agriculture is likely to explain the increase in
+deforestation rates. The political crisis of 2009 (Ploch \& Cook
+\protect\hyperlink{ref-Ploch2012}{2012}), followed by several years of
+political instability and weak governance could also explain the
+increase in the deforestation rate observed on the period 2005-2014
+(Smith \emph{et al.} \protect\hyperlink{ref-Smith2003}{2003}). These
+results show that despite the conservation policy in Madagascar
+(Freudenberger \protect\hyperlink{ref-Freudenberger2010}{2010}),
+deforestation has dramatically increased at the national level since
+2005. Results of this study, including recent spatially explicit forest
+cover change maps and forest cover estimates, should help implement new
+conservation strategies to save Madagascar natural tropical forests and
+their unique biodiversity.
+
+\hypertarget{authors-contribution}{%
+\section{Author's contribution}\label{authors-contribution}}
+
+All authors conceived the ideas and designed methodology; GV analysed
+the data and wrote the \texttt{R/GRASS} script; GV drafted the
+manuscript. All authors contributed critically to the drafts and gave
+final approval for publication.
+
+\hypertarget{acknowledgements}{%
+\section{Acknowledgements}\label{acknowledgements}}
+
+The authors thank Jean-François Bastin for useful comments on a previous
+version of the manuscript and Peter Vogt for useful advices on which
+metric to use to estimate forest fragmentation. This study is part of
+the Cirad's BioSceneMada project (\url{https://bioscenemada.cirad.fr})
+and the Joint Research Center's ReCaREDD project
 (\url{http://forobs.jrc.ec.europa.eu/recaredd}). The BioSceneMada
 project is funded by FRB (Fondation pour la Recherche sur la
-  Biodiversité) and the FFEM (Fond Français pour l'Environnement
-  Mondial) under the project agreement AAP-SCEN-2013 I. The ReCaREDD
-  project is funded by the European Commission.
-
-\end{acknowledgements}
-
-%% REFERENCES
-
-%% The reference list is compiled as follows:
-
-% \begin{thebibliography}{}
-
-% \bibitem[AUTHOR(YEAR)]{LABEL}
-% REFERENCE 1
-
-% \bibitem[AUTHOR(YEAR)]{LABEL}
-% REFERENCE 2
-
-% \end{thebibliography}
-
-%% Since the Copernicus LaTeX package includes the BibTeX style file copernicus.bst,
-%% authors experienced with BibTeX only have to include the following two lines:
-%%
-\bibliographystyle{copernicus}
-\bibliography{biblio}
-%%
-%% URLs and DOIs can be entered in your BibTeX file as:
-%%
-%% URL = {http://www.xyz.org/~jones/idx_g.htm}
-%% DOI = {10.5194/xyz}
-
-%% LITERATURE CITATIONS
-%%
-%% command                        & example result
-%% \citet{jones90}|               & Jones et al. (1990)
-%% \citep{jones90}|               & (Jones et al., 1990)
-%% \citep{jones90,jones93}|       & (Jones et al., 1990, 1993)
-%% \citep[p.~32]{jones90}|        & (Jones et al., 1990, p.~32)
-%% \citep[e.g.,][]{jones90}|      & (e.g., Jones et al., 1990)
-%% \citep[e.g.,][p.~32]{jones90}| & (e.g., Jones et al., 1990, p.~32)
-%% \citeauthor{jones90}|          & Jones et al.
-%% \citeyear{jones90}|            & 1990
-
-%% FIGURES
-
-%% When figures and tables are placed at the end of the MS (article in one-column style), please add \clearpage
-%% between bibliography and first table and/or figure as well as between each table and/or figure.
-
-%% ONE-COLUMN FIGURES
-
-%%f
-%\begin{figure}[t]
-%\includegraphics[width=8.3cm]{FILE NAME}
-%\caption{TEXT}
-%\end{figure}
-%
-%%% TWO-COLUMN FIGURES
-%
-%%f
-%\begin{figure*}[t]
-%\includegraphics[width=12cm]{FILE NAME}
-%\caption{TEXT}
-%\end{figure*}
-%
-%
-
-%%% TABLES
-%%%
-%%% The different columns must be seperated with a & command and should
-%%% end with \\ to identify the column brake.
-%
-%%% ONE-COLUMN TABLE
-%
-%%t
-%\begin{table}[t]
-%\caption{TEXT}
-%\begin{tabular}{column = lcr}
-%\tophline
-%
-%\middlehline
-%
-%\bottomhline
-%\end{tabular}
-%\belowtable{} % Table Footnotes
-%\end{table}
-
-%%% TWO-COLUMN TABLE
-%
-%%t
-%\begin{table*}[t]
-%\caption{TEXT}
-%\begin{tabular}{column = lcr}
-%\tophline
-%
-%\middlehline
-%
-%\bottomhline
-%\end{tabular}
-%\belowtable{} % Table Footnotes
-%\end{table*}
-
-%%% MATHEMATICAL EXPRESSIONS
-%
-%%% All papers typeset by Copernicus Publications follow the math typesetting regulations
-%%% given by the IUPAC Green Book (IUPAC: Quantities, Units and Symbols in Physical Chemistry,
-%%% 2nd Edn., Blackwell Science, available at: http://old.iupac.org/publications/books/gbook/green_book_2ed.pdf, 1993).
-%%%
-%%% Physical quantities/variables are typeset in italic font (t for time, T for Temperature)
-%%% Indices which are not defined are typeset in italic font (x, y, z, a, b, c)
-%%% Items/objects which are defined are typeset in roman font (Car A, Car B)
-%%% Descriptions/specifications which are defined by itself are typeset in roman font (abs, rel, ref, tot, net, ice)
-%%% Abbreviations from 2 letters are typeset in roman font (RH, LAI)
-%%% Vectors are identified in bold italic font using \vec{x}
-%%% Matrices are identified in bold roman font
-%%% Multiplication signs are typeset using the LaTeX commands \times (for vector products, grids, and exponential notations) or \cdot
-%%% The character * should not be applied as mutliplication sign
-%
-%
-%%% EQUATIONS
-%
-%%% Single-row equation
-%
-%\begin{equation}
-%
-%\end{equation}
-%
-%%% Multiline equation
-%
-%\begin{align}
-%& 3 + 5 = 8\\
-%& 3 + 5 = 8\\
-%& 3 + 5 = 8
-%\end{align}
-%
-%
-%%% MATRICES
-%
-%\begin{matrix}
-%x & y & z\\
-%x & y & z\\
-%x & y & z\\
-%\end{matrix}
-%
-%
-%%% ALGORITHM
-%
-%\begin{algorithm}
-%\caption{?}
-%\label{a1}
-%\begin{algorithmic}
-%?
-%\end{algorithmic}
-%\end{algorithm}
-%
-%
-%%% CHEMICAL FORMULAS AND REACTIONS
-%
-%%% For formulas embedded in the text, please use \chem{}
-%
-%%% The reaction environment creates labels including the letter R, i.e. (R1), (R2), etc.
-%
-%\begin{reaction}
-%%% \rightarrow should be used for normal (one-way) chemical reactions
-%%% \rightleftharpoons should be used for equilibria
-%%% \leftrightarrow should be used for resonance structures
-%\end{reaction}
-%
-%
-%%% PHYSICAL UNITS
-%%%
-%%% Please use \unit{} and apply the exponential notation
+Biodiversité) and the FFEM (Fond Français pour l'Environnement Mondial)
+under the project agreement AAP-SCEN-2013 I. The ReCaREDD project is
+funded by the European Commission. The authors declare that there are no
+conflicts of interest related to this article.
+
+\hypertarget{data-accessibility}{%
+\section{Data accessibility}\label{data-accessibility}}
+
+All the data and the script used for this study have been made
+permanently and publicly available on the Zenodo research data
+repository so that the results are entirely reproducible:
+
+\begin{itemize}
+\tightlist
+\item
+  Input data: \url{https://doi.org/10.5281/zenodo.1118955}
+\item
+  Script: \url{https://doi.org/10.5281/zenodo.1118484}
+\item
+  Output data: \url{https://doi.org/10.5281/zenodo.1145785}
+\end{itemize}
+
+\hypertarget{tables}{%
+\section{Tables}\label{tables}}
+
+\textbackslash{}begin\{table\}
+
+\textbackslash{}caption\{(\#tab:forest\_cover)Test\} \centering
+
+\begin{tabular}[t]{lrrrr}
+\toprule
+Year & Forest (Kha) & Unmap (Kha) & Annual defor. (Kha/yr) & Rate (\%/yr)\\
+\midrule
+1953 & 15,968 & 0 & - & -\\
+1973 & 14,243 & 3,317 & 86 & 0.6\\
+1990 & 10,762 & 0 & 205 & 1.6\\
+2000 & 9,879 & 0 & 88 & 0.8\\
+2005 & 9,668 & 0 & 42 & 0.4\\
+\addlinespace
+2010 & 9,320 & 0 & 70 & 0.7\\
+2014 & 8,925 & 0 & 99 & 1.1\\
+\bottomrule
+\end{tabular}
+
+\textbackslash{}end\{table\}
+
+\textbackslash{}begin\{table\}
+
+\textbackslash{}caption\{(\#tab:comp\_forest)Test\} \centering
+
+\begin{tabular}[t]{llrrrrrrrr}
+\toprule
+Forest type & Source & 1953 & 1973 & 1990 & 2000 & 2005 & 2010 & 2013 & 2014\\
+\midrule
+Total & Harper2007 & 15,996 & 14,173 & 10,606 & 8,982 & - & - & - & -\\
+ & MEFT2009 & - & - & 10,650 & 9,678 & 9,413 & - & - & -\\
+ & ONE2015 & - & - & - & - & 9,451 & 8,977 & 8,486 & -\\
+ & this study & 15,968 & 14,243 & 10,762 & 9,879 & 9,668 & 9,320 & 9,051 & 8,925\\
+Moist & Harper2007 & 8,766 & 6,876 & 5,234 & 4,167 & - & - & - & -\\
+\addlinespace
+ & MEFT2009 & - & - & 5,271 & 4,788 & 4,700 & - & - & -\\
+ & ONE2015 & - & - & - & - & 4,556 & 4,457 & 4,345 & -\\
+ & this study & 8,578 & 6,990 & 5,270 & 4,872 & 4,768 & 4,633 & 4,470 & 4,410\\
+Dry & Harper2007 & 4,252 & 4,028 & 2,712 & 2,457 & - & - & - & -\\
+ & MEFT2009 & - & - & 3,321 & 3,085 & 3,028 & - & - & -\\
+\addlinespace
+ & ONE2015 & - & - & - & - & 3,223 & 2,970 & 2,679 & -\\
+ & this study & 4,762 & 4,435 & 3,225 & 2,941 & 2,881 & 2,735 & 2,642 & 2,596\\
+Spiny & Harper2007 & 2,978 & 3,030 & 2,420 & 2,132 & - & - & - & -\\
+ & MEFT2009 & - & - & 2,124 & 1,872 & 1,757 & - & - & -\\
+ & ONE2015 & - & - & - & - & 1,682 & 1,559 & 1,467 & -\\
+\addlinespace
+ & this study & 2,463 & 2,583 & 2,055 & 1,858 & 1,811 & 1,744 & 1,731 & 1,713\\
+Mangroves & Harper2007 & - & - & 240 & 226 & - & - & - & -\\
+ & MEFT2009 & - & - & - & - & - & - & - & -\\
+ & ONE2015 & - & - & - & - & 174 & 171 & 170 & -\\
+ & this study & 143 & 200 & 181 & 178 & 177 & 177 & 177 & 177\\
+\bottomrule
+\end{tabular}
+
+\textbackslash{}end\{table\}
+
+\textbackslash{}begin\{table\}
+
+\textbackslash{}caption\{(\#tab:comp\_defor)Test\} \centering
+
+\begin{tabular}[t]{llrrrrrr}
+\toprule
+Forest type & Source & 1953-1973 & 1973-1990 & 1990-2000 & 2000-2005 & 2005-2010 & 2010-2013\\
+\midrule
+Total & Harper2007 & 91 (0.3) & 200 (1.7) & 81 (0.9) & - & - & -\\
+ & MEFT2009 & - & - & 97 (0.8) & 53 (0.5) & - & -\\
+ & ONE2015 & - & - & - & - & 95 (1.2) & 164 (1.5)\\
+ & this study & 86 (0.6) & 205 (1.6) & 88 (0.9) & 42 (0.4) & 70 (0.7) & 90 (1.0)\\
+Moist & Harper2007 & 94 (0.6) & 87 (1.7) & 32 (0.8) & - & - & -\\
+\addlinespace
+ & MEFT2009 & - & - & 48 (0.8) & 17 (0.4) & - & -\\
+ & ONE2015 & - & - & - & - & 20 (0.5) & 37 (0.9)\\
+ & this study & 79 (1.0) & 101 (1.6) & 40 (0.8) & 21 (0.4) & 27 (0.6) & 54 (1.2)\\
+Dry & Harper2007 & 11 (0.2) & 77 (1.9) & 20 (0.7) & - & - & -\\
+ & MEFT2009 & - & - & 24 (0.7) & 11 (0.4) & - & -\\
+\addlinespace
+ & ONE2015 & - & - & - & - & 51 (1.8) & 97 (2.3)\\
+ & this study & 16 (0.4) & 71 (1.9) & 28 (0.9) & 12 (0.4) & 29 (1.0) & 31 (1.1)\\
+Spiny & Harper2007 & -3 (-0.1) & 36 (1.2) & 28 (1.2) & - & - & -\\
+ & MEFT2009 & - & - & 25 (1.2) & 23 (1.2) & - & -\\
+ & ONE2015 & - & - & - & - & 25 (1.7) & 31 (1.7)\\
+\addlinespace
+ & this study & -6 (-0.2) & 31 (1.3) & 20 (1.0) & 9 (0.5) & 13 (0.7) & 4 (0.3)\\
+Mangroves & Harper2007 & - & - & 1 (0.2) & - & - & -\\
+ & MEFT2009 & - & - & - & - & - & -\\
+ & ONE2015 & - & - & - & - & 0 (0.3) & 0 (0.2)\\
+ & this study & -3 (-1.7) & 1 (0.6) & 0 (0.2) & 0 (0.0) & 0 (0.0) & 0 (0.0)\\
+\bottomrule
+\end{tabular}
+
+\textbackslash{}end\{table\}
+
+\begin{table}
+
+\caption{\label{tab:frag}Test}
+\centering
+\begin{tabular}[t]{lrrrrrr}
+\toprule
+Year & Forest (Kha) & 0-20 & 21-40 & 41-60 & 61-80 & 81-100\\
+\midrule
+1953 & 15,968 & 0 & 1 & 8 & 12 & 78\\
+1973 & 14,243 & 6 & 9 & 12 & 16 & 57\\
+1990 & 10,762 & 7 & 10 & 13 & 17 & 53\\
+2000 & 9,879 & 7 & 11 & 14 & 17 & 51\\
+2005 & 9,673 & 8 & 11 & 14 & 18 & 49\\
+\addlinespace
+2010 & 9,320 & 8 & 12 & 15 & 18 & 47\\
+2014 & 8,925 & 9 & 13 & 16 & 19 & 44\\
+\bottomrule
+\end{tabular}
+\end{table}
+
+\hypertarget{figures}{%
+\section{Figures}\label{figures}}
+
+\begin{figure}
+\centering
+\includegraphics{figs/ecoregion.png}
+\caption{\textbf{Ecoregions and forest types in Madagascar.} Madagascar
+can be divided into four climatic ecoregions with four forest types: the
+moist forest in the East (green), the dry forest in the West (orange),
+the spiny forest in the South (red), and the mangroves on the West coast
+(blue). Ecoregions were defined following climatic (Cornet
+\protect\hyperlink{ref-Cornet1974}{1974}) and vegetation (Ministère de
+l'Environnement \protect\hyperlink{ref-IEFN1996}{1996}) criteria. The
+dark grey areas represent the remaining natural forest cover for the
+year 2014.}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics{figs/fig_fcc.png}
+\caption{\textbf{Forest cover change on six decades from 1953 to 2014 in
+Madagascar.} Forest cover changes from \emph{c.} 1973 to 2014 are shown
+in the main figure, and forest cover in \emph{c.} 1953 is shown in the
+bottom-right inset. Two zooms in the western dry (left part) and eastern
+moist (right part) ecoregions present more detailed views of (from top
+to bottom): forest cover in 1950s, forest cover change from \emph{c.}
+1973 to 2014, forest fragmentation in 2014 and distance to forest edge
+in 2014. Data on water bodies (blue) and water seasonality (light blue
+for seasonal water to dark blue for permanent water) has been extracted
+from Pekel \emph{et al.} (\protect\hyperlink{ref-Pekel2016}{2016}).}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics{figs/dist.png}
+\caption{\textbf{Evolution of the distance to forest edge from 1953 to
+2014 in Madagascar.} Black dots represent the mean distance to forest
+edge for each year. Vertical dashed segments represent the 90\%
+quantiles (5\% and 95\%) of the distance to forest edge. Horizontal
+dashed grey line indicates a distance to forest edge of 100 m. Numbers
+at the bottom of each vertical segments are the percentage of forest at
+a distance to forest edge lower than 100 m for each year.}
+\end{figure}
+
+\hypertarget{references}{%
+\section*{References}\label{references}}
+\addcontentsline{toc}{section}{References}
+
+\hypertarget{refs}{}
+\leavevmode\hypertarget{ref-Achard2014}{}%
+Achard, F., Beuchle, R., Mayaux, P., Stibig, H.-J., Bodart, C., Brink,
+A., Carboni, S., Desclée, B., Donnay, F., Eva, H.D., Lupi, A., Raši, R.,
+Seliger, R. \& Simonetti, D. (2014) Determination of tropical
+deforestation rates and related carbon losses from 1990 to 2010.
+\emph{Global Change Biology}, \textbf{20}, 2540--2554.
+
+\leavevmode\hypertarget{ref-Aleman2017}{}%
+Aleman, J.C., Jarzyna, M.A. \& Staver, A.C. (2017) Forest extent and
+deforestation in tropical africa since 1900. \emph{Nature Ecology \&
+Evolution}.
+
+\leavevmode\hypertarget{ref-Ali2008}{}%
+Ali, J.R. \& Aitchison, J.C. (2008) Gondwana to Asia: Plate tectonics,
+paleogeography and the biological connectivity of the Indian
+sub-continent from the Middle Jurassic through latest Eocene (166--35
+Ma). \emph{Earth-Science Reviews}, \textbf{88}, 145--166.
+
+\leavevmode\hypertarget{ref-Allnutt2008}{}%
+Allnutt, T.F., Ferrier, S., Manion, G., Powell, G.V.N., Ricketts, T.H.,
+Fisher, B.L., Harper, G.J., Irwin, M.E., Kremen, C., Labat, J.-N., Lees,
+D.C., Pearce, T.A. \& Rakotondrainibe, F. (2008) A method for
+quantifying biodiversity loss and its application to a 50-year record of
+deforestation across Madagascar. \emph{Conservation Letters},
+\textbf{1}, 173--181.
+
+\leavevmode\hypertarget{ref-Armitage2015}{}%
+Armitage, A.R., Highfield, W.E., Brody, S.D. \& Louchouarn, P. (2015)
+The contribution of mangrove expansion to salt marsh loss on the Texas
+Gulf Coast. \emph{PloS One}, \textbf{10}, e0125404.
+
+\leavevmode\hypertarget{ref-Bastin2017}{}%
+Bastin, J.-F., Berrahmouni, N., Grainger, A., Maniatis, D., Mollicone,
+D., Moore, R., Patriarca, C., Picard, N., Sparrow, B., Abraham, E.M.,
+Aloui, K., Atesoglu, A., Attore, F., Bassüllü, Ç., Bey, A., Garzuglia,
+M., García-Montero, L.G., Groot, N., Guerin, G., Laestadius, L., Lowe,
+A.J., Mamane, B., Marchi, G., Patterson, P., Rezende, M., Ricci, S.,
+Salcedo, I., Diaz, A.S.-P., Stolle, F., Surappaeva, V. \& Castro, R.
+(2017) The extent of forest in dryland biomes. \emph{Science},
+\textbf{356}, 635--638.
+
+\leavevmode\hypertarget{ref-Brinck2017}{}%
+Brinck, K., Fischer, R., Groeneveld, J., Lehmann, S., Dantas De Paula,
+M., Pütz, S., Sexton, J.O., Song, D. \& Huth, A. (2017) High resolution
+analysis of tropical forest fragmentation and its impact on the global
+carbon cycle. \emph{Nature Communications}, \textbf{8}, 14855.
+
+\leavevmode\hypertarget{ref-Broadbent2008}{}%
+Broadbent, E.N., Asner, G.P., Keller, M., Knapp, D.E., Oliveira, P.J. \&
+Silva, J.N. (2008) Forest fragmentation and edge effects from
+deforestation and selective logging in the brazilian amazon.
+\emph{Biological Conservation}, \textbf{141}, 1745--1757.
+
+\leavevmode\hypertarget{ref-Brooks2002}{}%
+Brooks, T.M., Mittermeier, R.A., Mittermeier, C.G., Fonseca, G.A.B. da,
+Rylands, A.B., Konstant, W.R., Flick, P., Pilgrim, J., Oldfield, S.,
+Magin, G. \& Hilton-Taylor, C. (2002) Habitat loss and extinction in the
+hotspots of biodiversity. \emph{Conservation Biology}, \textbf{16},
+909--923.
+
+\leavevmode\hypertarget{ref-Burns2016}{}%
+Burns, S.J., Godfrey, L.R., Faina, P., McGee, D., Hardt, B.,
+Ranivoharimanana, L. \& Randrianasy, J. (2016) Rapid human-induced
+landscape transformation in Madagascar at the end of the first
+millennium of the Common Era. \emph{Quaternary Science Reviews},
+\textbf{134}, 92--99.
+
+\leavevmode\hypertarget{ref-Cornet1974}{}%
+Cornet, A. (1974) \emph{Essai de cartographie bioclimatique à
+Madagascar}. Orstom.
+
+\leavevmode\hypertarget{ref-Cox2012}{}%
+Cox, M.P., Nelson, M.G., Tumonggor, M.K., Ricaut, F.-X. \& Sudoyo, H.
+(2012) A small cohort of Island Southeast Asian women founded
+Madagascar. \emph{Proceedings of the Royal Society B: Biological
+Sciences}, \textbf{279}, 2761--2768.
+
+\leavevmode\hypertarget{ref-Crottini2012}{}%
+Crottini, A., Madsen, O., Poux, C., Strauß, A., Vieites, D.R. \& Vences,
+M. (2012) Vertebrate time-tree elucidates the biogeographic pattern of a
+major biotic change around the K--T boundary in Madagascar.
+\emph{Proceedings of the National Academy of Sciences}, \textbf{109},
+5358--5363.
+
+\leavevmode\hypertarget{ref-Eklund2016}{}%
+Eklund, J., Blanchet, F.G., Nyman, J., Rocha, R., Virtanen, T. \&
+Cabeza, M. (2016) Contrasting spatial and temporal trends of protected
+area effectiveness in mitigating deforestation in Madagascar.
+\emph{Biological Conservation}, \textbf{203}, 290--297.
+
+\leavevmode\hypertarget{ref-Freudenberger2010}{}%
+Freudenberger, K. (2010) Paradise Lost? Lessons from 25 years of USAID
+environment programs in Madagascar. \emph{International Resources Group,
+Washington DC}.
+
+\leavevmode\hypertarget{ref-Gibson2013}{}%
+Gibson, L., Lynam, A.J., Bradshaw, C.J., He, F., Bickford, D.P.,
+Woodruff, D.S., Bumrungsri, S. \& Laurance, W.F. (2013) Near-complete
+extinction of native small mammal fauna 25 years after forest
+fragmentation. \emph{Science}, \textbf{341}, 1508--1510.
+
+\leavevmode\hypertarget{ref-Goodman2005}{}%
+Goodman, S.M. \& Benstead, J.P. (2005) Updated estimates of biotic
+diversity and endemism for Madagascar. \emph{Oryx}, \textbf{39}, 73--77.
+
+\leavevmode\hypertarget{ref-Gorenflo2011}{}%
+Gorenflo, L.J., Corson, C., Chomitz, K.M., Harper, G., Honzák, M. \&
+Özler, B. (2011) \emph{Exploring the Association Between People and
+Deforestation in Madagascar} (eds R.P. Cincotta \& L.J. Gorenflo).
+Springer Berlin Heidelberg.
+
+\leavevmode\hypertarget{ref-Grinand2017}{}%
+Grinand, C., Le Maire, G., Vieilledent, G., Razakamanarivo, H.,
+Razafimbelo, T. \& Bernoux, M. (2017) Estimating temporal changes in
+soil carbon stocks at ecoregional scale in Madagascar using
+remote-sensing. \emph{International Journal of Applied Earth Observation
+and Geoinformation}, \textbf{54}, 1--14.
+
+\leavevmode\hypertarget{ref-Grinand2013}{}%
+Grinand, C., Rakotomalala, F., Gond, V., Vaudry, R., Bernoux, M. \&
+Vieilledent, G. (2013) Estimating deforestation in tropical humid and
+dry forests in Madagascar from 2000 to 2010 using multi-date Landsat
+satellite images and the Random Forests classifier. \emph{Remote Sensing
+of Environment}, \textbf{139}, 68--80.
+
+\leavevmode\hypertarget{ref-Grouzis2001}{}%
+Grouzis, M., Razanaka, S., Le Floc'h, E. \& Leprun, J.-C. (2001)
+Évolution de la végétation et de quelques paramètres édaphiques au cours
+de la phase post-culturale dans la région d'Analabo. \emph{Sociétés
+paysannes, transitions agraires et dynamiques écologiques dans le
+Sud-Ouest de Madagascar, Antananarivo, IRD/CNRE}, 327--337.
+
+\leavevmode\hypertarget{ref-Hansen2013}{}%
+Hansen, M.C., Potapov, P.V., Moore, R., Hancher, M., Turubanova, S.A.,
+Tyukavina, A., Thau, D., Stehman, S.V., Goetz, S.J., Loveland, T.R.,
+Kommareddy, A., Egorov, A., Chini, L., Justice, C.O. \& Townshend,
+J.R.G. (2013) High-Resolution Global Maps of 21st-Century Forest Cover
+Change. \emph{Science}, \textbf{342}, 850--853.
+
+\leavevmode\hypertarget{ref-Harper2007}{}%
+Harper, G.J., Steininger, M.K., Tucker, C.J., Juhn, D. \& Hawkins, F.
+(2007) Fifty years of deforestation and forest fragmentation in
+Madagascar. \emph{Environmental Conservation}, \textbf{34}, 325--333.
+
+\leavevmode\hypertarget{ref-Humbert1927}{}%
+Humbert, H. (1927) La destruction d'une flore insulaire par le feu.
+Principaux aspects de la végétation à Madagascar. \emph{Mémoires de
+l'Académie Malgache}, \textbf{5}, 1--80.
+
+\leavevmode\hypertarget{ref-Keenan2015}{}%
+Keenan, R.J., Reams, G.A., Achard, F., Freitas, J.V. de, Grainger, A. \&
+Lindquist, E. (2015) Dynamics of global forest area: Results from the
+FAO Global Forest Resources Assessment 2015. \emph{Forest Ecology and
+Management}, \textbf{352}, 9--20.
+
+\leavevmode\hypertarget{ref-Kim2014}{}%
+Kim, D.-H., Sexton, J.O., Noojipady, P., Huang, C., Anand, A., Channan,
+S., Feng, M. \& Townshend, J.R. (2014) Global, Landsat-based
+forest-cover change from 1990 to 2000. \emph{Remote Sensing of
+Environment}, \textbf{155}, 178--193.
+
+\leavevmode\hypertarget{ref-Klein2002}{}%
+Klein, J. (2002) Deforestation in the Madagascar Highlands --
+Established `truth' and scientific uncertainty. \emph{GeoJournal},
+\textbf{56}, 191--199.
+
+\leavevmode\hypertarget{ref-Kull2000}{}%
+Kull, C.A. (2000) Deforestation, erosion, and fire: degradation myths in
+the environmental history of Madagascar. \emph{Environment and History},
+\textbf{6}, 423--450.
+
+\leavevmode\hypertarget{ref-MEFT2009}{}%
+MEFT, USAID, and CI. (2009) \emph{Evolution de la couverture de forêts
+naturelles à Madagascar, 1990-2000-2005}. Antananarivo.
+
+\leavevmode\hypertarget{ref-IEFN1996}{}%
+Ministère de l'Environnement. (1996) \emph{IEFN: Inventaire Ecologique
+Forestier National}. Ministère de l'Environnement de Madagascar,
+Direction des Eaux et Forêts, DFS Deutsch Forest Service GmbH,
+Entreprise d'études de développement rural "Mamokatra", FTM.
+
+\leavevmode\hypertarget{ref-Minten2013}{}%
+Minten, B., Sander, K. \& Stifel, D. (2013) Forest management and
+economic rents: Evidence from the charcoal trade in Madagascar.
+\emph{Energy for Sustainable Development}, \textbf{17}, 106--115.
+
+\leavevmode\hypertarget{ref-Murcia1995}{}%
+Murcia, C. (1995) Edge effects in fragmented forests: Implications for
+conservation. \emph{Trends in Ecology \& Evolution}, \textbf{10},
+58--62.
+
+\leavevmode\hypertarget{ref-Neteler2008}{}%
+Neteler, M. \& Mitasova, H. (2008) \emph{Open source GIS: a GRASS GIS
+approach}. Springer.
+
+\leavevmode\hypertarget{ref-Olander2008}{}%
+Olander, L.P., Gibbs, H.K., Steininger, M., Swenson, J.J. \& Murray,
+B.C. (2008) Reference scenarios for deforestation and forest degradation
+in support of REDD: a review of data and methods. \emph{Environmental
+Research Letters}, \textbf{3}, 025011.
+
+\leavevmode\hypertarget{ref-Olofsson2014}{}%
+Olofsson, P., Foody, G.M., Herold, M., Stehman, S.V., Woodcock, C.E. \&
+Wulder, M.A. (2014) Good practices for estimating area and assessing
+accuracy of land change. \emph{Remote Sensing of Environment},
+\textbf{148}, 42--57.
+
+\leavevmode\hypertarget{ref-Olofsson2013}{}%
+Olofsson, P., Foody, G.M., Stehman, S.V. \& Woodcock, C.E. (2013) Making
+better use of accuracy data in land change studies: Estimating accuracy
+and area and quantifying uncertainty using stratified estimation.
+\emph{Remote Sensing of Environment}, \textbf{129}, 122--131.
+
+\leavevmode\hypertarget{ref-ONE2013}{}%
+ONE, DGF, FTM, MNP, and CI. (2013) \emph{Evolution de la couverture de
+forêts naturelles à Madagascar 2005-2010}. Antananarivo.
+
+\leavevmode\hypertarget{ref-ONE2015}{}%
+ONE, DGF, MNP, WCS, and Etc Terra. (2015) \emph{Changement de la
+couverture de forêts naturelles à Madagascar, 2005-2010-2013}.
+Antananarivo.
+
+\leavevmode\hypertarget{ref-Pearson2009}{}%
+Pearson, R.G. \& Raxworthy, C.J. (2009) The evolution of local endemism
+in Madagascar: watershed versus climatic gradient hypotheses evaluated
+by null biogeographic models. \emph{Evolution}, \textbf{63}, 959--967.
+
+\leavevmode\hypertarget{ref-Pekel2016}{}%
+Pekel, J.-F., Cottam, A., Gorelick, N. \& Belward, A.S. (2016)
+High-resolution mapping of global surface water and its long-term
+changes. \emph{Nature}, \textbf{540}, 418--422.
+
+\leavevmode\hypertarget{ref-Perrier1921}{}%
+Perrier de La Bâthie, H. (1921) \emph{La végétation malgache}. Musée
+Colonial.
+
+\leavevmode\hypertarget{ref-Ploch2012}{}%
+Ploch, L. \& Cook, N. (2012) Madagascar's Political Crisis.
+
+\leavevmode\hypertarget{ref-Puyravaud2003}{}%
+Puyravaud, J.P. (2003) Standardizing the calculation of the annual rate
+of deforestation. \emph{Forest Ecology and Management}, \textbf{177},
+593--596.
+
+\leavevmode\hypertarget{ref-Rakotomalala2015}{}%
+Rakotomala, F., Rabenandrasana, J., Andriambahiny, J., Rajaonson, R.,
+Andriamalala, F., Burren, C., Rakotoarijaona, J., Parany, B., Vaudry,
+R., Rakotoniaina, S., Ranaivosoa, R., Rahagalala, P., Randrianary, T. \&
+Grinand, C. (2015) Estimation de la déforestation des forêts humides à
+Madagascar entre 2005, 2010 et 2013: combinaison multi-date d'images
+LANDSAT, utilisation de l'algorithme Random Forest et procédure de
+validation. \emph{Revue Française de Photogrammétrie et de
+Télédétection}, 11--23.
+
+\leavevmode\hypertarget{ref-Saatchi2011}{}%
+Saatchi, S.S., Harris, N.L., Brown, S., Lefsky, M., Mitchard, E.T.A.,
+Salas, W., Zutta, B.R., Buermann, W., Lewis, S.L., Hagen, S., Petrova,
+S., White, L., Silman, M. \& Morel, A. (2011) Benchmark map of forest
+carbon stocks in tropical regions across three continents.
+\emph{Proceedings of the National Academy of Sciences}, \textbf{108},
+9899--9904.
+
+\leavevmode\hypertarget{ref-Saunders1991}{}%
+Saunders, D.A., Hobbs, R.J. \& Margules, C.R. (1991) Biological
+Consequences of Ecosystem Fragmentation: A Review. \emph{Conservation
+Biology}, \textbf{5}, 18--32.
+
+\leavevmode\hypertarget{ref-Scales2011}{}%
+Scales, I.R. (2011) Farming at the Forest Frontier: Land Use and
+Landscape Change in Western Madagascar, 1896-2005. \emph{Environment and
+History}, \textbf{17}, 499--524.
+
+\leavevmode\hypertarget{ref-Smith2003}{}%
+Smith, R.J., Muir, R.D.J., Walpole, M.J., Balmford, A. \&
+Leader-Williams, N. (2003) Governance and the loss of biodiversity.
+\emph{Nature}, \textbf{426}, 67--70.
+
+\leavevmode\hypertarget{ref-Tidd2001}{}%
+Tidd, S.T., Pinder, J. \& Ferguson, G.W. (2001) Deforestation and
+habitat loss for the Malagasy flat-tailed tortoise from 1963 through
+1993. \emph{Chelonian Conservation and Biology}, \textbf{4}, 59--65.
+
+\leavevmode\hypertarget{ref-Tropek2014}{}%
+Tropek, R., Sedláček, O., Beck, J., Keil, P., Musilová, Z., Šímová, I.
+\& Storch, D. (2014) Comment on "High-resolution global maps of
+21st-century forest cover change". \emph{Science}, \textbf{344},
+981--981.
+
+\leavevmode\hypertarget{ref-Tyukavina2015}{}%
+Tyukavina, A., Baccini, A., Hansen, M.C., Potapov, P.V., Stehman, S.V.,
+Houghton, R.A., Krylov, A.M., Turubanova, S. \& Goetz, S.J. (2015)
+Aboveground carbon loss in natural and managed tropical forests from
+2000 to 2012. \emph{Environmental Research Letters}, \textbf{10},
+074002.
+
+\leavevmode\hypertarget{ref-Tyukavina2017}{}%
+Tyukavina, A., Hansen, M.C., Potapov, P.V., Stehman, S.V.,
+Smith-Rodriguez, K., Okpa, C. \& Aguilar, R. (2017) Types and rates of
+forest disturbance in Brazilian Legal Amazon, 20002013. \emph{Science
+Advances}, \textbf{3}.
+
+\leavevmode\hypertarget{ref-UN2015}{}%
+United Nations. (2015) \emph{World Population Prospects: The 2015
+Revision, Key Findings and Advance Tables. Working Paper No.
+ESA/P/WP.241.}
+
+\leavevmode\hypertarget{ref-Verhegghen2016}{}%
+Verhegghen, A., Eva, H., Desclée, B. \& Achard, F. (2016) Review and
+Combination of Recent Remote Sensing Based Products for Forest Cover
+Change Assessments in Cameroon. \emph{International Forestry Review},
+\textbf{18}, 14--25.
+
+\leavevmode\hypertarget{ref-Vieilledent2016}{}%
+Vieilledent, G., Gardi, O., Grinand, C., Burren, C., Andriamanjato, M.,
+Camara, C., Gardner, C.J., Glass, L., Rasolohery, A., Rakoto Ratsimba,
+H., Gond, V. \& Rakotoarijaona, J.-R. (2016) Bioclimatic envelope models
+predict a decrease in tropical forest carbon stocks with climate change
+in Madagascar. \emph{Journal of Ecology}, \textbf{104}, 703--715.
+
+\leavevmode\hypertarget{ref-Vieilledent2013}{}%
+Vieilledent, G., Grinand, C. \& Vaudry, R. (2013) Forecasting
+deforestation and carbon emissions in tropical developing countries
+facing demographic expansion: a case study in Madagascar. \emph{Ecology
+and Evolution}, \textbf{3}, 1702--1716.
+
+\leavevmode\hypertarget{ref-Virah-Sawmy2009}{}%
+Virah-Sawmy, M. (2009) Ecosystem management in Madagascar during global
+change. \emph{Conservation Letters}, \textbf{2}, 163--170.
+
+\leavevmode\hypertarget{ref-Vorontsova2017}{}%
+Vorontsova, M.S., Besnard, G., Forest, F., Malakasi, P., Moat, J.,
+Clayton, W.D., Ficinski, P., Savva, G.M., Nanjarisoa, O.P., Razanatsoa,
+J., Randriatsara, F.O., Kimeu, J.M., Luke, W.R.Q., Kayombo, C. \&
+Linder, H.P. (2016) Madagascar's grasses and grasslands: anthropogenic
+or natural? \emph{Proceedings of the Royal Society of London B:
+Biological Sciences}, \textbf{283}.
 

### 2018-01-17 manuscript/manuscript.Rmd
Use Rnw in place of Rmd

@@
 ---
-title: Combining global tree cover loss data with historical national forest-cover
-  maps to look at six decades of deforestation and forest fragmentation in Madagascar
+title: "Combining global tree cover loss data with historical national forest-cover maps to look at six decades of deforestation and forest fragmentation in Madagascar"
 
@@
     toc: no
+    highlight: tango
 
 link-citations: yes
+# colorlinks: yes
+linkcolor: blue
+urlcolor: magenta
+citecolor: green
+
 bibliography: bib/biblio.bib
@@
                fig.show="hide", size="small",
+               fig.align="center",
                tidy=FALSE)
@@
 
-(*) /Email: ghislain.vieilledent@cirad.fr, /Phone: +39.329.457.2273
-
-*Running headline:* Six decades of deforestation in Madagascar
-
-# Highlights
-
-- We produced new 30m-resolution annual forest-cover maps for Madagascar for the period 2000-2014.
-- Madagascar has lost 44% of its natural forest-cover over the period 1953-2014.
-- Half of the tropical forest in Madagascar is now located at less than 100m from forest edge.
-- Annual deforestation rate has increased in Madagascar since 2005 to reach 1.08%/yr.
-- Conservation and development efforts must be intensified to save Madagascar forest and biodiversity.
-
-# Summary
-
-The island of Madagascar has a unique biodiversity, mainly located in the tropical forests of the island.
+**Abstract**: The island of Madagascar has a unique biodiversity, mainly located in the tropical forests of the island.
 This biodiversity is highly threatened by anthropogenic deforestation.
@@
 
-**Keywords**: biodiversity, climate-change, deforestation, forest-fragmentation, Madagascar, tropical forest.
+**Keywords**: biodiversity, climate-change, deforestation, forest-fragmentation, Madagascar, tropical forest
+
+**Running headline**: Six decades of deforestation in Madagascar
+
+($\star$) Corresponding author: <ghislain.vieilledent@cirad.fr>
+
+```{r highlights}
+# # Highlights
+# - We produced new 30m-resolution annual forest-cover maps for Madagascar for the period 2000-2014.
+# - Madagascar has lost 44% of its natural forest-cover over the period 1953-2014.
+# - Half of the tropical forest in Madagascar is now located at less than 100m from forest edge.
+# - Annual deforestation rate has increased in Madagascar since 2005 to reach 1.08%/yr.
+# - Conservation and development efforts must be intensified to save Madagascar forest and biodiversity.
+```
 
@@
 South, and mangroves on the Western coast
-(Fig. \ref{fig:ecoregions}). To define the forest domains, we used a map
+(Fig. \@ref(fig:ecoregions)). To define the forest domains, we used a map
 from the MEFT (*"Ministère de l'Environnement et des Forêts à
@@
 dropped to 8.9 Mha, corresponding to about 15% of the national
-territory (Fig.~\ref{fig:fcc} and
-Tab. \@ref(tab:forest_cover)). Madagascar has lost 44% and 37% of its
+territory (Fig. \@ref(fig:fcc) and
+Tab. \@ref(tab:forestcover)). Madagascar has lost 44% and 37% of its
 natural forests between 1953 and 2014, and between 1973 and 2014
-respectively (Fig.~\ref{fig:fcc} and Tab. \@ref(tab:forest_cover)). In
+respectively (Fig. \@ref(fig:fcc) and Tab. \@ref(tab:forestcover)). In
 2014 the remaining 8.9 Mha of natural forest were distributed as
@@
 of dry forest (29%), 1.7 Mha of spiny forest (19%) and 0.18 Mha
-(2%) of mangrove forest (Fig.~\ref{fig:ecoregion} and
-Tab. \@ref(tab:comp_forest)). Regarding the deforestation trend, we
+(2%) of mangrove forest (Fig. \@ref(fig:ecoregion) and
+Tab. \@ref(tab:compforest)). Regarding the deforestation trend, we
 observed a progressive decrease of the deforestation rate after 1990
@@
 ha/yr (0.4%/yr) over the period 2000-2005
-(Tab. \@ref(tab:forest_cover)). Then from 2005, the deforestation rate
+(Tab. \@ref(tab:forestcover)). Then from 2005, the deforestation rate
 has progressively increased and has more than doubled over the period
 2010-2014 (99,000 ha/yr, 1.1%/yr) compared to 2000-2005
-(Tab. \@ref(tab:forest_cover)). The deforestation trend, characterized
+(Tab. \@ref(tab:forestcover)). The deforestation trend, characterized
 by a progressive decrease of the deforestation rate over the period
@@
 is valid for all four forest types except the spiny forest
-(Tab. \@ref(tab:comp_defor)). For the spiny forest, the deforestation
+(Tab. \@ref(tab:compdefor)). For the spiny forest, the deforestation
 rate during the period 2010-2013 was lower than on the period
-2005-2010 (Tab. \@ref(tab:comp_defor)).
+2005-2010 (Tab. \@ref(tab:compdefor)).
 
@@
 wall-to-wall forest cover change maps from 1990 to 2014 for the full
-territory of Madagascar (Tab. \@ref(tab:forest_cover)). This allowed us
+territory of Madagascar (Tab. \@ref(tab:forestcover)). This allowed us
 to produce more robust estimates of forest cover and deforestation
@@
 (Pearson's correlation coefficient = 0.99) to estimates from the
-three previous studies (Tab. \@ref(tab:comp_forest)) with a RMSE of
+three previous studies (Tab. \@ref(tab:compforest)) with a RMSE of
 300,000 ha (6\% of the mean forest cover of 4.8 Mha when considering
@@
 significant differences in deforestation estimates
-(Tab. \@ref(tab:comp_defor)), a similar deforestation trend was
+(Tab. \@ref(tab:compdefor)), a similar deforestation trend was
 observed across studies with a decrease of deforestation rates over
@@
 _c._ 300 m in 2014 while it was previously _c._ 1.5 km in
-1973 (Fig.~\ref{fig:dist_edge}). Moreover, a large proportion (73%)
+1973 (Fig. \@ref(fig:distedge)). Moreover, a large proportion (73%)
 of the forest was located at a distance greater than 100 m in 1973,
 while almost half of the forest (46%) is now at a distance lower than
-100~m from forest edge in 2014 (Fig.~\ref{fig:dist_edge}). The
+100~m from forest edge in 2014 (Fig. \@ref(fig:distedge)). The
 percentage of lowly fragmented forest in Madagascar has continuously
@@
 
-```{r forest_cover}
+(ref:tabcap1) **Evolution of natural forest cover and deforestation rates from 1953 to 2014 in Madagascar.** Areas are provided in thousands of hectares (Kha).
+Forest map for the year 1973 has 3.3 Mha of unclassified areas due to the presence of clouds on satellite images.
+As a consequence, partial deforestation rates for the periods 1953-1973 and 1973-1990 are computed based on the available forest extent.
+The last two columns indicate the annual deforested areas and annual deforestation rates on the previous time-period (e.g. 1953-1973 for year 1973, 1973-1990 for year 1990, etc.).
+
+```{r forestcover}
 defor.df <- read.table(file="tabs/defor.txt", sep="\t", header=TRUE)
@@
 names(tab.1) <- c("Year", "Forest (Kha)", "Unmap (Kha)", "Annual defor. (Kha/yr)", "Rate (%/yr)")
-# Caption
-#cap <- "**Evolution of natural forest cover and deforestation rates from 1953 to 2014 in Madagascar.** Areas are provided in thousands of hectares (Kha).
-Forest map for the year 1973 has 3.3 Mha of unclassified areas due to the presence of clouds on satellite images.
-As a consequence, partial deforestation rates for the periods 1953-1973 and 1973-1990 are computed based on the available forest extent.
-The last two columns indicate the annual deforested areas and annual deforestation rates on the previous time-period (e.g. 1953-1973 for year 1973, 1973-1990 for year 1990, etc.)."
 # Kable
-knitr::kable(tab.1, format.args=list(big.mark=","), digits=1, booktabs=TRUE, caption="Test")
+knitr::kable(tab.1, format.args=list(big.mark=","), digits=1, booktabs=TRUE, caption="(ref:tabcap1)")
 ```
 
-```{r comp_forest}
+(ref:tabcap2) **Comparing Madagascar forest cover estimates with previous studies on the period 1953-2014.** We compared our estimates of forest cover with the estimates from three previous studies [@Harper2007; @MEFT2009; @ONE2015].
+Areas are provided in thousands of hectares (Kha).
+We obtained a Pearson's correlation coefficient of 0.99 between our forest cover estimates and forest cover estimates from previous studies.
+The increase in mangrove and spiny forest covers from _c._ 1953 to _c._ 1973 in @Harper2007 and our study is most probably due to differences in forest definition and mapping methods between the 1953 aerial-photography derived map and the 1973 Landsat image derived map.
+
+```{r compforest}
+# Import and format
 fcc.comp <- read.table(file="tabs/fcc_comp.txt", sep=" ", header=TRUE)
@@
 col.names <- c("Forest type","Source","1953","1973","1990","2000","2005","2010","2013","2014")
-# Caption
-#cap <- "**Comparing Madagascar forest cover estimates with previous studies on the period 1953-2014.** We compared our estimates of forest cover with the estimates from three previous studies [@Harper2007; @MEFT2009; @ONE2015].
-Areas are provided in thousands of hectares (Kha).
-We obtained a Pearson's correlation coefficient of 0.99 between our forest cover estimates and forest cover estimates from previous studies.
-The increase in mangrove and spiny forest covers from _c._ 1953 to _c._ 1973 in @Harper2007 and our study is most probably due to differences in forest definition and mapping methods between the 1953 aerial-photography derived map and the 1973 Landsat image derived map."
 # Kable
-knitr::kable(tab.2, col.names=col.names,align=c("l","l",rep("r",8)), format.args=list(big.mark=","), booktabs=TRUE, caption="Test")
+knitr::kable(tab.2, col.names=col.names, align=c("l","l",rep("r",8)), format.args=list(big.mark=","), booktabs=TRUE, caption="(ref:tabcap2)")
 ```
 
-```{r comp_defor}
+(ref:tabcap3) **Comparing Madagascar annual deforestation rates with previous studies on the period 1953-2014.** Annual deforested areas (in thousands of hectares per year, Kha/yr) and annual deforestation rates (second number in parenthesis, in %/yr) are provided.
+For deforestation rates in %/yr, exact same numbers as in scientific articles and reports from previous studies [@Harper2007; @MEFT2009; @ONE2015] have been reported.
+The way annual deforestation rates in %/yr have been computed in these previous studies can slightly differ from one study to another, but estimates always correct for the potential presences of clouds on satellite images and unclassified areas on forest maps.
+Annual deforested areas in Kha/yr have been recomputed from forest cover estimates in Tab. \@ref(tab:compforest) (except for @Harper2007 for the periods 1973-1990 and 1990-2000 for which annual deforested areas in Kha/yr were derived from numbers reported in the original publication, see methods) and do not correct for the potential presence of clouds.
+
+```{r compdefor}
+# Import and format
 tab.3 <- read.table(file="tabs/defor_comp.txt", sep=" ", header=TRUE)
@@
 col.names <- c("Forest type","Source","1953-1973","1973-1990","1990-2000","2000-2005","2005-2010","2010-2013")
-# Caption
-# cap <- "**Comparing Madagascar annual deforestation rates with previous studies on the period 1953-2014.** Annual deforested areas (in thousands of hectares per year, Kha/yr) and annual deforestation rates (second number in parenthesis, in %/yr) are provided.
-For deforestation rates in %/yr, exact same numbers as in scientific articles and reports from previous studies [@Harper2007; @MEFT2009; @ONE2015] have been reported.
-The way annual deforestation rates in %/yr have been computed in these previous studies can slightly differ from one study to another, but estimates always correct for the potential presences of clouds on satellite images and unclassified areas on forest maps.
-Annual deforested areas in Kha/yr have been recomputed from forest cover estimates in Tab. \\@ref(tab:comp_forest) (except for @Harper2007 for the periods 1973-1990 and 1990-2000 for which annual deforested areas in Kha/yr were derived from numbers reported in the original publication, see methods) and do not correct for the potential presence of clouds."
 # Kable
-knitr::kable(tab.3, col.names=col.names, align=c("l","l",rep("r",6)), booktabs=TRUE, caption="Test")
+knitr::kable(tab.3, col.names=col.names, align=c("l","l",rep("r",6)), booktabs=TRUE, caption="(ref:tabcap3)")
 ```
+
+(ref:tabcap4) **Evolution of the forest fragmentation from 1953 to 2014 in Madagascar.** Five forest fragmentation classes, based on the percentage of forest in the neighborhood, are defined: 0-20% (highly fragmented), 21-40%, 41-60%, 61-80% and 81-100% (lowly fragmented).
+The percentage of forest falling in each forest fragmentation class is reported for each year.
+Forest areas are provided in thousands of hectares (Kha).
 
@@
 names(tab.4) <- c("Year","Forest (Kha)", "0-20", "21-40", "41-60", "61-80", "81-100")
-# Caption
-#cap <- "**Evolution of the forest fragmentation from 1953 to 2014 in Madagascar.** Five forest fragmentation classes, based on the percentage of forest in the neighborhood, are defined: 0-20% (highly fragmented), 21-40%, 41-60%, 61-80% and 81-100% (lowly fragmented).
-The percentage of forest falling in each forest fragmentation class is reported for each year.
-Forest areas are provided in thousands of hectares (Kha)."
 # Kable
-knitr::kable(tab.4, digits=0, format.args=list(big.mark=","), booktabs=TRUE, caption="Test")
+knitr::kable(tab.4, digits=0, format.args=list(big.mark=","), booktabs=TRUE, caption="(ref:tabcap4)")
 ```
@@
 
-![**Ecoregions and forest types in Madagascar.** Madagascar can be divided into four climatic ecoregions with four forest types: the moist forest in the East (green), the dry forest in the West (orange), the spiny forest in the South (red), and the mangroves on the West coast (blue).
+(ref:figcap1) **Ecoregions and forest types in Madagascar.** Madagascar can be divided into four climatic ecoregions with four forest types: the moist forest in the East (green), the dry forest in the West (orange), the spiny forest in the South (red), and the mangroves on the West coast (blue).
 Ecoregions were defined following climatic [@Cornet1974] and vegetation [@IEFN1996] criteria.
 The dark grey areas represent the
-remaining natural forest cover for the year 2014.](figs/ecoregion.png)
+remaining natural forest cover for the year 2014.
 
-![**Forest cover change on six decades from 1953 to 2014 in Madagascar.** Forest cover changes from _c._ 1973 to 2014 are shown in the main figure, and forest cover in _c._ 1953 is shown in the bottom-right inset.
+``` {r ecoregion, fig.cap="(ref:figcap1)", out.width="11cm"}
+knitr::include_graphics("figs/ecoregion.png")
+```
+
+(ref:figcap2) **Forest cover change on six decades from 1953 to 2014 in Madagascar.** Forest cover changes from _c._ 1973 to 2014 are shown in the main figure, and forest cover in _c._ 1953 is shown in the bottom-right inset.
 Two zooms in the western dry (left part) and eastern moist (right part) ecoregions present more detailed views of (from top to bottom): forest cover in 1950s, forest cover change from _c._ 1973 to 2014, forest fragmentation in 2014 and distance to forest edge in 2014.
-Data on water bodies (blue) and water seasonality (light blue for seasonal water to dark blue for permanent water) has been extracted from @Pekel2016.](figs/fig_fcc.png)
+Data on water bodies (blue) and water seasonality (light blue for seasonal water to dark blue for permanent water) has been extracted from @Pekel2016.
 
-![**Evolution of the distance to forest edge from 1953 to 2014 in Madagascar.** Black dots represent the mean distance to forest edge for each year.
+``` {r fcc, fig.cap="(ref:figcap2)", out.width="\\textwidth"}
+knitr::include_graphics("figs/fig_fcc.png")
+```
+
+(ref:figcap3) **Evolution of the distance to forest edge from 1953 to 2014 in Madagascar.** Black dots represent the mean distance to forest edge for each year.
 Vertical dashed segments represent the 90% quantiles (5% and 95%) of the distance to forest edge.
 Horizontal dashed grey line indicates a distance to forest edge of 100 m.
-Numbers at the bottom of each vertical segments are the percentage of forest at a distance to forest edge lower than 100 m for each year.](figs/dist.png)
+Numbers at the bottom of each vertical segments are the percentage of forest at a distance to forest edge lower than 100 m for each year.
+
+``` {r distedge, fig.cap="(ref:figcap3)", out.width="12cm"}
+knitr::include_graphics("figs/dist.png")
+```
 

### 2018-01-17 manuscript/manuscript.tex
Use Rnw in place of Rmd

--- 2018-01-12 manuscript/manuscript.tex
@@
 
-\documentclass[a4paper, 12pt, leqno]{article} %leqno: numéro d'équation à gauche
+\documentclass[a4paper, 12pt, leqno]{article}\usepackage[]{graphicx}\usepackage[]{color}
+%% maxwidth is the original width if it is less than linewidth
+%% otherwise use linewidth (to make sure the graphics do not exceed the margin)
+\makeatletter
+\def\maxwidth{ %
+  \ifdim\Gin@nat@width>\linewidth
+    \linewidth
+  \else
+    \Gin@nat@width
+  \fi
+}
+\makeatother
+
+\definecolor{fgcolor}{rgb}{0.345, 0.345, 0.345}
+\newcommand{\hlnum}[1]{\textcolor[rgb]{0.686,0.059,0.569}{#1}}%
+\newcommand{\hlstr}[1]{\textcolor[rgb]{0.192,0.494,0.8}{#1}}%
+\newcommand{\hlcom}[1]{\textcolor[rgb]{0.678,0.584,0.686}{\textit{#1}}}%
+\newcommand{\hlopt}[1]{\textcolor[rgb]{0,0,0}{#1}}%
+\newcommand{\hlstd}[1]{\textcolor[rgb]{0.345,0.345,0.345}{#1}}%
+\newcommand{\hlkwa}[1]{\textcolor[rgb]{0.161,0.373,0.58}{\textbf{#1}}}%
+\newcommand{\hlkwb}[1]{\textcolor[rgb]{0.69,0.353,0.396}{#1}}%
+\newcommand{\hlkwc}[1]{\textcolor[rgb]{0.333,0.667,0.333}{#1}}%
+\newcommand{\hlkwd}[1]{\textcolor[rgb]{0.737,0.353,0.396}{\textbf{#1}}}%
+\let\hlipl\hlkwb
+
+\usepackage{framed}
+\makeatletter
+\newenvironment{kframe}{%
+ \def\at@end@of@kframe{}%
+ \ifinner\ifhmode%
+  \def\at@end@of@kframe{\end{minipage}}%
+  \begin{minipage}{\columnwidth}%
+ \fi\fi%
+ \def\FrameCommand##1{\hskip\@totalleftmargin \hskip-\fboxsep
+ \colorbox{shadecolor}{##1}\hskip-\fboxsep
+     % There is no \\@totalrightmargin, so:
+     \hskip-\linewidth \hskip-\@totalleftmargin \hskip\columnwidth}%
+ \MakeFramed {\advance\hsize-\width
+   \@totalleftmargin\z@ \linewidth\hsize
+   \@setminipage}}%
+ {\par\unskip\endMakeFramed%
+ \at@end@of@kframe}
+\makeatother
+
+\definecolor{shadecolor}{rgb}{.97, .97, .97}
+\definecolor{messagecolor}{rgb}{0, 0, 0}
+\definecolor{warningcolor}{rgb}{1, 0, 1}
+\definecolor{errorcolor}{rgb}{1, 0, 0}
+\newenvironment{knitrout}{}{} % an empty environment to be redefined in TeX
+
+\usepackage{alltt} %leqno: numéro d'équation à gauche
 \pagenumbering{arabic} % choose how to number the pages
@@
 \usepackage{longtable}
+\usepackage[table]{xcolor}
 \setlength{\LTleft}{-5cm plus 1 fill}
@@
 %\linenumbers
-\usepackage{caption2}
 \newcommand{\logit}{\text{logit}}
@@
 \title{Combining global tree cover loss data with historical national
-  forest-cover maps to look at six decades of deforestation and forest
+  forest cover maps to look at six decades of deforestation and forest
   fragmentation in Madagascar}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
+\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
 \begin{document}
@@
 
-\vspace{0.3cm}
-
-{\small
-   \begin{flushleft}
-     \textbf{Highlights:}\\
-     - We produced new 30 m resolution forest-cover maps for Madagascar for the period 2000-2014.\\
-     - Madagascar has lost 44\% of its natural forest-cover over the period 1953-2014.\\
-     - Annual deforestation rate has increased in Madagascar since 2005 to reach 1.1\%/yr in 2014.\\
-     - Half of the tropical forest in Madagascar is now located at less than 100~m from forest edge.\\
-     - Conservation efforts must be intensified to save Madagascar unique forests and biodiversity.\\
-   \end{flushleft}
- }
+% \vspace{0.3cm}
+
+% {\small
+%   \begin{flushleft}
+%     \textbf{Highlights:}\\
+%     - We produced new 30 m resolution forest cover maps for Madagascar for the period 2000-2014.\\
+%     - Madagascar has lost 44\% of its natural forest cover over the period 1953-2014.\\
+%     - Annual deforestation rate has increased in Madagascar since 2005 to reach 1.1\%/yr in 2014.\\
+%     - Half of the tropical forest in Madagascar is now located at less than 100~m from forest edge.\\
+%     - Conservation efforts must be intensified to save Madagascar unique forests and biodiversity.\\
+%   \end{flushleft}
+% }
 
@@
 
-  The island of Madagascar has an unparalleled biodiversity, mainly
-  located in the tropical forests of the island, which is highly
-  threatened by anthropogenic deforestation. Scattered forest maps
-  from past studies at national level with substantial gaps (due to
-  presence of cloud cover on satellite imagery) prevent the analysis
-  of long-term deforestation trends in Madagascar. In this study, we
-  propose a new approach combining historical (1953-2000) national
-  forest-cover maps with recent (2001-2014) global annual tree cover
-  loss data to look at six decades (1953-2014) of deforestation and
-  forest fragmentation in Madagascar. We produced new forest-cover
-  maps at 30~m resolution over the full territory of Madagascar for
-  the year 1990, and annually from 2000 to 2014. We estimated that
-  Madagascar has lost 44\% of its natural forest cover over the period
-  1953-2014 (including 37\% over the period 1973-2014). Natural
-  forests cover 8.9 Mha in 2014 (15\% of the national territory) which
-  are divided into 4.4 Mha (50\%) of moist forests, 2.6 Mha (29\%) of
-  dry forests, 1.7 Mha of spiny forests (19\%) and 177,000 ha (2\%) of
-  mangroves. Since 2005, the annual deforestation rate has
-  progressively increased in Madagascar to reach 99,000 ha/yr during
-  2010-2014 (corresponding to a rate of 1.1\%/yr). This increase is
-  probably due to rapid population growth (close to 3\%/yr) and to
-  poor law enforcement in the country. Around half of the forest
-  (46\%) is now located at less than 100~m from the forest
-  edge. Accurate forest-cover change maps can be used to assess the
+  The island of Madagascar has a unique biodiversity, mainly located
+  in the tropical forests of the island. This biodiversity is
+  highly threatened by anthropogenic deforestation. Existing
+  historical forest maps at national level are scattered and have
+  substantial gaps which prevent an exhaustive assessment of long-term
+  deforestation trends in Madagascar. In this study, we combine
+  historical national forest cover maps (covering the period
+  1953-2000) with a recent global annual tree cover loss dataset
+  (2001-2014) to look at six decades of deforestation and forest
+  fragmentation in Madagascar (from 1953 to 2014). We produced new
+  forest cover maps at 30~m resolution for the year 1990 and annually
+  from 2000 to 2014 over the full territory of Madagascar. We
+  estimated that Madagascar has lost 44\% of its natural forest cover
+  over the period 1953-2014 (including 37\% over the period
+  1973-2014). Natural forests cover 8.9 Mha in 2014 (15\% of the
+  national territory) and include 4.4 Mha (50\%) of moist forests, 2.6
+  Mha (29\%) of dry forests, 1.7 Mha of spiny forests (19\%) and
+  177,000 ha (2\%) of mangroves. Since 2005, the annual deforestation
+  rate has progressively increased in Madagascar to reach 99,000 ha/yr
+  during 2010-2014 (corresponding to a rate of 1.1\%/yr). This
+  increase is probably due to rapid population growth (close to
+  3\%/yr) and to poor law enforcement in the country. Around half of
+  the forest (46\%) is now located at less than 100~m from the forest
+  edge. Accurate forest cover change maps can be used to assess the
   effectiveness of past and current conservation programs and
@@
 \citep{Saatchi2011, Vieilledent2016}. As a consequence, accurate and
-spatially explicit maps of forest-cover and forest-cover change are
+spatially explicit maps of forest cover and forest cover change are
 necessary to monitor biodiversity loss and carbon emissions from
@@
 strategies for the future \citep{Vieilledent2013,
-  Vieilledent2016}. Simple time-series of forest-cover estimates, such
+  Vieilledent2016}. Simple time-series of forest cover estimates, such
 as those provided by the FAO Forest Resource Assessment report
@@
 
-Unfortunately, accurate and exhaustive forest-cover maps are not
-available for Madagascar for the last fifteen years (2000-2015).
-\citet{Harper2007} produced maps of forest cover and forest cover
-changes over Madagascar for the years \emph{c.}~1953, \emph{c.}~1973,
-1990 and 2000. The \emph{c.}~1953 forest map was derived from the
-visual interpretation of aerial photography at coarse scale
-(1/1,000,000). Forest maps for the years \emph{c.}~1973, 1990, and
-2000 were obtained from supervised classification of Landsat satellite
-images at 60~m resolution (for the year 1973) or 30~m resolution (for
-years 1990 and 2000) and can be used to derive more accurate estimates
-of forest cover (89.5\% accuracy reported for the forest/non-forest
-map of year 2000). Nonetheless, maps provided by \citet{Harper2007}
-are not exhaustive (due to the presence of clouds in the satellite
-imagery), e.g.~11 244~km2 are mapped as unknown cover type for the
-year 2000. Using a similar supervised classification approach as in
+Unfortunately, accurate and exhaustive forest cover maps are not
+available for Madagascar after year 2000. \citet{Harper2007} produced
+maps of forest cover and forest cover changes over Madagascar for the
+years \emph{c.}~1953, \emph{c.}~1973, 1990 and 2000. The
+\emph{c.}~1953 forest map was derived from the visual interpretation
+of aerial photography at coarse scale (1/1,000,000). Forest maps for
+the years \emph{c.}~1973, 1990, and 2000 were obtained from supervised
+classification of Landsat satellite images at 60~m resolution (for the
+year 1973) or 30~m resolution (for years 1990 and 2000) and can be
+used to derive more accurate estimates of forest cover (89.5\%
+accuracy reported for the forest/non-forest map of year
+2000). Nonetheless, maps provided by \citet{Harper2007} are not
+exhaustive (due to the presence of clouds in the satellite imagery),
+e.g.~11 244~km2 are mapped as unknown cover type for the year
+2000. Using a similar supervised classification approach as in
 \citet{Harper2007}, more recent maps have been produced for the
@@
 support of international environmental NGOs \citep{MEFT2009,
-  ONE2013}. Another set of recent forest-cover maps using an advanced
+  ONE2013}. Another set of recent forest cover maps using an advanced
 statistical tool for classification, the Random Forest classifier
@@
 and can be tested at the national scale for
-Madagascar. \citet{Kim2014} produced a global forest-cover change map
+Madagascar. \citet{Kim2014} produced a global forest cover change map
 from 1990 to 2000 (derived from Landsat imagery). This product was
 updated to cover the period 1975-2005
-(\url{http://glcf.umd.edu/data/landsatFCC/}) but forest-cover maps
+(\url{http://glcf.umd.edu/data/landsatFCC/}) but forest cover maps
 after 2005 were not produced. Moreover, the approach used in
@@
 products from \citet{Hansen2013} to derive annual wall-to-wall
-forest-cover change maps over the period 2000-2014 for Madagascar. We
-use the forest-cover map provided by \citet{Harper2007} for the year
+forest cover change maps over the period 2000-2014 for Madagascar. We
+use the forest cover map provided by \citet{Harper2007} for the year
 2000 (defining the land-use) with the tree cover loss product provided
@@
 
-\subsection{Creation of new forest-cover maps of Madagascar from
+\subsection{Creation of new forest cover maps of Madagascar from
 1953 to 2014}
@@
 map at 30~m resolution using a bilinear interpolation. We thus
-obtained a forest-cover map for the year 2000 covering the full
-territory of Madagascar. We then combined this forest-cover map of the
+obtained a forest cover map for the year 2000 covering the full
+territory of Madagascar. We then combined this forest cover map of the
 year 2000 with the annual tree cover loss maps from 2001 to 2014
-provided by \citet{Hansen2013} to create annual forest-cover maps from
+provided by \citet{Hansen2013} to create annual forest cover maps from
 2001 to 2014 at 30~m resolution. To do so, Hansen's tree cover loss
@@
 areas (due to the presence of clouds on satellite images) using our
-forest-cover map of year 2000. To do so, we assumed that if forest was
+forest cover map of year 2000. To do so, we assumed that if forest was
 present in 2000, the pixel was also forested in 1990. The remaining
@@
 for the year 2000. Similarly we completed the Harper's forest map of
-year 1973 by filling unclassified areas using our forest-cover map of
+year 1973 by filling unclassified areas using our forest cover map of
 the year 1990 assuming that if forest was present in 1990, it was also
@@
 pixels for year 1973 corresponded to a significant total area of 3.3
-million ha. We also reprojected the forest-cover map of year 1953 to a
-common projection in order to compare the forest-cover area in 1953
-with forest-cover areas at the following dates. This map was produced
+million ha. We also reprojected the forest cover map of year 1953 to a
+common projection in order to compare the forest cover area in 1953
+with forest cover areas at the following dates. This map was produced
 by scanning a paper map derived from aerial photos, and thus could not
@@
 processing of satellite imagery \citep{Harper2007}. Finally for all
-forest-cover maps from 1973, the isolated single non-forest pixels
+forest cover maps from 1973, the isolated single non-forest pixels
 (i.e.~fully surrounded by forest pixels) were removed, assuming that
@@
 fragmentation. All the resulting maps have been made permanently and
-freely available on Zenodo (a research data repository) at
+freely available on the Zenodo research data repository at
 \url{https://doi.org/10.5281/zenodo.1145785}.
 
-\subsection{Computing forest-cover areas and deforestation rates}
-
-From these new forest-cover maps, we calculated the total forest-cover
+\subsection{Computing forest cover areas and deforestation rates}
+
+From these new forest cover maps, we calculated the total forest cover
 area for seven available years (1953-1973-1990-2000-2005-2010-2014),
@@
 on six decades might not be fully appropriate. As a consequence, our
-estimates of the forest-cover and deforestation rates for mangroves in
+estimates of the forest cover and deforestation rates for mangroves in
 Madagascar must be considered with this limitation.
 
-\subsection{Comparing our forest-cover and deforestation rate
+\subsection{Comparing our forest cover and deforestation rate
 estimates with previous studies}
 
-We compared our estimates of forest-cover and deforestation rates with
+We compared our estimates of forest cover and deforestation rates with
 estimates from the three existing studies at the national scale for
 Madagascar: (i) \citep{Harper2007}, (ii) \citep{MEFT2009} and (iii)
-\citep{ONE2015}. \citet{Harper2007} provides forest-cover and
+\citep{ONE2015}. \citet{Harper2007} provides forest cover and
 deforestation estimates for the periods
-c. 1953-c. 1973-1990-2000. MEFT, USAID, and CI (2009) provides
-estimates for the periods 1990-2000-2005 and ONE, DGF, MNP, WCS, and
-Etc Terra (2015) provides estimates for the periods 2005-2010-2013. To
-compare our forest-cover and deforestation estimates over the same
-time periods, we consider an additional time-period in our study
-(2010-2013) by creating an extra forest-cover map for the year
-2013. We computed the Pearson's correlation coefficient and the root
-mean square error (RMSE) between our forest-cover estimates and
-forest-cover estimates from previous studies for all the dates and
-forest types (including also the total forest cover estimates). For
-previous studies, the computation of annual deforestation rates (in
-\%/yr) is not always detailed and might slightly differ from one study
-to another \citep[see][]{Puyravaud2003}. \citet{Harper2007} also
-provide total deforested areas for the two periods 1973-1990 and
-1990-2000. We converted these values into annual deforested area
-estimates. When annual deforested areas were not reported (for
-1953-1973 in \citet{Harper2007} and in \citet{MEFT2009} and
-\citet{ONE2015}), we computed them from the forest-cover estimates in
-each study. These estimates cannot be corrected from the potential
-bias due to the presence of residual clouds. Forest-cover and
-deforestation rates were then compared between all studies for the
-whole of Madagascar and the four ecoregions. The same ecoregion
-boundaries as in our study were used in \citet{ONE2015} but this was
-not the case for \citet{Harper2007} and \citet{MEFT2009}, which can
-explain a part of the differences between the estimates.
+c. 1953-c. 1973-1990-2000. \citet{MEFT2009} provides estimates for the
+periods 1990-2000-2005 and \citet{ONE2015} provides estimates for the
+periods 2005-2010-2013. To compare our forest cover and deforestation
+estimates over the same time periods, we consider an additional
+time-period in our study (2010-2013) by creating an extra forest cover
+map for the year 2013. We computed the Pearson's correlation
+coefficient and the root mean square error (RMSE) between our
+forest cover estimates and forest cover estimates from previous
+studies for all the dates and forest types (including also the total
+forest cover estimates). For previous studies, the computation of
+annual deforestation rates (in \%/yr) is not always detailed and might
+slightly differ from one study to another
+\citep[see][]{Puyravaud2003}. \citet{Harper2007} also provide total
+deforested areas for the two periods 1973-1990 and 1990-2000. We
+converted these values into annual deforested area estimates. When
+annual deforested areas were not reported (for 1953-1973 in
+\citet{Harper2007} and in \citet{MEFT2009} and \citet{ONE2015}), we
+computed them from the forest cover estimates in each study. These
+estimates cannot be corrected from the potential bias due to the
+presence of residual clouds. forest cover and deforestation rates were
+then compared between all studies for the whole of Madagascar and the
+four ecoregions. The same ecoregion boundaries as in our study were
+used in \citet{ONE2015} but this was not the case for
+\citet{Harper2007} and \citet{MEFT2009}, which can explain a part of
+the differences between the estimates.
 
@@
 
-\subsection{Dynamics of forest cover and deforestation
-intensity}
+\subsection{Forest cover change and deforestation rates}
 
@@
 respectively (Fig.~\ref{fig:fcc} and Tab.~\ref{tab:forest_cover}). In
-2014 the remaining 8.9 Mha of natural forest were distributed as follow: 4.4
-Mha of moist forest (50\% of total forest cover), 2.6 Mha of dry
-forest (29\%), 1.7 Mha of spiny forest (19\%) and 0.18 Mha (2\%) of
-mangrove forest (Fig.~\ref{fig:ecoregion} and
+2014 the remaining 8.9 Mha of natural forest were distributed as
+follow: 4.4 Mha of moist forest (50\% of total forest cover), 2.6 Mha
+of dry forest (29\%), 1.7 Mha of spiny forest (19\%) and 0.18 Mha
+(2\%) of mangrove forest (Fig.~\ref{fig:ecoregion} and
 Tab.~\ref{tab:comp_forest}). Regarding the deforestation trend, we
@@
 2010-2014 (99,000 ha/yr, 1.1\%/yr) compared to 2000-2005
-(Tab.~\ref{tab:forest_cover}). The deforestation trend characterized
+(Tab.~\ref{tab:forest_cover}). The deforestation trend, characterized
 by a progressive decrease of the deforestation rate over the period
-1990-2005 and a progressive increase of the deforestation after 2005
-is valid for all four ecoregions (Tab.~\ref{tab:comp_defor}), with the
-exception of the spiny forest domain for which the deforestation rate
-during the period 2010-2013 was lower than during 2005-2010
-(Tab.~\ref{tab:comp_defor}).
-
-\subsection{Comparison with previous forest-cover change studies
+1990-2005 and a progressive increase of the deforestation after 2005,
+is valid for all four forest types except the spiny forest
+(Tab.~\ref{tab:comp_defor}). For the spiny forest, the deforestation
+rate during the period 2010-2013 was lower than on the period
+2005-2010 (Tab.~\ref{tab:comp_defor}).
+
+\subsection{Comparison with previous forest cover change studies
 in Madagascar}
 
-Forest-cover maps provided by previous studies over Madagascar were
+Forest cover maps provided by previous studies over Madagascar were
 not exhaustive (unclassified areas) due to the presence of clouds on
@@
 \citet{MEFT2009} and \citet{ONE2015}. With our approach, we produced
-wall to wall forest-cover change maps from 1990 to 2014 for the full
+wall-to-wall forest cover change maps from 1990 to 2014 for the full
 territory of Madagascar (Tab.~\ref{tab:forest_cover}). This allowed us
-to produce more robust estimates of forest-cover and deforestation
-rates over this period. Our forest-cover estimates over the period
+to produce more robust estimates of forest cover and deforestation
+rates over this period. Our forest cover estimates over the period
 1953-2013 (considering forest cover estimates at national level and by
@@
 
-\subsection{Benefits of the combined use of recent global annual
-  tree cover loss data with historical national forest-cover maps}
+\subsection{Advantages of combining recent global annual tree cover loss
+  data with historical national forest cover maps}
 
@@
 loss data \citep{Hansen2013} with historical (1953-2000) national
-forest-cover maps \citep{Harper2007} to look at six decades
+forest cover maps \citep{Harper2007} to look at six decades
 (1953-2014) of deforestation and forest fragmentation in
-Madagascar. We produced annual forest-cover maps at 30~m resolution
+Madagascar. We produced annual forest cover maps at 30~m resolution
 covering Madagascar for the period 2000 to 2014. Our study extends the
-forest-cover monitoring on a six decades period (from 1953 to 2014)
+forest cover monitoring on a six decades period (from 1953 to 2014)
 while harmonizing the data from previous studies \citep{Harper2007,
@@
 forest/non-forest map \citep{Tropek2014}. We propose to use a
-historical national forest-cover map, based on a national forest
+historical national forest cover map, based on a national forest
 definition, as a forest cover mask. This approach could be easily
 extended to other regions or countries for which an accurate
-forest-cover map is available at any date within the period 2000-2014,
+forest cover map is available at any date within the period 2000-2014,
 but preferably at the beginning of the period to profit from the full
@@
 
-The accuracy of the derived forest-cover change maps depends directly
-on the accuracies of the historical forest-cover maps and the tree
+The accuracy of the derived forest cover change maps depends directly
+on the accuracies of the historical forest cover maps and the tree
 cover loss dataset. Using visual-interpretation of aerial images in
@@
 \citep{Achard2014} and they found a good agreement between the two
-sets of estimates. Therefore, our forest-cover change maps after 2000
+sets of estimates. Therefore, our forest cover change maps after 2000
 might be more accurate for the dense moist forest than for the dry and
@@
 spatial variability of the deforestation is well represented. A proper
-accuracy assessment of our forest-cover change maps should be
+accuracy assessment of our forest cover change maps should be
 performed to better estimate the uncertainty surrounding our
-forest-cover change estimates in Madagascar from year 2000
+forest cover change estimates in Madagascar from year 2000
 \citep{Olofsson2013,Olofsson2014}. Despite this limitation, we have
@@
 
-\subsection{Dynamics of forest-cover in Madagascar from 1953 to 2014}
-
-We estimated that natural forests in Madagascar cover 8.9 Mha in 2014
+\subsection{Natural forest cover change in Madagascar from 1953 to 2014}
+
+We estimated that natural forest in Madagascar covers 8.9 Mha in 2014
 (corresponding to 15\% of the country) and that Madagascar has lost
@@
 correction for the presence of clouds. Our deforestation rate
-estimates from 1990 to 2014 have been computed from wall to wall maps
+estimates from 1990 to 2014 have been computed from wall-to-wall maps
 at 30~m resolution and can be considered more accurate in comparison
-with estimates from these previous studies. Our forest-cover and
+with estimates from these previous studies. Our forest cover and
 deforestation rate estimates can be used as source of information for
@@
 dramatically increased at the national level since 2005. Results of
-this study, including recent spatially explicit forest-cover change
-maps and forest-cover estimates, should help implement new
+this study, including recent spatially explicit forest cover change
+maps and forest cover estimates, should help implement new
 conservation strategies to save Madagascar natural tropical forests
@@
 
-Authors thank Jean-François Bastin for useful comments on a previous
-version of the manuscript and Peter Vogt for fruitfull discussion on
-which metric to use to compute forest fragmentation. This study is
-part of the Cirad's BioSceneMada project
-(\url{https://bioscenemada.cirad.fr}) and the Joint Research Center's
-ReCaREDD project (\url{http://forobs.jrc.ec.europa.eu/recaredd}). The
-BioSceneMada project is funded by FRB (Fondation pour la Recherche sur
-la Biodiversité) and the FFEM (Fond Français pour l'Environnement
+The authors thank Jean-François Bastin for useful comments on a previous
+version of the manuscript and Peter Vogt for useful advices on which
+metric to use to estimate forest fragmentation. This study is part of
+the Cirad's BioSceneMada project (\url{https://bioscenemada.cirad.fr})
+and the Joint Research Center's ReCaREDD project
+(\url{http://forobs.jrc.ec.europa.eu/recaredd}). The BioSceneMada
+project is funded by FRB (Fondation pour la Recherche sur la
+Biodiversité) and the FFEM (Fond Français pour l'Environnement
 Mondial) under the project agreement AAP-SCEN-2013 I. The ReCaREDD
-project is funded by the European Commission.
+project is funded by the European Commission. The authors declare that there are no conflicts of interest related to this article.
 
@@
 All the data and the script used for this study have been made permanently and publicly available
-on Zenodo (a research data repository) so that the results are entirely reproducible:
+on the Zenodo research data repository so that the results are entirely reproducible:
 \begin{itemize}
@@
 Kim, D.H., Sexton, J.O., Noojipady, P., Huang, C., Anand, A., Channan, S.,
-  Feng, M. \& Townshend, J.R. (2014) {Global, Landsat-based forest-cover change
+  Feng, M. \& Townshend, J.R. (2014) {Global, Landsat-based forest cover change
   from 1990 to 2000}. \emph{Remote Sensing of Environment} \textbf{155},
@@
 \vfill
+\rowcolors{2}{gray!6}{white}
 \begin{table}[!h]
-  \begin{longtable}[]{@{}rrrrr@{}}
+
+\caption{\label{tab:forest_cover}\textbf{Evolution of natural forest cover and deforestation rates from 1953 to 2014 in Madagascar.} Areas are provided in thousands of hectares (Kha).
+Forest map for the year 1973 has 3.3 Mha of unclassified areas due to the presence of clouds on satellite images.
+As a consequence, partial deforestation rates for the periods 1953-1973 and 1973-1990 are computed based on the available forest extent.
+The last two columns indicate the annual deforested areas and annual deforestation rates on the previous time-period (e.g.~1953-1973 for year 1973, 1973-1990 for year 1990, etc.).}
+\centering
+\begin{tabular}[t]{lrrrr}
+\hiderowcolors
 \toprule
-    Year & Forest (ha) & Unmapped (ha) & Defor. area (ha/yr) & Rate
-    (\%/yr) \\
+Year & Forest (Kha) & Unmap (Kha) & Annual defor. (Kha/yr) & Rate (\%/yr)\\
 \midrule
-    \endhead
-    1953 & 15,968 & 0 & - & - \\
-    1973 & 14,243 & 3,317 & 86 & 0.6 \\
-    1990 & 10,762 & 0 & 205 & 1.6 \\
-    2000 & 9,879 & 0 & 88 & 0.8 \\
-    2005 & 9,668 & 0 & 42 & 0.4 \\
-    2010 & 9,320 & 0 & 70 & 0.7 \\
-    2014 & 8,925 & 0 & 99 & 1.1 \\
+\showrowcolors
+1953 & 15,968 & 0 & - & -\\
+1973 & 14,243 & 3,317 & 86 & 0.6\\
+1990 & 10,762 & 0 & 205 & 1.6\\
+2000 & 9,879 & 0 & 88 & 0.8\\
+2005 & 9,668 & 0 & 42 & 0.4\\
+2010 & 9,320 & 0 & 70 & 0.7\\
+2014 & 8,925 & 0 & 99 & 1.1\\
 \bottomrule
-  \end{longtable}
-  \addtocounter{table}{-1}
-
-  \caption{\textbf{Evolution of forest cover and deforestation rates
-      from 1953 to 2014 in Madagascar}. Areas are provided in
-    thousands of hectares (Kha). Forest map for the year 1973 has 3.3
-    Mha of unclassified areas due to the presence of clouds on
-    satellite images. As a consequence, partial deforestation rates
-    for the periods 1953-1973 and 1973-1990 are computed based on the
-    available forest extent. The last two columns indicate the annual
-    deforested areas and annual deforestation rates on the previous
-    time-period (e.g.~1953-1973 for year 1973, 1973-1990 for year
-    1990, etc.).}
-
-  \label{tab:forest_cover}
+\end{tabular}
 \end{table}
+\rowcolors{2}{white}{white}
+
 \vfill
@@
 \vspace*{\stretch{1}}
+\rowcolors{2}{gray!6}{white}
 \begin{table}[!h]
-  %{\small
-  \begin{longtable}[]{@{}llrrrrrrrr@{}}
-    \toprule
-    Forest type & Source & 1953 & 1973 & 1990 & 2000 & 2005 & 2010 & 2013 &
-    2014 \\
-    \midrule
-    \endhead
-    Total & Harper2007 & 15,996 & 14,173 & 10,606 & 8,982 & - & - & - & - \\
-          & MEFT2009 & - & - & 10,650 & 9,678 & 9,413 & - & - & - \\
-                & ONE2015 & - & - & - & - & 9,451 & 8,977 & 8,486 & - \\
-                & this study & 15,968 & 14,243 & 10,762 & 9,879 & 9,668 & 9,320 & 9,051
-                                                                          & 8,925 \\
-    Moist & Harper2007 & 8,766 & 6,876 & 5,234 & 4,167 & - & - & - &
-                                                                     - \\
-                & MEFT2009 & - & - & 5,271 & 4,788 & 4,700 & - & - & - \\
-                & ONE2015 & - & - & - & - & 4,556 & 4,457 & 4,345 & - \\
-                & this study & 8,578 & 6,990 & 5,270 & 4,872 & 4,768 & 4,633 & 4,470 &
-                                                                                       4,410 \\
-    Dry & Harper2007 & 4,252 & 4,028 & 2,712 & 2,457 & - & - & - &
-                                                                   - \\
-                & MEFT2009 & - & - & 3,321 & 3,085 & 3,028 & - & - & - \\
-                & ONE2015 & - & - & - & - & 3,223 & 2,970 & 2,679 & - \\
-                & this study & 4,762 & 4,435 & 3,225 & 2,941 & 2,881 & 2,735 & 2,642 &
-                                                                                       2,596 \\
-    Spiny & Harper2007 & 2,978 & 3,030 & 2,420 & 2,132 & - & - & - &
-                                                                     - \\
-                & MEFT2009 & - & - & 2,124 & 1,872 & 1,757 & - & - & - \\
-                & ONE2015 & - & - & - & - & 1,682 & 1,559 & 1,467 & - \\
-                & this study & 2,463 & 2,583 & 2,055 & 1,858 & 1,811 & 1,744 & 1,731 &
-                                                                                       1,713 \\
-    Mangroves & Harper2007 & - & - & 240 & 226 & - & - & - &
-                                                             - \\
-                & MEFT2009 & - & - & - & - & - & - & - & - \\
-                & ONE2015 & - & - & - & - & 174 & 171 & 170 & - \\
-                & this study & 143 & 200 & 181 & 178 & 177 & 177 & 177 &
-                                                                         177 \\
-    \bottomrule
-  \end{longtable}%}
-  \addtocounter{table}{-1}
-
-  \caption{\textbf{Comparing Madagascar forest-cover estimates (in
-      Kha) with previous studies on the period 1953-2014}. We compared
-    our estimates of forest-cover with the estimates from three
-    previous studies \citep{Harper2007, MEFT2009, ONE2015}. Areas are
-    provided in thousands of hectares (Kha). We obtained a Pearson's
-    correlation coefficient of 0.99 between our forest-cover estimates
-    and forest-cover estimates from previous studies. The increase in
+
+\caption{\label{tab:comp_forest}\textbf{Comparing Madagascar forest cover estimates with
+      previous studies on the period 1953-2014}. We compared our
+    estimates of forest cover with the estimates from three previous
+    studies \citep{Harper2007, MEFT2009, ONE2015}. Areas are provided
+    in thousands of hectares (Kha). We obtained a Pearson's
+    correlation coefficient of 0.99 between our forest cover estimates
+    and forest cover estimates from previous studies. The increase in
     mangrove and spiny forest covers from \emph{c.} 1953 to \emph{c.}
@@
     derived map.}
-
-  \label{tab:comp_forest}
+\centering
+\begin{tabular}[t]{llrrrrrrrr}
+\hiderowcolors
+\toprule
+Forest type & Source & 1953 & 1973 & 1990 & 2000 & 2005 & 2010 & 2013 & 2014\\
+\midrule
+\showrowcolors
+Total & Harper2007 & 15,996 & 14,173 & 10,606 & 8,982 & - & - & - & -\\
+ & MEFT2009 & - & - & 10,650 & 9,678 & 9,413 & - & - & -\\
+ & ONE2015 & - & - & - & - & 9,451 & 8,977 & 8,486 & -\\
+ & this study & 15,968 & 14,243 & 10,762 & 9,879 & 9,668 & 9,320 & 9,051 & 8,925\\
+Moist & Harper2007 & 8,766 & 6,876 & 5,234 & 4,167 & - & - & - & -\\
+ & MEFT2009 & - & - & 5,271 & 4,788 & 4,700 & - & - & -\\
+ & ONE2015 & - & - & - & - & 4,556 & 4,457 & 4,345 & -\\
+ & this study & 8,578 & 6,990 & 5,270 & 4,872 & 4,768 & 4,633 & 4,470 & 4,410\\
+Dry & Harper2007 & 4,252 & 4,028 & 2,712 & 2,457 & - & - & - & -\\
+ & MEFT2009 & - & - & 3,321 & 3,085 & 3,028 & - & - & -\\
+ & ONE2015 & - & - & - & - & 3,223 & 2,970 & 2,679 & -\\
+ & this study & 4,762 & 4,435 & 3,225 & 2,941 & 2,881 & 2,735 & 2,642 & 2,596\\
+Spiny & Harper2007 & 2,978 & 3,030 & 2,420 & 2,132 & - & - & - & -\\
+ & MEFT2009 & - & - & 2,124 & 1,872 & 1,757 & - & - & -\\
+ & ONE2015 & - & - & - & - & 1,682 & 1,559 & 1,467 & -\\
+ & this study & 2,463 & 2,583 & 2,055 & 1,858 & 1,811 & 1,744 & 1,731 & 1,713\\
+Mangroves & Harper2007 & - & - & 240 & 226 & - & - & - & -\\
+ & MEFT2009 & - & - & - & - & - & - & - & -\\
+ & ONE2015 & - & - & - & - & 174 & 171 & 170 & -\\
+ & this study & 143 & 200 & 181 & 178 & 177 & 177 & 177 & 177\\
+\bottomrule
+\end{tabular}
 \end{table}
+\rowcolors{2}{white}{white}
+
 \vspace*{\stretch{1}}
@@
 \vspace*{\stretch{1}}
+\rowcolors{2}{gray!6}{white}
 \begin{table}[!h]
-  %{\small
-  \begin{longtable}[]{@{}llrrrrrr@{}}
+
+\caption{\label{tab:comp_defor}\textbf{Comparing Madagascar annual deforestation rates
+      with previous studies on the period 1953-2014}. Annual
+    deforested areas (in thousands of hectares per year, Kha/yr) and
+    annual deforestation rates (second number in parenthesis, in
+    \%/yr) are provided. For deforestation rates in \%/yr, exact same
+    numbers as in scientific articles and reports from previous
+    studies \citep{Harper2007, MEFT2009, ONE2015} have been
+    reported. The way annual deforestation rates in \%/yr have been
+    computed in these previous studies can slightly differ from one
+    study to another, but estimates always correct for the potential
+    presences of clouds on satellite images and unclassified areas on
+    forest maps. Annual deforested areas in Kha/yr have been
+    recomputed from forest cover estimates in
+    Tab.~\ref{tab:comp_forest} (except for \citet{Harper2007} for the
+    periods 1973-1990 and 1990-2000 for which annual deforested areas
+    in Kha/yr were derived from numbers reported in the original
+    publication, see methods) and do not correct for the potential
+    presence of clouds.}
+\centering
+\begin{tabular}[t]{llrrrrrr}
+\hiderowcolors
 \toprule
-    Forest type & Source & 1953-1973 & 1973-1990 & 1990-2000 & 2000-2005 &
-    2005-2010 & 2010-2013 \\
+Forest type & Source & 1953-1973 & 1973-1990 & 1990-2000 & 2000-2005 & 2005-2010 & 2010-2013\\
 \midrule
-    \endhead
-    Total & Harper2007 & 91 (0.3) & 200 (1.7) & 81 (0.9) & - & - &
-                                                                   - \\
-                & MEFT2009 & - & - & 97 (0.8) & 53 (0.5) & - & - \\
-                & ONE2015 & - & - & - & - & 95 (1.2) & 164 (1.5) \\
-                & this study & 86 (0.6) & 205 (1.6) & 88 (0.9) & 42 (0.4) & 70 (0.7) &
-                                                                                       90 (1.0) \\
-    Moist & Harper2007 & 94 (0.6) & 87 (1.7) & 32 (0.8) & - & - &
-                                                                  - \\
-                & MEFT2009 & - & - & 48 (0.8) & 17 (0.4) & - & - \\
-                & ONE2015 & - & - & - & - & 20 (0.5) & 37 (0.9) \\
-                & this study & 79 (1.0) & 101 (1.6) & 40 (0.8) & 21 (0.4) & 27 (0.6) &
-                                                                                       54 (1.2) \\
-    Dry & Harper2007 & 11 (0.2) & 77 (1.9) & 20 (0.7) & - & - &
-                                                                - \\
-                & MEFT2009 & - & - & 24 (0.7) & 11 (0.4) & - & - \\
-                & ONE2015 & - & - & - & - & 51 (1.8) & 97 (2.3) \\
-                & this study & 16 (0.4) & 71 (1.9) & 28 (0.9) & 12 (0.4) & 29 (1.0) & 31
-                                                                                      (1.1) \\
-    Spiny & Harper2007 & -3 (-0.1) & 36 (1.2) & 28 (1.2) & - & - &
-                                                                   - \\
-                & MEFT2009 & - & - & 25 (1.2) & 23 (1.2) & - & - \\
-                & ONE2015 & - & - & - & - & 25 (1.7) & 31 (1.7) \\
-                & this study & -6 (-0.2) & 31 (1.3) & 20 (1.0) & 9 (0.5) & 13 (0.7) & 4
-                                                                                      (0.3) \\
-    Mangroves & Harper2007 & - & - & 1 (0.2) & - & - & - \\
-                & MEFT2009 & - & - & - & - & - & - \\
-                & ONE2015 & - & - & - & - & 0 (0.3) & 0 (0.2) \\
-                & this study & -3 (-1.7) & 1 (0.6) & 0 (0.2) & 0 (0.0) & 0 (0.0) & 0
-                                                                                   (0.0) \\
+\showrowcolors
+Total & Harper2007 & 91 (0.3) & 200 (1.7) & 81 (0.9) & - & - & -\\
+ & MEFT2009 & - & - & 97 (0.8) & 53 (0.5) & - & -\\
+ & ONE2015 & - & - & - & - & 95 (1.2) & 164 (1.5)\\
+ & this study & 86 (0.6) & 205 (1.6) & 88 (0.9) & 42 (0.4) & 70 (0.7) & 90 (1.0)\\
+Moist & Harper2007 & 94 (0.6) & 87 (1.7) & 32 (0.8) & - & - & -\\
+ & MEFT2009 & - & - & 48 (0.8) & 17 (0.4) & - & -\\
+ & ONE2015 & - & - & - & - & 20 (0.5) & 37 (0.9)\\
+ & this study & 79 (1.0) & 101 (1.6) & 40 (0.8) & 21 (0.4) & 27 (0.6) & 54 (1.2)\\
+Dry & Harper2007 & 11 (0.2) & 77 (1.9) & 20 (0.7) & - & - & -\\
+ & MEFT2009 & - & - & 24 (0.7) & 11 (0.4) & - & -\\
+ & ONE2015 & - & - & - & - & 51 (1.8) & 97 (2.3)\\
+ & this study & 16 (0.4) & 71 (1.9) & 28 (0.9) & 12 (0.4) & 29 (1.0) & 31 (1.1)\\
+Spiny & Harper2007 & -3 (-0.1) & 36 (1.2) & 28 (1.2) & - & - & -\\
+ & MEFT2009 & - & - & 25 (1.2) & 23 (1.2) & - & -\\
+ & ONE2015 & - & - & - & - & 25 (1.7) & 31 (1.7)\\
+ & this study & -6 (-0.2) & 31 (1.3) & 20 (1.0) & 9 (0.5) & 13 (0.7) & 4 (0.3)\\
+Mangroves & Harper2007 & - & - & 1 (0.2) & - & - & -\\
+ & MEFT2009 & - & - & - & - & - & -\\
+ & ONE2015 & - & - & - & - & 0 (0.3) & 0 (0.2)\\
+ & this study & -3 (-1.7) & 1 (0.6) & 0 (0.2) & 0 (0.0) & 0 (0.0) & 0 (0.0)\\
 \bottomrule
-  \end{longtable}%}
-  \addtocounter{table}{-1}
-
-  \caption{\textbf{Comparing Madagascar annual deforestation rates
-      with previous studies on the period 1953-2014}. Annual
-    deforestated areas (in Kha/yr) and annual deforestation rates
-    (second number in parenthesis, in \%/yr) are provided. For
-    deforestation rates in \%/yr, exact same numbers as in scientific
-    articles and reports from previous studies \citep{Harper2007,
-      MEFT2009, ONE2015} have been reported. The way annual
-    deforestation rates in \%/yr have been computed in these previous
-    studies can slightly differ from one study to another, but
-    estimates always correct for the potential presences of clouds on
-    satellite images and unclassified areas on forest maps. Annual
-    deforested areas in Kha/yr have been recomputed from forest-cover
-    estimates in Tab.~\ref{tab:comp_forest} (except for
-    \citet{Harper2007} for the periods 1973-1990 and 1990-2000 for
-    which annual deforested areas in Kha/yr were derived from numbers
-    reported in the original publication, see methods) and do not
-    correct for the potential presence of clouds.  }
-
-  \label{tab:comp_defor}
+\end{tabular}
 \end{table}
+\rowcolors{2}{white}{white}
+
 \vspace*{\stretch{1}}
@@
 
-  \caption{\textbf{Forest-cover change on six decades from 1953 to
-      2014 in Madagascar.} Forest cover changes from \emph{c.} 1973 to
+  \caption{\textbf{Forest cover change on six decades from 1953 to
+      2014 in Madagascar.} forest cover changes from \emph{c.} 1973 to
     2014 are shown in the main figure, and forest cover in \emph{c.}
@@
     dry (left part) and eastern moist (right part) ecoregions present
-    more detailed views of (from top to bottom): forest-cover in
-    1950s, forest-cover change from \emph{c.} 1973 to 2014, forest
+    more detailed views of (from top to bottom): forest cover in
+    1950s, forest cover change from \emph{c.} 1973 to 2014, forest
     fragmentation in 2014 and distance to forest edge in 2014. Data on

### 2018-01-17 manuscript/manuscript.Rnw
Rmd to Rnw

--- 2018-01-17 manuscript/manuscript.tex
@@
 
-\documentclass[a4paper, 12pt, leqno]{article}\usepackage[]{graphicx}\usepackage[]{color}
-%% maxwidth is the original width if it is less than linewidth
-%% otherwise use linewidth (to make sure the graphics do not exceed the margin)
-\makeatletter
-\def\maxwidth{ %
-  \ifdim\Gin@nat@width>\linewidth
-    \linewidth
-  \else
-    \Gin@nat@width
-  \fi
-}
-\makeatother
-
-\definecolor{fgcolor}{rgb}{0.345, 0.345, 0.345}
-\newcommand{\hlnum}[1]{\textcolor[rgb]{0.686,0.059,0.569}{#1}}%
-\newcommand{\hlstr}[1]{\textcolor[rgb]{0.192,0.494,0.8}{#1}}%
-\newcommand{\hlcom}[1]{\textcolor[rgb]{0.678,0.584,0.686}{\textit{#1}}}%
-\newcommand{\hlopt}[1]{\textcolor[rgb]{0,0,0}{#1}}%
-\newcommand{\hlstd}[1]{\textcolor[rgb]{0.345,0.345,0.345}{#1}}%
-\newcommand{\hlkwa}[1]{\textcolor[rgb]{0.161,0.373,0.58}{\textbf{#1}}}%
-\newcommand{\hlkwb}[1]{\textcolor[rgb]{0.69,0.353,0.396}{#1}}%
-\newcommand{\hlkwc}[1]{\textcolor[rgb]{0.333,0.667,0.333}{#1}}%
-\newcommand{\hlkwd}[1]{\textcolor[rgb]{0.737,0.353,0.396}{\textbf{#1}}}%
-\let\hlipl\hlkwb
-
-\usepackage{framed}
-\makeatletter
-\newenvironment{kframe}{%
- \def\at@end@of@kframe{}%
- \ifinner\ifhmode%
-  \def\at@end@of@kframe{\end{minipage}}%
-  \begin{minipage}{\columnwidth}%
- \fi\fi%
- \def\FrameCommand##1{\hskip\@totalleftmargin \hskip-\fboxsep
- \colorbox{shadecolor}{##1}\hskip-\fboxsep
-     % There is no \\@totalrightmargin, so:
-     \hskip-\linewidth \hskip-\@totalleftmargin \hskip\columnwidth}%
- \MakeFramed {\advance\hsize-\width
-   \@totalleftmargin\z@ \linewidth\hsize
-   \@setminipage}}%
- {\par\unskip\endMakeFramed%
- \at@end@of@kframe}
-\makeatother
-
-\definecolor{shadecolor}{rgb}{.97, .97, .97}
-\definecolor{messagecolor}{rgb}{0, 0, 0}
-\definecolor{warningcolor}{rgb}{1, 0, 1}
-\definecolor{errorcolor}{rgb}{1, 0, 0}
-\newenvironment{knitrout}{}{} % an empty environment to be redefined in TeX
-
-\usepackage{alltt} %leqno: numéro d'équation à gauche
+\documentclass[a4paper, 12pt, leqno]{article} %leqno: numéro d'équation à gauche
 \pagenumbering{arabic} % choose how to number the pages
@@
 
+<<options, echo=FALSE, results="hide">>=
+library(knitr)
+library(kableExtra)
+opts_chunk$set(echo=FALSE, cache=FALSE,
+               #results="hide", 
+               warning=FALSE,
+               message=FALSE, highlight=TRUE,
+               fig.show="hide", size="small",
+               fig.align="center",
+               tidy=FALSE)
+options(knitr.kable.NA="-")
+@
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
+
 \begin{document}
@@
 \vfill
-\rowcolors{2}{gray!6}{white}
-\begin{table}[!h]
-
-\caption{\label{tab:forest_cover}\textbf{Evolution of natural forest cover and deforestation rates from 1953 to 2014 in Madagascar.} Areas are provided in thousands of hectares (Kha).
+<<forest_cover, results="asis">>=
+# Import and format
+defor.df <- read.table(file="tabs/defor.txt", sep="\t", header=TRUE)
+perc.1 <- round(100*(defor.df$area[defor.df$Year==1973]-defor.df$area[defor.df$Year==2014])/defor.df$area[defor.df$Year==1973])
+perc.2 <- round(100*(defor.df$area[defor.df$Year==1953]-defor.df$area[defor.df$Year==2014])/defor.df$area[defor.df$Year==1953])
+d.1 <- defor.df$theta[defor.df$Year==1990]
+d.2 <- defor.df$theta[defor.df$Year==2010]
+tab.1 <- defor.df
+tab.1[,c(2:4)] <- round(tab.1[,c(2:4)]/1000)
+tab.1$Year <- as.character(tab.1$Year)
+names(tab.1) <- c("Year", "Forest (Kha)", "Unmap (Kha)", "Annual defor. (Kha/yr)", "Rate (%/yr)")
+# Caption
+tabcap1 <- "\\textbf{Evolution of natural forest cover and deforestation rates from 1953 to 2014 in Madagascar.} Areas are provided in thousands of hectares (Kha).
 Forest map for the year 1973 has 3.3 Mha of unclassified areas due to the presence of clouds on satellite images.
 As a consequence, partial deforestation rates for the periods 1953-1973 and 1973-1990 are computed based on the available forest extent.
-The last two columns indicate the annual deforested areas and annual deforestation rates on the previous time-period (e.g.~1953-1973 for year 1973, 1973-1990 for year 1990, etc.).}
-\centering
-\begin{tabular}[t]{lrrrr}
-\hiderowcolors
-\toprule
-Year & Forest (Kha) & Unmap (Kha) & Annual defor. (Kha/yr) & Rate (\%/yr)\\
-\midrule
-\showrowcolors
-1953 & 15,968 & 0 & - & -\\
-1973 & 14,243 & 3,317 & 86 & 0.6\\
-1990 & 10,762 & 0 & 205 & 1.6\\
-2000 & 9,879 & 0 & 88 & 0.8\\
-2005 & 9,668 & 0 & 42 & 0.4\\
-2010 & 9,320 & 0 & 70 & 0.7\\
-2014 & 8,925 & 0 & 99 & 1.1\\
-\bottomrule
-\end{tabular}
-\end{table}
-\rowcolors{2}{white}{white}
-
+The last two columns indicate the annual deforested areas and annual deforestation rates on the previous time-period (e.g.~1953-1973 for year 1973, 1973-1990 for year 1990, etc.)."
+# Kable
+knitr::kable(tab.1, digits=1, format.args=list(big.mark=","), booktabs=TRUE, format="latex", 
+             caption=tabcap1, align=c("l","r","r","r","r"), linesep="") %>% 
+  kable_styling(latex_options=c("striped","hold_position"))
+@
 \vfill
@@
 \vspace*{\stretch{1}}
-\rowcolors{2}{gray!6}{white}
-\begin{table}[!h]
-
-\caption{\label{tab:comp_forest}\textbf{Comparing Madagascar forest cover estimates with
+<<comp_forest, results="asis">>=
+# Import and format
+fcc.comp <- read.table(file="tabs/fcc_comp.txt", sep=" ", header=TRUE)
+tab.2 <- fcc.comp
+tab.2 <- tab.2[,c(2,1,3:10)]
+tab.2[,c(3:10)] <- round(tab.2[,c(3:10)]/1000)
+tab.2$ForestType <- c("Total",rep("",3),"Moist",rep(" ",3),"Dry",rep(" ",3),"Spiny",rep(" ",3),"Mangroves",rep(" ",3))
+col.names <- c("Forest type","Source","1953","1973","1990","2000","2005","2010","2013","2014")
+# Caption
+tabcap2 <- "\\textbf{Comparing Madagascar forest cover estimates with
       previous studies on the period 1953-2014}. We compared our
     estimates of forest cover with the estimates from three previous
-    studies \citep{Harper2007, MEFT2009, ONE2015}. Areas are provided
+    studies \\citep{Harper2007, MEFT2009, ONE2015}. Areas are provided
     in thousands of hectares (Kha). We obtained a Pearson's
@@
     and forest cover estimates from previous studies. The increase in
-    mangrove and spiny forest covers from \emph{c.} 1953 to \emph{c.}
-    1973 in \citet{Harper2007} and our study is most probably due to
+    mangrove and spiny forest covers from \\emph{c.} 1953 to \\emph{c.}
+    1973 in \\citet{Harper2007} and our study is most probably due to
     differences in forest definition and mapping methods between the
     1953 aerial-photography derived map and the 1973 Landsat image
-    derived map.}
-\centering
-\begin{tabular}[t]{llrrrrrrrr}
-\hiderowcolors
-\toprule
-Forest type & Source & 1953 & 1973 & 1990 & 2000 & 2005 & 2010 & 2013 & 2014\\
-\midrule
-\showrowcolors
-Total & Harper2007 & 15,996 & 14,173 & 10,606 & 8,982 & - & - & - & -\\
- & MEFT2009 & - & - & 10,650 & 9,678 & 9,413 & - & - & -\\
- & ONE2015 & - & - & - & - & 9,451 & 8,977 & 8,486 & -\\
- & this study & 15,968 & 14,243 & 10,762 & 9,879 & 9,668 & 9,320 & 9,051 & 8,925\\
-Moist & Harper2007 & 8,766 & 6,876 & 5,234 & 4,167 & - & - & - & -\\
- & MEFT2009 & - & - & 5,271 & 4,788 & 4,700 & - & - & -\\
- & ONE2015 & - & - & - & - & 4,556 & 4,457 & 4,345 & -\\
- & this study & 8,578 & 6,990 & 5,270 & 4,872 & 4,768 & 4,633 & 4,470 & 4,410\\
-Dry & Harper2007 & 4,252 & 4,028 & 2,712 & 2,457 & - & - & - & -\\
- & MEFT2009 & - & - & 3,321 & 3,085 & 3,028 & - & - & -\\
- & ONE2015 & - & - & - & - & 3,223 & 2,970 & 2,679 & -\\
- & this study & 4,762 & 4,435 & 3,225 & 2,941 & 2,881 & 2,735 & 2,642 & 2,596\\
-Spiny & Harper2007 & 2,978 & 3,030 & 2,420 & 2,132 & - & - & - & -\\
- & MEFT2009 & - & - & 2,124 & 1,872 & 1,757 & - & - & -\\
- & ONE2015 & - & - & - & - & 1,682 & 1,559 & 1,467 & -\\
- & this study & 2,463 & 2,583 & 2,055 & 1,858 & 1,811 & 1,744 & 1,731 & 1,713\\
-Mangroves & Harper2007 & - & - & 240 & 226 & - & - & - & -\\
- & MEFT2009 & - & - & - & - & - & - & - & -\\
- & ONE2015 & - & - & - & - & 174 & 171 & 170 & -\\
- & this study & 143 & 200 & 181 & 178 & 177 & 177 & 177 & 177\\
-\bottomrule
-\end{tabular}
-\end{table}
-\rowcolors{2}{white}{white}
-
+    derived map."
+# Kable
+knitr::kable(tab.2, col.names=col.names, align=c("l","l",rep("r",8)), format.args=list(big.mark=","), 
+             booktabs=TRUE, format="latex", caption=tabcap2, linesep="") %>% 
+  kable_styling(latex_options=c("striped","hold_position"))
+@
 \vspace*{\stretch{1}}
@@
 \vspace*{\stretch{1}}
-\rowcolors{2}{gray!6}{white}
-\begin{table}[!h]
-
-\caption{\label{tab:comp_defor}\textbf{Comparing Madagascar annual deforestation rates
+<<comp_defor, results="asis">>=
+# Import and format
+tab.3 <- read.table(file="tabs/defor_comp.txt", sep=" ", header=TRUE)
+tab.3$ForestType <- c("Total",rep("",3),"Moist",rep(" ",3),"Dry",rep(" ",3),"Spiny",rep(" ",3),"Mangroves",rep(" ",3))
+col.names <- c("Forest type","Source","1953-1973","1973-1990","1990-2000","2000-2005","2005-2010","2010-2013")
+# Caption
+tabcap3 <- "\\textbf{Comparing Madagascar annual deforestation rates
       with previous studies on the period 1953-2014}. Annual
@@
     annual deforestation rates (second number in parenthesis, in
-    \%/yr) are provided. For deforestation rates in \%/yr, exact same
+    \\%/yr) are provided. For deforestation rates in \\%/yr, exact same
     numbers as in scientific articles and reports from previous
-    studies \citep{Harper2007, MEFT2009, ONE2015} have been
-    reported. The way annual deforestation rates in \%/yr have been
+    studies \\citep{Harper2007, MEFT2009, ONE2015} have been
+    reported. The way annual deforestation rates in \\%/yr have been
     computed in these previous studies can slightly differ from one
@@
     recomputed from forest cover estimates in
-    Tab.~\ref{tab:comp_forest} (except for \citet{Harper2007} for the
+    Tab.~\\ref{tab:comp_forest} (except for \\citet{Harper2007} for the
     periods 1973-1990 and 1990-2000 for which annual deforested areas
@@
     publication, see methods) and do not correct for the potential
-    presence of clouds.}
-\centering
-\begin{tabular}[t]{llrrrrrr}
-\hiderowcolors
-\toprule
-Forest type & Source & 1953-1973 & 1973-1990 & 1990-2000 & 2000-2005 & 2005-2010 & 2010-2013\\
-\midrule
-\showrowcolors
-Total & Harper2007 & 91 (0.3) & 200 (1.7) & 81 (0.9) & - & - & -\\
- & MEFT2009 & - & - & 97 (0.8) & 53 (0.5) & - & -\\
- & ONE2015 & - & - & - & - & 95 (1.2) & 164 (1.5)\\
- & this study & 86 (0.6) & 205 (1.6) & 88 (0.9) & 42 (0.4) & 70 (0.7) & 90 (1.0)\\
-Moist & Harper2007 & 94 (0.6) & 87 (1.7) & 32 (0.8) & - & - & -\\
- & MEFT2009 & - & - & 48 (0.8) & 17 (0.4) & - & -\\
- & ONE2015 & - & - & - & - & 20 (0.5) & 37 (0.9)\\
- & this study & 79 (1.0) & 101 (1.6) & 40 (0.8) & 21 (0.4) & 27 (0.6) & 54 (1.2)\\
-Dry & Harper2007 & 11 (0.2) & 77 (1.9) & 20 (0.7) & - & - & -\\
- & MEFT2009 & - & - & 24 (0.7) & 11 (0.4) & - & -\\
- & ONE2015 & - & - & - & - & 51 (1.8) & 97 (2.3)\\
- & this study & 16 (0.4) & 71 (1.9) & 28 (0.9) & 12 (0.4) & 29 (1.0) & 31 (1.1)\\
-Spiny & Harper2007 & -3 (-0.1) & 36 (1.2) & 28 (1.2) & - & - & -\\
- & MEFT2009 & - & - & 25 (1.2) & 23 (1.2) & - & -\\
- & ONE2015 & - & - & - & - & 25 (1.7) & 31 (1.7)\\
- & this study & -6 (-0.2) & 31 (1.3) & 20 (1.0) & 9 (0.5) & 13 (0.7) & 4 (0.3)\\
-Mangroves & Harper2007 & - & - & 1 (0.2) & - & - & -\\
- & MEFT2009 & - & - & - & - & - & -\\
- & ONE2015 & - & - & - & - & 0 (0.3) & 0 (0.2)\\
- & this study & -3 (-1.7) & 1 (0.6) & 0 (0.2) & 0 (0.0) & 0 (0.0) & 0 (0.0)\\
-\bottomrule
-\end{tabular}
-\end{table}
-\rowcolors{2}{white}{white}
-
+    presence of clouds."
+# Kable
+knitr::kable(tab.3, col.names=col.names, align=c("l","l",rep("r",6)), booktabs=TRUE, 
+             caption=tabcap3, linesep="") %>% 
+  kable_styling(latex_options=c("striped","hold_position"))
+@
 \vspace*{\stretch{1}}

### 2018-01-18 manuscript/manuscript.Rnw
Format Rnw

@@
 \begin{center}
-{\large
   Ghislain Vieilledent$^{1,2,\star}$, Clovis Grinand$^{3}$,
@@
   Thomas F. Allnutt$^{5,6}$, and Frédéric Achard$^{1}$\\
-}
 \end{center}
@@
              booktabs=TRUE, format="latex", caption=tabcap2, linesep="") %>% 
-  kable_styling(latex_options=c("striped","hold_position"))
+  kable_styling(latex_options=c("striped","hold_position","scale_down"))
 @
@@
              caption=tabcap3, linesep="") %>% 
+  kable_styling(latex_options=c("striped","hold_position","scale_down"))
+@
+\vspace*{\stretch{1}}
+
+\newpage
+
+\vspace*{\stretch{1}}
+<<frag, results="asis">>=
+# Import and format
+frag.df <- read.table(file="tabs/fordens.txt", sep="\t", header=TRUE)
+tab.4 <- frag.df
+tab.4$Year <- as.character(tab.4$Year)
+tab.4$forest <- round(tab.4$forest/1000)
+names(tab.4) <- c("Year","Forest (Kha)", "0-20", "21-40", "41-60", "61-80", "81-100")
+# Caption
+tabcap4 <- "\\textbf{Evolution of the forest fragmentation from 1953 to
+      2014 in Madagascar}. Five forest fragmentation classes, based on
+    the percentage of forest in the neighborhood, are defined: 0-20\\%
+    (highly fragmented), 21-40\\%, 41-60\\%, 61-80\\% and 81-100\\% (lowly
+    fragmented). The percentage of forest falling in each forest
+    fragmentation class is reported for each year. Forest areas are
+    provided in thousands of hectares (Kha)."
+# Kable
+knitr::kable(tab.4, digits=0, format.args=list(big.mark=","), booktabs=TRUE,
+             caption=tabcap4, linesep="") %>%
   kable_styling(latex_options=c("striped","hold_position"))
 @
-\vspace*{\stretch{1}}
-
-\newpage
-
-\vspace*{\stretch{1}}
-\begin{table}[!h]
-
-  %{\small
-  \begin{longtable}[]{@{}rrrrrrrr@{}}
-    \toprule
-    Year & Forest (Kha) & 0-20\% & 21-40\% & 41-60\% &
-    61-80\% & 81-100\% \\
-    \midrule
-    \endhead
-    1953  & 15,968  & 0  & 1  & 8  & 12  & 78 \\
-    1973  & 14,243  & 6  & 9  & 12  & 16  & 57 \\
-    1990  & 10,762  & 7  & 10  & 13  & 17  & 53 \\
-    2000  & 9,879  & 7  & 11  & 14  & 17  & 51 \\
-    2005  & 9,673  & 8  & 11  & 14  & 18  & 49 \\
-    2010  & 9,320  & 8  & 12  & 15  & 18  & 47 \\
-    2014  & 8,925  & 9  & 13  & 16  & 19  & 44 \\
-    \bottomrule
-  \end{longtable}%}
-  \addtocounter{table}{-1}
-
-  \caption{\textbf{Evolution of the forest fragmentation from 1953 to
-      2014 in Madagascar}. Five forest fragmentation classes, based on
-    the percentage of forest in the neighborhood, are defined: 0-20\%
-    (highly fragmented), 21-40\%, 41-60\%, 61-80\% and 81-100\% (lowly
-    fragmented). The percentage of forest falling in each forest
-    fragmentation class is reported for each year. Forest areas are
-    provided in thousands of hectares (Kha).}
-
-  \label{tab:frag}
-\end{table}
 \vspace*{\stretch{1}}
@@
 
-  \includegraphics[width=\textwidth]{figs/fig_fcc.png}
+  \includegraphics[width=16cm]{figs/fig_fcc.png}
 

### 2018-01-18 manuscript/manuscript.tex
Format Rnw

@@
 \centering
-\begin{tabular}[t]{llrrrrrrrr}
+\resizebox{\linewidth}{!}{\begin{tabular}[t]{llrrrrrrrr}
 \hiderowcolors
@@
 \bottomrule
-\end{tabular}
+\end{tabular}}
 \end{table}
@@
 \centering
-\begin{tabular}[t]{llrrrrrr}
+\resizebox{\linewidth}{!}{\begin{tabular}[t]{llrrrrrr}
 \hiderowcolors
@@
 \vspace*{\stretch{1}}
+\rowcolors{2}{gray!6}{white}
 \begin{table}[!h]
 
-  %{\small
-  \begin{longtable}[]{@{}rrrrrrrr@{}}
-    \toprule
-    Year & Forest (Kha) & 0-20\% & 21-40\% & 41-60\% &
-    61-80\% & 81-100\% \\
-    \midrule
-    \endhead
-    1953  & 15,968  & 0  & 1  & 8  & 12  & 78 \\
-    1973  & 14,243  & 6  & 9  & 12  & 16  & 57 \\
-    1990  & 10,762  & 7  & 10  & 13  & 17  & 53 \\
-    2000  & 9,879  & 7  & 11  & 14  & 17  & 51 \\
-    2005  & 9,673  & 8  & 11  & 14  & 18  & 49 \\
-    2010  & 9,320  & 8  & 12  & 15  & 18  & 47 \\
-    2014  & 8,925  & 9  & 13  & 16  & 19  & 44 \\
-    \bottomrule
-  \end{longtable}%}
-  \addtocounter{table}{-1}
-
-  \caption{\textbf{Evolution of the forest fragmentation from 1953 to
+\caption{\label{tab:frag}\textbf{Evolution of the forest fragmentation from 1953 to
       2014 in Madagascar}. Five forest fragmentation classes, based on
@@
     provided in thousands of hectares (Kha).}
-
-  \label{tab:frag}
+\centering
+\begin{tabular}[t]{lrrrrrr}
+\hiderowcolors
+\toprule
+Year & Forest (Kha) & 0-20 & 21-40 & 41-60 & 61-80 & 81-100\\
+\midrule
+\showrowcolors
+1953 & 15,968 & 0 & 1 & 8 & 12 & 78\\
+1973 & 14,243 & 6 & 9 & 12 & 16 & 57\\
+1990 & 10,762 & 7 & 10 & 13 & 17 & 53\\
+2000 & 9,879 & 7 & 11 & 14 & 17 & 51\\
+2005 & 9,673 & 8 & 11 & 14 & 18 & 49\\
+2010 & 9,320 & 8 & 12 & 15 & 18 & 47\\
+2014 & 8,925 & 9 & 13 & 16 & 19 & 44\\
+\bottomrule
+\end{tabular}
 \end{table}
+\rowcolors{2}{white}{white}
+
 \vspace*{\stretch{1}}

### 2018-01-18 manuscript/manuscript.Rnw
Update gitignore
Update and numbering bibliography section

@@
 \usepackage[round,sort]{natbib} %% Natbib is a popular style for formatting references.
+\usepackage[nottoc,numbib]{tocbibind} %% Referecnes as section
 %\usepackage{multibib}
@@
   fragmentation in Madagascar}
+\author{}
 \date{}
@@
 
-%%\bibliographystyle{besjournals}
-%%\bibliography{biblio}
-
-\begin{thebibliography}{60}
-\providecommand{\natexlab}[1]{#1}
-
-\bibitem[{Achard \emph{et~al.}(2014)Achard, Beuchle, Mayaux, Stibig, Bodart,
-  Brink, Carboni, Desclée, Donnay, Eva, Lupi, Raši, Seliger \&
-  Simonetti}]{Achard2014}
-Achard, F., Beuchle, R., Mayaux, P., Stibig, H.J., Bodart, C., Brink, A.,
-  Carboni, S., Desclée, B., Donnay, F., Eva, H.D., Lupi, A., Raši, R.,
-  Seliger, R. \& Simonetti, D. (2014) Determination of tropical deforestation
-  rates and related carbon losses from 1990 to 2010. \emph{Global Change
-  Biology} \textbf{20}, 2540--2554.
-
-\bibitem[{Aleman \emph{et~al.}(2017)Aleman, Jarzyna \& Staver}]{Aleman2017}
-Aleman, J.C., Jarzyna, M.A. \& Staver, A.C. (2017) Forest extent and
-  deforestation in tropical africa since 1900. \emph{Nature Ecology \&
-  Evolution} .
-
-\bibitem[{Ali \& Aitchison(2008)}]{Ali2008}
-Ali, J.R. \& Aitchison, J.C. (2008) {Gondwana to Asia: Plate tectonics,
-  paleogeography and the biological connectivity of the Indian sub-continent
-  from the Middle Jurassic through latest Eocene (166--35 Ma)}.
-  \emph{Earth-Science Reviews} \textbf{88}, 145--166.
-
-\bibitem[{Allnutt \emph{et~al.}(2008)Allnutt, Ferrier, Manion, Powell,
-  Ricketts, Fisher, Harper, Irwin, Kremen, Labat, Lees, Pearce \&
-  Rakotondrainibe}]{Allnutt2008}
-Allnutt, T.F., Ferrier, S., Manion, G., Powell, G.V.N., Ricketts, T.H., Fisher,
-  B.L., Harper, G.J., Irwin, M.E., Kremen, C., Labat, J.N., Lees, D.C., Pearce,
-  T.A. \& Rakotondrainibe, F. (2008) {A method for quantifying biodiversity
-  loss and its application to a 50-year record of deforestation across
-  Madagascar}. \emph{Conservation Letters} \textbf{1}, 173--181.
-
-\bibitem[{Armitage \emph{et~al.}(2015)Armitage, Highfield, Brody \&
-  Louchouarn}]{Armitage2015}
-Armitage, A.R., Highfield, W.E., Brody, S.D. \& Louchouarn, P. (2015) {The
-  contribution of mangrove expansion to salt marsh loss on the Texas Gulf
-  Coast}. \emph{PloS One} \textbf{10}, e0125404.
-
-\bibitem[{Bastin \emph{et~al.}(2017)Bastin, Berrahmouni, Grainger, Maniatis,
-  Mollicone, Moore, Patriarca, Picard, Sparrow, Abraham, Aloui, Atesoglu,
-  Attore, Bass{\"u}ll{\"u}, Bey, Garzuglia, Garc{\'\i}a-Montero, Groot, Guerin,
-  Laestadius, Lowe, Mamane, Marchi, Patterson, Rezende, Ricci, Salcedo, Diaz,
-  Stolle, Surappaeva \& Castro}]{Bastin2017}
-Bastin, J.F., Berrahmouni, N., Grainger, A., Maniatis, D., Mollicone, D.,
-  Moore, R., Patriarca, C., Picard, N., Sparrow, B., Abraham, E.M., Aloui, K.,
-  Atesoglu, A., Attore, F., Bass{\"u}ll{\"u}, {\c C}., Bey, A., Garzuglia, M.,
-  Garc{\'\i}a-Montero, L.G., Groot, N., Guerin, G., Laestadius, L., Lowe, A.J.,
-  Mamane, B., Marchi, G., Patterson, P., Rezende, M., Ricci, S., Salcedo, I.,
-  Diaz, A.S.P., Stolle, F., Surappaeva, V. \& Castro, R. (2017) The extent of
-  forest in dryland biomes. \emph{Science} \textbf{356}, 635--638.
-
-\bibitem[{Brinck \emph{et~al.}(2017)Brinck, Fischer, Groeneveld, Lehmann,
-  Dantas De~Paula, Pütz, Sexton, Song \& Huth}]{Brinck2017}
-Brinck, K., Fischer, R., Groeneveld, J., Lehmann, S., Dantas De~Paula, M.,
-  Pütz, S., Sexton, J.O., Song, D. \& Huth, A. (2017) High resolution analysis
-  of tropical forest fragmentation and its impact on the global carbon cycle.
-  \emph{Nature Communications} \textbf{8}, 14855.
-
-\bibitem[{Broadbent \emph{et~al.}(2008)Broadbent, Asner, Keller, Knapp,
-  Oliveira \& Silva}]{Broadbent2008}
-Broadbent, E.N., Asner, G.P., Keller, M., Knapp, D.E., Oliveira, P.J. \& Silva,
-  J.N. (2008) Forest fragmentation and edge effects from deforestation and
-  selective logging in the brazilian amazon. \emph{Biological Conservation}
-  \textbf{141}, 1745 -- 1757.
-
-\bibitem[{Brooks \emph{et~al.}(2002)Brooks, Mittermeier, Mittermeier,
-  da~Fonseca, Rylands, Konstant, Flick, Pilgrim, Oldfield, Magin \&
-  Hilton-Taylor}]{Brooks2002}
-Brooks, T.M., Mittermeier, R.A., Mittermeier, C.G., da~Fonseca, G.A.B.,
-  Rylands, A.B., Konstant, W.R., Flick, P., Pilgrim, J., Oldfield, S., Magin,
-  G. \& Hilton-Taylor, C. (2002) Habitat loss and extinction in the hotspots of
-  biodiversity. \emph{Conservation Biology} \textbf{16}, 909--923.
-
-\bibitem[{Burns \emph{et~al.}(2016)Burns, Godfrey, Faina, McGee, Hardt,
-  Ranivoharimanana \& Randrianasy}]{Burns2016}
-Burns, S.J., Godfrey, L.R., Faina, P., McGee, D., Hardt, B., Ranivoharimanana,
-  L. \& Randrianasy, J. (2016) {Rapid human-induced landscape transformation in
-  Madagascar at the end of the first millennium of the Common Era}.
-  \emph{Quaternary Science Reviews} \textbf{134}, 92 -- 99.
-
-\bibitem[{Cornet(1974)}]{Cornet1974}
-Cornet, A. (1974) {Essai de cartographie bioclimatique à Madagascar}. Tech.
-  rep., Orstom.
-
-\bibitem[{Cox \emph{et~al.}(2012)Cox, Nelson, Tumonggor, Ricaut \&
-  Sudoyo}]{Cox2012}
-Cox, M.P., Nelson, M.G., Tumonggor, M.K., Ricaut, F.X. \& Sudoyo, H. (2012) {A
-  small cohort of Island Southeast Asian women founded Madagascar}.
-  \emph{Proceedings of the Royal Society B: Biological Sciences} \textbf{279},
-  2761--2768.
-
-\bibitem[{Crottini \emph{et~al.}(2012)Crottini, Madsen, Poux, Strau{\ss},
-  Vieites \& Vences}]{Crottini2012}
-Crottini, A., Madsen, O., Poux, C., Strau{\ss}, A., Vieites, D.R. \& Vences, M.
-  (2012) {Vertebrate time-tree elucidates the biogeographic pattern of a major
-  biotic change around the K--T boundary in Madagascar}. \emph{Proceedings of
-  the National Academy of Sciences} \textbf{109}, 5358--5363.
-
-\bibitem[{Eklund \emph{et~al.}(2016)Eklund, Blanchet, Nyman, Rocha, Virtanen \&
-  Cabeza}]{Eklund2016}
-Eklund, J., Blanchet, F.G., Nyman, J., Rocha, R., Virtanen, T. \& Cabeza, M.
-  (2016) {Contrasting spatial and temporal trends of protected area
-  effectiveness in mitigating deforestation in Madagascar}. \emph{Biological
-  Conservation} \textbf{203}, 290 -- 297.
-
-\bibitem[{Freudenberger(2010)}]{Freudenberger2010}
-Freudenberger, K. (2010) {Paradise Lost? Lessons from 25 years of USAID
-  environment programs in Madagascar}. \emph{International Resources Group,
-  Washington DC} .
-
-\bibitem[{Gibson \emph{et~al.}(2013)Gibson, Lynam, Bradshaw, He, Bickford,
-  Woodruff, Bumrungsri \& Laurance}]{Gibson2013}
-Gibson, L., Lynam, A.J., Bradshaw, C.J., He, F., Bickford, D.P., Woodruff,
-  D.S., Bumrungsri, S. \& Laurance, W.F. (2013) Near-complete extinction of
-  native small mammal fauna 25 years after forest fragmentation. \emph{Science}
-  \textbf{341}, 1508--1510.
-
-\bibitem[{Goodman \& Benstead(2005)}]{Goodman2005}
-Goodman, S.M. \& Benstead, J.P. (2005) {Updated estimates of biotic diversity
-  and endemism for Madagascar}. \emph{Oryx} \textbf{39}, 73--77.
-
-\bibitem[{Gorenflo \emph{et~al.}(2011)Gorenflo, Corson, Chomitz, Harper,
-  Honzák \& {\"O}zler}]{Gorenflo2011}
-Gorenflo, L.J., Corson, C., Chomitz, K.M., Harper, G., Honzák, M. \&
-  {\"O}zler, B. (2011) \emph{{Exploring the Association Between People and
-  Deforestation in Madagascar}}, vol. 1650. Springer Berlin Heidelberg.
-
-\bibitem[{Grinand \emph{et~al.}(2017)Grinand, Le~Maire, Vieilledent,
-  Razakamanarivo, Razafimbelo \& Bernoux}]{Grinand2017}
-Grinand, C., Le~Maire, G., Vieilledent, G., Razakamanarivo, H., Razafimbelo, T.
-  \& Bernoux, M. (2017) {Estimating temporal changes in soil carbon stocks at
-  ecoregional scale in Madagascar using remote-sensing}. \emph{International
-  Journal of Applied Earth Observation and Geoinformation} \textbf{54}, 1--14.
-
-\bibitem[{Grinand \emph{et~al.}(2013)Grinand, Rakotomalala, Gond, Vaudry,
-  Bernoux \& Vieilledent}]{Grinand2013}
-Grinand, C., Rakotomalala, F., Gond, V., Vaudry, R., Bernoux, M. \&
-  Vieilledent, G. (2013) {Estimating deforestation in tropical humid and dry
-  forests in Madagascar from 2000 to 2010 using multi-date Landsat satellite
-  images and the Random Forests classifier}. \emph{Remote Sensing of
-  Environment} \textbf{139}, 68--80.
-
-\bibitem[{Grouzis \emph{et~al.}(2001)Grouzis, Razanaka, Le~Floc’h \&
-  Leprun}]{Grouzis2001}
-Grouzis, M., Razanaka, S., Le~Floc’h, E. \& Leprun, J.C. (2001)
-  {{\'E}volution de la v{\'e}g{\'e}tation et de quelques param{\`e}tres
-  {\'e}daphiques au cours de la phase post-culturale dans la r{\'e}gion
-  d’Analabo}. \emph{Soci{\'e}t{\'e}s paysannes, transitions agraires et
-  dynamiques {\'e}cologiques dans le Sud-Ouest de Madagascar, Antananarivo,
-  IRD/CNRE} pp. 327--337.
-
-\bibitem[{Hansen \emph{et~al.}(2013)Hansen, Potapov, Moore, Hancher,
-  Turubanova, Tyukavina, Thau, Stehman, Goetz, Loveland, Kommareddy, Egorov,
-  Chini, Justice \& Townshend}]{Hansen2013}
-Hansen, M.C., Potapov, P.V., Moore, R., Hancher, M., Turubanova, S.A.,
-  Tyukavina, A., Thau, D., Stehman, S.V., Goetz, S.J., Loveland, T.R.,
-  Kommareddy, A., Egorov, A., Chini, L., Justice, C.O. \& Townshend, J.R.G.
-  (2013) {High-Resolution Global Maps of 21st-Century Forest Cover Change}.
-  \emph{Science} \textbf{342}, 850--853.
-
-\bibitem[{Harper \emph{et~al.}(2007)Harper, Steininger, Tucker, Juhn \&
-  Hawkins}]{Harper2007}
-Harper, G.J., Steininger, M.K., Tucker, C.J., Juhn, D. \& Hawkins, F. (2007)
-  {Fifty years of deforestation and forest fragmentation in Madagascar}.
-  \emph{Environmental Conservation} \textbf{34}, 325--333.
-
-\bibitem[{Humbert(1927)}]{Humbert1927}
-Humbert, H. (1927) {La destruction d'une flore insulaire par le feu. Principaux
-  aspects de la végétation à Madagascar}. \emph{Mémoires de l'Académie
-  Malgache} \textbf{5}, 1--80.
-
-\bibitem[{Keenan \emph{et~al.}(2015)Keenan, Reams, Achard, de~Freitas, Grainger
-  \& Lindquist}]{Keenan2015}
-Keenan, R.J., Reams, G.A., Achard, F., de~Freitas, J.V., Grainger, A. \&
-  Lindquist, E. (2015) {Dynamics of global forest area: Results from the FAO
-  Global Forest Resources Assessment 2015}. \emph{Forest Ecology and
-  Management} \textbf{352}, 9 -- 20, changes in Global Forest Resources from
-  1990 to 2015.
-
-\bibitem[{Kim \emph{et~al.}(2014)Kim, Sexton, Noojipady, Huang, Anand, Channan,
-  Feng \& Townshend}]{Kim2014}
-Kim, D.H., Sexton, J.O., Noojipady, P., Huang, C., Anand, A., Channan, S.,
-  Feng, M. \& Townshend, J.R. (2014) {Global, Landsat-based forest cover change
-  from 1990 to 2000}. \emph{Remote Sensing of Environment} \textbf{155},
-  178--193.
-
-\bibitem[{Klein(2002)}]{Klein2002}
-Klein, J. (2002) {Deforestation in the Madagascar Highlands -- Established
-  `truth' and scientific uncertainty}. \emph{GeoJournal} \textbf{56}, 191--199.
-
-\bibitem[{Kull(2000)}]{Kull2000}
-Kull, C.A. (2000) {Deforestation, erosion, and fire: degradation myths in the
-  environmental history of Madagascar}. \emph{Environment and History}
-  \textbf{6}, 423--450.
-
-\bibitem[{MEFT \emph{et~al.}(2009)}]{MEFT2009}
-{MEFT, USAID, and CI} (2009) \emph{{Evolution de la couverture de forêts
-  naturelles à Madagascar, 1990-2000-2005}}.
-
-\bibitem[{{Ministère de l'Environnement}(1996)}]{IEFN1996}
-{Ministère de l'Environnement} (1996) {IEFN: Inventaire Ecologique Forestier
-  National}. Tech. rep., {Ministère de l'Environnement de Madagascar,
-  Direction des Eaux et Forêts, DFS Deutsch Forest Service GmbH, Entreprise
-  d'études de développement rural "Mamokatra", FTM}.
-
-\bibitem[{Minten \emph{et~al.}(2013)Minten, Sander \& Stifel}]{Minten2013}
-Minten, B., Sander, K. \& Stifel, D. (2013) {Forest management and economic
-  rents: Evidence from the charcoal trade in Madagascar}. \emph{Energy for
-  Sustainable Development} \textbf{17}, 106 -- 115, special Issue on Charcoal.
-
-\bibitem[{Murcia(1995)}]{Murcia1995}
-Murcia, C. (1995) Edge effects in fragmented forests: implications for
-  conservation. \emph{Trends in Ecology \& Evolution} \textbf{10}, 58 -- 62.
-
-\bibitem[{Neteler \& Mitasova(2008)}]{Neteler2008}
-Neteler, M. \& Mitasova, H. (2008) \emph{{Open source GIS: a GRASS GIS
-  approach}}. Springer.
-
-\bibitem[{Olander \emph{et~al.}(2008)Olander, Gibbs, Steininger, Swenson \&
-  Murray}]{Olander2008}
-Olander, L.P., Gibbs, H.K., Steininger, M., Swenson, J.J. \& Murray, B.C.
-  (2008) {Reference scenarios for deforestation and forest degradation in
-  support of REDD: a review of data and methods}. \emph{Environmental Research
-  Letters} \textbf{3}, 025011.
-
-\bibitem[{Olofsson \emph{et~al.}(2014)Olofsson, Foody, Herold, Stehman,
-  Woodcock \& Wulder}]{Olofsson2014}
-Olofsson, P., Foody, G.M., Herold, M., Stehman, S.V., Woodcock, C.E. \& Wulder,
-  M.A. (2014) Good practices for estimating area and assessing accuracy of land
-  change. \emph{Remote Sensing of Environment} \textbf{148}, 42 -- 57.
-
-\bibitem[{Olofsson \emph{et~al.}(2013)Olofsson, Foody, Stehman \&
-  Woodcock}]{Olofsson2013}
-Olofsson, P., Foody, G.M., Stehman, S.V. \& Woodcock, C.E. (2013) Making better
-  use of accuracy data in land change studies: Estimating accuracy and area and
-  quantifying uncertainty using stratified estimation. \emph{Remote Sensing of
-  Environment} \textbf{129}, 122 -- 131.
-
-\bibitem[{ONE \emph{et~al.}(2013)}]{ONE2013}
-{ONE, DGF, FTM, MNP, and CI} (2013) \emph{{Evolution de la couverture de
-  forêts naturelles à Madagascar 2005-2010}}. \url{https://www.pnae.mg/couverture-de-forets-naturelles-2005-2010}
-
-\bibitem[{ONE \emph{et~al.}(2015)}]{ONE2015}
-{ONE, DGF, MNP, WCS, and Etc Terra} (2015) \emph{{Changement de la couverture
-  de forêts naturelles à Madagascar, 2005-2010-2013}}. \url{https://www.pnae.mg/couverture-de-forets-naturelles-2005-2010-2013}
-
-\bibitem[{Pearson \& Raxworthy(2009)}]{Pearson2009}
-Pearson, R.G. \& Raxworthy, C.J. (2009) {The evolution of local endemism in
-  Madagascar: watershed versus climatic gradient hypotheses evaluated by null
-  biogeographic models}. \emph{Evolution} \textbf{63}, 959--967.
-
-\bibitem[{Pekel \emph{et~al.}(2016)Pekel, Cottam, Gorelick \&
-  Belward}]{Pekel2016}
-Pekel, J.F., Cottam, A., Gorelick, N. \& Belward, A.S. (2016) {High-resolution
-  mapping of global surface water and its long-term changes}. \emph{Nature}
-  \textbf{540}, 418--422.
-
-\bibitem[{Perrier~de La~B{\^a}thie(1921)}]{Perrier1921}
-Perrier~de La~B{\^a}thie, H. (1921) \emph{{La v{\'e}g{\'e}tation malgache}},
-  vol.~23. Mus{\'e}e Colonial.
-
-\bibitem[{Ploch \& Cook(2012)}]{Ploch2012}
-Ploch, L. \& Cook, N. (2012) {Madagascar's Political Crisis}.
-
-\bibitem[{Potapov \emph{et~al.}(2017)Potapov, Hansen, Laestadius, Turubanova,
-  Yaroshenko, Thies, Smith, Zhuravleva, Komarova, Minnemeyer \&
-  Esipova}]{Potapov2017}
-Potapov, P., Hansen, M.C., Laestadius, L., Turubanova, S., Yaroshenko, A.,
-  Thies, C., Smith, W., Zhuravleva, I., Komarova, A., Minnemeyer, S. \&
-  Esipova, E. (2017) {The last frontiers of wilderness: Tracking loss of intact
-  forest landscapes from 2000 to 2013}. \emph{Science Advances} \textbf{3}.
-
-\bibitem[{Puyravaud(2003)}]{Puyravaud2003}
-Puyravaud, J.P. (2003) Standardizing the calculation of the annual rate of
-  deforestation. \emph{Forest Ecology and Management} \textbf{177}, 593--596.
-
-\bibitem[{Rakotomala \emph{et~al.}(2015)Rakotomala, Rabenandrasana,
-  Andriambahiny, Rajaonson, Andriamalala, Burren, Rakotoarijaona, Parany,
-  Vaudry, Rakotoniaina, Ranaivosoa, Rahagalala, Randrianary \&
-  Grinand}]{Rakotomalala2015}
-Rakotomala, F., Rabenandrasana, J., Andriambahiny, J., Rajaonson, R.,
-  Andriamalala, F., Burren, C., Rakotoarijaona, J., Parany, B., Vaudry, R.,
-  Rakotoniaina, S., Ranaivosoa, R., Rahagalala, P., Randrianary, T. \& Grinand,
-  C. (2015) {Estimation de la d{\'e}forestation des for{\^e}ts humides {\`a}
-  Madagascar entre 2005, 2010 et 2013: combinaison multi-date d'images LANDSAT,
-  utilisation de l'algorithme Random Forest et proc{\'e}dure de validation}.
-  \emph{Revue Fran{\c{c}}aise de Photogramm{\'e}trie et de
-  T{\'e}l{\'e}d{\'e}tection} pp. 11--23.
-
-\bibitem[{Riitters \emph{et~al.}(2000)Riitters, Wickham, O'Neill, Jones \&
-  Smith}]{Riitters2000}
-Riitters, K., Wickham, J., O'Neill, R., Jones, B. \& Smith, E. (2000)
-  Global-scale patterns of forest fragmentation. \emph{Conservation Ecology}
-  \textbf{4}, 3.
-
-\bibitem[{Saatchi \emph{et~al.}(2011)Saatchi, Harris, Brown, Lefsky, Mitchard,
-  Salas, Zutta, Buermann, Lewis, Hagen, Petrova, White, Silman \&
-  Morel}]{Saatchi2011}
-Saatchi, S.S., Harris, N.L., Brown, S., Lefsky, M., Mitchard, E.T.A., Salas,
-  W., Zutta, B.R., Buermann, W., Lewis, S.L., Hagen, S., Petrova, S., White,
-  L., Silman, M. \& Morel, A. (2011) Benchmark map of forest carbon stocks in
-  tropical regions across three continents. \emph{Proceedings of the National
-  Academy of Sciences} \textbf{108}, 9899--9904.
-
-\bibitem[{Saunders \emph{et~al.}(1991)Saunders, Hobbs \&
-  Margules}]{Saunders1991}
-Saunders, D.A., Hobbs, R.J. \& Margules, C.R. (1991) {Biological Consequences
-  of Ecosystem Fragmentation: A Review}. \emph{Conservation Biology}
-  \textbf{5}, 18--32.
-
-\bibitem[{Scales(2011)}]{Scales2011}
-Scales, I.R. (2011) {Farming at the Forest Frontier: Land Use and Landscape
-  Change in Western Madagascar, 1896-2005}. \emph{Environment and History}
-  \textbf{17}, 499--524.
-
-\bibitem[{Smith \emph{et~al.}(2003)Smith, Muir, Walpole, Balmford \&
-  Leader-Williams}]{Smith2003}
-Smith, R.J., Muir, R.D.J., Walpole, M.J., Balmford, A. \& Leader-Williams, N.
-  (2003) Governance and the loss of biodiversity. \emph{Nature} \textbf{426},
-  67--70.
-
-\bibitem[{Tidd \emph{et~al.}(2001)Tidd, Pinder \& Ferguson}]{Tidd2001}
-Tidd, S.T., Pinder, J. \& Ferguson, G.W. (2001) {Deforestation and habitat loss
-  for the Malagasy flat-tailed tortoise from 1963 through 1993}.
-  \emph{Chelonian Conservation and Biology} \textbf{4}, 59--65.
-
-\bibitem[{Tropek \emph{et~al.}(2014)Tropek, Sedl{\'a}{\v c}ek, Beck, Keil,
-  Musilov{\'a}, {\v S}{\'\i}mov{\'a} \& Storch}]{Tropek2014}
-Tropek, R., Sedl{\'a}{\v c}ek, O., Beck, J., Keil, P., Musilov{\'a}, Z., {\v
-  S}{\'\i}mov{\'a}, I. \& Storch, D. (2014) {Comment on "High-resolution global
-  maps of 21st-century forest cover change"}. \emph{Science} \textbf{344},
-  981--981.
-
-\bibitem[{Tyukavina \emph{et~al.}(2015)Tyukavina, Baccini, Hansen, Potapov,
-  Stehman, Houghton, Krylov, Turubanova \& Goetz}]{Tyukavina2015}
-Tyukavina, A., Baccini, A., Hansen, M.C., Potapov, P.V., Stehman, S.V.,
-  Houghton, R.A., Krylov, A.M., Turubanova, S. \& Goetz, S.J. (2015)
-  Aboveground carbon loss in natural and managed tropical forests from 2000 to
-  2012. \emph{Environmental Research Letters} \textbf{10}, 074002.
-
-\bibitem[{Tyukavina \emph{et~al.}(2017)Tyukavina, Hansen, Potapov, Stehman,
-  Smith-Rodriguez, Okpa \& Aguilar}]{Tyukavina2017}
-Tyukavina, A., Hansen, M.C., Potapov, P.V., Stehman, S.V., Smith-Rodriguez, K.,
-  Okpa, C. \& Aguilar, R. (2017) {Types and rates of forest disturbance in
-  Brazilian Legal Amazon, 2000{\textendash}2013}. \emph{Science Advances}
-  \textbf{3}.
-
-\bibitem[{{United Nations}(2015)}]{UN2015}
-{United Nations} (2015) \emph{{World Population Prospects: The 2015 Revision,
-  Key Findings and Advance Tables. Working Paper No. ESA/P/WP.241.}}
-
-\bibitem[{Verhegghen \emph{et~al.}(2016)Verhegghen, Eva, Desclée \&
-  Achard}]{Verhegghen2016}
-Verhegghen, A., Eva, H., Desclée, B. \& Achard, F. (2016) {Review and
-  Combination of Recent Remote Sensing Based Products for Forest Cover Change
-  Assessments in Cameroon}. \emph{International Forestry Review} \textbf{18},
-  14--25.
-
-\bibitem[{Vieilledent \emph{et~al.}(2016)Vieilledent, Gardi, Grinand, Burren,
-  Andriamanjato, Camara, Gardner, Glass, Rasolohery, Rakoto~Ratsimba, Gond \&
-  Rakotoarijaona}]{Vieilledent2016}
-Vieilledent, G., Gardi, O., Grinand, C., Burren, C., Andriamanjato, M., Camara,
-  C., Gardner, C.J., Glass, L., Rasolohery, A., Rakoto~Ratsimba, H., Gond, V.
-  \& Rakotoarijaona, J.R. (2016) {Bioclimatic envelope models predict a
-  decrease in tropical forest carbon stocks with climate change in Madagascar}.
-  \emph{Journal of Ecology} \textbf{104}, 703--715.
-
-\bibitem[{Vieilledent \emph{et~al.}(2013)Vieilledent, Grinand \&
-  Vaudry}]{Vieilledent2013}
-Vieilledent, G., Grinand, C. \& Vaudry, R. (2013) {Forecasting deforestation
-  and carbon emissions in tropical developing countries facing demographic
-  expansion: a case study in Madagascar}. \emph{Ecology and Evolution}
-  \textbf{3}, 1702--1716.
-
-\bibitem[{Virah-Sawmy(2009)}]{Virah-Sawmy2009}
-Virah-Sawmy, M. (2009) {Ecosystem management in Madagascar during global
-  change}. \emph{Conservation Letters} \textbf{2}, 163--170.
-
-\bibitem[{Vorontsova \emph{et~al.}(2016)Vorontsova, Besnard, Forest, Malakasi,
-  Moat, Clayton, Ficinski, Savva, Nanjarisoa, Razanatsoa, Randriatsara, Kimeu,
-  Luke, Kayombo \& Linder}]{Vorontsova2017}
-Vorontsova, M.S., Besnard, G., Forest, F., Malakasi, P., Moat, J., Clayton,
-  W.D., Ficinski, P., Savva, G.M., Nanjarisoa, O.P., Razanatsoa, J.,
-  Randriatsara, F.O., Kimeu, J.M., Luke, W.R.Q., Kayombo, C. \& Linder, H.P.
-  (2016) {Madagascar's grasses and grasslands: anthropogenic or natural?}
-  \emph{Proceedings of the Royal Society of London B: Biological Sciences}
-  \textbf{283}.
-
-\end{thebibliography}
+\bibliographystyle{bib/besjournals}
+\bibliography{bib/biblio}
 
@@
 knitr::kable(tab.2, col.names=col.names, align=c("l","l",rep("r",8)), format.args=list(big.mark=","), 
-             booktabs=TRUE, format="latex", caption=tabcap2, linesep="") %>% 
+             booktabs=TRUE, format="latex", caption=tabcap2, linesep=c("","","","\\addlinespace")) %>% 
   kable_styling(latex_options=c("striped","hold_position","scale_down"))
@@
 knitr::kable(tab.3, col.names=col.names, align=c("l","l",rep("r",6)), booktabs=TRUE, 
-             caption=tabcap3, linesep="") %>% 
+             caption=tabcap3, linesep=c("","","","\\addlinespace")) %>% 
   kable_styling(latex_options=c("striped","hold_position","scale_down"))

### 2018-01-18 deforestmap.R

@@ -983,23 +983,16 @@ opts_chunk$set(echo=FALSE, cache=FALSE,
                message=FALSE, highlight=TRUE,
                fig.show="hide", size="small",
                tidy=FALSE)
-options(knitr.kable.NA = "-")
+options(knitr.kable.NA="-")
+opts_knit$set(root.dir="manuscript")
 
-## Knit and translate to html and pdf
-dir.create("report")
-## Main document
-render("deforestmap.Rmd")
+## Knit
+knitr::knit2pdf("manuscript/manuscript.Rnw", output="manuscript/manuscript.tex")
+
 ## Cover letter
-render("coverletter.md",output_format=c("pdf_document"),output_dir="report") # pdf output
-
-##========================
-## Pandoc for word
-##========================
-
-system("pandoc manuscript/manuscript.tex --bibliography=manuscript/bib/biblio.bib \\
-       --csl=manuscript/bib/journal-of-applied-ecology.csl -o manuscript/manuscript.md")
+rmarkdown::render("manuscript/coverletter.md", output_format=c("pdf_document"),
+                  output_dir="manuscript") # pdf output
 
 ##===========================================================================
 ## End of script
 ##===========================================================================
-

### 2018-01-18 manuscript/manuscript.Rnw
Cover letter

@@
 available forest extent. Area and rate estimates are produced at the
-national scale and for the four forest ecosystems present in
+national scale and for the four forest types present in
 Madagascar: moist forest in the East, dry forest in the West, spiny
 forest in the South, and mangroves on the Western coast
-(Fig.~\ref{fig:ecoregion}). To define the forest domains, we used a
+(Fig.~\ref{fig:ecoregion}). To define the forest types, we used a
 map from the MEFT (\emph{``Ministère de l'Environnement et des Forêts
@@
 
-The authors thank Jean-François Bastin for useful comments on a previous
-version of the manuscript and Peter Vogt for useful advices on which
+The authors thank two anonymous reviewers for useful comments on a previous
+version of the manuscript. They also thank Peter Vogt for useful advices on which
 metric to use to estimate forest fragmentation. This study is part of
@@
       vegetation \citep{IEFN1996} criteria. The dark grey areas
-      represent the remaining natural forest cover for the year 2014.}
+      represent the remaining natural forest cover for the year 2014.
+      Forest types are defined on the basis of their appartenance to
+      one of the four ecoregions.}
 

### 2018-01-22 manuscript/manuscript.Rnw
Update cover letters

@@
 \usepackage[round,sort]{natbib} %% Natbib is a popular style for formatting references.
-\usepackage[nottoc,numbib]{tocbibind} %% Referecnes as section
 %\usepackage{multibib}
@@
 estimates from the three existing studies at the national scale for
-Madagascar: (i) \citep{Harper2007}, (ii) \citep{MEFT2009} and (iii)
-\citep{ONE2015}. \citet{Harper2007} provides forest cover and
+Madagascar: (i) \citet{Harper2007}, (ii) \citet{MEFT2009} and (iii)
+\citet{ONE2015}. \citet{Harper2007} provides forest cover and
 deforestation estimates for the periods
@@
 
+\renewcommand{\bibname}{References}
+\renewcommand{\bibsection}{\section{\bibname}}
 \bibliographystyle{bib/besjournals}
@@
 
-    \includegraphics[width=11cm]{figs/ecoregion.png}
+    \includegraphics[width=10cm]{figs/ecoregion.png}
 

### 2018-01-22 deforestmap.R

@@ -978,6 +978,7 @@ save(list=SavedObjects,file="deforestmap.rda")
 ##========================
 
 ## Set knitr chunk default options
+library(knitr)
 opts_chunk$set(echo=FALSE, cache=FALSE,
                results="hide", warning=FALSE,
                message=FALSE, highlight=TRUE,

### 2018-01-23 manuscript/manuscript.Rnw
Update cover letter

@@
 
-\documentclass[a4paper, 12pt, leqno]{article} %leqno: numéro d'équation à gauche
+\documentclass[a4paper, 12pt, leqno, dvipsnames]{article} %leqno: numéro d'équation à gauche
 \pagenumbering{arabic} % choose how to number the pages
@@
 \setlength{\LTright}{-5cm plus 1 fill}
-\usepackage[colorlinks=true,citecolor=blue]{hyperref} %% Gère les hyperliens
+\usepackage[colorlinks=true,citecolor=Blue,urlcolor=Maroon]{hyperref} %% Gère les hyperliens
 \usepackage{lineno} %% Numérotation des lignes
@@
   substantial gaps which prevent an exhaustive assessment of long-term
-  deforestation trends in Madagascar. In this study, we combine
+  deforestation trends in Madagascar. In this study, we combined
   historical national forest cover maps (covering the period
@@
   and estimates can be used in the REDD+ framework which aims at
-  ``Reducing Emissions from Deforestation and Forest Degradation'' and
+  ``Reducing Emissions from Deforestation and forest Degradation'' and
   for optimizing the current protected area network.
@@
 forest cover map of year 2000. To do so, we assumed that if forest was
-present in 2000, the pixel was also forested in 1990. The remaining
+present in 2000, the pixel was also forested in 1990. Indeed, there is little evidence of natural forest
+regeneration in Madagascar \citep{Grouzis2001, Harper2007}, especially over such a short period of time. The remaining
 unclassified pixels were limited to a relatively small total area of
@@
 pixels for year 1973 corresponded to a significant total area of 3.3
-million ha. We also reprojected the forest cover map of year 1953 to a
+million ha which was left as is. We also reprojected the forest cover map of year 1953 to a
 common projection in order to compare the forest cover area in 1953
@@
 record and derive long-term estimates of deforestation. Moreover, this
-approach can be repeated in the future if and when the global tree
-cover product is updated. We have made the \R{}/GRASS code used for this
+approach could be repeated in the future with the release of updated tree cover loss data. 
+We have made the \R{}/GRASS code used for this
 study freely available in a GitHub repository (see Data availability

### 2018-01-23 deforestmap.R

@@ -489,6 +489,53 @@ system("r.out.gdal --o input=dist_edge_2014_mask nodata=9999 createopt='compress
        output=outputs/dist_edge_2014_mask.tif")
 system("r.mask -r")
 
+##==========================
+## Histograms forest edge
+##==========================
+
+Year <- c(1953,1973,1990,2000,2005,2010,2014) 
+year <- length(Year)
+# Bins of distance
+bins <- seq(0,5000,by=100)
+nbins <- length(bins)-1
+# Data-frame to store results
+r <- data.frame()
+
+# Loop on years
+for (i in 1:length(Year)) {
+  # Message
+  cat(paste("Year: ",Year[i],"\n",sep=""))
+  # Histogram
+  his <- system(paste0("r.stats -nc input=dist_edge_",Year[i]), intern=TRUE)
+  d <- data.frame(matrix(as.numeric(unlist(strsplit(his," "))), ncol=2, byrow=TRUE))
+  names(d) <- c("bin","ncell")
+  # Number of cells per bin of distance
+  hist.df <- data.frame(bins=1:nbins,ncell=NA)
+  for (i in 1:nbins) {
+    w <- which(d$bin>=bins[i] & d$bin<bins[i+1])
+    hist.df$ncell[i] <- sum(d$ncell[w])
+  }
+  # Proportion
+  tot.cell <- sum(hist.df$ncell)
+  hist.df$prop <- 100*hist.df$ncell/tot.cell
+  # Store results
+  r <- rbind(r,hist.df)
+}
+# Year as factor
+r$Year <- as.factor(rep(Year,each=nbins))
+# Reformat bins
+r$bins <- r$bins*100-50
+# Save results
+write.table(r,file="outputs/histdist.txt",row.names=FALSE,sep="\t")
+
+# Plot with ggplot
+pp <- ggplot(data=r, aes(x=bins, y=prop, colour=Year)) + 
+  xlab("Distance to forest edge (m)") +
+  ylab("Proportion of forest (%)") +
+  xlim(c(0,5000)) +
+  geom_line()
+ggsave("outputs/histdist.png",pp)
+
 ##========================
 ## Statistics whole forest
 ##========================

### 2018-01-25 manuscript/manuscript.Rnw
Update cover letter

@@
 maps of forest cover and forest cover changes over Madagascar for the
-years \emph{c.}~1953, \emph{c.}~1973, 1990 and 2000. The
-\emph{c.}~1953 forest map was derived from the visual interpretation
-of aerial photography at coarse scale (1/1,000,000). Forest maps for
-the years \emph{c.}~1973, 1990, and 2000 were obtained from supervised
-classification of Landsat satellite images at 60~m resolution (for the
-year 1973) or 30~m resolution (for years 1990 and 2000) and can be
-used to derive more accurate estimates of forest cover (89.5\%
-accuracy reported for the forest/non-forest map of year
-2000). Nonetheless, maps provided by \citet{Harper2007} are not
-exhaustive (due to the presence of clouds in the satellite imagery),
+years 1953, 1973, 1990 and 2000. The
+1953 forest map is a vector map derived from the visual interpretation
+of aerial photographies. Forest maps for
+the years 1973, 1990, and 2000 were obtained from the supervised
+classification of Landsat satellite images and can be
+used to derive more accurate estimates of forest cover than those from the FAO
+Forest Resource Assessment report. Nonetheless, maps provided by 
+\citet{Harper2007} are not exhaustive (due to the presence of clouds in the satellite imagery),
 e.g.~11 244~km2 are mapped as unknown cover type for the year
@@
 \citet{Harper2007}, we also assess trends in deforestation rates and
-forest fragmentation from \emph{c.} 1953 to 2014. We finally discuss
+forest fragmentation from 1953 to 2014. We finally discuss
 the possibility to extend our approach to other tropical countries or
-repeat it in the future. We also discuss how our results could help
-assess the effectiveness of current conservation strategies, and
-implementation new conservation strategies for the future in
-Madagascar.
+repeat it in the future for Madagascar. We also discuss how our results could help
+assess the effectiveness of past and current conservation strategies in Madagascar, and
+implement new strategies in the future.
 
@@
 
-We produced annual forest/non-forest maps at 30~m resolution for the
-full territory of Madagascar for the period 2000-2014 by combining the
-forest map of year 2000 from \citet{Harper2007}, and the tree cover
-percentage and annual tree cover loss maps over the period 2000-2014
-from \citet{Hansen2013}. The 2000 Harper's forest map includes 208,000
-ha of unclassified areas due to the presence of clouds on satellite
-images, mostly (88\%) within the moist forest domain which covered
+Original 1990-2000 forest cover change map for Madagascar from
+\citet{Harper2007} is a raster map at 28.5 m resolution. It was
+derived from the supervised classification of Landsat TM (Thematic Mapper) and 
+ETM+ (Enhanced Thematic Mapper Plus) satellite images. For our study, this map 
+has been resampled at 30~m resolution 
+using a nearest-neighbor interpolation and reprojected in the WGS 84/UTM zone 
+38S projected coordinate system.
+
+The 2000 Harper's forest map includes 208,000 ha of unclassified areas due to 
+the presence of clouds on satellite images, mostly (88\%) within the moist 
+forest domain which covered
 4.17 Mha in total in 2000. To provide a label (forest or non-forest)
@@
 obtained a forest cover map for the year 2000 covering the full
-territory of Madagascar. We then combined this forest cover map of the
+territory of Madagascar.
+
+We then combined the forest cover map of the
 year 2000 with the annual tree cover loss maps from 2001 to 2014
-provided by \citet{Hansen2013} to create annual forest cover maps from
+from \citet{Hansen2013} to create annual forest cover maps from
 2001 to 2014 at 30~m resolution. To do so, Hansen's tree cover loss
 maps were resampled on the same grid as the original Harper's map at
-30~m resolution using a nearest-neighbor interpolation. We also
-completed the Harper's forest map of year 1990 by filling unclassified
+30~m resolution using a nearest-neighbor interpolation. We also completed 
+the Harper's forest map of year 1990 by filling unclassified
 areas (due to the presence of clouds on satellite images) using our
 forest cover map of year 2000. To do so, we assumed that if forest was
-present in 2000, the pixel was also forested in 1990. Indeed, there is little evidence of natural forest
-regeneration in Madagascar \citep{Grouzis2001, Harper2007}, especially over such a short period of time. The remaining
-unclassified pixels were limited to a relatively small total area of
-\emph{c.} 8,000 ha. We labeled these residual pixels as non-forest, as
-for the year 2000. Similarly we completed the Harper's forest map of
-year 1973 by filling unclassified areas using our forest cover map of
-the year 1990 assuming that if forest was present in 1990, it was also
+present in 2000, the pixel was also forested in 1990. Indeed, there is little 
+evidence of natural forest regeneration in Madagascar 
+\citep{Grouzis2001, Harper2007}, especially over such a short period of time. 
+The remaining unclassified pixels were limited to a relatively small total area 
+of about 8,000 ha. We labeled these residual pixels as non-forest, as
+for the year 2000.
+
+The 1973 forest cover map for Madagascar from \citet{Harper2007} is a 
+raster map at 57 m resolution derived from the supervised classification of 
+Landsat MSS (Multispectral Scanner System) satellite images. We resampled this 
+map at 30~m resolution using a nearest-neighbor interpolation on the same grid 
+as the forest cover maps for years 1990 and 2000. We completed the Harper's 
+forest map of year 1973 by filling unclassified areas using our forest cover 
+map of the year 1990 assuming that if forest was present in 1990, it was also
 present in 1973. Contrary to the year 1990, the remaining unclassified
 pixels for year 1973 corresponded to a significant total area of 3.3
-million ha which was left as is. We also reprojected the forest cover map of year 1953 to a
-common projection in order to compare the forest cover area in 1953
-with forest cover areas at the following dates. This map was produced
-by scanning a paper map derived from aerial photos, and thus could not
-be perfectly aligned with the other maps produced through digital
-processing of satellite imagery \citep{Harper2007}. Finally for all
-forest cover maps from 1973, the isolated single non-forest pixels
-(i.e.~fully surrounded by forest pixels) were removed, assuming that
-single non-forest pixels inside a forest patch were not corresponding
-to deforestation (they might correspond to selective logging
-activities). This allowed us to avoid counting very small scale events
-(\textless 0.1 ha, such as selective logging) as forest
-fragmentation. All the resulting maps have been made permanently and
+millions ha which was left as is.
+
+The 1953 forest cover map from \citet{Harper2007} is a vector map produced by scanning 
+the paper map of \citet{Humbert1965} which was derived from the visual interpretation 
+of aerial photographies. We reprojected the forest cover map of year 1953 in the WGS 84/UTM zone 
+38S projected coordinate system. Because of the methodology used to derive the 1953 forest 
+cover map, it was not possible to perfectly aligned this map with the forest cover 
+maps of later years which were produced through digital processing of satellite imagery. 
+As a consequence, the 1953 cannot be merged with the map of laters years to identify 
+precisely the location of the deforested areas. Nonetheless, the 1953 forest cover 
+map can be used to have a rough estimate of the forest cover and forest fragmentation at this date.
+To do so, the map was rasterized at 30 m resolution on the same grid 
+as the forest cover maps for years 1973, 1990 and 2000.
+
+Finally for all forest cover maps from 1973, isolated single non-forest pixels
+(i.e.~fully surrounded by forest pixels) were recategorized as forest pixels.
+Doing so, forest cover increased from about 95,000 ha for year 1953 to
+about 600,000 ha for year 2010. This allowed us to avoid counting very small scale
+events (\textless 0.1 ha, such as selective logging or windthrow) as
+deforestation. It also prevents us from underestimating forest cover and overestimating 
+forest fragmentation. All the resulting maps have been made permanently and
 freely available on the Zenodo research data repository at
@@
 
-Because of the large unclassified area (3.3 million ha) in 1973, the
+Because of the large unclassified area (3.3 millions ha) in 1973, the
 annual deforestation areas and rates for the two periods 1953-1973 and
@@
 2014 in Madagascar. The mean distance to forest edge has decreased to
-\emph{c.} 300~m in 2014 while it was previously \emph{c.} 1.5~km in
+about 300~m in 2014 while it was of about 1.5~km in
 1973 (Fig.~\ref{fig:dist_edge}). Moreover, a large proportion (73\%)
 of the forest was located at a distance greater than 100~m in 1973,
-while almost half of the forest (46\%) is now at a distance lower than
+while almost half of the forest (46\%) is at a distance lower than
 100~m from forest edge in 2014 (Fig.~\ref{fig:dist_edge}). The
@@
 decreased since 1953. The percentage of forest belonging to the lowly
-fragmented category has fallen from 57\% in 1973 to 44\% in 2014. In
+fragmented class has fallen from 57\% in 1973 to 44\% in 2014. In
 2014, 22\% of the forest belonged to the two highest
 fragmented forest classes (less than 40\% of forest cover in the
-neighborhood) while 15\% of the forest belonged to these two
-classes in 1973 (Tab.~\ref{tab:frag}).
+neighborhood) while only 15\% of the forest belonged to these two
+fragmentation classes in 1973 (Tab.~\ref{tab:frag}).
 
@@
 of the species. The additional deforestation we observed over the
-period 2000-2014 (around 1Mha of natural forest) worsen this
+period 2000-2014 (around 1 Mha of natural forest) worsen this
 result. Regarding carbon emissions, using the 2010 aboveground forest
@@
 Biodiversité) and the FFEM (Fond Français pour l'Environnement
-Mondial) under the project agreement AAP-SCEN-2013 I. The ReCaREDD
+Mondial) under the project agreement AAP-SCEN-2013~I. The ReCaREDD
 project is funded by the European Commission. The authors declare that there are no conflicts of interest related to this article.
@@
     and forest cover estimates from previous studies. The increase in
-    mangrove and spiny forest covers from \\emph{c.} 1953 to \\emph{c.}
-    1973 in \\citet{Harper2007} and our study is most probably due to
+    mangrove and spiny forest covers from 1953 to 1973 in \\citet{Harper2007} and our study is most probably due to
     differences in forest definition and mapping methods between the
@@
       represent the remaining natural forest cover for the year 2014.
-      Forest types are defined on the basis of their appartenance to
+      Forest types are defined on the basis of their belonging to
       one of the four ecoregions.}
@@
   \caption{\textbf{Forest cover change on six decades from 1953 to
-      2014 in Madagascar.} forest cover changes from \emph{c.} 1973 to
-    2014 are shown in the main figure, and forest cover in \emph{c.}
-    1953 is shown in the bottom-right inset. Two zooms in the western
+      2014 in Madagascar.} forest cover changes from 1973 to
+    2014 are shown in the main figure, and forest cover in 1953 is shown 
+    in the bottom-right inset. Two zooms in the western
     dry (left part) and eastern moist (right part) ecoregions present
     more detailed views of (from top to bottom): forest cover in
-    1950s, forest cover change from \emph{c.} 1973 to 2014, forest
+    1953, forest cover change from 1973 to 2014, forest
     fragmentation in 2014 and distance to forest edge in 2014. Data on

### 2018-01-25 R/plotfcc.R

@@ -136,7 +136,7 @@ plot_zoom_frag <- function(rast_file,palette=frag.palette) {
     geom_raster(aes(fill=factor(value))) + ylab("Fragmentation 2014") +
     scale_fill_manual(values=palette,breaks=c(1:5),
                       na.value="transparent",
-                      guide=guide_legend(title="Frag. index",title.position="top",
+                      guide=guide_legend(title="Frag. class",title.position="top",
                                          label.position="right")) +
     theme_bw() + theme_zoom + coord_equal() +
     theme(legend.position="bottom", 
@@ -212,7 +212,7 @@ ggsave("outputs/ecoregion.png", plot.ecoregion, width=10, height=15, units="cm")
 for1953 <- raster("outputs/fcc_for1953.tif")
 for1953.plot <- gplot(for1953,maxpixels=res) +
   geom_raster(aes(fill=factor(value))) +
-  xlab("Cover 1950s") +
+  xlab("Cover 1953") +
   scale_fill_manual(values=fcc_for1953.palette,na.value="transparent") +
   theme_bw() + theme_base + 
   theme(panel.background=element_rect(fill="transparent"),
@@ -255,7 +255,7 @@ fcc.combi <- fcc.plot + theme(legend.position="null") +
   annotation_custom(grob=legend,xmin=400000,xmax=760000,ymin=8200000,ymax=8800000)
 
 ## Zooms for1953
-zw.for1953 <- plot_zoom_fcc("outputs/fcc_for1953_zoom1.tif",fcc_for1953.palette) + ggtitle("Western zoom") + ylab("Cover 1950s")
+zw.for1953 <- plot_zoom_fcc("outputs/fcc_for1953_zoom1.tif",fcc_for1953.palette) + ggtitle("Western zoom") + ylab("Cover 1953")
 ze.for1953 <- plot_zoom_fcc("outputs/fcc_for1953_zoom2.tif",fcc_for1953.palette) + ggtitle("Eastern zoom") + ylab("")
 
 ## Zooms fcc

### 2018-01-25 deforestmap.R

@@ -17,14 +17,6 @@
 ## GRASS GIS 7.2.x is also needed to run this script
 ## https://grass.osgeo.org/
 
-## Read argument for download
-## Set "down" to TRUE if you want to download the sources. Otherwise, the data already provided in the gisdata repository will be used.
-arg <- commandArgs(trailingOnly=TRUE)
-down <- FALSE
-if (length(arg)>0) {
-  down <- arg[1]
-}
-
 ##= Libraries
 pkg <- c("broom","sp","rgdal","raster","ggplot2","gridExtra","RColorBrewer",
          "rasterVis","knitr","rmarkdown","rgeos","rgdal","rgrass7","xtable")
@@ -67,6 +59,7 @@ initGRASS(gisBase="/usr/lib/grass72",home=tempdir(),
 
 ##==================================================================
 ## Download data from Zenodo: https://doi.org/10.5281/zenodo.1118955
+down <- FALSE
 if (down) {
   d <- "https://zenodo.org/record/1118956/files/gisdata.zip"
   download.file(url=d,destfile="gisdata.zip",method="curl",quiet=FALSE)
@@ -327,6 +320,11 @@ SavedObjects <- c(SavedObjects, "cloud.1973")
 #= Export
 system("r.out.gdal --o input=for1973 createopt='compress=lzw,predictor=2' type=Byte output=outputs/for1973.tif")
 
+#= We rename for1973 (with cloud class 5) as for1973_5
+#= for1973 will be without clouds
+system("g.rename raster=for1973,for1973_5")
+system("r.mapcalc --o 'for1973 = if(!isnull(for1973_5) &&& for1973_5==1, 1, null())'")
+
 #====================================================================================
 # Reproject c.1953
 
@@ -361,7 +359,7 @@ system("r.out.gdal --o input=for1953 createopt='compress=lzw,predictor=2' \\
 # Remove salt and pepper
 # =======================================
 
-Year <- c(1953,1973,1990,2000,2005,2010,2013,2014)
+Year <- c(1953,1973,1990,2000,2005,2010,2014)
 system("g.region rast=harper -ap")
 system("r.mask --o raster=harper")
 for (i in 1:length(Year)) {
@@ -371,15 +369,31 @@ for (i in 1:length(Year)) {
   system(paste0("r.neighbors --o input=for",Year[i]," output=for",Year[i],"_sum method=sum size=3"))
[…]

### 2018-03-11 deforestmap.R
Histogram TC

@@ -1028,6 +1028,44 @@ carbon_emissions <- acd_2010-acd_2014
 SavedObjects <- c(SavedObjects,"carbon_emissions")
 
 ##========================
+## Histogram of tree cover
+##========================
+
+system("r.mapcalc 'for2000_moist = if(for2000==1 && ecoregion==3,1,null())'")
+system("r.mask --o raster=for2000_moist")
+system("r.mapcalc --o 'tc2000_moist = tc2000'")
+system("r.mask -r")
+hist_tc2000_moist <- system("r.stats -c input=tc2000_moist nsteps=100", intern=TRUE)
+df_tc2000_moist <- data.frame(matrix(unlist(strsplit(hist_tc2000_moist, " ")), ncol=2, byrow=TRUE))
+df_tc2000_moist <- df_tc2000_moist[-c(101),]
+names(df_tc2000_moist) <- c("tc","npix")
+df_tc2000_moist$npix <- as.numeric(as.character(df_tc2000_moist$npix))
+df_tc2000_moist$tc <- as.numeric(as.character(df_tc2000_moist$tc))
+df_tc2000_moist$nha <- df_tc2000_moist$npix*30*30/10000
+df_tc2000_moist$freq <- 100*df_tc2000_moist$npix/sum(df_tc2000_moist$npix)
+# Cumulative proportion
+for (i in 100:1) {
+  df_tc2000_moist$cum_prop[i] <- 100*sum(df_tc2000_moist$npix[100:i])/sum(df_tc2000_moist$npix)
+}
+# How much moist forest has a TC over 75%
+prop_tc75 <- round(df_tc2000_moist$cum_prop[df_tc2000_moist$tc==75]) # 90%
+prop_tc70 <- round(df_tc2000_moist$cum_prop[df_tc2000_moist$tc==70]) # 92%
+
+# Plot pdf
+pdf("outputs/tc_threshold_moist.pdf", width=8, height=8)
+par(mar=c(5,4,2,1),cex=1.4)
+plot(df_tc2000_moist$tc,df_tc2000_moist$cum_prop, type="l",
+     lwd=2,
+     xlab="Tree cover (%)",
+     ylab="Percentage of moist forest",
+     ylim=c(0,100), axes=FALSE)
+axis(1, at=seq(0,100,25), labels=seq(0,100,25))
+axis(2, at=seq(0,100,10), labels=seq(0,100,10))
+abline(v=75, lty=2)
+abline(h=90, lty=2)
+dev.off()
+
[…]

### 2018-03-13 manuscript/manuscript.Rnw
Histogram TC
Writting third cover letter

@@
   rate has progressively increased in Madagascar to reach 99,000 ha/yr
-  during 2010-2014 (corresponding to a rate of 1.1\%/yr). This
-  increase is probably due to rapid population growth (close to
-  3\%/yr) and to poor law enforcement in the country. Around half of
+  during 2010-2014 (corresponding to a rate of 1.1\%/yr). Around half of
   the forest (46\%) is now located at less than 100~m from the forest
-  edge. Accurate forest cover change maps can be used to assess the
+  edge. Our approach could be replicated to other developping countries with tropical forest.
+  Accurate forest cover change maps can be used to assess the
   effectiveness of past and current conservation programs and
@@
 air moisture) can be dramatically modified, with consequences on the
-abundance and distribution of species \citep{Murcia1995}. Forest
+abundance and distribution of species \citep{Murcia1995, Broadbent2008, Gibson2013}. Forest
 fragmentation can also have substantial effects on forest carbon
@@
 mis-classification in specific areas, especially in the dry and spiny
-forest domain for which the spectral answer has a strong seasonal
-behavior due to the deciduousness of such forests (overall accuracy is
+forest domain for which the spectral signal shows strong seasonal
+variations due to the deciduousness of such forests (overall accuracy is
 lower than 0.8 for the dry and spiny forests for the maps produced by
@@
 
-The 2000 Harper's forest map includes 208,000 ha of unclassified areas due to 
-the presence of clouds on satellite images, mostly (88\%) within the moist 
-forest domain which covered
-4.17 Mha in total in 2000. To provide a label (forest or non-forest)
-to these unclassified pixels, we used the 2000 tree cover percentage
-map of \citet{Hansen2013} by selecting a threshold of 75\% tree cover
-to define forest cover as recommended by other studies for the moist
-domain \citep{Achard2014, Aleman2017}. To do so, the Hansen's 2000
-tree cover map was resampled on the same grid as the original Harper's
-map at 30~m resolution using a bilinear interpolation. We thus
-obtained a forest cover map for the year 2000 covering the full
-territory of Madagascar.
+The 2000 Harper's forest map includes 208,000 ha of unclassified areas due to the presence of clouds on satellite images.
+Unclassified areas were mostly (88\%) present within the moist forest domain which covered 4.17 Mha in 2000.
+To provide a label (forest or non-forest) to these unclassified pixels, we used the 2000 tree cover percentage map of \citet{Hansen2013} and selected a tree cover threshold of 75\% to define the forest \citep{Achard2014, Aleman2017}.
+This threshold allows to characterize properly the moist forest in Madagascar as 90\% of the moist forest in 2000 in \citet{Harper2007} has a tree cover greater than 75\% (Fig.~\ref{fig:tc_threshold}).
+For this step, the Hansen's 2000 tree cover map was resampled on the same grid as the original Harper's map at 30~m resolution using a bilinear interpolation.
+We thus obtained a forest cover map for the year 2000 covering the full territory of Madagascar.
 
@@
 the years 1953, 1973, 1990, 2000, 2005, 2010 and 2014 at 30~m
-resolution. We used a moving window of $51 \times 51$ pixels centered
+resolution. We used a moving window of $51 \times 51$ pixels (corresponding to an area of about 2.34~km$^2$) centered
 on each forest pixel to compute the percentage of forest pixels in the
 neighborhood. We used this percentage as an indication of the forest
-fragmentation. The size of the moving windows was based on a
+fragmentation \citep{Riitters2012, Vogt2017}. The size of the moving windows was based on a
 compromise: a sufficiently high number of cells (here 2601) had to be
@@
 number of cells had to be choosen to have a local estimate of the
-fragmentation. Computations were done using the function
+fragmentation.
+Water bodies were not masked when computing the percentage of forest pixels, meaning that forest located near a water body was considered as fragmented.
+Computations were done using the function
 \texttt{r.neighbors} of the GRASS GIS software
@@
 quantiles (5\% and 95\%) of the distance to forest edge and looked at
-the evolution of these values with time. Previous studies have shown
+the variation of these values over time. Previous studies have shown
 that forest micro-habitats were mainly altered within the first 100~m
-of the forest edge \citep{Brinck2017, Gibson2013, Murcia1995,
+of the forest edge \citep{Brinck2017, Murcia1995,
   Broadbent2008}. Consequently, we also estimated the percentage of
 forest within the first 100~m of the forest edge for each year and
-looked at the evolution of this percentage over the six decades.
+looked at the variation of this percentage over the six decades.
 
@@
 
-Natural forests in Madagascar covered 16.0 Mha in 1953, about 27\% of
-the national territory of 587,041 km2. In 2014, the forest cover
-dropped to 8.9 Mha, corresponding to about 15\% of the national
-territory (Fig.~\ref{fig:fcc} and
-Tab.~\ref{tab:forest_cover}). Madagascar has lost 44\% and 37\% of its
-natural forests between 1953 and 2014, and between 1973 and 2014
-respectively (Fig.~\ref{fig:fcc} and Tab.~\ref{tab:forest_cover}). In
-2014 the remaining 8.9 Mha of natural forest were distributed as
-follow: 4.4 Mha of moist forest (50\% of total forest cover), 2.6 Mha
-of dry forest (29\%), 1.7 Mha of spiny forest (19\%) and 0.18 Mha
-(2\%) of mangrove forest (Fig.~\ref{fig:ecoregion} and
-Tab.~\ref{tab:comp_forest}). Regarding the deforestation trend, we
+Natural forests in Madagascar covered 16.0 Mha in 1953, about 27\% of the national territory of 587,041 km2.
+In 2014, the forest cover dropped to 8.9 Mha, corresponding to about 15\% of the national territory (Fig.~\ref{fig:fcc} and Tab.~\ref{tab:forest_cover}).
+Madagascar has lost 44\% of its natural forest between 1953 and 2014, including 37\% between 1973 and 2014 (Fig.~\ref{fig:fcc} and Tab.~\ref{tab:forest_cover}).
+In 2014 the remaining 8.9 Mha of natural forest were distributed as follow: 4.4 Mha of moist forest (50\% of total forest cover), 2.6 Mha of dry forest (29\%), 1.7 Mha of spiny forest (19\%) and 0.18 Mha (2\%) of mangrove forest (Fig.~\ref{fig:ecoregion} and Tab.~\ref{tab:comp_forest}).
+
+The forest cover change map produced on the period 1953-2014 (Fig.~\ref{fig:fcc}) allows to identify hot-spots of deforestation.
+Among the many recent hot-spots of deforestation visible on the map for the period 2000-2014 , one is located at the south of the CAZ (\emph{``Corridor Ankeniheny Zahamena''}) protected area, in the moist forest at the east of Madagascar (see eastern zoom in Fig.~\ref{fig:fcc}).
+Another major hot-spot of deforestation is located around the Ranobe-PK32 new protected area, in the dry forest at the south-west of Madagascar (see western zoom in Fig.~\ref{fig:fcc}).
+
+Regarding the deforestation trend, we
 observed a progressive decrease of the deforestation rate after 1990
@@
 
-\subsection{Evolution of forest fragmentation with time}
+\subsection{Variation of forest fragmentation over time}
 
-In parallel to the dynamics of deforestation, forest fragmentation has
-progressively increased since 1953 in Madagascar. We observed a
+Forest fragmentation has progressively increased since 1953 in Madagascar. We observed a
 continuous decrease of the mean distance to forest edge from 1953 to
@@
 global tree cover dataset from \citet{Hansen2013} into a
-forest/non-forest map \citep{Tropek2014}. We propose to use a
+forest/non-forest map \citep{Tropek2014}. We propose the use of an
 historical national forest cover map, based on a national forest
@@
 but preferably at the beginning of the period to profit from the full
-record and derive long-term estimates of deforestation. Moreover, this
+record of tree cover loss and derive long-term estimates of deforestation. Moreover, this
 approach could be repeated in the future with the release of updated tree cover loss data. 
@@
 than doubled on the period 2010-2014 compared to 2000-2005. Our
-results are confirmed by previous studies \citep{Harper2007, MEFT2009,
+results are supported by previous studies \citep{Harper2007, MEFT2009,
   ONE2015} despite differences in the methodologies regarding (i)
@@
 at 30~m resolution and can be considered more accurate in comparison
-with estimates from these previous studies. Our forest cover and
+with estimates from these previous studies. Our natural forest cover and
 deforestation rate estimates can be used as source of information for
@@
 # Caption
-tabcap1 <- "\\textbf{Evolution of natural forest cover and deforestation rates from 1953 to 2014 in Madagascar.} Areas are provided in thousands of hectares (Kha).
+tabcap1 <- "\\textbf{Change in natural forest cover and deforestation rates from 1953 to 2014 in Madagascar.} Areas are provided in thousands of hectares (Kha).
 Forest map for the year 1973 has 3.3 Mha of unclassified areas due to the presence of clouds on satellite images.
@@
 # Caption
-tabcap4 <- "\\textbf{Evolution of the forest fragmentation from 1953 to
+tabcap4 <- "\\textbf{Change in forest fragmentation from 1953 to
       2014 in Madagascar}. Five forest fragmentation classes, based on
@@
 
-  \caption{\textbf{Evolution of the distance to forest edge from 1953
+  \caption{\textbf{Change in distance to forest edge from 1953
       to 2014 in Madagascar.} Black dots represent the mean distance
@@
 
+\newpage
+
+\section{Supplementary material}
+
+\setcounter{figure}{0}
+\renewcommand{\thefigure}{A\arabic{figure}}
+
+\vfill
+\begin{figure}[h!]
+  \centering
+
+  \includegraphics[width=12cm]{figs/tc_threshold_moist.pdf}
+
+  \caption{\textbf{Selection of the tree cover threshold to define the land type for unclassified areas in 2000.} We considered the forest in 2000 \citep{Harper2007} in the moist ecoregion (see Fig.~\ref{fig:ecoregion}).
+We plotted the percentage of forest having a tree cover greater or equal to a given value specified by the x-axis.
+From this figure, we see that 90\% of the moist forest in 2000 has a tree cover $\geq$ 75.}
+
+  \label{fig:tc_threshold}
+
+\end{figure}
+\vfill
+
 \end{document}

### 2018-03-13 deforestmap.R

@@ -179,6 +179,47 @@ ha.cloud.2000.moist <- sum(report.harper.moist$area[report.harper.moist$code %in
 perc.cloud.moist <- 100*ha.cloud.2000.moist/ha.cloud.2000 # 88%
 SavedObjects <- c(SavedObjects, "ha.cloud.2000.moist", "perc.cloud.moist")
 
+##========================
+## Histogram of tree cover
+##========================
+
+# Moist forest 2000 in Harper
+system("r.mapcalc --o 'for2000_Harper_moist = if((harper==111 || harper==112 || harper==115 || \\
+       harper==775 || harper==777) && ecoregion==3, 1, null())'")
+system("r.mask --o raster=for2000_Harper_moist")
+system("r.mapcalc --o 'tc2000_moist = tc2000'")
+system("r.mask -r")
+hist_tc2000_moist <- system("r.stats -c input=tc2000_moist nsteps=100", intern=TRUE)
+df_tc2000_moist <- data.frame(matrix(unlist(strsplit(hist_tc2000_moist, " ")), ncol=2, byrow=TRUE))
+df_tc2000_moist <- df_tc2000_moist[-c(101),]
+names(df_tc2000_moist) <- c("tc","npix")
+df_tc2000_moist$npix <- as.numeric(as.character(df_tc2000_moist$npix))
+df_tc2000_moist$tc <- as.numeric(as.character(df_tc2000_moist$tc))
+df_tc2000_moist$nha <- df_tc2000_moist$npix*30*30/10000
+df_tc2000_moist$freq <- 100*df_tc2000_moist$npix/sum(df_tc2000_moist$npix)
+# Cumulative proportion
+for (i in 100:1) {
+  df_tc2000_moist$cum_prop[i] <- 100*sum(df_tc2000_moist$npix[100:i])/sum(df_tc2000_moist$npix)
+}
+write.table(df_tc2000_moist, file="outputs/df_tc2000_moist.txt", sep="\t", row.names=FALSE)
+# How much moist forest has a TC over 75%
+prop_tc75 <- round(df_tc2000_moist$cum_prop[df_tc2000_moist$tc==75]) # 90%
+prop_tc70 <- round(df_tc2000_moist$cum_prop[df_tc2000_moist$tc==70]) # 92%
+
+# Plot pdf
+pdf("outputs/tc_threshold_moist.pdf", width=8, height=8)
+par(mar=c(5,4,2,1),cex=1.4)
[…]

### 2018-03-14 manuscript/manuscript.Rnw
Update cover letter

@@
 compared to old-growth natural forests in terms of biodiversity and
-carbon storage.
+carbon storage \citep{Martin2013}.
 

### 2018-03-14 manuscript/manuscript.Rnw
Update cover letter

@@
 
-Regarding the deforestation trend, we
-observed a progressive decrease of the deforestation rate after 1990
-from 205,000 ha/yr (1.6\%/yr) over the period 1973-1990 to 42,000
-ha/yr (0.4\%/yr) over the period 2000-2005
-(Tab.~\ref{tab:forest_cover}). Then from 2005, the deforestation rate
-has progressively increased and has more than doubled over the period
-2010-2014 (99,000 ha/yr, 1.1\%/yr) compared to 2000-2005
-(Tab.~\ref{tab:forest_cover}). The deforestation trend, characterized
-by a progressive decrease of the deforestation rate over the period
-1990-2005 and a progressive increase of the deforestation after 2005,
-is valid for all four forest types except the spiny forest
-(Tab.~\ref{tab:comp_defor}). For the spiny forest, the deforestation
-rate during the period 2010-2013 was lower than on the period
-2005-2010 (Tab.~\ref{tab:comp_defor}).
+Regarding the deforestation trend, we observed a progressive decrease of the deforestation rate after 1990 from 205,000 ha/yr (1.6\%/yr) over the period 1973-1990 to 42,000 ha/yr (0.4\%/yr) over the period 2000-2005 (Tab.~\ref{tab:forest_cover}).
+Then from 2005, the deforestation rate has progressively increased and has more than doubled over the period 2010-2014 (99,000 ha/yr, 1.1\%/yr) compared to 2000-2005 (Tab.~\ref{tab:forest_cover}).
+The deforestation trend, characterized by a progressive decrease of the deforestation rate over the period 1990-2005 and a progressive increase of the deforestation after 2005, is valid for all four forest types except the spiny forest (Tab.~\ref{tab:comp_defor}).
+For the spiny forest, the deforestation rate during the period 2010-2013 was lower than on the period 2005-2010 (Tab.~\ref{tab:comp_defor}).
 
@@
 wall-to-wall forest cover change maps from 1990 to 2014 for the full
-territory of Madagascar (Tab.~\ref{tab:forest_cover}). This allowed us
+territory of Madagascar (Fig.~\ref{fig:fcc}). This allowed us
 to produce more robust estimates of forest cover and deforestation
-rates over this period. Our forest cover estimates over the period
+rates over this period (Tab.~\ref{tab:forest_cover}). Our forest cover estimates over the period
 1953-2013 (considering forest cover estimates at national level and by
@@
 (corresponding to 15\% of the country) and that Madagascar has lost
-44\% of its natural forest since 1953 (37\% since 1973). There is
-ongoing scientific debate about the extent of the ``original'' forest
+44\% of its natural forest since 1953 (37\% since 1973). If there are no doubts 
+about the direct causes of deforestation in Madagascar, attributable to human 
+activities such as slash-and-burn agriculture and pasture \citep{Scales2011}, there 
+is ongoing scientific debate about the extent of the ``original'' forest
 cover in Madagascar, and the extent to which humans have altered the
@@
 Madagascar. Around 90\% of Madagascar's species are forest dependent
-\citep{Allnutt2008, Goodman2005} and \citet{Allnutt2008} estimated
-that deforestation between 1953 and 2000 led to an extinction of 9\%
+\citep{Allnutt2008, Goodman2005}. Based on occurrence data for 
+2243 species of plants and invertebrates, \citet{Allnutt2008} estimated
+that deforestation between 1953 and 2000 has led to an extinction of 9\%
 of the species. The additional deforestation we observed over the
@@
 
-All authors conceived the ideas and designed methodology; GV analysed
+All authors conceived the ideas and designed the methodology; GV analysed
 the data and wrote the {\R}/GRASS script; GV drafted the manuscript. All

### 2018-03-16 manuscript/manuscript.Rnw
Update cover letter and manuscript

@@
   the forest (46\%) is now located at less than 100~m from the forest
-  edge. Our approach could be replicated to other developping countries with tropical forest.
+  edge. Our approach could be replicated to other developing countries with tropical forest.
   Accurate forest cover change maps can be used to assess the
@@
 agriculture and pasture \citep{Scales2011}. Tropical forests in
-Madagascar also store a large amount of carbon \citep{Vieilledent2016}
-and high rates of deforestation in Madagascar are responsible for
-large CO$_2$ emissions in the atmosphere
-\citep{Achard2014}. Deforestation threatens species survival by
+Madagascar also store a large amount of carbon (136 MgC.ha$^{-1}$ in the moist
+forest, \citealp{Vieilledent2016}) and high rates of deforestation in Madagascar
+(1.4--4.7~\%/yr, \citealp{Achard2002}) are responsible for large CO$_2$ emissions 
+in the atmosphere. Deforestation threatens species survival by
 directly reducing their available habitat \citep{Brooks2002,
@@
 fragmentation can also have substantial effects on forest carbon
-storage capacity, as carbon stocks are much lower at the forest edge
+storage capacity, as carbon stocks are about 50\% lower at the forest edge
 than under a closed canopy \citep{Brinck2017}. Moreover, forest
@@
 1953 forest map is a vector map derived from the visual interpretation
-of aerial photographies. Forest maps for
+of aerial photographs. Forest maps for
 the years 1973, 1990, and 2000 were obtained from the supervised
@@
 mis-classification in specific areas, especially in the dry and spiny
-forest domain for which the spectral signal shows strong seasonal
+forest domain, for which the spectral signal shows strong seasonal
 variations due to the deciduousness of such forests (overall accuracy is
@@
 the paper map of \citet{Humbert1965} which was derived from the visual interpretation 
-of aerial photographies. We reprojected the forest cover map of year 1953 in the WGS 84/UTM zone 
+of aerial photographs. We reprojected the forest cover map of year 1953 in the WGS 84/UTM zone 
 38S projected coordinate system. Because of the methodology used to derive the 1953 forest 
@@
 maps of later years which were produced through digital processing of satellite imagery. 
-As a consequence, the 1953 cannot be merged with the map of laters years to identify 
+As a consequence, the 1953 cannot be merged with the map of later years to identify 
 precisely the location of the deforested areas. Nonetheless, the 1953 forest cover 
@@
 about 600,000 ha for year 2010. This allowed us to avoid counting very small scale
-events (\textless 0.1 ha, such as selective logging or windthrow) as
+events (\textless 0.1 ha, such as selective logging or wind-throw) as
 deforestation. It also prevents us from underestimating forest cover and overestimating 
@@
 considered to be able to compute a percentage and a reasonably low
-number of cells had to be choosen to have a local estimate of the
+number of cells had to be chosen to have a local estimate of the
 fragmentation.
@@
 definition, as a forest cover mask. This approach could be easily
-extended to other regions or countries for which an accurate
-forest cover map is available at any date within the period 2000-2014,
-but preferably at the beginning of the period to profit from the full
-record of tree cover loss and derive long-term estimates of deforestation. Moreover, this
-approach could be repeated in the future with the release of updated tree cover loss data. 
-We have made the \R{}/GRASS code used for this
+extended to other tropical regions or countries for which an accurate
+forest cover map is available at any date within the period 2000-2014 
+(but preferably at the beginning of the period to profit from the full
+record of tree cover loss and derive long-term estimates of deforestation). For example, 
+forest cover maps are available at 20~m resolution for Cameroon and Central 
+African Republic for years 2000 and 2010 \citep{Gross2017}. When high resolution
+forest cover maps are not available, coarser resolution forest cover maps
+(leading to coarser deforestation estimates) could be extracted from global
+land cover products such as GLC2000 at 1~km resolution \citep{Bartholome2005}
+or CCI Land Cover at 300~m resolution \citep{Li2018}.
+Moreover, this approach could be repeated in the future with the release of
+updated tree cover loss data. We have made the \R{}/GRASS code used for this
 study freely available in a GitHub repository (see Data availability
@@
 The accuracy of the derived forest cover change maps depends directly
-on the accuracies of the historical forest cover maps and the tree
+on the accuracy of the historical forest cover maps and the tree
 cover loss dataset. Using visual-interpretation of aerial images in
@@
 and 16.9\% of false negatives (see Tab. S5 in
-\citet{Hansen2013}). These numbers rise at 20.7\% and 20.6\%
+\citealt{Hansen2013}). These numbers rise at 20.7\% and 20.6\%
 respectively for the subtropical biome. In the subtropical biome, the
@@
 spiny forest. In another study assessing the accuracy of the tree
-cover loss product accross the tropics \citep{Tyukavina2015}, authors
+cover loss product across the tropics \citep{Tyukavina2015}, authors
 reported 4\% of false positives and 48\% of false negatives in
-Sub-Saharian Africa. They showed that 85\% of missing loss occured on
+Sub-Saharan Africa. They showed that 85\% of missing loss occurred on
 the edges of other loss patches. This means that tree cover loss might
-be underestimated in Sub-Saharian Africa, probably due to the
+be underestimated in Sub-Saharan Africa, probably due to the
 prevalence of small-scale disturbance which is hard to map at 30~m,
@@
 Madagascar. Around 90\% of Madagascar's species are forest dependent
-\citep{Allnutt2008, Goodman2005}. Based on occurrence data for 
-2243 species of plants and invertebrates, \citet{Allnutt2008} estimated
+\citep{Allnutt2008, Goodman2005}. Based on occurrence data for 2243 plant and
+invertebrate species, \citet{Allnutt2008} estimated
 that deforestation between 1953 and 2000 has led to an extinction of 9\%
@@
 deforestation rate observed on the period 2005-2014
-\citep{Smith2003}. These results show that despite the conservation
+(see \citealp{Smith2003} for a discussion on the link between governance and
+forest cover loss). These results show that despite the conservation
 policy in Madagascar \citep{Freudenberger2010}, deforestation has
@@
 The authors thank two anonymous reviewers for useful comments on a previous
-version of the manuscript. They also thank Peter Vogt for useful advices on which
+version of the manuscript. They also thank Peter Vogt for useful advice on which
 metric to use to estimate forest fragmentation. This study is part of
@@
 tabcap3 <- "\\textbf{Comparing Madagascar annual deforestation rates
-      with previous studies on the period 1953-2014}. Annual
+      with previous studies on the period 1953-2013}. Annual
     deforested areas (in thousands of hectares per year, Kha/yr) and

### 2018-03-16 manuscript/manuscript.Rnw
Arrange text

@@
 species and endemism in many taxonomic groups \citep{Crottini2012,
-  Goodman2005}. Most of the biodiversity in Madagascar is
-concentrated in the tropical forests of the island which can be
-divided into four types: the moist forest in the East, the dry forest
-in the West, the spiny forest in the South and the mangroves on the
-West coast \citep{Vieilledent2016}. This unparalleled biodiversity is
-severely threatened by deforestation \citep{Harper2007,
-  Vieilledent2013} associated with human activities such as slash-and-burn
-agriculture and pasture \citep{Scales2011}. Tropical forests in
-Madagascar also store a large amount of carbon (136 MgC.ha$^{-1}$ in the moist
-forest, \citealp{Vieilledent2016}) and high rates of deforestation in Madagascar
-(1.4--4.7~\%/yr, \citealp{Achard2002}) are responsible for large CO$_2$ emissions 
-in the atmosphere. Deforestation threatens species survival by
-directly reducing their available habitat \citep{Brooks2002,
-  Tidd2001}. Forest fragmentation can also lead to species extinction
-by isolating populations from each other and creating forest patches
-too small to maintain viable populations
+  Goodman2005}. Most of the biodiversity in Madagascar is concentrated
+in the tropical forests of the island which can be divided into four
+types: the moist forest in the East, the dry forest in the West, the
+spiny forest in the South and the mangroves on the West coast
+\citep{Vieilledent2016}. This unparalleled biodiversity is severely
+threatened by deforestation \citep{Harper2007, Vieilledent2013}
+associated with human activities such as slash-and-burn agriculture
+and pasture \citep{Scales2011}. Tropical forests in Madagascar also
+store a large amount of carbon (136 MgC.ha$^{-1}$ in the moist forest,
+\citealp{Vieilledent2016}) and high rates of deforestation in
+Madagascar (1.4--4.7~\%/yr, \citealp{Achard2002}) are responsible for
+large CO$_2$ emissions in the atmosphere. Deforestation threatens
+species survival by directly reducing their available habitat
+\citep{Brooks2002, Tidd2001}. Forest fragmentation can also lead to
+species extinction by isolating populations from each other and
+creating forest patches too small to maintain viable populations
 \citep{Saunders1991}. Fragmentation also increases forest edge where
@@
 air moisture) can be dramatically modified, with consequences on the
-abundance and distribution of species \citep{Murcia1995, Broadbent2008, Gibson2013}. Forest
-fragmentation can also have substantial effects on forest carbon
-storage capacity, as carbon stocks are about 50\% lower at the forest edge
-than under a closed canopy \citep{Brinck2017}. Moreover, forest
-carbon stocks vary spatially due to climate or soil factors
-\citep{Saatchi2011, Vieilledent2016}. As a consequence, accurate and
-spatially explicit maps of forest cover and forest cover change are
-necessary to monitor biodiversity loss and carbon emissions from
-deforestation and forest fragmentation, assess the efficiency of
-present conservation strategies \citep{Eklund2016}, and implement new
-strategies for the future \citep{Vieilledent2013,
-  Vieilledent2016}. Simple time-series of forest cover estimates, such
-as those provided by the FAO Forest Resource Assessment report
-\citep{Keenan2015} are not sufficient.
+abundance and distribution of species \citep{Murcia1995,
+  Broadbent2008, Gibson2013}. Forest fragmentation can also have
+substantial effects on forest carbon storage capacity, as carbon
+stocks are about 50\% lower at the forest edge than under a closed
+canopy \citep{Brinck2017}. Moreover, forest carbon stocks vary
+spatially due to climate or soil factors \citep{Saatchi2011,
+  Vieilledent2016}. As a consequence, accurate and spatially explicit
+maps of forest cover and forest cover change are necessary to monitor
+biodiversity loss and carbon emissions from deforestation and forest
+fragmentation, assess the efficiency of present conservation
+strategies \citep{Eklund2016}, and implement new strategies for the
+future \citep{Vieilledent2013, Vieilledent2016}. Simple time-series of
+forest cover estimates, such as those provided by the FAO Forest
+Resource Assessment report \citep{Keenan2015} are not sufficient.
 
@@
 maps of forest cover and forest cover changes over Madagascar for the
-years 1953, 1973, 1990 and 2000. The
-1953 forest map is a vector map derived from the visual interpretation
-of aerial photographs. Forest maps for
-the years 1973, 1990, and 2000 were obtained from the supervised
-classification of Landsat satellite images and can be
-used to derive more accurate estimates of forest cover than those from the FAO
-Forest Resource Assessment report. Nonetheless, maps provided by 
-\citet{Harper2007} are not exhaustive (due to the presence of clouds in the satellite imagery),
-e.g.~11 244~km2 are mapped as unknown cover type for the year
-2000. Using a similar supervised classification approach as in
-\citet{Harper2007}, more recent maps have been produced for the
-periods 2000-2005-2010 by national institutions, with the technical
-support of international environmental NGOs \citep{MEFT2009,
+years 1953, 1973, 1990 and 2000. The 1953 forest map is a vector map
+derived from the visual interpretation of aerial photographs. Forest
+maps for the years 1973, 1990, and 2000 were obtained from the
+supervised classification of Landsat satellite images and can be used
+to derive more accurate estimates of forest cover than those from the
+FAO Forest Resource Assessment report. Nonetheless, maps provided by
+\citet{Harper2007} are not exhaustive (due to the presence of clouds
+in the satellite imagery), e.g.~11 244~km2 are mapped as unknown cover
+type for the year 2000. Using a similar supervised classification
+approach as in \citet{Harper2007}, more recent maps have been produced
+for the periods 2000-2005-2010 by national institutions, with the
+technical support of international environmental NGOs \citep{MEFT2009,
   ONE2013}. Another set of recent forest cover maps using an advanced
@@
 forest domain, for which the spectral signal shows strong seasonal
-variations due to the deciduousness of such forests (overall accuracy is
-lower than 0.8 for the dry and spiny forests for the maps produced by
-\citet{ONE2015}). Moreover, the production of such forest maps from a
-supervised classification approach requires significant resources,
+variations due to the deciduousness of such forests (overall accuracy
+is lower than 0.8 for the dry and spiny forests for the maps produced
+by \citet{ONE2015}). Moreover, the production of such forest maps from
+a supervised classification approach requires significant resources,
 especially regarding the image selection step (required to minimize
@@
 historical forest maps from \citet{Harper2007} and more recent global
-products from \citet{Hansen2013} to derive annual wall-to-wall
-forest cover change maps over the period 2000-2014 for Madagascar. We
-use the forest cover map provided by \citet{Harper2007} for the year
-2000 (defining the land-use) with the tree cover loss product provided
-by \citet{Hansen2013} that we apply only inside forest areas
-identified by \citet{Harper2007}. Similar to the approach of
-\citet{Harper2007}, we also assess trends in deforestation rates and
-forest fragmentation from 1953 to 2014. We finally discuss
-the possibility to extend our approach to other tropical countries or
-repeat it in the future for Madagascar. We also discuss how our results could help
-assess the effectiveness of past and current conservation strategies in Madagascar, and
-implement new strategies in the future.
+products from \citet{Hansen2013} to derive annual wall-to-wall forest
+cover change maps over the period 2000-2014 for Madagascar. We use the
+forest cover map provided by \citet{Harper2007} for the year 2000
+(defining the land-use) with the tree cover loss product provided by
+\citet{Hansen2013} that we apply only inside forest areas identified
+by \citet{Harper2007}. Similar to the approach of \citet{Harper2007},
+we also assess trends in deforestation rates and forest fragmentation
+from 1953 to 2014. We finally discuss the possibility to extend our
+approach to other tropical countries or repeat it in the future for
+Madagascar. We also discuss how our results could help assess the
+effectiveness of past and current conservation strategies in
+Madagascar, and implement new strategies in the future.
 
@@
 Original 1990-2000 forest cover change map for Madagascar from
-\citet{Harper2007} is a raster map at 28.5 m resolution. It was
-derived from the supervised classification of Landsat TM (Thematic Mapper) and 
-ETM+ (Enhanced Thematic Mapper Plus) satellite images. For our study, this map 
-has been resampled at 30~m resolution 
-using a nearest-neighbor interpolation and reprojected in the WGS 84/UTM zone 
+\citet{Harper2007} is a raster map at 28.5~m resolution. It was
+derived from the supervised classification of Landsat TM (Thematic
+Mapper) and ETM+ (Enhanced Thematic Mapper Plus) satellite images. For
+our study, this map has been resampled at 30~m resolution using a
+nearest-neighbor interpolation and reprojected in the WGS 84/UTM zone
 38S projected coordinate system.
 
-The 2000 Harper's forest map includes 208,000 ha of unclassified areas due to the presence of clouds on satellite images.
-Unclassified areas were mostly (88\%) present within the moist forest domain which covered 4.17 Mha in 2000.
-To provide a label (forest or non-forest) to these unclassified pixels, we used the 2000 tree cover percentage map of \citet{Hansen2013} and selected a tree cover threshold of 75\% to define the forest \citep{Achard2014, Aleman2017}.
-This threshold allows to characterize properly the moist forest in Madagascar as 90\% of the moist forest in 2000 in \citet{Harper2007} has a tree cover greater than 75\% (Fig.~\ref{fig:tc_threshold}).
-For this step, the Hansen's 2000 tree cover map was resampled on the same grid as the original Harper's map at 30~m resolution using a bilinear interpolation.
-We thus obtained a forest cover map for the year 2000 covering the full territory of Madagascar.
+The 2000 Harper's forest map includes 208,000 ha of unclassified areas
+due to the presence of clouds on satellite images. Unclassified areas
+were mostly (88\%) present within the moist forest domain which
+covered 4.17 Mha in 2000. To provide a label (forest or non-forest) to
+these unclassified pixels, we used the 2000 tree cover percentage map
+of \citet{Hansen2013} and selected a tree cover threshold of 75\% to
+define the forest \citep{Achard2014, Aleman2017}. This threshold
+allows to characterize properly the moist forest in Madagascar as 90\%
+of the moist forest in 2000 in \citet{Harper2007} has a tree cover
+greater than 75\% (Fig.~\ref{fig:tc_threshold}). For this step, the
+Hansen's 2000 tree cover map was resampled on the same grid as the
+original Harper's map at 30~m resolution using a bilinear
+interpolation. We thus obtained a forest cover map for the year 2000
+covering the full territory of Madagascar.
 
-We then combined the forest cover map of the
-year 2000 with the annual tree cover loss maps from 2001 to 2014
-from \citet{Hansen2013} to create annual forest cover maps from
-2001 to 2014 at 30~m resolution. To do so, Hansen's tree cover loss
-maps were resampled on the same grid as the original Harper's map at
-30~m resolution using a nearest-neighbor interpolation. We also completed 
-the Harper's forest map of year 1990 by filling unclassified
-areas (due to the presence of clouds on satellite images) using our
-forest cover map of year 2000. To do so, we assumed that if forest was
-present in 2000, the pixel was also forested in 1990. Indeed, there is little 
-evidence of natural forest regeneration in Madagascar 
-\citep{Grouzis2001, Harper2007}, especially over such a short period of time. 
-The remaining unclassified pixels were limited to a relatively small total area 
-of about 8,000 ha. We labeled these residual pixels as non-forest, as
-for the year 2000.
+We then combined the forest cover map of the year 2000 with the annual
+tree cover loss maps from 2001 to 2014 from \citet{Hansen2013} to
+create annual forest cover maps from 2001 to 2014 at 30~m
+resolution. To do so, Hansen's tree cover loss maps were resampled on
+the same grid as the original Harper's map at 30~m resolution using a
+nearest-neighbor interpolation. We also completed the Harper's forest
+map of year 1990 by filling unclassified areas (due to the presence of
+clouds on satellite images) using our forest cover map of year
+2000. To do so, we assumed that if forest was present in 2000, the
+pixel was also forested in 1990. Indeed, there is little evidence of
+natural forest regeneration in Madagascar \citep{Grouzis2001,
+  Harper2007}, especially over such a short period of time.  The
+remaining unclassified pixels were limited to a relatively small total
+area of about 8,000 ha. We labeled these residual pixels as
+non-forest, as for the year 2000.
 
 The 1973 forest cover map for Madagascar from \citet{Harper2007} is a
-raster map at 57 m resolution derived from the supervised classification of 
-Landsat MSS (Multispectral Scanner System) satellite images. We resampled this 
-map at 30~m resolution using a nearest-neighbor interpolation on the same grid 
-as the forest cover maps for years 1990 and 2000. We completed the Harper's 
-forest map of year 1973 by filling unclassified areas using our forest cover 
-map of the year 1990 assuming that if forest was present in 1990, it was also
+raster map at 57~m resolution derived from the supervised
+classification of Landsat MSS (Multispectral Scanner System) satellite
+images. We resampled this map at 30~m resolution using a
+nearest-neighbor interpolation on the same grid as the forest cover
+maps for years 1990 and 2000. We completed the Harper's forest map of
+year 1973 by filling unclassified areas using our forest cover map of
+the year 1990 assuming that if forest was present in 1990, it was also
 present in 1973. Contrary to the year 1990, the remaining unclassified
@@
 
-The 1953 forest cover map from \citet{Harper2007} is a vector map produced by scanning 
-the paper map of \citet{Humbert1965} which was derived from the visual interpretation 
-of aerial photographs. We reprojected the forest cover map of year 1953 in the WGS 84/UTM zone 
-38S projected coordinate system. Because of the methodology used to derive the 1953 forest 
-cover map, it was not possible to perfectly aligned this map with the forest cover 
-maps of later years which were produced through digital processing of satellite imagery. 
-As a consequence, the 1953 cannot be merged with the map of later years to identify 
-precisely the location of the deforested areas. Nonetheless, the 1953 forest cover 
-map can be used to have a rough estimate of the forest cover and forest fragmentation at this date.
-To do so, the map was rasterized at 30 m resolution on the same grid 
-as the forest cover maps for years 1973, 1990 and 2000.
+The 1953 forest cover map from \citet{Harper2007} is a vector map
+produced by scanning the paper map of \citet{Humbert1965} which was
+derived from the visual interpretation of aerial photographs. We
+reprojected the forest cover map of year 1953 in the WGS 84/UTM zone
+38S projected coordinate system. Because of the methodology used to
+derive the 1953 forest cover map, it was not possible to perfectly
+aligned this map with the forest cover maps of later years which were
+produced through digital processing of satellite imagery. As a
+consequence, the 1953 cannot be merged with the map of later years to
+identify precisely the location of the deforested areas. Nonetheless,
+the 1953 forest cover map can be used to have a rough estimate of the
+forest cover and forest fragmentation at this date.  To do so, the map
+was rasterized at 30~m resolution on the same grid as the forest cover
+maps for years 1973, 1990 and 2000.
 
-Finally for all forest cover maps from 1973, isolated single non-forest pixels
-(i.e.~fully surrounded by forest pixels) were recategorized as forest pixels.
-Doing so, forest cover increased from about 95,000 ha for year 1953 to
-about 600,000 ha for year 2010. This allowed us to avoid counting very small scale
-events (\textless 0.1 ha, such as selective logging or wind-throw) as
-deforestation. It also prevents us from underestimating forest cover and overestimating 
-forest fragmentation. All the resulting maps have been made permanently and
-freely available on the Zenodo research data repository at
-\url{https://doi.org/10.5281/zenodo.1145785}.
+Finally for all forest cover maps from 1973, isolated single
+non-forest pixels (i.e.~fully surrounded by forest pixels) were
+recategorized as forest pixels.  Doing so, forest cover increased from
+about 95,000~ha for year 1953 to about 600,000~ha for year 2010. This
+allowed us to avoid counting very small scale events (\textless 0.1
+ha, such as selective logging or wind-throw) as deforestation. It also
+prevents us from underestimating forest cover and overestimating
+forest fragmentation. All the resulting maps have been made
+permanently and freely available on the Zenodo research data
+repository at \url{https://doi.org/10.5281/zenodo.1145785}.
 
@@
 the years 1953, 1973, 1990, 2000, 2005, 2010 and 2014 at 30~m
-resolution. We used a moving window of $51 \times 51$ pixels (corresponding to an area of about 2.34~km$^2$) centered
-on each forest pixel to compute the percentage of forest pixels in the
+resolution. We used a moving window of $51 \times 51$ pixels
+(corresponding to an area of about 2.34~km$^2$) centered on each
+forest pixel to compute the percentage of forest pixels in the
 neighborhood. We used this percentage as an indication of the forest
-fragmentation \citep{Riitters2012, Vogt2017}. The size of the moving windows was based on a
-compromise: a sufficiently high number of cells (here 2601) had to be
-considered to be able to compute a percentage and a reasonably low
-number of cells had to be chosen to have a local estimate of the
-fragmentation.
-Water bodies were not masked when computing the percentage of forest pixels, meaning that forest located near a water body was considered as fragmented.
-Computations were done using the function
-\texttt{r.neighbors} of the GRASS GIS software
+fragmentation \citep{Riitters2012, Vogt2017}. The size of the moving
+windows was based on a compromise: a sufficiently high number of cells
+(here 2601) had to be considered to be able to compute a percentage
+and a reasonably low number of cells had to be chosen to have a local
+estimate of the fragmentation. Water bodies were not masked when
+computing the percentage of forest pixels, meaning that forest located
+near a water body was considered as fragmented. Computations were done
+using the function \texttt{r.neighbors} of the GRASS GIS software
 \citep{Neteler2008}. Using the density of forest in the neighborhood,
@@
 
-Natural forests in Madagascar covered 16.0 Mha in 1953, about 27\% of the national territory of 587,041 km2.
-In 2014, the forest cover dropped to 8.9 Mha, corresponding to about 15\% of the national territory (Fig.~\ref{fig:fcc} and Tab.~\ref{tab:forest_cover}).
-Madagascar has lost 44\% of its natural forest between 1953 and 2014, including 37\% between 1973 and 2014 (Fig.~\ref{fig:fcc} and Tab.~\ref{tab:forest_cover}).
-In 2014 the remaining 8.9 Mha of natural forest were distributed as follow: 4.4 Mha of moist forest (50\% of total forest cover), 2.6 Mha of dry forest (29\%), 1.7 Mha of spiny forest (19\%) and 0.18 Mha (2\%) of mangrove forest (Fig.~\ref{fig:ecoregion} and Tab.~\ref{tab:comp_forest}).
+Natural forests in Madagascar covered 16.0 Mha in 1953, about 27\% of
+the national territory of 587,041 km2. In 2014, the forest cover
+dropped to 8.9 Mha, corresponding to about 15\% of the national
+territory (Fig.~\ref{fig:fcc} and
+Tab.~\ref{tab:forest_cover}). Madagascar has lost 44\% of its natural
+forest between 1953 and 2014, including 37\% between 1973 and 2014
+(Fig.~\ref{fig:fcc} and Tab.~\ref{tab:forest_cover}). In 2014 the
+remaining 8.9 Mha of natural forest were distributed as follow: 4.4
+Mha of moist forest (50\% of total forest cover), 2.6 Mha of dry
+forest (29\%), 1.7 Mha of spiny forest (19\%) and 0.18 Mha (2\%) of
+mangrove forest (Fig.~\ref{fig:ecoregion} and
+Tab.~\ref{tab:comp_forest}).
 
-The forest cover change map produced on the period 1953-2014 (Fig.~\ref{fig:fcc}) allows to identify hot-spots of deforestation.
-Among the many recent hot-spots of deforestation visible on the map for the period 2000-2014 , one is located at the south of the CAZ (\emph{``Corridor Ankeniheny Zahamena''}) protected area, in the moist forest at the east of Madagascar (see eastern zoom in Fig.~\ref{fig:fcc}).
-Another major hot-spot of deforestation is located around the Ranobe-PK32 new protected area, in the dry forest at the south-west of Madagascar (see western zoom in Fig.~\ref{fig:fcc}).
+The forest cover change map produced on the period 1953-2014
+(Fig.~\ref{fig:fcc}) allows to identify hot-spots of
+deforestation. Among the many recent hot-spots of deforestation
+visible on the map for the period 2000-2014 , one is located at the
+south of the CAZ (\emph{``Corridor Ankeniheny Zahamena''}) protected
+area, in the moist forest at the east of Madagascar (see eastern zoom
+in Fig.~\ref{fig:fcc}). Another major hot-spot of deforestation is
+located around the Ranobe-PK32 new protected area, in the dry forest
+at the south-west of Madagascar (see western zoom in
+Fig.~\ref{fig:fcc}).
 
-Regarding the deforestation trend, we observed a progressive decrease of the deforestation rate after 1990 from 205,000 ha/yr (1.6\%/yr) over the period 1973-1990 to 42,000 ha/yr (0.4\%/yr) over the period 2000-2005 (Tab.~\ref{tab:forest_cover}).
-Then from 2005, the deforestation rate has progressively increased and has more than doubled over the period 2010-2014 (99,000 ha/yr, 1.1\%/yr) compared to 2000-2005 (Tab.~\ref{tab:forest_cover}).
-The deforestation trend, characterized by a progressive decrease of the deforestation rate over the period 1990-2005 and a progressive increase of the deforestation after 2005, is valid for all four forest types except the spiny forest (Tab.~\ref{tab:comp_defor}).
-For the spiny forest, the deforestation rate during the period 2010-2013 was lower than on the period 2005-2010 (Tab.~\ref{tab:comp_defor}).
+Regarding the deforestation trend, we observed a progressive decrease
+of the deforestation rate after 1990 from 205,000 ha/yr (1.6\%/yr)
+over the period 1973-1990 to 42,000 ha/yr (0.4\%/yr) over the period
+2000-2005 (Tab.~\ref{tab:forest_cover}). Then from 2005, the
+deforestation rate has progressively increased and has more than
+doubled over the period 2010-2014 (99,000 ha/yr, 1.1\%/yr) compared to
+2000-2005 (Tab.~\ref{tab:forest_cover}). The deforestation trend,
+characterized by a progressive decrease of the deforestation rate over
+the period 1990-2005 and a progressive increase of the deforestation
+after 2005, is valid for all four forest types except the spiny forest
+(Tab.~\ref{tab:comp_defor}). For the spiny forest, the deforestation
+rate during the period 2010-2013 was lower than on the period
+2005-2010 (Tab.~\ref{tab:comp_defor}).
 
@@
 wall-to-wall forest cover change maps from 1990 to 2014 for the full
-territory of Madagascar (Fig.~\ref{fig:fcc}). This allowed us
-to produce more robust estimates of forest cover and deforestation
-rates over this period (Tab.~\ref{tab:forest_cover}). Our forest cover estimates over the period
-1953-2013 (considering forest cover estimates at national level and by
-ecoregions for all the available dates) were well correlated
-(Pearson's correlation coefficient $=$ 0.99) to estimates from the
-three previous studies (Tab.~\ref{tab:comp_forest}) with a RMSE of
-300,000 ha (6\% of the mean forest cover of 4.8 Mha when considering
-all dates and forest types together). These small differences can be
-partly attributed to differences in ecoregion boundaries. Despite
-significant differences in deforestation estimates
-(Tab.~\ref{tab:comp_defor}), a similar deforestation trend was
-observed across studies with a decrease of deforestation rates over
-the period 1990-2005, followed by a progressive increase of the
-deforestation after 2005.
+territory of Madagascar (Fig.~\ref{fig:fcc}). This allowed us to
+produce more robust estimates of forest cover and deforestation rates
+over this period (Tab.~\ref{tab:forest_cover}). Our forest cover
+estimates over the period 1953-2013 (considering forest cover
+estimates at national level and by ecoregions for all the available
+dates) were well correlated (Pearson's correlation coefficient $=$
+0.99) to estimates from the three previous studies
+(Tab.~\ref{tab:comp_forest}) with a RMSE of 300,000 ha (6\% of the
+mean forest cover of 4.8 Mha when considering all dates and forest
+types together). These small differences can be partly attributed to
+differences in ecoregion boundaries. Despite significant differences
+in deforestation estimates (Tab.~\ref{tab:comp_defor}), a similar
+deforestation trend was observed across studies with a decrease of
+deforestation rates over the period 1990-2005, followed by a
+progressive increase of the deforestation after 2005.
 
@@
 
-Forest fragmentation has progressively increased since 1953 in Madagascar. We observed a
-continuous decrease of the mean distance to forest edge from 1953 to
-2014 in Madagascar. The mean distance to forest edge has decreased to
-about 300~m in 2014 while it was of about 1.5~km in
-1973 (Fig.~\ref{fig:dist_edge}). Moreover, a large proportion (73\%)
-of the forest was located at a distance greater than 100~m in 1973,
-while almost half of the forest (46\%) is at a distance lower than
-100~m from forest edge in 2014 (Fig.~\ref{fig:dist_edge}). The
-percentage of lowly fragmented forest in Madagascar has continuously
-decreased since 1953. The percentage of forest belonging to the lowly
-fragmented class has fallen from 57\% in 1973 to 44\% in 2014. In
-2014, 22\% of the forest belonged to the two highest
-fragmented forest classes (less than 40\% of forest cover in the
-neighborhood) while only 15\% of the forest belonged to these two
+Forest fragmentation has progressively increased since 1953 in
+Madagascar. We observed a continuous decrease of the mean distance to
+forest edge from 1953 to 2014 in Madagascar. The mean distance to
+forest edge has decreased to about 300~m in 2014 while it was of about
+1.5~km in 1973 (Fig.~\ref{fig:dist_edge}). Moreover, a large
+proportion (73\%) of the forest was located at a distance greater than
+100~m in 1973, while almost half of the forest (46\%) is at a distance
+lower than 100~m from forest edge in 2014
+(Fig.~\ref{fig:dist_edge}). The percentage of lowly fragmented forest
+in Madagascar has continuously decreased since 1953. The percentage of
+forest belonging to the lowly fragmented class has fallen from 57\% in
+1973 to 44\% in 2014. In 2014, 22\% of the forest belonged to the two
+highest fragmented forest classes (less than 40\% of forest cover in
+the neighborhood) while only 15\% of the forest belonged to these two
 fragmentation classes in 1973 (Tab.~\ref{tab:frag}).
@@
 (but preferably at the beginning of the period to profit from the full
-record of tree cover loss and derive long-term estimates of deforestation). For example, 
-forest cover maps are available at 20~m resolution for Cameroon and Central 
-African Republic for years 2000 and 2010 \citep{Gross2017}. When high resolution
-forest cover maps are not available, coarser resolution forest cover maps
-(leading to coarser deforestation estimates) could be extracted from global
-land cover products such as GLC2000 at 1~km resolution \citep{Bartholome2005}
-or CCI Land Cover at 300~m resolution \citep{Li2018}.
-Moreover, this approach could be repeated in the future with the release of
-updated tree cover loss data. We have made the \R{}/GRASS code used for this
-study freely available in a GitHub repository (see Data availability
-statement) to facilitate application to other study areas or repeat
-the analysis in the future for Madagascar.
+record of tree cover loss and derive long-term estimates of
+deforestation). For example, forest cover maps are available at 20~m
+resolution for Cameroon and Central African Republic for years 2000
+and 2010 \citep{Gross2017}. When high resolution forest cover maps are
+not available, coarser resolution forest cover maps (leading to
+coarser deforestation estimates) could be extracted from global land
+cover products such as GLC2000 at 1~km resolution
+\citep{Bartholome2005} or CCI Land Cover at 300~m resolution
+\citep{Li2018}.  Moreover, this approach could be repeated in the
+future with the release of updated tree cover loss data. We have made
+the \R{}/GRASS code used for this study freely available in a GitHub
+repository (see Data availability statement) to facilitate application
+to other study areas or repeat the analysis in the future for
+Madagascar.
 
 The accuracy of the derived forest cover change maps depends directly
-on the accuracy of the historical forest cover maps and the tree
-cover loss dataset. Using visual-interpretation of aerial images in
-342 areas distributed among all forest types, \citet{Harper2007}
-estimated an overall 89.5\% accuracy in identifying forest/non-forest
-classes for the year 2000. The accuracy assessment of the tree cover
-loss dataset for the tropical biome reported 13\% of false positives
-and 16.9\% of false negatives (see Tab. S5 in
-\citealt{Hansen2013}). These numbers rise at 20.7\% and 20.6\%
-respectively for the subtropical biome. In the subtropical biome, the
-lower density tree cover canopy makes it difficult to detect change
-from tree cover to bare ground. For six countries in Central Africa,
-with a majority of moist dense forest, \citet{Verhegghen2016} have
-compared deforestation estimates derived from the global tree cover
-loss dataset \citep{Hansen2013} with results derived from
-semi-automated supervised classification of Landsat satellite images
-\citep{Achard2014} and they found a good agreement between the two
-sets of estimates. Therefore, our forest cover change maps after 2000
-might be more accurate for the dense moist forest than for the dry and
-spiny forest. In another study assessing the accuracy of the tree
-cover loss product across the tropics \citep{Tyukavina2015}, authors
-reported 4\% of false positives and 48\% of false negatives in
-Sub-Saharan Africa. They showed that 85\% of missing loss occurred on
-the edges of other loss patches. This means that tree cover loss might
-be underestimated in Sub-Saharan Africa, probably due to the
-prevalence of small-scale disturbance which is hard to map at 30~m,
-but that areas of large-scale deforestation are well identified and
-spatial variability of the deforestation is well represented. A proper
-accuracy assessment of our forest cover change maps should be
-performed to better estimate the uncertainty surrounding our
-forest cover change estimates in Madagascar from year 2000
-\citep{Olofsson2013,Olofsson2014}. Despite this limitation, we have
-shown that the deforestation trend we observed for Madagascar, with a
-doubling deforestation on the period 2010-2014 compared to 2000-2005,
-was consistent with the other studies at the national scale
-\citep{ONE2015, MEFT2009}.
+on the accuracy of the historical forest cover maps and the tree cover
+loss dataset. Using visual-interpretation of aerial images in 342
+areas distributed among all forest types, \citet{Harper2007} estimated
+an overall 89.5\% accuracy in identifying forest/non-forest classes
+for the year 2000. The accuracy assessment of the tree cover loss
+dataset for the tropical biome reported 13\% of false positives and
+16.9\% of false negatives (see Tab. S5 in \citealt{Hansen2013}). These
+numbers rise at 20.7\% and 20.6\% respectively for the subtropical
+biome. In the subtropical biome, the lower density tree cover canopy
+makes it difficult to detect change from tree cover to bare
+ground. For six countries in Central Africa, with a majority of moist
+dense forest, \citet{Verhegghen2016} have compared deforestation
+estimates derived from the global tree cover loss dataset
+\citep{Hansen2013} with results derived from semi-automated supervised
+classification of Landsat satellite images \citep{Achard2014} and they
+found a good agreement between the two sets of estimates. Therefore,
+our forest cover change maps after 2000 might be more accurate for the
+dense moist forest than for the dry and spiny forest. In another study
+assessing the accuracy of the tree cover loss product across the
+tropics \citep{Tyukavina2015}, authors reported 4\% of false positives
+and 48\% of false negatives in Sub-Saharan Africa. They showed that
+85\% of missing loss occurred on the edges of other loss patches. This
+means that tree cover loss might be underestimated in Sub-Saharan
+Africa, probably due to the prevalence of small-scale disturbance
+which is hard to map at 30~m, but that areas of large-scale
+deforestation are well identified and spatial variability of the
+deforestation is well represented. A proper accuracy assessment of our
+forest cover change maps should be performed to better estimate the
+uncertainty surrounding our forest cover change estimates in
+Madagascar from year 2000 \citep{Olofsson2013,Olofsson2014}. Despite
+this limitation, we have shown that the deforestation trend we
+observed for Madagascar, with a doubling deforestation on the period
+2010-2014 compared to 2000-2005, was consistent with the other studies
+at the national scale \citep{ONE2015, MEFT2009}.
 
@@
 (corresponding to 15\% of the country) and that Madagascar has lost
-44\% of its natural forest since 1953 (37\% since 1973). If there are no doubts 
-about the direct causes of deforestation in Madagascar, attributable to human 
-activities such as slash-and-burn agriculture and pasture \citep{Scales2011}, there 
-is ongoing scientific debate about the extent of the ``original'' forest
-cover in Madagascar, and the extent to which humans have altered the
-natural forest landscapes since their large-scale settlement around
-800 CE \citep{Burns2016, Cox2012}. Early French naturalists stated
-that the full island was originally covered by forest
-\citep{Humbert1927, Perrier1921}, leading to the common statement that
-90\% of the natural forests have disappeared since the arrival of
-humans on the island \citep{Kull2000}. More recent studies
-counter-balanced that point of view saying that extensive areas of
-grassland existed in Madagascar long before human arrival and were
-determined by climate, natural grazing and other natural factors
-\citep{Vorontsova2017, Virah-Sawmy2009}. Other authors have questioned
-the entire narrative of extensive alteration of the landscape by early
-human activity which, through legislation, has severe consequences on
-local people \citep{Klein2002, Kull2000}. Whatever the original
-proportion of natural forests and grasslands in Madagascar, our
-results demonstrate that human activities since the 1950s have
-profoundly impacted the natural tropical forests and that conservation
-and development programs in Madagascar have failed to stop
-deforestation in the recent years. Deforestation has strong
-consequences on biodiversity and carbon emissions in
-Madagascar. Around 90\% of Madagascar's species are forest dependent
-\citep{Allnutt2008, Goodman2005}. Based on occurrence data for 2243 plant and
-invertebrate species, \citet{Allnutt2008} estimated
-that deforestation between 1953 and 2000 has led to an extinction of 9\%
-of the species. The additional deforestation we observed over the
-period 2000-2014 (around 1 Mha of natural forest) worsen this
-result. Regarding carbon emissions, using the 2010 aboveground forest
-carbon map by \citet{Vieilledent2016}, we estimated that deforestation
-on the period 2010-2014 has led to 40.2 Mt C of carbon emissions in
-the atmosphere (10 Mt C /yr) and that the remaining aboveground forest
-carbon stock in 2014 is 832.8 Mt C. Associated to deforestation, we
-showed that the remaining forests of Madagascar are highly fragmented
-with 46\% of the forest being at less than 100~m of the forest
-edge. Small forest fragments do not allow to maintain viable
-populations and ``edge effects'' at forest/non-forest interfaces have
-impacts on both carbon emissions \citep{Brinck2017} and biodiversity
-loss \citep{Gibson2013, Murcia1995}.
+44\% of its natural forest since 1953 (37\% since 1973). If there are
+no doubts about the direct causes of deforestation in Madagascar,
+attributable to human activities such as slash-and-burn agriculture
+and pasture \citep{Scales2011}, there is ongoing scientific debate
+about the extent of the ``original'' forest cover in Madagascar, and
+the extent to which humans have altered the natural forest landscapes
+since their large-scale settlement around 800 CE \citep{Burns2016,
+  Cox2012}. Early French naturalists stated that the full island was
+originally covered by forest \citep{Humbert1927, Perrier1921}, leading
+to the common statement that 90\% of the natural forests have
+disappeared since the arrival of humans on the island
+\citep{Kull2000}. More recent studies counter-balanced that point of
+view saying that extensive areas of grassland existed in Madagascar
+long before human arrival and were determined by climate, natural
+grazing and other natural factors \citep{Vorontsova2017,
+  Virah-Sawmy2009}. Other authors have questioned the entire narrative
+of extensive alteration of the landscape by early human activity
+which, through legislation, has severe consequences on local people
+\citep{Klein2002, Kull2000}. Whatever the original proportion of
+natural forests and grasslands in Madagascar, our results demonstrate
+that human activities since the 1950s have profoundly impacted the
+natural tropical forests and that conservation and development
+programs in Madagascar have failed to stop deforestation in the recent
+years. Deforestation has strong consequences on biodiversity and
+carbon emissions in Madagascar. Around 90\% of Madagascar's species
+are forest dependent \citep{Allnutt2008, Goodman2005}. Based on
+occurrence data for 2243 plant and invertebrate species,
+\citet{Allnutt2008} estimated that deforestation between 1953 and 2000
+has led to an extinction of 9\% of the species. The additional
+deforestation we observed over the period 2000-2014 (around 1 Mha of
+natural forest) worsen this result. Regarding carbon emissions, using
+the 2010 aboveground forest carbon map by \citet{Vieilledent2016}, we
+estimated that deforestation on the period 2010-2014 has led to 40.2
+Mt C of carbon emissions in the atmosphere (10 Mt C /yr) and that the
+remaining aboveground forest carbon stock in 2014 is 832.8 Mt
+C. Associated to deforestation, we showed that the remaining forests
+of Madagascar are highly fragmented with 46\% of the forest being at
+less than 100~m of the forest edge. Small forest fragments do not
+allow to maintain viable populations and ``edge effects'' at
+forest/non-forest interfaces have impacts on both carbon emissions
+\citep{Brinck2017} and biodiversity loss \citep{Gibson2013,
+  Murcia1995}.
 
@@
 at 30~m resolution and can be considered more accurate in comparison
-with estimates from these previous studies. Our natural forest cover and
-deforestation rate estimates can be used as source of information for
-the next FAO Forest Resources Assessment \citep{Keenan2015}. Current
-rates of deforestation can also be used to build reference scenarios
-for deforestation in Madagascar and contribute to the implementation
-of deforestation mitigation activities in the framework of REDD+
-\citep{Olander2008}.
+with estimates from these previous studies. Our natural forest cover
+and deforestation rate estimates can be used as source of information
+for the next FAO Forest Resources Assessment
+\citep{Keenan2015}. Current rates of deforestation can also be used to
+build reference scenarios for deforestation in Madagascar and
+contribute to the implementation of deforestation mitigation
+activities in the framework of REDD+ \citep{Olander2008}.
 
@@
 instability and weak governance could also explain the increase in the
-deforestation rate observed on the period 2005-2014
-(see \citealp{Smith2003} for a discussion on the link between governance and
-forest cover loss). These results show that despite the conservation
-policy in Madagascar \citep{Freudenberger2010}, deforestation has
-dramatically increased at the national level since 2005. Results of
-this study, including recent spatially explicit forest cover change
-maps and forest cover estimates, should help implement new
-conservation strategies to save Madagascar natural tropical forests
-and their unique biodiversity.
+deforestation rate observed on the period 2005-2014 (see
+\citealp{Smith2003} for a discussion on the link between governance
+and forest cover loss). These results show that despite the
+conservation policy in Madagascar \citep{Freudenberger2010},
+deforestation has dramatically increased at the national level since
+2005. Results of this study, including recent spatially explicit
+forest cover change maps and forest cover estimates, should help
+implement new conservation strategies to save Madagascar natural
+tropical forests and their unique biodiversity.
 
@@
 
-All authors conceived the ideas and designed the methodology; GV analysed
-the data and wrote the {\R}/GRASS script; GV drafted the manuscript. All
-authors contributed critically to the drafts and gave final approval for
-publication.
+All authors conceived the ideas and designed the methodology; GV
+analysed the data and wrote the {\R}/GRASS script; GV drafted the
+manuscript. All authors contributed critically to the drafts and gave
+final approval for publication.
 
@@
 
-The authors thank two anonymous reviewers for useful comments on a previous
-version of the manuscript. They also thank Peter Vogt for useful advice on which
-metric to use to estimate forest fragmentation. This study is part of
-the Cirad's BioSceneMada project (\url{https://bioscenemada.cirad.fr})
-and the Joint Research Center's ReCaREDD project
-(\url{http://forobs.jrc.ec.europa.eu/recaredd}). The BioSceneMada
-project is funded by FRB (Fondation pour la Recherche sur la
-Biodiversité) and the FFEM (Fond Français pour l'Environnement
+The authors thank two anonymous reviewers for useful comments on a
+previous version of the manuscript. They also thank Peter Vogt for
+useful advice on which metric to use to estimate forest
+fragmentation. This study is part of the Cirad's BioSceneMada project
+(\url{https://bioscenemada.cirad.fr}) and the Joint Research Center's
+ReCaREDD project (\url{http://forobs.jrc.ec.europa.eu/recaredd}). The
+BioSceneMada project is funded by FRB (Fondation pour la Recherche sur
+la Biodiversité) and the FFEM (Fond Français pour l'Environnement
 Mondial) under the project agreement AAP-SCEN-2013~I. The ReCaREDD
-project is funded by the European Commission. The authors declare that there are no conflicts of interest related to this article.
+project is funded by the European Commission. The authors declare that
+there are no conflicts of interest related to this article.
 
@@
 
-  \caption{\textbf{Selection of the tree cover threshold to define the land type for unclassified areas in 2000.} We considered the forest in 2000 \citep{Harper2007} in the moist ecoregion (see Fig.~\ref{fig:ecoregion}).
-We plotted the percentage of forest having a tree cover greater or equal to a given value specified by the x-axis.
-From this figure, we see that 90\% of the moist forest in 2000 has a tree cover $\geq$ 75.}
+  \caption{\textbf{Selection of the tree cover threshold to define the
+      land type for unclassified areas in 2000.} We considered the
+    forest in 2000 \citep{Harper2007} in the moist ecoregion (see
+    Fig.~\ref{fig:ecoregion}). We plotted the percentage of forest
+    having a tree cover greater or equal to a given value specified by
+    the x-axis. From this figure, we see that 90\% of the moist forest
+    in 2000 has a tree cover $\geq$ 75.}
 

### 2018-03-19 manuscript/manuscript.Rnw
Link to Cirad Dataverse

@@
     $[1]$ \textbf{Joint Research Center of the European Commission} -- Bio-economy Unit, I-21027 Ispra (VA), ITALY\\
-    $[2]$ \textbf{Cirad} -- UPR Forêts et Sociétés, F-34398 Montpellier, FRANCE\\
+    $[2]$ \textbf{CIRAD} -- UPR Forêts et Sociétés, F-34398 Montpellier, FRANCE\\
     $[3]$ \textbf{ETC Terra}, F-75020 Paris, FRANCE\\
@@
 prevents us from underestimating forest cover and overestimating
-forest fragmentation. All the resulting maps have been made
-permanently and freely available on the Zenodo research data
-repository at \url{https://doi.org/10.5281/zenodo.1145785}.
+forest fragmentation.
 
@@
 useful advice on which metric to use to estimate forest
-fragmentation. This study is part of the Cirad's BioSceneMada project
+fragmentation. This study is part of the CIRAD's BioSceneMada project
 (\url{https://bioscenemada.cirad.fr}) and the Joint Research Center's
@@
 
-All the data and the script used for this study have been made permanently and publicly available
-on the Zenodo research data repository so that the results are entirely reproducible:
+Data and code used for this study have been made permanently and publicly available
+on the CIRAD Dataverse repository so that the results are entirely reproducible:
 \begin{itemize}
-\item Input data: \url{https://doi.org/10.5281/zenodo.1118955}
-\item Script: \url{https://doi.org/10.5281/zenodo.1118484}
-\item Output data: \url{https://doi.org/10.5281/zenodo.1145785}
+\item Input data: \url{http://dx.doi.org/10.18167/DVN1/2FP7LR}
+\item Code: \url{http://dx.doi.org/10.18167/DVN1/275TDF}
+\item Output data: \url{http://dx.doi.org/10.18167/DVN1/AUBRRC}
 \end{itemize}

### 2018-03-19 deforestmap.R

@@ -19,7 +19,7 @@
 
 ##= Libraries
 pkg <- c("broom","sp","rgdal","raster","ggplot2","gridExtra","RColorBrewer",
-         "rasterVis","knitr","rmarkdown","rgeos","rgdal","rgrass7","xtable")
+         "rasterVis","knitr","rmarkdown","kableExtra","rgeos","rgdal","rgrass7")
 ## broom: to convert map into data-frame with tiny()
 ## gridExtra: to combine several ggplots
 ## rasterVis: for gplot()
@@ -1093,8 +1093,8 @@ opts_knit$set(root.dir="manuscript")
 knitr::knit2pdf("manuscript/manuscript.Rnw", output="manuscript/manuscript.tex")
 
 ## Cover letter
-rmarkdown::render("manuscript/coverletter.md", output_format=c("pdf_document"),
-                  output_dir="manuscript") # pdf output
+rmarkdown::render("manuscript/coverletter/coverletter3.md", output_format=c("pdf_document"),
+                  output_dir="manuscript/coverletter") # pdf output
 
 ##===========================================================================
 ## End of script

### 2018-03-20 deforestmap.R
Link to Cirad Dataverse

@@ -19,6 +19,7 @@
 
 ##= Libraries
 pkg <- c("broom","sp","rgdal","raster","ggplot2","gridExtra","RColorBrewer",
+         "dataverse",
          "rasterVis","knitr","rmarkdown","kableExtra","rgeos","rgdal","rgrass7")
 ## broom: to convert map into data-frame with tiny()
 ## gridExtra: to combine several ggplots
@@ -58,12 +59,16 @@ initGRASS(gisBase="/usr/lib/grass72",home=tempdir(),
           override=TRUE)
 
 ##==================================================================
-## Download data from Zenodo: https://doi.org/10.5281/zenodo.1118955
+## Download data from Cirad Dataverse: http://dx.doi.org/10.18167/DVN1/2FP7LR
 down <- FALSE
 if (down) {
-  d <- "https://zenodo.org/record/1118956/files/gisdata.zip"
-  download.file(url=d,destfile="gisdata.zip",method="curl",quiet=FALSE)
-  unzip("gisdata.zip", exdir="test")
+  library(dataverse)
+  Sys.setenv("DATAVERSE_SERVER"="dataverse.cirad.fr")
+  dataverse::get_dataset("doi:10.18167/DVN1/2FP7LR")
+  f <- dataverse::get_file(file="gisdata.zip",
+                           dataset="doi:10.18167/DVN1/2FP7LR")
+  writeBin(f, "gisdata.zip")
+  unzip("gisdata.zip", exdir="gisdata")
 }
 
 ##===========================================================

### 2018-10-11 R/plotfcc.R
Update to 2017

@@ -7,12 +7,12 @@
 # license         :GPLv3
 # ==============================================================================
 
-library(ggplot2)
-library(rasterVis)
-library(gridExtra)
-library(grid)
-library(rgdal)
-library(ggspatial)  # for geom_spatial()
+require(ggplot2)
+require(rasterVis)
+require(gridExtra)
+require(grid)
+require(rgdal)
+require(ggspatial)  # for geom_spatial()
 
 # Zooms
 zoom.w <- list(xmin=346000,xmax=439000,ymin=7387000,ymax=7480000)
@@ -22,7 +22,7 @@ zooms <- list(zoom.w,zoom.e)
 # Compute zooms
 for (z in 1:length(zooms)) {
   ExtentZ <- paste(zooms[[z]]$xmin,zooms[[z]]$ymin,zooms[[z]]$xmax,zooms[[z]]$ymax)
-  Input <- c("fcc.tif","fcc_for1953.tif","fordens2014_class_mask.tif","dist_edge_2014_mask.tif")
+  Input <- c("fcc.tif","fcc_for1953.tif","fordens2017_class_mask.tif","dist_edge_2017_mask.tif")
   nodata <- c(rep(255,3),9999)
   Output <- paste0(sub(".tif","",Input),"_zoom",z,".tif")
   for (i in 1:length(Input)) {
@@ -45,9 +45,10 @@ get_legend <- function(myggplot) {
 
 # ======
 # Colors
-for2014.col = rgb(34/255,139/255,34/255) # forest green
-defor00_14.col = rgb(255/255,0/255,0/255) # red
-defor90_00.col = rgb(255/255,165/255,0/255) # orange
+for2017.col = rgb(34/255,139/255,34/255) # forest green
+defor10_17.col = rgb(240/255,59/255,32/255) # red
+defor00_10.col = rgb(254/255,178/255,76/255) # orange
+defor90_00.col = rgb(255/255,237/255,160/255) # light orange
 defor73_90.col = rgb(100/255,100/255,100/255) # grey
 nofor73.col = rgb(243/255,243/255,220/255) # light yellow
 w1.col = rgb(153/255,217/255,234/255) # light blue
@@ -58,12 +59,12 @@ water.col <-colorRampPalette(c(w1.col,w2.col))(12)
 # Palette
 # fcc (discrete)
 fcc.col <- c(nofor73.col,water.col,nofor73.col,defor73_90.col,nofor73.col,
[…]

### 2018-10-11 deforestmap.R

@@ -58,6 +58,10 @@ initGRASS(gisBase="/usr/lib/grass72",home=tempdir(),
           location="deforestmap",mapset="PERMANENT",
           override=TRUE)
 
+# TEMP!
+execute <- FALSE
+if (execute) {
+
 ##==================================================================
 ## Download data from Cirad Dataverse: http://dx.doi.org/10.18167/DVN1/2FP7LR
 down <- FALSE
@@ -632,7 +636,7 @@ defor.df$clouds[defor.df$Year==1973] <- cloud.1973
 write.table(defor.df,"outputs/defor.txt",row.names=FALSE,sep="\t")
 
 #= Deforestation statistics for comparison
-Year <- c(1953,1973,1990,2000,2005,2010,2013,2015,2017)  # Here we add 2013
+Year <- c(1953,1973,1990,2000,2005,2010,2013)  # Here we add 2013
 defor.df <- data.frame(Year=Year,area=NA,ann.defor=NA,theta=NA)
 # Areas
 for (i in 1:length(Year)) {
@@ -673,6 +677,7 @@ write.table(fordens.df,"outputs/fordens.txt",row.names=FALSE,sep="\t")
 Year <- c(1953,1973,1990,2000,2005,2010,2015,2017)
 frag.df <- data.frame(Year=Year,forest=NA,patch=NA,transitional=NA,
                       edge=NA,perforated=NA,interior=NA,undetermined=NA)
+Res <- "30"
 # Areas
 for (i in 1:length(Year)) {
   # Message
@@ -785,7 +790,6 @@ for (i in 1:length(Year)) {
 }
 write.table(perc.100m.dist.df,"outputs/perc_100m_dist.txt",row.names=FALSE,sep="\t")
 
-
 ##========================
 ## Statistics by ecoregion
 ##========================
@@ -793,14 +797,14 @@ write.table(perc.100m.dist.df,"outputs/perc_100m_dist.txt",row.names=FALSE,sep="
 # Ecoregions
 # 1: spiny, 2: mangroves, 3: moist, 4: dry
 ecoregion <- c("spiny","mangroves","moist","dry")
+Res <- 30
 
 # Loop on ecoregions
-#for (j in 1:length(ecoregion)) {
-for (j in 1:3) {
+for (j in 1:length(ecoregion)) {
   # Message
   cat(paste0("Ecoregion: ",ecoregion[j]))
   # Mask
-  system(paste0("r.mask --o raster=ecoregion maskcats=",j,"'"))
[…]

### 2018-10-11 update_2017/update_2017.Rnw
Update 2017

--- 2018-03-19 manuscript/manuscript.Rnw
@@
 \usepackage[colorlinks=true,citecolor=Blue,urlcolor=Maroon]{hyperref} %% Gère les hyperliens
-\usepackage{lineno} %% Numérotation des lignes
+%\usepackage{lineno} %% Numérotation des lignes
 %\linenumbers
@@
 
-\title{Combining global tree cover loss data with historical national
+\title{Update to: Combining global tree cover loss data with historical national
   forest cover maps to look at six decades of deforestation and forest
-  fragmentation in Madagascar}
+  fragmentation in Madagascar$^{\S}$}
 \author{}
@@
 \begin{center}
-  Ghislain Vieilledent$^{1,2,\star}$, Clovis Grinand$^{3}$,
-  Fety A. Rakotomalala$^{3}$,\\
-  Rija Ranaivosoa$^{4}$, Jean-Roger Rakotoarijaona$^{4}$,\\
-  Thomas F. Allnutt$^{5,6}$, and Frédéric Achard$^{1}$\\
+  Ghislain Vieilledent$^{1,\star}$, Clovis Grinand$^{2}$,
+  Fety A. Rakotomalala$^{2}$,\\
+  Rija Ranaivosoa$^{3}$, Jean-Roger Rakotoarijaona$^{3}$,\\
+  Thomas F. Allnutt$^{4,5}$, and Frédéric Achard$^{6}$\\
 \end{center}
@@
 {\small
+\noindent$^{\S}$ This article was initially published as:
+
+\noindent\textbf{Vieilledent G., C.
+Grinand, F.
+A.
+Rakotomalala, R.
+Ranaivosoa, J.-R.
+Rakotoarijaona, T.
+F.
+Allnutt, and F.
+Achard.} 2018.
+Combining global tree cover loss data with historical national forest-cover maps to look at six decades of deforestation and forest fragmentation in Madagascar.
+\emph{Biological Conservation.} \textbf{222}: 189-197.
+[doi: \href{https://doi.org/10.1016/j.biocon.2018.04.008}{10.1016/j.biocon.2018.04.008}].}
+
+\vspace{0.3cm}
+
+{\small
   \begin{flushleft}  
-    $[1]$ \textbf{Joint Research Center of the European Commission} -- Bio-economy Unit, I-21027 Ispra (VA), ITALY\\
-    $[2]$ \textbf{CIRAD} -- UPR Forêts et Sociétés, F-34398 Montpellier, FRANCE\\
-    $[3]$ \textbf{ETC Terra}, F-75020 Paris, FRANCE\\
-    $[4]$ \textbf{Office National pour l'Environnement}, 101 Antananarivo, MADAGASCAR\\
-    $[5]$ \textbf{Wildlife Conservation Society}, 101 Antananarivo, MADAGASCAR\\
-    $[6]$ \textbf{GreenInfo Network}, Oakland, California, USA\\
+    $[1]$ \textbf{CIRAD} -- UPR Forêts et Sociétés, F-34398 Montpellier, FRANCE\\
+    $[2]$ \textbf{Nitid\ae}, F-75020 Paris, FRANCE\\
+    $[3]$ \textbf{Office National pour l'Environnement}, 101 Antananarivo, MADAGASCAR\\
+    $[4]$ \textbf{Wildlife Conservation Society}, 101 Antananarivo, MADAGASCAR\\
+    $[5]$ \textbf{GreenInfo Network}, Oakland, California, USA\\
+    $[6]$ \textbf{Joint Research Center of the European Commission} -- Bio-economy Unit, I-21027 Ispra (VA), ITALY\\
     ~\\
@@
     \textbackslash{E-mail}:~ghislain.vieilledent@cirad.fr
-    \textbackslash{Phone}:~+39.033.278.3516\\
   \end{flushleft}}
@@
 \doublespacing
-\linenumbers
+%\linenumbers
 
@@
 
-  The island of Madagascar has a unique biodiversity, mainly located
-  in the tropical forests of the island. This biodiversity is
-  highly threatened by anthropogenic deforestation. Existing
-  historical forest maps at national level are scattered and have
-  substantial gaps which prevent an exhaustive assessment of long-term
-  deforestation trends in Madagascar. In this study, we combined
-  historical national forest cover maps (covering the period
-  1953-2000) with a recent global annual tree cover loss dataset
-  (2001-2014) to look at six decades of deforestation and forest
-  fragmentation in Madagascar (from 1953 to 2014). We produced new
-  forest cover maps at 30~m resolution for the year 1990 and annually
-  from 2000 to 2014 over the full territory of Madagascar. We
-  estimated that Madagascar has lost 44\% of its natural forest cover
-  over the period 1953-2014 (including 37\% over the period
-  1973-2014). Natural forests cover 8.9 Mha in 2014 (15\% of the
-  national territory) and include 4.4 Mha (50\%) of moist forests, 2.6
-  Mha (29\%) of dry forests, 1.7 Mha of spiny forests (19\%) and
-  177,000 ha (2\%) of mangroves. Since 2005, the annual deforestation
-  rate has progressively increased in Madagascar to reach 99,000 ha/yr
-  during 2010-2014 (corresponding to a rate of 1.1\%/yr). Around half of
-  the forest (46\%) is now located at less than 100~m from the forest
-  edge. Our approach could be replicated to other developing countries with tropical forest.
-  Accurate forest cover change maps can be used to assess the
-  effectiveness of past and current conservation programs and
-  implement new strategies for the future. In particular, forest maps
-  and estimates can be used in the REDD+ framework which aims at
-  ``Reducing Emissions from Deforestation and forest Degradation'' and
-  for optimizing the current protected area network.
+Blabla
 
@@
 
-Separated from the African continent and the Indian plate about 165
-and 88 million years ago respectively \citep{Ali2008}, the flora and
-fauna of Madagascar followed its own evolutionary path. Isolation
-combined with a high number of micro-habitats \citep{Pearson2009} has
-led to Madagascar's exceptional biodiversity both in term of number of
-species and endemism in many taxonomic groups \citep{Crottini2012,
-  Goodman2005}. Most of the biodiversity in Madagascar is concentrated
-in the tropical forests of the island which can be divided into four
-types: the moist forest in the East, the dry forest in the West, the
-spiny forest in the South and the mangroves on the West coast
-\citep{Vieilledent2016}. This unparalleled biodiversity is severely
-threatened by deforestation \citep{Harper2007, Vieilledent2013}
-associated with human activities such as slash-and-burn agriculture
-and pasture \citep{Scales2011}. Tropical forests in Madagascar also
-store a large amount of carbon (136 MgC.ha$^{-1}$ in the moist forest,
-\citealp{Vieilledent2016}) and high rates of deforestation in
-Madagascar (1.4--4.7~\%/yr, \citealp{Achard2002}) are responsible for
-large CO$_2$ emissions in the atmosphere. Deforestation threatens
-species survival by directly reducing their available habitat
-\citep{Brooks2002, Tidd2001}. Forest fragmentation can also lead to
-species extinction by isolating populations from each other and
-creating forest patches too small to maintain viable populations
-\citep{Saunders1991}. Fragmentation also increases forest edge where
-ecological conditions (such as air temperature, light intensity and
-air moisture) can be dramatically modified, with consequences on the
-abundance and distribution of species \citep{Murcia1995,
-  Broadbent2008, Gibson2013}. Forest fragmentation can also have
-substantial effects on forest carbon storage capacity, as carbon
-stocks are about 50\% lower at the forest edge than under a closed
-canopy \citep{Brinck2017}. Moreover, forest carbon stocks vary
-spatially due to climate or soil factors \citep{Saatchi2011,
-  Vieilledent2016}. As a consequence, accurate and spatially explicit
-maps of forest cover and forest cover change are necessary to monitor
-biodiversity loss and carbon emissions from deforestation and forest
-fragmentation, assess the efficiency of present conservation
-strategies \citep{Eklund2016}, and implement new strategies for the
-future \citep{Vieilledent2013, Vieilledent2016}. Simple time-series of
-forest cover estimates, such as those provided by the FAO Forest
-Resource Assessment report \citep{Keenan2015} are not sufficient.
-
-Unfortunately, accurate and exhaustive forest cover maps are not
-available for Madagascar after year 2000. \citet{Harper2007} produced
-maps of forest cover and forest cover changes over Madagascar for the
-years 1953, 1973, 1990 and 2000. The 1953 forest map is a vector map
-derived from the visual interpretation of aerial photographs. Forest
-maps for the years 1973, 1990, and 2000 were obtained from the
-supervised classification of Landsat satellite images and can be used
-to derive more accurate estimates of forest cover than those from the
-FAO Forest Resource Assessment report. Nonetheless, maps provided by
-\citet{Harper2007} are not exhaustive (due to the presence of clouds
-in the satellite imagery), e.g.~11 244~km2 are mapped as unknown cover
-type for the year 2000. Using a similar supervised classification
-approach as in \citet{Harper2007}, more recent maps have been produced
-for the periods 2000-2005-2010 by national institutions, with the
-technical support of international environmental NGOs \citep{MEFT2009,
-  ONE2013}. Another set of recent forest cover maps using an advanced
-statistical tool for classification, the Random Forest classifier
-\citep{Grinand2013, Rakotomalala2015}, was produced for the periods
-2005-2010-2013 \citep{ONE2015}. However, these maps are either too old
-to give recent estimates of deforestation \citep{MEFT2009, ONE2013},
-include large areas of missing information due to images with high
-percentage of cloud cover \citep{ONE2013}, or show large
-mis-classification in specific areas, especially in the dry and spiny
-forest domain, for which the spectral signal shows strong seasonal
-variations due to the deciduousness of such forests (overall accuracy
-is lower than 0.8 for the dry and spiny forests for the maps produced
-by \citet{ONE2015}). Moreover, the production of such forest maps from
-a supervised classification approach requires significant resources,
-especially regarding the image selection step (required to minimize
-cloud cover) and the training step (visual interpretation of a large
-number of polygons needed to train the classification algorithm)
-\citep{Rakotomalala2015}. Most of this work of image selection and
-visual interpretation would need to be repeated to produce new forest
-maps in the future using a similar approach.
-
-Global forest or tree cover products have also been published recently
-and can be tested at the national scale for
-Madagascar. \citet{Kim2014} produced a global forest cover change map
-from 1990 to 2000 (derived from Landsat imagery). This product was
-updated to cover the period 1975-2005
-(\url{http://glcf.umd.edu/data/landsatFCC/}) but forest cover maps
-after 2005 were not produced. Moreover, the approach used in
-\citet{Kim2014} did not accurately map the forests in the dry and
-spiny ecosystems of Madagascar (see Fig. 8 in \citet{Kim2014}).
-\citet{Hansen2013} mapped tree cover percentage, annual tree cover
-loss and gain from 2000 to 2012 at global scale at 30 m
-resolution. This product has since been updated and is now available
-up to the year 2014 \citep{Hansen2013}. To map forest cover from the
-\citet{Hansen2013} product, a tree cover threshold must be selected
-(that defines forest cover). Selecting such a threshold is not
-straightforward as the accuracy of the global tree cover map strongly
-varies between forest types, and is substantially lower for dry
-forests than for moist forests \citep{Bastin2017}. Moreover, the
-\citet{Hansen2013} product does not provide information on
-land-use. In particular the global tree cover map does not separate
-tree plantations such as oil palm or eucalyptus plantations from
-natural forests \citep{Tropek2014}. Thus, the global tree cover map
-from \citet{Hansen2013} cannot be used alone to produce a map of
-forest cover \citep{Tyukavina2017}.
-
-In this study, we present a simple approach which combines the
-historical forest maps from \citet{Harper2007} and more recent global
-products from \citet{Hansen2013} to derive annual wall-to-wall forest
-cover change maps over the period 2000-2014 for Madagascar. We use the
-forest cover map provided by \citet{Harper2007} for the year 2000
-(defining the land-use) with the tree cover loss product provided by
-\citet{Hansen2013} that we apply only inside forest areas identified
-by \citet{Harper2007}. Similar to the approach of \citet{Harper2007},
-we also assess trends in deforestation rates and forest fragmentation
-from 1953 to 2014. We finally discuss the possibility to extend our
-approach to other tropical countries or repeat it in the future for
-Madagascar. We also discuss how our results could help assess the
-effectiveness of past and current conservation strategies in
-Madagascar, and implement new strategies in the future.
-
-\newpage
-
-\section{Materials and Methods}
-\label{materials-and-methods}
-
-\subsection{Creation of new forest cover maps of Madagascar from
-1953 to 2014}
-
-Original 1990-2000 forest cover change map for Madagascar from
-\citet{Harper2007} is a raster map at 28.5~m resolution. It was
-derived from the supervised classification of Landsat TM (Thematic
-Mapper) and ETM+ (Enhanced Thematic Mapper Plus) satellite images. For
-our study, this map has been resampled at 30~m resolution using a
-nearest-neighbor interpolation and reprojected in the WGS 84/UTM zone
-38S projected coordinate system.
-
-The 2000 Harper's forest map includes 208,000 ha of unclassified areas
-due to the presence of clouds on satellite images. Unclassified areas
-were mostly (88\%) present within the moist forest domain which
-covered 4.17 Mha in 2000. To provide a label (forest or non-forest) to
-these unclassified pixels, we used the 2000 tree cover percentage map
-of \citet{Hansen2013} and selected a tree cover threshold of 75\% to
-define the forest \citep{Achard2014, Aleman2017}. This threshold
-allows to characterize properly the moist forest in Madagascar as 90\%
-of the moist forest in 2000 in \citet{Harper2007} has a tree cover
-greater than 75\% (Fig.~\ref{fig:tc_threshold}). For this step, the
-Hansen's 2000 tree cover map was resampled on the same grid as the
-original Harper's map at 30~m resolution using a bilinear
-interpolation. We thus obtained a forest cover map for the year 2000
-covering the full territory of Madagascar.
-
-We then combined the forest cover map of the year 2000 with the annual
-tree cover loss maps from 2001 to 2014 from \citet{Hansen2013} to
-create annual forest cover maps from 2001 to 2014 at 30~m
-resolution. To do so, Hansen's tree cover loss maps were resampled on
-the same grid as the original Harper's map at 30~m resolution using a
-nearest-neighbor interpolation. We also completed the Harper's forest
-map of year 1990 by filling unclassified areas (due to the presence of
-clouds on satellite images) using our forest cover map of year
-2000. To do so, we assumed that if forest was present in 2000, the
-pixel was also forested in 1990. Indeed, there is little evidence of
-natural forest regeneration in Madagascar \citep{Grouzis2001,
-  Harper2007}, especially over such a short period of time.  The
-remaining unclassified pixels were limited to a relatively small total
-area of about 8,000 ha. We labeled these residual pixels as
-non-forest, as for the year 2000.
-
-The 1973 forest cover map for Madagascar from \citet{Harper2007} is a
-raster map at 57~m resolution derived from the supervised
-classification of Landsat MSS (Multispectral Scanner System) satellite
-images. We resampled this map at 30~m resolution using a
-nearest-neighbor interpolation on the same grid as the forest cover
-maps for years 1990 and 2000. We completed the Harper's forest map of
-year 1973 by filling unclassified areas using our forest cover map of
-the year 1990 assuming that if forest was present in 1990, it was also
-present in 1973. Contrary to the year 1990, the remaining unclassified
-pixels for year 1973 corresponded to a significant total area of 3.3
-millions ha which was left as is.
-
-The 1953 forest cover map from \citet{Harper2007} is a vector map
-produced by scanning the paper map of \citet{Humbert1965} which was
-derived from the visual interpretation of aerial photographs. We
-reprojected the forest cover map of year 1953 in the WGS 84/UTM zone
-38S projected coordinate system. Because of the methodology used to
-derive the 1953 forest cover map, it was not possible to perfectly
-aligned this map with the forest cover maps of later years which were
-produced through digital processing of satellite imagery. As a
-consequence, the 1953 cannot be merged with the map of later years to
-identify precisely the location of the deforested areas. Nonetheless,
-the 1953 forest cover map can be used to have a rough estimate of the
-forest cover and forest fragmentation at this date.  To do so, the map
-was rasterized at 30~m resolution on the same grid as the forest cover
-maps for years 1973, 1990 and 2000.
-
-Finally for all forest cover maps from 1973, isolated single
-non-forest pixels (i.e.~fully surrounded by forest pixels) were
-recategorized as forest pixels. Doing so, forest cover increased from
-about 95,000~ha for year 1953 to about 600,000~ha for year 2010. This
-allowed us to avoid counting very small scale events (\textless 0.1
-ha, such as selective logging or wind-throw) as deforestation. It also
-prevents us from underestimating forest cover and overestimating
-forest fragmentation.
-
-\subsection{Computing forest cover areas and deforestation rates}
-
-From these new forest cover maps, we calculated the total forest cover
-area for seven available years (1953-1973-1990-2000-2005-2010-2014),
-and the annual deforested area and annual deforestation rate for the
-corresponding six time periods between 1953 and 2014. The annual
-deforestation rates were calculated using Eq.~\ref{eq:theta}
-\citep{Puyravaud2003, Vieilledent2013}:
-
-\begin{equation}
-  \label{eq:theta}
-  \theta = 100 \times [1-(1-(F_{t_2}-F_{t_1})/F_{t_1})^{(1/(t_2-t_1))}]
-\end{equation}
-
-In Eq.~\ref{eq:theta}, $\theta$ is the annual deforestation rate (in \%/yr),
-$F_{t_2}$ and $F_{t_1}$ are the forest cover free of clouds at both
-dates $t_2$ and $t_1$, and $t_2-t_1$ is the time-interval (in
-years) between the two dates.
-
-Because of the large unclassified area (3.3 millions ha) in 1973, the
-annual deforestation areas and rates for the two periods 1953-1973 and
-1973-1990 are only partial estimates computed on the basis of the
-available forest extent. Area and rate estimates are produced at the
-national scale and for the four forest types present in
-Madagascar: moist forest in the East, dry forest in the West, spiny
-forest in the South, and mangroves on the Western coast
-(Fig.~\ref{fig:ecoregion}). To define the forest types, we used a
-map from the MEFT (\emph{``Ministère de l'Environnement et des Forêts
-  à Madagascar''}) with the boundaries of the four ecoregions in
-Madagascar. Ecoregions were defined on the basis of climatic and
-vegetation criteria using the climate classification by
-\citet{Cornet1974} and the vegetation classification from the 1996
-IEFN national forest inventory \citep{IEFN1996}. Because mangrove
-forests are highly dynamic ecosystems that can expand or contract on
-decadal scales depending on changes in environmental factors
-\citep{Armitage2015}, a fixed delimitation of the mangrove ecoregion
-on six decades might not be fully appropriate. As a consequence, our
-estimates of the forest cover and deforestation rates for mangroves in
-Madagascar must be considered with this limitation.
-
-\subsection{Comparing our forest cover and deforestation rate
-estimates with previous studies}
-
-We compared our estimates of forest cover and deforestation rates with
-estimates from the three existing studies at the national scale for
-Madagascar: (i) \citet{Harper2007}, (ii) \citet{MEFT2009} and (iii)
-\citet{ONE2015}. \citet{Harper2007} provides forest cover and
-deforestation estimates for the periods
-c. 1953-c. 1973-1990-2000. \citet{MEFT2009} provides estimates for the
-periods 1990-2000-2005 and \citet{ONE2015} provides estimates for the
-periods 2005-2010-2013. To compare our forest cover and deforestation
-estimates over the same time periods, we consider an additional
-time-period in our study (2010-2013) by creating an extra forest cover
-map for the year 2013. We computed the Pearson's correlation
-coefficient and the root mean square error (RMSE) between our
-forest cover estimates and forest cover estimates from previous
-studies for all the dates and forest types (including also the total
-forest cover estimates). For previous studies, the computation of
-annual deforestation rates (in \%/yr) is not always detailed and might
-slightly differ from one study to another
-\citep[see][]{Puyravaud2003}. \citet{Harper2007} also provide total
-deforested areas for the two periods 1973-1990 and 1990-2000. We
-converted these values into annual deforested area estimates. When
-annual deforested areas were not reported (for 1953-1973 in
-\citet{Harper2007} and in \citet{MEFT2009} and \citet{ONE2015}), we
-computed them from the forest cover estimates in each study. These
-estimates cannot be corrected from the potential bias due to the
-presence of residual clouds. Forest cover and deforestation rates were
-then compared between all studies for the whole of Madagascar and the
-four ecoregions. The same ecoregion boundaries as in our study were
-used in \citet{ONE2015} but this was not the case for
-\citet{Harper2007} and \citet{MEFT2009}, which can explain a part of
-the differences between the estimates.
-
-\subsection{Fragmentation}
-
-We also conducted an analysis of changes in forest fragmentation for
-the years 1953, 1973, 1990, 2000, 2005, 2010 and 2014 at 30~m
-resolution. We used a moving window of $51 \times 51$ pixels
-(corresponding to an area of about 2.34~km$^2$) centered on each
-forest pixel to compute the percentage of forest pixels in the
-neighborhood. We used this percentage as an indication of the forest
-fragmentation \citep{Riitters2012, Vogt2017}. The size of the moving
-windows was based on a compromise: a sufficiently high number of cells
-(here 2601) had to be considered to be able to compute a percentage
-and a reasonably low number of cells had to be chosen to have a local
-estimate of the fragmentation. Water bodies were not masked when
-computing the percentage of forest pixels, meaning that forest located
-near a water body was considered as fragmented. Computations were done
-using the function \texttt{r.neighbors} of the GRASS GIS software
-\citep{Neteler2008}. Using the density of forest in the neighborhood,
-we defined five forest fragmentation classes: 0-20\% (highly
-fragmented), 21-40\%, 41-60\%, 61-80\% and 81-100\% (lowly
-fragmented). We reported the percentage of forest falling in each
-fragmentation class for the six years and analyzed the dynamics of
-fragmentation over the six decades.
-
-We also computed the distance to forest edge for all forest
-pixels for the years 1953, 1973, 1990, 2000, 2005, 2010 and 2014. For
-that, we used the function \texttt{gdal\_proximity.py} of the GDAL
-library (\url{http://www.gdal.org/}). We computed the mean and 90\%
-quantiles (5\% and 95\%) of the distance to forest edge and looked at
-the variation of these values over time. Previous studies have shown
-that forest micro-habitats were mainly altered within the first 100~m
-of the forest edge \citep{Brinck2017, Murcia1995,
-  Broadbent2008}. Consequently, we also estimated the percentage of
-forest within the first 100~m of the forest edge for each year and
-looked at the variation of this percentage over the six decades.
-
-\newpage
-
-\section{Results}
-\label{results}
-
-\subsection{Forest cover change and deforestation rates}
-
-Natural forests in Madagascar covered 16.0 Mha in 1953, about 27\% of
-the national territory of 587,041 km2. In 2014, the forest cover
-dropped to 8.9 Mha, corresponding to about 15\% of the national
-territory (Fig.~\ref{fig:fcc} and
-Tab.~\ref{tab:forest_cover}). Madagascar has lost 44\% of its natural
-forest between 1953 and 2014, including 37\% between 1973 and 2014
-(Fig.~\ref{fig:fcc} and Tab.~\ref{tab:forest_cover}). In 2014 the
-remaining 8.9 Mha of natural forest were distributed as follow: 4.4
-Mha of moist forest (50\% of total forest cover), 2.6 Mha of dry
-forest (29\%), 1.7 Mha of spiny forest (19\%) and 0.18 Mha (2\%) of
-mangrove forest (Fig.~\ref{fig:ecoregion} and
-Tab.~\ref{tab:comp_forest}).
-
-The forest cover change map produced on the period 1953-2014
-(Fig.~\ref{fig:fcc}) allows to identify hot-spots of
-deforestation. Among the many recent hot-spots of deforestation
-visible on the map for the period 2000-2014 , one is located at the
-south of the CAZ (\emph{``Corridor Ankeniheny Zahamena''}) protected
-area, in the moist forest at the east of Madagascar (see eastern zoom
-in Fig.~\ref{fig:fcc}). Another major hot-spot of deforestation is
-located around the Ranobe-PK32 new protected area, in the dry forest
-at the south-west of Madagascar (see western zoom in
-Fig.~\ref{fig:fcc}).
-
-Regarding the deforestation trend, we observed a progressive decrease
-of the deforestation rate after 1990 from 205,000 ha/yr (1.6\%/yr)
-over the period 1973-1990 to 42,000 ha/yr (0.4\%/yr) over the period
-2000-2005 (Tab.~\ref{tab:forest_cover}). Then from 2005, the
-deforestation rate has progressively increased and has more than
-doubled over the period 2010-2014 (99,000 ha/yr, 1.1\%/yr) compared to
-2000-2005 (Tab.~\ref{tab:forest_cover}). The deforestation trend,
-characterized by a progressive decrease of the deforestation rate over
-the period 1990-2005 and a progressive increase of the deforestation
-after 2005, is valid for all four forest types except the spiny forest
-(Tab.~\ref{tab:comp_defor}). For the spiny forest, the deforestation
-rate during the period 2010-2013 was lower than on the period
-2005-2010 (Tab.~\ref{tab:comp_defor}).
-
-\subsection{Comparison with previous forest cover change studies
-in Madagascar}
-
-Forest cover maps provided by previous studies over Madagascar were
-not exhaustive (unclassified areas) due to the presence of clouds on
-satellite images used to produce such maps. In \citet{Harper2007}, the
-maps of years 1990 and 2000 include 0.5 and 1.12 Mha of unknown cover
-type respectively. Proportions of unclassified areas are not reported
-in the two other existing studies at the national level by
-\citet{MEFT2009} and \citet{ONE2015}. With our approach, we produced
-wall-to-wall forest cover change maps from 1990 to 2014 for the full
-territory of Madagascar (Fig.~\ref{fig:fcc}). This allowed us to
-produce more robust estimates of forest cover and deforestation rates
-over this period (Tab.~\ref{tab:forest_cover}). Our forest cover
-estimates over the period 1953-2013 (considering forest cover
-estimates at national level and by ecoregions for all the available
-dates) were well correlated (Pearson's correlation coefficient $=$
-0.99) to estimates from the three previous studies
-(Tab.~\ref{tab:comp_forest}) with a RMSE of 300,000 ha (6\% of the
-mean forest cover of 4.8 Mha when considering all dates and forest
-types together). These small differences can be partly attributed to
-differences in ecoregion boundaries. Despite significant differences
-in deforestation estimates (Tab.~\ref{tab:comp_defor}), a similar
-deforestation trend was observed across studies with a decrease of
-deforestation rates over the period 1990-2005, followed by a
-progressive increase of the deforestation after 2005.
-
-\subsection{Variation of forest fragmentation over time}
-
-Forest fragmentation has progressively increased since 1953 in
-Madagascar. We observed a continuous decrease of the mean distance to
-forest edge from 1953 to 2014 in Madagascar. The mean distance to
-forest edge has decreased to about 300~m in 2014 while it was of about
-1.5~km in 1973 (Fig.~\ref{fig:dist_edge}). Moreover, a large
-proportion (73\%) of the forest was located at a distance greater than
-100~m in 1973, while almost half of the forest (46\%) is at a distance
-lower than 100~m from forest edge in 2014
-(Fig.~\ref{fig:dist_edge}). The percentage of lowly fragmented forest
-in Madagascar has continuously decreased since 1953. The percentage of
-forest belonging to the lowly fragmented class has fallen from 57\% in
-1973 to 44\% in 2014. In 2014, 22\% of the forest belonged to the two
-highest fragmented forest classes (less than 40\% of forest cover in
-the neighborhood) while only 15\% of the forest belonged to these two
-fragmentation classes in 1973 (Tab.~\ref{tab:frag}).
-
-\newpage
-
-\section{Discussion}
-\label{discussion}
-
-\subsection{Advantages of combining recent global annual tree cover loss
-  data with historical national forest cover maps}
-
-In this study, we combined recent (2001-2014) global annual tree cover
-loss data \citep{Hansen2013} with historical (1953-2000) national
-forest cover maps \citep{Harper2007} to look at six decades
-(1953-2014) of deforestation and forest fragmentation in
-Madagascar. We produced annual forest cover maps at 30~m resolution
-covering Madagascar for the period 2000 to 2014. Our study extends the
-forest cover monitoring on a six decades period (from 1953 to 2014)
-while harmonizing the data from previous studies \citep{Harper2007,
-  MEFT2009, ONE2015}. We propose a generic approach to solve the
-problem of forest definition which is needed to transform the 2000
-global tree cover dataset from \citet{Hansen2013} into a
-forest/non-forest map \citep{Tropek2014}. We propose the use of an
-historical national forest cover map, based on a national forest
-definition, as a forest cover mask. This approach could be easily
-extended to other tropical regions or countries for which an accurate
-forest cover map is available at any date within the period 2000-2014
-(but preferably at the beginning of the period to profit from the full
-record of tree cover loss and derive long-term estimates of
-deforestation). For example, forest cover maps are available at 20~m
-resolution for Cameroon and Central African Republic for years 2000
-and 2010 \citep{Gross2017}. When high resolution forest cover maps are
-not available, coarser resolution forest cover maps (leading to
-coarser deforestation estimates) could be extracted from global land
-cover products such as GLC2000 at 1~km resolution
-\citep{Bartholome2005} or CCI Land Cover at 300~m resolution
-\citep{Li2018}.  Moreover, this approach could be repeated in the
-future with the release of updated tree cover loss data. We have made
-the \R{}/GRASS code used for this study freely available in a GitHub
-repository (see Data availability statement) to facilitate application
-to other study areas or repeat the analysis in the future for
-Madagascar.
-
-The accuracy of the derived forest cover change maps depends directly
-on the accuracy of the historical forest cover maps and the tree cover
-loss dataset. Using visual-interpretation of aerial images in 342
-areas distributed among all forest types, \citet{Harper2007} estimated
-an overall 89.5\% accuracy in identifying forest/non-forest classes
-for the year 2000. The accuracy assessment of the tree cover loss
-dataset for the tropical biome reported 13\% of false positives and
-16.9\% of false negatives (see Tab. S5 in \citealt{Hansen2013}). These
-numbers rise at 20.7\% and 20.6\% respectively for the subtropical
-biome. In the subtropical biome, the lower density tree cover canopy
-makes it difficult to detect change from tree cover to bare
-ground. For six countries in Central Africa, with a majority of moist
-dense forest, \citet{Verhegghen2016} have compared deforestation
-estimates derived from the global tree cover loss dataset
-\citep{Hansen2013} with results derived from semi-automated supervised
-classification of Landsat satellite images \citep{Achard2014} and they
-found a good agreement between the two sets of estimates. Therefore,
-our forest cover change maps after 2000 might be more accurate for the
-dense moist forest than for the dry and spiny forest. In another study
-assessing the accuracy of the tree cover loss product across the
-tropics \citep{Tyukavina2015}, authors reported 4\% of false positives
-and 48\% of false negatives in Sub-Saharan Africa. They showed that
-85\% of missing loss occurred on the edges of other loss patches. This
-means that tree cover loss might be underestimated in Sub-Saharan
-Africa, probably due to the prevalence of small-scale disturbance
-which is hard to map at 30~m, but that areas of large-scale
-deforestation are well identified and spatial variability of the
-deforestation is well represented. A proper accuracy assessment of our
-forest cover change maps should be performed to better estimate the
-uncertainty surrounding our forest cover change estimates in
-Madagascar from year 2000 \citep{Olofsson2013,Olofsson2014}. Despite
-this limitation, we have shown that the deforestation trend we
-observed for Madagascar, with a doubling deforestation on the period
-2010-2014 compared to 2000-2005, was consistent with the other studies
-at the national scale \citep{ONE2015, MEFT2009}.
-
-Consistent with \citet{Harper2007}, we did not consider potential
-forest regrowth in Madagascar (although \citet{Hansen2013} provided a
-tree cover gains layer for the period 2001-2013) for several
-reasons. First, the tree gain layer of \citet{Hansen2013} includes and
-catches more easily tree plantations than natural forest regrowth
-\citep{Tropek2014}. Second, there is little evidence of natural forest
-regeneration in Madagascar \citep{Grouzis2001, Harper2007}. This can
-be explained by several ecological processes following burning
-practice such as soil erosion \citep{Grinand2017} and reduced seed
-bank due to fire and soil loss \citep{Grouzis2001}. Moreover, in areas
-where forest regeneration is ecologically possible, young forest
-regrowth are more easily re-burnt for agriculture and pasture. Third,
-young secondary forests provide more limited ecosystem services
-compared to old-growth natural forests in terms of biodiversity and
-carbon storage \citep{Martin2013}.
-
-\subsection{Natural forest cover change in Madagascar from 1953 to 2014}
-
-We estimated that natural forest in Madagascar covers 8.9 Mha in 2014
-(corresponding to 15\% of the country) and that Madagascar has lost
-44\% of its natural forest since 1953 (37\% since 1973). If there are
-no doubts about the direct causes of deforestation in Madagascar,
-attributable to human activities such as slash-and-burn agriculture
-and pasture \citep{Scales2011}, there is ongoing scientific debate
-about the extent of the ``original'' forest cover in Madagascar, and
-the extent to which humans have altered the natural forest landscapes
-since their large-scale settlement around 800 CE \citep{Burns2016,
-  Cox2012}. Early French naturalists stated that the full island was
-originally covered by forest \citep{Humbert1927, Perrier1921}, leading
-to the common statement that 90\% of the natural forests have
-disappeared since the arrival of humans on the island
-\citep{Kull2000}. More recent studies counter-balanced that point of
-view saying that extensive areas of grassland existed in Madagascar
-long before human arrival and were determined by climate, natural
-grazing and other natural factors \citep{Vorontsova2017,
-  Virah-Sawmy2009}. Other authors have questioned the entire narrative
-of extensive alteration of the landscape by early human activity
-which, through legislation, has severe consequences on local people
-\citep{Klein2002, Kull2000}. Whatever the original proportion of
-natural forests and grasslands in Madagascar, our results demonstrate
-that human activities since the 1950s have profoundly impacted the
-natural tropical forests and that conservation and development
-programs in Madagascar have failed to stop deforestation in the recent
-years. Deforestation has strong consequences on biodiversity and
-carbon emissions in Madagascar. Around 90\% of Madagascar's species
-are forest dependent \citep{Allnutt2008, Goodman2005}. Based on
-occurrence data for 2243 plant and invertebrate species,
-\citet{Allnutt2008} estimated that deforestation between 1953 and 2000
-has led to an extinction of 9\% of the species. The additional
-deforestation we observed over the period 2000-2014 (around 1 Mha of
-natural forest) worsen this result. Regarding carbon emissions, using
-the 2010 aboveground forest carbon map by \citet{Vieilledent2016}, we
-estimated that deforestation on the period 2010-2014 has led to 40.2
-Mt C of carbon emissions in the atmosphere (10 Mt C /yr) and that the
-remaining aboveground forest carbon stock in 2014 is 832.8 Mt
-C. Associated to deforestation, we showed that the remaining forests
-of Madagascar are highly fragmented with 46\% of the forest being at
-less than 100~m of the forest edge. Small forest fragments do not
-allow to maintain viable populations and ``edge effects'' at
-forest/non-forest interfaces have impacts on both carbon emissions
-\citep{Brinck2017} and biodiversity loss \citep{Gibson2013,
-  Murcia1995}.
-
-\subsection{Deforestation trend and impacts on conservation and
-  development policies}
-
-In our study, we have shown that the progressive decrease of the
-deforestation rate on the period 1990-2005 was followed by a
-continuous increase in the deforestation rate on the period
-2005-2014. In particular, we showed that deforestation rate has more
-than doubled on the period 2010-2014 compared to 2000-2005. Our
-results are supported by previous studies \citep{Harper2007, MEFT2009,
-  ONE2015} despite differences in the methodologies regarding (i)
-forest definition (associated to independent visual interpretations of
-observation polygons to train the classifier), (ii) classification
-algorithms, (iii) deforestation rate computation method, and (iv)
-correction for the presence of clouds. Our deforestation rate
-estimates from 1990 to 2014 have been computed from wall-to-wall maps
-at 30~m resolution and can be considered more accurate in comparison
-with estimates from these previous studies. Our natural forest cover
-and deforestation rate estimates can be used as source of information
-for the next FAO Forest Resources Assessment
-\citep{Keenan2015}. Current rates of deforestation can also be used to
-build reference scenarios for deforestation in Madagascar and
-contribute to the implementation of deforestation mitigation
-activities in the framework of REDD+ \citep{Olander2008}.
-
-The increase of deforestation rates after 2005 can be explained by
-population growth and political instability in the country. Nearly
-90\% of Madagascar's population relies on biomass for their daily
-energy needs \citep{Minten2013} and the link between population size
-and deforestation has previously been demonstrated in Madagascar
-\citep{Vieilledent2013, Gorenflo2011}. With a mean demographic growth
-rate of about 2.8\%/yr and a population which has increased from 16 to
-24 million people on the period 2000-2015 \citep{UN2015}, the
-increasing demand in wood-fuel and space for agriculture is likely to
-explain the increase in deforestation rates. The political crisis of
-2009 \citep{Ploch2012}, followed by several years of political
-instability and weak governance could also explain the increase in the
-deforestation rate observed on the period 2005-2014 (see
-\citealp{Smith2003} for a discussion on the link between governance
-and forest cover loss). These results show that despite the
-conservation policy in Madagascar \citep{Freudenberger2010},
-deforestation has dramatically increased at the national level since
-2005. Results of this study, including recent spatially explicit
-forest cover change maps and forest cover estimates, should help
-implement new conservation strategies to save Madagascar natural
-tropical forests and their unique biodiversity.
-
-\newpage
-
-\section{Author's contribution}
-\label{authors-contribution}
-
-All authors conceived the ideas and designed the methodology; GV
-analysed the data and wrote the {\R}/GRASS script; GV drafted the
-manuscript. All authors contributed critically to the drafts and gave
-final approval for publication.
-
-\section{Acknowledgements}
-\label{acknowledgements}
-
-The authors thank two anonymous reviewers for useful comments on a
-previous version of the manuscript. They also thank Peter Vogt for
-useful advice on which metric to use to estimate forest
-fragmentation. This study is part of the CIRAD's BioSceneMada project
-(\url{https://bioscenemada.cirad.fr}) and the Joint Research Center's
-ReCaREDD project (\url{http://forobs.jrc.ec.europa.eu/recaredd}). The
-BioSceneMada project is funded by FRB (Fondation pour la Recherche sur
-la Biodiversité) and the FFEM (Fond Français pour l'Environnement
-Mondial) under the project agreement AAP-SCEN-2013~I. The ReCaREDD
-project is funded by the European Commission. The authors declare that
-there are no conflicts of interest related to this article.
+Blabla
 
@@
 # Caption
-tabcap1 <- "\\textbf{Change in natural forest cover and deforestation rates from 1953 to 2014 in Madagascar.} Areas are provided in thousands of hectares (Kha).
+tabcap1 <- "\\textbf{Change in natural forest cover and deforestation rates from 1953 to 2017 in Madagascar.} Areas are provided in thousands of hectares (Kha).
 Forest map for the year 1973 has 3.3 Mha of unclassified areas due to the presence of clouds on satellite images.
@@
 tab.2$ForestType <- c("Total",rep("",3),"Moist",rep(" ",3),"Dry",rep(" ",3),"Spiny",rep(" ",3),"Mangroves",rep(" ",3))
-col.names <- c("Forest type","Source","1953","1973","1990","2000","2005","2010","2013","2014")
+col.names <- c("Forest type","Source","1953","1973","1990","2000","2005","2010","2013","2017")
 # Caption
 tabcap2 <- "\\textbf{Comparing Madagascar forest cover estimates with
-      previous studies on the period 1953-2014}. We compared our
+      previous studies on the period 1953-2017}. We compared our
     estimates of forest cover with the estimates from three previous
@@
 # Kable
-knitr::kable(tab.3, col.names=col.names, align=c("l","l",rep("r",6)), booktabs=TRUE, 
+knitr::kable(tab.3, col.names=col.names, align=c("l","l",rep("r",6)), booktabs=TRUE, format="latex",
              caption=tabcap3, linesep=c("","","","\\addlinespace")) %>% 
@@
 tabcap4 <- "\\textbf{Change in forest fragmentation from 1953 to
-      2014 in Madagascar}. Five forest fragmentation classes, based on
+      2017 in Madagascar}. Five forest fragmentation classes, based on
     the percentage of forest in the neighborhood, are defined: 0-20\\%
@@
 # Kable
-knitr::kable(tab.4, digits=0, format.args=list(big.mark=","), booktabs=TRUE,
+knitr::kable(tab.4, digits=0, format.args=list(big.mark=","), booktabs=TRUE, format="latex",
              caption=tabcap4, linesep="") %>%
@@
       vegetation \citep{IEFN1996} criteria. The dark grey areas
-      represent the remaining natural forest cover for the year 2014.
+      represent the remaining natural forest cover for the year 2017.
       Forest types are defined on the basis of their belonging to
@@
   \caption{\textbf{Forest cover change on six decades from 1953 to
-      2014 in Madagascar.} forest cover changes from 1973 to
-    2014 are shown in the main figure, and forest cover in 1953 is shown 
+      2017 in Madagascar.} forest cover changes from 1973 to
+    2017 are shown in the main figure, and forest cover in 1953 is shown 
     in the bottom-right inset. Two zooms in the western
@@
     more detailed views of (from top to bottom): forest cover in
-    1953, forest cover change from 1973 to 2014, forest
-    fragmentation in 2014 and distance to forest edge in 2014. Data on
+    1953, forest cover change from 1973 to 2017, forest
+    fragmentation in 2017 and distance to forest edge in 2017. Data on
     water bodies (blue) and water seasonality (light blue for seasonal
@@
   \caption{\textbf{Change in distance to forest edge from 1953
-      to 2014 in Madagascar.} Black dots represent the mean distance
+      to 2017 in Madagascar.} Black dots represent the mean distance
     to forest edge for each year. Vertical dashed segments represent
@@
 
-\newpage
-
-\section{Supplementary material}
-
-\setcounter{figure}{0}
-\renewcommand{\thefigure}{A\arabic{figure}}
-
-\vfill
-\begin{figure}[h!]
-  \centering
-
-  \includegraphics[width=12cm]{figs/tc_threshold_moist.pdf}
-
-  \caption{\textbf{Selection of the tree cover threshold to define the
-      land type for unclassified areas in 2000.} We considered the
-    forest in 2000 \citep{Harper2007} in the moist ecoregion (see
-    Fig.~\ref{fig:ecoregion}). We plotted the percentage of forest
-    having a tree cover greater or equal to a given value specified by
-    the x-axis. From this figure, we see that 90\% of the moist forest
-    in 2000 has a tree cover $\geq$ 75.}
-
-  \label{fig:tc_threshold}
-
-\end{figure}
-\vfill
-
 \end{document}

### 2018-10-11 deforestmap.R

@@ -58,10 +58,6 @@ initGRASS(gisBase="/usr/lib/grass72",home=tempdir(),
           location="deforestmap",mapset="PERMANENT",
           override=TRUE)
 
-# TEMP!
-execute <- FALSE
-if (execute) {
-
 ##==================================================================
 ## Download data from Cirad Dataverse: http://dx.doi.org/10.18167/DVN1/2FP7LR
 down <- FALSE
@@ -854,9 +850,6 @@ for (j in 1:length(ecoregion)) {
   system("r.mask -r")
 }
 
-# TEMP!
-} # End if (execute) 
-
 ##========================================================
 ## Comparison with previous studies on forest-cover change
 ##========================================================
@@ -878,18 +871,18 @@ fcc.comp[fcc.comp$ForestType=="Moist" & fcc.comp$Source=="this study", 3:9] <- d
 fcc.comp[fcc.comp$ForestType=="Dry" & fcc.comp$Source=="this study", 3:9] <- defor_dry_for_comp$area
 fcc.comp[fcc.comp$ForestType=="Spiny" & fcc.comp$Source=="this study", 3:9] <- defor_spiny_for_comp$area
 fcc.comp[fcc.comp$ForestType=="Mangroves" & fcc.comp$Source=="this study", 3:9] <- defor_mangroves_for_comp$area
-# This study for year 2014
+# This study for year 2017
 defor <- read.table("outputs/defor.txt",header=TRUE)
 defor_moist <- read.table("outputs/defor_moist.txt",header=TRUE)
 defor_dry <- read.table("outputs/defor_dry.txt",header=TRUE)
 defor_spiny <- read.table("outputs/defor_spiny.txt",header=TRUE)
 defor_mangroves <- read.table("outputs/defor_mangroves.txt",header=TRUE)
-# Complete data for year 2014
-fcc.comp$y2014[fcc.comp$ForestType=="Total" & fcc.comp$Source=="this study"] <- defor$area[defor$Year==2014]
-fcc.comp$y2014[fcc.comp$ForestType=="Moist" & fcc.comp$Source=="this study"] <- defor_moist$area[defor_moist$Year==2014]
-fcc.comp$y2014[fcc.comp$ForestType=="Dry" & fcc.comp$Source=="this study"] <- defor_dry$area[defor_dry$Year==2014]
[…]

### 2018-10-12 update_2017/update_2017.Rnw
Update 2017

@@
 {\small
-\noindent$^{\S}$ This article was initially published as:
+\noindent$^{\S}$ Article originally published as:
 
@@
 
-\doublespacing
+%\doublespacing
 %\linenumbers
 
-\newcommand{\keywords}[1]{\par\noindent
-{\small{\em Keywords\/}: #1}}
+\section{Summary}
 
-\begin{abstract}
+In a recent paper \citep{Vieilledent2018}, we have published estimates of deforestation and forest fragmentation in Madagascar for the period 1953-2014.
+We had used a method combining historical forest cover maps (1953-2000) with annual tree cover loss data (2000-2014) to derive new forest cover maps for the period 2000-2014.
+We had shown that deforestation rate had increased since 2005 to reach about 100,000 ha/yr (1\%/yr) for the period 2010-2014 and that remaining forest was highly fragmented, 46\% of the remaining forest in 2014 (covering 8.9 Mha) being at less than 100 m from forest edge or a forest clearing.
 
-Blabla
+In this follow-up study, we have used version 1.5 of the Global Forest Change data \citep{Hansen2013}, providing updated tree cover loss data for the period 2014-2017 (\url{http://earthenginepartners.appspot.com/science-2013-global-forest}), to provide updated estimates of deforestation and forest fragmentation in Madagascar on the periods 2010-2015 and 2015-2017.
+To do so, we used the exact same methodology as the one described in \citet{Vieilledent2018}.
 
-\vspace{0.5cm}
+We provide new forest cover and forest fragmentation maps for Madagascar for the years 2015 and 2017 (Fig.~\ref{fig:fcc}).
+We estimate that tropical forests cover 8.4 Mha in 2017 in Madagascar.
+We show that the deforestation trend observed on the period 2005-2014 in \citet{Vieilledent2018} has been accelerating on the period 2015-2017.
+During these two years, 324,000 hectares have been deforested at the national scale in Madagascar (Tab.~\ref{tab:forest_cover}).
+This corresponds to a 56\% increase in the annual deforestation rate compared to the period 2010-2015 (Tab.~\ref{tab:forest_cover}).
+Associated to deforestation, remaining forest keeps on being highly fragmented (Tab.~\ref{tab:frag}).
+At the end of the year 2017, almost half (49\%) of the forest in Madagascar was located at a distance lower than 100 m from forest edge or a forest clearing (Fig.~\ref{fig:dist_edge}).
 
-\keywords{biodiversity, climate-change, deforestation, forest-fragmentation, Madagascar, tropical forest}
+Some limitations about using version 1.5 of the tree cover loss data to analyze the deforestation trend on the full period 2000-2017 are discussed on the Global Forest Change website.
+Only the satellite data from 2011 onward have been reprocessed in this new version, so that comparing deforestation on the period 2000-2012 with deforestation on the period 2011-2017 should be performed with caution.
+In our case, we can confidently compare deforestation on the period 2010-2015 with deforestation on the period 2015-2017.
 
-\end{abstract}
+Increase in the deforestation rates in Madagascar is most likely associated to demographic growth and poor environmental law enforcement.
+These news results are alarming.
+If nothing is done to rapidly curb deforestation in the country, Madagascar risks loosing almost all its tropical forest and associated biodiversity in the next 50 years.
 
 \newpage
-
-\section{Introduction}
-\label{introduction}
-
-Blabla
 
@@
 
-Data and code used for this study have been made permanently and publicly available
-on the CIRAD Dataverse repository so that the results are entirely reproducible:
+Results of this study are entirely reproducible. Data and code used in the analysis have been archived on the CIRAD Dataverse repository:
 \begin{itemize}
@@
 \end{itemize}
-
-\newpage
-\singlespacing
 
@@
 
-\vspace*{\stretch{1}}
-<<comp_forest, results="asis">>=
-# Import and format
-fcc.comp <- read.table(file="tabs/fcc_comp.txt", sep=" ", header=TRUE)
-tab.2 <- fcc.comp
-tab.2 <- tab.2[,c(2,1,3:10)]
-tab.2[,c(3:10)] <- round(tab.2[,c(3:10)]/1000)
-tab.2$ForestType <- c("Total",rep("",3),"Moist",rep(" ",3),"Dry",rep(" ",3),"Spiny",rep(" ",3),"Mangroves",rep(" ",3))
-col.names <- c("Forest type","Source","1953","1973","1990","2000","2005","2010","2013","2017")
-# Caption
-tabcap2 <- "\\textbf{Comparing Madagascar forest cover estimates with
-      previous studies on the period 1953-2017}. We compared our
-    estimates of forest cover with the estimates from three previous
-    studies \\citep{Harper2007, MEFT2009, ONE2015}. Areas are provided
-    in thousands of hectares (Kha). We obtained a Pearson's
-    correlation coefficient of 0.99 between our forest cover estimates
-    and forest cover estimates from previous studies. The increase in
-    mangrove and spiny forest covers from 1953 to 1973 in \\citet{Harper2007} and our study is most probably due to
-    differences in forest definition and mapping methods between the
-    1953 aerial-photography derived map and the 1973 Landsat image
-    derived map."
-# Kable
-knitr::kable(tab.2, col.names=col.names, align=c("l","l",rep("r",8)), format.args=list(big.mark=","), 
-             booktabs=TRUE, format="latex", caption=tabcap2, linesep=c("","","","\\addlinespace")) %>% 
-  kable_styling(latex_options=c("striped","hold_position","scale_down"))
-@
-\vspace*{\stretch{1}}
-
-\newpage
-
-\vspace*{\stretch{1}}
-<<comp_defor, results="asis">>=
-# Import and format
-tab.3 <- read.table(file="tabs/defor_comp.txt", sep=" ", header=TRUE)
-tab.3$ForestType <- c("Total",rep("",3),"Moist",rep(" ",3),"Dry",rep(" ",3),"Spiny",rep(" ",3),"Mangroves",rep(" ",3))
-col.names <- c("Forest type","Source","1953-1973","1973-1990","1990-2000","2000-2005","2005-2010","2010-2013")
-# Caption
-tabcap3 <- "\\textbf{Comparing Madagascar annual deforestation rates
-      with previous studies on the period 1953-2013}. Annual
-    deforested areas (in thousands of hectares per year, Kha/yr) and
-    annual deforestation rates (second number in parenthesis, in
-    \\%/yr) are provided. For deforestation rates in \\%/yr, exact same
-    numbers as in scientific articles and reports from previous
-    studies \\citep{Harper2007, MEFT2009, ONE2015} have been
-    reported. The way annual deforestation rates in \\%/yr have been
-    computed in these previous studies can slightly differ from one
-    study to another, but estimates always correct for the potential
-    presences of clouds on satellite images and unclassified areas on
-    forest maps. Annual deforested areas in Kha/yr have been
-    recomputed from forest cover estimates in
-    Tab.~\\ref{tab:comp_forest} (except for \\citet{Harper2007} for the
-    periods 1973-1990 and 1990-2000 for which annual deforested areas
-    in Kha/yr were derived from numbers reported in the original
-    publication, see methods) and do not correct for the potential
-    presence of clouds."
-# Kable
-knitr::kable(tab.3, col.names=col.names, align=c("l","l",rep("r",6)), booktabs=TRUE, format="latex",
-             caption=tabcap3, linesep=c("","","","\\addlinespace")) %>% 
-  kable_styling(latex_options=c("striped","hold_position","scale_down"))
-@
-\vspace*{\stretch{1}}
-
-\newpage
+% \vspace*{\stretch{1}}
+% <<comp_forest, results="asis">>=
+% # Import and format
+% fcc.comp <- read.table(file="tabs/fcc_comp.txt", sep=" ", header=TRUE)
+% tab.2 <- fcc.comp
+% tab.2 <- tab.2[,c(2,1,3:10)]
+% tab.2[,c(3:10)] <- round(tab.2[,c(3:10)]/1000)
+% tab.2$ForestType <- c("Total",rep("",3),"Moist",rep(" ",3),"Dry",rep(" ",3),"Spiny",rep(" ",3),"Mangroves",rep(" ",3))
+% col.names <- c("Forest type","Source","1953","1973","1990","2000","2005","2010","2013","2017")
+% # Caption
+% tabcap2 <- "\\textbf{Comparing Madagascar forest cover estimates with
+%       previous studies on the period 1953-2017}. We compared our
+%     estimates of forest cover with the estimates from three previous
+%     studies \\citep{Harper2007, MEFT2009, ONE2015}. Areas are provided
+%     in thousands of hectares (Kha). We obtained a Pearson's
+%     correlation coefficient of 0.99 between our forest cover estimates
+%     and forest cover estimates from previous studies. The increase in
+%     mangrove and spiny forest covers from 1953 to 1973 in \\citet{Harper2007} and our study is most probably due to
+%     differences in forest definition and mapping methods between the
+%     1953 aerial-photography derived map and the 1973 Landsat image
+%     derived map."
+% # Kable
+% knitr::kable(tab.2, col.names=col.names, align=c("l","l",rep("r",8)), format.args=list(big.mark=","), 
+%              booktabs=TRUE, format="latex", caption=tabcap2, linesep=c("","","","\\addlinespace")) %>% 
+%   kable_styling(latex_options=c("striped","hold_position","scale_down"))
+% @
+% \vspace*{\stretch{1}}
+% 
+% \newpage
+% 
+% \vspace*{\stretch{1}}
+% <<comp_defor, results="asis">>=
+% # Import and format
+% tab.3 <- read.table(file="tabs/defor_comp.txt", sep=" ", header=TRUE)
+% tab.3$ForestType <- c("Total",rep("",3),"Moist",rep(" ",3),"Dry",rep(" ",3),"Spiny",rep(" ",3),"Mangroves",rep(" ",3))
+% col.names <- c("Forest type","Source","1953-1973","1973-1990","1990-2000","2000-2005","2005-2010","2010-2013")
+% # Caption
+% tabcap3 <- "\\textbf{Comparing Madagascar annual deforestation rates
+%       with previous studies on the period 1953-2013}. Annual
+%     deforested areas (in thousands of hectares per year, Kha/yr) and
+%     annual deforestation rates (second number in parenthesis, in
+%     \\%/yr) are provided. For deforestation rates in \\%/yr, exact same
+%     numbers as in scientific articles and reports from previous
+%     studies \\citep{Harper2007, MEFT2009, ONE2015} have been
+%     reported. The way annual deforestation rates in \\%/yr have been
+%     computed in these previous studies can slightly differ from one
+%     study to another, but estimates always correct for the potential
+%     presences of clouds on satellite images and unclassified areas on
+%     forest maps. Annual deforested areas in Kha/yr have been
+%     recomputed from forest cover estimates in
+%     Tab.~\\ref{tab:comp_forest} (except for \\citet{Harper2007} for the
+%     periods 1973-1990 and 1990-2000 for which annual deforested areas
+%     in Kha/yr were derived from numbers reported in the original
+%     publication, see methods) and do not correct for the potential
+%     presence of clouds."
+% # Kable
+% knitr::kable(tab.3, col.names=col.names, align=c("l","l",rep("r",6)), booktabs=TRUE, format="latex",
+%              caption=tabcap3, linesep=c("","","","\\addlinespace")) %>% 
+%   kable_styling(latex_options=c("striped","hold_position","scale_down"))
+% @
+% \vspace*{\stretch{1}}
+% 
+% \newpage
 
@@
 
+% \vfill
+% \begin{figure}[h!]
+%   \centering
+%     
+%     \includegraphics[width=10cm]{figs/ecoregion.png}
+% 
+%     \caption{\textbf{Ecoregions and forest types in Madagascar.}
+%       Madagascar can be divided into four climatic ecoregions with
+%       four forest types: the moist forest in the East (green), the dry
+%       forest in the West (orange), the spiny forest in the South
+%       (red), and the mangroves on the West coast (blue). Ecoregions
+%       were defined following climatic \citep{Cornet1974} and
+%       vegetation \citep{IEFN1996} criteria. The dark grey areas
+%       represent the remaining natural forest cover for the year 2017.
+%       Forest types are defined on the basis of their belonging to
+%       one of the four ecoregions.}
+%    
+%     \label{fig:ecoregion}
+%     
+% \end{figure}
+% \vfill
+% 
+% \newpage
+
+\enlargethispage{\baselineskip}
 \vfill
-\begin{figure}[h!]
-  \centering
-
-    \includegraphics[width=10cm]{figs/ecoregion.png}
-
-    \caption{\textbf{Ecoregions and forest types in Madagascar.}
-      Madagascar can be divided into four climatic ecoregions with
-      four forest types: the moist forest in the East (green), the dry
-      forest in the West (orange), the spiny forest in the South
-      (red), and the mangroves on the West coast (blue). Ecoregions
-      were defined following climatic \citep{Cornet1974} and
-      vegetation \citep{IEFN1996} criteria. The dark grey areas
-      represent the remaining natural forest cover for the year 2017.
-      Forest types are defined on the basis of their belonging to
-      one of the four ecoregions.}
-
-    \label{fig:ecoregion}
-
-\end{figure}
-\vfill
-
-\newpage
-
-\vspace*{\stretch{1}}
 \begin{figure}[h!]
@@
 \end{figure}
-\vspace*{\stretch{1}}
+\vfill
 
